### 2021-08-27 paper/Paper.Rnw
cp[ying files

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}

%external documents
\externaldocument[SI-]{SupMat}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\DeclareMathOperator{\E}{E}% expected value
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\cor}{cor}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\prob}{P}

%attempt 1 at nat and sharp
%\newcommand{\nat}{\mathlarger{\natural}}
%\newcommand{\shp}{\mathlarger{\sharp}}

%attempt 2 at nat and sharp
%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
%\newcommand{\shp}{\#}

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\bs}{\backslash}

%header material for paper
\title{Relationships in the extremes and their influence on competition and coexistence}
\date{}

%Put other possible titles below, commented, and we will select the best one at the end
%Relationships between extreme events and their influence on competition and coexistence
%Coexistence and consequences of extreme environmental relationships
%Coexistence and the consequences of extreme environmental relationships
%Coexistence and consequences of environmental relationships in their extremes
%Coexistence and implications of asymmetric environmental relationships
%Extreme [climatic] events and their implications for biodiversity
%Deductions to coexistence, extreme events, and asymmetric relationships
%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence
%Coexistence and relationships between environmental variables in their extremes

\author[a]{Pimsupa Jasmin Albert}
\author[a,b]{Daniel C. Reuman}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}

%***Need to indicate corresponding authors

\begin{document}
\SweaveOpts{concordance=TRUE}
\SweaveOpts{concordance=TRUE}
\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text

<<eval=T,echo=F,message=F,warning=F>>=
#system(paste("biber", sub("\\.Rnw$", "", current_input())))

#place R code here for loading in necessary variables from Results
#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
#allregres<-matrix(rnorm(10),2,5)
fig2data <- load(file="../results_numeric/fig2data.rda")
M <- load(file="../results_numeric/M.rda")
rho <- load(file="../results_numeric/rho.rda")
se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
@

\maketitle

\begin{abstract}
Place abstract here
\end{abstract}

%***Add to these
\keywords{kw1, kw2,kw3}

%Here are the paragraph-level "rules" of writing, for reference: 
%Every paragraph needs to have a clear topic sentence.
%It needs to be one of the first three sentences.
%If it is not the first sentence, then the sentences preceding it are for background, in order to make it possible to state the topic sentence.
%The topic sentence is the main *result* or *claim* of the paragraph, not just a statement of what the topic is.
%All sentences in the paragraph support and validate the claim made in the topic sentence.

\section{Introduction}\label{section:introduction}

%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
%biodiversity (Chesson, 1994).
%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
%(more than the abstracts if you want), and consider using them as references here.
The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
the other and drive it extinction \citep{Hardin_1960}.
Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
framework for understanding how stable species coexistence 
nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{}, 
%***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
%in a similar way to how I filled in the Angert example.
%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
%right column on the first page of the Barabas review (Ecol Monogr, 2018).
even though it is mathematically complex in the eyes
of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
%more detail than you previously gave to do that.

%Old version, some comments interspersed: 
%Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
%***DAN: "Blanketed" is a weird verb in this context. 
%***DAN: Also, MCT is a conceptually deep theory, and ATA is a fairly simple observation that happens to have some consequences. I would not be confortable with phrasing that seems to put them on the same footing! 
%Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
%More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
%Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
%Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
%For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
%The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
%Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
%In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events. Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Ghosh_2020_Eco, Ghosh_2020_AER}.
%Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
%New version:
Every ecological discipline involves the study of relationships among biological and environmental variables,
but standard methods for examining the association between variables provide only a small portion of the overall information content contained in the
relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
the three pairs of variables in figure \ref{fig:noise} all have the same
Pearson correlation; and likewise have the same Spearman correlation. 
%***DAN: Jasmin, when you make the figure, please display the Pearson and Spearman correlations on each panel, as in 
%Fig. 1 of Ghosh et al (2020, Ecology and Evolution).
But each panel of figure \ref{fig:noise} nevertheless
depicts a very different type of association. Other metrics of association are used, in addition to correlations
(e.g., various common forms of regression), but correlations are extremely common, and other common
methods also provide only a small portion of the available information \citep{Nelsen2006}.
The variables in figure \ref{fig:noise}A,C 
%***DAN: Jasmin, when you make the figure, put the two asymmetric cases in A and C and put the symmetric case in B, in the middle.
exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
they are more strongly related to each
other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
%***DAN: Jasmin, please insert bib entries for the references Anderson et al., 2018; de Valpine et al., 2014; Popovic et al., 2019
%from the Ghosh et al (2020; Adv Ecol Res) paper, and cite them here.
our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
%***DAN: Again note the similarity between the old and new versions, the new is just a bit more fleshed out, and maybe a
%bit more precise in some of the wordings in places.

%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
why ATAs may be relevant to them.
%***DAN: I have inserted some comments below that might help.
Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018} ].
%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects. 
In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}. 
In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low. 
Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones. 
For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.   
Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.

%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
%Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
%While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
%For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
%***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
%We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive.

%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.

%***DAN: stopped here 2021 07 20

%[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation). 
%There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities. Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013} 
Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004, Ummenhofer_2017}. Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition. Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme drought \citep{White_2000, White_2001, Hoover_2014}. Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.  \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics. Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease. Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}. Increased variability of environmental fluctuations is a newer consideration for studying environmental changes in comparison to changes in means; however, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}. Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}. This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.

Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (similar, correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. This is a change in ATA not just from the event alone, but from the relationship between species. Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environment (but see Ghosh for impact of ATA on community aggregate properties).

To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model. Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. We simulated two different communites up to 100,000 time steps to hopefully make it clear to the reader than ATA makes a difference (see figure 2). The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not. Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, almost seven times as much....

To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}). The community follows the criteria required for temporal storage effects to operate. The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season. The grass dominates and has high fecundity when there is sufficeint rain in the late season. Competition is influenced by resource comsumption of growing juvenile plants. Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment. When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year. The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated. However, if environmental responses become asymmetrically associated by, say, extreme rainfall events coexistence can be altered. If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated. Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated. We see the possible consequences of ATA in Figure 2. In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.

%Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
%Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
%Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
%Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.

%If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes. 
%Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics. If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events. Both species may become more alike (similar, correlated), an important consideration for competition and coexistence. For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution. This is a change in ATA not just from the event alone, but from the relationship between species. Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.

%move this paragraph to discussion
%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers. 
%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be. 
%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].

This paper has the following goals. 
G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
We address how ATA contributes to these components. 
We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
are likely to be negligible or meaningfully large. 
%G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
%***DAN: This is a good additional goal. We will either have to have already defined the "ATA rescue and ATA exlusion sets of trait space"
%or else readers won't understand this as it was written, above (now commented). These are complex concepts, so it may be better to 
%save formal definitions for later and reword this. Here's a try at rewording: 
G3) To explore, using our simple model, the likelihood that species can coexist with
ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
may be negligible or meaninfully large.
G4) To provide an empirical example of species coexistence, and to highlight the role ATAs plays in coexistence for this example.
Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climatic changes, 
and a means for beginning to understand its importance for real systems.

\section{Theory}\label{section:theory}

We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
plays a role.
%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.

The lottery model is as follows. 
%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year.
For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
properties, corresponding to various tail association cases. Note that $N$ is constant through time.

Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
%***DAN: Search for REF and replace all with appropriate citations
quantify the contributions of 
different mechanisms to species coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
We now define storage effects for a general two species model, and specifically 
for the lottery model, using an approach and notation
similar to those of \cite{Ellner_2016;Ellner_2019}; see those papers for additional details.

For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. As for the lottery model example, we here assume, for simpicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
Species growth rates for the lottery model can then be written 
$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.

In general, the storage effect for species $i$ is the contribution of covariation between $E$ and 
$C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
Specifically, storage effects for species $i$ are 
\begin{equation}
\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
\end{equation}
where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' (IGR) of species $i$, that is, the mean growth rate of 
species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
$\olr_{j \bs i}$ refers to the mean steady-state 
growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
influences of environment and competition on the invader or resident have been rendered independent.  
Following \cite{Ellner_2016,Ellner_2019}, a subscript 
$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
The \emph{scaling factor}
$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$. 
Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions. Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.

For the lottery model, when $i$ is an invader, the competition experienced by either 
species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
Likewise, 
$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
though we already knew this equals $0$. 
An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
We show in SI section \ref{SI-SIsect:scaling_factors}
that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
Since we assumed log-normally distributed $B_i$, 
$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SI-SIsect:scaling_factors}.
%Say this here if it turns out to be true, or say something like it:
%These alternatives turn out to make very little difference to our results. 
Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between invader and resident mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
having removed the effects of ATAs.
The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decompostion within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term within the decomposition provided by the MTC \citep{Ellner_2019} -- see below for details.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume, for simplicity, that the cummulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
in one of 
four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
described in the Introduction.
%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
%the *right* way
For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.

For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further1}) for the lottery model using the simulation 
approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.

The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas asymmetries of tail associations are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ which is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
coexistence opperating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of the MTC which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
the magnitude of the new ATA term in this expression, $\Delta_i^{[E \nat C]}$, relative to the magnitudes of the other terms.

\section{Methods}\label{section:methods}

%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
%from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
%are likely to be negligible or meaningfully large.

Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
% and $r_1-q_{12}r_2$?
against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
for left-tail associated noise. For each combination of $\delta=0.2, 0.5, 0.8, 1.0$
and $\mu_1-\mu_2=0.0,-0.3,-0.6,-0.8$,
we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
the contributions of other mechanisms.
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
SI section \ref{SI-sect:efficient_computation}.

%G3) To explore, using our simple model, the likelihood that species can coexist with
%ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
%may be negligible or meaninfully large.

To acheive goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
ATAs but not without, or vice versa, we plotted the IGR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
as well as the
IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$. 
For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
we were interested in the values of $\mu_1-\mu_2$ for which one of these
quantities is positive and the other is negative; this set of value characterizes the species which can coexist
with ATAs, but not without, or vice versa. The one dimensional space of values $\mu$ is a species trait space,
and $\mu_1$ and $\mu_2$ can be regarded as two points in it. But we instead define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
Points of the trait-difference space characterize possible combinations of traits of the two competing species in our model. For fixed values of 
$\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of 
values $\mu_1 - \mu_2$ for which the the IGR, $\olr_{i \bs i}$, is positive. This 
is the subspace for which the two species in the lottery model coexist in the presence of ATAs.
We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of values
$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$ is positive.
This is the subspace for which the two species in the lottery model coexist in the absence of 
ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ is intended to indicate the
absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
values $\mu_1 - \mu_2$ for which model species coexist in the presence of ATAs, but not in the absence of ATAs. 
Similarly, we define ``ATA exclusion space'' to be the set difference 
$\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$
against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
$\mu_1 - \mu_2$ for all combinations of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.

%***DAN: Jasmin, I edited the below to get the above. Kept the below so I could add one comment to it, and in case it is needed again in
%the near future
%to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR], 
%***DAN: Note that this is not the IGR. The IGR is just r_{1\1}. So I abandoned this terminology and used "coexistence metric" instead
%$\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$. The same mathematical simplifications described above hold, such that $\mu_1$ and $\mu_2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$. We henceforward call this difference the species trait difference and sets of multiple $\mu_1-\mu_2$ are sets of trait space range. Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space. Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the set of points between $\mu_1$ and $\mu_2$ is a subset of trait space, and the difference, $\mu_1-\mu_2$, is the range of the subset of trait space. We then define $\psi_{ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is included (coexistence with ATA) and $\psi_{\sim ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is not included (coexistence without ATA). We then consider the difference between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ to make the following conclusions. 
%If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue. 
%If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion. 
%In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA. We highlighted and color coded these areas in our plots. We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.

%***DAN: Some version of the below text ought to go in the figure caption, do that later, Dan
%The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space. If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space. If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.

%Ideas while driving to pick up Max:
%1) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell. I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
%2) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing. However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model. Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way. Threshold relationships can produce tail association, as we showed elsewhere. Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.

\section{Results}\label{section:results}

%G2

%Para on the main goal, whether the ATA contribution is meaningfully large
%ATA effects tend to oppose "the rest" of storage effects

%Para on how it's usually negative, whereas \Delta_[EC] is positive, so except when \delta=1, ATA effects tend to weaken storage effects
%Note that when mu1-mu2=2, we are in the case where only storage effects allow for coexistence

%Para on when delta=1 (annual species) - in this case, coexistence is only possible when mu1-mu2 is close to 0, and in that case the only
%thing allowing for coexistence at all is storage effects. Interetingly (and this is the topic sentence), "the rest" of storage effects are
%negative, and it's the ATA component of storage effects which are allowing the species to coexist. Suggests the possibility that for 
%annual species whcih are close competitors, ATAs might be especially important? Jasmin ought to check on this to make sure it's accurate.
%What I'm suspicious about is that for annual species which are close competitors (say, mu1-mu2=0), and in the symmetric case, storage
%effects in the lotery model actually don't opperate, and there is no coexistence. That sounds weird to me. Can we get some independent
%verification of that? Why would annual species be different in this respect? There must be some delta between 0.8 and 1 where \Delta^[EC]
%switches from positive to negative.

%***DAN: Jasmin, you computed standard errors of everything, but they are not currently reported. The best way to handle this would
%be if we could say in captions in appropriate places "standard errors were always less than xyz", where xyz is some extremely 
%small number. Standard errors are smalled for larger

%from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 3). When species are the same, $\mu_1 - \mu_2=0$, only storage effect mechanisms contribute to IGR and other mechanisms (in black) stay at zero. When $\delta=1$, other mechanisms stay relatively more stable with $\Delta_i^0$ set at a negative value and other non-storage effect values being at zero or extremely close to zero. Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases. The contribution of ATA captured by $\Delta_i^{[E \nat C]}$ (red) is not small. In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$. We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.

%G3

While the contribution of ATA is definitively not absent, in most cases it does not make a difference for coexistence or comepetitive exclusion. In most panels of Figure 3, the magnitude of negative contribution from ATA does not effect the sign of the IGR. The absense of ATA contribution will increase IGR in those cases, but the outcome of coexistence would likely not be different. %maybe move or expand more on this point in discussion 
%cite sources that say the sign of IGR matter more than the magnitude 
However, there are cases where ATA makes a difference for coexistence or competitive exclusion where the maginitude of $\Delta_i^{[E \nat C]}$ is large and other mechanisms do not counteract resulting in a sign change for IGR. We call these cases ATA rescue or ATA exclusion, the exploration of which is at the core of G3.

To acheive G3, we assessed the effect of ATA by comparing and plotting (Figure 4) the IGR computed for G2 and the IGR if there were no ATA (IGR - $\Delta_i^{[E \nat C]}$). To find areas of ATA exclusion or ATA rescue we looked at where these line cross y=0 and highlighted and color coded ATA exclusion trait space and ATA rescue trait space (see Methods). In these spaces, contribution of ATA is the mechanism soley responsible for a positive or negative IGR. Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical. Other $\delta$ values are more indicative of ATA exclusion events.  When $\delta<1$ ATA exlusion space increase in range and magnitude as $\sigma$ increases. There is indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however changes in range have a more complex trend. Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$. Here, the range is limited by an increasing $\Delta_i^{[E \nat C]}$ as trait space increases and the two IGR's are close to converging.

\section{Discussion}\label{section:discussion}

Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.

Need a para about multi-species models - how would things change? One difference is there are more different
things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.

Might be interesting to study how tail associations of environmental variables are changing and how this might 
alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
research program which presumably would start with a more detailed study of how tail associations of environmental
variables may be changing under global change.

Search for comments in the rest of the text indicating things that need to be done in the Discussion

Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
the importance of ATA for coexistence mediated by these factors.

Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
how to say something about how future work may be able to explore the importance of ATA in that context?

Also need to talk about how this stuff will apply to data.

In the Discussion, I intend to talk about how possible changes in tail association 
patterns of env vars in nature may influence coexistence of species. It won't influence the 
other mechanisms, though changes in the extreme values of environmental variables, known
to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
%to also look at the other mechanisms. May still need a sup mat laying out all the math.

\clearpage
\newpage

%you store your bibliographic info in refs.bib, see that file
\bibliographystyle{ecology_letters2}
\bibliography{refs}

\clearpage
\newpage

\begin{figure}
%\includegraphics[width=\textwidth]{"../results_figs/fig1.pdf"}
\caption{
%Figure displaying the noise. Three panels, left-, symmetric, and right-tail assoc noise. Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
Simple example of bivariate noise with standard normal marginals. A) is left-tail associated, B) is symmetric, and C) is right-tail associated. All three cases have the same Pearson correlation up to sampling variation, yet show different symmetry of tail associaton. Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution. The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
\label{fig:noise}}
\end{figure}

\begin{figure}
%\includegraphics[width=\textwidth]{"../results_figs/figure2.pdf"}
\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example. This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet. So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive. You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes. We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
Populations inlfuenced by asymmetric tail association recover from low densities less often/ spend more time on average under one-species dominance. The top panel, A, is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise defined by $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$. The bottom panel, B, is like A, but the population is influenced by bivariate noise with left-tailed asymmetric association defined by the same parameters. Both simulations were ran for total time of \Sexpr{M}. Noticeable population density was defined at a threshold of 0.95 of total community space. Periods of noticable coexistence are marked in grey. The community in A had mean noticable coexistence time of \Sexpr{fig2$coSym} and a mean one-species dominance time of \Sexpr{fig2$domSym}. The community in B had mean noticable coexistence time of \Sexpr{fig2$coATA} and a mean one-species dominance time of \Sexpr{fig2$domATA}. While mean periods of coexistence are more similar, community B had much longer mean periods of one-species dominace. The population in B also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.     
\label{fig:pedagog}}
\end{figure}

\begin{figure}
%\includegraphics[width=\textwidth]{"../results_figs/fig3.pdf"}
\caption{The contribution of ATA is non-negligible. All panels show the breakdown of contributing mechanisms of coexistence in accordance to \cite{Ellner_2019} new decomposition method, against increasing $\sigma$. Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$. The cases where $\delta=1$ are not shown because they are mathematically straighforward to see that IGR $=\mu_1 - \mu_2$ and is influenced by $\sigma$ (see SupMat). In most cases, the contribution of ATA (red) is negatively; however when $\delta=1$, ATA contribution is positive. The contribution of ATA is non-negligible and in special cases result in a sign change of IGR (orange). Standard error was computed but not displayed because the maximum standard error of all computations was \Sexpr{round(se3,5)}. 
\label{fig:fig3}}
\end{figure}

\begin{figure}
%\includegraphics[width=\textwidth]{"../results_figs/fig4.pdf"}
\caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 
IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference 
$\mu_1-\mu_2$ in order to demonstrate ATA rescue and ATA exclusion. The red line, $\Delta_i^{[E \nat C]}$, 
%***DAN: I would normally have used \nat instead of ||, above, but \nat seems to not work in figure captions for some reason,
%it crashes the latex on my machine at least. If we can fix this, we should, but it's not the end of the world if we can't.
%Jasmin, could you compile on your machine when || above was \nat?
is the difference between the two IGR's. Each panel is defined by a unique set of parameters $\sigma$ 
and $\delta$. ATA rescue is marked by regions in the panels where IGR wihtout ATA contribution is less 
than IGR at zero, denoted by regions of dotted yellow. ATA exclusion is marked by regions where IGR 
without ATA contribution is greater than IGR at zero, denoted by regions of dotted pink. ATA rescue 
is only seen in cases where $\delta=1$ (left column). All other cases shown here have subsets of 
species trait difference space where ATA exclusion occurs. Despite panels in the top row not 
containing the color coded region, these cases still contain ATA exlusion spaces; however, the space 
is too small to appropriately denote while still maintaining consistent x-axes across all panels. 
For $\delta<1$, ATA exclusion space increases down rows as $\sigma$ increases. ATA exclusion space also 
seems to increase across columns as $\delta$ decreases; however this trend is broken in the last row, 
indicating a more compex relationship between ATA exclusion space, $\delta$, and $\sigma$. Standard error was computed but not displayed because the maximum standard error of all computations was \Sexpr{round(se4,5)}.  
}\label{fig:fig4}
\end{figure}

\end{document}

### 2021-08-27 paper/SupMat.Rnw
cp[ying files

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}
\usepackage{mathrsfs}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{accents}

%new commands and so on
\DeclareRobustCommand{\bbone}{\text{\usefont{U}{bbold}{m}{n}1}}

\DeclareMathOperator{\EX}{E}% expected value
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\VarX}{var}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\CovX}{cov}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\Prob}{P}

\newcommand{\sC}{\mathscr{C}} %for a script C like Ellner et al (2016), Ecol Lett
\newcommand{\bs}{\backslash}

%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decide not to use)
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\olep}{\overline{\epsilon}}

%external documents
\externaldocument[MT-]{Paper}

%header material for paper
\title{Relationships in the extremes and their influence on competition and coexistence: Supporting information}
\date{}

\author[a]{Pimsupa Jasmin Albert}
\author[a,b]{Daniel C. Reuman}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}

%***Need to indicate corresponding authors

\begin{document}
%\SweaveOpts{concordance=TRUE}

%The following is where you load in the numeric results that will be embedded in the text
%<<eval=T,echo=F,message=F,warning=F>>=
%system(paste("biber", sub("\\.Rnw$", "", current_input())))

%#place R code here for loading in necessary variables from Results
%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
%allregres<-matrix(rnorm(10),2,5)
%@

\maketitle

%probably want table of contents, figure and table lists

\section{Scaling factors}\label{SIsect:scaling_factors}

Following \cite{Ellner_2016}, we define
\begin{equation}
\sC_j = -r_j(\overline{E}_j,C_j),
\end{equation}
and 
\begin{equation}
\sC_{j \bs i} = -r_j(\overline{E}_j,C_{j \bs i}),
\end{equation}
i.e., $\sC_{j \bs i}$ is the negative of the growth rate of species $j$ when $E_j$ is at its mean value
and when $i$ is at negligible abundance. Then $q_{ij}$ is defined by computing
\begin{equation}
\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}},
\end{equation}
and evaluating it at $\sC_{j \bs i}=0$ [see \cite{Chesson_1994} and \cite{Ellner_2016} for background
on this definition]. To make sense of this definition, it must be possible to write $\sC_{i \bs i}$, explicitly 
or implicitly, uniquely as a function of $\sC_{j \bs i}$ \citep{Ellner_2016}.

For the lottery model, for $i \neq j$, we have
\begin{align}
\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
\end{align}
Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it calculate $q_{ij}$,
but we instead compute $q_{ij}$ through implicit differentiation:
\begin{align}
\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
\end{align}
Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
\begin{align}
q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
\end{align}
as stated in the main text.

%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor;
%Steve reads this and thinks it makes sense to change it because something is factually or presentationally 
%sub-optimal; or Dan or Jasmin actually develops a more rigorous understanding of scaling factors and based
%on that thinks something here is sub-optimal. Anyway, the "do both" approach seems like it can't go wrong,
%especially if Steve's prediction is correct that it won't make much difference.
%***DAN: You probably need to spend a bit more time with the cited papers to better understand this. 
We also use $q_{ij}=1$ as an alternative value, based on communication with Steve Ellner, for the following reasons.
First, as discussed in the supporting information of \cite{Ellner_2016}, and as further elaborated by 
\cite{Barabas_2018}, it takes very special circumstances for the scaling factors to exist and be unique, 
especially with size-structured models. Second, in empirically parameterized models, scaling factor values are 
often not robust to arbitrary modelling choices. 
Although the lottery model is not size structured and we do not empirically parameterize it, the above shortcomings of 
scaling factors, generally, suggest we should find alternatives to their use. The alternative 
suggested by Ellner for empirical applications is to do a series of pairwise invader-resident comparisons, each with an 
equal weighting of invader and resident $r$ values; for our context this amounts to setting scaling factors to $1$. 
We consider both approaches because the above-mentioned shortcomings do not apply to our case, but it seems 
possible that future applications will often set scaling factors to $1$, and we want our ideas to also be considered 
by researchers making that choice.

\section{Mechanisms other than storage effects}\label{SIsect:other_mechanisms}

We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
original decompositions of \cite{Chesson_1994}; and then we include our extensions of the decomposition.

We begin by decomposing 
the invasion growth rate,
\begin{equation}
\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})].
\end{equation}
The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
\begin{align}
\epsilon_i^0 = r_i(\overline{E}_i,\overline{C}_{i \bs i}).
\end{align}
The additional contribution of variation in $E_i$ to the invasion growth rate is then
\begin{equation}
\olep_i^E = \E[r_i(E_i,\overline{C}_{i \bs i})]-\epsilon_i^0,
\end{equation}
whereas the additional contribution of variation in $C_{i \bs i}$ is 
\begin{equation}
\olep_i^C=\E[r_i(\overline{E}_i,C_{i \bs i})]-\epsilon_i^0.
\end{equation}
We then define
\begin{align}
\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0.
\end{align}
Thus we have the primary decomposition 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC}.
\end{equation}
But some of $\olep_i^{EC}$ comes from the fact that $E_i$ and $C_{i \bs i}$ are
both varying, and some comes from covariation in these two quantities. So we next separate these
contributions in a secondary decomposition of $\olep_i^{EC}$.

Let
\begin{equation}
\olep_i^{(E\shp C)}=\E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0
\end{equation}
and let
\begin{equation}
\olep_i^{(EC)} = \olep_i^{EC}-\olep_i^{(E\shp C)},
\end{equation}
so
\begin{equation}
\olep_i^{(EC)}=\E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)].
\end{equation}
The parentheses in the superscripts of $\olep_i^{(EC)}$ and $\olep_i^{(E\shp C)}$
indicate that we are talking about a term in the secondary decomposition, i.e., in the decomposition of 
$\olep_i^{EC}$. We now have
\begin{align}
\olr_{i \bs i} &= \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{EC} \\
\olep_i^{EC} &= \olep_i^{(E\shp C)}+\olep_i^{(EC)},
\end{align}
i.e.,
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C +\olep_i^{(E\shp C)}+\olep_i^{(EC)}. 
\end{equation}
The term $\olep_i^{(EC)}$ is due to covariation between $E_i$ and $C_{i \bs i}$, whereas the term 
$\olep_i^{(E\shp C)}$ is due to variation \emph{per se} in both $E_i$ and $C_{i \bs i}$, after the effects of 
covariation itself are removed. 
The term $\olep_i^{(EC)}$ relates to storage effects, as we will see.

As in the main text, we can now define the tertiary decomposition
\begin{align}
\olep_i^{(EC)} &= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \\
&= \left[ \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)] \right]+\left[ \E[r_i(E_i^\nat,C_{i \bs i}^\nat)]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \right] \\
&= \olep_i^{[E \nat C]}+\olep_i^{[EC]}.
\end{align}
The brackets in the superscripts of $\olep_i^{[EC]}$ and $\olep_i^{[E \nat C]}$ indicate that we are talking about a term in the
tertiary decomposition, i.e., in the decomposition of $\olep_i^{(EC)}$. 
The term $\olep_i^{[E \nat C]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
and the the term $\olep_i^{[EC]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
The symbol $\nat$ denotes the removal of asymmetries of tail association. We therefore have 
\begin{equation}
\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[EC]}+\olep_i^{[E \nat C]},\label{eq:decomp_overall_IGR}
\end{equation}
our overall decompsition of $r_{i \bs i}$.
The same derivation also applies to provide 
\begin{equation}
\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[EC]}+\olep_j^{[E \nat C]},\label{eq:decomp_overall_rjbsi}
\end{equation}
an analogous decomposition of $\olr_{j \bs i}$.

Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$ and assuming without loss of generality that $\mu_i \leq \mu_j$ so that
species $i$ is the weaker competitor and $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ can be regarded as a coexistence metric, we have
\begin{equation}
\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[EC]}+\Delta_i^{[E\nat C]},\label{eq:decomp_overall_delta}
\end{equation}
where $\Delta_i^0=\epsilon_i^0- q_{ij} \epsilon_j^0$;
$\Delta_i^E=\olep_i^E -q_{ij}\olep_j^E$ is the contribution to coexistence of environmental variation;
$\Delta_i^C=\olep_i^C -q_{ij}\olep_j^C$ is the contribution to coexistence of variation in competitive pressure;
$\Delta_i^{(E\shp C)}=\olep_i^{(E\shp C)} - q_{ij}\olep_j^{(E\shp C)}$ is the contribution to coexistence of 
variation \emph{per se} in both environment and competition, not including the effects of covariation of these quantities;
$\Delta_i^{[EC]}=\olep_i^{[EC]} - q_{ij} \olep_j^{[EC]}$ is the contribution to coexistence of $EC$ covariation \emph{per se},
not including the effects of asymmetric tail associations between environment and competition;
and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ is the contribution to coexistence of 
asymmetric tail assocations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.

\section{Noise} \label{sect:noise}

We created two cases for $(b_1,b_2)$, left- and right-tail association. 
To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N\left( \vec{0}, \Sigma \right)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. 
Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -|a_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$. 
We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. 
Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value. 
For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values. Right-tail associated noise was generated in an analogous way. 
Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
noise was generated from the left- or right-tail association cases.

When needed, noise with symmetric tail association was generated by taking a very large sample (100000) 
%***DAN: Dan, later, check whether 100000 is what was actually used. I'm thinking we could probably do a million.
of left-tail associated noise generated as 
described above, and then calculating the Pearson correaltion coefficient, $\rho$, of the sample.
Symmetric noise $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$ was then generated 
from the bivariate normal distribution $N\left(\vec{0}, \Sigma \right)$, where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, 
and then we set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.

\section{Mathematical simplifications for plotting}\label{sect:math_simplifications}

We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the lottery model. 
One can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that
\begin{equation}
E_i=B_i=\exp(b_i)
\end{equation}
\begin{equation}
C_i=\frac{B_1 N_1 + B_2 N_2}{\delta N}
\end{equation}
\begin{equation}
C_{i \bs i} = B_j/\delta=\exp(b_j)/\delta
\end{equation}
\begin{equation}
r_i=\ln(1-\delta+E_i/C_i)
\end{equation}
%IGR
\begin{align}
\olr_{i \bs i} &= \E[\ln(1-\delta+E_i/C_{i \bs i})] \\
&= \E[\ln(1-\delta+\delta B_i/B_j)] \\
&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]\label{eq:base_IGR}
\end{align}
\begin{equation}
\overline{E}_i = \exp(\mu_i + \sigma^2/2)
\end{equation}
\begin{equation}
\overline{C}_{i \bs i}=\exp(\mu_j+\sigma^2/2)/\delta
\end{equation}
%baseline
\begin{equation}
\epsilon_i^0 = \ln(1-\delta+\delta \exp(\mu_i - \mu_j))\label{eq:decomp:baseline}
\end{equation}
%contrib of variation in E
and
\begin{align}
\olep_i^E &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i+\mu_i)/\exp(\mu_j+\sigma^2/2))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}
\end{align}
where $\tilde{b}_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%contrib of variation in C
\begin{align}
\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(b_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma \tilde{b}_j +\mu_j))]-\epsilon_i^0 \\
&= \E[\ln(1-\delta + \delta \exp(-\sigma \tilde{b}_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 
\end{align}
where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
\begin{align}
\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(b_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i +\mu_i)/\exp(\sigma \tilde{b}_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}
\end{align}
where $(\tilde{b}_i, \tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$. Continuing,
%contrib of variation in E and C, per se  
\begin{align}
\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(b_i^\shp)/\exp(b_j^\shp))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i^\shp +\mu_i)/\exp(\sigma \tilde{b}_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
&= \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}
\end{align}
where $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
%storage effects
\begin{equation}
\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}
\end{equation}
where $(\tilde{b}_i,\tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$, and 
$\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normal and independent. Continuing,
%contrib of ATA
\begin{equation}
\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
\end{equation}
where $(\ddot{b}_i,\ddot{b}_j)$ have the same (normal) marginals and covariance as $(b_i,b_j)$ but are 
bivariate normal. Thus
\begin{equation}
\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
\end{equation}
where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
from expression (\ref{eq:decomp:storage_effects}), i.e., 
\begin{equation}
\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
\end{equation}
This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.

We next decompose $\olr_{j \bs i}$ for the lottery model. We have 
\begin{equation}
E_j = B_j = \exp(b_j)
\end{equation}
\begin{equation}
C_j=\frac{B_1 N_1 + B_2 N_2}{\delta N}
\end{equation}
\begin{equation}
C_{j \bs i} = B_j/\delta = \exp(b_j)/\delta
\end{equation}
\begin{equation}
r_j=\ln(1-\delta+E_j/C_j)
\end{equation}
\begin{align}
\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
&= 0
\end{align}
\begin{equation}
\overline{E}_j = \exp(\mu_j+\sigma^2/2)
\end{equation}
\begin{equation}
\overline{C}_{j \bs i} = \exp(\mu_j + \sigma^2/2)/\delta
\end{equation}
%baseline
\begin{align}
\epsilon_j^0 &= r_j(\overline{E}_j,\overline{C}_{j \bs i}) \\
&= \ln \left( 1-\delta+\delta \frac{\exp(\mu_j + \sigma^2/2)}{\exp(\mu_j + \sigma^2/2)} \right) \\
&= 0 \label{eq:decomp:baseline_j}
\end{align}
and
%contrib of variation in E
\begin{align}
\olep_j^E &= \E[r_j(E_j,\overline{C}_{j \bs i})]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta B_j/\overline{B}_j)]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma \tilde{b}_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma \tilde{b}_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}
\end{align}
where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%contrib of variation in C
\begin{align}
\olep_j^C &= \E[r_j(\overline{E}_j,C_{j \bs i})]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\overline{B}_j}{B_j} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(b_j)} \right) \right]-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma \tilde{b}_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(-\sigma \tilde{b}_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}
\end{align}
where again $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
%last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
\begin{align}
\olep_j^{EC} &= \E[r_j(E_j,C_{j \bs i})]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta B_j/B_j)]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= -\olep_j^E-\olep_j^C-\epsilon_j^0.\label{eq:decomp:j:EC}
\end{align}
Continuing,
%contrib of variation in E and C, per se  
\begin{align}
\olep_j^{(E\shp C)} &= \E[r_j(E_j^\shp , C_{j \bs i}^\shp )]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{B_j^\shp}{B_j} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j^\shp)}{\exp(b_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma \tilde{b}_j^\shp + \mu_j)}{\exp(\sigma \tilde{b}_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
&= \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}
\end{align}
where $(E_j^\shp , C_{j \bs i}^\shp)$ has the same marginals as $(E_j,C_{j \bs i})$ but has independent components;
$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; $b_j^\shp$ is distributed in the same
way as $b_j$ but is independent of it; and $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and 
independent of each other. Continuing,
%storage effects
\begin{align}
\olep_j^{(EC)} &= \E[r_j(E_j,C_{j \bs i})]-\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
&= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
\end{align}
these results obtained from (\ref{eq:decomp:j:EC}) and (\ref{eq:decomp:j:EshpC}). Again, $\tilde{b}_j$ and 
$\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
\begin{align}
\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
&= 0 \label{eq:decomp:j:EnatC}
\end{align}
because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
\begin{align}
\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
\end{align}
where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.

Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
$\delta$, $\sigma$, and $\mu_i - \mu_j$, but not separately on $\mu_i$ or $\mu_j$. The terms of the decomposition
of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\mu_i$ or $\mu_j$ at all.
As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
\begin{align}
q_{ij} &= \frac{\exp(\mu_i)}{(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)} \\
&= \frac{\exp(-\mu_i)\exp(\mu_i)}{\exp(-\mu_i)[(1-\delta) \exp(\mu_j)+\delta \exp(\mu_i)]} \\
&= \frac{1}{(1-\delta)\exp(-(\mu_i-\mu_j))+\delta}.\label{eq:decomp:qnot1}
\end{align}
These also either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
$\mu_i-\mu_j$. Thus each of the terms of the decomposition (\ref{MT-eq:full_decomp}) from the main text
depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.

We now want to show that, for the lottery model, 
all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
as $q_{ij}$, and therefore all elements of our decomposition (\ref{eq:decomp_overall_delta}), are the same 
for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(b_i^{(l)}),\exp(b_j^{(l)})/\delta)$ and 
$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(b_i^{(r)}),\exp(b_j^{(r)})/\delta)$
for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and likewise the marginals of 
$(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
So any component of (\ref{eq:decomp_overall_IGR}), or $q_{ij}$, will not depend on the choice
of our left- or right-tail associated noise cases as long as it depends only on marginal distributions.
But it is easy to see that $q_{ij}$ depends only on marginals, since $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$  
depends only on $\mu_i$, $\mu_j$ and $\delta$, and these are parameters of the marginal distributions of 
$(b_i,b_j)$. The alternative value $q_{ij}=1$ is constant, so also takes the same value for both our
left- and right-tail association noise cases. It is also easy to see from the definitions of $\epsilon_i^0$, 
$\olep_i^E$, $\olep_i^C$, and $\olep_i^{(E \shp C)}$ that these quantities only depend on marginals.

Whereas $\olep_i^{[EC]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
cases because the bivariate random variables $(E_i^\nat , C_{i \bs i}^\nat )$ and 
$(E_i^\shp ,C_{i \bs i}^\shp)$ have had their tail association removed, and the only differences 
between $(E_i^{(l)},C_{i \bs i}^{(l)})$ and $(E_i^{(r)},C_{i \bs i}^{(r)})$ are in their tail association
properties (i.e., these two bivariate distributions have the same marginals and covariances).
In other words, $(E_i^\nat , C_{i \bs i}^\nat )$ is distributed in the same way for the left- and right-tail-association cases,
as is $(E_i^\shp ,C_{i \bs i}^\shp)$.

We now examine $\olep_i^{[E \nat C ]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
\begin{equation}
\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
\end{equation}
But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
of the choice between our left- and right-tail-associated noise cases. Now examining 
$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
$(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
$(b_1^{(l)},b_2^{(l)})$. But then we have 
\begin{align}
\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
\end{align}
which completes the proof that $\olep_i^{[E \nat C]}$ is independent of left- versus right-tail associated noise.
For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 
above, in this section, and note that all those terms depend only on marginals.

\section{Efficient computation}\label{sect:efficient_computation}

We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the lottery model.

First, for a large integer $M$ (we used $M=$),
%***DAN: Jasmin, please fill in the value, 1000000 or bigger if you can swing it
carry out the following steps once.
\begin{enumerate}
\item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
\item Generate $M$ independent points $(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$, as in SI section \ref{sect:noise}, which are left-tail associated
and which have standard-normal marginals. 
\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(\tilde{\ddot{b}}_i^{(k)},\tilde{\ddot{b}}_j^{(k)})$, 
$k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
$(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$.
\end{enumerate}
These data are stored and used repeatedly below. This provides computational efficiency because 
random number generation is computationally expensive.

Next, given values of the parameters $\delta$, $\mu_i - \mu_j$, and $\sigma$, proceed as follows to get estimates of the 
terms of the decomposition (\ref{eq:decomp_overall_IGR}).
\begin{enumerate}
\item Compute $\epsilon_i^0$ following (\ref{eq:decomp:baseline}), which is a simple formula of $\delta$ and 
$\mu_i - \mu_j$, not requiring use of the random samples described above. This gives an exact value of
$\epsilon_i^0$, not an estimate.
\item Estimate $\olep_i^E$, following (\ref{eq:decomp:contrib_of_var_E}), as 
$\widehat{\olep_i^E}=\mean_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0$, an estimate with 
standard error $\se \left( \widehat{\olep_i^E} \right) = \sd_k [\ln(1-\delta+\delta \exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]/\sqrt{M}$,
where $\sd(\cdot)$ denotes sample standard deviation.
\item Estimate $\olep_i^C$, following (\ref{eq:decomp:contrib_of_var_C}), as 
$\widehat{\olep_i^C}=\mean_k [\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-\epsilon_i^0$, which has 
standard error $\se \left( \widehat{\olep_i^C} \right) = \sd_k[\ln(1-\delta+\delta \exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_i^\shp-\tilde{b}_j^\shp$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) is normally distributed
with mean $0$ and variance $2$, so we can estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:decomp:contrib_of_EC_var_per_se}),
as $\widehat{\olep_i^{(E \shp C)}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0.$ The standard error of this expression is 
$\se \left( \widehat{\olep_i^{(E \shp C)}} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma \sqrt{2} u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_EC}), as
$\widehat{\olep_i^{[EC]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$. The standard error of this expression is 
$\se \left( \widehat{\olep_i^{[EC]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
$\widehat{\olep_i^{[E \nat C]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
error of this expression is $\se \left( \widehat{\olep_i^{[E \nat C]}} \right)=\sqrt{s_3^2+s_1^2}$, where
$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
which has standard error $s_3$ from above.
\end{enumerate}
It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
\widehat{\olep_i^{[EC]}}+\widehat{\olep_i^{[E \nat C]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
at the level of the estimators, as well

Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
\begin{enumerate}
\item Use $\epsilon_j^0=0$, as in (\ref{eq:decomp:baseline_j}), and this is the exact value.
\item Estimate $\olep_j^E$, following (\ref{eq:decomp:contrib_of_var_E_j}), as
$\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]-\epsilon_j^0$, an estimate
with standard error $\se \left( \widehat{\olep_j^E} \right) = \sd_k [\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^C$, following (\ref{eq:decomp:contrib_of_var_C_j}), as
$\widehat{\olep_j^C}=\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]-\epsilon_j^0$, which has 
standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
$\widehat{\olep_j^{(E \shp C)}} = \mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0$.
The standard error of this expression is $\se \left(  \widehat{\olep_j^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))-\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\olep_j^{[EC]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[EC]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[EC]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
\item Use $\olep_j^{[E \nat C]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
\item Finally, use $\olr_{j \bs i} = 0$, again exact. 
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi}) applies for the 
estimates as well as for the quantities being estimated.

Next, $q_{ij}$ can be taken (see SI section \ref{SIsect:scaling_factors}) to be, alternatively, $1$ or 
the quantity displayed in (\ref{eq:decomp:qnot1}). Hence the terms of the decomposition 
(\ref{eq:decomp_overall_delta}) can straightforwardly be etimated using our estimates elaborated 
above, as follows. There are two alternative versions of the estiamtors for this decomposition, corresponding 
to the alternative values for $q_{ij}$.
\begin{enumerate}
\item Compute $\Delta_i^0 = \epsilon_i^0 - q_{ij} \epsilon_j^0$, which is computed exactly, since 
the quantities described above for the components $\epsilon_i^0$, $\epsilon_j^0$ and $q_{ij}$ are exact, as opposed to being
estimates. Simplifying, since $\epsilon_j^0=0$, we have $\Delta_i^0 = \epsilon_i^0$.
\item Estimate $\Delta_i^E$ as $\widehat{\Delta_i^E}=\widehat{\olep_i^E}-q_{ij}\widehat{\olep_j^E}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))]-\epsilon_i^0-
q_{ij} \mean_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))],$ which has standard error 
$\se\left( \widehat{\Delta_i^E} \right)=\sd_k[\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
q_{ij} \ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))]/\sqrt{M}.$
\item Estimate $\Delta_i^C$ as $\widehat{\Delta_i^C}=\widehat{\olep_i^C}-q_{ij}\widehat{\olep_j^C}=
\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]-
\epsilon_i^0-q_{ij}\mean_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))],$ which has standard error
$\se \left( \widehat{\Delta_i^C} \right)=\sd_k[\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\Delta_i^{(E \shp C)}$ as $\widehat{\Delta_i^{(E \shp C)}} = \widehat{\olep_i^{(E \shp C)}}-q_{ij} \widehat{\olep_j^{(E \shp C)}}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0-
q_{ij}[\mean_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\olep_j^E}-\widehat{\olep_j^C}-\epsilon_j^0].$
This simplifies to $\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]-
q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]-\widehat{\Delta_i^E}-\widehat{\Delta_i^C}-\Delta_i^0$, 
which can be straightforwardly shown has standard error
$\se \left( \widehat{\Delta_i^{(E \shp C)}} \right) = 
\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-
\ln(1-\delta+\delta\exp(\sigma u^{(k)}+\mu_i-\mu_j-\sigma^2/2))-
\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))-
q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))+
q_{ij}\ln(1-\delta+\delta\exp(\sigma u^{(k)}-\sigma^2/2))+
q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
\item Estimate $\Delta_i^{[EC]}$ as $\widehat{\Delta_i^{[EC]}}=\widehat{\olep_i^{[EC]}}-q_{ij}\widehat{\olep_j^{[EC]}}=
\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-
\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]+
q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))],$ which one can straightforwardly show has standard error
$\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$ 
and $s_2=\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]/\sqrt{M}$.
\item For $\Delta_i^{[E \nat C]}$, we recall that $\olep_j^{[E \nat C]}=0$, so $\Delta_i^{[E \nat C]}=\olep_i^{[E \nat C]}$. So we use
$\widehat{\olep_i^{[E \nat C]}}$ for $\widehat{\Delta_i^{[E \nat C]}}$.
\item Finally, we can also estimate $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ simply as $\widehat{\olr_{i \bs i}}$
since $\olr_{j \bs i}=0$.
\end{enumerate}
As for (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), it is again the case that (\ref{eq:decomp_overall_delta})
applies for the estimators as well as for the quantities being estimated.

%***DAN: Most but not all of the below is obsolete now, replaced and upgraded by stuff above this point

\section{Invasion growth rate of species 1, and storage effects} \label{sect:IGR}

We consider the invasion growth rate of species 1, $\bar r_1=\EX(\ln (\frac{N_1(t+1)}{N_1(t)}))$, when $N_1$ is close to 0 and $N_2$ is close to $N$.
Hence $\EX(\cdot)$ represents expected value.
We have

\begin{equation}
\frac{N_1(t+1)}{N_1(t)}=(1-\delta)+\delta N \frac{B_1(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)},  \label{sp1_Nt+1_over_Nt}
\end{equation}

\noindent but setting $N_1=0$ and $N_2=N$ here gives the result that

\begin{equation}
\bar r_1=\EX(\ln [(1-\delta)+\delta \exp(b_1-b_2)]).  \label{rbar1_E}
\end{equation}

\noindent We considered the various cases for the joint distribution $(b_1,b_2)$, specified in the main text and in the previous section.

To study storage effect for the model, we also consider the contribution of EC covariance to the difference between the invader's and the resident's mean growth rates, to be denoted $\Delta I$. 
Following REF, we define

\begin{equation}
\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp (b_1^\#-b_2)]),  \label{rbar1sharp_E}
\end{equation}

\noindent where $b_1^\#$ is distributed in the same way as $b_1$, but is independent of it and $b_2$. 
Then $\bar r_1 - \bar r_1^\#$ is the contribution of EC covariance to the mean growth rate species 1, the invader. 
We know, a priori, that the mean growth rate of the resident, species 2, must be 0, $\bar r_2 =0$, but equation (\ref{model_eq}) also implies that

\begin{equation}
\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta N \frac{B_2(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  \label{sp2_Nt+1_over_Nt}
\end{equation}

\noindent Setting $N_2=N$ and $N_1=0$ then gives

\begin{equation}
\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta=1  \label{sp2_Nt+1_over_Nt_simp}
\end{equation}

\noindent so

\begin{equation}
\EX(\ln (\frac{N_2(t+1)}{N_2(t)}))=0, \label{r2_is_zero}
\end{equation}

\noindent as expected. 
Again, following REF, we also define

\begin{equation}
\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(b_2^\#-b_2)]),  \label{rbar2sharp_E}
\end{equation}

\noindent where $b_2^\#$ is distributed in the same way as $b_2$, but is independent of it and of $b_1$. 
Thus $\bar r_2- \bar r_2^\# = -\bar r_2^\#$ is the contribution of EC covariance to the mean growth rate of species 2, the resident. 
And so

\begin{equation}
\Delta I=\left(\bar r_1 - \bar r_1^\#\right)-\left(\bar r_2- \bar r_2^\#\right)=\bar r_1 - \bar r_1^\#+\bar r_2^\#  \label{DeltaI}
\end{equation}

\noindent has been defined in terms of expected values of elementary expressions of the random variables $\left(b_1,b_2\right)$, $\left(b_1^\#,b_2\right)$, and $\left(b_2^\#, b_2\right)$.

Having already specified that $b_1$ (and therefore also $b_1^\#$) is distributed as $N\left(\mu_1, \sigma^2\right)$, and that $b_2$ (and therefore also $b_2^\#$) is distributed as $N\left(\mu_2, \sigma^2\right)$, we know, using $\Sigma$ as notation for a covariance matrix that

\begin{equation}
\left(b_1^\#, b_2\right) \sim N\left((\mu_1,\mu_2),\Sigma \right) \label{b1sharpb2_distribution}
\end{equation}

\noindent and

\begin{equation}
\left(b_2^\#, b_2\right) \sim N(( \mu_2,\mu_2), \Sigma) \label{b2sharpb2_distribution}
\end{equation}

\noindent where

\begin{equation}
\Sigma = \begin{pmatrix} \sigma^2&0\\0&\sigma^2\end{pmatrix}. \label{Sigmamat_cov0}
\end{equation}

\noindent Thus the difference $b_1^\#-b_2$ that occurs in the expression for $\bar r_1^\#$ (\ref{rbar1sharp_E}) is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance $2\sigma^2$. 
Thus

\begin{equation}
\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar1sharp_E_usub}
\end{equation}

\noindent where $u \sim N(\mu_1-\mu_2,2\sigma^2)$. 
Likewise, the difference $b_2^\#-b_2$ in (\ref{rbar2sharp_E}) is a normally distributed random variable with mean $\mu_2-\mu_2=0$ and variance $2\sigma^2$, so

\begin{equation}
\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar2sharp_E_usub}
\end{equation}

\noindent where now $u \sim N(0,2\sigma^2)$. 
The expression (\ref{rbar1_E}) for $\bar r_1$ can also be further simplified, in a similar way, when we are in the symmetric tail association case considered in section \ref{sect:noise}. 
In that case,

\begin{equation}
(b_1,b_2) \sim  N(( \mu_1,\mu_2), \Sigma) \label{b1b2_distribution_sym}
\end{equation}

\noindent where now

\begin{equation}
\Sigma = \begin{pmatrix} \sigma^2&\rho\\\rho&\sigma^2\end{pmatrix} \label{Sigmamat_rho}
\end{equation}

\noindent and $rho=put in value$ was determined as in section \ref{sect:noise}.
Therefore, the expression $b_1-b_2$ that occurred in \ref{rbar1_E}, in the symmetric tail association case, is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance

\begin{align}
\VarX(b_1-b_2)&=\CovX(b_1-b_2, b_1-b_2) \\
                        &=\CovX(b_1,b_1)+\CovX(b_2,b_2)-\CovX(b_1,b_2)-\CovX(b_2,b_1) \\
                        &=2\sigma^2-2\rho.
\end{align}

\noindent Thus, again only in the symmetric tail association case,

\begin{equation}
\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), \label{rbar1_E_usub}
\end{equation}

\noindent where here $u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$. 
As for all we know there is no simplicifaction of the expression \ref{rbar1_E}  for $\bar r_1$ in the left and right-tail association cases.

Thus for in summary,

\begin{equation}
\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho), \label{sym_rbar1}
\end{equation}

\noindent for the symmetric case, and

\begin{equation}
\bar r_1^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2), \label{sym_rbar1sharp}
\end{equation}

\begin{equation}
\bar r_2^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(0, 2\sigma^2). \label{sym_rbar2sharp}
\end{equation}

\noindent We next use the expressions to develop methods to estimate $\bar r_1$ and $\Delta I$ for any given paramters $\delta,\mu_1,\mu_2,\sigma$.

\section{Estimating $\bar r_1$ and $\Delta I$}

First, for a large integer $M$, carry out the following steps once:

\begin{enumerate}
\item Generate left-tail assocated noise $\left(\tilde b_{l1}^{(i)}, \tilde b_{l2}^{(i)}\right)$, $i=1,...,M$, as described in the section \ref{sect:noise}. Hence $\tilde b_{l1}^{i)}$ and $\tilde b_{l2}^{(i)}$ are distributed as $N(0,1)$. 
\item Generate right-tail associated noise $\left(\tilde b_{r1}^{(i)}, \tilde b_{r2}^{(i)}\right)$, $i=1,...,M$, via the analogous procedure. Again, $\tilde b_{r1}^{i)}$ and $\tilde b_{r2}^{(i)}$ are distributed as $N(0,1)$. 
\item Generate $M$ points $\tilde u_i$, $i = 1,...,M$ from a standard normal distribution. 
\end{enumerate}

\noindent Next, given values of parameters $\delta, \mu_1, \mu_2$, and $\sigma$, proceed as follows to get estimates of $\bar r_1, \bar r_1^{\#}, \bar r_2^{\#}$, and $\Delta I$.

\begin{enumerate}

\item Estimate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.

\item Estiate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$. The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.

\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat{\bar r_1}=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$. The standard error of this estimation is $\se(\hat{\bar r_1}) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.

\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_1^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$. The standard error is $\se(\hat{\bar r_1^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_2^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$. The standard error is $\se(\hat{\bar r_2^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.

\item $\hat {\Delta I} = \hat{\bar r_1} - \hat{\bar r_1^{\#}} + \hat{\bar r_2^{\#}}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases. We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat{\bar r_1}, \hat{\bar r_1^{\#}}$, and $\hat{\bar r_2^{\#}}$.

\end{enumerate}

\noindent For large enough $M$, standard error should be very small. Since random variables are generated once only, computation of these quantitaties for a large number of parameter sets should be fast.

\section{Additional quantities influencing invasion success} \label{sect:addquant}

Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspects of the distribution of
\begin{equation}
r_1 = \ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1_distribution}
\end{equation}

We here elaborate how some other aspects of the distribution of $r_1$ were quantified, for our three tail association cases and for any of our parameters $\delta, \mu_1, \mu_2, \sigma$.

First, we considered $\sd(r_1)$, the standard deviation. If this is large enough, then $r_1$ can occasionally be positive even if $\bar r_1$ is strongly negative, and this invasion can occur under the right environmental circumstances.

We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
\begin{align}
\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2) > 1] \\
&= \Prob[\exp (b_1-b_2)>1] \\
&= \Prob[b_1>b_2],
\end{align}

\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.

We also consider the mean value of the possible part of the distribution of $r_1$, $\EX(r_1|r_1>0)$. 
This characterizes how quickly invasion may happen, when it happens, with larger values corresponding to faster invasion. 
If it takes several time steps for the population of the invader to rise from negligible level at which species 1 is a noticeable presence in the community, then environmental conditions must be suitable for invasion for several sequential time steps. 
Thus large values of $\EX(r_1|r_1>0)$ should also correspond to a greater probability of invasion because they should correspond to faster invasions and therefore to reduced need for sequential periods of suitable environmental conditions.

Finally, for a few choices of parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.

\section{Symmetries and parameter reduction}

The quantities we consider to address invasion prospects for the weaker competitor, species 1 (see sections \ref{sect:IGR} and \ref{sect:addquant}) depend only on $\mu_1-\mu_2$ and not, independtly, on $\mu_1$ and $\mu_2$, and also are the same for our left- and right-tailed association cases. We prove these statements in this section.

Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all. 
Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases.
But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get

\begin{align}
\bar r_1 &= \EX(\ln[1-\delta+\delta \exp(\beta_1-\eta-(\beta_2-\eta))])\\
&= \EX(\ln[1-\delta+\delta \exp(\beta_1-\beta_2)]).
\end{align}
\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$, and thus $\bar r_1$ in the asymmetric tail associations also depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.

In section \ref{sect:addquant}, we introduce several other metrics. These are all based on the distribution 
\begin{equation}
r_1=\ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1distribution}
\end{equation}

\noindent But, again making the substitutions $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$, we can see via the same resoning as above that the whole distribution $r_1$ depends only on $\mu_1 - \mu_2$, and not on $\mu_1$ and $\mu_2$ independently, for all three of our tail association cases.

To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$. 
Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. 
It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed.
From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. 
Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in \ref{r1distribution}, we get

\begin{align}
r_1 &= \ln[1-\delta+\delta \exp(-\beta_2+\mu_1+\mu_2-(\beta_1+\mu_1+\mu_2))]\\
&= \ln[1-\delta+\delta \exp(\beta_1-\beta_2)].
\end{align}

\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ (\ref{r1distribution}) are the same for left- and right-tail associated noise.

\section{Decomposition}\label{Decomposition}

To quantify the contribution of different drivers to the difference in growth rate between the invader and the resident, the amount variance due to that driver is considered. We consider the contributions of the null growth rate, $\epsilon^0$, the contribution of variance in E (environment), $\bar \epsilon^E$, the contribution of the variance in C (competitive pressure), $\bar \epsilon^C$, the contrubition of E and C co-varying $\bar \epsilon^{(EC)}$, and the the contribtion of E and C varying, but independently, $\bar \epsilon^{(E\#C)}$, for the invader and the resident. The difference between invader and resident is computed term by term to get a full decomposition as follows,

\begin{equation}
\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label{IGR}
\end{equation}

\noindent where each $\Delta$ denotes a drivers contribution the the invader growth rate subtracted by the drivers contribution to the resident's groeth rate.

The null growth rate is computed as the growrth rate when both drivers (E and C) are set at their mean. Recalling the growth rate eqaution from equation \ref{rbar1_E},

\begin{align}
\Delta^0 &= \epsilon_1^0 - \epsilon_2^0\\
&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - \ln(1-\delta + \delta \frac{\bar B_2}{\bar B_2})\\
&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - 0. \\
%&=\ln(1-\delta + \delta \exp(\bar b_1-\bar b_2)
\end{align}

The contribution of variance in E to a species growth rate is computed by letting E vary and setting C to its mean, subtracted by the null. The contribution of variance in C to a species' growth rate is computed in an analogous way, but reverse.

\begin{align}
\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
&=\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
\end{align}

\begin{align}
\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
&=\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
\end{align}

The contribution of E and C both varying but not together is computed by letting both terms vary but decoupling any association by substituting in a new independent varibable for one of the varibales. Here we use C. The term is subtracted by the previous contributions.

\begin{align}
\Delta_1^{(E\#C)} &= \bar \epsilon_1^{(E\#C)} - \bar \epsilon_2^{(E\#C)}\\
&=[\EX(\ln[1-\delta + \delta \frac{B_1}{B_2^{\#}}]) -[\epsilon_1^0 +\bar \epsilon_1^E + \bar \epsilon_1^C]] - [\EX(\ln[1-\delta + \delta \frac{B_2}{B_2^{\#}}]) -[\epsilon_2^0 +\bar \epsilon_2^E + \bar \epsilon_2^C]]\\
\end{align}

The contribution of E and C co-varying is the storage effect. Computed by,

\begin{equation}
\Delta^{(EC)} = \bar \epsilon_1^{(EC)} - \bar \epsilon_2^{(EC)}.
\end{equation}

\noindent Where,

\begin{align}
\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C].
\end{align}

By substituting these expansions into \ref{IGR}, we get,

Recall from main text that the goal of this research is to investigate the implications of extreme events and thus asymmetric tail associations on coexistence. Hence, we compare IGR and mechanism values for each noise type. Because only association structure and not means and standard deviation is differ across noise types, different noise types will not result in differences in any of the decomposition terms outside of the storage effect term. All other terms do not consider the associative relationship between the two variables. The null growth grate is a constant that depends on the mean of both variables, the contribution of variance in one variable only depends on that one variable scaled by the mean of the other, and the varying but independent term while it does consider both variable they are set to be completely uncorrelated. These four terms will not differ across noise types.

\section{new decomposition term}

A new decomposition term was proposed to quantify the effect of asymmetric associations to coexistence. Since asymmetry only differs in the storage effect term, $\bar \epsilon^{(EC)}$, the contribution of asymmetry associations between the species will be a further decomposition.

\begin{equation}
\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{( {EC_0})}
\end{equation}

The contribution of asymmetry to the storage effect is the observed storage effect subtracted by a null storage effect if the association between species was completely symmetrical. 
The descrition of $\bar \epsilon^{(\ddot {EC})}$ is given in section 3 and equations 9 and 21-23.

\begin{equation}
\bar \epsilon^{( {EC_0})} = \bar r_1 - \bar r_1^{\#} + \bar r_2^{\#},
\end{equation}

\noindent where $ \bar r_1, \bar r_1^{\#}$ and $\bar r_2^{\#}$ are in the symmetric case. In our model,

\begin{equation}
\bar \epsilon^{( {EC_0})} = \EX(\ln[1-\delta+\delta \exp(u_1)]) - \EX(\ln[1-\delta+\delta \exp(u_1^{\#})]) + \EX(\ln[1-\delta+\delta \exp(u_2^{\#})]),
\end{equation}

\noindent where $u_1 \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$, $u_1^{\#} \sim N(\mu_1-\mu_2, 2\sigma^2)$, and $u_2^{\#} \sim N(0, 2\sigma^2)$

\noindent *Note that the contribution of asymmetry to storage effects is not the same as the storage effect that we computed above for asymmetriv cases.

In generalized terms, this further decomposition could be applied to test for contributions of asymmetry between any two variables. Here we tested the asymmetry between $b_1$ and $b_2$. Why did we not look at E and C? Do they have to be normally distrubuted?

This break down would only work if the variables are normally distributed. Thus one can convert their distribution to normal to impliment this decomposition.

\clearpage
\newpage

%you store your bibliographic info in refs.bib, see that file
\bibliographystyle{ecology_letters2}
\bibliography{refs}

\end{document}

### 2021-09-01 paper/Paper.Rnw
transferred a few minor comment changes I had made but not committed to the earlier version of the repo

@@
 
+%***DAN: drop the delta =1 case from here
 %Para on when delta=1 (annual species) - in this case, coexistence is only possible when mu1-mu2 is close to 0, and in that case the only
@@
 %switches from positive to negative.
+
+%***DAN: consider possile para on mu1-mu2=0 case
+%***DAN: For discussion: is it true that ATA always weakens storage effects?
+Same as saying the ATA terms is always negative?
+The actual question is, for what models are ATA effects negative, versus positive?
 
@@
 \section{Discussion}\label{section:discussion}
+
+I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
+where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
+similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
+there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
+are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
 

### 2021-09-01 paper/Paper.Rnw
various changes to make the code load from te right place and save to the right place. sub optimal at present, discussed already with jasmin

@@
 \SweaveOpts{concordance=TRUE}
-\SweaveOpts{concordance=TRUE}
-\SweaveOpts{concordance=TRUE}
 
@@
 #allregres<-matrix(rnorm(10),2,5)
-fig2data <- load(file="../results_numeric/fig2data.rda")
-M <- load(file="../results_numeric/M.rda")
-rho <- load(file="../results_numeric/rho.rda")
-se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
-se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
+#fig2data <- load(file="../results_numeric/fig2data.rda")
+#M <- load(file="../results_numeric/M.rda")
+#rho <- load(file="../results_numeric/rho.rda")
+#se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
+#se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
 @
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{"../results_figs/fig1.pdf"}
+\includegraphics[width=\textwidth]{"../results_figs/fig1_vertical.pdf"}
 \caption{
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{"../results_figs/figure2.pdf"}
+\includegraphics[width=\textwidth]{"../results_figs/figure2.pdf"}
 \caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{"../results_figs/fig3.pdf"}
+\includegraphics[width=\textwidth]{"../results_figs/fig3.2.pdf"}
 \caption{The contribution of ATA is non-negligible.
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{"../results_figs/fig4.pdf"}
+\includegraphics[width=\textwidth]{"../results_figs/fig4.2.pdf"}
 \caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 

### 2021-09-01 code/Figure1.R

@@ -1,4 +1,4 @@
-load("results_numeric/noise_etc.RData")
+load("../results_numeric/noise_etc.RData")
 
 
 P <- c(cor(b_tilde$l, method='pearson')[1,2], cor(b_tilde$s, method='pearson')[1,2], cor(b_tilde$r, method='pearson')[1,2])
@@ -8,7 +8,7 @@ PL <- round(P[1],4)
 #col2rgb("darkblue")
 blue <- rgb(0,0,0.545,0.3)
 
-pdf("results_figs/fig1.pdf", width=7, height=3)
+pdf("../results_figs/fig1.pdf", width=7, height=3)
 #png("fig1.png", height=560, width=1600,pointsize=22)
 par(mar=c(0.5,0.5,0,0), oma=c(5,5,2,2))
 layout(matrix(c(4,10,10,6,11,11,8,12,
@@ -56,7 +56,7 @@ plot(d$y, d$x, xlab=NA, ylab=NA, sub=NA, main=NA, bty='n', type='l',xaxt='n', ya
 dev.off()
 
 
-pdf("results_figs/fig1_vertical.pdf", width=2.8, height=7)
+pdf("../results_figs/fig1_vertical.pdf", width=2.8, height=7)
 
 par(mar=c(0.5,0.5,0,0), oma=c(5,5,2,2))
 

### 2021-09-01 code/Figure2.R

@@ -7,27 +7,27 @@ source("./simsPlot_fxn.R")
 #( r w/o ATA is positive and r with ATA is negative - ATA causes r to become negative )
 
 # *** use this one for paper *** #
-source("./makenoise.R")
-pdf("./results_figs/figure2.pdf") 
+#source("./makenoise.R")
+pdf("../results_figs/figure2.pdf") 
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 fig2 <- simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
 dev.off()
-save(fig2, file="results_numeric/fig2data.rda")
+save(fig2, file="../results_numeric/fig2data.rda")
 
 
-png("./results_figs/figure2.PNG", width=1000, height=600, pointsize=15) 
+png("../results_figs/figure2.PNG", width=1000, height=600, pointsize=15) 
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
 dev.off()
 
-png("figure2.1.2.PNG", width=1000, height=600, pointsize=15)
+png("../results_figs/figure2.1.2.PNG", width=1000, height=600, pointsize=15)
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 simsPlot(mudif=-0.6, delta=0.8, sigma=4) # r_woATA = 0.0185, r = -0.0602, [E||C] = -0.0787
 dev.off()
 
 #### ---- FACILITATING ---- ####
 #( r w/o ATA is negative and r with ATA is positive - ATA causes r to become positive )
-pdf("figure2.2.pdf")
+pdf("../results_figs/figure2.2.pdf")
 #png("figure2.2.1.PNG", width=1000, height=600, pointsize=15)
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 simsPlot(mudif=0, delta=1, sigma=7) # r_woATA = -0.0071, r = 0.00036, [E||C] = 0.0074
@@ -56,7 +56,7 @@ dev.off()
 
 #most negative [E||C] - criteria for old sim figure
[…]

### 2021-09-01 code/Figure3.2.R

@@ -1,4 +1,4 @@
-load("results_numeric/params.RData")
+load("../results_numeric/params.RData")
 source("./decomp1_plotting.R")
 mudif <- c(0, -0.5, -2, -4)
 #FIGURE 3#
@@ -15,7 +15,7 @@ midx <- (pw*(nd/2))/sum(wd) #middle of figure panels; horizontal
 midy <- (ph*(nm/2))/sum(ht) #middle of figure panels; vertical 
 
 #########################################################################################
-pdf("results_figs/fig3.2.pdf")
+pdf("../results_figs/fig3.2.pdf")
 
 par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
 
@@ -69,10 +69,10 @@ mtext(expression(sigma), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig3.2maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
 cat("maximum standard error in figure three is", fig3.2maxse, "\n(M=", M, ")\n")
-saveRDS(fig3.2maxse, file="results_numeric/fig3.2maxse.RDS")
+saveRDS(fig3.2maxse, file="../results_numeric/fig3.2maxse.RDS")
 #Sys.time()
 #####################################################################################
-pdf("results_figs/fig3.2_qij.pdf")
+pdf("../results_figs/fig3.2_qij.pdf")
 
 par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2))
 
@@ -126,5 +126,5 @@ mtext(expression(sigma), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig3.2qijmaxse <- max(unlist(lapply(res, function(X){X$Dq_se})), na.rm = TRUE)
 cat("maximum standard error in figure three (qij) is", fig3.2qijmaxse, "\n(M=", M, ")\n")
-saveRDS(fig3.2qijmaxse, file="results_numeric/fig3.2qijmaxse.RDS")
+saveRDS(fig3.2qijmaxse, file="../results_numeric/fig3.2qijmaxse.RDS")
 #Sys.time()

### 2021-09-01 code/Figure4.2.R

@@ -1,6 +1,6 @@
 source("./decomp2_plotting.R")
 mudif = seq(0.0, -4.5, length.out = 100)
-load("results_numeric/params.RData")
+load("../results_numeric/params.RData")
 #FIGURE 4# 
 #plotting IGR and IGR-ATA to find ATA exclusion and ATA rescue in trait space (plot against mudif)
 #for different sigma and delta
@@ -15,7 +15,7 @@ midx <- (pw*(nd/2))/sum(wd) #middle of figure panels; horizontal
 midy <- (ph*(ns/2))/sum(ht) #middle of figure panels; vertical 
 
 #########################################################################################
-pdf("results_figs/fig4.2.pdf")
+pdf("../results_figs/fig4.2.pdf")
 
 par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
 
@@ -68,5 +68,5 @@ mtext(expression(mu[1]-mu[2]), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig4.2maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
 cat("maximum standard error in figure four is", fig4.2maxse, "\n(M=", M, ")\n")
-saveRDS(fig4.2maxse, file="results_numeric/fig4.2maxse.RDS")
+saveRDS(fig4.2maxse, file="../results_numeric/fig4.2maxse.RDS")
 Sys.time()

### 2021-09-01 code/MASTER.R

@@ -1,4 +1,6 @@
-#MASTER
+#MASTER - assumes the R working directory is the directory this file is in
+
+rm(list=ls())
 
 source("./makenoise.R") #make standard noise, change M here 
 
@@ -9,9 +11,9 @@ source("./Figure1.R")
 #FIGURE 2
 source("./Figure2.R")
 #FIGURE 3
-source("./Figure3.R")
+source("./Figure3.2.R")
 #FIGURE 4
-source("/Figure4.R")
+source("./Figure4.2.R")
 
 #source("./parameter_loop.R") #loop through parameters to get results #see section 4
 

### 2021-09-01 code/decomp1_plotting.R

@@ -2,7 +2,7 @@
 source("./decomposition_fxn.R")
 
 dePlot1 <- function(mudif, delta, qij=FALSE, legend=FALSE,...){
-  load("results_numeric/noise_etc.RData")
+  load("../results_numeric/noise_etc.RData")
   store <- vector(mode='list', length=7)
   sigma <- seq(0,7,1)
   

### 2021-09-01 code/decomp2_plotting.R

@@ -9,7 +9,7 @@ source("./decomposition_fxn.R")
 #delta  an integer value of delta
 
 dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, qij=FALSE, legend=FALSE,...){
-  load("results_numeric/noise_etc.RData")
+  load("../results_numeric/noise_etc.RData")
 
   store <- vector(mode='list', length=length(mudif))
   

### 2021-09-01 code/makenoise.R

@@ -8,10 +8,10 @@ rho <- cor(b_tilde$l)[1,2]
 
 cat("M=",M, " rho=", rho)
 
-if(dir.exists("results_numeric")==FALSE){
-  dir.create("results_numeric")
+if(dir.exists("../results_numeric")==FALSE){
+  dir.create("../results_numeric")
 }
 
-save(b_tilde,u_tilde, file = "results_numeric/noise_etc.RData")
-save(M, file="results_numeric/M.rda")
-save(rho, file="results_numeric/rho.rda")
+save(b_tilde,u_tilde, file = "../results_numeric/noise_etc.RData")
+save(M, file="../results_numeric/M.rda")
+save(rho, file="../results_numeric/rho.rda")

### 2021-09-01 code/parameters.R

@@ -19,9 +19,9 @@ delta <- c(0.2, 0.4, 0.6, 0.8)
 sigma <- c(2, 4, 5, 6)
 mudif <- c(0.0, -0.3, -0.6, -0.8)
 
-if(dir.exists("results_numeric")==FALSE){
-  dir.create("results_numeric")
+if(dir.exists("../results_numeric")==FALSE){
+  dir.create("../results_numeric")
 }
 
 
-save(delta, mu, mudif, sigma, file = "results_numeric/params.RData")
+save(delta, mu, mudif, sigma, file = "../results_numeric/params.RData")

### 2021-09-01 code/simsPlot_fxn.R

@@ -41,7 +41,7 @@ co.pPlot <- function(co.p, pop, start, end,...){
 }
 
 simsPlot <- function(mudif, delta, sigma, start=1, end=500){
-  load("results_numeric/noise_etc.RData")
+  load("../results_numeric/noise_etc.RData")
   
   #noise
   n <- transform(b_tilde, u_tilde, rho, sigma=sigma, mudif=mudif, b_s=T)

### 2021-09-01 code/Figure2.R
trying to get Sweave to work, and other small things

@@ -12,7 +12,7 @@ pdf("../results_figs/figure2.pdf")
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 fig2 <- simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
 dev.off()
-save(fig2, file="../results_numeric/fig2data.rda")
+save(fig2, file="../results_numeric/fig2data.RData")
 
 
 png("../results_figs/figure2.PNG", width=1000, height=600, pointsize=15) 

### 2021-09-01 code/Figure2.R
make noise_loc an argument

@@ -1,4 +1,5 @@
 source("./simsPlot_fxn.R")
+noise_loc <- "../results_numeric/noise.RData"
 fig_loc <- "../results_figs/"
 if(dir.exists(fig_loc)==FALSE){
   dir.create(fig_loc)
@@ -10,7 +11,7 @@ fig2data_loc <- "../results_numeric/fig2data.RDS"
 pdf(fig2_loc) 
 par(mfrow=c(2,1), mar=c(3,3,1,1), oma=c(2,1,1,1))
 
-fig2 <- simsPlot(mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
+fig2 <- simsPlot(noise_loc, mudif=-0.8, delta=0.8, sigma=5, end=1000) # r_woATA = 0.098, r = -0.0373, [E||C] = -0.1354
 
 dev.off()
 saveRDS(fig2, file=fig2data_loc)

### 2021-09-01 code/decomp1_plotting.R
make noise_loc an argument

@@ -1,8 +1,8 @@
 #THIS IS A FUNCTION
 source("./decomposition_fxn.R")
 
-dePlot1 <- function(mudif, delta, qij=FALSE, legend=FALSE,...){
-  load("../results_numeric/noise_etc.RData")
+dePlot1 <- function(noise_loc, mudif, delta, qij=FALSE, legend=FALSE,...){
+  load(noise_loc)
   store <- vector(mode='list', length=7)
   sigma <- seq(0,7,1)
   
@@ -45,26 +45,3 @@ dePlot1 <- function(mudif, delta, qij=FALSE, legend=FALSE,...){
 }
 
 
-#dePlot1(-0.3,1.0)
-
-
-
-
-
-#x0 <- decomp(0.8, c(0.3,0.3), 0.5, M)
-#x1 <- decomp(1.6, c(0.3,0.3), 0.5, M)
-#x2 <- decomp(3.2, c(0.3,0.3), 0.5, M)
-
-#plot(0, xlab="", ylab="", ylim=c(-0.08,0.35), xlim=c(0.8,3.2), col="white")
-#polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DEC_[1],x1$DEC_[1], x2$DEC_[1],0)
-#,col="darkred", border=NA)
-#polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$Dnull[1],x1$Dnull[1], x2$Dnull[1],0)
-#,col="darkblue", border=NA)
-#polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DE[1],x1$DE[1], x2$DE[1],0)
-#,col="darkorange1", border=NA)
-#polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DC[1],x1$DC[1], x2$DC[1],0)
-#,col="gold", border=NA)
-#polygon(c(0.8,0.8,1.6,3.2,3.2),c(0,x0$DECsharp[1],x1$DECsharp[1], x2$DECsharp[1],0)
-#,col="darkgreen", border=NA)
-#lines(c(0.8,1.6,3.2), c(sum(x0[1,]), sum(x1[1,]), sum(x2[1,])), lwd=3)
-#abline(h=0, lty=2)

### 2021-09-01 code/Figure4.R
fixed directory calling

@@ -1,6 +1,22 @@
 source("./decomp2_plotting.R")
+
+numeric_results_loc <- "../results_numeric"
+if(dir.exists(numeric_results_loc)==FALSE){
+  dir.create(numeric_results_loc)
+}
+params_loc <- paste(numeric_results_loc, "/params.RData", sep="")
+noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
+
+fig_loc <- "../results_figs/"
+if(dir.exists(fig_loc)==FALSE){
+  dir.create(fig_loc)
+}
+fig4_loc <- paste(fig_loc,"fig4.pdf",sep="")
+
+load(params_loc)
+
 mudif = seq(0.0, -4.5, length.out = 100)
-load("../results_numeric/params.RData")
+
 #FIGURE 4# 
 #plotting IGR and IGR-ATA to find ATA exclusion and ATA rescue in trait space (plot against mudif)
 #for different sigma and delta
@@ -15,7 +31,7 @@ midx <- (pw*(nd/2))/sum(wd) #middle of figure panels; horizontal
 midy <- (ph*(ns/2))/sum(ht) #middle of figure panels; vertical 
 
 #########################################################################################
-pdf("../results_figs/fig4.2.pdf")
+pdf(fig4.pdf)
 
 par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
 
@@ -54,7 +70,6 @@ for (s in 1:ns){
     }
     if(m==12){
       abline(v=-0.8, col="red", lty=2)
-      #lines(rep(-0.8,2), c(-1.5, 1),col="red", lty=2)
     }
     m <- m+1
   }
@@ -66,7 +81,8 @@ mtext(expression(delta), side=3, outer=TRUE, line=-2, font=2, cex=1.5, at=midx)
 mtext(expression(mu[1]-mu[2]), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 
 dev.off()
-fig4.2maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
-cat("maximum standard error in figure four is", fig4.2maxse, "\n(M=", M, ")\n")
-saveRDS(fig4.2maxse, file="../results_numeric/fig4.2maxse.RDS")
+fig4maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
[…]

### 2021-09-01 paper/Paper.Rnw

@@
 #allregres<-matrix(rnorm(10),2,5)
-#fig2data <- load(file="../results_numeric/fig2data.rda")
-#M <- load(file="../results_numeric/M.rda")
-#rho <- load(file="../results_numeric/rho.rda")
-#se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
-#se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
+load(file="../results_numeric/fig2data.RData")
+load(file="../results_numeric/M.rda")
+load(file="../results_numeric/rho.rda")
+se3 <- readRDS(file="../results_numeric/fig3.2maxse.RDS")
+se4 <- readRDS(file="../results_numeric/fig4.2maxse.RDS")
 @
@@
 Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution.
-The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.
-\label{fig:noise}}
+The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.\label{fig:noise}}
 \end{figure}
@@
 We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
-Populations inlfuenced by asymmetric tail association recover from low densities less often/ spend more time on average under one-species dominance.
+Populations inlfuenced by asymmetric tail association recover from low densities less often spend more time on average under one-species dominance.
 The top panel, A, is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise defined by $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$.

### 2021-09-01 code/makenoise.R
fixing directory calls

@@ -1,5 +1,15 @@
 source("./makenoise_fxn.R")
 
+numeric_results_loc <- "../results_numeric"
+
+if(dir.exists(numeric_results_loc)==FALSE){
+  dir.create(numeric_results_loc)
+}
+
+noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
+M_loc <- paste(numeric_results_loc, "/M.RDS", sep = "")
+rho_loc <- paste(numeric_results_loc, "/rho.RDS", sep = "")
+
 M <- 1000000
 b_tilde <- makenoise(M)
 u_tilde <- rnorm(M)
@@ -8,10 +18,8 @@ rho <- cor(b_tilde$l)[1,2]
 
 cat("M=",M, " rho=", rho)
 
-if(dir.exists("../results_numeric")==FALSE){
-  dir.create("../results_numeric")
-}
 
-save(b_tilde,u_tilde, file = "../results_numeric/noise_etc.RData")
-save(M, file="../results_numeric/M.rda")
-save(rho, file="../results_numeric/rho.rda")
+save(b_tilde,u_tilde, file = noise_loc)
+save(M, file=M_loc)
+save(rho, file=rho_loc)
+

### 2021-09-02 paper/SupMat.Rnw
a few more changes to finally make Sweave work on my machine. depends on jasmin changing the pdf file names so they don't include a . other than the one right bfore the extension
added section to sup mat for delta=1

@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 &= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
-&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
+&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:pECp}
 \end{align}
@@
 above, in this section, and note that all those terms depend only on marginals.
+
+\section{The $\delta=1$ case}\label{sect:delta1}
+
+%The i terms
+Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
+\begin{equation}
+\epsilon_i^0 = \mu_i - \mu_j.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
+\begin{align}
+\olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
+&= -\sigma^2/2.
+\end{align}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
+\begin{align}
+\olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
+&= \sigma^2/2.
+\end{align}
+Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
+\begin{align}
+\olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
+&= 0.
+\end{align}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
+\begin{align}
+\olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
+&= 0.
+\end{align}
+Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
+\begin{align}
+\olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
+&= 0.
+\end{align}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
+\begin{equation}
+\olep_i^{[E \nat C]}=0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
+\begin{equation}
+\olep_i^{[EC]}=0.
+\end{equation}
+%The j terms
+Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
+\begin{equation}
+\epsilon_j^0 = 0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
+\begin{equation}
+\olep_j^E = -\sigma^2/2.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
+\begin{equation}
+\olep_j^C = \sigma^2/2.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
+\begin{equation}
+\olep_j^{EC} = 0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
+\begin{equation}
+\olep_j^{(E \shp C)} = 0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
+\begin{equation}
+\olep_j^{(EC)} = 0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
+\begin{equation}
+\olep_j^{[E \nat C]} = 0.
+\end{equation}
+Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
+\begin{equation}
+\olep_j^{[EC]}=0.
+\end{equation}
+
+For $\delta=1$, the second value of $q_{ij}$ we considered, namely
+$q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
+options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
+equations using $q_{ij}=1$ gives
+\begin{align}
+\Delta_i^0 &= \mu_i - \mu_j \\
+\Delta_i^E &= 0  \\
+\Delta_i^C &= 0  \\
+\Delta_i^{(E \shp C)} &=  0 \\
+\Delta_i^{[E \nat C]} &= 0  \\
+\Delta_i^{[EC]} &=  0. 
+\end{align}
+Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
+which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
+without loss of generality).
+So there is no robust coexistence in the $\delta=1$ case for our simple model.
 

### 2021-09-02 code/MASTER.R
changedfigure  script names

@@ -11,10 +11,10 @@ source("./Figure1.R")
 #FIGURE 2
 source("./Figure2.R")
 #FIGURE 3
-source("./Figure3.2.R")
+source("./Figure3.R")
 #FIGURE 4
-source("./Figure4.2.R")
+source("./Figure4.R")
 
-#source("./parameter_loop.R") #loop through parameters to get results #see section 4
+
 
 

### 2021-09-02 code/Figure4.R
new var mudif_4

@@ -15,7 +15,7 @@ fig4_loc <- paste(fig_loc,"fig4.pdf",sep="")
 
 load(params_loc)
 
-mudif = seq(0.0, -4.5, length.out = 100)
+mudif_4 = seq(range(mudif)[2], range(mudif)[1]-0.5, length.out = 100)
 
 #FIGURE 4# 
 #plotting IGR and IGR-ATA to find ATA exclusion and ATA rescue in trait space (plot against mudif)

### 2021-09-02 paper/Paper.Rnw
a few more changes to finally make Sweave work on my machine. depends on jasmin changing the pdf file names so they don't include a . other than the one right bfore the extension
edits to main text

@@
 G3) To explore, using our simple model, the likelihood that species can coexist with
-ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
+ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
 may be negligible or meaninfully large.
-G4) To provide an empirical example of species coexistence, and to highlight the role ATAs plays in coexistence for this example.
-Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climatic changes, 
-and a means for beginning to understand its importance for real systems.
+G4) To provide an empirical example of species coexistence, and to highlight the role ATAs play in coexistence for the example.
+In the Discussion, we revisit the potential that climatic changes will influence
+the roles of ATAs in species coexistence and the maintenance of biodiversity.
+%***DAN: I added the above sentence. Not a fully fledged "goal" but a serious thing
+%to discuss.
+Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
+and a means for understanding its importance for real systems.
 
@@
 We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SI-SIsect:scaling_factors}.
-%Say this here if it turns out to be true, or say something like it:
+%***DAN: Say this here if it turns out to be true, or say something like it:
 %These alternatives turn out to make very little difference to our results. 
@@
 \ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
-\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ which is similar to and motivated by the 
-original decomposition of \cite{Chesson_1994}, and which we use; the decomposition is how the mechanisms of 
-coexistence opperating in a given case are illuminated. The decomposition is 
+\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
+which equals the IGR. The decomposition, which we use below, is similar to and motivated by the 
+original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
+coexistence operating in a given case are illuminated. The decomposition is 
 \begin{equation}
@@
 for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
-for left-tail associated noise. For each combination of $\delta=0.2, 0.5, 0.8, 1.0$
-and $\mu_1-\mu_2=0.0,-0.3,-0.6,-0.8$,
-we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. Relative importance of contributions was judged from the 
+for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
+and $\mu_1-\mu_2=0,-0.5,-2,-4$,
+we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
+Relative importance of contributions was judged from the 
 plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
 the contributions of other mechanisms.
+Mathematical simplications for the $\delta=1$ case make it straightforward to see that
+coexistence is impossible and that the case is relatively trivial (SI 
+section \ref{SI-sect:delta1}). So that case is not considered further.  
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
@@
 set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
-values $\mu_1 - \mu_2$ for which model species coexist in the presence of ATAs, but not in the absence of ATAs. 
+values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
 Similarly, we define ``ATA exclusion space'' to be the set difference 
@@
 against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
-$\mu_1 - \mu_2$ for all combinations of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
+$\mu_1 - \mu_2$ for all combinations of $\delta=0.2, 0.4, 0.6, 0.8$ and $\sigma=2,4,5,6$.
 
-%***DAN: Jasmin, I edited the below to get the above. Kept the below so I could add one comment to it, and in case it is needed again in
-%the near future
-%to identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model, we plotted the invasion growth rate [IGR],
-%***DAN: Note that this is not the IGR. The IGR is just r_{1\1}. So I abandoned this terminology and used "coexistence metric" instead
-%$\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, and the IGR without ATA contributribution, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}- \Delta_i^{[E \nat C]}$, against the difference between species, $\mu_1-\mu_2$.
-The same mathematical simplifications described above hold, such that $\mu_1$ and $\mu_2$ are the only ways that the species differ and computations for coexistence in the lottery model only depend on the difference, $\mu_1-\mu_2$.
-We henceforward call this difference the species trait difference and sets of multiple $\mu_1-\mu_2$ are sets of trait space range.
-Letting the trait $\mu$ vary within a one dimensional space, this space is then a trait space.
-Letting there be two points $\mu_1$ and $\mu_2$ within that trait space, then the set of points between $\mu_1$ and $\mu_2$ is a subset of trait space, and the difference, $\mu_1-\mu_2$, is the range of the subset of trait space.
-We then define $\psi_{ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is included (coexistence with ATA) and $\psi_{\sim ATA}$ to be the subset of trait space bound by a $\mu_1$ and $\mu_2$ whose range defines the maximal species trait difference that allows a postive IGR when ATA is not included (coexistence without ATA).
-We then consider the difference between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ to make the following conclusions.
-%If the range of $\psi_{ATA}$ > range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{\sim ATA}$ that result coexistence with ATA but failure to coexist without ATA; hence ATA rescue.
-%If the range of $\psi_{ATA}$ < the range of $\psi_{\sim ATA}$ then the continuous set of numbers in between the range of $\psi_{ATA}$ and the range of $\psi_{\sim ATA}$ defines the bounds of extensions to $\psi_{ATA}$ that result coexistence without ATA but failure to coexist with ATA; hence ATA exclusion.
-%In other words, ATA rescue is a subset of all $\mu_1-\mu_2$ that results in coexistence with ATA but failure to coexist without ATA, and ATA exclusion is a subset of all $\mu_1-\mu_2$ that results in coexistence without ATA, but failure to coexist with ATA.
-We highlighted and color coded these areas in our plots.
-We made the plots for values of $\delta=0.5, 0.8, 1.0$ and $\sigma=2.0,4.0,5.0, 6.0$.
+%***DAN: Methods paras here for G4, if we accomplish G4
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig1_vertical.pdf"}
+\includegraphics[width=.5\textwidth]{../results_figs/fig1_vertical.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/figure2.pdf"}
+\includegraphics[width=\textwidth]{../results_figs/figure2.pdf}
 \caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig3.2.pdf"}
+\includegraphics[width=\textwidth]{../results_figs/fig3p2.pdf}
 \caption{The contribution of ATA is non-negligible.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{"../results_figs/fig4.2.pdf"}
+\includegraphics[width=\textwidth]{../results_figs/fig4p2.pdf}
 \caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 
 IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference 
-$\mu_1-\mu_2$ in order to demonstrate ATA rescue and ATA exclusion. The red line, $\Delta_i^{[E \nat C]}$, 
+$\mu_1-\mu_2$ in order to demonstrate ATA rescue and ATA exclusion. The red line, $\Delta_i^{[E || C]}$, 
 %***DAN: I would normally have used \nat instead of ||, above, but \nat seems to not work in figure captions for some reason,

### 2021-09-02 code/decomp2_plotting.R
make noise_loc an arguemnt

@@ -8,8 +8,8 @@ source("./decomposition_fxn.R")
 #sigma  an integer value of sigma
 #delta  an integer value of delta
 
-dePlot2 <- function(mudif = seq(0, -0.8, -0.01), sigma, delta, qij=FALSE, legend=FALSE,...){
-  load("../results_numeric/noise_etc.RData")
+dePlot2 <- function(noise_loc, mudif, sigma, delta, qij=FALSE, legend=FALSE,...){
+  load(noise_loc)
 
   store <- vector(mode='list', length=length(mudif))
   

### 2021-09-02 code/Figure3.R
fixing loading and saving locations

@@ -82,8 +82,8 @@ mtext(expression(sigma), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig3maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
 cat("maximum standard error in figure three is", fig3maxse, "\n(M=", M, ")\n")
-fig3maxse_loc <- paste(numeric_results_loc, "fig3maxse.RDS", sep="")
-saveRDS(fig3.2maxse, file=fig3maxse_loc)
+fig3maxse_loc <- paste(numeric_results_loc, "/fig3maxse.RDS", sep="")
+saveRDS(fig3maxse, file=fig3maxse_loc)
 #Sys.time()
 #####################################################################################
 pdf(fig3qij_loc)
@@ -138,6 +138,6 @@ mtext(expression(sigma), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig3qijmaxse <- max(unlist(lapply(res, function(X){X$Dq_se})), na.rm = TRUE)
 cat("maximum standard error in figure three (qij) is", fig3qijmaxse, "\n(M=", M, ")\n")
-fig3qijmaxse_loc <- paste(numeric_results_loc, "fig3maxse.RDS", sep="")
+fig3qijmaxse_loc <- paste(numeric_results_loc, "/fig3maxse.RDS", sep="")
 saveRDS(fig3qijmaxse, file=fig3maxse_loc)
 #Sys.time()

### 2021-09-02 code/Figure4.R

@@ -31,7 +31,7 @@ midx <- (pw*(nd/2))/sum(wd) #middle of figure panels; horizontal
 midy <- (ph*(ns/2))/sum(ht) #middle of figure panels; vertical 
 
 #########################################################################################
-pdf(fig4.pdf)
+pdf(fig4_loc)
 
 par(mgp=c(3,0.5,0), mar = c(1,1,1,1), oma=c(3,5,3,2), xpd=TRUE)
 
@@ -58,7 +58,7 @@ res <- vector(mode='list',length=1)
 m <- 1
 for (s in 1:ns){
   for (d in 1:nd){
-    res <- append(res,dePlot2(mudif,sigma[s], delta[d], xaxt="n"))
+    res <- append(res,dePlot2(noise_loc, mudif_4,sigma[s], delta[d], xaxt="n"))
     axis(1, labels=ifelse(m>12, yes=TRUE, no=FALSE), tick=TRUE)
     mtext(LETTERS[m], side=3, line=-1.45, at=-4.5, adj=1)
     
@@ -84,5 +84,5 @@ dev.off()
 fig4maxse <- max(unlist(lapply(res, function(X){X$D_se})), na.rm = TRUE)
 cat("maximum standard error in figure four is", fig4maxse, "\n(M=", M, ")\n")
 fig4maxse_loc <- paste(numeric_results_loc, "/fig4maxse.RDS", sep="")
-saveRDS(fig4.2maxse, file=fig4maxse_loc)
+saveRDS(fig4maxse, file=fig4maxse_loc)
 Sys.time()

### 2021-09-02 code/makenoise.R
fix save to saveRDS

@@ -20,6 +20,6 @@ cat("M=",M, " rho=", rho)
 
 
 save(b_tilde,u_tilde, file = noise_loc)
-save(M, file=M_loc)
-save(rho, file=rho_loc)
+saveRDS(M, file=M_loc)
+saveRDS(rho, file=rho_loc)
 

### 2021-09-02 paper/Paper.Rnw
fixing figures and objects

@@
 #allregres<-matrix(rnorm(10),2,5)
-load(file="../results_numeric/fig2data.RData")
-load(file="../results_numeric/M.rda")
-load(file="../results_numeric/rho.rda")
-se3 <- readRDS(file="../results_numeric/fig3.2maxse.RDS")
-se4 <- readRDS(file="../results_numeric/fig4.2maxse.RDS")
+fig2 <- readRDS(file="../results_numeric/fig2data.RDS")
+M <- readRDS(file="../results_numeric/M.RDS")
+rho <- readRDS(file="../results_numeric/rho.RDS")
+se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
+se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
 @
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{../results_figs/fig1_vertical.pdf}
+\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/figure2.pdf}
+\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
 \caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig3p2.pdf}
+\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
 \caption{The contribution of ATA is non-negligible.
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig4p2.pdf}
+\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
 \caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 

### 2021-09-02 paper/SupMat.Rnw
uncommentedout sweave and put in code chunck to load M and rho and place them in where needed

@@
 \SweaveOpts{concordance=TRUE}
-%\SweaveOpts{concordance=TRUE}
+\SweaveOpts{concordance=TRUE}
 
 %The following is where you load in the numeric results that will be embedded in the text
-%<<eval=T,echo=F,message=F,warning=F>>=
-%system(paste("biber", sub("\\.Rnw$", "", current_input())))
-
-%#place R code here for loading in necessary variables from Results
-%#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-%allregres<-matrix(rnorm(10),2,5)
-%@
+<<eval=T,echo=F,message=F,warning=F>>=
+#system(paste("biber", sub("\\.Rnw$", "", current_input())))
+
+#place R code here for loading in necessary variables from Results
+#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
+#allregres<-matrix(rnorm(10),2,5)
+M <- readRDS(file="../results_numeric/M.RDS")
+rho <- readRDS(file="../results_numeric/rho.RDS")
+@
 
@@
 
-When needed, noise with symmetric tail association was generated by taking a very large sample (100000) 
+When needed, noise with symmetric tail association was generated by taking a very large sample (\Sexpr{M}) 
 %***DAN: Dan, later, check whether 100000 is what was actually used. I'm thinking we could probably do a million.
@@
 
-First, for a large integer $M$ (we used $M=$),
-%***DAN: Jasmin, please fill in the value, 1000000 or bigger if you can swing it
+First, for a large integer $M$ (we used $M=\Sexpr{M}$),
 carry out the following steps once.

### 2021-09-02 code/decomposition_fxn.R
deleting old&uneccesary

@@ -1,10 +1,10 @@
 decompose <- function(mudif,sigma,delta,b_tilde,u) {
   
-  bi_til<-b_tilde$l[,1]
-  bj_til<-b_tilde$l[,2]
-  bi_um <-b_tilde$s[,1]
-  bj_um <-b_tilde$s[,2]
-  M <- nrow(b_tilde$s)
+  bi_til<-b_tilde$l[,1] #left ATA_i
+  bj_til<-b_tilde$l[,2] #left ATA_j
+  bi_um <-b_tilde$s[,1] #symmetric_i
+  bj_um <-b_tilde$s[,2] #symmetric_j
+  M <- nrow(b_tilde$s) 
   
   #invader i
   #epsilon^0_i
@@ -156,116 +156,5 @@ decompose <- function(mudif,sigma,delta,b_tilde,u) {
 #round(dec,3)
 
 
-#decomposition function #need to check
-decomp <- function(sigma, mu, delta, M){
-  
-  #exponentiate noise to get B
-  load("./noise_etc.RData")
-  rv <- transform(b_tilde, u_tilde, rho, sigma, mu, b_s=TRUE) #makes random vars defined by parameters
-  Bl <- cbind(exp(rv$b_l1), exp(rv$b_l2))
-  Br <- cbind(exp(rv$b_r1), exp(rv$b_r2))
-  Bs <- cbind(exp(rv$b_s1), exp(rv$b_s2))
-  
-  #make B sharp
-  bsharp_tilde <- makenoise(M)
-  rho <- cor(bsharp_tilde$l)[1,2]
-  rvsharp <- transform(bsharp_tilde, u_tilde, rho, sigma, mu, b_s=TRUE)
-  Bl2sharp <- exp(rvsharp$b_l2)
-  Br2sharp <- exp(rvsharp$b_r2)
-  Bs2sharp <- exp(rvsharp$b_s2)
-  
-  #epsilon_0
-  null_l <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
-  null_r <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
-  null_s <- log(1-delta+delta*(mean(Bl[,1])/mean(Bl[,2])))
-  
-  #epsilon_C
[…]

### 2021-09-02 code/makenoise.R
set seed (1360)

@@ -9,6 +9,8 @@ if(dir.exists(numeric_results_loc)==FALSE){
 noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
 M_loc <- paste(numeric_results_loc, "/M.RDS", sep = "")
 rho_loc <- paste(numeric_results_loc, "/rho.RDS", sep = "")
+#-----------------------------------------------------------------#
+set.seed(1360)
 
 M <- 1000000
 b_tilde <- makenoise(M)

### 2021-09-02 paper/Paper.Rnw
comments, figure refs, Sexpr, commas, brackets

@@
 applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{},
-%***DAN: Jasmin, fill in Y and Y with names of systems, and accompanying reference(s), 
-%in a similar way to how I filled in the Angert example.
+%***DAN: 
 %Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
@@
 the three pairs of variables in figure \ref{fig:noise} all have the same
-Pearson correlation; and likewise have the same Spearman correlation. 
-%***DAN: Jasmin, when you make the figure, please display the Pearson and Spearman correlations on each panel, as in 
-%Fig. 1 of Ghosh et al (2020, Ecology and Evolution).
+Pearson correlation. 
+%***
+
 But each panel of figure \ref{fig:noise} nevertheless
@@
 The variables in figure \ref{fig:noise}A,C 
-%***DAN: Jasmin, when you make the figure, put the two asymmetric cases in A and C and put the symmetric case in B, in the middle.
+%***
 exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
@@
 considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
-%***DAN: Jasmin, please insert bib entries for the references Anderson et al., 2018; de Valpine et al., 2014; Popovic et al., 2019
-%from the Ghosh et al (2020; Adv Ecol Res) paper, and cite them here.
+%***
+%
 our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
@@
 Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
-In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018} ].
+In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
 %Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
@@
 In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low.
+%**JAS: should we introduce invader analysis or terminology here?
 Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
@@
 %discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
-%For clarification, we explain the how storage effects operate in maintaining biodiversity and the conditions that are required for it to do so.
-%***DAN: the above sentence will be superceded by the new topic sentence, above, of the combined paragraph, so you can cut this.
+%
+%***
 %We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
@@
 Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
-Environmental fluctuations are becoming more influenced by extreme climatic events (Hansen et al 2012) as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004, Ummenhofer_2017}.
+Environmental fluctuations are becoming more influenced by extreme climatic events \citep{Hansen_2012} as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004, Ummenhofer_2017}.
 Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition.
@@
 Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
-Increased variability of environmental fluctuations is a newer consideration for studying environmental changes in comparison to changes in means; however, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Increased variability of environmental fluctuations is a newer consideration for studying environmental changes in comparison to changes in means (vitation); however, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
 Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
@@
 Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-We simulated two different communites up to 100,000 time steps to hopefully make it clear to the reader than ATA makes a difference (see figure 2).
+We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader than ATA makes a difference (see figure \ref{fig:pedagog}).
 The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
-Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, almost seven times as much....
+Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, almost X times as much....
 
@@
 The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
-However, if environmental responses become asymmetrically associated by, say, extreme rainfall events coexistence can be altered.
+However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
 If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
 Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
-We see the possible consequences of ATA in Figure 2.
+We see the possible consequences of ATA in Figure \ref{fig:pedagog}.
 In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.
@@
 plays a role.
-%***DAN: make sure you define MTC and ATA once (probably in the Intro) and then use the abbreviations onle after that.
+
 For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.
@@
 Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019}
-%***DAN: Search for REF and replace all with appropriate citations
+
 quantify the contributions of 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' (IGR) of species $i$, that is, the mean growth rate of 
+where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' [IGR] of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;

### 2021-09-03 paper/Paper.Rnw
editing to G2 results and figure captions

@@
 
-%G2
-
-%Para on the main goal, whether the ATA contribution is meaningfully large
-%ATA effects tend to oppose "the rest" of storage effects
-
-%Para on how it's usually negative, whereas \Delta_[EC] is positive, so except when \delta=1, ATA effects tend to weaken storage effects
-%Note that when mu1-mu2=2, we are in the case where only storage effects allow for coexistence
-
-%***DAN: drop the delta =1 case from here
-%Para on when delta=1 (annual species) - in this case, coexistence is only possible when mu1-mu2 is close to 0, and in that case the only
-%thing allowing for coexistence at all is storage effects. Interetingly (and this is the topic sentence), "the rest" of storage effects are
-%negative, and it's the ATA component of storage effects which are allowing the species to coexist. Suggests the possibility that for 
-%annual species whcih are close competitors, ATAs might be especially important? Jasmin ought to check on this to make sure it's accurate.
-%What I'm suspicious about is that for annual species which are close competitors (say, mu1-mu2=0), and in the symmetric case, storage
-%effects in the lotery model actually don't opperate, and there is no coexistence. That sounds weird to me. Can we get some independent
-%verification of that? Why would annual species be different in this respect? There must be some delta between 0.8 and 1 where \Delta^[EC]
-%switches from positive to negative.
-
-%***DAN: consider possile para on mu1-mu2=0 case
-%***DAN: For discussion: is it true that ATA always weakens storage effects?
-Same as saying the ATA terms is always negative?
-The actual question is, for what models are ATA effects negative, versus positive?
-
-%***DAN: Jasmin, you computed standard errors of everything, but they are not currently reported. The best way to handle this would
-%be if we could say in captions in appropriate places "standard errors were always less than xyz", where xyz is some extremely 
-%small number. Standard errors are smalled for larger
-
-%from methods : Relative importance of contributions was judged from the plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to the contributions of other mechanisms.
-To explore the relative magnitudes of the contributions to coexistence that come from ATA versus other mechanisms in our simple two-species lottery model (G2), we plotted the decomposing mechanisms against increasing $\sigma$ for various values of $\mu_1 - \mu_2$ and $\delta$ (Figure 3).
-When species are the same, $\mu_1 - \mu_2=0$, only storage effect mechanisms contribute to IGR and other mechanisms (in black) stay at zero.
-When $\delta=1$, other mechanisms stay relatively more stable with $\Delta_i^0$ set at a negative value and other non-storage effect values being at zero or extremely close to zero.
-Otherwise, all mechanisms tend to increase in magnitude as $\sigma$ increases.
-The contribution of ATA captured by $\Delta_i^{[E \nat C]}$ (red) is not small.
-In almost all cases, aside from when $\delta=1$, ATA contributes negatively to coexistence while symmetric associations, $\Delta_i^{[EC]}$ (blue) contribute postively, and the former decreases with $\sigma$ while the latter increases with $\sigma$.
-We see the opposite when $\delta=1$, where the contribution of ATA are positve and the contribution of symmetric association are negative and the trends are reversed.
+%G2 and related side results
+Fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
+of the contributions to coexistence that come from ATAs compared to other mechanisms), Fig.
+\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
+magnitude comparable to the magnitudes of other terms of the extended MTC decomposition. 
+Magnitudes of $\Delta_i^{[E \nat C]}$ tended to be larger, relative to the magnitudes of the 
+other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
+$\mu_2$, i.e., for closer-to-equal competitors. 
+For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
+line and the $x$ axis for some values of $\sigma$, e.g., see panel H. 
+Such cases correspond to parameter combinations
+for which the presence or absence of ATAs determine whether the species coexist or not. In the 
+case of panel H, species 1 is excluded with ATA but the species can coexist without them. 
+In the $\mu_1=\mu_2$ case (panels A-D), for which the two species were equal 
+competitors, storage effects (blue plus red
+lines on Fig. \ref{fig:fig3}) were the only means by which coexisitence could occur, and ATA effects
+tended to be negative, weakening total storage effects.
 
@@
 \includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
-\caption{
-%Figure displaying the noise.
-Three panels, left-, symmetric, and right-tail assoc noise.
-Use standard normal marginals.Explain in caption that alternatives for $(b_1,b_2)$ are obtained by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.Jasmin to make and insert the figure here using the includegraphics command, and then draft a caption along the lines described here.
-Simple example of bivariate noise with standard normal marginals.
-A) is left-tail associated, B) is symmetric, and C) is right-tail associated.
-All three cases have the same Pearson correlation up to sampling variation, yet show different symmetry of tail associaton.
-Asymmetrically tail associated cases are perfectly correlated in the lower (A)  or upper (C) halfs of the distribution.
-The variables $(b_1,b_2)$ are transformed by multipling by $\sigma$ and then adding $(\mu_1,\mu_2)$.\label{fig:noise}}
+\caption{Simple example of bivariate noise with standard normal marginals. A) is left-tail 
+associated, B) is symmetric, and C) is right-tail associated noise. All three cases have 
+the same Pearson correlation, up to sampling variation, yet show different very different 
+patterns of association. We used asymmetrically tail associated cases which are perfectly 
+correlated in the lower (A)  or upper (C) halfs of the distributions, though it is also 
+possible to generate asymmetrically tail associated noise with imperfect correlations in 
+both halves of the distributions. The variables $(b_1,b_2)$ were obtianed by generating 
+noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$ (SI 
+section \ref{SI-sect:noise}).\label{fig:noise}}
 \end{figure}
+%***DAN: Dan has edited this caption at least once.
 
@@
 \end{figure}
+%***DAN: Dan has not yet edited this caption, will do so when it is time to edit the Intro
 
@@
 \includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
-\caption{The contribution of ATA is non-negligible.
-All panels show the breakdown of contributing mechanisms of coexistence in accordance to \cite{Ellner_2019} new decomposition method, against increasing $\sigma$.
-Each panel is defined by a unique set of parameters $\mu_1-\mu_2$ and $\delta$.
-The cases where $\delta=1$ are not shown because they are mathematically straighforward to see that IGR $=\mu_1 - \mu_2$ and is influenced by $\sigma$ (see SupMat).
-In most cases, the contribution of ATA (red) is negatively; however when $\delta=1$, ATA contribution is positive.
-The contribution of ATA is non-negligible and in special cases result in a sign change of IGR (orange).
-Standard error was computed but not displayed because the maximum standard error of all computations was \Sexpr{round(se3,5)}.
-\label{fig:fig3}}
+\caption{The contribution of ATAs to species coexistence is meaningfully large compared to 
+other components of the MTC decomposition. All panels show the IGR and the decomposition of
+the IGR provided by the extended MTC, equation (\ref{eq:full_decomp}). For each panel, 
+$\delta$ and $\mu_1-\mu_2$ 
+are fixed at values specified in the margins, and the IGR and components of the decomposition
+are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
+has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
+to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
+betweed the ATA line (red)
+and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
+coexistence and exclusion of the weaker competitor. Standard error was never greater than
+\Sexpr{round(se3,5)}, a small value compared to y axis extents and comparisons between lines,
+so standard errors were not displayed. \label{fig:fig3}}
 \end{figure}
+%***DAN: Dan has edited this caption
 

### 2021-09-03 code/Figure4.R
fixed legend

@@ -44,7 +44,7 @@ layout(matrix(c(2,3,4,5,1,
 
 plot.new() #1
 legend("topright", 
-       legend=c(expression(IGR),expression(IGR-Delta["[E||C]"]), expression(Delta["[E||C]"])),
+       legend=c(expression(IGR),expression(IGR-Delta[i]^"[E||C]"), expression(Delta[i]^"[E||C]")),
        col = c("orange","navy", "red"),
        lty = c(1,1,1), bty="n", cex=1.5, inset=c(-0.25,-0.042),
        y.intersp = 1.25, x.intersp = 0.1, seg.len=1)

### 2021-09-03 paper/Paper.Rnw
edits, mostly to results, and also removing a lot of obsolete text from the sup mat and putting it in a scraps doc

@@
 %save formal definitions for later and reword this. Here's a try at rewording: 
-G3) To explore, using our simple model, the likelihood that species can coexist with
+G3) To explore, using our simple model, whether species can coexist with
 ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
@@
 $\mu_2$, i.e., for closer-to-equal competitors. 
-For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
-line and the $x$ axis for some values of $\sigma$, e.g., see panel H. 
-Such cases correspond to parameter combinations
-for which the presence or absence of ATAs determine whether the species coexist or not. In the 
-case of panel H, species 1 is excluded with ATA but the species can coexist without them. 
 In the $\mu_1=\mu_2$ case (panels A-D), for which the two species were equal 
@@
 %G3
+Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
+can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
+make the difference between coexistence and exclusion: both ``ATA rescue'' and ``ATA
+exclusion'' (see Methods) are possible.
+For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
+line and the $x$ axis for some values of $\sigma$, e.g., see panel H. 
+Such cases correspond to parameter combinations
+for which the presence or absence of ATAs determine whether the species coexist or not. In the 
+case of panel H, for values of $\sigma$ around $4$, 
+species 1 is excluded with ATAs but the two species can coexist without ATAs. 
+Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
+indicate the IGR when the effects of ATAs are included, and blue lines are the IGR when
+the effects of ATAs are neglected. These two lines frequently crossed the $x$ axis in 
+different places, showing that both ATA rescue space and ATA exclusion space could be non-empty
+subsets of our trait-difference space (see Methods). For the parameters we considered, 
+both ATA rescue and exclusion were more common for larger values of $\sigma$.
+ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
+more common for larger $\delta$. Although bands of ATA rescue or exlusion on Fig.
+\ref{fig:fig4} appear narrow, they nevertheless do occur,
 
-While the contribution of ATA is definitively not absent, in most cases it does not make a difference for coexistence or comepetitive exclusion.
-In most panels of Figure 3, the magnitude of negative contribution from ATA does not effect the sign of the IGR.
-The absense of ATA contribution will increase IGR in those cases, but the outcome of coexistence would likely not be different.
-%maybe move or expand more on this point in discussion
-%cite sources that say the sign of IGR matter more than the magnitude 
-However, there are cases where ATA makes a difference for coexistence or competitive exclusion where the maginitude of $\Delta_i^{[E \nat C]}$ is large and other mechanisms do not counteract resulting in a sign change for IGR.
-We call these cases ATA rescue or ATA exclusion, the exploration of which is at the core of G3.
+Althoug ATA rescue or exclusion
+occurs only for the minority of parameter combinations considered, this does not mean
+ATAs are unimportant for species coexistance and biodiversity maintenance - the same is
+probably also true of
 
-To acheive G3, we assessed the effect of ATA by comparing and plotting (Figure 4) the IGR computed for G2 and the IGR if there were no ATA (IGR - $\Delta_i^{[E \nat C]}$).
-To find areas of ATA exclusion or ATA rescue we looked at where these line cross y=0 and highlighted and color coded ATA exclusion trait space and ATA rescue trait space (see Methods).
-In these spaces, contribution of ATA is the mechanism soley responsible for a positive or negative IGR.
  Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical.
@@
 and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. Standard error was never greater than
-\Sexpr{round(se3,5)}, a small value compared to y axis extents and comparisons between lines,
-so standard errors were not displayed. \label{fig:fig3}}
+coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
+were never greater than \Sexpr{round(se3,5)}, so are not displayed. \label{fig:fig3}}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
-\caption{ATA rescue and ATA exclusion can be defined within subsets species trait difference spaces. 
-IGR (ATA included; yellow) and IGR without ATA (violet) are plotted against species trait difference 
-$\mu_1-\mu_2$ in order to demonstrate ATA rescue and ATA exclusion. The red line, $\Delta_i^{[E || C]}$, 
-%***DAN: I would normally have used \nat instead of ||, above, but \nat seems to not work in figure captions for some reason,
-%it crashes the latex on my machine at least. If we can fix this, we should, but it's not the end of the world if we can't.
-%Jasmin, could you compile on your machine when || above was \nat?
-is the difference between the two IGR's. Each panel is defined by a unique set of parameters $\sigma$ 
-and $\delta$. ATA rescue is marked by regions in the panels where IGR wihtout ATA contribution is less 
-than IGR at zero, denoted by regions of dotted yellow. ATA exclusion is marked by regions where IGR 
-without ATA contribution is greater than IGR at zero, denoted by regions of dotted pink. ATA rescue 
-is only seen in cases where $\delta=1$ (left column). All other cases shown here have subsets of 
-species trait difference space where ATA exclusion occurs. Despite panels in the top row not 
-containing the color coded region, these cases still contain ATA exlusion spaces; however, the space 
-is too small to appropriately denote while still maintaining consistent x-axes across all panels. 
-For $\delta<1$, ATA exclusion space increases down rows as $\sigma$ increases. ATA exclusion space also 
-seems to increase across columns as $\delta$ decreases; however this trend is broken in the last row, 
-indicating a more compex relationship between ATA exclusion space, $\delta$, and $\sigma$.
-Standard error was computed but not displayed because the maximum standard error of all computations was \Sexpr{round(se4,5)}.
-}\label{fig:fig4}
+\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
+competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
+All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
+(\ref{eq:full_decomp})], and the difference of these quantities. 
+For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
+margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
+``trait-difference space'' corresponding to ATA rescue 
+($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
+or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
+see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
+$\text{IGR}-\Delta_i^{[E || C]}$ is negative correspond to ATA rescue (shaded yellow), whereas
+values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[E || C]}$ is positive 
+correspond to ATA exclusion (shaded violet). Top panels also include domains
+of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
+probably of minimal practical importance. 
+Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
+so are not displayed. The dashed red line on panel L corresponds to the parameters
+used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
+using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section ).}\label{fig:fig4}
 \end{figure}
+%***DAN: Dan has edited this caption
 

### 2021-09-03 paper/SupMat.Rnw
edits, mostly to results, and also removing a lot of obsolete text from the sup mat and putting it in a scraps doc

@@
 
-%***DAN: Most but not all of the below is obsolete now, replaced and upgraded by stuff above this point
-
-\section{Invasion growth rate of species 1, and storage effects} \label{sect:IGR}
-
-We consider the invasion growth rate of species 1, $\bar r_1=\EX(\ln (\frac{N_1(t+1)}{N_1(t)}))$, when $N_1$ is close to 0 and $N_2$ is close to $N$.
-Hence $\EX(\cdot)$ represents expected value.
-We have
-
-\begin{equation}
-\frac{N_1(t+1)}{N_1(t)}=(1-\delta)+\delta N \frac{B_1(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)},  \label{sp1_Nt+1_over_Nt}
-\end{equation}
-
-\noindent but setting $N_1=0$ and $N_2=N$ here gives the result that
-
-\begin{equation}
-\bar r_1=\EX(\ln [(1-\delta)+\delta \exp(b_1-b_2)]).  \label{rbar1_E}
-\end{equation}
-
-\noindent We considered the various cases for the joint distribution $(b_1,b_2)$, specified in the main text and in the previous section.
-
-To study storage effect for the model, we also consider the contribution of EC covariance to the difference between the invader's and the resident's mean growth rates, to be denoted $\Delta I$. 
-Following REF, we define
-
-\begin{equation}
-\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp (b_1^\#-b_2)]),  \label{rbar1sharp_E}
-\end{equation}
-
-\noindent where $b_1^\#$ is distributed in the same way as $b_1$, but is independent of it and $b_2$. 
-Then $\bar r_1 - \bar r_1^\#$ is the contribution of EC covariance to the mean growth rate species 1, the invader. 
-We know, a priori, that the mean growth rate of the resident, species 2, must be 0, $\bar r_2 =0$, but equation (\ref{model_eq}) also implies that
-
-\begin{equation}
-\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta N \frac{B_2(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}.  \label{sp2_Nt+1_over_Nt}
-\end{equation}
-
-\noindent Setting $N_2=N$ and $N_1=0$ then gives
-
-\begin{equation}
-\frac{N_2(t+1)}{N_2(t)}=(1-\delta)+\delta=1  \label{sp2_Nt+1_over_Nt_simp}
-\end{equation}
-
-\noindent so
-
-\begin{equation}
-\EX(\ln (\frac{N_2(t+1)}{N_2(t)}))=0, \label{r2_is_zero}
-\end{equation}
-
-\noindent as expected. 
-Again, following REF, we also define
-
-\begin{equation}
-\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(b_2^\#-b_2)]),  \label{rbar2sharp_E}
-\end{equation}
-
-\noindent where $b_2^\#$ is distributed in the same way as $b_2$, but is independent of it and of $b_1$. 
-Thus $\bar r_2- \bar r_2^\# = -\bar r_2^\#$ is the contribution of EC covariance to the mean growth rate of species 2, the resident. 
-And so
-
-\begin{equation}
-\Delta I=\left(\bar r_1 - \bar r_1^\#\right)-\left(\bar r_2- \bar r_2^\#\right)=\bar r_1 - \bar r_1^\#+\bar r_2^\#  \label{DeltaI}
-\end{equation}
-
-\noindent has been defined in terms of expected values of elementary expressions of the random variables $\left(b_1,b_2\right)$, $\left(b_1^\#,b_2\right)$, and $\left(b_2^\#, b_2\right)$.
-
-Having already specified that $b_1$ (and therefore also $b_1^\#$) is distributed as $N\left(\mu_1, \sigma^2\right)$, and that $b_2$ (and therefore also $b_2^\#$) is distributed as $N\left(\mu_2, \sigma^2\right)$, we know, using $\Sigma$ as notation for a covariance matrix that
-
-\begin{equation}
-\left(b_1^\#, b_2\right) \sim N\left((\mu_1,\mu_2),\Sigma \right) \label{b1sharpb2_distribution}
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-\left(b_2^\#, b_2\right) \sim N(( \mu_2,\mu_2), \Sigma) \label{b2sharpb2_distribution}
-\end{equation}
-
-\noindent where
-
-\begin{equation}
-\Sigma = \begin{pmatrix} \sigma^2&0\\0&\sigma^2\end{pmatrix}. \label{Sigmamat_cov0}
-\end{equation}
-
-\noindent Thus the difference $b_1^\#-b_2$ that occurs in the expression for $\bar r_1^\#$ (\ref{rbar1sharp_E}) is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance $2\sigma^2$.
-Thus
-
-\begin{equation}
-\bar r_1^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar1sharp_E_usub}
-\end{equation}
-
-\noindent where $u \sim N(\mu_1-\mu_2,2\sigma^2)$. 
-Likewise, the difference $b_2^\#-b_2$ in (\ref{rbar2sharp_E}) is a normally distributed random variable with mean $\mu_2-\mu_2=0$ and variance $2\sigma^2$, so
-
-\begin{equation}
-\bar r_2^\#=\EX(\ln [(1-\delta)+\delta \exp(u)]),  \label{rbar2sharp_E_usub}
-\end{equation}
-
-\noindent where now $u \sim N(0,2\sigma^2)$. 
-The expression (\ref{rbar1_E}) for $\bar r_1$ can also be further simplified, in a similar way, when we are in the symmetric tail association case considered in section \ref{sect:noise}. 
-In that case,
-
-\begin{equation}
-(b_1,b_2) \sim  N(( \mu_1,\mu_2), \Sigma) \label{b1b2_distribution_sym}
-\end{equation}
-
-\noindent where now
-
-\begin{equation}
-\Sigma = \begin{pmatrix} \sigma^2&\rho\\\rho&\sigma^2\end{pmatrix} \label{Sigmamat_rho}
-\end{equation}
-
-\noindent and $rho=put in value$ was determined as in section \ref{sect:noise}.
-Therefore, the expression $b_1-b_2$ that occurred in \ref{rbar1_E}, in the symmetric tail association case, is a normally distributed random variable with mean $\mu_1-\mu_2$ and variance
-
-\begin{align}
-\VarX(b_1-b_2)&=\CovX(b_1-b_2, b_1-b_2) \\
-                        &=\CovX(b_1,b_1)+\CovX(b_2,b_2)-\CovX(b_1,b_2)-\CovX(b_2,b_1) \\
-                        &=2\sigma^2-2\rho.
-\end{align}
-
-\noindent Thus, again only in the symmetric tail association case,
-
-\begin{equation}
-\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), \label{rbar1_E_usub}
-\end{equation}
-
-\noindent where here $u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$. 
-As for all we know there is no simplicifaction of the expression \ref{rbar1_E}  for $\bar r_1$ in the left and right-tail association cases.
-
-Thus for in summary,
-
-\begin{equation}
-\bar r_1 = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho), \label{sym_rbar1}
-\end{equation}
-
-\noindent for the symmetric case, and
-
-\begin{equation}
-\bar r_1^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(\mu_1-\mu_2, 2\sigma^2), \label{sym_rbar1sharp}
-\end{equation}
-
-\begin{equation}
-\bar r_2^{\#} = \EX(\ln[1-\delta+\delta \exp(u)]), u \sim N(0, 2\sigma^2). \label{sym_rbar2sharp}
-\end{equation}
-
-\noindent We next use the expressions to develop methods to estimate $\bar r_1$ and $\Delta I$ for any given paramters $\delta,\mu_1,\mu_2,\sigma$.
-
-\section{Estimating $\bar r_1$ and $\Delta I$}
-
-First, for a large integer $M$, carry out the following steps once:
-
-\begin{enumerate}
-\item Generate left-tail assocated noise $\left(\tilde b_{l1}^{(i)}, \tilde b_{l2}^{(i)}\right)$, $i=1,...,M$, as described in the section \ref{sect:noise}.
-Hence $\tilde b_{l1}^{i)}$ and $\tilde b_{l2}^{(i)}$ are distributed as $N(0,1)$.
-\item Generate right-tail associated noise $\left(\tilde b_{r1}^{(i)}, \tilde b_{r2}^{(i)}\right)$, $i=1,...,M$, via the analogous procedure.
-Again, $\tilde b_{r1}^{i)}$ and $\tilde b_{r2}^{(i)}$ are distributed as $N(0,1)$.
-\item Generate $M$ points $\tilde u_i$, $i = 1,...,M$ from a standard normal distribution. 
-\end{enumerate}
-
-\noindent Next, given values of parameters $\delta, \mu_1, \mu_2$, and $\sigma$, proceed as follows to get estimates of $\bar r_1, \bar r_1^{\#}, \bar r_2^{\#}$, and $\Delta I$.
-
-\begin{enumerate}
-
-\item Estimate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{l1}^{(i)}-b_{l2}^{(i)})])$ where $b_{l1}^{(i)} = \sigma \tilde b_{l1}^{(i)} + \mu_1$ and $b_{l2}^{(i)} = \sigma \tilde b_{l2}^{(i)} + \mu_2$.
-The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{l1}^{(i)}-b_{l2}^{(i)})])}{\sqrt M}$, where $\sd _i(\cdot)$ is standard deviation.
-
-\item Estiate $\bar r_1$ for the left-tail associated case as $\hat{\bar r_1} = \mean_i (\ln [1-\delta+\delta\exp(b_{r1}^{(i)}-b_{r2}^{(i)})])$ where $b_{r1}^{(i)} = \sigma \tilde b_{r1}^{(i)} + \mu_1$ and $b_{r2}^{(i)} = \sigma \tilde b_{r2}^{(i)} + \mu_2$.
-The standard error of this estimate is $\se (\hat{\bar r_1}) =\frac{\sd _i (\ln[1-\delta +\delta \exp(b_{r1}^{(i)}-b_{r2}^{(i)})])}{\sqrt M}$.
-
-\item Estimate $\bar r_1$ in the symmetric tail association case as $\hat{\bar r_1}=\mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{(2\sigma^2-2\rho)}\tilde u_i + \mu_1 - \mu_2$.
-The standard error of this estimation is $\se(\hat{\bar r_1}) = \frac {\sd_i(\ln[1-\delta+\delta \exp(u_i)])}{\sqrt M}$.
-
-\item Estimate $\bar r_1^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_1^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i + \mu_1 - \mu_2$.
-The standard error is $\se(\hat{\bar r_1^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
-
-\item Estimate $\bar r_2^{\#}$ (for all three tail-assocaition cases) as $\hat{\bar r_2^{\#}} = \mean_i(\ln[1-\delta+\delta \exp(u_i)])$, where $u_i = \sqrt{2\sigma^2}\tilde u_i$.
-The standard error is $\se(\hat{\bar r_2^{\#}}) = \sd_i \frac{ln[1-\delta+\delta \exp(u_i)]}{\sqrt M}$.
-
-\item $\hat {\Delta I} = \hat{\bar r_1} - \hat{\bar r_1^{\#}} + \hat{\bar r_2^{\#}}$, estimated separately for the left-tail associated, right-tail associated and symmetric noise cases.
-We obtained conservative standard error estimates by adding the standard error of the estimates of $\hat{\bar r_1}, \hat{\bar r_1^{\#}}$, and $\hat{\bar r_2^{\#}}$.
-
-\end{enumerate}
-
-\noindent For large enough $M$, standard error should be very small.
-Since random variables are generated once only, computation of these quantitaties for a large number of parameter sets should be fast.
-
-\section{Additional quantities influencing invasion success} \label{sect:addquant}
-
-Not only does the expected value $\bar r_1$ influence the capacity for species 1 to invade where rare, but so do other aspects of the distribution of
-\begin{equation}
-r_1 = \ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1_distribution}
-\end{equation}
-
-We here elaborate how some other aspects of the distribution of $r_1$ were quantified, for our three tail association cases and for any of our parameters $\delta, \mu_1, \mu_2, \sigma$.
-
-First, we considered $\sd(r_1)$, the standard deviation.
-If this is large enough, then $r_1$ can occasionally be positive even if $\bar r_1$ is strongly negative, and this invasion can occur under the right environmental circumstances.
-
-We also considered the probability $\Prob[r_1 >0]$, which is, explicitly, the probability that environmental variables that permit invasion. We have
-\begin{align}
-\Prob[r_1>0] &= \Prob [1-\delta+\delta \exp(b_1-b_2) > 1] \\
-&= \Prob[\exp (b_1-b_2)>1] \\
-&= \Prob[b_1>b_2],
-\end{align}
-
-\noindent so this was computed for each considered parameter set of $\delta, \mu_1, \mu_2, \sigma$, and for each of our three tail association cases.
-
-We also consider the mean value of the possible part of the distribution of $r_1$, $\EX(r_1|r_1>0)$. 
-This characterizes how quickly invasion may happen, when it happens, with larger values corresponding to faster invasion. 
-If it takes several time steps for the population of the invader to rise from negligible level at which species 1 is a noticeable presence in the community, then environmental conditions must be suitable for invasion for several sequential time steps.
-Thus large values of $\EX(r_1|r_1>0)$ should also correspond to a greater probability of invasion because they should correspond to faster invasions and therefore to reduced need for sequential periods of suitable environmental conditions.
-
-Finally, for a few choices of parameters $\delta, \mu_1, \mu_2, \sigma$, we displayed the whole distribution of $r_1$, values for each of our tail association cases, to make visual comparisons.
-
-\section{Symmetries and parameter reduction}
-
-The quantities we consider to address invasion prospects for the weaker competitor, species 1 (see sections \ref{sect:IGR} and \ref{sect:addquant}) depend only on $\mu_1-\mu_2$ and not, independtly, on $\mu_1$ and $\mu_2$, and also are the same for our left- and right-tailed association cases.
-We prove these statements in this section.
-
-Equations \ref{sym_rbar1} - \ref{sym_rbar2sharp} obviously depend only on $\mu_1-\mu_2$, and not independently on $\mu_1$ and $\mu_2$, since only $\mu_1-\mu_2$ appears in these expressions if $\mu_1$ or $\mu_2$ appear at all.
-Equation \ref{rbar1_E} is $\bar r_1$ in the asymmetric tail association cases.
-But letting $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$ and substituting into \ref{rbar1_E}, we get
-
-\begin{align}
-\bar r_1 &= \EX(\ln[1-\delta+\delta \exp(\beta_1-\eta-(\beta_2-\eta))])\\
-&= \EX(\ln[1-\delta+\delta \exp(\beta_1-\beta_2)]).
-\end{align}
-\noindent And so altering the mean of both $b_1$ and $b_2$ distributions by the same amount has no affect of $\bar r_1$, and thus $\bar r_1$ in the asymmetric tail associations also depends only on $\mu_1 - \mu_2$ and not independtly on $\mu_1$ or $\mu_2$.
-
-In section \ref{sect:addquant}, we introduce several other metrics. These are all based on the distribution 
-\begin{equation}
-r_1=\ln[1-\delta+\delta \exp(b_1-b_2)]. \label{r1distribution}
-\end{equation}
-
-\noindent But, again making the substitutions $\beta_1 = b_1 + \eta$ and $\beta_2 = b_2 + \eta$, we can see via the same resoning as above that the whole distribution $r_1$ depends only on $\mu_1 - \mu_2$, and not on $\mu_1$ and $\mu_2$ independently, for all three of our tail association cases.
-
-To see that our various metrics are the same for out left- and right-tail associated noise processes, let $(b_1,b_2)$ denote our left-tail associated random variable with parameters $\mu_1,\mu_2,\sigma$.
-Then define $\beta_1 = -b_2 + \mu_1 +\mu_2$, $\beta_2=-b_1+\mu_1+\mu_2$. 
-It is easy to see that $\EX(\beta_1)=\mu_1$, $\EX(\beta_2)=\mu_2$, $\sd(\beta_1)=\sd(\beta_2)=\sigma$, so $\beta_1$ and $\beta_2$ are normally distributed.
-From there it is easy to see that $\beta_1$ and $beta_2$ are distributed in the same wat as our right-tail associated noise with parameter $\mu_1,\mu_2$, and $\sigma$. 
-Making the substitutions $b_2 = -\beta_1 + \mu_1 + \mu_2$ and $b_1 = -\beta_2 + \mu_1+\mu_2$ in \ref{r1distribution}, we get
-
-\begin{align}
-r_1 &= \ln[1-\delta+\delta \exp(-\beta_2+\mu_1+\mu_2-(\beta_1+\mu_1+\mu_2))]\\
-&= \ln[1-\delta+\delta \exp(\beta_1-\beta_2)].
-\end{align}
-
-\noindent And since $(\beta_1, \beta_2)$ is distributed in the same way as our right-tail associated noise, this proves all metrics based on the distribution of $r_1$ (\ref{r1distribution}) are the same for left- and right-tail associated noise.
-
-\section{Decomposition}\label{Decomposition}
-
-To quantify the contribution of different drivers to the difference in growth rate between the invader and the resident, the amount variance due to that driver is considered.
-We consider the contributions of the null growth rate, $\epsilon^0$, the contribution of variance in E (environment), $\bar \epsilon^E$, the contribution of the variance in C (competitive pressure), $\bar \epsilon^C$, the contrubition of E and C co-varying $\bar \epsilon^{(EC)}$, and the the contribtion of E and C varying, but independently, $\bar \epsilon^{(E\#C)}$, for the invader and the resident.
-The difference between invader and resident is computed term by term to get a full decomposition as follows,
-
-\begin{equation}
-\bar r_1-\bar r_2 = \Delta^0+\Delta^E+\Delta^C+\Delta^{(EC)}+\Delta^{(E\#C)}, \label{IGR}
-\end{equation}
-
-\noindent where each $\Delta$ denotes a drivers contribution the the invader growth rate subtracted by the drivers contribution to the resident's groeth rate.
-
-The null growth rate is computed as the growrth rate when both drivers (E and C) are set at their mean. Recalling the growth rate eqaution from equation \ref{rbar1_E},
-
-\begin{align}
-\Delta^0 &= \epsilon_1^0 - \epsilon_2^0\\
-&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - \ln(1-\delta + \delta \frac{\bar B_2}{\bar B_2})\\
-&=\ln(1-\delta + \delta \frac{\bar B_1}{\bar B_2}) - 0. \\
-%&=\ln(1-\delta + \delta \exp(\bar b_1-\bar b_2)
-\end{align}
-
-The contribution of variance in E to a species growth rate is computed by letting E vary and setting C to its mean, subtracted by the null.
-The contribution of variance in C to a species' growth rate is computed in an analogous way, but reverse.
-
-\begin{align}
-\Delta^E &= \bar \epsilon_1^E - \bar \epsilon_2^E\\
-&=[\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) - \epsilon_2^0]\\
-&=\EX (\ln[1-\delta + \delta \frac{B_1}{\bar B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{ B_2}{\bar B_2}]) \\
-\end{align}
-
-\begin{align}
-\Delta^C &= \bar \epsilon_1^C - \bar \epsilon_2^C\\
-&=[\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0] - [\EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) - \epsilon_2^0]\\
-&=\EX (\ln[1-\delta + \delta \frac{\bar B_1}{B_2}]) - \epsilon_1^0 - \EX(\ln[1-\delta + \delta \frac{\bar B_2}{B_2}]) \\
-\end{align}
-
-The contribution of E and C both varying but not together is computed by letting both terms vary but decoupling any association by substituting in a new independent varibable for one of the varibales.
-Here we use C.
-The term is subtracted by the previous contributions.
-
-\begin{align}
-\Delta_1^{(E\#C)} &= \bar \epsilon_1^{(E\#C)} - \bar \epsilon_2^{(E\#C)}\\
-&=[\EX(\ln[1-\delta + \delta \frac{B_1}{B_2^{\#}}]) -[\epsilon_1^0 +\bar \epsilon_1^E + \bar \epsilon_1^C]] - [\EX(\ln[1-\delta + \delta \frac{B_2}{B_2^{\#}}]) -[\epsilon_2^0 +\bar \epsilon_2^E + \bar \epsilon_2^C]]\\
-\end{align}
-
-The contribution of E and C co-varying is the storage effect. Computed by,
-
-\begin{equation}
-\Delta^{(EC)} = \bar \epsilon_1^{(EC)} - \bar \epsilon_2^{(EC)}.
-\end{equation}
-
-\noindent Where,
-
-\begin{align}
-\bar \epsilon_i^{(EC)}&=\bar \epsilon_i^{EC}-\bar \epsilon_i^{(E\#C)}\\
-&= \EX(\ln(1-\delta + \delta \frac{B_i}{B_j}))-[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C]\\
-&- \EX(\ln(1-\delta + \delta \frac{B_i}{B_j^{\#}})) +[\epsilon_i^0 +\bar \epsilon_i^E + \bar \epsilon_i^C].
-\end{align}
-
-By substituting these expansions into \ref{IGR}, we get,
-
-Recall from main text that the goal of this research is to investigate the implications of extreme events and thus asymmetric tail associations on coexistence.
-Hence, we compare IGR and mechanism values for each noise type.
-Because only association structure and not means and standard deviation is differ across noise types, different noise types will not result in differences in any of the decomposition terms outside of the storage effect term.
-All other terms do not consider the associative relationship between the two variables.
-The null growth grate is a constant that depends on the mean of both variables, the contribution of variance in one variable only depends on that one variable scaled by the mean of the other, and the varying but independent term while it does consider both variable they are set to be completely uncorrelated.
-These four terms will not differ across noise types.
-
-\section{new decomposition term}
-
-A new decomposition term was proposed to quantify the effect of asymmetric associations to coexistence.
-Since asymmetry only differs in the storage effect term, $\bar \epsilon^{(EC)}$, the contribution of asymmetry associations between the species will be a further decomposition.
-
-\begin{equation}
-\bar \epsilon^{(\dot {EC})} = \bar \epsilon^{(EC)} - \bar \epsilon^{( {EC_0})}
-\end{equation}
-
-The contribution of asymmetry to the storage effect is the observed storage effect subtracted by a null storage effect if the association between species was completely symmetrical. 
-The descrition of $\bar \epsilon^{(\ddot {EC})}$ is given in section 3 and equations 9 and 21-23.
-
-\begin{equation}
-\bar \epsilon^{( {EC_0})} = \bar r_1 - \bar r_1^{\#} + \bar r_2^{\#},
-\end{equation}
-
-\noindent where $ \bar r_1, \bar r_1^{\#}$ and $\bar r_2^{\#}$ are in the symmetric case. In our model,
-
-\begin{equation}
-\bar \epsilon^{( {EC_0})} = \EX(\ln[1-\delta+\delta \exp(u_1)]) - \EX(\ln[1-\delta+\delta \exp(u_1^{\#})]) + \EX(\ln[1-\delta+\delta \exp(u_2^{\#})]),
-\end{equation}
-
-\noindent where $u_1 \sim N(\mu_1-\mu_2, 2\sigma^2-2\rho)$, $u_1^{\#} \sim N(\mu_1-\mu_2, 2\sigma^2)$, and $u_2^{\#} \sim N(0, 2\sigma^2)$
-
-\noindent *Note that the contribution of asymmetry to storage effects is not the same as the storage effect that we computed above for asymmetriv cases.
-
-In generalized terms, this further decomposition could be applied to test for contributions of asymmetry between any two variables.
-Here we tested the asymmetry between $b_1$ and $b_2$.
-Why did we not look at E and C?
-Do they have to be normally distrubuted?
-
-This break down would only work if the variables are normally distributed. Thus one can convert their distribution to normal to impliment this decomposition.
-
 \clearpage

### 2021-09-03 paper/Paper.Rnw
edits

@@
 \usepackage{graphicx}
-\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
+\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
 \usepackage{url}
@@
 We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
-plays a role.
-
-For pedgogical clarity, we develop each step of the theory in general and for a simple example model, the lottery model.
-
+plays a role. For pedgogical clarity, we develop each step of the theory in general and for 
+a simple example model, the lottery model. 
 The lottery model is as follows. 
@@
 Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
-
 quantify the contributions of 
@@
 for the lottery model, using an approach and notation
-similar to those of \cite{Ellner_2016;Ellner_2019}; see those papers for additional details.
+similar to those of \cite{Ellner_2016,Ellner_2019}; see those papers for additional details.
 
@@
 \includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
-\caption{The contribution of ATAs to species coexistence is meaningfully large compared to 
+\caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
 other components of the MTC decomposition. All panels show the IGR and the decomposition of
@@
 coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
-were never greater than \Sexpr{round(se3,5)}, so are not displayed. \label{fig:fig3}}
+were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
+analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) was generated 
+using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
+in detail, main conclusions were the same. \label{fig:fig3}}
 \end{figure}
@@
 used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
-using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section ).}\label{fig:fig4}
+using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
 \end{figure}

### 2021-09-03 paper/SupMat.Rnw
edits

@@
 \SweaveOpts{concordance=TRUE}
-\SweaveOpts{concordance=TRUE}
 
@@
 \maketitle
+
+\tableofcontents
+\listoffigures
 
@@
 
+\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
+
+%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/fig3_qij.pdf}
+\caption[Analogous to Fig. \ref{MT-fig:fig3} but making a different choice for $q_{ij}$]
+{Analogous to Fig. \ref{MT-fig:fig3} of the main text, but using 
+$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. See the caption of
+that figure for details. \label{fig:fig3_alt}}
+\end{figure}
+%***DAN: Jasmin, in the caption of Fig. 3 in the main text there is a statment about standard
+%error. We need another such statement here, but for the standard errors that applied here.
+
 \clearpage

### 2021-09-03 paper/Paper.Rnw
edits, done for now with Methods and Results

@@
 
-%G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
-%from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
-%are likely to be negligible or meaningfully large.
-
+%G2
 Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
@@
 $\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
-% and $r_1-q_{12}r_2$?
 against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
-values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed that, for the 
+values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed 
+that, for the 
 lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
@@
 
-%G3) To explore, using our simple model, the likelihood that species can coexist with
-%ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions biodiversity maintenance 
-%may be negligible or meaninfully large.
-
+%G3
 To acheive goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
@@
 
-%***DAN: Some version of the below text ought to go in the figure caption, do that later, Dan
-%The spaces in between the lines where they intersect the x-axis at y=0, indicate either an ATA rescue trait space or an ATA exclusion trait space.
-If the line plotting IGR is to the left of the line plotting IGR without ATA, then this space is an ATA rescue trait space.
-If the line plotting IGR is to the right of the line plotting IGR without ATA, then this space is an ATA exclusion trait space.
-
 %Ideas while driving to pick up Max:
@@
 
-%G2 and related side results
+%G2 
 Fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
@@
 can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
-make the difference between coexistence and exclusion: both ``ATA rescue'' and ``ATA
+make the difference between coexistence and exclusion; both ``ATA rescue'' and ``ATA
 exclusion'' (see Methods) are possible.
@@
 Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
-indicate the IGR when the effects of ATAs are included, and blue lines are the IGR when
-the effects of ATAs are neglected. These two lines frequently crossed the $x$ axis in 
-different places, showing that both ATA rescue space and ATA exclusion space could be non-empty
+represent the IGR when the effects of ATAs are included, and blue lines are the IGR when
+the effects of ATAs are removed. These two lines frequently crossed the $x$ axis in 
+different places, showing that both ATA rescue space and ATA exclusion space can be non-empty
 subsets of our trait-difference space (see Methods). For the parameters we considered, 
@@
 more common for larger $\delta$. Although bands of ATA rescue or exlusion on Fig.
-\ref{fig:fig4} appear narrow, they nevertheless do occur,
-
-Althoug ATA rescue or exclusion
-occurs only for the minority of parameter combinations considered, this does not mean
-ATAs are unimportant for species coexistance and biodiversity maintenance - the same is
-probably also true of
-
- Occurence of ATA rescue is slim and only occurs when $\delta=1$ and species traits are extremely similar or identical.
-Other $\delta$ values are more indicative of ATA exclusion events.
-When $\delta<1$ ATA exlusion space increase in range and magnitude as $\sigma$ increases.
-There is indication that ATA exlcusion space also increases in magnitude as $\delta$ decreases, however changes in range have a more complex trend.
-Range increases across decreasing $\delta$ for all $\sigma$ values except $\sigma=6$.
-Here, the range is limited by an increasing $\Delta_i^{[E \nat C]}$ as trait space increases and the two IGR's are close to converging.
+\ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
+difference between coexistence and extirpation of one species in some cases.
 
@@
 using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
-in detail, main conclusions were the same. \label{fig:fig3}}
+in some details, main conclusions were the same. \label{fig:fig3}}
 \end{figure}

### 2021-09-03 code/Figure3.R
fixed typo in saving maxse for qij version

@@ -138,6 +138,6 @@ mtext(expression(sigma), outer=TRUE, side=1, line=1, cex.lab=1.3, at=midx)
 dev.off()
 fig3qijmaxse <- max(unlist(lapply(res, function(X){X$Dq_se})), na.rm = TRUE)
 cat("maximum standard error in figure three (qij) is", fig3qijmaxse, "\n(M=", M, ")\n")
-fig3qijmaxse_loc <- paste(numeric_results_loc, "/fig3maxse.RDS", sep="")
+fig3qijmaxse_loc <- paste(numeric_results_loc, "/fig3qijmaxse.RDS", sep="")
 saveRDS(fig3qijmaxse, file=fig3maxse_loc)
 #Sys.time()

### 2021-09-03 paper/SupMat.Rnw
basic idea for section 7 (qij not matter for fig4) and put max se in caption of fig 3 alternate

@@
 rho <- readRDS(file="../results_numeric/rho.RDS")
+se3qij <- readRDS(file="../results_numeric/fig3qijmaxse.RDS")
 @
@@
 
+In fig \ref{fig:fig4} we plot IGR , $\Delta_i^{[E \nat C]$, and IGR $-\Delta_i^{[E \nat C]$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.
+
+From section \ref{sect:math_simplifications}, 
+\begin{align}
+\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
+&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
+&= 0,
+\end{align}
+and
+\begin{align}
+\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
+&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
+&= 0. %\label{eq:decomp:j:EnatC}
+\end{align}
+
+Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[E \nat C]$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ are zero so
+\begin{equation}
+\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
+\end{equation}
+and 
+\begin{equation}
+\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}=\olep_i^{[E \nat C]}
+\end{equation}
+
+Similarly for IGR $-\Delta_i^{[E \nat C]$, 
+\begin{equation}
+(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]} = \olr_{i \bs i}-\olep_i^{[E \nat C]}.
+\end{equation}
+
+Thus $q_{ij}$ does not matter for anything computed for Figure \ref{fig:fig4}.
+
 \clearpage
@@
 $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. See the caption of
-that figure for details. \label{fig:fig3_alt}}
+that figure for details. Standard errors of plotted quantities 
+were never greater than \Sexpr{round(se3qij,5)}, so are not displayed. \label{fig:fig3_alt}}
 \end{figure}

### 2021-09-07 paper/Paper.Rnw
working on lit review in intro

@@
 Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
-Environmental fluctuations are becoming more influenced by extreme climatic events \citep{Hansen_2012} as extreme events increase in frequency, duration, and intensity \citep{Meehl_2004, Ummenhofer_2017}.
-Extreme events (heat-waves, floods, hurricanes, etc.) alter the means and variation of environmental fluctuations, and thus have implications for coexistence and competition.
-Competitive output and community composition of grassland plants has been shown to be impacted by heat-waves and extreme drought \citep{White_2000, White_2001, Hoover_2014}.
-Studies on the direct effect of extreme events on competing species are slim (but see above); however, the consequences of change in environmental means, which may be due to extreme events or gradual climate change in general, on coexistence is more commonly studied.
+
+Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and exent \citep{Meehl_2004, Ummenhofer_2017}.
+The rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc.
+are a product of climate change and have implications for coexistence and competition.
+Changes in means can be attributed to extreme events; however changes in means are moreso a result of gradual increases or decreases over time in an environmental variable.
+Extreme events push environmental variables to their extremes, increasing variance.
+Effects on competition and coexistence due to climate change are more widley researched through studing changes in mean environmental changes.
+For example, decreased precipitation decreased the competitive ability of dominant forest tree species \citep{Clark_2011}.
+This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity as dominant species become excluded \citep{Mariotte_2013}.
+At the same time, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
 \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
 Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
+In contrast, increased environmental variance has been less commmonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
+Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
+However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
+Morever, in regard to invasive species, increase climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
 Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
-Increased variability of environmental fluctuations is a newer consideration for studying environmental changes in comparison to changes in means (vitation); however, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
 Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
@@
 Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
-If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
+If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events.
+Both species may become more alike (correlated), an important consideration for competition and coexistence.
 For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.

### 2021-09-07 paper/Paper.Rnw
intro editing

@@
 To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model.
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader than ATA makes a difference (see figure \ref{fig:pedagog}).
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
+We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see figure \ref{fig:pedagog}).
 The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
-Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, almost X times as much....
+Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
 

### 2021-09-08 paper/Paper.Rnw
editing intro

@@
 Pearson correlation. 
-%***
-
 But each panel of figure \ref{fig:noise} nevertheless
@@
 Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
-For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overeall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
+For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
 Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
@@
 Extreme events push environmental variables to their extremes, increasing variance.
-Effects on competition and coexistence due to climate change are more widley researched through studing changes in mean environmental changes.
-For example, decreased precipitation decreased the competitive ability of dominant forest tree species \citep{Clark_2011}.
-This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity as dominant species become excluded \citep{Mariotte_2013}.
-At the same time, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+Effects on competition and coexistence due to climate change are more widley researched through studing changes in the mean of environmental variables.
+For example, decreased growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
+This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
+Conversley, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
 \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
@@
 However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
-Morever, in regard to invasive species, increase climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
+Morever, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
 Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
@@
 For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-This is a change in ATA not just from the event alone, but from the relationship between species.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environment (but see Ghosh for impact of ATA on community aggregate properties).
+A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see Ghosh for impact of ATA on community aggregate properties).
 
-To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model.
+To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model, and below (in the next paragraph), we describe a hypothetical example.
 Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see figure \ref{fig:pedagog}).
+We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
 The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
@@
 When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
-The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated.
+The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and, following our simulated example (Figure \ref{fig:pedagog}), the species persist in coexistence.
 However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
-If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental relationship may become more correlated in the higher values or right-tail associated.
+If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the higher values or right-tail associated.
 Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
-We see the possible consequences of ATA in Figure \ref{fig:pedagog}.
-In simulating their populations,  ATA can cause species to have longer periods of extreme dominance and lower growth rates overall.
-
-%Interacting environmental drivers can be asymmetrically related in the extreme values (as shown in \cite{Ghosh_2020_AER}). 
-%Environmental drivers directly influence competitive pressure experienced by populations through environmental response; competition is always a function of environmental response \citep{Chesson_2012, Ellner_2016}.
-%Thus we have implemented our tools for the study of ATAs to explore the importance of environmental relationships in their extremes on competition and coexistence. 
-%Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we continue/move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-Moreover, the lottery model within the framework of coexistence theory provides a suitable opportunity to explore relationships in their extremes and the consequences to coexistence.
-
-%If extreme events are more temporally extensive/expansive than regular environmental fluctuations, extreme events may cause the environmental response of two species who depend on temporally separated indicators of growth to become more correlated in the extremes.
-%Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
-If extreme events are more temporally extensive (last longer, have higher duration) than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (similar, correlated), an important consideration for competition and coexistence.
-For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-This is a change in ATA not just from the event alone, but from the relationship between species.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition, they do not directly consider ATA between species response caused by environment.
+Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
 
@@
 %Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
-
+We have demonstrated that ATAs exist and are common in nature \citep{Ghosh_2020_AER}, and we have elucidated how ATAs between environmental response variables might arise via extreme events.
+ATA in environmental response directly feeds into an ATA effect on the EC covariance contribution to the storage effect mechanism of coexistence.
+Thus we seek to explore the consequences of ATAs on coexistence, competition, and the maintenance of biodiversity.
 This paper has the following goals. 
@@
 Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
-and a means for understanding its importance for real systems.
+and a means for understanding its importance for real systems. This is not a good ending sentence.
 

### 2021-09-08 paper/Paper.Rnw
added refs

@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{},
+applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020},
 %***DAN: 
@@
 A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see Ghosh for impact of ATA on community aggregate properties).
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 

### 2021-09-13 paper/SupMat.Rnw
fixed curly bracket things in section 4

@@
 
-In fig \ref{fig:fig4} we plot IGR , $\Delta_i^{[E \nat C]$, and IGR $-\Delta_i^{[E \nat C]$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.
-
+In fig \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[E \nat C]}$, and IGR $-\Delta_i^{[E \nat C]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
 From section \ref{sect:math_simplifications}, 
@@
 \end{align}
-
-Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[E \nat C]$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ are zero so
+Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[E \nat C]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ are zero so
 \begin{equation}
@@
 \end{equation}
-
-Similarly for IGR $-\Delta_i^{[E \nat C]$, 
+Similarly for IGR $-\Delta_i^{[E \nat C]}$, 
 \begin{equation}
@@
 \end{equation}
-
-Thus $q_{ij}$ does not matter for anything computed for Figure \ref{fig:fig4}.
+Thus $q_{ij}$ does not matter for anything computed for Figure \ref{MT-fig:fig4}.
 

### 2021-09-13 paper/Paper.Rnw
starting edits to Intro

@@
 %***Add to these
-\keywords{kw1, kw2,kw3}
+\keywords{competition, coexistence, biodiversity, modern theory of coexistence, storage effects, asymmetric tail association}
 
@@
 
-%Old version, prior to my edits: Central to ecology is the maintenance of biodiversity in competitive ecosystems. Populations subject to strong
-%competitive pressure may succumb to competitive exclusion - an idea which postulates if species within the same
-%trophic level and similar niches exist, one will eventually outcompete the other (Hardin, 1960). Mediation of
-%competitive pressure and exclusion allow for species to stably coexist. The criteria for coexistence and competitive
-%exclusion can be broken down into contributing mechanisms provided by modern coexistence theory, a framework
-%which seeks to assess and quantify mechanisms that result in coexisting species and thus the maintenance of
-%biodiversity (Chesson, 1994).
-%Now the new, edited version, with some additional changes requested, to be made by Jasmin, please:
-The maintenance of biodiversity in competitive ecosystems is a research topic at the absolute core of modern ecology \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
-%***DAN: Jasmin, add a few classic references. See the first sentence of Ellner et al 2019, read the abstracts of those papers
-%(more than the abstracts if you want), and consider using them as references here.
-The competitive exclusion principle, covered, for instance, in most or all introductory ecology courses and texts, states 
-that two species competing for the same limiting resource cannot coexist at constant densities - one will outcompete 
-the other and drive it extinction \citep{Hardin_1960}.
+%Intro to competition, coexistence, MCT, etc.
+Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
+\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
+The competitive exclusion principle states 
+that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
+the other and drive it to extinction \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
-framework for understanding how stable species coexistence 
-nevertheless occurs, and biodiversity in maintained, in real ecosystems, through stabilizing mechanisms and niche 
+framework for understanding how species nevertheless coexist, and biodiversity in maintained, in real ecosystems. 
+Coexistence occurs through stabilizing factors and niche 
 differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
-MCT breaks down coexistence into the contributions of various mechanisms, with name such as `relative nonlinearity
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as `relative nonlinearity
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to quite a few empirical systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020},
-%***DAN: 
-%Please also read the abstract of the Angert reference to make sure it is appropriate to cite it in this way (and so that
-%you become aware of that reference). One place to look for more references to put in for Y and Z is the top of the 
-%right column on the first page of the Barabas review (Ecol Monogr, 2018).
-even though it is mathematically complex in the eyes
-of most ecologists (but see \cite{Ellner_2016,Ellner_2019} for efforts to make it more readily applied).
-%***DAN: Jasmin, note the differences carefully, please, between your version (commented, above) and the edits I've 
-%made. It's the same in main structure, making all the same points. It's just a bit more fleshed out. One of the purposes
-%of an Intro is to link the reader to the appropriate background lit. You need a fair number of references and a little
-%more detail than you previously gave to do that.
+applied to multiple systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
+Although the MTC is mathematically complex to most ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
 
-%Old version, some comments interspersed: 
-%Equally central to ecology is the study of relationships between ecological variables; however, it is almost exclusively blanketed by simple measures of associativity, i.e., correlation coefficients.
-%***DAN: "Blanketed" is a weird verb in this context. 
-%***DAN: Also, MCT is a conceptually deep theory, and ATA is a fairly simple observation that happens to have some consequences.
-I would not be confortable with phrasing that seems to put them on the same footing!
-%Although two sets of variables may have equal correlation coefficients, they can show different association structures (see Figure).
-%More specifically, the association within the tail ends of data and the asymmetry between the tail associations, termed asymmetric tail associations (ATAs), tell a more interesting story about the relationship between variables...and?...
-%Recent work has shown that ATAs have important considerations for ecological phenomena; although, the implications of ATAs for competition and coexistence have not yet been investigated. 
-%Asymmetry of tail associations within a dataset of variables describes uneven correlations between the tails of a dataset.
-%For a simple example, a dataset with more association in the lower (upper) half and lesser association in the upper (lower) half is termed left (right) tail associated. 
-%The application of ATAs in ecology is relatively new, and in exploration of ATAs in ecology, Gosh et al (AER) found ATAs in all eight ecological datasets they considered.
-%Thus, the existence and functioning role of ATAs in nature is reasonable to assume and question. 
-%In fact, ATAs can have influence on ecosystem service stability, Taylor’s law, and extinction risks via extreme climatic events.
-Conversely, ATA’s can be influenced by Moran Effect and Leibeig’s Law of Minimum \citep{Ghosh_2020_Eco, Ghosh_2020_AER}.
-%Despite the frequent occurrence of ATAs in ecological data, the implications of tail associations for competition and coexistence, to our knowledge, has not yet been examined; doing so is the main goal of this paper.
-%New version:
+%Intro to ATAs
 Every ecological discipline involves the study of relationships among biological and environmental variables,
-but standard methods for examining the association between variables provide only a small portion of the overall information content contained in the
+but standard methods for examining the association between two variables provide only a small 
+portion of the overall information content of the
 relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
 the three pairs of variables in figure \ref{fig:noise} all have the same
-Pearson correlation. 
-But each panel of figure \ref{fig:noise} nevertheless
-depicts a very different type of association. Other metrics of association are used, in addition to correlations
+Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
+of association are also used, in addition to correlations
 (e.g., various common forms of regression), but correlations are extremely common, and other common
-methods also provide only a small portion of the available information \citep{Nelsen2006}.
+methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
 The variables in figure \ref{fig:noise}A,C 
-%***
 exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
@@
 other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
-termed `right-tail association') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
+termed ``right-tail association'') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
 considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
-%***
-%
-our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can be mechanistically important for 
+our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
 many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
 synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
-%***DAN: Again note the similarity between the old and new versions, the new is just a bit more fleshed out, and maybe a
-%bit more precise in some of the wordings in places.
 
-%***DAN: Please combine the following two paragraphs into one single paragraph with this topic sentence:
 Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
@@
 why ATAs may be relevant to them.
-%***DAN: I have inserted some comments below that might help.
+%***
 Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
@@
 the roles of ATAs in species coexistence and the maintenance of biodiversity.
-%***DAN: I added the above sentence. Not a fully fledged "goal" but a serious thing
-%to discuss.
 Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 

### 2021-09-14 paper/Paper.Rnw
continued editing to Intro

@@
 
+%Intro to storage effects and why they might be influenced by ATAs
 Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
 have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
-why ATAs may influence one mechanism of coexistence, in particular: storage effects. We review, conceptually, what storage effects are and we explain 
-why ATAs may be relevant to them.
+why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018} and we explain why ATAs may be relevant to them.
+Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
+whichever species is currently rare; this is the characteristic that separates so-called \emph{stabilizing} mechanisms of coexistence from
+\emph{equalizing} mechanisms, which instead reduce average fitness differences between comeptitors \citep{Chesson_2000}.
+Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive covariation between the quality of the environment (E)
+for a species and the strength of competition (C) experienced by that species (called EC covariance). 
+%This is reasonable assumption for non-rare species, since 
+%a better environment leads to greater numbers which leads to more intraspecific competition, but a rare species may be essentially unaffected
+%by intraspecific competition. 
+Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
+same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
+in times when the environment is poor; environmentally insensitive adult stages and dormant stages such as seed banks are
+example mechanisms \citep{Chesson_2012}. 
+When present, these factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
+Storage effects are quantified by measuring to what extent EC covariance contributes to differences between 
+growth rates of species when they are at low ``invader'' densities, and growth rates of the abundant ``resident'' species \citep{Chesson_1994,Ellner_2016}. 
+Thus ATAs may influence storage effects because storage effects involve EC covariation. ATAs between E and C may,
+in principle, mediate the strength of storage effects, though to our knowledge this possibility has not previously been investigated.
 %***
-Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
-In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
+%Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
+%In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
 %Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
-The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
-In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
-In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low.
+%The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
+%In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
+%In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low.
 %**JAS: should we introduce invader analysis or terminology here?
-Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
-Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
-Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
-For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
-Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
-Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
-
-%***DAN: You don't need to explain why it is reasonable to suspect something, that belabors the point too much. Just say indicate that storage effects
-%involve environment-coexistence associations, and hence there is the potential for ATAs to influence storage effects.
-%Uniquely, storage effect is the only mechanism of modern coexistence theory which depends on the variance between two drivers rather than one.
-%While we aimed to investigate the importance of ATAs on coexistence in general and all the mechanisms which compose the modern theory of coexistence, it was quickly revealed that ATAs are specifically important for storage effects and not for other mechanisms of coexistence.
-%***DAN: Now that you have written this, I see it belabors the point too much. It sounds defensive to explain this early on that none of the 
-%other mechanisms are influenced. I think instead it's enough to just explain, as laid out in 
-%new, proposed topic sentence above, that storage effects are influenced by ATAs. That justifies the study, and then the reader can 
-%discover later that only storage effects are influenced. If a referee complains, I think we will be able to effectively respond after the fact. 
-%
-%***
-%We emphasize that although this paper focuses on storage effects, we do consider other mechanisms of modern coexistence theory, though we quickly realize the importance of ATAs is solely for storage effects; hence, it is still a paper generally about the importance of ATAs for coexistence because.
-%***DAN: Protesting too much, sounds defensive. Maybe we can return to this in the Discussion, after we've made our points, but it 
-%currently seems to me even that won't be necessary. We can just point out in the first para of the Discussion something like "We have learned
-%that ATAs do meaningfully influence species coexistence, and they do so through storage effects." Less defensive.
-
-%***DAN: This explanation of storage effects is too long. See if you can cut it to approximately half its current length.
-
-%***DAN: stopped here 2021 07 20
+%Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
+%Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
+%Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
+%For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
+%Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
+%Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
 

### 2021-09-14 paper/Paper.Rnw
continuing edits to Intro

@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
-Although the MTC is mathematically complex to most ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+Although the MTC is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
 
@@
 other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
-termed ``right-tail association'') portions of the variables' distributions. Although ATAs and related ideas had previously been very seldom 
-considered in ecology and related areas (but see \citep{Anderson_2019, deValpine_2014, Popovic_2019}), 
+termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
+and related ideas had previously been very seldom 
+considered in ecology and related areas (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
 our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
@@
 why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018} and we explain why ATAs may be relevant to them.
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
 Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
-whichever species is currently rare; this is the characteristic that separates so-called \emph{stabilizing} mechanisms of coexistence from
-\emph{equalizing} mechanisms, which instead reduce average fitness differences between comeptitors \citep{Chesson_2000}.
-Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive covariation between the quality of the environment (E)
+whichever species is currently rare; this is the characteristic that make storage effects a so-called \emph{stabilizing} mechanism 
+of coexistence \citep{Chesson_2000}.
+Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
+covariation between the fluctuating quality of the environment (E)
 for a species and the strength of competition (C) experienced by that species (called EC covariance). 
@@
 same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
-in times when the environment is poor; environmentally insensitive adult stages and dormant stages such as seed banks are
-example mechanisms \citep{Chesson_2012}. 
-When present, these factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
-Storage effects are quantified by measuring to what extent EC covariance contributes to differences between 
-growth rates of species when they are at low ``invader'' densities, and growth rates of the abundant ``resident'' species \citep{Chesson_1994,Ellner_2016}. 
-Thus ATAs may influence storage effects because storage effects involve EC covariation. ATAs between E and C may,
-in principle, mediate the strength of storage effects, though to our knowledge this possibility has not previously been investigated.
-%***
+in times when the environment is poor; dormant stages such as are in seed banks are an
+example buffering mechanisms \citep{Chesson_2012}. 
+When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
+Importantly, storage effects are, in fact, quantified by measuring to what extent EC covariance contributes to differences between 
+invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
+Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, so ATAs 
+between between E and C may mediate the strength of storage effects.
+%***Old text below, kept for now, but should be pretty much replaced by the above
 %Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
@@
 %Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
-%Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population growth rate in both directions. 
+%Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population 
+%growth rate in both directions. 
 %Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
@@
 %Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
+
+%***DAN: I think the above paragraphs are pretty solid. I am still working on the below.
 
@@
 
+%***DAN: Jasmin, there are a lot of disjointed sentences in the below paragraph, at least they appear disjointed to me.
+%Can you please say in a comment right below this one what the topic sentence is? As usual, if you are having trouble 
+%determining what the topic sentence it, it may mean you don't have one, and you need to determine what you're doing 
+%with this paragraph and make a topic sentence. 
 Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and exent \citep{Meehl_2004, Ummenhofer_2017}.
@@
 
-Extreme events can have important ecological impacts for controphic competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
+%***DAN: I am also having a hard time identifying the topic sentence of this paragraph, can you tell me which one it is, please,
+%in a comment right under this one?
+Extreme events can have important ecological impacts for controphic 
+%***DAN: I don't know what "controphic" means and when I google it I get nothing. Perhaps you intended a different word?
+competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
 If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events.
@@
 
-%move this paragraph to discussion
-%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
-%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
 We have demonstrated that ATAs exist and are common in nature \citep{Ghosh_2020_AER}, and we have elucidated how ATAs between environmental response variables might arise via extreme events.
@@
 
+%Need something like the following, maybe
+%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
+%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
+%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
+%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+
 I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 

### 2021-09-17 paper/Paper.Rnw
typos

@@
 %with this paragraph and make a topic sentence. 
-Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and exent \citep{Meehl_2004, Ummenhofer_2017}.
+Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
 The rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc.
 are a product of climate change and have implications for coexistence and competition.
-Changes in means can be attributed to extreme events; however changes in means are moreso a result of gradual increases or decreases over time in an environmental variable.
+Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual increases or decreases over time in an environmental variable.
 Extreme events push environmental variables to their extremes, increasing variance.
-Effects on competition and coexistence due to climate change are more widley researched through studing changes in the mean of environmental variables.
+Effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables.
 For example, decreased growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
 This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
-Conversley, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
 \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
 Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
-In contrast, increased environmental variance has been less commmonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
+In contrast, increased environmental variance has been less commonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
 Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
 However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
-Morever, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
+Moreover, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
 Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.

### 2021-09-17 paper/Paper.Rnw
parahraph 4 of intro
topic sentence for paragraph 5

@@
 %with this paragraph and make a topic sentence. 
+%***JAS: I am trying to relay the message that climate change and extreme events have consequences for coexistence and competition
+% Maybe a better way to start this paragraph is:
+% EC covariance, and thus the storage effect, may be altered by changes in the environment. 
 Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
-The rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc.
-are a product of climate change and have implications for coexistence and competition.
+Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc.
+implications for coexistence and competition.
 Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual increases or decreases over time in an environmental variable.
@@
 %in a comment right under this one?
-Extreme events can have important ecological impacts for controphic 
+%**JAS: I intended for the first sentence to be the topic sentence 
+% a better one:
+% Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
+Extreme events can have important ecological impacts for
 %***DAN: I don't know what "controphic" means and when I google it I get nothing. Perhaps you intended a different word?
+% can cut out 
 competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
@@
 A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in gerneral (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 

### 2021-09-17 paper/Paper.Rnw
edits to para 4

@@
 %with this paragraph and make a topic sentence. 
-%***JAS: I am trying to relay the message that climate change and extreme events have consequences for coexistence and competition
+%***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition
 % Maybe a better way to start this paragraph is:
@@
 Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
-Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc.
-implications for coexistence and competition.
-Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual increases or decreases over time in an environmental variable.
-Extreme events push environmental variables to their extremes, increasing variance.
-Effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables.
-For example, decreased growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
+Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
+Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
+Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables, thus less likely to be a result of extreme events.
+For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
 This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
@@
 \cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
-Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as rainfall levels are forecasted to decrease.
+Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
 In contrast, increased environmental variance has been less commonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.

### 2021-09-17 paper/Paper.Rnw
comments for paragraph 4
changes made to Intro in meeting with Jasmin

@@
 between between E and C may mediate the strength of storage effects.
-%***Old text below, kept for now, but should be pretty much replaced by the above
-%Storage effects allow for species to fluctuate over time while escaping extinction, continously recovering from low densities. 
-%In order for storage effects to contribute as a stabilizing mechanism to coexistence, there must first be covariation between competitive pressures and environmental fluctuations [EC covariance \citep{Chesson_1994, Barabas_2018}].
-%Because the contributive value of storage effects relies on the relationship between environment and competition \citep{Chesson_1994, Barabas_2018}, it reasonably follows to suspect the contribution of storage effectsto be altered by an association between environment and competition that is asymmetric, i.e., ATAs between environment and competition.
-%The reliance of storage effect's magnitude on the degree of relationship between environment and competition indicate that storage effects involve environment-coexistence associations, and hence there is potential for ATAs to influence storage effects.
-%In addition to EC covariance, functioning temporal storage effects also require that heterospecific populations respond differently to environmental fluctuations, and different life stages must also vary in responses \citep{}.
-%In a simple two species system, differences in environmental response allow for one species to grow to high densities in its favorable environemnt while the other is very low.
-%**JAS: should we introduce invader analysis or terminology here?
-%Growth is stable for the high density species due to positive EC covariance -- high intraspecific competition limits its growth rate in a favorable environment. 
-%Conversely, EC covariance is lesser for the low density species, leading to large fluctuations in population 
-%growth rate in both directions. 
-%Although, this may appear like a scenario for extiction, it can, in fact, be advantageous due to differences between life stages, buffering species against extinction by limiting negative growth rates, but not positive ones.
-%For example, insensitive adults or dormants seeds can still persist when in an unfavorable environment and overall growth rate is negative \citep{Chesson_2012}; environment and competiton do not effect the buffered life stages.
-%Thus, averaged over multiple generations, higher fluctuation in growth rate can lead to a positive low density growth rate and coexistence is possible \citep{Chesson_2000}.
+
+%***DAN: Jasmin to edit the below as follows:
+%1) Topic sentence: Another potentially important reason to quatify the importance of ATA to species coexistence is that there may
+%be reason to believe that climate change can alter ATAs of environmental variables implicated in coexistence, as follows. 
+%2) Just adapt below text, maybe try to simplify a bit.
+%3) This para could just focus on the mechanics of how ATAs change for two variables measured at different times. 
+%***DAN: Keeping this sentence so Jasmin can put it somwhere, not here.
 %Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
-
-%***DAN: I think the above paragraphs are pretty solid. I am still working on the below.
-
-%[As we know], anthropogenic climate change threatens almost all aspects of life including via extreme climatic events, which have only become more extreme - increasing in magnitude and frequency - and have contributed to both the increase in global means as well as an increase in environmental variability (citation).
-%There are many studies on the influence of environmental extremes and variability on competition and coexistence \citep{Adler_2006,  Parepa_2013, Diez_2012}; however, the study of the influence of relationships and interactions between drivers in their extremes on competition and coexistence, is lacking in the scientific literature.
-%;we review these studies here  In one of the first empirical tests for storage effect and its requirements, \cite{Adler_2006} showed that climate variability - including extreme events - promoted the coexistence of grass species in Kansas prairie, specifically their ability to recover from low densities.
-Moreover, in studies on the invasibility of plants, the density and competitive balance of invasive Japanese knotweed increases as resource and thus climate variability increased.\citep{Parepa_2013}
-
-%***DAN: Jasmin, there are a lot of disjointed sentences in the below paragraph, at least they appear disjointed to me.
-%Can you please say in a comment right below this one what the topic sentence is? As usual, if you are having trouble 
-%determining what the topic sentence it, it may mean you don't have one, and you need to determine what you're doing 
-%with this paragraph and make a topic sentence. 
-%***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition
-% Maybe a better way to start this paragraph is:
-% EC covariance, and thus the storage effect, may be altered by changes in the environment. 
-Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
-Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
-Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
-Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables, thus less likely to be a result of extreme events.
-For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
-This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
-Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-\cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
-Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
-In contrast, increased environmental variance has been less commonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
-Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
-However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
-Moreover, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
-Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
-Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
-Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
-This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
-
 %***DAN: I am also having a hard time identifying the topic sentence of this paragraph, can you tell me which one it is, please,
@@
 Extreme events can have important ecological impacts for
-%***DAN: I don't know what "controphic" means and when I google it I get nothing. Perhaps you intended a different word?
-% can cut out 
 competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
@@
 A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
+%***DAN: Leave the following for the next para
 Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 
-To demonstrate plausibility of ATA altering coexistence between species, we carried out long time scale simulations using a simple two-species lottery model, and below (in the next paragraph), we describe a hypothetical example.
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
-The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
-Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
-
+%***DAN:
+%1) Topic sentence: To make the above ideas more concrete, we review a relevant example from the literature that may be
+%influenced as described above. 
+%2) Make it concrete - name the system, name the organisms, name the seasons. 
 To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
@@
 
+%***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition
+% Maybe a better way to start this paragraph is:
+% EC covariance, and thus the storage effect, may be altered by changes in the environment. 
+% need an indication that we are more interested in increased variance and extreme events because we are looking we are looking for changes in a co-varying relationship??
+%***DAN: Possible topic sentence like thing after talking to Jasmin on 9/17: 
+%Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied,
+%the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, 
+%have been studied in several systems. 
+%***DAN: Do it like you have it but shorter.
+Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
+Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
+%Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here... 
+Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
+Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables, thus less likely to be a result of extreme events.
+For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
+This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
+Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+\cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
+Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
+%...and variance...
+In contrast, increased environmental variance has been less commonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
+Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
+However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
+Moreover, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
+%...and extreme events (though this blends together with changes in variance)
+Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
+Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
+This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
+%DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
+%(means, varinace, and extreme events).
+
+%Move up the lottery model para from Theory to here, suture up the rift that this will produce in theory
+To further demonstrate plausibility of ATA altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model.
+We introduce the lottery model and describe simulations based on it that clearly show that ATAs can influence
+species coexistence, though the details of this influence are the subject of the rest of the paper.
+%Lotetry model intro
+%very succinct description of sims you do (mostly it's in the figure caption)
+%1-2 sentences on the differences observed
+lottery model, and below (in the next paragraph), we describe a hypothetical example.
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
+We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
+The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
+Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
+
+%Goal para
 We have demonstrated that ATAs exist and are common in nature \citep{Ghosh_2020_AER}, and we have elucidated how ATAs between environmental response variables might arise via extreme events.
@@
 I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
+%The second one is the one to cite
 %2) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.

### 2021-09-20 paper/Paper.Rnw
new para4: mechanics of extreme events and ATA

@@
 %in a comment right under this one?
-%**JAS: I intended for the first sentence to be the topic sentence 
-% a better one:
 % Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
-Extreme events can have important ecological impacts for
-competing communities where species growth and the maintenance of biodiversity depends on distinct temporal events/dynamics.
-If extreme events are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (correlated), an important consideration for competition and coexistence.
-For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
-%***DAN: Leave the following for the next para
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
-
+Furthermore, an additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
+The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events, i.e., heat-waves, floods, hurricanes, fires, etc.
+Temporally extensive extreme events can have important ecological impacts for
+competing populations where species growth and the maintenance of biodiversity is dependent of phenologically separated environmental signals.
+A temporally extensive event can equalize the phenologically separated environmental signals, making those phenological signals more correlated.
+For example, consider two precipitations signals, one in early spring and one in late summer.
+If a long flood occurs, it's likely that both signals will be close to equal in that year.
+If extreme floods occur in more years than not, then the two signal will become more correlated in higher precipitation values, a right-tailed ATA.
+Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA.
+The same can be said about spatially separated signals and more spatially extensive extreme events.
+ATA of environmental signals can then introduce or change ATAs in the fitness or traits of species that depend on the signals and potentially affect their coexistence.
 %***DAN:
@@
 Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
+
+are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events.
+Both species may become more alike (correlated), an important consideration for competition and coexistence.
+For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
+A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
+%***DAN: Leave the following for the next para
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 

### 2021-09-20 paper/Paper.Rnw
desert annual as example

@@
 %***DAN:
-%1) Topic sentence: To make the above ideas more concrete, we review a relevant example from the literature that may be
-%influenced as described above. 
+%1) Topic sentence: 
+To make the above ideas more concrete, we review a relevant example from the literature that may be
+influenced as previously described. 
+We introduce a simplified example of an Arizona desert annual plant community from \cite{Chesson_2012}. 
 %2) Make it concrete - name the system, name the organisms, name the seasons. 
-To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, we introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
 The community follows the criteria required for temporal storage effects to operate.
+Different species dominate and have high germination rates depending on rainfall timing.
+For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season.
+Eriogonum abertianum dominates is sufficient rainfall is reached in the late germination season.
+Competition is influenced by resource consumption of seed-producing adult plants.
+Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment.
+When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year.
+The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated.
+However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
+If the community is disturbed by frequent and long floods with some years of normalcy, the desert annual response relationship may become more correlated in the higher values or right-tail associated.
+Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the desert annual environmental response relationship may become more correlated in the lower values or left tail associated.
+
+To make the above ideas more concrete, we review a relevant example from the literature that may be
+influenced as previously described. 
+We introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
+%2) Make it concrete - name the system, name the organisms, name the seasons. 
+To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, The community follows the criteria required for temporal storage effects to operate.
 The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.

### 2021-09-21 paper/Paper.Rnw
editing intro

@@
 Furthermore, an additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
-The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events, i.e., heat-waves, floods, hurricanes, fires, etc.
+The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
 Temporally extensive extreme events can have important ecological impacts for
@@
 %***DAN: Do it like you have it but shorter.
-Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}, influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}.
-Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
+Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied, the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, have been studied in several systems.
+Environmental fluctuations are directly impacted by climate change through changes in means and increased variance, both of which can be attributed to extreme events \citep{Hansen_2012}.
+Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables. 
+Increased variance is less studied. 
+% thus less likely to be a result of extreme events. 
+%influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}. 
+%Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here... 
 Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
-Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables, thus less likely to be a result of extreme events.
+
 For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
@@
 To further demonstrate plausibility of ATA altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model. 
+Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, 
+hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. 
 We introduce the lottery model and describe simulations based on it that clearly show that ATAs can influence 
 species coexistence, though the details of this influence are the subject of the rest of the paper.
-%Lotetry model intro
+The lottery model is as follows.
+Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
+model equations are
+\begin{equation}
+N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
+\end{equation}
+for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
+The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+in proportion to the reproductive outputs of the two species that year.
+For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
+time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
+that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
+However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
+properties, corresponding to various tail association cases. Note that $N$ is constant through time.
 %very succinct description of sims you do (mostly it's in the figure caption)
 %1-2 sentences on the differences observed
-lottery model, and below (in the next paragraph), we describe a hypothetical example.
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence.
-We simulated two different communites up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
-The only distinction between the two simulations is that one depends on asymmterically tail associated temporal dynamics and one does not.
-Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverablility when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
+We simulated two different communities up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
+The only distinction between the two simulations is that one depends on asymmetrically tail associated temporal dynamics and one does not.
+Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverability when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
 
@@
 plays a role. For pedgogical clarity, we develop each step of the theory in general and for 
-a simple example model, the lottery model. 
-The lottery model is as follows. 
+a simple example model, the lottery model (see \ref{section:introduction}).
+
 %***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
-Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
-model equations are
-\begin{equation}
-N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
-\end{equation}
-for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
-The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
-in proportion to the reproductive outputs of the two species that year.
-For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
-time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
-that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
-However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
-bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
-properties, corresponding to various tail association cases. Note that $N$ is constant through time.
 
@@
 
+Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
+
 \clearpage

### 2021-09-22 paper/Paper.Rnw
editing lit review para-mean shifts

@@
 Furthermore, an additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
-The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
+The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
 Temporally extensive extreme events can have important ecological impacts for
@@
 %***DAN: Do it like you have it but shorter.
+%the RAMPAGE of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
 Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied, the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, have been studied in several systems.
-Environmental fluctuations are directly impacted by climate change through changes in means and increased variance, both of which can be attributed to extreme events \citep{Hansen_2012}.
-Currently, effects on competition and coexistence due to climate change are more widely researched through studying changes in the mean of environmental variables. 
-Increased variance is less studied. 
-% thus less likely to be a result of extreme events. 
-%influenced by extreme climatic events as they increase in frequency, intensity, duration, and extent \citep{Meehl_2004, Ummenhofer_2017}. 
-%Thus, the rampage of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
+Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}.
+While both changes can be attributed to extreme events, shifts in mean are more-so a result of gradual trend effects, and increased variance is a result of extreme event effects \citep{Jentsch_2007} .
+Impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables; impacts of increased variance is less studied, although the rate of study has increased.
+\citep{Jentsch_2007}  %importance has been realized and variability is becoming more commonly studied now
+
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
-Changes in means can be attributed to extreme events; however changes in means are more-so a result of gradual shifts over time, whereas increased variance is more likely a product of extreme events as they push environmental variables to their extremes.
-
-For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}.
-This may be detrimental if the dominant species is crucial to the ecosystem or it may not be as rare species can increase in diversity while dominant species become excluded \citep{Mariotte_2013}.
-Conversely, changes in environmental means can increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-\cite{Descombes_2020}, through experimental manipulation, found that range shifts, as a product of climate change, favored coexistence between alpine plants while \cite{Usinowicz_2021} predict increases in competitive pressures as the climate change induced range shifts lead populations to interact with new conspecifics.
-Similarly, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
+For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and as such, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
+On the contrary, dominant species exclusion may be beneficial as rare species can increase in diversity\citep{Mariotte_2013}.
+Changes in environmental means can also increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+As a result of changing means, range shifts can favor plant coexistence by introducing new herbivores\cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
 %...and variance...
@@
 However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
+\cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
 Moreover, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.

### 2021-09-22 paper/Paper.Rnw
edit to lit review

@@
 \citep{Jentsch_2007}  %importance has been realized and variability is becoming more commonly studied now
-
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
 For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and as such, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
-On the contrary, dominant species exclusion may be beneficial as rare species can increase in diversity\citep{Mariotte_2013}.
+On the contrary, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}.
 Changes in environmental means can also increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
@@
 %...and variance...
-In contrast, increased environmental variance has been less commonly studied; however, more research on the effect of increased variance has been urged and received recent focus through the study of extreme climatic events \citep{Jentsch_2007, Smith_2011}.
-Extreme drought and heat waves impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
+Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentch_2007}, the most commonly manipulated extremes being temperature and rainfall.
+Extreme drought and heat waves have shown impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
 However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
 \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
-Moreover, in regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
+In regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
 %...and extreme events (though this blends together with changes in variance)
-Extreme events constitute large disturbances, quantitatively amounting to large fluctuations in environmental drivers and their ecological responses \citep{Smith_2011}.
-Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes; environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
+Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
 Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.

### 2021-09-23 paper/Paper.Rnw
some more ediing to intro

@@
 % Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
-Furthermore, an additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
-The consequence of climate change that has the most potential to influence ATAs in the environment the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
+An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
+Among the various implications of climate change, the most potentially influential to ATAs in the environment is the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
 Temporally extensive extreme events can have important ecological impacts for
-competing populations where species growth and the maintenance of biodiversity is dependent of phenologically separated environmental signals.
-A temporally extensive event can equalize the phenologically separated environmental signals, making those phenological signals more correlated.
-For example, consider two precipitations signals, one in early spring and one in late summer.
+competing populations where species growth and the maintenance of biodiversity is dependent on phenologically separated environmental signals.
+A temporally extensive event can equalize the phenologically separated environmental signals, making them more correlated.
+For example, consider two precipitations signals, one in early spring and one in late summer, that are usually different.
 If a long flood occurs, it's likely that both signals will be close to equal in that year.
-If extreme floods occur in more years than not, then the two signal will become more correlated in higher precipitation values, a right-tailed ATA.
-Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA.
+If extreme floods occur in more years than not, then the two signals will become more correlated in higher precipitation values, a right-tailed ATA (figure \ref{fig:noise}C).
+Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A).
 The same can be said about spatially separated signals and more spatially extensive extreme events.
@@
 For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season.
-Eriogonum abertianum dominates is sufficient rainfall is reached in the late germination season.
+Eriogonum abertianum dominates if sufficient rainfall is reached in the late germination season.
 Competition is influenced by resource consumption of seed-producing adult plants.
-Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment.
+Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment.
 When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year.
-The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated.
-However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
-If the community is disturbed by frequent and long floods with some years of normalcy, the desert annual response relationship may become more correlated in the higher values or right-tail associated.
-Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the desert annual environmental response relationship may become more correlated in the lower values or left tail associated.
+The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated...
 
@@
 influenced as previously described. 
-We introduce an example of a two species grass-forb plant community (inspired by \cite{Chesson_2012}).
-%2) Make it concrete - name the system, name the organisms, name the seasons. 
-To contextualize idea of how ATA in environmental response might manifest and modfiy the outcome of coexistence, The community follows the criteria required for temporal storage effects to operate.
-The forb dominates and has high fecundity when a sufficient threhold of rain is reached in the early season.
-The grass dominates and has high fecundity when there is sufficeint rain in the late season.
-Competition is influenced by resource comsumption of growing juvenile plants.
-Dormant seeds are unaffected by competiton and can persist in the community in an unfavorable environment.
+We introduce a hypothetical example of the commonly studied two species grass-forb plant community \citep{Chesson_2012, Hallett_2019}.
+The community follows the criteria required for temporal storage effects to operate.
+The forb dominates and has high fecundity when a sufficient threshold of rain is reached in the early season.
+The grass dominates and has high fecundity when there is sufficient rain in the late season.
+Competition is influenced by resource consumption of growing juvenile plants.
+Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment.
 When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
-The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and, following our simulated example (Figure \ref{fig:pedagog}), the species persist in coexistence.
+The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and the species persist in coexistence...
+
 However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
-If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the higher values or right-tail associated.
-Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb environmental response relationship may become more correlated in the lower values or left tail associated.
-Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
+If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb/desert annual
+environmental response relationship may become more correlated in the higher values, a right-tailed ATA (figure \ref{fig:noise}C).
+Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb/desert annual environmental response relationship may become more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A).
+The subsequent potential changes in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA). 
+%Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
 
-are more temporally extensive than moderate disturbances, then it is more likely that two species will be influenced by the same events.
-Both species may become more alike (correlated), an important consideration for competition and coexistence.
-For example, an extreme flood event that lasts across the growth season of two different species and increases recruitment in both species will cause recruitment of the two species to become more correlated in the upper tail end of the distribution.
-A subsequent change in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
 %***DAN: Leave the following for the next para
@@
 % need an indication that we are more interested in increased variance and extreme events because we are looking we are looking for changes in a co-varying relationship??
-%***DAN: Possible topic sentence like thing after talking to Jasmin on 9/17: 
-%Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied,
-%the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, 
-%have been studied in several systems. 
+
 %***DAN: Do it like you have it but shorter.
 %the RAMPAGE of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
+%***DAN: Possible topic sentence like thing after talking to Jasmin on 9/17: 
 Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied, the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, have been studied in several systems.
@@
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
-For example, decreased mean growth season precipitation can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and as such, \cite{diPaola_2012} predict the collapse of coexistence between Mediterranean tree species as mean rainfall levels are forecasted to decrease.
+Decreased mean growth season precipitation, for instance, can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and coexistence is predicted to collapse between Mediterranean tree species in accordance to forecasted mean rainfall level decrease \cite{diPaola_2012}.
 On the contrary, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}.

### 2021-09-24 paper/Paper.Rnw
some more edits to intro

@@
 The subsequent potential changes in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
-%Again, we see the possible consequences of ATA in Figure \ref{fig:pedagog}, i.e., less or no coexistence.
-
-%***DAN: Leave the following for the next para
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 
@@
 While both changes can be attributed to extreme events, shifts in mean are more-so a result of gradual trend effects, and increased variance is a result of extreme event effects \citep{Jentsch_2007} .
-Impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables; impacts of increased variance is less studied, although the rate of study has increased.
-\citep{Jentsch_2007}  %importance has been realized and variability is becoming more commonly studied now
+Impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables. 
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
@@
 %...and variance...
-Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentch_2007}, the most commonly manipulated extremes being temperature and rainfall.
-Extreme drought and heat waves have shown impact competitive output and community composition of grassland plants\citep{White_2000, White_2001, Hoover_2014}.
+Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
+Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}.
 However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
 \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
-In regard to invasive species, increased climate and resource variability allowed for greater success of invasive plants in terms of density and competitive equilibrium \citep{Parepa_2013}.
+For invasive species, increased climate and resource variability allowed for greater success of invasibility in terms of density and competitive equilibrium \citep{Parepa_2013}.
 %...and extreme events (though this blends together with changes in variance)
 Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
-Moreover, increased variability has been found to promote coexistence in plants communities \citep{Adler_2006,  Angert_2009}, but disrupt coexistence in birds \citep{Stenseth_2002, Stenseth_2015}.
+Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the storage effect \citep{Adler_2006,  Angert_2009}.
 This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 %DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
@@
 %1-2 sentences on the differences observed
-We simulated two different communities up to \Sexpr{M} time steps to hopefully make it clear to the reader that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog}).
+We simulated two different communities up to \Sexpr{M} time steps to hopefully make it clear to the that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog} for more details).
 The only distinction between the two simulations is that one depends on asymmetrically tail associated temporal dynamics and one does not.

### 2021-09-24 paper/Paper.Rnw
word and reference

@@
 % Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
-An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables implicated in coexistence, as follows.
+An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables required in coexistence, as follows.
 Among the various implications of climate change, the most potentially influential to ATAs in the environment is the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.

### 2021-09-24 paper/Paper.Rnw
small changes to intro

@@
 Temporally extensive extreme events can have important ecological impacts for
-competing populations where species growth and the maintenance of biodiversity is dependent on phenologically separated environmental signals.
-A temporally extensive event can equalize the phenologically separated environmental signals, making them more correlated.
-For example, consider two precipitations signals, one in early spring and one in late summer, that are usually different.
+competing populations where species growth and the maintenance of biodiversity is dependent on
+phenologically separated environmental signals, for instance, equalizing the signals, making them more correlated. 
+Consider two precipitations signals, one in early spring and one in late summer, that are usually different. 
 If a long flood occurs, it's likely that both signals will be close to equal in that year. 
@@
 %1) Topic sentence:
+
 To make the above ideas more concrete, we review a relevant example from the literature that may be
@@
 Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}.
-While both changes can be attributed to extreme events, shifts in mean are more-so a result of gradual trend effects, and increased variance is a result of extreme event effects \citep{Jentsch_2007} .
-Impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables. 
+While both changes can be attributed to extreme events, shifts in mean are more-so a result of gradual trend effects, and increased variance is a result of extreme event effects \citep{Jentsch_2007}.
+However, impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables. 
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
 Decreased mean growth season precipitation, for instance, can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and coexistence is predicted to collapse between Mediterranean tree species in accordance to forecasted mean rainfall level decrease \cite{diPaola_2012}.
-On the contrary, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}.
-Changes in environmental means can also increase fitness of a dominant species and in turn reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-As a result of changing means, range shifts can favor plant coexistence by introducing new herbivores\cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
+Alternatively, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+As a product of changing means, range shifts can favor plant coexistence by introducing new herbivores \cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
 %...and variance...
 Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
-Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}.
-However, complete exclusion of dominant species due to extreme events is uncommon \citep{Lloret_2012}.
+Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although, complete exclusion of dominant species is uncommon \citep{Lloret_2012}.
 \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
@@
 %...and extreme events (though this blends together with changes in variance)
-Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence \citep{Chesson_1981, Warner_1985}.
-Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the storage effect \citep{Adler_2006,  Angert_2009}.
+Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e.
+MCT \citep{Chesson_1981, Warner_1985}.
+Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the aforementioned storage effect \citep{Adler_2006,  Angert_2009}.
 This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general (see above), they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
+Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general, they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
 %DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
@@
 
-%Move up the lottery model para from Theory to here, suture up the rift that this will produce in theory
+%Move upl the lottery model para from Theory to here, suture up the rift that this will produce in theory
 To further demonstrate plausibility of ATA altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model. 
@@
 the roles of ATAs in species coexistence and the maintenance of biodiversity.
-Overall, our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
-and a means for understanding its importance for real systems. This is not a good ending sentence.
+Our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
+and a means for understanding its importance for real systems. Need better ending sentence.
 

### 2021-09-24 paper/Paper.Rnw
editted eexample paragraph and split para 6 in two

@@
 
-To make the above ideas more concrete, we review a relevant example from the literature that may be
-influenced as previously described. 
-We introduce a simplified example of an Arizona desert annual plant community from \cite{Chesson_2012}. 
-%2) Make it concrete - name the system, name the organisms, name the seasons. 
-The community follows the criteria required for temporal storage effects to operate. 
-Different species dominate and have high germination rates depending on rainfall timing. 
-For example, Baileya multiradiata dominates if sufficient rainfall is reached in the early germination season.
-Eriogonum abertianum dominates if sufficient rainfall is reached in the late germination season.
-Competition is influenced by resource consumption of seed-producing adult plants.
-Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment.
-When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the desert annual community also fluctuate year to year.
-The result is a relationship between the desert annuals' environmental responses (germination) that is symmetrically associated...
-
-To make the above ideas more concrete, we review a relevant example from the literature that may be
-influenced as previously described. 
-We introduce a hypothetical example of the commonly studied two species grass-forb plant community \citep{Chesson_2012, Hallett_2019}.
-The community follows the criteria required for temporal storage effects to operate.
-The forb dominates and has high fecundity when a sufficient threshold of rain is reached in the early season.
-The grass dominates and has high fecundity when there is sufficient rain in the late season.
-Competition is influenced by resource consumption of growing juvenile plants.
-Dormant seeds are unaffected by competition and can persist in the community in an unfavorable environment.
-When rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the grass-forb community also fluctuate year to year.
-The result is a relationship between the grass's and forb's environmental responses (fecundity) that is symmetrically associated, and the species persist in coexistence...
-
-However, if environmental responses become asymmetrically associated by, say, extreme rainfall events, coexistence can be altered.
-If the community is disturbed by frequent and long floods with some years of normalcy, the grass-forb/desert annual
-environmental response relationship may become more correlated in the higher values, a right-tailed ATA (figure \ref{fig:noise}C).
-Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the grass-forb/desert annual environmental response relationship may become more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A).
+To make the above ideas more concrete, we review a relevant example that may be influenced as previously described. 
+Imagine a community where different species fitness depended on phenologically separated environmental signals.
+For instance, \cite{Chesson_2012}, provides two examples: an Australian two species grass-forb community and an Arizona desert community of multiple desert annual species.
+The communities follow the criteria required for temporal storage effects to operate.
+In both examples, species have different periods of germination, competition is influenced by resource consumption, and growth is buffered by overlapping generations or via seed banks in the annual case.
+Temporal differences in germination rates suggest the possibility that competition and growth may be influenced by a temporally fluctuating environmental variable, such as rainfall, 
+such that germination of different species is contingent on rainfall signals at different times during a growth season, 
+and the timing of a rainfall event predicts maximal germination and dominance of one species. 
+If that is the case, one might expect that when rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the communities also fluctuate year to year. 
+The result is a relationship between the species' environmental responses (germination) that is symmetrically associated, and the species persist in coexistence. 
+However, if environmental responses become asymmetrically associated by extreme rainfall events, coexistence can be altered. 
+If the community is disturbed by frequent and long floods with some years of normalcy, the environmental response relationship may become more correlated in the higher values, a right-tailed ATA (figure \ref{fig:noise}C).
+Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the environmental response relationship may become more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A).
 The subsequent potential changes in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
@@
 %***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition
-% Maybe a better way to start this paragraph is:
 % EC covariance, and thus the storage effect, may be altered by changes in the environment. 
@@
 As a product of changing means, range shifts can favor plant coexistence by introducing new herbivores \cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
+
 %...and variance...

### 2021-09-28 paper/Paper.Rnw

@@
 %...and variance...
-Although, the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
-Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although, complete exclusion of dominant species is uncommon \citep{Lloret_2012}.
+While the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
+Despite research on climate variability in relation to competition and coexistence being only a portion of that, we review some studies here.
+Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although complete exclusion of dominant species is uncommon \citep{Lloret_2012}.
 \cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
@@
 %...and extreme events (though this blends together with changes in variance)
-Although, increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e.
+Although increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e.
 MCT \citep{Chesson_1981, Warner_1985}.

### 2021-09-28 paper/Paper.Rnw
edited through para 5 of the Intro

@@
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
-framework for understanding how species nevertheless coexist, and biodiversity in maintained, in real ecosystems. 
+framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
 Coexistence occurs through stabilizing factors and niche 
@@
 in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
-applied to multiple systems, including annual plants \citep{Angert_2009, Hallett_2019}, perennial plants \citep{Adler_2010, Adler_2006}, 
+applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
@@
 termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
-and related ideas had previously been very seldom 
-considered in ecology and related areas (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
+and related ideas had previously been seldom 
+considered in ecology (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
 our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
@@
 Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
-have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. It is straightforward to see, intuitively, 
+have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
+It is straightforward to see, intuitively, 
 why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
@@
 Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
-whichever species is currently rare; this is the characteristic that make storage effects a so-called \emph{stabilizing} mechanism 
+whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
 of coexistence \citep{Chesson_2000}.
@@
 When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
-Importantly, storage effects are, in fact, quantified by measuring to what extent EC covariance contributes to differences between 
+Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
 invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
-Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, so ATAs 
-between between E and C may mediate the strength of storage effects.
+Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
+occur between environment and competition variables, can be an aspect of EC covariation.
 
-%***DAN: Jasmin to edit the below as follows:
-%1) Topic sentence: Another potentially important reason to quatify the importance of ATA to species coexistence is that there may
-%be reason to believe that climate change can alter ATAs of environmental variables implicated in coexistence, as follows. 
-%2) Just adapt below text, maybe try to simplify a bit.
-%3) This para could just focus on the mechanics of how ATAs change for two variables measured at different times. 
-%***DAN: Keeping this sentence so Jasmin can put it somwhere, not here.
-%Given that storage effect depends on the relationship between E and C, we question the consequences of EC relationships characterized by ATA, the presence of which may be increasing as climate change brings about more extreme climatic events.
-%***DAN: I am also having a hard time identifying the topic sentence of this paragraph, can you tell me which one it is, please,
-%in a comment right under this one?
-% Extreme events can have important ecological impacts for storage effect and coexistence by introducing ATAs. 
-An additional driving reason to quantify the significance of ATAs for species coexistence is the potential for climate change to alter ATAs of environmental variables required in coexistence, as follows.
-Among the various implications of climate change, the most potentially influential to ATAs in the environment is the increase in intensity, duration, extent, and frequency of extreme climatic events \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}, i.e., heat-waves, floods, hurricanes, fires, etc.
-Temporally extensive extreme events can have important ecological impacts for
-competing populations where species growth and the maintenance of biodiversity is dependent on
-phenologically separated environmental signals, for instance, equalizing the signals, making them more correlated. 
-Consider two precipitations signals, one in early spring and one in late summer, that are usually different. 
-If a long flood occurs, it's likely that both signals will be close to equal in that year. 
-If extreme floods occur in more years than not, then the two signals will become more correlated in higher precipitation values, a right-tailed ATA (figure \ref{fig:noise}C). 
-Similarly, if extreme droughts occur in more years than not, then the signals will be more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A). 
-The same can be said about spatially separated signals and more spatially extensive extreme events. 
-ATA of environmental signals can then introduce or change ATAs in the fitness or traits of species that depend on the signals and potentially affect their coexistence.
-%***DAN:
-%1) Topic sentence:
+%Climate change may alter ATAs of coexistence-relevant ATAs
+An additional reason to quantify the importance of ATAs for species coexistence is that there is reason to believe climate change may
+alter ATAs of environmental variables which impact coexistence, as follows.
+It is well known that climate change can or may alter the intensity, duration, extent and frequency of extreme environmental events, including
+heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}.  
+If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
+may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
+For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
+whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
+contributing to their coexistence, or failure to coexist, can boil down to covariation 
+between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
+to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
+longer lasting increases the degree of right-tail association between early- and late-spring rainfall. 
+Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
+left-tail association between these variables. As we will show in this study, changes in ATAs can then
+result in different competitive outcomes. Although we focus here on temporal coexistence
+mechanisms, ATAs between environmental variables measured in different places may also be altered 
+by climate change, if more intensive climatic events are also more spatially extensive, potentially 
+influencing spatial coexistence mechanisms.
 
-To make the above ideas more concrete, we review a relevant example that may be influenced as previously described. 
-Imagine a community where different species fitness depended on phenologically separated environmental signals.
-For instance, \cite{Chesson_2012}, provides two examples: an Australian two species grass-forb community and an Arizona desert community of multiple desert annual species.
-The communities follow the criteria required for temporal storage effects to operate.
-In both examples, species have different periods of germination, competition is influenced by resource consumption, and growth is buffered by overlapping generations or via seed banks in the annual case.
-Temporal differences in germination rates suggest the possibility that competition and growth may be influenced by a temporally fluctuating environmental variable, such as rainfall, 
-such that germination of different species is contingent on rainfall signals at different times during a growth season, 
-and the timing of a rainfall event predicts maximal germination and dominance of one species. 
-If that is the case, one might expect that when rain fall dynamics temporally fluctuate from year to year, relative abundance and dominance of the communities also fluctuate year to year. 
-The result is a relationship between the species' environmental responses (germination) that is symmetrically associated, and the species persist in coexistence. 
-However, if environmental responses become asymmetrically associated by extreme rainfall events, coexistence can be altered. 
-If the community is disturbed by frequent and long floods with some years of normalcy, the environmental response relationship may become more correlated in the higher values, a right-tailed ATA (figure \ref{fig:noise}C).
-Conversely, if the community is disturbed by frequent and long drought with some years of normalcy, the environmental response relationship may become more correlated in the lower values, a left-tailed ATA (figure \ref{fig:noise}A).
-The subsequent potential changes in competitive ability and coexistence is not just attributed to the extreme event alone, but from the resulting extreme relationship between species (ATA).
+%make the above abstract ideas more concrete
+To make the above ideas more concrete, we briefly mention examples from prior work to 
+which the ideas may apply. 
+\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
+are likely influenced differently by phenologically distict environmental variables: an Australian 
+grass-forb community, and an Arizona community of desert annuals. 
+The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
+examples, distinct species have different periods of germination, competition is influenced by resource 
+consumption, and population growth is buffered by overlapping generations or via seed banks.
+Temporal differences between species in germination rates suggest the possibility that 
+population growth of distinct species may be driven by rainfall signals in distinct parts
+of the growing season. Thus all the ingredients seem to be present by which climatic changes may
+alter ATAs between environmental variables which are relevant for the coexistence (or not)
+of species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
 

### 2021-09-28 paper/Paper.Rnw
continued editing Intro

@@
 of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-alter ATAs between environmental variables which are relevant for the coexistence (or not)
+alter ATAs between environmental variables which are relevant for the coexistence
 of species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
@@
 %***DAN: Possible topic sentence like thing after talking to Jasmin on 9/17:
-Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied, the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, have been studied in several systems.
+
+Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+not been studied, the importance for coexistence of other aspects of climate change, including changes 
+in means and variability of environmental variables, has been studied in several systems, so we briefly 
+review some of that related literature.
+For instance, decreased growth season precipitation can decrease the competitive ability of dominant 
+forest tree species \citep{Clark_2011}, and coexistence is predicted to collapse between Mediterranean 
+tree species under forecasted mean rainfall scenarios \cite{diPaola_2012}.
+
+Alternatively, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
+
+As a product of changing means, range shifts can favor plant coexistence by introducing new herbivores \cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
+
 Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}.
@@
 %Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
-Decreased mean growth season precipitation, for instance, can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and coexistence is predicted to collapse between Mediterranean tree species in accordance to forecasted mean rainfall level decrease \cite{diPaola_2012}.
-Alternatively, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-As a product of changing means, range shifts can favor plant coexistence by introducing new herbivores \cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
 

### 2021-09-29 paper/Paper.Rnw
finished editing for now, through the Results, but no Discussion yet

--- 2021-09-28 paper/Paper.Rnw
@@
 
-\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
+\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research and Kansas Biological Survey, University of Kansas}
 \affil[b]{Laboratory of Populations, Rockefeller University}
@@
 %***Add to these
-\keywords{competition, coexistence, biodiversity, modern theory of coexistence, storage effects, asymmetric tail association}
-
-%Here are the paragraph-level "rules" of writing, for reference: 
-%Every paragraph needs to have a clear topic sentence.
-%It needs to be one of the first three sentences.
-%If it is not the first sentence, then the sentences preceding it are for background, in order to make it possible to state the topic sentence.
-%The topic sentence is the main *result* or *claim* of the paragraph, not just a statement of what the topic is.
-%All sentences in the paragraph support and validate the claim made in the topic sentence.
+\keywords{asymmetric tail association, biodiversity, coexistence, competition, modern coexistence theory, storage effects}
 
@@
 differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as `relative nonlinearity
-in competition' and `storage effects' \citep{Chesson_1994} which will be familiar to many readers. The MCT has been
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
+in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
-Although the MTC is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
 
@@
 relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
-the three pairs of variables in figure \ref{fig:noise} all have the same
+the three pairs of variables in Fig. \ref{fig:noise} all have the same
 Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
-of association are also used, in addition to correlations
-(e.g., various common forms of regression), but correlations are extremely common, and other common
+of association are also often used, in addition to correlations
+(e.g., various forms of regression), but correlations are extremely common, and other common
 methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
-The variables in figure \ref{fig:noise}A,C 
+The variables in Fig. \ref{fig:noise}A,C 
 exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
 they are more strongly related to each
-other in the left (figure \ref{fig:noise}A, termed ``left-tail association'') or right (figure \ref{fig:noise}C,
+other in the left (Fig. \ref{fig:noise}A, termed ``left-tail association'') or right (Fig. \ref{fig:noise}C,
 termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
@@
 for a species and the strength of competition (C) experienced by that species (called EC covariance). 
-%This is reasonable assumption for non-rare species, since 
-%a better environment leads to greater numbers which leads to more intraspecific competition, but a rare species may be essentially unaffected
-%by intraspecific competition. 
 Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
 same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
-in times when the environment is poor; dormant stages such as are in seed banks are an
-example buffering mechanisms \citep{Chesson_2012}. 
+in times when the environment is poor. Dormant stages such as seed banks are an
+example buffering mechanism \citep{Chesson_2012}. 
 When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
@@
 
+%Lottery model and results based on it
+To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
+simulations using a simple two-species model, the lottery model. The simulations described in this
+paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
+precise nature of that influence is then the subject of the rest of the paper.
+The lottery model is as follows.
+Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
+model equations are
+\begin{equation}
+N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
+\end{equation}
+for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
+The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
+For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically distributed (iid) through
+time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
+that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
+We do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
+bivariate normal distribution. 
+For proof of concept, we simulated two communities for \Sexpr{M} time 
+steps each, using symmetric and asymmetric tail associations (Fig. \ref{fig:noise}) between $b_1$ and $b_2$, respectively, in the
+two simulations. Fig. \ref{fig:pedagog} gives details of the 
+simulations and shows that ATAs substantially influenced aspects of species coexistence for this model setup.
+
 %Climate change may alter ATAs of coexistence-relevant ATAs
-An additional reason to quantify the importance of ATAs for species coexistence is that there is reason to believe climate change may
+An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
 alter ATAs of environmental variables which impact coexistence, as follows.
@@
 left-tail association between these variables. As we will show in this study, changes in ATAs can then
-result in different competitive outcomes. Although we focus here on temporal coexistence
+result in different competitive outcomes between species. Although we focus here on temporal coexistence
 mechanisms, ATAs between environmental variables measured in different places may also be altered 
-by climate change, if more intensive climatic events are also more spatially extensive, potentially 
-influencing spatial coexistence mechanisms.
+by climate change, if more intensive climatic events are also more spatially extensive.
 
@@
 of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-alter ATAs between environmental variables which are relevant for the coexistence (or not)
-of species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
+alter ATAs between environmental variables which are relevant for the coexistence
+of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
 
-%***JAS: I am trying to relay the message that climate change and specifically extreme events have consequences for coexistence and competition
-% EC covariance, and thus the storage effect, may be altered by changes in the environment. 
-% need an indication that we are more interested in increased variance and extreme events because we are looking we are looking for changes in a co-varying relationship??
+%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
+Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+not been studied, the importance for coexistence of other aspects of climate change, including changes 
+in means and measures of variability of environmental variables, has been studied in many systems, so we briefly 
+review some of that large, related literature.
+For instance, \cite{diPaola_2012} concluded that coexistence of drought sensitive,
+flood resistant species with flood sensitive, drought resistant species in 
+Mediteranean coastal swamp forests of central Italy would be jeopardized
+by projected changes in mean rainfall for the region. In an interesting example
+involving climate-change effects on herbivore-mediated competition in plants,
+\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
+higher altitudes of herbivore populations than of plant populations, newly arrived
+herbivores at high altitudes can inflate species richness by reducing dominant plant species 
+in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
+alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
+all species they considered, but that changes in competition would more than offset these direct benefits,
+and, as a result, some species would be extirpated. Numerous studies also take an ``event-focussed'' 
+perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
+form of extreme weather events, can alter competitive interactions. For instance, 
+\cite{White_2001} showed that single extreme heating events can alter competition 
+between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
+nutrient availability for plants, which is mediated by soil moisture content, can alter 
+the potential for invasive species to enter an ecosystem and establish, a process that involves
+the invader overcoming competition from resident species. 
+Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
+communities, via storage effects \citep{Adler_2006,  Angert_2009}.
+Changes in variability and in extreme events are distinct from the changes in ATAs we 
+consider here, because ATAs pertain to \emph{relationships} between environmental and other
+variables in the extremes, which are distinct from and can be altered independently
+of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
-%***DAN: Do it like you have it but shorter.
-%the RAMPAGE of extreme events such as heat-waves, floods, hurricanes, fires, etc. implications for coexistence and competition. 
-%***DAN: Possible topic sentence like thing after talking to Jasmin on 9/17: 
-Although, to our knowledge, the potential importance of climatic changes in ATAs on coexistence has not been studied, the importance for coexistence of other aspects of climate change, including changes in means and variability of environmental variables, have been studied in several systems.
-Environmental fluctuations are directly impacted by climate change through changes in means and increased variance \citep{Hansen_2012}.
-While both changes can be attributed to extreme events, shifts in mean are more-so a result of gradual trend effects, and increased variance is a result of extreme event effects \citep{Jentsch_2007}.
-However, impacts on competition and coexistence of climate change have been more commonly researched via studying changes in the mean of environmental variables. 
-%Jasmin tried to split this para up into effects on coexistence of changes in mean, starting here...
-Decreased mean growth season precipitation, for instance, can decrease the competitive ability of dominant forest tree species \citep{Clark_2011}, and coexistence is predicted to collapse between Mediterranean tree species in accordance to forecasted mean rainfall level decrease \cite{diPaola_2012}.
-Alternatively, dominant species exclusion may be beneficial as rare species can increase in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-As a product of changing means, range shifts can favor plant coexistence by introducing new herbivores \cite{Descombes_2020} or disfavor coexistence through increased competition \cite{Usinowicz_2021}.
-
-%...and variance...
-While the body of literature on the impact of increased environmental variability has expanded, event-based studies accounted for only one-fifth of experimental climate change research as of 2006 \citep{Jentsch_2007}, the most commonly manipulated extremes being temperature and rainfall.
-Despite research on climate variability in relation to competition and coexistence being only a portion of that, we review some studies here.
-Extreme drought and heat waves have shown to impact competitive output, rearranging species composition and dominance of grass-forb communities \citep{White_2000, White_2001, Hoover_2014}; although complete exclusion of dominant species is uncommon \citep{Lloret_2012}.
-\cite{Fay_2003} experimentally manipulated rainfall for more frequent and intense precipitation while keeping constant the total volume of rain over Kansas tallgrass prairie and found that while some species had reduced fitness, overall community structure remained the same.
-For invasive species, increased climate and resource variability allowed for greater success of invasibility in terms of density and competitive equilibrium \citep{Parepa_2013}.
-%...and extreme events (though this blends together with changes in variance)
-Although increased variability of environmental fluctuations is a newer consideration for studying environmental changes, environmental variability has been analytically shown to be important for classical coexistence, i.e.
-MCT \citep{Chesson_1981, Warner_1985}.
-Moreover, by using MCT, increased variability has been shown to promote coexistence in plants communities, via the aforementioned storage effect \citep{Adler_2006,  Angert_2009}.
-This present study has applications to extreme events and climate change as we consider relationships in their extremes, potentially caused by extreme events, and their implications for competition and coexistence.
-Though previous literature considers impacts of extreme events, change in means, and increased variability on species coexistence and competition and community dynamics in general, they do not directly consider ATA between species response caused by environmental changes -- extreme relationships caused by extreme events (but see \cite{Ghosh_2021} for impact of ATA on community aggregate properties).
-%DAN: So what jasmin was trying to do was summarize a lot of lit on effects of climate change on coexistence, in these three parts
-%(means, varinace, and extreme events).
-
-%Move upl the lottery model para from Theory to here, suture up the rift that this will produce in theory
-To further demonstrate plausibility of ATA altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model. 
-Environmental variability has been classically shown to promote coexistence via storage mechanisms in a temporal lottery model of competition \citep{Chesson_1981, Warner_1985}, 
-hence we move forward with this simple model to examine the effect of extreme relationships between environmental responses on the contribution of storage effects to coexistence. 
-We introduce the lottery model and describe simulations based on it that clearly show that ATAs can influence 
-species coexistence, though the details of this influence are the subject of the rest of the paper.
-The lottery model is as follows.
-Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
-model equations are
-\begin{equation}
-N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
-\end{equation}
-for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
-The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
-in proportion to the reproductive outputs of the two species that year.
-For each $i$, we assume the random variables $B_i(t)$ are independent and identically distributed (iid) through
-time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
-that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
-However, we do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
-bivariate normal distribution. We will consider various distributions of $\left(b_1, b_2\right)$ with the above 
-properties, corresponding to various tail association cases. Note that $N$ is constant through time.
-%very succinct description of sims you do (mostly it's in the figure caption)
-%1-2 sentences on the differences observed
-We simulated two different communities up to \Sexpr{M} time steps to hopefully make it clear to the that ATA makes a difference for competition and coexistence between species (see Figure \ref{fig:pedagog} for more details).
-The only distinction between the two simulations is that one depends on asymmetrically tail associated temporal dynamics and one does not.
-Markedly, the simulation with ATA has longer periods of one-species dominance than the simulation without, implying less recoverability when rare and thus less coexistence \citep{Chesson_2000} when environmentally dependent variables are asymmetrically tail associated.
+%***DAN: Jasmin, I commented the below sentence because that body of work is famously 
+%complex and sophisticated, and I did not feel we could do it justice in short space.
+%decreased growth season precipitation can decrease the competitive ability of dominant 
+%forest tree species \citep{Clark_2011}, 
+%***DAN: Jasmin, I commented the below sentence because I'm not sure the two studies 
+%cited there are actually directly about coexistence, in the MCT sense. One of them is
+%a general review on invasive species, and the other has to do with competition,
+%but coexistence is not mentioned. I searched both papers for "coexist" and found
+%no hits. OK, both are relevant, but maybe not so directly, to what we are talking about,
+%and since this paragraph could anyway stand to be shortened a bit, I took these out.
+%Alternatively, dominant species exclusion may be beneficial as rare species can increase 
+%in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can 
+%reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
 
 %Goal para
-We have demonstrated that ATAs exist and are common in nature \citep{Ghosh_2020_AER}, and we have elucidated how ATAs between environmental response variables might arise via extreme events.
-ATA in environmental response directly feeds into an ATA effect on the EC covariance contribution to the storage effect mechanism of coexistence.
-Thus we seek to explore the consequences of ATAs on coexistence, competition, and the maintenance of biodiversity.
-This paper has the following goals. 
-G1) To augment the modern theory of coexistence (MTC) to show how ATA plays a role. 
-The MTC decomposes a measure of coexistence into components due to each of several mechanisms. 
+We previously demonstrated \citep{Ghosh_2020_AER} that ATAs exist and are common in nature, and we argued above 
+that it is plausible both that climate change may alter environmental ATAs, and that ATAs may influence species 
+coexistence. Thus we seek to explore the consequences of ATAs for competition, coexistence, and the maintenance of 
+biodiversity, by means of the following specific goals. 
+G1) To augment MCT to show formally how ATA plays a role in coexistence. 
+MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
 We address how ATA contributes to these components. 
 We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
-G2) To explore, using a simple model, the relative magnitudes of the contributions to coexistence that come
-from ATAs compared to other mechanisms. This is to help illuminate whether ATA contributions to coexistence 
+G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
+from ATAs compared to other, known mechanisms of coexistence. This is to help illuminate whether ATA contributions to coexistence 
 are likely to be negligible or meaningfully large. 
-%G3) To identify and define the ATA rescue or ATA exlusion set of trait space range for which species will coexist or become competitively excluded in our model. 
-%***DAN: This is a good additional goal. We will either have to have already defined the "ATA rescue and ATA exlusion sets of trait space"
-%or else readers won't understand this as it was written, above (now commented). These are complex concepts, so it may be better to 
-%save formal definitions for later and reword this. Here's a try at rewording: 
 G3) To explore, using our simple model, whether species can coexist with
@@
 may be negligible or meaninfully large.
-G4) To provide an empirical example of species coexistence, and to highlight the role ATAs play in coexistence for the example.
-In the Discussion, we revisit the potential that climatic changes will influence
-the roles of ATAs in species coexistence and the maintenance of biodiversity.
-Our study presents a new mechanism of coexistence and biodiversity maintenance which may be influenced by climate change, 
-and a means for understanding its importance for real systems. Need better ending sentence.
+G4) To provide an empirical example of species coexistence which highlights the role ATAs can play.
+In the Discussion, we revisit the possibility that climate change will alter
+ATAs of environmental variables and thereby influence competition and species coexistence.
+Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
+and a means of understanding its theoretical and importance and its importance for real ecosystems.
 
@@
 
-We here pursue goal G1 from the Introduction, to augment the MTC to show how ATA
-plays a role. For pedgogical clarity, we develop each step of the theory in general and for 
-a simple example model, the lottery model (see \ref{section:introduction}).
-
-%***DAN: Note the inclusion of the model here means we will have to edit the "Model" section of sup mat, since we don't need to intro the model twice. Do later.
-
-Modern coexistence theory \citep{Chesson_2000} and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
+plays a role in coexistence. For pedgogical clarity, we develop each step of our new theory for a general 
+two-species model, and for the lottery model (Introduction).
+MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
 quantify the contributions of 
-different mechanisms to species coexistence. But we show below
+multiple mechanisms to coexistence. But we show below
 that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-We now define storage effects for a general two species model, and specifically 
-for the lottery model, using an approach and notation
-similar to those of \cite{Ellner_2016,Ellner_2019}; see those papers for additional details.
+So we begin by defining storage effects for a general two species model and 
+for the lottery model. Our mathematical notation parallels
+that of \cite{Ellner_2016,Ellner_2019}.
 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. As for the lottery model example, we here assume, for simpicity, 
+competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simpicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
@@
 
-In general, the storage effect for species $i$ is the contribution of covariation between $E$ and 
+For a general model, the storage effect for species $i$ is the contribution of covariation between $E$ and 
 $C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
@@
 We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SI-SIsect:scaling_factors}.
-%***DAN: Say this here if it turns out to be true, or say something like it:
-%These alternatives turn out to make very little difference to our results. 
 Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
@@
 (\ref{eq:storage_effects_further}) is a decompostion within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effects, which are themselves a term within the decomposition provided by the MTC \citep{Ellner_2019} -- see below for details.
+of storage effects, which are themselves a term within the decomposition provided by MCT \citep{Ellner_2019} -- see 
+below for additional details.
 
@@
 marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
-$(b_i,b_j)$ itself need not be. Given these definitions, it is straightforward to compute the 
+$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
 component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further1}) for the lottery model using the simulation 
@@
 
-The other mechanisms in modern coexistence theory that apply in our modelling context of purely temporal variation
-make no reference to covariation between quantities, whereas asymmetries of tail associations are 
+The other mechanisms in MCT that apply in our modelling context of purely temporal variation
+make no reference to covariation between quantities, whereas ATAs are 
 defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
-for which asymmetries of tail association are relevant. Mathematical details supporting this claim are in SI section 
+for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
 \ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
@@
 \end{equation}
-which is the extension of the MTC which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
+which is the extension of MCT which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
 the magnitude of the new ATA term in this expression, $\Delta_i^{[E \nat C]}$, relative to the magnitudes of the other terms.
@@
 
+All computations were done in R on an ordinary laptop. Complete computer codes for the project are stored at
+\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
+
 %Ideas while driving to pick up Max:
@@
 of the contributions to coexistence that come from ATAs compared to other mechanisms), Fig.
-\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
-magnitude comparable to the magnitudes of other terms of the extended MTC decomposition. 
+\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) often, but not always, has 
+magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
 Magnitudes of $\Delta_i^{[E \nat C]}$ tended to be larger, relative to the magnitudes of the 
@@
 
-I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
-where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
-similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
-there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
-are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
+%I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
+%where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
+%similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
+%there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
+%are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
 
-Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
-via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
-study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
+%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
+%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
+%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
 
-Need a para about multi-species models - how would things change? One difference is there are more different
-things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
-just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
+%Need a para about multi-species models - how would things change? One difference is there are more different
+%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
+%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
 
-Might be interesting to study how tail associations of environmental variables are changing and how this might 
-alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
-research program which presumably would start with a more detailed study of how tail associations of environmental
-variables may be changing under global change.
+%Might be interesting to study how tail associations of environmental variables are changing and how this might 
+%alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
+%research program which presumably would start with a more detailed study of how tail associations of environmental
+%variables may be changing under global change.
 
-Search for comments in the rest of the text indicating things that need to be done in the Discussion
+%Search for comments in the rest of the text indicating things that need to be done in the Discussion
 
-Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
-mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
-the importance of ATA for coexistence mediated by these factors.
+%Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
+%mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
+%the importance of ATA for coexistence mediated by these factors.
 
-Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
-solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
-how to say something about how future work may be able to explore the importance of ATA in that context?
+%Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
+%solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
+%how to say something about how future work may be able to explore the importance of ATA in that context?
 
-Also need to talk about how this stuff will apply to data.
+%Also need to talk about how this stuff will apply to data.
 
-In the Discussion, I intend to talk about how possible changes in tail association 
-patterns of env vars in nature may influence coexistence of species. It won't influence the 
-other mechanisms, though changes in the extreme values of environmental variables, known
-to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
+%In the Discussion, I intend to talk about how possible changes in tail association 
+%patterns of env vars in nature may influence coexistence of species. It won't influence the 
+%other mechanisms, though changes in the extreme values of environmental variables, known
+%to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
 %though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
@@
 
-Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
+%Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
+
+\section{Acknowledgments}
+
+The authors thank Shyamolina Ghosh for initially suggesting the research problem, and Lawrence Sheppard, Max Castorani,
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Steve Ellner, and Robin Snyder for helpful discussions. 
+The authors were partly supported by US National
+Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, and the California Department of Fish and
+Wildlife Delta Science Program. 
+%***DAN: Jasmim, please add any other acknolwedgments here - anyone else you have discussed this with or anything like that.
+%Be generous here, add people even if they only helped a tiny bit. If Steve consents to be coauthor, we take him
+%off this list and add him to the author list.
 
@@
 \includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
-\caption{Simple example of bivariate noise with standard normal marginals. A) is left-tail 
-associated, B) is symmetric, and C) is right-tail associated noise. All three cases have 
-the same Pearson correlation, up to sampling variation, yet show different very different 
+\caption{Simple example of bivariate random variables with standard normal marginals showing
+alternative patterns of association in the tails. A) is ``left-tail 
+associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
+the same Pearson correlation, $\rho$, up to sampling variation, yet show very different 
 patterns of association. We used asymmetrically tail associated cases which are perfectly 
-correlated in the lower (A)  or upper (C) halfs of the distributions, though it is also 
+correlated in the lower (A)  or upper (C) halves of the distributions, though it is also 
 possible to generate asymmetrically tail associated noise with imperfect correlations in 
-both halves of the distributions. The variables $(b_1,b_2)$ were obtianed by generating 
-noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$ (SI 
-section \ref{SI-sect:noise}).\label{fig:noise}}
+both halves of the distributions \citep{Ghosh_2020_AER,Nelsen2006}. The variables $(b_1,b_2)$ described in
+the Introduction and Methods were obtained by generating 
+noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$.\label{fig:noise}}
 \end{figure}
@@
 \includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
-\caption{%Pedagogical figure somehow displaying the different outcomes of competition in ATA and symmetric cases for an example.
-This will be cited from the Intro, so needs to be understandable to someone who does not know any of the ideas we introduce yet.
-So in particular you cannot use IGR as your metric of coexistence, you need to use something more intuitive.
-You will pick model parameter values and do two sims, one with ATA noise and one with symmtric noise, and show different outcomes.
-We should discuss and possibly take several iterations to get this figure right since it's a pedagogically important figure - if the reader is not convinced in the Intro that we are studying something that makes a difference, they are not going to slog through the Theory section, which is inevitably going to be tough going for many readers.
-Populations inlfuenced by asymmetric tail association recover from low densities less often spend more time on average under one-species dominance.
-The top panel, A, is a simulation over time of a population within a two species lottery model whose environmental response and competition is influenced by symmetric bivariate noise defined by $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$.
-The bottom panel, B, is like A, but the population is influenced by bivariate noise with left-tailed asymmetric association defined by the same parameters.
-Both simulations were ran for total time of \Sexpr{M}.
-Noticeable population density was defined at a threshold of 0.95 of total community space.
-Periods of noticable coexistence are marked in grey.
-The community in A had mean noticable coexistence time of \Sexpr{fig2$coSym} and a mean one-species dominance time of \Sexpr{fig2$domSym}.
-The community in B had mean noticable coexistence time of \Sexpr{fig2$coATA} and a mean one-species dominance time of \Sexpr{fig2$domATA}.
-While mean periods of coexistence are more similar, community B had much longer mean periods of one-species dominace.
-The population in B also recovered from low densities less often and thus has a lower ability to recover when rare and thus that community has lower coexistence.
+\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
+spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
+species coexistence. Panel A shows species 1, from the lottery 
+model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}B). Panel B shows species 1 from the 
+same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}A).
+Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
+``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
+total abundance, are marked in grey. The average duration of periods of noticiable coexistence 
+was \Sexpr{fig2$coSym} time steps for panel A and \Sexpr{fig2$coATA} for panel B, over the full length of the simulations; 
+whereas the average duration of periods of
+one-species dominance was \Sexpr{fig2$domSym} time steps for panel A and \Sexpr{fig2$domATA} for B. Thus, while periods of 
+noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
+on average, for the ATA case. Parameters were $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, 
+%***DAN: Jasmin, at this point of the ms, the reader does not know that only the difference mu1-mu2 matters, so this
+%will confuse them. Please replace the above specfication of mu1-mu2 by a specification of which values of both
+%mu1 and mu2 were actually used for this simulation. 
+$\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$. See Introduction for model details. 
 \label{fig:pedagog}}
 \end{figure}
-%***DAN: Dan has not yet edited this caption, will do so when it is time to edit the Intro
+%***DAN: Dan has edited this caption
 
@@
 \caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
-other components of the MTC decomposition. All panels show the IGR and the decomposition of
-the IGR provided by the extended MTC, equation (\ref{eq:full_decomp}). For each panel, 
+other components of the MCT decomposition. All panels show the IGR and the decomposition of
+the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
 $\delta$ and $\mu_1-\mu_2$ 

### 2021-09-29 paper/SupMat.Rnw
finished editing for now, through the Results, but no Discussion yet

@@
 Right-tail associated noise was generated in an analogous way.
-Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
-noise was generated from the left- or right-tail association cases.
 
 When needed, noise with symmetric tail association was generated by taking a very large sample (\Sexpr{M}) 
-%***DAN: Dan, later, check whether 100000 is what was actually used. I'm thinking we could probably do a million.
 of left-tail associated noise generated as 
@@
 and then we set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
+
+Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
+noise was generated from the symmetric, left- or right-tail associated cases.
 

### 2021-09-30 paper/Paper.Rnw
spelling and grammar in intro

@@
 of association are also often used, in addition to correlations
-(e.g., various forms of regression), but correlations are extremely common, and other common
+(e.g., various forms of regression), but correlations are extremely common, and other
 methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
@@
 all species they considered, but that changes in competition would more than offset these direct benefits,
-and, as a result, some species would be extirpated. Numerous studies also take an ``event-focussed'' 
+and, as a result, some species would be extirpated. Numerous studies also take an ``event-focused'' 
 perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
@@
 biodiversity, by means of the following specific goals. 
-G1) To augment MCT to show formally how ATA plays a role in coexistence. 
+G1) To augment MCT to show formally how ATAs plays a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
@@
 ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
-may be negligible or meaninfully large.
+may be negligible or meaningfully large.
 G4) To provide an empirical example of species coexistence which highlights the role ATAs can play.
@@
 Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
-and a means of understanding its theoretical and importance and its importance for real ecosystems.
+and a means of understanding its theoretical importance and its importance for real ecosystems.
 

### 2021-09-30 paper/Paper.Rnw
spelling fixes in Theory

@@
 biodiversity, by means of the following specific goals. 
-G1) To augment MCT to show formally how ATAs plays a role in coexistence. 
+G1) To augment MCT to show formally how ATA plays a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
@@
 We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
-plays a role in coexistence. For pedgogical clarity, we develop each step of our new theory for a general 
+plays a role in coexistence. For pedagogical clarity, we develop each step of our new theory for a general 
 two-species model, and for the lottery model (Introduction).
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simpicity, 
+competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
@@
 The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
-(\ref{eq:storage_effects_further}) is a decompostion within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
+(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
 of storage effects, which are themselves a term within the decomposition provided by MCT \citep{Ellner_2019} -- see 
@@
 How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? We assume, for simplicity, that the cummulative distribution functions (CDFs) of $E_i$ and 
+correlation unchanged? We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
 $C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
@@
 %the *right* way
-For those familar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
+For those familiar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 %G3
-To acheive goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
+To achieve goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
 ATAs but not without, or vice versa, we plotted the IGR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
@@
 we were interested in the values of $\mu_1-\mu_2$ for which one of these
-quantities is positive and the other is negative; this set of value characterizes the species which can coexist
+quantities is positive and the other is negative; this set of values characterizes the species which can coexist
 with ATAs, but not without, or vice versa. The one dimensional space of values $\mu$ is a species trait space,

### 2021-09-30 paper/Paper.Rnw
finished revisions

@@
 competitors, storage effects (blue plus red
-lines on Fig. \ref{fig:fig3}) were the only means by which coexisitence could occur, and ATA effects
+lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
 tended to be negative, weakening total storage effects.
@@
 Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
-can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
+can coexist with ATAs, but not without, or vice versa), our results show that ATA can, indeed, sometimes
 make the difference between coexistence and exclusion; both ``ATA rescue'' and ``ATA
@@
 ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
-more common for larger $\delta$. Although bands of ATA rescue or exlusion on Fig.
+more common for larger $\delta$. Although bands of ATA rescue or exclusion on Fig.
 \ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
@@
 %Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
+
+%Higher sigma means a more fluctuating and variable environment.
+Then, if environmental fluctuations become more variable, the effect of ATAs becomes more important and non-negligible (rescue or exclusion).
+%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
+
+%Knowledge of stochastic structure (Shoemaker et al 2020)
 

### 2021-09-30 paper/Paper.Rnw
added acknowledgement

@@
 The authors thank Shyamolina Ghosh for initially suggesting the research problem, and Lawrence Sheppard, Max Castorani,
-Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Steve Ellner, and Robin Snyder for helpful discussions. 
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Steve Ellner, and Robin Snyder for helpful discussions. 
 The authors were partly supported by US National

### 2021-09-30 paper/Paper.Rnw
fixed mu's in fig 2 caption and spelling and added Sexpr{rho} to fig 1 caption

@@
 associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
-the same Pearson correlation, $\rho$, up to sampling variation, yet show very different 
+the same Pearson correlation, $\rho = \Sexpr{rho}$, up to sampling variation, yet show very different 
 patterns of association. We used asymmetrically tail associated cases which are perfectly 
@@
 ``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
-total abundance, are marked in grey. The average duration of periods of noticiable coexistence 
+total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
 was \Sexpr{fig2$coSym} time steps for panel A and \Sexpr{fig2$coATA} for panel B, over the full length of the simulations; 
@@
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case. Parameters were $\mu_1-\mu_2 = \Sexpr{fig2$params[1]}$, 
-%***DAN: Jasmin, at this point of the ms, the reader does not know that only the difference mu1-mu2 matters, so this
-%will confuse them. Please replace the above specfication of mu1-mu2 by a specification of which values of both
-%mu1 and mu2 were actually used for this simulation. 
-$\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$. See Introduction for model details. 
+on average, for the ATA case.
+Parameters were $\mu_1 = \Sexpr{fig2$params[1]} + \Sexpr{fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$.
+See Introduction for model details.
 \label{fig:pedagog}}
@@
 to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
-betweed the ATA line (red)
+between the ATA line (red)
 and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species

### 2021-09-30 paper/Paper.Rnw
rho not needed

@@
 M <- readRDS(file="../results_numeric/M.RDS")
-rho <- readRDS(file="../results_numeric/rho.RDS")
 se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
@@
 associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
-the same Pearson correlation, $\rho = \Sexpr{rho}$, up to sampling variation, yet show very different 
+the same Pearson correlation, $\rho$, up to sampling variation, yet show very different 
 patterns of association. We used asymmetrically tail associated cases which are perfectly 

### 2021-10-01 paper/Paper.Rnw
added abstract

@@
 
-\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research and Kansas Biological Survey, University of Kansas}
+\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
 \affil[b]{Laboratory of Populations, Rockefeller University}
@@
 \begin{abstract}
-Place abstract here
+\noindent Understanding the coexistence of competing species is a topic of very broad 
+interest in ecology. The modern coexistence
+theory (MCT) of Chesson identifies several mechanisms of species coexistence,
+one of which, storage effects, hinges on relationships between fluctuations
+in environmental and competitive pressures on species populations. However, such relationships
+have typically been measured using covariance methods, which do not account for 
+the possibility that environmental and competitive influences may be more correlated
+when these influences are strong than when they are weak, or vice versa. 
+Recent work showed that such `asymmetric tail associations' (ATA) are extremely common
+between ecological variables, and are important for a wide range of phenomena including 
+extinction risk and the stability of ecosystem functioning. 
+Here we extend the MCT to take into account the influence of ATAs. Storage
+effects can be decomposed into two components, one of which quantifies the
+influence of ATAs on species coexistence, and is a novel contribution of this study; and 
+the other of which represents
+the influence of environment-competition correlation \emph{per se}, without 
+the influence of ATAs. Simulations of a simple model
+%and an empirical example
+illustrate that ATA influences on coexistence can be comparable in magnitude to other
+mechanisms of coexistence, and that the presence or absence of ATAs can make the 
+difference between species coexisting and competitive exclusion. 
 \end{abstract}

### 2021-10-01 paper/Paper.Rnw
s in abstract

@@
 when these influences are strong than when they are weak, or vice versa. 
-Recent work showed that such `asymmetric tail associations' (ATA) are extremely common
+Recent work showed that such `asymmetric tail associations' (ATAs) are extremely common
 between ecological variables, and are important for a wide range of phenomena including 

### 2021-10-06 paper/Paper.Rnw
start of revising notation for ATA contribution start revision of notation for ATA contribution

--- 2021-10-01 paper/Paper.Rnw
@@
 [(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
-&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
+&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
 \end{align}
-The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
+The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
 ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
+The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
+correlation \emph{per se}, remaining after having removed the effects of ATAs.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
@@
 The new symbol should correspondingly be pronounced ``partial sharp''.
-The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
-having removed the effects of ATAs.
 The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
@@
 which is the extension of MCT which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
-the magnitude of the new ATA term in this expression, $\Delta_i^{[E \nat C]}$, relative to the magnitudes of the other terms.
+the magnitude of the new ATA term in this expression, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
 
@@
 as well as the
-IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$. 
+IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$. 
 For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
@@
 We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of values
-$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$ is positive.
+$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
 This is the subspace for which the two species in the lottery model coexist in the absence of 
@@
 i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
-On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$
+On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
 against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
@@
 magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
-Magnitudes of $\Delta_i^{[E \nat C]}$ tended to be larger, relative to the magnitudes of the 
+Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
 other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 

### 2021-10-06 paper/SupMat.Rnw
added abstract
start revising notationfor ATA contribution

@@
 &= \left[ \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)] \right]+\left[ \E[r_i(E_i^\nat,C_{i \bs i}^\nat)]-\E[r_i(E_i^\shp,C_{i \bs i}^\shp)] \right] \\
-&= \olep_i^{[E \nat C]}+\olep_i^{[EC]}.
+&= \olep_i^{[EC]}+\olep_i^{[E \nat C]}.
 \end{align}
@@
 tertiary decomposition, i.e., in the decomposition of $\olep_i^{(EC)}$. 
-The term $\olep_i^{[E \nat C]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
-and the the term $\olep_i^{[EC]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
+The term $\olep_i^{[EC]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
+and the the term $\olep_i^{[E \nat C]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
 The symbol $\nat$ denotes the removal of asymmetries of tail association. We therefore have 
 \begin{equation}
-\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[EC]}+\olep_i^{[E \nat C]},\label{eq:decomp_overall_IGR}
+\olr_{i \bs i} = \epsilon_i^0 + \olep_i^E + \olep_i^C + \olep_i^{(E\shp C)}+\olep_i^{[E \nat C]}+\olep_i^{[EC]},\label{eq:decomp_overall_IGR}
 \end{equation}
@@
 \begin{equation}
-\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[EC]}+\olep_j^{[E \nat C]},\label{eq:decomp_overall_rjbsi}
+\olr_{j \bs i} = \epsilon_j^0 + \olep_j^E + \olep_j^C + \olep_j^{(E\shp C)}+\olep_j^{[E \nat C]}+\olep_j^{[EC]},\label{eq:decomp_overall_rjbsi}
 \end{equation}
@@
 \begin{equation}
-\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[EC]}+\Delta_i^{[E\nat C]},\label{eq:decomp_overall_delta}
+\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]}+\Delta_i^{[EC]},\label{eq:decomp_overall_delta}
 \end{equation}
@@
 variation \emph{per se} in both environment and competition, not including the effects of covariation of these quantities;
-$\Delta_i^{[EC]}=\olep_i^{[EC]} - q_{ij} \olep_j^{[EC]}$ is the contribution to coexistence of $EC$ covariation \emph{per se},
+ $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$is the contribution to coexistence of $EC$ covariation \emph{per se},
 not including the effects of asymmetric tail associations between environment and competition;
-and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ is the contribution to coexistence of 
-asymmetric tail assocations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.
+and $\Delta_i^{[EC]}=\olep_i^{[EC]} - q_{ij} \olep_j^{[EC]}$ is the contribution to coexistence of 
+asymmetric tail associations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.
 

### 2021-10-06 paper/SupMat.Rnw
finished revision for section 1-5

@@
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))] - \E[\ln(1-\delta+\delta\exp(b_i^\shp-b_j^\shp))],\label{eq:decomp:contrib_EC_1}
 \end{equation}
@@
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
-\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
+\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]
+, \label{eq:decomp:contrib_EC}
+\end{equation}
+where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
+$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$ and $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent.
+Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_EC})
+from expression (\ref{eq:decomp:storage_effects}), i.e., 
+\begin{equation}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
+\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_2}
 \end{equation}
@@
 $(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
-Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
-from expression (\ref{eq:decomp:storage_effects}), i.e., 
-\begin{equation}
-\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
-\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
-\end{equation}
 This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.
@@
 \begin{align}
-\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
+\olep_j^{[E \nat C]} &= \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] - \E[r_j(E_j^\shp , C_{j \bs i}^\shp)]\\
+&= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
+&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
+\label{eq:decomp:j:EChard}
+\end{align}
+where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
+Continuing,
+\begin{align}
+\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
+&= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)]
 &= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
@@
 because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
-$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
-\begin{align}
-\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
-&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
-\end{align}
-where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
+$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other.
 This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.
@@
 
-Whereas $\olep_i^{[EC]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
+Whereas $\olep_i^{[E \nat C]} = \E[r_i(E_i^\nat , C_{i \bs i}^\nat )]-\E[r_i(E_i^\shp ,C_{i \bs i}^\shp)]$ depends on
 more than just marginals, this quantity is still independent of the choice between our left- and right-tail-association 
@@
 
-We now examine $\olep_i^{[E \nat C ]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
+We now examine $\olep_i^{[EC ]}$, which, by (\ref{eq:decomp:contrib_ATA_2}), equals 
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
 \end{equation}
@@
 \end{align}
-which completes the proof that $\olep_i^{[E \nat C]}$ is independent of left- versus right-tail associated noise.
+which completes the proof that $\olep_i^{[EC]}$ is independent of left- versus right-tail associated noise.
 For the terms of the decomposition (\ref{eq:decomp_overall_rjbsi}), it suffices to look at the equations for those terms, 

### 2021-10-08 paper/Paper.Rnw
discussion

@@
 
-%Need something like the following, maybe
-%Previously, storage effects were described as the contribution of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-%However, the description of storage effects as the covariance between drivers is a (potentially dangerous) misnomer in terminology, and we argue that storage effects, instead, should be described as the contribution of the relationship between drivers.
-%All aspects and contributing values of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the strucutre of associativity throuhgout the distribution of the relationship and any asymmetries there may be.
-%Thus, aligning with the decomposition presented in \cite{Ellner_2019}, we propose storage effect can be broken down further into the contribution of ATA [to storage effect/coexistence] and the contribution of symmetric association [to storage effect/coexistence].
+Outline
 
+\underline{Paragraph summarizing results of G1 and how it compares and builds on prior theory.} 
+We provided a farther tertiary decomposition to MCT, 
+extending the framework put forward by \citep{Ellner_2019}. 
+We decompose the storage effect into two new terms, 
+$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, the contribution of correlation \emph{per se}. 
+Previously, storage effects were described as the contribution
+of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
+However, the description of storage effects as the covariance between drivers
+is a misnomer in terminology, and we propose that storage effects, instead, 
+should be described as the contribution of the relationship between drivers. 
+All aspects and of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the structure of associativity throughout the distribution of the relationship and any asymmetries there may be.
+Thus, it is misleading to describe the contribution of the whole EC relationship 
+as the contribution of EC covariance.
+In addressing this inconsistency, we have separated the mechanisms of storage effect into the contribution of ATA [to storage effect/coexistence] and the contribution of correlation \emph{per se} [to storage effect/coexistence].
+
+\underline{paragraph summarizing results of G2 and how it compares and builds on prior theory}
+Furthermore, we have shown not only that is it possible to decompose storage effect into the effect of correlation \emph{per se} and the effect of ATAs, but this new term that holds the contribution of not only the correlation but the structure of the correlation is a non-negligible term, in the lottery model.
+Our result adds coexistence to the list of ecological theory/phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
+%Higher sigma means a more fluctuating and variable environment.
+Then, if environmental fluctuations become more variable, the effect of ATAs becomes more important and non-negligible (rescue or exclusion).
+%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
+
+\underline{paragraph summarizing results of G3 and how it compares and builds on prior theory}
+Not only is the contribution of ATA non-neglible but in some cases the presence of absence of ATA can mediate the result of competition between species, i.e., coexistence or extirpation. 
+Our investigation is a simple one; 
+the ATA present is a perfect correlation is exactly one half of the relationship, 
+but varying degrees of ATA may produce a similar result as well. 
+%maybe the above sentence should go in caveats/assumptions paragraph 
+ATA specialist? Does this go in this paragraph
 %I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
@@
 %are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
+
+\underline{paragraph summarizing results of G4 and how it compares and builds on prior theory}
+
+\underline{how results apply to or inform patterns or management questions in real life}
+ATA specialists?
+climate change and ATA
+
+\underline{caveats/assumptions that could change result}
+
+\underline{conclude: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
+We provided another showcase of the importance of ATAs in ecology, 
+building on previous studies.
+ATAs clearly have implications for coexistence and competition;
+though more research is needed in exploring more complex models 
+including spatiotemporal and multispecies models.
+Hope to see this tertiary decomposition of MCT be applied to empirical research.
+More investigations should be done to explore implications of ATAs 
+to other ecological theories.
 
@@
 %Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
-
-%Higher sigma means a more fluctuating and variable environment.
-Then, if environmental fluctuations become more variable, the effect of ATAs becomes more important and non-negligible (rescue or exclusion).
-%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
 

### 2021-10-08 code/Figure3.R
fix legend

@@ -45,7 +45,7 @@ plot.new() #1
 legend("topright", 
        legend=c(expression(Delta[i]^0), expression(Delta[i]^E),
                 expression(Delta[i]^C), expression(Delta[i]^"(E#C)"),
-                expression(Delta[i]^"[EC]"), expression(Delta[i]^"[E||C]")
+                expression(Delta[i]^"[E||C]"), expression(Delta[i]^"[EC]")
                 ,expression(IGR)), 
        col = c("black","black","black","black","blue","red","orange"),
        lty = c(1,2,4,3,1,1,1), bty="n", cex=1.5, inset=c(0.1,-0.042),
@@ -101,7 +101,7 @@ plot.new() #1
 legend("topright", 
        legend=c(expression(Delta[i]^0), expression(Delta[i]^E),
                 expression(Delta[i]^C), expression(Delta[i]^"(E#C)"),
-                expression(Delta[i]^"[EC]"), expression(Delta[i]^"[E||C]")
+                expression(Delta[i]^"[E||C]"), expression(Delta[i]^"[EC]")
                 ,expression(IGR)), 
        col = c("black","black","black","black","blue","red","orange"),
        lty = c(1,2,4,3,1,1,1), bty="n", cex=1.5, inset=c(0.1,-0.025),

### 2021-10-08 code/Figure4.R

@@ -45,7 +45,7 @@ layout(matrix(c(2,3,4,5,1,
 
 plot.new() #1
 legend("topright", 
-       legend=c(expression(IGR),expression(IGR-Delta[i]^"[E||C]"), expression(Delta[i]^"[E||C]")),
+       legend=c(expression(IGR),expression(IGR-Delta[i]^"[EC]"), expression(Delta[i]^"[EC]")),
        col = c("orange","navy", "red"),
        lty = c(1,1,1), bty="n", cex=1.5, inset=c(-0.25,-0.042),
        y.intersp = 1.25, x.intersp = 0.1, seg.len=1)

### 2021-10-28 paper/Paper.Rnw
sentence in discussion
fixed parameter description in fig 2 caption

@@
 Our result adds coexistence to the list of ecological theory/phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
+Seemingly consistent with fluctuation dependent mechanisms, as \sigma increased, so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
 %Higher sigma means a more fluctuating and variable environment.
@@
 on average, for the ATA case.
-Parameters were $\mu_1 = \Sexpr{fig2$params[1]} + \Sexpr{fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\delta = \Sexpr{fig2$params[2]}$, and $\sigma = \Sexpr{fig2$params[3]}$.
+Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$.
 See Introduction for model details.

### 2021-10-28 paper/SupMat.Rnw
finished revising notation in SupMat

@@
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))] - \E[\ln(1-\delta+\delta\exp(b_i^\shp-b_j^\shp))],\label{eq:decomp:contrib_EC_1}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
 \end{equation}
@@
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
-\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]
-, \label{eq:decomp:contrib_EC}
-\end{equation}
-where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
-$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$ and $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent.
-Continuing, $\olep_i^{[EC]}=\olep_i^{(EC)}-\olep_i^{[E \nat C]}$ is available by subtracting expression (\ref{eq:decomp:contrib_EC})
-from expression (\ref{eq:decomp:storage_effects}), i.e., 
-\begin{equation}
 \olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
-\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_2}
+\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
 \end{equation}
@@
 $(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
+Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
+from expression (\ref{eq:decomp:storage_effects}), i.e., 
+\begin{equation}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
+\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
+\end{equation}
 This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.
@@
 \begin{align}
-\olep_j^{[E \nat C]} &= \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] - \E[r_j(E_j^\shp , C_{j \bs i}^\shp)]\\
-&= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
-&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],
-\label{eq:decomp:j:EChard}
-\end{align}
-where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
-Continuing,
-\begin{align}
-\olep_j^{[EC]} &= \olep_j^{(EC)} - \olep_j^{[E \nat C]} \\
-&= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)]
+\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
 &= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
@@
 because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
-$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other.
+$E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
+\begin{align}
+\olep_j^{[E \nat C]} &= \olep_j^{(EC)} - \olep_j^{[EC]} \\
+&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
+\end{align}
+where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
+
 This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.
@@
 
-We now examine $\olep_i^{[EC ]}$, which, by (\ref{eq:decomp:contrib_ATA_2}), equals 
+We now examine $\olep_i^{[EC]}$, which, by (\ref{eq:decomp:contrib_ATA_1}), equals 
 \begin{equation}
@@
 \ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\mu_i-\mu_j+\sigma^2/2))]/\sqrt{M}$. 
-\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_EC}), as
-$\widehat{\olep_i^{[EC]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$.
+\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
+$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$.
 The standard error of this expression is
-$\se \left( \widehat{\olep_i^{[EC]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
+$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
 and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
-\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
-$\widehat{\olep_i^{[E \nat C]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
+\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
+$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
 \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
-error of this expression is $\se \left( \widehat{\olep_i^{[E \nat C]}} \right)=\sqrt{s_3^2+s_1^2}$, where
+error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
 $s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
@@
 It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
-\widehat{\olep_i^{[EC]}}+\widehat{\olep_i^{[E \nat C]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
+\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
 at the level of the estimators, as well
@@
 \ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
-\item Estimate $\olep_j^{[EC]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[EC]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[EC]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
-\item Use $\olep_j^{[E \nat C]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
+\item Estimate $\olep_j^{[E \nat C]}$, following (\ref{eq:decomp:j:EChard}), as $\widehat{\olep_j^{[E \nat C]}}=\mean_k [-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))],$ which has $\se \left( \widehat{\olep_j^{[E \nat C]}} \right) = \sd_k[-\ln(1-\delta+\delta \exp(\sigma \sqrt{2} u^{(k)}))]/\sqrt{M}$.
+\item Use $\olep_j^{[EC]}=0$, following (\ref{eq:decomp:j:EnatC}). This is exact.
 \item Finally, use $\olr_{j \bs i} = 0$, again exact. 
@@
 q_{ij}\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
-\item Estimate $\Delta_i^{[EC]}$ as $\widehat{\Delta_i^{[EC]}}=\widehat{\olep_i^{[EC]}}-q_{ij}\widehat{\olep_j^{[EC]}}=
+\item Estimate $\Delta_i^{[E \nat C]}$ as $\widehat{\Delta_i^{[E \nat C]}}=\widehat{\olep_i^{[E \nat C]}}-q_{ij}\widehat{\olep_j^{[E \nat C]}}=
 \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-
@@
 and $s_2=\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]/\sqrt{M}$.
-\item For $\Delta_i^{[E \nat C]}$, we recall that $\olep_j^{[E \nat C]}=0$, so $\Delta_i^{[E \nat C]}=\olep_i^{[E \nat C]}$. So we use
-$\widehat{\olep_i^{[E \nat C]}}$ for $\widehat{\Delta_i^{[E \nat C]}}$.
+\item For $\Delta_i^{[EC]}$, we recall that $\olep_j^{[EC]}=0$, so $\Delta_i^{[EC]}=\olep_i^{[EC]}$. So we use
+$\widehat{\olep_i^{[EC]}}$ for $\widehat{\Delta_i^{[EC]}}$.
 \item Finally, we can also estimate $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ simply as $\widehat{\olr_{i \bs i}}$
@@
 
-In fig \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[E \nat C]}$, and IGR $-\Delta_i^{[E \nat C]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
+In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
 From section \ref{sect:math_simplifications}, 
@@
 \begin{align}
-\olep_j^{[E \nat C]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
+\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
 &= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
@@
 \end{align}
-Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[E \nat C]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}$ are zero so
+Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
 \begin{equation}
@@
 \begin{equation}
-\Delta_i^{[E\nat C]}=\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]}=\olep_i^{[E \nat C]}
+\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
 \end{equation}
-Similarly for IGR $-\Delta_i^{[E \nat C]}$, 
+Similarly for IGR $-\Delta_i^{[EC]}$, 
 \begin{equation}
-(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-\olep_i^{[E \nat C]}-q_{ij} \olep_j^{[E \nat C]} = \olr_{i \bs i}-\olep_i^{[E \nat C]}.
+(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
 \end{equation}
-Thus $q_{ij}$ does not matter for anything computed for Figure \ref{MT-fig:fig4}.
+Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.
 

### 2021-10-29 paper/Paper.Rnw
trivial

--- 2021-10-06 paper/Paper.Rnw
@@
 when these influences are strong than when they are weak, or vice versa. 
-Recent work showed that such `asymmetric tail associations' (ATAs) are extremely common
+Recent work showed that such `asymmetric tail associations' (ATA) are extremely common
 between ecological variables, and are important for a wide range of phenomena including 
@@
 the other of which represents
-the influence of environment-competition correlation \emph{per se}, without 
-the influence of ATAs. Simulations of a simple model
+the influence of environment-competition correlation \emph{per se}, 
+without the influence of ATAs. Simulations of a simple model
 %and an empirical example
@@
 [(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
-&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
+&= \Delta_i^{[E \nat C]} + \Delta_i^{[EC]}.\label{eq:storage_effects_further}
 \end{align}
-The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
+The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to 
 ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
-The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
-correlation \emph{per se}, remaining after having removed the effects of ATAs.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
@@
 The new symbol should correspondingly be pronounced ``partial sharp''.
+The term $\Delta_i^{[EC]}$ is the component of storage effects due to $EC$ correlation \emph{per se}, remaining after 
+having removed the effects of ATAs.
 The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
@@
 which is the extension of MCT which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
-the magnitude of the new ATA term in this expression, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
+the magnitude of the new ATA term in this expression, $\Delta_i^{[E \nat C]}$, relative to the magnitudes of the other terms.
 
@@
 as well as the
-IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$. 
+IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$. 
 For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
@@
 We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of values
-$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
+$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$ is positive.
 This is the subspace for which the two species in the lottery model coexist in the absence of 
@@
 i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
-On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
+On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[E \nat C]}$
 against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
@@
 magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
-Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
+Magnitudes of $\Delta_i^{[E \nat C]}$ tended to be larger, relative to the magnitudes of the 
 other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 

### 2021-10-29 paper/Paper.Rnw
trivial

--- 2021-10-28 paper/Paper.Rnw
@@
 Our result adds coexistence to the list of ecological theory/phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
-Seemingly consistent with fluctuation dependent mechanisms, as \sigma increased, so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
+Seemingly consistent with fluctuation dependent mechanisms, as $\sigma$ increased, so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
 %Higher sigma means a more fluctuating and variable environment.

### 2021-12-14 paper/Paper.Rnw
contributed to methods for G4 and added figure 5

@@
 %***DAN: Methods paras here for G4, if we accomplish G4
+To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species in chemostat batch experiments by Descamps-Julien & Ganzalez 2005.
+In their experiments, measurements of growth of Cyclotella pseudostelligera and Fragilaria crotonensis over a range of temperatures was used to parametrize a competitive resource model as follows:
+put model here
+Descamps-Julien & Gonzalez used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d) and Ellner 2016 and 2018 broke down their coexistence into the contrirbuting mechanisms.
+Here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of C.
+pseudostelligera and F.
+crotonensis to account for contributions of ATAs and ATA effects as described above.
+
+Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model with $E=$ and $C=$ against independent parameters of temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
+For each plot, we set two of the parameters constant at the original parameters which coexistence was found in Descamps-Julien & Gonzalez, and let the remaining temperature parameter vary.
+%say which scenario this is in Ellner 2018
+We let $\theta_0$ vary between $[x,y]$, $a$ vary between $[x,y]$, and $P$ vary between $[x,y]$. 
+Relative importance of contributions was judged from the 
+plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
+the contributions of other mechanisms. 
+The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
+
+To further explore the parameter space, we generated 2 dimensional surface plots which reveal the subsets of parameter space where coexistence is possible with ranging degrees of ATA contributions and where ATA rescue occurs.
+We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/IGR$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
+Positive values of $\Delta_i^{[EC]}/IGR$ indicate coexistence and negative values indicate failure of coexistence.
+Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and higher color intensity indicate greater contribution of ATAs to the IGR and positive cases where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
 
@@
 
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
+\caption{The contribution of ATAs to diatom coexistence is meaningfully large compared to 
+other components of the MCT decomposition. All panels show the IGR and the decomposition of
+the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
+two of the three parameters defining temperature oscillation, amplitude, period, and mean, 
+is fixed at the original coexistence values found by Descamps-Julien & Gonzalez ($a=6, P=60, Tbar=18$). The IGR and components of the decomposition
+are plotted against one varying parameter.
+The original value of the varying parameter and coexistence quantities at that value is indicated by the vertical dashed magenta lines.
+The contribution of ATAs (red line)
+has magnitude comparable to other components and, in these parameters, is always positive, indicating that ATAs can contribute meaningfully
+to diatom coexistence. In some instances, the IGR line (orange) falls 
+between the ATA line (red)
+and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
+coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.  \label{fig:fig5}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/fig6.pdf}
+\caption{} \label{fig:fig6}
+\end{figure}
+
 \end{document}

### 2021-12-14 code/diatom/Figure5.R
figure 5 yea

@@ -1,0 +1,59 @@
+# Figure 5
+dat_loc <- "../results_numeric/"
+fig_loc <- "../results_figs/"
+if(dir.exists(fig_loc)==FALSE){
+  dir.create(fig_loc)
+}
+fig5 <- paste(fig_loc,"fig5.pdf",sep="")
+
+vAmp <- readRDS(paste(dat_loc,'varyingAmplitude.RDS',sep='')) #from explorespace.R
+vPer <- readRDS(paste(dat_loc,'varyingPerioda-6.RDS',sep=''))
+vTbr <- readRDS(paste(dat_loc,'varyTbar.RDS',sep=''))
+
+
+col <- c(rep("black",4), "red", "blue", "orange")
+line <- c(1:4, 1,1,1)
+
+varylist <- list(vAmp, vPer, vTbr)
+range <- sapply(varylist, function(X){range(X[,1:7])})
+ymin <- min(range)
+ymax <- max(range)
+orig <- c(a=6, P=60, Tbar=18)
+
+#default is c(5, 4, 4, 2) + 0.1
+pdf(fig5, height=5, width=15)
+par(mfrow=c(1,3), oma=c(0,3,0,0), mar=c(5,1,2,1) )
+for (i in 1:3){
+	vdat <- varylist[[i]]
+	
+	#empty box
+	plot(0, yaxt='n', xlim=range(vdat[,8]), ylim=c(ymin, ymax), col='white', xlab=colnames(varylist[[i]])[8], ylab='', cex.lab=1.3, cex.axis=1.3)
+	
+	#axis
+	if (i==1){
+		axis(2)
+	}
+	
+	#loop for lines
+	for (j in 1:7){
+		lines(vdat[,8], vdat[,j], col=col[j], lty=line[j])	
+	}
+	abline(h=0, col="lightgrey") #zero
+	abline(v=orig[i], col='magenta', lty=3)
+	
+	#label
+	mtext(LETTERS[i], 3, -1.5, adj=0.985)
+	
+	#legend
+	if (i==1){
+		legend("topleft", legend=c(expression(Delta[i]^0), expression(Delta[i]^E), 				expression(Delta[i]^C), expression(Delta[i]^"(E#C)"), expression(Delta[i]^"[E||C]"), expression(Delta[i]^"[EC]") ,expression(IGR)), col = col,lty = line, bty="n", cex=1.2, inset=c(0,0))
+	}
+}
[…]

### 2021-12-14 code/diatom/explorespace.R
file for making results to make fig 5

@@ -1,0 +1,75 @@
+source("getEpsilons_fxn.R")
+dat_loc <- "../results_numeric/"
+#loop through amplitude? 
+ep1<- get_epsilons(3, 60, "Voftemp", 500, 1)
+ep2<- get_epsilons(3, 60, "Voftemp", 500, 2)
+Delta1 <- ep1-ep2
+sum(Delta1)-(Delta1$epsEC+Delta1$epsECpar)
+
+a <- seq(0.5,7,0.5)
+ep1 <- get_epsilons(a[1], 60, "temp", 500, 1)
+ep2 <- get_epsilons(a[1], 60, "temp", 500, 2)
+Delta1 <- ep1-ep2
+
+for (i in 2:length(a)){
+	ep1 <- get_epsilons(a[i], 60, "temp", 500, 1)
+	ep2 <- get_epsilons(a[i], 60, "temp", 500, 2)
+	Delta1[i,] <- ep1-ep2
+	print(i)
+}
+Delta1 <- Delta1[,-c(4,6)]
+Delta1$IGR <- rowSums(Delta1)
+Delta1$amplitude <- a
+
+saveRDS(Delta1, paste(dat_loc,'varyingAmplitude.RDS',sep=''))
+round(Delta1[,c(5,7,8)],5)
+#######################################################################
+
+ep1 <- get_epsilons(4.415, 60, "temp", 500, 1)
+ep2 <- get_epsilons(4.415, 60, "temp", 500, 2)
+D <- ep1-ep2
+D <- D[-c(4,6)]
+D[7] <- sum(D)
+D[7]-D[5];D[7]
+#4.29-4.41 have ATA rescue
+#######################################################################
+
+
+#vary P at a=6
+P <- seq(20,120,10)
+ep1 <- get_epsilons(6, P[1], "temp", 500, 1)
+ep2 <- get_epsilons(6, P[1], "temp", 500, 2)
+DeltaP <- ep1-ep2
+
+for (i in 2:length(P)){
+	ep1 <- get_epsilons(6, P[i], "temp", 500, 1)
+	ep2 <- get_epsilons(6, P[i], "temp", 500, 2)
+	DeltaP[i,] <- ep1-ep2
+	print(i)
+}
+DeltaP <- DeltaP[,-c(4,6)]
+DeltaP$IGR <- rowSums(DeltaP)
+
+round(DeltaP[,c(5,7,8)],5)
[…]

### 2021-12-20 paper/Paper.Rnw
adding results and caption fig 6

@@
 
+%G4
+Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
+show that the ATA contributions have magnitude comparable to the 
+magnitudes of other terms of the extended MCT decomposition and 
+parameters subspaces exist where ATA rescue occurs. 
+Within the investigated parameters of Fig.\ref{fig:fig5}, ATA contribution, $\Delta_i^{[EC]}$, 
+is always positive, even in cases where IGR is negative, and is almost always second to 
+the contribution of varying environment, $\Delta_i^{E}$, 
+including for the original coexistence parameters specified in Gonzalez Decamps.  
+Fig.\ref{fig:fig5} plots the various contributions of coexistence (see eq.
+\ref{eq:full_decomp}) for different varying parameters with two of the three parameters set constant in each plot.
+Increasing amplitude and mean temperature showed increasing trends in the IGR and decomposed quantities, while increasing periods showed less change.
+
 \section{Discussion}\label{section:discussion}
@@
 see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
-$\text{IGR}-\Delta_i^{[E || C]}$ is negative correspond to ATA rescue (shaded yellow), whereas
-values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[E || C]}$ is positive 
+$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
+values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
 correspond to ATA exclusion (shaded violet). Top panels also include domains
@@
 \includegraphics[width=\textwidth]{../results_figs/fig6.pdf}
-\caption{} \label{fig:fig6}
+\caption{The degree of ATA contribution to the IGR change with parameters of temperature oscillation.
+The ratio $\Delta_i^{[EC]}/IGR$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions.
+In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/IGR$ only result from positive $\Delta_i^{[EC]}$ and positive $IGR$ and negative $\Delta_i^{[EC]}/IGR$ only result from positive $\Delta_i^{[EC]}$ and negative $IGR$.
+Thus, cyan areas indicate failure to coexist and magenta areas indicate coexistence with ATA contribution increasing with color intensity.
+The most intense areas of magenta indicate parameter subspaces where ATA rescue occurs.
+For simplicity and readability magnitudes of $\Delta_i^{[EC]}/IGR$ >1 are capped at $\abs{1}$.
+} \label{fig:fig6}
 \end{figure}

### 2021-12-30 code/diatom/workflow.R
all functions in one script

@@ -1,0 +1,131 @@
+source("diatom/DataAndVKQfuns.R")
+source("diatom/partialSharp_fxns.R")
+require(deSolve)
+
+
+getep3 <- function(a, P, Tbar, time, reps, sp, invader=1){
+	
+	parms <- c(Tbar=Tbar, a=a, P=P, D=0.09, S=35)
+	time <- round(time/P,0)*P
+	
+	y0 <- c(R=.1,x1=0,x2=20)
+	times <- seq(0,time,by=.1)
+	out1i <- ode(y0,times,func=forceChemo,parms=parms) 
+	e <- times > max(times-signif(floor(max(times)/2.5),2)) 
+	out1i <- out1i[e,] 
+	temp <- out1i[,5]
+	R <- out1i[,2]
+	
+	if (sp==2){
+		Vfun <- V2modfun; Kfun <- K2flatfun
+	} else {Vfun <- V1quad; Kfun <- K1flatfun}
+
+	E <- temp
+	r <- function(E, C, parms){Vfun(E)/C - parms["D"]}
+	
+	C <- (Kfun(temp)+R)/R
+	
+	#special E's and C's
+	E0 <- mean(E)
+	C0 <- mean(C)
+	
+	EC <- expand.grid(E, C)
+	Esharp <- EC[,1]; Csharp <- EC[,2]
+	rm(EC)
+	
+	ECpsharp <- makePsharp(E, C, reps)
+	Epsharp <- ECpsharp[,1,]; Cpsharp <- ECpsharp[,2,]
+	
+	#special r's
+	rsharp <- mean(r(Esharp, Csharp, parms))
+	
+	rpsharpsims <- apply(ECpsharp, MARGIN=3, function(EC)			{ mean(r(EC[,1], EC[,2], parms)) } );
+	rpsharp <- median(rpsharpsims)
+	
+	rbar <- mean(r(E, C, parms))
+	
+	#epsilons
+	eps0 <- r(E0, C0, parms)
+	epsE <- mean(r(E, C0, parms)) - eps0
+	epsC <- mean(r(E0, C, parms)) - eps0
+	#epsEC <- rbar - epsE - epsC - eps0
+	epsEsharpC <- rsharp - epsE - epsC - eps0
+	#epsECpar <- rbar - rsharp
+	epsECbrk <- rbar - rpsharp
+	epsEpsharpC <- rpsharp - rsharp
+	
[…]

### 2021-12-30 paper/Paper.Rnw

@@
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 %***DAN: Methods paras here for G4, if we accomplish G4
-To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species in chemostat batch experiments by Descamps-Julien & Ganzalez 2005.
-In their experiments, measurements of growth of Cyclotella pseudostelligera and Fragilaria crotonensis over a range of temperatures was used to parametrize a competitive resource model as follows:
+To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species in chemostat batch experiments by \cite{Descamps_2005}.
+In their experiments, measurements of growth of \emph{Cyclotella pseudostelligera} and \emph{Fragilaria crotonensis} over a range of temperatures was used to parametrize a competitive resource model as follows:
 put model here
-Descamps-Julien & Gonzalez used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d) and Ellner 2016 and 2018 broke down their coexistence into the contrirbuting mechanisms.
-Here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of C.
-pseudostelligera and F.
-crotonensis to account for contributions of ATAs and ATA effects as described above.
+\cite{Descamps_2005} used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d) and \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms.
+Here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of \emph{C.
+pseudostelligera} and \emph{F.
+crotonensis} to account for contributions of ATAs and ATA effects as described above, using decomposition defined by $E=...$ and $C=...$.
 
 Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model with $E=$ and $C=$ against independent parameters of temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
-For each plot, we set two of the parameters constant at the original parameters which coexistence was found in Descamps-Julien & Gonzalez, and let the remaining temperature parameter vary.
+For each plot, we set two of the parameters constant at the original parameters which coexistence was found in \cite{Descamps_2005}, and let the remaining temperature parameter vary.
 %say which scenario this is in Ellner 2018
@@
 To further explore the parameter space, we generated 2 dimensional surface plots which reveal the subsets of parameter space where coexistence is possible with ranging degrees of ATA contributions and where ATA rescue occurs.
-We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/IGR$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
-Positive values of $\Delta_i^{[EC]}/IGR$ indicate coexistence and negative values indicate failure of coexistence.
+We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
+Positive values of $\Delta_i^{[EC]}/\text{IGR}$ indicate coexistence and negative values indicate failure of coexistence.
 Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and higher color intensity indicate greater contribution of ATAs to the IGR and positive cases where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
@@
 two of the three parameters defining temperature oscillation, amplitude, period, and mean, 
-is fixed at the original coexistence values found by Descamps-Julien & Gonzalez ($a=6, P=60, Tbar=18$). The IGR and components of the decomposition
+is fixed at the original coexistence values found by \cite{Descamps_2005} ($a=6, P=60, \theta_0=18$). The IGR and components of the decomposition
 are plotted against one varying parameter.
@@
 and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.  \label{fig:fig5}
+coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.}  \label{fig:fig5}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig6.pdf}
+\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
 \caption{The degree of ATA contribution to the IGR change with parameters of temperature oscillation.
-The ratio $\Delta_i^{[EC]}/IGR$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions.
-In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/IGR$ only result from positive $\Delta_i^{[EC]}$ and positive $IGR$ and negative $\Delta_i^{[EC]}/IGR$ only result from positive $\Delta_i^{[EC]}$ and negative $IGR$.
+The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions.
+In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and negative IGR.
 Thus, cyan areas indicate failure to coexist and magenta areas indicate coexistence with ATA contribution increasing with color intensity.
 The most intense areas of magenta indicate parameter subspaces where ATA rescue occurs.
-For simplicity and readability magnitudes of $\Delta_i^{[EC]}/IGR$ >1 are capped at $\abs{1}$.
+For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ >1 were capped at $|1|$.
 } \label{fig:fig6}

### 2021-12-30 paper/Paper.Rnw
some writing

@@
 put model here
-\cite{Descamps_2005} used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d) and \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms.
+\cite{Descamps_2005} used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d and \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms.
 Here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of \emph{C.
 pseudostelligera} and \emph{F.
-crotonensis} to account for contributions of ATAs and ATA effects as described above, using decomposition defined by $E=...$ and $C=...$.
+crotonensis} to account for contributions of ATAs and ATA effects as described above.
+Our decomposition is defined by $E(t)=\theta(t)$, $C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$.
+We first considered the scenario where \emph{F.
+crotonensis} is the invader and emph{C.
+pseudostelligera} is the resident to compute the coexistence decomposition for \emph{F.
+crotonensis}.
 
-Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model with $E=$ and $C=$ against independent parameters of temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
-For each plot, we set two of the parameters constant at the original parameters which coexistence was found in \cite{Descamps_2005}, and let the remaining temperature parameter vary.
-%say which scenario this is in Ellner 2018
-We let $\theta_0$ vary between $[x,y]$, $a$ vary between $[x,y]$, and $P$ vary between $[x,y]$. 
+Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model against independent parameters defining temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
+For each plot, we set two of the parameters constant at the original parameters which coexistence was experimentally found in \cite{Descamps_2005}, and let the remaining temperature parameter vary.
+We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
 Relative importance of contributions was judged from the 
@@
 
-To further explore the parameter space, we generated 2 dimensional surface plots which reveal the subsets of parameter space where coexistence is possible with ranging degrees of ATA contributions and where ATA rescue occurs.
+To further explore the parameter space, we generated two dimensional surface plots which reveal the subsets of parameter space where coexistence is possible with ranging degrees of ATA contributions and where ATA rescue occurs.
 We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
@@
 \underline{paragraph summarizing results of G2 and how it compares and builds on prior theory}
-Furthermore, we have shown not only that is it possible to decompose storage effect into the effect of correlation \emph{per se} and the effect of ATAs, but this new term that holds the contribution of not only the correlation but the structure of the correlation is a non-negligible term, in the lottery model.
-Our result adds coexistence to the list of ecological theory/phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
+Furthermore, we have shown not only that is it possible to decompose storage effect into the effect of correlation \emph{per se} and the effect of ATAs, but this new term that holds the contribution of not only the correlation but the structure of the correlation is a non-negligible term, in a simple lottery model.
+Our result amends coexistence to the list of ecological theory and phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
 Seemingly consistent with fluctuation dependent mechanisms, as $\sigma$ increased, so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
@@
 ATA specialist? Does this go in this paragraph
+
+The parameter subspaces of ATA rescue and ATA exclusion suggest that some species may be ATA specialists  - only persisting in habitat patches where ATAs contribute to its coexistence. 
 %I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
@@
 \underline{paragraph summarizing results of G4 and how it compares and builds on prior theory}
+The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
+
+Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
+Previous studies report storage effect as a contribution to coexistence of EC covariance; yet compute it as the contribution of the EC relationship (covariance and beyond; symmetric and asymmetric).
 

### 2022-01-06 paper/Paper.Rnw
edditing methods and results, added modified fig6

@@
 %***DAN: Methods paras here for G4, if we accomplish G4
-To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species in chemostat batch experiments by \cite{Descamps_2005}.
-In their experiments, measurements of growth of \emph{Cyclotella pseudostelligera} and \emph{Fragilaria crotonensis} over a range of temperatures was used to parametrize a competitive resource model as follows:
+To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species established by chemostat batch experiments \citep{Descamps_2005}.
+\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive resource model as follows:
 put model here
-\cite{Descamps_2005} used simulations to show that coexistence occurs with oscillating temperature fluctuations with parameters $\theta_0 = 18, a = 6, P = 60$d and \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms.
-Here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of \emph{C.
+Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C.
+pseudostelligera} and \emph{F.
+crotonensis} occurs with oscillating temperature fluctuations, $\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, defined by parameters $\theta_0 = 18, a = 6, P = 60$d.
+\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, and here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of \emph{C.
 pseudostelligera} and \emph{F.
@@
 crotonensis}.
-
+%put reverse in sup mat?
 Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model against independent parameters defining temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
@@
 We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
-Positive values of $\Delta_i^{[EC]}/\text{IGR}$ indicate coexistence and negative values indicate failure of coexistence.
 Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and higher color intensity indicate greater contribution of ATAs to the IGR and positive cases where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
@@
 Increasing amplitude and mean temperature showed increasing trends in the IGR and decomposed quantities, while increasing periods showed less change.
+
+Fig.\ref{fig:fig6} displays the subsets of environmental parameters for which coexistence occurs or fails and the degree of ATA contribution to the IGR.
+Specific to this model and parameter ranges, because $\Delta_i^{[EC]}$ is always positive, all positive values of $\Delta_i^{[EC]}/\text{IGR}$ indicate coexistence and negative values of $\Delta_i^{[EC]}/\text{IGR}$ indicate failure of coexistence.
+Futhermore, ATA exlusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>IGR$.
+To preserve simplicity and readibility of the plots, |$\Delta_i^{[EC]}/\text{IGR}$>1| were capped at |1|.
 

### 2022-01-06 paper/Paper.Rnw
trying to make discussion

@@
 Our decomposition is defined by $E(t)=\theta(t)$, $C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$.
+Thus in our exploration, environmental fluctuation is temperature fluctuation.
 We first considered the scenario where \emph{F.
@@
 %put reverse in sup mat?
+%do we need to explain how we separated the asymmetries using the sims??
 Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model against independent parameters defining temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
@@
 Futhermore, ATA exlusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>IGR$.
-To preserve simplicity and readibility of the plots, |$\Delta_i^{[EC]}/\text{IGR}$>1| were capped at |1|.
+To preserve simplicity and readibility of the plots, $|\Delta_i^{[EC]}/\text{IGR}>1|$ were capped at $|1|$.
+Higher magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ tended to occur as the IGR became smaller and smaller, thus creating the sharp contrast in the plots where highest intensity positive spaces quickly change into highest intensity negative spaces with decreasing parameters.
+Like the result of Fig.\ref{fig:fig5}, we see in Fig.\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence.
+Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{IGR}$ changes from negative to positive.
+The effect of varying periodicity of environmental fluctuations is...
 
@@
 \underline{Paragraph summarizing results of G1 and how it compares and builds on prior theory.} 
-We provided a farther tertiary decomposition to MCT, 
+We provided a further tertiary decomposition to MCT, 
 extending the framework put forward by \citep{Ellner_2019}. 
@@
 In addressing this inconsistency, we have separated the mechanisms of storage effect into the contribution of ATA [to storage effect/coexistence] and the contribution of correlation \emph{per se} [to storage effect/coexistence].
+%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
+%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
+%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
+
+%Need a para about multi-species models - how would things change? One difference is there are more different
+%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
+%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
+We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
+Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
 
@@
 Our result amends coexistence to the list of ecological theory and phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
-Seemingly consistent with fluctuation dependent mechanisms, as $\sigma$ increased, so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
-%Higher sigma means a more fluctuating and variable environment.
-Then, if environmental fluctuations become more variable, the effect of ATAs becomes more important and non-negligible (rescue or exclusion).
+
+Seemingly consistent with fluctuation dependent mechanisms, 
+as $\sigma$, the variation in the environment, increased, 
+so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
+This may suggest that, if environmental fluctuations become more variable, 
+the effect of ATAs becomes more important and non-negligible (rescue or exclusion). 
+{Might be interesting to study how tail associations of environmental variables are changing and how this might 
+alter competition, coexistence, and the maintenance of biodiversity generally.} 
+%This would be a whole potential 
+%research program which presumably would start with a more detailed study of how tail associations of environmental
+%variables may be changing under global change.
+
 %Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
@@
 \underline{caveats/assumptions that could change result}
+We used the most extreme version of asymmetric noise, i.e., perfect correlation changing to very low correlation at exactly the half point of the distribution, and a simple lottery model.
 
@@
 to other ecological theories.
-
-%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
-%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
-%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
-
-%Need a para about multi-species models - how would things change? One difference is there are more different
-%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
-%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
-
-%Might be interesting to study how tail associations of environmental variables are changing and how this might 
-%alter competition, coexistence, and the maintenance of biodiversity generally. This would be a whole potential 
-%research program which presumably would start with a more detailed study of how tail associations of environmental
-%variables may be changing under global change.
 

### 2022-01-07 paper/Paper.Rnw
small edit to caption
making some discussion attempts

@@
 
+We produced and applied a mathematical extension to MCT that, for the first time, 
+considers ATAs for the coexistence of species and maintenance of biodiversity. 
+The results of our application to a simple lottery model and 
+to a competitive diatom model show that the contribution of ATAs to coexistence 
+is non-negligible and meaningfully contributes to the coexistence of species. 
+For some parameters, the contribution of ATAs was the driving mechanism of 
+coexistence – the IGR would be negative without it – deeming the scenario a case of 
+ATA rescue. For other parameter values ATA exclusion occurs where ATAs are the 
+driving mechanism of the failure of coexistence and the lack of ATAs would result in 
+a positive IGR. 
+Subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists}.
+Our results reinforce that common methods of correlation fail to reflect and include 
+important information on the structure of the relationship between variables and amends
+ coexistence to the list of ecological theory and phenomena that ATAs have been found to 
+ be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
+
+ We provided a further tertiary decomposition to MCT, 
+extending the framework put forward by \citep{Ellner_2019}. 
+We decompose the storage effect into two new terms, 
+$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, 
+the contribution of correlation \emph{per se}. 
+Previously, storage effects were described as the contribution
+of covariance between the environment and competition to 
+coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
+However, the description of storage effects as the covariance between drivers
+is a misnomer in terminology, and we propose that storage effects, instead, 
+should be described as the contribution of the relationship between drivers. 
+All aspects and of the relationship between environmental response and competitive pressure 
+are not fully captured by only the covariance between them because a measure of covariance 
+does not capture the structure of associativity throughout the distribution of the relationship 
+and any asymmetries there may be. 
+Thus, it is misleading to describe the contribution of the whole EC relationship 
+as the contribution of EC covariance.
+In addressing this inconsistency, we have separated the mechanisms of storage effect 
+into the contribution of ATA [to storage effect/coexistence] and the contribution of 
+correlation \emph{per se} [to storage effect/coexistence].
+
+Consistent with the mechanisms of biodiversity maintenance and specifically 
+fluctuation-dependent mechanisms (cite chesson), higher values of environmental fluctuation 
+resulted in higher IGR’s. Both of our new terms, $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$ 
+also increased in magnitude with increased environmental fluctuation. Increased environmental 
+fluctuation is reflected in increased $\sigma$ in the lottery model and increased $a$ in the 
+competitive diatom model.   mean temp and period?
+
+Distinct increase in contribution of ATAs to coexistence as environmental fluctuations increase may suggest that ATAs may become more important to consider for coexistence and possibly other ecological phenomena as the environment becomes more variable, a consequence of climate change and extreme climatic events.
+
 Outline
@@
 \underline{Paragraph summarizing results of G1 and how it compares and builds on prior theory.}
-We provided a further tertiary decomposition to MCT, 
-extending the framework put forward by \citep{Ellner_2019}. 
-We decompose the storage effect into two new terms, 
-$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, the contribution of correlation \emph{per se}. 
-Previously, storage effects were described as the contribution
-of covariance between the environment and competition to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-However, the description of storage effects as the covariance between drivers
-is a misnomer in terminology, and we propose that storage effects, instead, 
-should be described as the contribution of the relationship between drivers. 
-All aspects and of the relationship between environmental response and competitive pressure are not fully captured by only the covariance between them because a measure of covariance does not capture the structure of associativity throughout the distribution of the relationship and any asymmetries there may be.
-Thus, it is misleading to describe the contribution of the whole EC relationship 
-as the contribution of EC covariance.
-In addressing this inconsistency, we have separated the mechanisms of storage effect into the contribution of ATA [to storage effect/coexistence] and the contribution of correlation \emph{per se} [to storage effect/coexistence].
+
 %Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
@@
 Furthermore, we have shown not only that is it possible to decompose storage effect into the effect of correlation \emph{per se} and the effect of ATAs, but this new term that holds the contribution of not only the correlation but the structure of the correlation is a non-negligible term, in a simple lottery model.
-Our result amends coexistence to the list of ecological theory and phenomena that ATAs have been found to be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
 
@@
 \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-\caption{The degree of ATA contribution to the IGR change with parameters of temperature oscillation.
+\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature oscillation.
 The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions.

### 2022-01-09 paper/Paper.Rnw
shortened the abstract to the 150 word limit of Ecol Lett

@@
 
+%Currently 148 of an allowed 150 words for Ecology Letters
 \begin{abstract}
-\noindent Understanding the coexistence of competing species is a topic of very broad 
-interest in ecology. The modern coexistence
-theory (MCT) of Chesson identifies several mechanisms of species coexistence,
+\noindent Species coexistence attracts wide 
+interest in ecology. Modern coexistence
+theory (MCT) identifies mechanisms of coexistence,
 one of which, storage effects, hinges on relationships between fluctuations
-in environmental and competitive pressures on species populations. However, such relationships
-have typically been measured using covariance methods, which do not account for 
-the possibility that environmental and competitive influences may be more correlated
-when these influences are strong than when they are weak, or vice versa. 
-Recent work showed that such `asymmetric tail associations' (ATAs) are extremely common
-between ecological variables, and are important for a wide range of phenomena including 
-extinction risk and the stability of ecosystem functioning. 
-Here we extend the MCT to take into account the influence of ATAs. Storage
-effects can be decomposed into two components, one of which quantifies the
-influence of ATAs on species coexistence, and is a novel contribution of this study; and 
-the other of which represents
-the influence of environment-competition correlation \emph{per se}, 
-without the influence of ATAs. Simulations of a simple model
-%and an empirical example
-illustrate that ATA influences on coexistence can be comparable in magnitude to other
-mechanisms of coexistence, and that the presence or absence of ATAs can make the 
-difference between species coexisting and competitive exclusion. 
+in environmental and competitive pressures. However, such relationships
+are typically measured using covariance, which does not account for 
+the possibility that environmental and competitive influences may be more related to each other
+when they are strong than when weak, or vice versa. 
+Recent work showed that such `asymmetric tail associations' (ATAs) are common
+between ecological variables, and are important for 
+extinction risk, ecosystem stability, and other phenomena. 
+We extend the MCT to account for ATAs. Storage
+effects are decomposed to show the influence of ATAs. Simulations of a simple model
+and an empirical example using diatoms
+illustrate that ATA influences can be comparable in magnitude to other
+mechanisms of coexistence, and that ATAs can make the 
+difference between species coexistence and competitive exclusion. 
 \end{abstract}
@@
 %***Add to these
-\keywords{asymmetric tail association, biodiversity, coexistence, competition, modern coexistence theory, storage effects}
+\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effects}
 

### 2022-01-09 paper/Paper.Rnw
edits to Intro

@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
+%***DAN: Jasmin, Descamps_2005 does not appear in refs.bib, I think. Please add it there or subtract the callout here, whatever is most appropriate.
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
-The variables in Fig. \ref{fig:noise}A,C 
+The variables in Fig. \ref{fig:noise}a,c 
 exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
 they are more strongly related to each
-other in the left (Fig. \ref{fig:noise}A, termed ``left-tail association'') or right (Fig. \ref{fig:noise}C,
+other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
 termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
@@
 example buffering mechanism \citep{Chesson_2012}. 
-When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
+When present, these three factors combine so that EC covariance provides a relative benefit to whatever species are currently rare.
 Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
@@
 alter ATAs of environmental variables which impact coexistence, as follows.
-It is well known that climate change can or may alter the intensity, duration, extent and frequency of extreme environmental events, including
-heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017}.  
+It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
+heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017, Lyon2019, Keelings2020}.  
 If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
@@
 of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
+
+%***DAN: Dan made it to here on 2022 01 09 editing, continue tomorrow.
 
@@
 Higher magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ tended to occur as the IGR became smaller and smaller, thus creating the sharp contrast in the plots where highest intensity positive spaces quickly change into highest intensity negative spaces with decreasing parameters.
-Like the result of Fig.\ref{fig:fig5}, we see in Fig.\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence.
+Like the result of Fig.
+\ref{fig:fig5}, we see in Fig.
+\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence.
 Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{IGR}$ changes from negative to positive.
@@
 \caption{Simple example of bivariate random variables with standard normal marginals showing
-alternative patterns of association in the tails. A) is ``left-tail 
-associated,'' B) is symmetric, and C) is ``right-tail associated'' noise. All three cases have 
-the same Pearson correlation, $\rho$, up to sampling variation, yet show very different 
-patterns of association. We used asymmetrically tail associated cases which are perfectly 
-correlated in the lower (A)  or upper (C) halves of the distributions, though it is also 
+alternative patterns of association in the tails. The varables in (a) are ``left-tail 
+associated,'' those in (b) are symmetrically associated, and those in (c) are ``right-tail associated''. All three cases have 
+the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these very different 
+patterns of association. We used asymmetrically tail associated cases (a, c) which are perfectly 
+correlated in the lower or upper halves of the distributions, though it is also 
 possible to generate asymmetrically tail associated noise with imperfect correlations in 
@@
 \end{figure}
-%***DAN: Dan has edited this caption at least once.
+%***DAN: Jasmin, I appologize for the following pedantic request, I explain below why it is necessary. Can you please change panel
+%labels on this and all figure from the format "A", "B", "C" to the format "(a)", "(b)", "(c)" (i.e., lower case panel labels
+%with parentheses around them). The reason for this is that this is Ecology Letters style. If we do not follow their style then their
+%copy editors will make the change for us. I would prefer not to have copy editors messing around with our figures. Though Ecology Letters
+%copy editors are usually pretty good, I have had many bad experiences with copy editors, generally, and I have had at least a few bad
+%experiences with Ecology Letters copy editors. What happens is, they introduce errors into a paper while typesetting it, and one has to
+%check their work very carefully to avoid those mistakes going to press. It's hard to blame them since it must be the most boring job 
+%in the world, but anyway we do not want to have errors in our published paper as a result. The best way to prevent them introducing 
+%errors into our figures is to not given them any reason to make any changes at all to our figures. The same change has to be 
+%made to all main text and sup mat figures, and then all callouts to figures have to also be so changed, please. I will do these
+%as much as possible while I edit, but if you could please also check it will help, and I am relying on you to change the panel labels
+%in the figured themselves. I am sorry again for the tedium, here. Maybe some day they will improve the publication process so we do not 
+%have to worry about nonsense like this.
 
@@
 spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
-species coexistence. Panel A shows species 1, from the lottery 
-model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}B). Panel B shows species 1 from the 
-same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}A).
+species coexistence. Panel (a) shows species 1, from the lottery 
+model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
+same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}a).
 Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
@@
 total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
-was \Sexpr{fig2$coSym} time steps for panel A and \Sexpr{fig2$coATA} for panel B, over the full length of the simulations; 
+was \Sexpr{fig2$coSym} time steps for panel (a) and \Sexpr{fig2$coATA} for panel (b), over the full length of the simulations; 
 whereas the average duration of periods of
-one-species dominance was \Sexpr{fig2$domSym} time steps for panel A and \Sexpr{fig2$domATA} for B. Thus, while periods of 
+one-species dominance was \Sexpr{fig2$domSym} time steps for panel (a) and \Sexpr{fig2$domATA} for (b). Thus, while periods of 
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
@@
 Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$.
-See Introduction for model details.
+See Introduction and Fig.
+\ref{fig:noise} caption for model details.
 \label{fig:pedagog}}
 \end{figure}
-%***DAN: Dan has edited this caption
 
@@
 \end{figure}
-%***DAN: Dan has edited this caption
 

### 2022-01-10 paper/Paper.Rnw
edits leading up to editing Jasmin's new Methods content on the empirical system, also added new lines to that content to make it better tracked by git. Will edit that content next.

@@
 
-%***DAN: Dan made it to here on 2022 01 09 editing, continue tomorrow.
-
 %Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
 Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-not been studied, the importance for coexistence of other aspects of climate change, including changes 
-in means and measures of variability of environmental variables, has been studied in many systems, so we briefly 
-review some of that large, related literature.
-For instance, \cite{diPaola_2012} concluded that coexistence of drought sensitive,
+not been studied, the importance for coexistence of changes 
+in means and variability of environmental variables has been studied in many systems. We briefly 
+review some of that literature.
+\cite{diPaola_2012} concluded that coexistence of drought sensitive,
 flood resistant species with flood sensitive, drought resistant species in 
 Mediteranean coastal swamp forests of central Italy would be jeopardized
-by projected changes in mean rainfall for the region. In an interesting example
-involving climate-change effects on herbivore-mediated competition in plants,
+by projected changes in mean rainfall. 
 \cite{Descombes_2020} found that if warming trends promote faster expansion toward 
 higher altitudes of herbivore populations than of plant populations, newly arrived
-herbivores at high altitudes can inflate species richness by reducing dominant plant species 
+herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
 in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
 alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
-all species they considered, but that changes in competition would more than offset these direct benefits,
-and, as a result, some species would be extirpated. Numerous studies also take an ``event-focused'' 
+all species they considered, but that changes in competition would more than offset these direct benefits. 
+Numerous studies take an ``event-focused'' 
 perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
 form of extreme weather events, can alter competitive interactions. For instance, 
-\cite{White_2001} showed that single extreme heating events can alter competition 
+\cite{White_2001} showed that single extreme heat events can alter competition 
 between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
-nutrient availability for plants, which is mediated by soil moisture content, can alter 
+nutrient availability for plants, mediated by soil moisture content, can alter 
 the potential for invasive species to enter an ecosystem and establish, a process that involves
@@
 %Goal para
-We previously demonstrated \citep{Ghosh_2020_AER} that ATAs exist and are common in nature, and we argued above 
+We previously demonstrated \citep{Ghosh_2020_AER} that ATAs are common in nature, and we argued above 
 that it is plausible both that climate change may alter environmental ATAs, and that ATAs may influence species 
@@
 We address how ATA contributes to these components. 
-We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; spatial variation can be considered in future work.
+We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
+spatial variation can be considered in future work.
 G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
-from ATAs compared to other, known mechanisms of coexistence. This is to help illuminate whether ATA contributions to coexistence 
+from ATAs compared to other, previously known mechanisms of coexistence. This is to help 
+illuminate whether ATA contributions to coexistence 
 are likely to be negligible or meaningfully large. 
-G3) To explore, using our simple model, whether species can coexist with
+G3) To explore, using the lottery model, whether species can coexist with
 ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
@@
 Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
-$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions. Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
+$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions.
+Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
 An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
@@
 Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
+Though some readers may find it slightly confusing to make simultaneous use of the variable 
+$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard and very common, so it would 
+be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
+all operators, $\E(\cdot)$ is rendered upright.
 
@@
 Points of the trait-difference space characterize possible combinations of traits of the two competing species in our model. For fixed values of 
-$\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of 
+$\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
+trait-difference space to be the set of 
 values $\mu_1 - \mu_2$ for which the the IGR, $\olr_{i \bs i}$, is positive. This 
@@
 %***DAN: Methods paras here for G4, if we accomplish G4
-To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species coexistence which highlights the role ATAs can play, we consider the coexistence of diatom species established by chemostat batch experiments \citep{Descamps_2005}.
-\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive resource model as follows:
+To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
+coexistence which highlights the role ATAs can play, we consider the coexistence of diatom 
+species established by chemostat batch experiments \citep{Descamps_2005}.
+\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
+and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
+resource model as follows:
 put model here
-Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C.
-pseudostelligera} and \emph{F.
-crotonensis} occurs with oscillating temperature fluctuations, $\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, defined by parameters $\theta_0 = 18, a = 6, P = 60$d.
-\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, and here, we consider varying parameters of temperature fluctuations and further decompose the coexistence of \emph{C.
-pseudostelligera} and \emph{F.
-crotonensis} to account for contributions of ATAs and ATA effects as described above.
-Our decomposition is defined by $E(t)=\theta(t)$, $C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$.
-Thus in our exploration, environmental fluctuation is temperature fluctuation.
-We first considered the scenario where \emph{F.
-crotonensis} is the invader and emph{C.
-pseudostelligera} is the resident to compute the coexistence decomposition for \emph{F.
-crotonensis}.
+Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C. pseudostelligera} 
+and \emph{F. crotonensis} occurs with oscillating temperature fluctuations, 
+$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
+\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, 
+and here, we consider varying parameters of temperature fluctuations and further decompose the 
+coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
+of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
+$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
+environmental fluctuation is temperature fluctuation. We first considered the scenario where 
+\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
+the coexistence decomposition for \emph{F. crotonensis}. 
 %put reverse in sup mat?
 %do we need to explain how we separated the asymmetries using the sims??
-Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat model against independent parameters defining temperature oscillation, $\theta_0, a$, and $P$, a total of three plots.
-For each plot, we set two of the parameters constant at the original parameters which coexistence was experimentally found in \cite{Descamps_2005}, and let the remaining temperature parameter vary.
+Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
+model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
+$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
+original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
+and let the remaining temperature parameter vary. 
 We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
 Relative importance of contributions was judged from the 
-plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
+plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
+large for the lottery model, compared to 
 the contributions of other mechanisms. 
@@
 
-To further explore the parameter space, we generated two dimensional surface plots which reveal the subsets of parameter space where coexistence is possible with ranging degrees of ATA contributions and where ATA rescue occurs.
-We evaluated coexistence under two varying parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total IGR, with a color bar and contour lines.
-Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and higher color intensity indicate greater contribution of ATAs to the IGR and positive cases where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
+To further explore the parameter space, we generated two dimensional surface plots which reveal 
+the subsets of parameter space where coexistence is possible with ranging degrees of ATA 
+contributions and where ATA rescue occurs. We evaluated coexistence under two varying 
+parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total 
+IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and 
+higher color intensity indicate greater contribution of ATAs to the IGR and positive cases 
+where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
 
-All computations were done in R on an ordinary laptop. Complete computer codes for the project are stored at
+All computations were done in R on an ordinary laptop running Ubuntu Linux. 
+Complete computer codes for the project are stored at
 \url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
-
-%Ideas while driving to pick up Max:
-%1) In this paper with Jasmin you need to cite the two papers you just got from the professor at University of Massachusetts, Lowell.
-I mean the papers about changes in the special extent of heat waves: https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2020GL087097 and https://iopscience.iop.org/article/10.1088/1748-9326/ab4b41
-%The second one is the one to cite
-%2) Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
-However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
-Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
-Threshold relationships can produce tail association, as we showed elsewhere.
-Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
 
@@
 \section{Discussion}\label{section:discussion}
+
+%Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
+However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
+Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
+Threshold relationships can produce tail association, as we showed elsewhere.
+Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
 

### 2022-01-10 paper/SupMat.Rnw
edits leading up to editing Jasmin's new Methods content on the empirical system, also added new lines to that content to make it better tracked by git. Will edit that content next.

@@
 $B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
-in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it calculate $q_{ij}$,
+in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
 but we instead compute $q_{ij}$ through implicit differentiation:
@@
 
-%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor;
+%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor - easy
+%just means a change in wording at the top bit;
 %Steve reads this and thinks it makes sense to change it because something is factually or presentationally 
@@
 %especially if Steve's prediction is correct that it won't make much difference.
-%***DAN: You probably need to spend a bit more time with the cited papers to better understand this. 
+%***DAN: Dan, you probably need to spend a bit more time with the cited papers to better understand this. 
 We also use $q_{ij}=1$ as an alternative value, based on communication with Steve Ellner, for the following reasons.
@@
 where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
-
 This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.

### 2022-01-10 paper/Paper.Rnw
formatting change for figure numbers in the sup mat
edits to G4 methods content and new figure caption, not yet done though

@@
 put model here
+%***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
 Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C. pseudostelligera} 
 and \emph{F. crotonensis} occurs with oscillating temperature fluctuations, 
-$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
+$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, 
+%***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
+defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
 \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, 
@@
 %put reverse in sup mat?
+%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to just consider 1 invading 2 for 
+%the lottery model, but cannot really do that here, right?
 %do we need to explain how we separated the asymmetries using the sims??
-Like for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
+%***DAN: I am not sure what you mean by this question, Jasmin
+As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
 model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
@@
 
-To further explore the parameter space, we generated two dimensional surface plots which reveal 
-the subsets of parameter space where coexistence is possible with ranging degrees of ATA 
+To further explore model parameter space, we generated two dimensional surface plots which reveal 
+some subsets of parameter space where coexistence is possible with ranging degrees of ATA 
 contributions and where ATA rescue occurs. We evaluated coexistence under two varying 
@@
 \end{figure}
-%***DAN: Dan has edited this caption
 
+%***DAN: Jasmine, some changes are needed to the figure itself. Please make the text in the legend bigger.
+%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
+%(that's what it says in the legend),
+%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
+%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
+%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
+%"A".
 \begin{figure}
 \includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{The contribution of ATAs to diatom coexistence is meaningfully large compared to 
+\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} 
+is meaningfully large compared to 
 other components of the MCT decomposition. All panels show the IGR and the decomposition of
-the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
-two of the three parameters defining temperature oscillation, amplitude, period, and mean, 
-is fixed at the original coexistence values found by \cite{Descamps_2005} ($a=6, P=60, \theta_0=18$). The IGR and components of the decomposition
-are plotted against one varying parameter.
+the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
+%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
+%invading which, and, at the end of the 
+%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
+For each panel, 
+two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
+and mean, $\theta_0$) are fixed at the original coexistence values used by 
+\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, \theta_0=18^\circ$C). 
+%***DAN: I added units for a  and \theta_0 here, please check these are correct 
+The IGR and components of the decomposition
+are then plotted against the other parameter.
 The original value of the varying parameter and coexistence quantities at that value is indicated by the vertical dashed magenta lines.
@@
 and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.}  \label{fig:fig5}
+coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.
+%***DAN: Don't forget to fill in the standard error stuff
+}  \label{fig:fig5}
 \end{figure}

### 2022-01-10 paper/Paper.Rnw
edits to some of the G4 material

@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
-%***DAN: Jasmin, Descamps_2005 does not appear in refs.bib, I think. Please add it there or subtract the callout here, whatever is most appropriate.
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 %G3
-To achieve goal G3 from the Introduction i.e., to explore, using the lottery model, the likelihood that species can coexist with
+To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
 ATAs but not without, or vice versa, we plotted the IGR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
@@
 
-%***DAN: Methods paras here for G4, if we accomplish G4
+%***DAN: Methods paras here for G4
 To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
-coexistence which highlights the role ATAs can play, we consider the coexistence of diatom 
-species established by chemostat batch experiments \citep{Descamps_2005}.
+coexistence which highlights the role ATAs can play, we use a laboratory system of freshwater diatoms
+explored by \cite{Descamps_2005}.
 \citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
@@
 %***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
-Via simulations, \cite{Descamps_2005} showed that coexistence of \emph{C. pseudostelligera} 
-and \emph{F. crotonensis} occurs with oscillating temperature fluctuations, 
+After using lab experiments to parameterize the model, \cite{Descamps_2005} use simulations to
+show that coexistence of \emph{C. pseudostelligera} 
+and \emph{F. crotonensis} occurs when temperature fluctuates periodically, 
 $\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, 
 %***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
-defined by parameters $\theta_0 = 18, a = 6, P = 60$d. 
-\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contrirbuting mechanisms, 
+with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
+%***DAN: I added units. Please check.
+\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contributing mechanisms, 
 and here, we consider varying parameters of temperature fluctuations and further decompose the 
@@
 %put reverse in sup mat?
-%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to just consider 1 invading 2 for 
-%the lottery model, but cannot really do that here, right?
+%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
+%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
 %do we need to explain how we separated the asymmetries using the sims??
@@
 
-%***DAN: Jasmine, some changes are needed to the figure itself. Please make the text in the legend bigger.
+%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
 %It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
@@
 %notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
-%"A".
+%"A". Make the panel labels bigger, as well as the axis numbers and labels. 
 \begin{figure}
@@
 two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
-and mean, $\theta_0$) are fixed at the original coexistence values used by 
-\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, \theta_0=18^\circ$C). 
+and mean, $\theta_0$; see Methods) are fixed at the original values used by 
+\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
 %***DAN: I added units for a  and \theta_0 here, please check these are correct 
-The IGR and components of the decomposition
-are then plotted against the other parameter.
-The original value of the varying parameter and coexistence quantities at that value is indicated by the vertical dashed magenta lines.
-The contribution of ATAs (red line)
-has magnitude comparable to other components and, in these parameters, is always positive, indicating that ATAs can contribute meaningfully
-to diatom coexistence. In some instances, the IGR line (orange) falls 
+the IGR and components of the MCT decomposition
+are then plotted against the other parameter. The original value of the varying parameter 
+is indicated by the vertical dashed magenta lines. ATA contributions 
+had magnitude comparable to other components, 
+%***DAN: Make sure this is still true after the correct version of C is put in, and after we
+%resolve the potential problem with the wiggles on the plot. Reword if not.
+indicating that ATAs can contribute meaningfully
+to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
 between the ATA line (red)
-and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities ?.
+and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
+coexistence and exclusion of the weaker competitor. 
+%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
+%for these plots? If I'm right about that then please reword accordingly. 
+Standard errors of plotted quantities ?.
 %***DAN: Don't forget to fill in the standard error stuff
+%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
+%species as the invader.
 }  \label{fig:fig5}

### 2022-01-11 code/diatom/workflow.R
edit

@@ -63,7 +63,8 @@ getDelt <- function(a, P, Tbar, time, sims){
 	ep1 <- getep3(a, P, Tbar, time, reps=sims, sp=1)
 	ep2 <- getep3(a, P, Tbar, time, reps=sims, sp=2)
 	Delta1 <- ep1-ep2
-	Delta1$IGR <- sum(Delta1[,-7])
+	Delta1 <- Delta1[,-7]
+	Delta1$IGR <- sum(Delta1)
 	Delta1$map <- Delta1$epsECbrk/Delta1$IGR
 	Delta1$time <- ep1$time
 	return(Delta1)

### 2022-01-11 code/diatom/Figure5.R
pushing fig 5 stuff

@@ -7,7 +7,7 @@ if(dir.exists(fig_loc)==FALSE){
 fig5 <- paste(fig_loc,"fig5.pdf",sep="")
 
 vAmp <- readRDS(paste(dat_loc,'varyingAmplitude.RDS',sep='')) #from explorespace.R
-vPer <- readRDS(paste(dat_loc,'varyingPerioda-6.RDS',sep=''))
+vPer <- readRDS(paste(dat_loc,'varyingPeriod.RDS',sep=''))
 vTbr <- readRDS(paste(dat_loc,'varyTbar.RDS',sep=''))
 
 
@@ -27,7 +27,7 @@ for (i in 1:3){
 	vdat <- varylist[[i]]
 	
 	#empty box
-	plot(0, yaxt='n', xlim=range(vdat[,8]), ylim=c(ymin, ymax), col='white', xlab=colnames(varylist[[i]])[8], ylab='', cex.lab=1.3, cex.axis=1.3)
+	plot(0, yaxt='n', xlim=range(vdat[,10]), ylim=c(ymin, ymax), col='white', xlab=colnames(varylist[[i]])[10], ylab='', cex.lab=1.3, cex.axis=1.3)
 	
 	#axis
 	if (i==1){
@@ -36,7 +36,7 @@ for (i in 1:3){
 	
 	#loop for lines
 	for (j in 1:7){
-		lines(vdat[,8], vdat[,j], col=col[j], lty=line[j])	
+		lines(vdat[,10], vdat[,j], col=col[j], lty=line[j])	
 	}
 	abline(h=0, col="lightgrey") #zero
 	abline(v=orig[i], col='magenta', lty=3)

### 2022-01-11 code/diatom/makefig5dat.R

@@ -1,0 +1,60 @@
+source("diatom/workflow.R")
+dat_loc <- "../results_numeric/"
+
+#amplitude
+a <- seq(1,6,length.out=100)
+Delta_a <- getDelt(a[1], 60, 18, 3000, 200)
+
+for (i in 2:length(a)){
+	Delta_a[i,] <- getDelt(a[i], 60, 18, 3000, 200)
+	print(i)
+}
+
+Delta_a$a <- a
+Delta_aa <- Delta_a[,-c(8,9)]
+saveRDS(Delta1, paste(dat_loc,'varyingAmplitude.RDS',sep=''))
+
+
+#######################################################################
+
+#vary P at a=6
+P <- seq(50,220,length.out = 100)
+
+DeltaP <- getDelt(6, P[1], 18, 2400, 200)
+
+for (i in 2:length(P)){
+  DeltaP[i,] <- getDelt(6, P[i], 18, 2400, 200)
+	print(i)
+}
+
+
+DeltaP$P <- P
+
+saveRDS(DeltaP, paste(dat_loc,'varyingPeriod.RDS',sep=''))
+
+#######################################################################
+
+Tbar <- seq(16,18,length.out = 100)
+
+DeltaT <- getDelt(6, 60, Tbar[1], 3000, 200)
+
+for (i in 2:length(Tbar)){
+	DeltaT[i,] <- getDelt(6, 60, Tbar[i], 3000, 200)
+	print(i)
+}
+
+
+DeltaT$Tbar <- Tbar
+
+saveRDS(DeltaT, paste(dat_loc,'varyTbar.RDS',sep=''))
+
+#######################################################################
+
+ep1 <- get_epsilons(4.415, 60, "temp", 500, 1)
+ep2 <- get_epsilons(4.415, 60, "temp", 500, 2)
+D <- ep1-ep2
+D <- D[-c(4,6)]
+D[7] <- sum(D)
+D[7]-D[5];D[7]
+#4.29-4.41 have ATA rescue
+

### 2022-01-11 paper/Paper.Rnw
addressing Dan's comments

@@
 and \emph{F. crotonensis} occurs when temperature fluctuates periodically, 
-$\theta(t)=\theta_0 (t) +  a\sin(2\pi t/P)$, 
+$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
 %***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
+%***JASMIN: You're right, it should not. Corrected
 with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
 %***DAN: I added units. Please check.
+%***JASMIN: Checked. Done. 
 \cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contributing mechanisms, 
@@
 %the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
+%***JASMIN: Right. Species in the lottery model only differed in mu but here they differ in they're VKQ functions.
+%*** JASMIN: Note to to run the reverse invader-resident scenario and put in SUpMat!
 %do we need to explain how we separated the asymmetries using the sims??
 %***DAN: I am not sure what you mean by this question, Jasmin
+%***JASMIN: I mean should we explain the methodology of the code we wrote that is implemented to compute the 
+% ATA contribution. The functions alignranks, normcor...  (in partialSharp_fxns.R) 
 As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
@@
 %***DAN: I added units for a  and \theta_0 here, please check these are correct 
+%***JASMIN: Done 
 the IGR and components of the MCT decomposition
@@
 and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. 
+coexistence and exclusion of the invader, \emph{F. crotonensis}. 
 %***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
 %for these plots? If I'm right about that then please reword accordingly. 
+%***JASMIN: Right, Done. 
 Standard errors of plotted quantities ?.
@@
 %species as the invader.
+An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
+\emph{F. crotonensis} is the resident.
 }  \label{fig:fig5}
@@
 \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature oscillation.
-The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger magnitudes signifying larger ATA contributions.
-In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and negative IGR.
-Thus, cyan areas indicate failure to coexist and magenta areas indicate coexistence with ATA contribution increasing with color intensity.
-The most intense areas of magenta indicate parameter subspaces where ATA rescue occurs.
-For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ >1 were capped at $|1|$.
-} \label{fig:fig6}
+\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature 
+oscillation. The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger 
+magnitudes signifying larger ATA contributions. In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/
+\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only 
+result from positive $\Delta_i^{[EC]}$ and negative IGR. Thus, cyan areas indicate failure to coexist and magenta areas 
+indicate coexistence with ATA contribution increasing with color intensity. The most intense areas of magenta indicate 
+parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ 
+>1 were capped at $|1|$. 
+An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
+\emph{F. crotonensis} is the resident.} \label{fig:fig6}
 \end{figure}

### 2022-01-11 paper/Paper.Rnw

@@
 and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
-resource model as follows:
-put model here
+resource model as follows.
+Letting $\frac{dS}{dt}$ and $\frac{dx_j}{dt}$ denote the rates of resource uptake and reproduction for $j=1,2$ the
+model equations are
+\begin{equation}
+\frac{dS}{dt} = D(S_0 - S) - Q_1x_1 \frac{V_1S}{K_1 + S} - Q_2x_2 \frac{V_2S}{K_2 + S}
+ \frac{dx_j}{dt} = x_j \frac{V_jS}{K_j+S}-D_xj, $j=1,2$.
+ \end{equation}
+ Here, $S$ is a extracellular silicate concentration, $S_0$ is the inflow concentration, $D$ is the outflow rate, and $x_j$ 
+ are the diatom densities. The temperature-dependent variables for resource consumption, $Q_j$, reproduction rate, 
+ $V_j$, and half-saturation constant, $K_j$ were obtained by fitting lines and curves to values obtained from the chemostat batch experiments \cite{Descamps_2005, Ellner_2016}. 
 %***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
+%***JASMIN: Done
 After using lab experiments to parameterize the model, \cite{Descamps_2005} use simulations to
@@
 \includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} 
-is meaningfully large compared to 
+\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} when 
+\emph{F. crotonensis} is the invader is meaningfully large compared to 
 other components of the MCT decomposition. All panels show the IGR and the decomposition of
@@
 %caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
+%***JASMIN: Done - ish. Need to make sup mat section 
 For each panel, 
@@
 %***DAN: Don't forget to fill in the standard error stuff
+
 %***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
 %species as the invader.
+%***JASMIN: Done
 An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 

### 2022-01-11 code/Figure1.R
changed panel labels

@@ -11,7 +11,7 @@ load(noise_loc)
 
 nt <- c("l", "s", "r") #nt = noise type
 blue <- rgb(0,0,0.545,0.3) #col2rgb("darkblue")
-
+labels <- c("(a)","(b)", "(c)")
 ##############################################################################################
 pdf(fig1_vertical_loc, width=2.8, height=7)
 
@@ -36,7 +36,7 @@ for (panel in 1:3){
   plot(noise[1:800,1], noise[1:800,2], col=blue, 
        ylab=NA, xlab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
   text(x=4,y=-3.8,labels=bquote(~ rho == .(round(P,5))), adj=1)
-  mtext(LETTERS[panel], side=3, line=-1.5, at=-3.7)
+  mtext(labels[panel], side=3, line=-1.6, at=-3.5)
   
   #marginals
   d1 <- density(noise[,1])
@@ -69,7 +69,7 @@ for (panel in 1:3){
   plot(noise[1:800,1], noise[1:800,2], col=blue, 
        ylab=NA, xlab=NA, xlim=c(-4,4), ylim=c(-4,4), pch=16, cex=1.5)
   text(x=4,y=-3.8,labels=bquote(~ rho == .(round(P,5))), adj=1)
-  mtext(LETTERS[panel], side=3, line=-1.5, at=-3.7)
+  mtext(labels[panel], side=3, line=-1.6, at=-3.5)
   
   #marginals
   d1 <- density(noise[,1])

### 2022-01-11 code/Figure3.R

@@ -62,7 +62,7 @@ for (m in 1:nm){
   for (d in 1:nd){
     res <- append(res,dePlot1(noise_loc, mudif[m], delta[d], xaxt="n"))
     axis(1, labels=ifelse(n>12, yes=TRUE, no=FALSE), tick=TRUE)
-    mtext(LETTERS[n], side=3, line=-1.45, at=0.5)
+    mtext(paste0("(", letters[n],")"), side=3, line=-1.45, at=0.7)
     
     if(n<5){
       mtext(paste(delta[d]), side=3, line=0.75, font=2, cex=0.8)
@@ -118,7 +118,7 @@ for (m in 1:nm){
   for (d in 1:nd){
     res <- append(res,dePlot1(noise_loc, mudif[m], delta[d], qij=TRUE, xaxt="n"))
     axis(1, labels=ifelse(n>12, yes=TRUE, no=FALSE), tick=TRUE)
-    mtext(LETTERS[n], side=3, line=-1.45, at=0.5)
+    mtext(paste0("(", letters[n],")"), side=3, line=-1.45, at=0.7)
     
     if(n<5){
       mtext(paste(delta[d]), side=3, line=0.75, font=2, cex=0.8)

### 2022-01-11 code/Figure4.R

@@ -62,7 +62,7 @@ for (s in 1:ns){
     res_loc <- paste(fig4numres_loc,m,".RDS", sep="")
     res <- append(res, list(dePlot2(noise_loc, res_exist=ifelse(file.exists(res_loc), yes=TRUE, no=FALSE), res_loc=res_loc, mudif_4,sigma[s], delta[d], xaxt="n")))
     axis(1, labels=ifelse(m>12, yes=TRUE, no=FALSE), tick=TRUE)
-    mtext(LETTERS[m], side=3, line=-1.45, at=-4.5, adj=1)
+    mtext(paste0("(", letters[m],")"), side=3, line=-1.45, at=-4.5, adj=1)
     
     if(m<5){
       mtext(paste(delta[d]), side=3, line=0.75, font=2, cex=0.8)

### 2022-01-19 code/diatom/workflow.R
add comment and fix typo

@@ -82,7 +82,7 @@ require(parallel)
 
 mapspace <- function(parmlist, sims, time){
 	
-	names(parmlist) <- c('a','P','Tbar')
+	#names(parmlist) <- c('a','P','Tbar')
 	#len <- unlist(lapply(parmlist, length))
 	#varlen <- len[len!=1]
 	#rows <- varlen[[1]]*varlen[[2]]
@@ -106,9 +106,9 @@ mapspace <- function(parmlist, sims, time){
 		}
 	}
 	#res <- data.frame(res)
-	resList <- mclapply(argsList, wrapDelt, mc.cores=7)
+	resList <- mclapply(argsList, wrapDelt, mc.cores=7) ###change cores here###
 	resdf <- data.frame(t(sapply(resList, function(X){c(X$IGR, X$epsECbrk, X$time, X$map)})))
-	parmsdf <- data.frame(matrix(unlist(argsList), ncol=5, byrow=T)[,-3:-4])
+	parmsdf <- data.frame(matrix(unlist(argsList), ncol=5, byrow=T)[,-4:-5])
 	
 	res <- cbind(parmsdf, resdf)
 	colnames(res) <- c("a", "P", "Tbar", "IGR", "ATA", "time", "map")

### 2022-01-19 paper/Paper.Rnw

@@
 
-%Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
-However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
-Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
-Threshold relationships can produce tail association, as we showed elsewhere.
-Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
-
+\underline{summary - include 1-2 sentences about misnomer}
 We produced and applied a mathematical extension to MCT that, for the first time, 
 considers ATAs for the coexistence of species and maintenance of biodiversity. 
+Implementing new methods of MCT \citep{Ellner_2019, Ellner_2016}, we provided a further tertiary 
+decomposition to MCT and decompose the storage effect into two new terms, 
+$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, the contribution of correlation \emph{per se}. 
 The results of our application to a simple lottery model and 
-to a competitive diatom model show that the contribution of ATAs to coexistence 
-is non-negligible and meaningfully contributes to the coexistence of species. 
+to a competitive diatom model show that ATA contributions to storage effects
+are non-negligible and meaningfully contribute to the coexistence of species and thus the maintenance of biodiversity. 
+Our results reinforce that common methods of correlation fail to reflect and include 
+important information on the structure of the relationship between variables and amends
+coexistence to the list of ecological theory and phenomena that ATAs have been found to 
+be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
+Futhermore, though storage effects are typically described as the contribution
+of covariance between drivers to coexistence, commonly referred to as \emph{EC covariance},
+\citep{Chesson_2012, Ellner_2016, Ellner_2019}, we suggest that storage effects, instead, 
+be described as the contribution of the \emph{relationship} between the drivers.
+Measures of covariance do not fully capture all aspects of associativity throughout the bivariate 
+distribution of the variables; thus, it may be misleading to describe the contribution of the whole 
+EC (or other drivers') relationship as the contribution of EC (or other drivers') covariance.
+Accordingly, we have separated the mechanisms of storage effect 
+into contributions of ATAs and contributions of correlation \emph{per se}.
+
+The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
+
+Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
+Previous studies report storage effect as a contribution to coexistence of EC covariance; yet compute it as the contribution of the EC relationship (covariance and beyond; symmetric and asymmetric).
+
+\underline{ATA specialization}
+Not only is the contribution of ATA non-neglible but in some cases the presence of absence of ATA 
+can mediate the result of competition between species, i.e., coexistence or extirpation. 
 For some parameters, the contribution of ATAs was the driving mechanism of 
@@
 Subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists}.
-Our results reinforce that common methods of correlation fail to reflect and include 
-important information on the structure of the relationship between variables and amends
- coexistence to the list of ecological theory and phenomena that ATAs have been found to 
- be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.
-
- We provided a further tertiary decomposition to MCT, 
-extending the framework put forward by \citep{Ellner_2019}. 
-We decompose the storage effect into two new terms, 
-$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, 
-the contribution of correlation \emph{per se}. 
-Previously, storage effects were described as the contribution
-of covariance between the environment and competition to 
-coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019}. 
-However, the description of storage effects as the covariance between drivers
-is a misnomer in terminology, and we propose that storage effects, instead, 
-should be described as the contribution of the relationship between drivers. 
-All aspects and of the relationship between environmental response and competitive pressure 
-are not fully captured by only the covariance between them because a measure of covariance 
-does not capture the structure of associativity throughout the distribution of the relationship 
-and any asymmetries there may be. 
-Thus, it is misleading to describe the contribution of the whole EC relationship 
-as the contribution of EC covariance.
-In addressing this inconsistency, we have separated the mechanisms of storage effect 
-into the contribution of ATA [to storage effect/coexistence] and the contribution of 
-correlation \emph{per se} [to storage effect/coexistence].
-
+The parameter subspaces of ATA rescue and ATA exclusion suggest that some species may be ATA specialists  - only persisting in habitat patches where ATAs contribute to its coexistence. 
+%I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
+%where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
+%similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
+%there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
+%are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
+
+\underline{future work: study ATAs in environmental variables and how theyre changing - tie in with specialist}
+%In the Discussion, I intend to talk about how possible changes in tail association 
+%patterns of env vars in nature may influence coexistence of species. It won't influence the 
+%other mechanisms, though changes in the extreme values of environmental variables, known
+%to be occurring, might. 
+%Should also mention this second point in the Discussion as something that merits further research.
+%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
+%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
+%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
+%to also look at the other mechanisms. May still need a sup mat laying out all the math.
+
+%Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
 Consistent with the mechanisms of biodiversity maintenance and specifically 
@@
 
-Outline
-
-\underline{Paragraph summarizing results of G1 and how it compares and builds on prior theory.}
-
-%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
-%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
-%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
-
-%Need a para about multi-species models - how would things change? One difference is there are more different
-%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
-%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
-We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
-Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
-
-\underline{paragraph summarizing results of G2 and how it compares and builds on prior theory}
-Furthermore, we have shown not only that is it possible to decompose storage effect into the effect of correlation \emph{per se} and the effect of ATAs, but this new term that holds the contribution of not only the correlation but the structure of the correlation is a non-negligible term, in a simple lottery model.
-
-Seemingly consistent with fluctuation dependent mechanisms, 
-as $\sigma$, the variation in the environment, increased, 
-so did the the magnitudes of $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$.
 This may suggest that, if environmental fluctuations become more variable, 
@@
 
-%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
-
-\underline{paragraph summarizing results of G3 and how it compares and builds on prior theory}
-Not only is the contribution of ATA non-neglible but in some cases the presence of absence of ATA can mediate the result of competition between species, i.e., coexistence or extirpation. 
+\underline{E and vital rates}
+%Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
+However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
+Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
+Threshold relationships can produce tail association, as we showed elsewhere.
+Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
+
+\underline{the new effects are not "weak noise" effects}
+if sigma really small, ATA would not matter?
+%go back to emails
+%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
+%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
+%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
+From proposal: 
+Classical analytic derivations of MCT using Taylor expansion approximations completely miss out
+on any additions to coexistence beyond simple correlations. 
+In fact, storage effects in analytic MCT is computationally equal to my new term, Δ^([E||C]), 
+the contributions of correlation per se. In newer, simulation based MCT, storage effect is 
+the sum of my two new terms, Δ^((EC))=Δ^([E||C])+Δ^([EC]), obscuring the effects of ATAs
+
+\underline{multispecies models}
+%Need a para about multi-species models - how would things change? One difference is there are more different
+%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
+%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
+We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
+Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
+
+\underline{caveats and future directions}
+%Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
+%mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
+%the importance of ATA for coexistence mediated by these factors.
+%Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
+%solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
+%how to say something about how future work may be able to explore the importance of ATA in that context?
+
+%Also need to talk about how this stuff will apply to data.
+we used simple/extreme ATA
+simple model and simple lab system
+future work needed to assess the extent of how much ATA are important for complex field systems
+what systems might be good? suggest systems and their important E vars.
+
 Our investigation is a simple one; 
@@
 but varying degrees of ATA may produce a similar result as well.
-%maybe the above sentence should go in caveats/assumptions paragraph 
-ATA specialist? Does this go in this paragraph
-
-The parameter subspaces of ATA rescue and ATA exclusion suggest that some species may be ATA specialists  - only persisting in habitat patches where ATAs contribute to its coexistence. 
-%I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
-%where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
-%similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
-%there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
-%are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
-
-\underline{paragraph summarizing results of G4 and how it compares and builds on prior theory}
-The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
-
-Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
-Previous studies report storage effect as a contribution to coexistence of EC covariance; yet compute it as the contribution of the EC relationship (covariance and beyond; symmetric and asymmetric).
-
-\underline{how results apply to or inform patterns or management questions in real life}
-ATA specialists?
-climate change and ATA
-
-\underline{caveats/assumptions that could change result}
+
 We used the most extreme version of asymmetric noise, i.e., perfect correlation changing to very low correlation at exactly the half point of the distribution, and a simple lottery model.
+
+%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
 
@@
 %Search for comments in the rest of the text indicating things that need to be done in the Discussion
-
-%Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
-%mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
-%the importance of ATA for coexistence mediated by these factors.
-
-%Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
-%solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
-%how to say something about how future work may be able to explore the importance of ATA in that context?
-
-%Also need to talk about how this stuff will apply to data.
-
-%In the Discussion, I intend to talk about how possible changes in tail association 
-%patterns of env vars in nature may influence coexistence of species. It won't influence the 
-%other mechanisms, though changes in the extreme values of environmental variables, known
-%to be occurring, might. Should also mention this second point in the Discussion as something that merits further research.
-%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
-%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
-%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
-%to also look at the other mechanisms. May still need a sup mat laying out all the math.
-
-%Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
 

### 2022-01-20 paper/Paper.Rnw
fixed some IGR that needed \text{}

@@
 parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total 
-IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/IGR$ and 
+IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ and 
 higher color intensity indicate greater contribution of ATAs to the IGR and positive cases 
-where $\Delta_i^{[EC]}>IGR$ indicate ATA rescue.
+where $\Delta_i^{[EC]}>\text{IGR}$ indicate ATA rescue.
 
@@
 Specific to this model and parameter ranges, because $\Delta_i^{[EC]}$ is always positive, all positive values of $\Delta_i^{[EC]}/\text{IGR}$ indicate coexistence and negative values of $\Delta_i^{[EC]}/\text{IGR}$ indicate failure of coexistence.
-Futhermore, ATA exlusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>IGR$.
+Futhermore, ATA exlusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>\text{IGR}$.
 To preserve simplicity and readibility of the plots, $|\Delta_i^{[EC]}/\text{IGR}>1|$ were capped at $|1|$.

### 2022-01-20 paper/SupMat.Rnw
starting new SI section on methods for quantifying ATA in empirical applicatoins
added empirical applications section

@@
 
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{section}{0}
+\renewcommand{\thesection}{S\arabic{section}}
+
 %new commands and so on
@@
 
-%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decide not to use)
+%attempt 3 at nat and sharp (see Paper.Rnw for attempts 1 and 2, which I decided not to use)
 \newcommand{\nat}{%
@@
 \newcommand{\olep}{\overline{\epsilon}}
+\newcommand{\olE}{\overline{E}}
+\newcommand{\olC}{\overline{C}}
 
@@
 
+\section{Decomposing storage effect in empirical applications}\label{sect:Empircal_app}
+
+The decomposition of storage effects and quantification of contributions of ATAs depends on 
+the removal of asymmetric tail associations from the relationship between
+coexistence drivers while maintaining their correlation and marginal distributions. 
+In the paper section \ref{MT-section:theory}, we describe analytical methods for generating the partial sharp variables, $E_i^\nat$ and $C_{i \bs i}^\nat$, which are distributed in the same way as the defined $E_i$ and $C_{i \bs i}$, but are symmetric in the tails and have a normal copula structure.
+The protocol for samples of coexistence drivers in empirical applications is as follows.
+Let $\olE_i$ be a sample of a environmental fluctuation and $\olC_{i \bs i}$ be a sample of competitive pressure.
+Let $\Theta_{\olE}$ and $\Theta_{\olC}$ be standard normal distributions obtained by normalizing the ranks of $\olE_i$ and $\olC_{i \bs i}$,
+such that $\Theta_{\olE} = \frac{rank(\olE_i)}{n_{\olE_i}+1} \circ \varphi^{-1}$ and $\Theta_{\olC} = \frac{rank(\olC_{i \bs i})}{n_{\olC_{i \bs i}}+1} \circ \varphi^{-1}$, where $\varphi$ is the CDF of a standard normal distribution.
+The normalized correlation of $\olE_i$ and $\olC_{i \bs i}$, denoted by $r$, is the covariance of $\Theta_{\olE}$ and $\Theta_{\olC}$.
+We then simulate an $S$ number of bivariate normal distributions $N_s\left(\vec{0}, \Sigma \right)$
+with covariance matrix $\Sigma = \begin{pmatrix} 1&\ddot{\rho} \\ \ddot{\rho}&1\end{pmatrix}$. 
+The columns of $(\olE_{i}, \olC_{i \bs i})$ are permutated and saved $S$ amount of times, 
+such that the rank of each column matches with the rank of each column of $N_s$. 
+To obtain $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$, the median of the permuations is computed. 
+Here, $\ddot\rho$ can be chosen a number of ways such that the Pearson, Spearman, or Kendall 
+correlation of $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$ matches that of $(\olE_{i}, \olC_{i \bs i})$
+or such that $\ddot\rho = r$. The code for this protocol is in \url{"diatom/partialSharp_fxns.R"} and was used in our application to the diatom chemostat system.
+
 \clearpage

### 2022-01-20 paper/Paper.Rnw
writing discussion: paragraph 3: future work in changign ATAs

@@
 driving mechanism of the failure of coexistence and the lack of ATAs would result in 
-a positive IGR. 
-Subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists}.
-The parameter subspaces of ATA rescue and ATA exclusion suggest that some species may be ATA specialists  - only persisting in habitat patches where ATAs contribute to its coexistence. 
+a positive IGR. Parameter
+subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists} - species which only persist in habitat patches where ATAs contribute to its coexistence.
 %I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
@@
 \underline{future work: study ATAs in environmental variables and how theyre changing - tie in with specialist}
+
+Our study highlights the importance of ATAs for ecological phenomena 
+and specifically for coexistence and the maintenance of biodiversity. 
+The occurence of ATAs among environmental variables themselves, 
+their contribution to coexistence, and their changes through space and time
+can and should be explored further. 
+Distinct increases in the magnitude of ATA contribution with increasing
+environmental fluctuations, as shown in our results and
+figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), suggests that ATAs and their effects
+may become more important to consider for coexistence and possibly other 
+ecological phenomena as environmental conditions evidently
+become more variable.
+As described in section \ref{section:introduction}, changes in the ATAs between environmental
+variables may occur as the climate becomes more variable and extreme events become more 
+expansive and extensive, causing relationships between separated events to become tighter at one end or another and thus asymmetric. Additionally, 
+if ATA specialist exist, their persistance may
+become at stake as the mechanism driving their coexistence changes with the climate.
+Thus it is of interest to further study how tail associations of environmental variables are changing and how this might 
+alter competition, coexistence, and the maintenance of biodiversity generally.
+Our investigation brings to attention the consequences of such
+possible \emph{relationships} in the extremes; however, 
+simpler changes in extreme events themselves such as
+intensity and frequency and their influence on coexistence also merit future work. 
+Possible changes in tail association
+patterns of environmental variables in nature can
+influence the coexistence of species through the storage effect and not other mechanisms, but
+extreme and changing climate is probable to affect other mechanims in other ways.
 %In the Discussion, I intend to talk about how possible changes in tail association 
@@
 %Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
-Consistent with the mechanisms of biodiversity maintenance and specifically 
-fluctuation-dependent mechanisms (cite chesson), higher values of environmental fluctuation 
-resulted in higher IGR’s. Both of our new terms, $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$ 
-also increased in magnitude with increased environmental fluctuation. Increased environmental 
-fluctuation is reflected in increased $\sigma$ in the lottery model and increased $a$ in the 
-competitive diatom model.   mean temp and period?
-
-Distinct increase in contribution of ATAs to coexistence as environmental fluctuations increase may suggest that ATAs may become more important to consider for coexistence and possibly other ecological phenomena as the environment becomes more variable, a consequence of climate change and extreme climatic events.
-
-This may suggest that, if environmental fluctuations become more variable, 
-the effect of ATAs becomes more important and non-negligible (rescue or exclusion). 
-{Might be interesting to study how tail associations of environmental variables are changing and how this might 
-alter competition, coexistence, and the maintenance of biodiversity generally.} 
+%onsistent with the mechanisms of biodiversity maintenance and specifically 
+%luctuation-dependent mechanisms (cite chesson), higher values of environmental fluctuation 
+%esulted in higher IGR’s. Both of our new terms, $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$ 
+%lso increased in magnitude with increased environmental fluctuation. Increased environmental 
+%luctuation is reflected in increased $\sigma$ in the lottery model and increased $a$ in the 
+%ompetitive diatom model.   mean temp and period?
+
 %This would be a whole potential 
@@
 %study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
+
 From proposal:
+
 Classical analytic derivations of MCT using Taylor expansion approximations completely miss out
 on any additions to coexistence beyond simple correlations. 
-In fact, storage effects in analytic MCT is computationally equal to my new term, Δ^([E||C]), 
+In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E\natC]}$, 
 the contributions of correlation per se. In newer, simulation based MCT, storage effect is 
-the sum of my two new terms, Δ^((EC))=Δ^([E||C])+Δ^([EC]), obscuring the effects of ATAs
+the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E\natC]}+\Delta^{[EC]}$, obscuring the effects of ATAs.
 

### 2022-01-25 paper/Paper.Rnw
edited caption of fig 5 and 6 to indicate coexistence only depends on sp1 IGR
edits to G4 methods paragraphs complete for now, over to Jasmin to review and compleete correspind SI sections

@@
 \author[a]{Pimsupa Jasmin Albert}
-\author[a,b]{Daniel C. Reuman}
+\author[a,b,*]{Daniel C. Reuman}
 
@@
 \affil[b]{Laboratory of Populations, Rockefeller University}
-
-%***Need to indicate corresponding authors
+\affil[*]{Corresponding author, reuman@ku.edu}
 
@@
 
-%***Add to these
 \keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effects}
@@
 $E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
-$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated in the sense
-described in the Introduction.
-%***DAN: This means you have to introduce this well, and explain why the way we define tail assoc is
-%the *right* way
+$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
 For those familiar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
@@
 against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
-values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. First, we showed 
-that, for the 
+values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. 
+First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
+it was then sufficient to decompose the invasion growth rate of the weaker competitor, species 1. 
+Second, we showed that, for the 
 lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
 instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
-against $\mu_1 - \mu_2$. Second, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
+against $\mu_1 - \mu_2$. 
+Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
 for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
@@
 as well as the
-IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$. 
+IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
 For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
@@
 To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
-coexistence which highlights the role ATAs can play, we use a laboratory system of freshwater diatoms
-explored by \cite{Descamps_2005}.
-\citep{Descamps_2005} used measurements of growth of \emph{Cyclotella pseudostelligera} 
-and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a competitive 
-resource model as follows.
-Letting $\frac{dS}{dt}$ and $\frac{dx_j}{dt}$ denote the rates of resource uptake and reproduction for $j=1,2$ the
-model equations are
-\begin{equation}
-\frac{dS}{dt} = D(S_0 - S) - Q_1x_1 \frac{V_1S}{K_1 + S} - Q_2x_2 \frac{V_2S}{K_2 + S}
- \frac{dx_j}{dt} = x_j \frac{V_jS}{K_j+S}-D_xj, $j=1,2$.
- \end{equation}
- Here, $S$ is a extracellular silicate concentration, $S_0$ is the inflow concentration, $D$ is the outflow rate, and $x_j$ 
- are the diatom densities. The temperature-dependent variables for resource consumption, $Q_j$, reproduction rate, 
- $V_j$, and half-saturation constant, $K_j$ were obtained by fitting lines and curves to values obtained from the chemostat batch experiments \cite{Descamps_2005, Ellner_2016}. 
-%***DAN: Jasmin, please add the model. Hard to evaluate the text without it.
-%***JASMIN: Done
-After using lab experiments to parameterize the model, \cite{Descamps_2005} use simulations to
-show that coexistence of \emph{C. pseudostelligera} 
-and \emph{F. crotonensis} occurs when temperature fluctuates periodically, 
+coexistence which highlights the role ATAs can play, we use a laboratory chemostat
+system of freshwater diatoms explored by \cite{Descamps_2005}.
+Those authors used measurements of growth rates of \emph{Cyclotella pseudostelligera} 
+and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a resource
+competition model. The model is 
+\begin{align}
+\frac{dS}{dt} &= D(S_0 - S) - Q_1x_1 \frac{V_1 S}{K_1 + S} - Q_2 x_2 \frac{V_2 S}{K_2 + S} \\
+\frac{dx_j}{dt} &= x_j \frac{V_j S}{K_j+S}-D x_j, 
+\end{align}
+for $j=1,2$, where $S$ is a extracellular silicate concentration in the chemostat, $x_1$ is the population density
+of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
+$D$ is the chemostat outflow rate, $S_0$ is concentration of silicates in the chemostat
+inflow, the $Q_j$ are amounts of silicate per cell, the $V_j$ are maximum population growth rates, and
+the $K_j$ are half-saturation constants for the dependence of growth rates on nutrient concentration.
+The temperature-dependent parameters $Q_j$, $V_j$, and $K_j$ were obtained by 
+\cite{Descamps_2005} through batch experiments and curve fitting or interpolation.
+\cite{Descamps_2005} showed that permanent coexistence of the two species occurs when 
+temperature fluctuates periodically, 
 $\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
-%***DAN: Jasmin, should \theta_0 depend on t? It should not, right? Please fix if I'm right.
-%***JASMIN: You're right, it should not. Corrected
 with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
-%***DAN: I added units. Please check.
-%***JASMIN: Checked. Done. 
-\cite{Ellner_2016, Ellner_2019} broke down their coexistence into the contributing mechanisms, 
-and here, we consider varying parameters of temperature fluctuations and further decompose the 
-coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
-of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
-$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
-environmental fluctuation is temperature fluctuation. We first considered the scenario where 
-\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
-the coexistence decomposition for \emph{F. crotonensis}. 
-%put reverse in sup mat?
-%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
-%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
-%***JASMIN: Right. Species in the lottery model only differed in mu but here they differ in they're VKQ functions.
-%*** JASMIN: Note to to run the reverse invader-resident scenario and put in SUpMat!
-%do we need to explain how we separated the asymmetries using the sims??
-%***DAN: I am not sure what you mean by this question, Jasmin
-%***JASMIN: I mean should we explain the methodology of the code we wrote that is implemented to compute the 
-% ATA contribution. The functions alignranks, normcor...  (in partialSharp_fxns.R) 
-As for G2, we first plotted the terms of equation \ref{eq:full_decomp} for the chemostat 
-model against independent parameters defining temperature oscillation, $\theta_0, a$, and 
-$P$, a total of three plots. For each plot, we set two of the parameters constant at the 
-original parameters which coexistence was experimentally found in \cite{Descamps_2005}, 
-and let the remaining temperature parameter vary. 
-We let $\theta_0$ vary in $[16,19]$, $a$ vary in $[1,6]$, and $P$ vary in $[20,120]$. 
-Relative importance of contributions was judged from the 
-plots to assess whether and to what extent ATA effects on coexistence were meaningfully 
-large for the lottery model, compared to 
-the contributions of other mechanisms. 
-The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
-
-To further explore model parameter space, we generated two dimensional surface plots which reveal 
-some subsets of parameter space where coexistence is possible with ranging degrees of ATA 
-contributions and where ATA rescue occurs. We evaluated coexistence under two varying 
-parameters and indicated $\Delta_i^{[EC]}/\text{IGR}$, a ratio of ATA contribution to total 
-IGR, with a color bar and contour lines. Greater magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ and 
-higher color intensity indicate greater contribution of ATAs to the IGR and positive cases 
-where $\Delta_i^{[EC]}>\text{IGR}$ indicate ATA rescue.
+\cite{Ellner_2016, Ellner_2019} broke down coexistence into contributing mechanisms
+according to (\ref{eq:previous_decomp}) for these same values of
+$\theta_0$, $a$ and $P$. We further decomponse
+coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
+of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
+made it sufficient to decompose only the invasion growth rate of species 1, for this 
+example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
+$i=2$, $j=1$. Further details of the model setup are in SI section \ref{SI-sect:diatom_model}
+and specifics of how our decomposition extends to this empirical case are in 
+SI section \ref{SI-sect:Empircal_app}.
+%***DAN: Jasmin, you had already created and started on the second of these SI sections, but the first one
+%is new, and needs to be filled in. Please make a start at that when you get a chance. I've put some 
+%notes there indicating what the content of that section should be.
+
+%***DAN: Jasmin, I edited the below paragraph pretty heavily, please check,
+%especially the quantitative statements at the end.
+To display results, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
+$\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
+used by \cite{Descamps_2005}. We also generated two-dimensional plots, where
+two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
+at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
+$\Delta_i^{[EC]}/\text{IGR}$ was displayed using color, to convey the 
+importance of the ATA contribution relative to the whole IGR. Larger magnitudes
+of this quantity indicate greater importance of ATAs. The case
+$\Delta_i^{[EC]}>\text{IGR}>0$ (which produces $\Delta_i^{[EC]}/\text{IGR}>1$)
+corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{IGR}<0$
+(which also produces $\Delta_i^{[EC]}/\text{IGR}>1$) corresponds to ATA exclusion.
+The condition $\Delta_i^{[EC]}/\text{IGR}>1$ encompasses the cases for which ATA
+exclusion or rescue occurred, though the condition does not delineate, by itself,
+which of these was the case.
 
@@
 on any additions to coexistence beyond simple correlations. 
-In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E\natC]}$, 
+In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
 the contributions of correlation per se. In newer, simulation based MCT, storage effect is 
-the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E\natC]}+\Delta^{[EC]}$, obscuring the effects of ATAs.
+the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, obscuring the effects of ATAs.
 
@@
 %***DAN: Don't forget to fill in the standard error stuff
-
 %***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
@@
 %***JASMIN: Done
-An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
-\emph{F. crotonensis} is the resident.
-}  \label{fig:fig5}
+Coexistence in this system and within 
+our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
+show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
+\emph{F. crotonensis} is the resident.}  \label{fig:fig5}
 \end{figure}
@@
 >1 were capped at $|1|$. 
-An analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
+Coexistence in this system and within 
+our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
+show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
 \emph{F. crotonensis} is the resident.} \label{fig:fig6}

### 2022-01-25 paper/SupMat.Rnw
edits to G4 methods paragraphs complete for now, over to Jasmin to review and compleete correspind SI sections
one more item needs to be added to the new sup mat section

@@
 %***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.
+%It has been added but Dan needs to edit it.
 
@@
 
-\section{Decomposing storage effect in empirical applications}\label{sect:Empircal_app}
+\section{Details of diatom model setup}\label{sect:diatom_model}
+
+%***DAN: Jasmin, please start on this section when you get a chance. Here are some topics that need
+%to be covered:
+
+%1) Precise delineation of the functions Q_i, V_i, K_i which were used. There are different options considered
+%all over the place in the code and by Ellner et al (2019) and by the original diatom study, so we have to
+%be very clear here. To give one example of the need for care, here, Ellner et al (2019) state on p. 6
+%of their paper that K_1 and K_2 were constant for the range of temperature considered, and they state
+%that they took constant values for these things. However, we used a larger range of temperatures. 
+%Depending on which of K1fun/K2fun versus K1flatfun/K2flatfun (from DDataAndVKQfuns.R) was actually used
+%may make a difference, so we have to specify. Note that even K1flatfun/K2flatfun were not actually flat
+%for the whole temperature range 6-24 degrees C, though they were for the range of temperature Ellner
+%et al considered. We should not say we used flat K1 and K2 functions unless we are sure we only ever
+%used temperatures in the flat ranges of these functions. Better just to describe the functions we
+%actually used, where the best way to describe them *may* be by plotting them. Anyway, what you need to
+%do is make sure you know what versions of each of these 6 functions your code actually used, and then
+%describe the functions precisely in this section.
+
+%2) You need to specify what choices were made for E and C. The text below, which I cut-and-pasted
+%from the main text, does some of this. Please use an edited version of that text here, edited as you
+%see fit if clarity needs to be improved and to fit the text in with the rest of this new section.
+%You ought to add a bit more exlanation to the text, so one understand why the choice of C make sense
+%and why r is what you've written there.
+
+%3) You need to outline how the quantities E_i and C_i\i were obtained from simulations.
+
+%***DAN: Some text and notes immediately below were cut and pasted from the main text, in case 
+%having this material helps in writing this new SI section
+%
+%and here, we consider varying parameters of temperature fluctuations and further decompose the 
+%coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
+%of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
+%$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
+%environmental fluctuation is temperature fluctuation. We first considered the scenario where 
+%\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
+%the coexistence decomposition for \emph{F. crotonensis}. 
+%put reverse in sup mat?
+%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
+%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
+%***JASMIN: Right. Species in the lottery model only differed in mu but here they differ in they're VKQ functions.
+%*** JASMIN: Note to to run the reverse invader-resident scenario and put in SUpMat!
+%do we need to explain how we separated the asymmetries using the sims??
+%***DAN: I am not sure what you mean by this question, Jasmin
+%***JASMIN: I mean should we explain the methodology of the code we wrote that is implemented to compute the 
+% ATA contribution. The functions alignranks, normcor...  (in partialSharp_fxns.R) 
+%***DAN: Yes, we are going to have to do that
+
+\section{Decomposing storage effects in the diatom case study}\label{sect:Empircal_app}
 
 The decomposition of storage effects and quantification of contributions of ATAs depends on 
-the removal of asymmetric tail associations from the relationship between
-coexistence drivers while maintaining their correlation and marginal distributions. 
+the removal of asymmetric tail associations from the bivariate distribution
+$(E_i,C_{i \bs i})$ while maintaining its correlation and marginal distributions.
+In the Theory section of the main text, we decribed methods for this removal process
+when the distributions of $E_i$ and $C_{i \bs i})$ are explicitly known, but for the 
+diatom example explored for goal G4 of the main text (see also SI section \ref{}), 
+we only have samples of $E_i$ and $C_{i \bs i})$
+
+;
+a modified distribution $E_i^\nat$ and $C_{i \bs i}^\nat$
+
 In the paper section \ref{MT-section:theory}, we describe analytical methods for generating the partial sharp variables, $E_i^\nat$ and $C_{i \bs i}^\nat$, which are distributed in the same way as the defined $E_i$ and $C_{i \bs i}$, but are symmetric in the tails and have a normal copula structure.
@@
 \end{figure}
-%***DAN: Jasmin, in the caption of Fig. 3 in the main text there is a statment about standard
-%error. We need another such statement here, but for the standard errors that applied here.
 

### 2022-01-25 paper/SupMat.Rnw
working on edits of SI section 9, not done yet

@@
 The decomposition of storage effects and quantification of contributions of ATAs depends on 
-the removal of asymmetric tail associations from the bivariate distribution
-$(E_i,C_{i \bs i})$ while maintaining its correlation and marginal distributions.
-In the Theory section of the main text, we decribed methods for this removal process
+the removal of asymmetric tail associations from the bivariate random variable
+$(E_i,C_{i \bs i})$, while maintaining its correlation and marginal distributions, to produce a new
+random variable $(E_i^\nat,C_{i \bs i}^\nat)$.
+In the Theory section of the main text, we decribed this removal process
 when the distributions of $E_i$ and $C_{i \bs i})$ are explicitly known, but for the 
-diatom example explored for goal G4 of the main text (see also SI section \ref{}), 
-we only have samples of $E_i$ and $C_{i \bs i})$
-
-;
-a modified distribution $E_i^\nat$ and $C_{i \bs i}^\nat$
+diatom example explored for goal G4 of the main text (see also SI section \ref{sect:diatom_model}), 
+we only have samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to
+produce a sample from $(E_i^\nat,C_{i \bs i}^\nat)$ in this modified context.
 

### 2022-01-26 paper/SupMat.Rnw
edited SI section 9 and a few other minor changes

@@
 \DeclareMathOperator{\Prob}{P}
+\DeclareMathOperator{\rk}{rank}
+\DeclareMathOperator{\nrk}{nrank}
 
@@
 The decomposition of storage effects and quantification of contributions of ATAs depends on 
-the removal of asymmetric tail associations from the bivariate random variable
+removing asymmetric tail associations from the bivariate random variable
 $(E_i,C_{i \bs i})$, while maintaining its correlation and marginal distributions, to produce a new
@@
 In the Theory section of the main text, we decribed this removal process
-when the distributions of $E_i$ and $C_{i \bs i})$ are explicitly known, but for the 
+when the distributions of $E_i$ and $C_{i \bs i}$ are known exactly, but for the 
 diatom example explored for goal G4 of the main text (see also SI section \ref{sect:diatom_model}), 
-we only have samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to
+we only had samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to
 produce a sample from $(E_i^\nat,C_{i \bs i}^\nat)$ in this modified context. 
-
-In the paper section \ref{MT-section:theory}, we describe analytical methods for generating the partial sharp variables, $E_i^\nat$ and $C_{i \bs i}^\nat$, which are distributed in the same way as the defined $E_i$ and $C_{i \bs i}$, but are symmetric in the tails and have a normal copula structure.
-The protocol for samples of coexistence drivers in empirical applications is as follows.
-Let $\olE_i$ be a sample of a environmental fluctuation and $\olC_{i \bs i}$ be a sample of competitive pressure.
-Let $\Theta_{\olE}$ and $\Theta_{\olC}$ be standard normal distributions obtained by normalizing the ranks of $\olE_i$ and $\olC_{i \bs i}$,
-such that $\Theta_{\olE} = \frac{rank(\olE_i)}{n_{\olE_i}+1} \circ \varphi^{-1}$ and $\Theta_{\olC} = \frac{rank(\olC_{i \bs i})}{n_{\olC_{i \bs i}}+1} \circ \varphi^{-1}$, where $\varphi$ is the CDF of a standard normal distribution.
-The normalized correlation of $\olE_i$ and $\olC_{i \bs i}$, denoted by $r$, is the covariance of $\Theta_{\olE}$ and $\Theta_{\olC}$.
-We then simulate an $S$ number of bivariate normal distributions $N_s\left(\vec{0}, \Sigma \right)$
-with covariance matrix $\Sigma = \begin{pmatrix} 1&\ddot{\rho} \\ \ddot{\rho}&1\end{pmatrix}$. 
-The columns of $(\olE_{i}, \olC_{i \bs i})$ are permutated and saved $S$ amount of times, 
-such that the rank of each column matches with the rank of each column of $N_s$. 
-To obtain $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$, the median of the permuations is computed. 
-Here, $\ddot\rho$ can be chosen a number of ways such that the Pearson, Spearman, or Kendall 
-correlation of $(\olE_{i}^\nat, \olC_{i \bs i}^\nat)$ matches that of $(\olE_{i}, \olC_{i \bs i})$
-or such that $\ddot\rho = r$. The code for this protocol is in \url{"diatom/partialSharp_fxns.R"} and was used in our application to the diatom chemostat system.
+Let $(E_{i,k},C_{i \bs i,k})$, $k=1,\ldots,T$ denote the samples from $(E_i,C_{i \bs i})$.
+Let $\rk(E_{i,k})$ be the rank of $E_{i,k}$ in the set 
+of $E_{i,k}$, $k=1,\ldots,T$, where the smallest value has rank $1$. Likewise 
+define $\rk(C_{i \bs i,k})$ as the rank of $C_{i \bs i,k}$ in the set
+of $C_{i \bs i,k}$, $k=1,\ldots,T$. Define the ``normalized rank''
+$\nrk(E_{i,k})$ as $\rk(E_{i,k})/(T+1)$, and likewise define 
+$\nrk(C_{i \bs i,k})$ as $\rk(C_{i \bs i,k})/(T+1)$. Letting $\varphi$ be the CDF
+of a standard normal distribution, the samples $\varphi^{-1}(\nrk(E_{i,k}))$, $k=1,\ldots,T$
+are standard normally distributed, as are the samples 
+$\varphi^{-1}(\nrk(C_{i \bs i,k}))$, $k=1,\ldots,T$. Let $\rho_i$ be the covariance 
+of these samples. We then define a bivariate normal random variable 
+$(e_i,c_{i \bs i})$ such that $\E(e_i)=\E(c_{i \bs i})=0$, $\var(e_i)=\var(c_{i \bs i})=1$,
+and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We used $\ddot{\rho}_i = \rho_i$.
+Generating a sample $(e_{i,k},c_{i \bs i,k})$, $k=1,\ldots,T$ from this distribution,
+we permute the ordered set $(E_{i,1},E_{i,2},\ldots,E_{i,T})$ so that the ranks of the 
+permuted ordered set equal the ranks of the corresponding elements of the ordered
+set $(e_{i,1},e_{i,2},\ldots,e_{i,T})$. We likewise permute the ordered set 
+$(C_{i \bs i,1},C_{i \bs i,2},\ldots,C_{i \bs i,T})$ so the ranks of the permuted result
+match the ranks of $(c_{i \bs i,1},c_{i \bs i,2},\ldots,c_{i \bs i,T})$.
+This step works as long as there are no ties among the $E_{i,k}$ or among
+the $C_{i \bs i,k}$. There were no ties when we applied these methods to the diatom system. 
+The two permuted ordered sets, here denoted $(E_{i,1}^\nat,E_{i,2}^\nat,\ldots,E_{i,T}^\nat)$
+and $(C_{i \bs i,1}^\nat,C_{i \bs i,2}^\nat,\ldots,C_{i \bs i,T}^\nat)$,
+when combine to make ordered pairs, form the desired sample from 
+$(E_i^\nat,C_{i \bs i}^\nat)$. Because $(E_{i,1}^\nat,\ldots,E_{i,T}^\nat)$
+is a permutation of $(E_{i,1},\ldots,E_{i,T})$ and 
+$(C_{i \bs i,1}^\nat,\ldots,C_{i \bs i,T}^\nat)$ is a permutation of
+$(C_{i \bs i,1},\ldots,C_{i \bs i,T})$, marginal distributions of our
+sample from $(E_i^\nat,C_{i \bs i}^\nat)$ were the same as those of our original 
+sample from $(E_i, C_{i \bs i})$. The choice of bivariate normal
+$(e_i,c_{i \bs i})$ ensures that our resulting sample shows symmetric tail association.
+Finally, we caution that sampling variation means any quantity computed from our resulting sample 
+$(E_{i,k}^\nat,C_{i \bs i,k}^\nat)$, $k=1,\ldots,T$ should be recomputed for 
+multiple such samples, all computed via the same process, and a median of 
+the resulting distribution of values should be used.
 

### 2022-01-26 paper/SupMat.Rnw
SM model setup

--- 2022-01-25 paper/SupMat.Rnw
@@
 %to be covered:
+Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for cell quota, reproduction rate, and half saturation, respectively, were used to paramaterize the competing diatom model outline in the paper.
+Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
+Descamps and Julien obtained estimated values for $Q_i$, $V_i$, and $K_i$ from single species nine-day batch experiments at four different temperatures (6, 12, 18, 24) (see Table 1 of Descamps and Julien) and further computed a linear interpolation to define temperature dependence functions of $Q_i(t)$, $V_i(t)$, and $K_i(t)$.
+These values and functions are indicated in Figure \ref{fig:VKQ} by open circles and dashed lines.
+Ellner 2016 and 2019 chose to slightly modify some values and functions, indicated in Figure \ref{fig:VKQ} by closed circles and solid lines.
 
@@
 
+\begin{figure}
+\includegraphics[width=.66\textwidth]{../results_figs/VKQfig.pdf}
+\caption[caption]
+{caption \label{fig:VKQ}}
+\end{figure}
+
 \clearpage

### 2022-01-26 paper/Paper.Rnw

@@
 these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\var(e_i)=\var(c_{i \bs i})=1$, and we denote $\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
+$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
+$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$

### 2022-01-26 paper/SupMat.Rnw
writing about E and C choices

@@
 %to be covered:
-Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for cell quota, reproduction rate, and half saturation, respectively, were used to paramaterize the competing diatom model outline in the paper.
+Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for 
+cell quota, reproduction rate, and half saturation, respectively, were used to 
+paramaterize the competing diatom model outline in the paper. 
 Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
-Descamps and Julien obtained estimated values for $Q_i$, $V_i$, and $K_i$ from single species nine-day batch experiments at four different temperatures (6, 12, 18, 24) (see Table 1 of Descamps and Julien) and further computed a linear interpolation to define temperature dependence functions of $Q_i(t)$, $V_i(t)$, and $K_i(t)$.
-These values and functions are indicated in Figure \ref{fig:VKQ} by open circles and dashed lines.
-Ellner 2016 and 2019 chose to slightly modify some values and functions, indicated in Figure \ref{fig:VKQ} by closed circles and solid lines.
+\cite{Descamps_2005} obtained estimated values for $Q_i$, $V_i$, and $K_i$ from 
+single species nine-day batch experiments at four different temperatures ($6^\circ$C, 
+$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and further
+computed a linear interpolation to define temperature dependence 
+functions of $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
+indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
+\cite{Ellne_2016, Ellner_2019} chose to slightly modify some values and functions 
+(to better align model with competition experiments), 
+%why did they make K_i's constant and do we need to adjust so that it
+%is constant over the range of temperatures considered 
+indicated in Figure \ref{fig:VKQ} by closed circles and solid lines. 
+These modifications do not result in meaningful differences when evaluating 
+coexistence at the parameter values used by \cite{Descamps_2005}.
+We chose to implement the modified functiones in our methods to quantify and 
+decompose coexistence.
 
-%1) Precise delineation of the functions Q_i, V_i, K_i which were used. There are different options considered
-%all over the place in the code and by Ellner et al (2019) and by the original diatom study, so we have to
 %be very clear here. To give one example of the need for care, here, Ellner et al (2019) state on p. 6
@@
 %and why r is what you've written there.
+
+Our decomposition is defined by $E(t)=\theta(t)$, 
+$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. 
+We note that our definition and selection of our "features" of interest are
+options among many and features need not be limited to $E$ and $C$ \citep{Ellner_2019}. 
+For example \cite{Ellner_2019} decomposes coexistence with respect to fluctuation in
+inflow silicate concentration, $S$, and temperature, and \cite{Ellner_2016} decomposes
+coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$. 
+Choice of drivers and their definitions are up to the investigator,
+but should real variables that influence population growth.
+Thus in our exploration, environmental fluctuation is temperature fluctuation.
 

### 2022-01-26 paper/SupMat.Rnw
edits... stopping for now

@@
 $12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and further
-computed a linear interpolation to define temperature dependence 
-functions of $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
+computed linear interpolations to
+functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
 indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
@@
 Our decomposition is defined by $E(t)=\theta(t)$, 
-$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. 
+$C_i(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_i)=\frac{V_i(E)}{C_i} -D$. 
 We note that our definition and selection of our "features" of interest are
@@
 coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$. 
+Like \cite{Ellner_2016}, we set $C_i(t)=(K_i(t)+S(t))/S(t)$ to reflect 
+potential versus actual growth of the diatom populations. 
 Choice of drivers and their definitions are up to the investigator,
-but should real variables that influence population growth.
-Thus in our exploration, environmental fluctuation is temperature fluctuation.
+but should refelct real variables that influence population growth.
 
 %3) You need to outline how the quantities E_i and C_i\i were obtained from simulations. 
+To obtain $E_i$ and $C_{i \bs i}$, the model was simulated for a $T=3600$ number of timesteps, setting initial population size of invader i to be zero and initial population of the resident, $r=20$.
 
@@
 and $(C_{i \bs i,1}^\nat,C_{i \bs i,2}^\nat,\ldots,C_{i \bs i,T}^\nat)$,
-when combine to make ordered pairs, form the desired sample from 
+when combined to make ordered pairs, form the desired sample from 
 $(E_i^\nat,C_{i \bs i}^\nat)$. Because $(E_{i,1}^\nat,\ldots,E_{i,T}^\nat)$

### 2022-01-27 paper/Paper.Rnw
small things and some useful comments

@@
 and $\mu_1$ and $\mu_2$ can be regarded as two points in it. But we instead define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
-Points of the trait-difference space characterize possible combinations of traits of the two competing species in our model. For fixed values of 
+Points of the trait-difference space characterize possible combinations of traits of the two 
+competing species in our model. For fixed values of 
 $\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
@@
 $\mu_2$, i.e., for closer-to-equal competitors. 
-In the $\mu_1=\mu_2$ case (panels A-D), for which the two species were equal 
+In the $\mu_1=\mu_2$ case (panels a-d), for which the two species were equal 
 competitors, storage effects (blue plus red
@@
 Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
-can coexist with ATAs, but not without, or vice versa), our results show that ATA can, indeed, sometimes
-make the difference between coexistence and exclusion; both ``ATA rescue'' and ``ATA
+can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
+make the difference between coexistence and exclusion: both ``ATA rescue'' and ``ATA
 exclusion'' (see Methods) are possible.
 For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
-line and the $x$ axis for some values of $\sigma$, e.g., see panel H. 
+line and the $x$ axis for some values of $\sigma$, e.g., see panel h. 
 Such cases correspond to parameter combinations
 for which the presence or absence of ATAs determine whether the species coexist or not. In the 
-case of panel H, for values of $\sigma$ around $4$, 
+case of panel h, for values of $\sigma$ around $4$, 
 species 1 is excluded with ATAs but the two species can coexist without ATAs. 
@@
 ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
-more common for larger $\delta$. Although bands of ATA rescue or exclusion on Fig.
+more common for larger $\delta$. Although some bands of ATA rescue or exclusion on Fig.
 \ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
@@
 \caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
-competitor; both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
-All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[E || C]}$ in equation 
+competitor: both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
+All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[EC]}$ in equation 
 (\ref{eq:full_decomp})], and the difference of these quantities. 
@@
 Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
-so are not displayed. The dashed red line on panel L corresponds to the parameters
+so are not displayed. The dashed red line on panel (l) corresponds to the parameters
 used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
@@
 %"A". Make the panel labels bigger, as well as the axis numbers and labels. 
+%***DAN: If I interpret correctly that the blue lines on Fig. 5 are the ATA contribution (needs to be 
+%changed because you use a red line on other figs), there are large areas on panels a and c where the 
+%IGR is between the ATA-contribution line and then x axis. Those are areas of ATA rescue (though you only
+%know this after looking at the analogous plot for the other species). I think in 
+%the next version of the figure you should consider highlighting them in a similar way to how you did 
+%in Fig. 4. Those are a couple nice wide zones of ATA rescue! Let me know if you want to talk about this.
 \begin{figure}

### 2022-01-27 code/MASTER.R
line to make VKQ plots

@@ -19,6 +19,8 @@ source("./diatom/Figure5.R")
 #FIGURE 6
 source("./diatom/Figure6.R")
 
+#VKQ figure (SI)
+source("./diatom/VKQplots.R")
 
 
 

### 2022-01-27 code/diatom/VKQplots.R
added lebels and legend

@@ -17,24 +17,38 @@ plotVKQ <- function(var, Tvals, dat1, dat2, dat1mod, dat2mod, fun1, fun2, fun1mo
   par(mgp=c(2, 0.5, 0),tcl=-0.3)
   
   plot(0, xlab='', ylab=ylab, xlim=range(Tvals), ylim=ylim, xaxt='n', cex.lab=1.5, cex.axis=1.3)
-  points(Tvals,dat1, cex=1.3)
-  points(Tvals,dat1mod, pch=19, cex=0.5)
-  lines(t, fun1(t), lty=2)
-  lines(t, fun1mod(t))
-
-  points(Tvals, dat2, col='red', cex=1.3)
-  points(Tvals,dat2mod, pch=19, cex=0.5, col='red')
-  lines(t, fun2(t), lty=2, col='red')
-  lines(t, fun2mod(t), col='red')
+  col <- "#2a2a8c" #sp1
+  points(Tvals,dat1, cex=1.3, col=col)
+  points(Tvals,dat1mod, pch=19, cex=0.5, col=col)
+  lines(t, fun1(t), lty=2, col="#6060bf")
+  lines(t, fun1mod(t), col=col)
+  
+  col <- "#8c2a2a" #sp2
+  points(Tvals, dat2, col=col, cex=1.3)
+  points(Tvals,dat2mod, pch=19, cex=0.5, col=col)
+  lines(t, fun2(t), lty=2, col='#bf6060')
+  lines(t, fun2mod(t), col=col)
 }
 
 pdf(VKQfig, height=7, width=5)
 par(mfrow=c(3,1), mar=c(0.7, 4, 1, 1), oma=c(3, 0,0,0))
 plotVKQ("V", Tvals, V1data, V2data, V1data, V2dataMod, V1fun, V2fun, V1quad, V2modfun)
+mtext("(a)", 3, -1.5, at=6)
+legend("bottomleft", legend=c("sp1","sp2"),  fill=c("#2a2a8c", "#8c2a2a"), inset=c(0,.3),
+       bty='n', border='white')
+legend("bottomleft", legend=c("original data","modified data","original function","modified function"), 
+       lty=c(NA, NA, 2, 1), pch=c(1, 19, NA, NA), bty='n', pt.cex=c(1.3, .5,NA,NA))
 plotVKQ("K", Tvals, K1data, K2data, K1flat, K2flat, K1fun, K2fun, K1flatfun, K2flatfun)
+mtext("(b)", 3, -1.5, at=6)
[…]

### 2022-01-27 code/diatom/DataAndVKQfuns.R
option to make forceChemo run with unmodified VKQ funs

@@ -44,6 +44,7 @@ V2modfun <-  approxfun(Tvals,V2dataMod,rule = 2)
 forceChemo <- function(t,y,parms) {
 	temp <- parms["Tbar"] + parms["a"]*sin(2*pi*t/parms["P"]); 
 	V1 <- V1quad(temp); V2<- V2modfun(temp); K1<- K1flatfun(temp); K2 <-  K2flatfun(temp) 
+	#V1 <- V1fun(temp); V2<- V2fun(temp); K1<- K1fun(temp); K2 <-  K2fun(temp) 
 	Q1 <- Q1fun(temp); Q2 <-  Q2fun(temp);
 	R <- y[1]; x1<- y[2]; x2<- y[3]; 
 	up1 <- V1*R/(K1 + R); 

### 2022-01-27 code/Figure3.R
fix xpd

@@ -105,7 +105,7 @@ legend("topright",
                 ,expression(IGR)), 
        col = c("black","black","black","black","blue","red","orange"),
        lty = c(1,2,4,3,1,1,1), bty="n", cex=1.8, inset=c(0,-0.05),
-       y.intersp = 1.1, x.intersp = 0.1, seg.len=0.8, lwd=1.5)
+       y.intersp = 1.1, x.intersp = 0.1, seg.len=0.8, lwd=1.5, xpd=NA)
 #2-5
 for (d in 1:nd){
   plot.new()
@@ -116,7 +116,7 @@ res <- vector(mode='list',length=1)
 n <- 1
 for (m in 1:nm){
   for (d in 1:nd){
-    res <- append(res,dePlot1(noise_loc, mudif[m], delta[d], qij=TRUE, xaxt="n", cex.axis=1.2, xpd=T))
+    res <- append(res,dePlot1(noise_loc, mudif[m], delta[d], qij=TRUE, xaxt="n", cex.axis=1.2))
     axis(1, labels=ifelse(n>12, yes=TRUE, no=FALSE), tick=TRUE, cex.axis=1.3)
     mtext(paste0("(", letters[n],")"), side=3, line=-1.7, at=0, cex=1.3, adj=0)
     

### 2022-01-27 code/diatom/Figure5.R
fixed line color

@@ -18,43 +18,47 @@ vTbr <- readRDS(paste(dat_loc,'varyTbar.RDS',sep=''))
 
 col <- c(rep("black",4), "red", "blue", "orange")
 line <- c(1:4, 1,1,1)
+lwd <- c(rep(2,6),3)
 
 varylist <- list(vAmp, vPer, vTbr)
 range <- sapply(varylist, function(X){range(X[,1:7])})
 ymin <- min(range)
 ymax <- max(range)
 orig <- c(a=6, P=60, Tbar=18)
+xlab <- c("a", "P", expression(theta[0]))
 
 #default is c(5, 4, 4, 2) + 0.1
 pdf(fig5, height=5, width=15)
-par(mfrow=c(1,3), oma=c(0,3,0,0), mar=c(5,1,2,1) )
+par(mfrow=c(1,3), oma=c(0,4,0,0), mar=c(5,1,2,1) )
 for (i in 1:3){
-	vdat <- varylist[[i]]
-	
-	#empty box
-	plot(0, yaxt='n', xlim=range(vdat[,10]), ylim=c(ymin, ymax), col='white', xlab=colnames(varylist[[i]])[10], ylab='', cex.lab=1.3, cex.axis=1.3)
-	
-	#axis
+  vdat <- varylist[[i]]
+  
+  #empty box
+  plot(0, yaxt='n', xlim=range(vdat[,10]), ylim=c(ymin, ymax), col='white', xlab="", ylab='', cex.lab=1.8, cex.axis=1.8)
+  title(xlab=xlab[i], cex.lab=2.3, line=3.5)
+  #axis
+  if (i==1){
+    axis(2, cex.axis=1.8)
+  }
+  
+  #loop for lines
+  for (j in 1:7){
+    lines(vdat[,10], vdat[,j], col=col[j], lty=line[j], lwd=lwd[j])	
+  }
+  abline(h=0, col="lightgrey") #zero
+  abline(v=orig[i], col='magenta', lty=3, lwd=2)
+  
+  #label
+  mtext(paste0("(", letters[i],")"), 3, -2.5, adj=0.985, cex=1.8)
+  
+  #legend
 	if (i==1){
-		axis(2)
-	}
-	
-	#loop for lines
-	for (j in 1:7){
-		lines(vdat[,10], vdat[,j], col=col[j], lty=line[j])	
-	}
-	abline(h=0, col="lightgrey") #zero
-	abline(v=orig[i], col='magenta', lty=3)
-	
-	#label
[…]

### 2022-02-10 paper/SupMat.Rnw
finish sm section 'Details of diatom model setup'

@@
 cell quota, reproduction rate, and half saturation, respectively, were used to 
-paramaterize the competing diatom model outline in the paper. 
+paramaterize the competing diatom model outlined in the paper. 
 Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
@@
 indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
-\cite{Ellne_2016, Ellner_2019} chose to slightly modify some values and functions 
-(to better align model with competition experiments), 
-%why did they make K_i's constant and do we need to adjust so that it
-%is constant over the range of temperatures considered 
-indicated in Figure \ref{fig:VKQ} by closed circles and solid lines. 
-These modifications do not result in meaningful differences when evaluating 
-coexistence at the parameter values used by \cite{Descamps_2005}.
-We chose to implement the modified functiones in our methods to quantify and 
-decompose coexistence.
+\cite{Ellne_2016, Ellner_2019} used slightly modified values and functions,
+indicated in \ref{fig:VQK} by filled circles and solid lines. 
+Our investigation follows that of \cite{Ellne_2016, Ellner_2019} and 
+uses the modified functions of  $Q_i(t)$, $V_i(t)$, and $K_i(t)$ to analyze coexistence.
 
@@
 Our decomposition is defined by $E(t)=\theta(t)$, 
-$C_i(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_i)=\frac{V_i(E)}{C_i} -D$. 
+$C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_{i \bs i})=\frac{V_i(E)}{C_{i \bs i}} -D$. 
 We note that our definition and selection of our "features" of interest are
@@
 coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$. 
-Like \cite{Ellner_2016}, we set $C_i(t)=(K_i(t)+S(t))/S(t)$ to reflect 
+Like \cite{Ellner_2016}, we set $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ to reflect 
 potential versus actual growth of the diatom populations. 
@@
 %3) You need to outline how the quantities E_i and C_i\i were obtained from simulations. 
-To obtain $E_i$ and $C_{i \bs i}$, the model was simulated for a $T=3600$ number of timesteps, setting initial population size of invader i to be zero and initial population of the resident, $r=20$.
+To obtain $E_i$ and $C_{i \bs i}$, the competiton model was simulated for a $T=3000$ number of timesteps to obtain $S$, the extracellular silicate concentration.
+The initial population size of invaderwas set at zero and initial population of the resident was set at 20.
+Environmental temperature fluctuations were defined by the paramters $a$, $P$, and $\theta_0$ for $\theta(t)=asin(2\pi t/P)+\theta_0$.
+Dilution was set to $D=0.09$ and inflow silicate concentration was set at $S_0=35$.
+A burn in time of $\frac{2}{3}T$ was cut from the resulting $S$ and $E_i(t)=\theta(t)$.
+Competitive pressure was calculated using $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$.
 

### 2022-02-10 paper/SupMat.Rnw
figures and captions

@@
 \includegraphics[width=.66\textwidth]{../results_figs/VKQfig.pdf}
-\caption[caption]
-{caption \label{fig:VKQ}}
+\caption[Species growth rate, silica uptake, and half sturation responses to changes in temperature.]
+{Species maximum growth rate (a), half saturation constant (b), and silica uptake (c) responses to changes in temperature.
+Species 1 is \emph{F.
+crotonensis}, shown in blue, and species 2 is \emph{C.
+pseudostelligera}, shown in red.
+Open circles and dashed lines indicate estimated data and interpolations from \cite{Gonzalez_2005} and closed circles and solid lines indicate values and function used in our analysis as defined in \cite{Ellner_2016, Ellner_2019}.
+\label{fig:VKQ}}
 \end{figure}
+
+%\begin{figure}
+%\includegraphics[width=\textwidth]{../results_figs/fig5_2.pdf}
+%\caption[Analogous to Fig. \ref{MT-fig:fig5} but reverse invader-resident scenario.]
+%{Analogous to Fig. \ref{MT-fig:fig5} but \emph{C. pseudostelligera} is the invader and 
+%\emph{F. crotonensis} is the resident. See the caption of
+%that figure for details. Standard errors..\label{fig:fig5_2}}
+%\end{figure}
+
+%\begin{figure}
+%\includegraphics[width=.38\textwidth]{../results_figs/fig6_2.pdf}
+%\caption[Analogous to Fig. \ref{MT-fig:fig6} but reverse invader-resident scenario.]
+%{Analogous to Fig. \ref{MT-fig:fig6} but \emph{C. pseudostelligera} is the invader and 
+%\emph{F. crotonensis} is the resident. See the caption of
+%that figure for details. Standard errors..\label{fig:fig6_2}}
+%\end{figure}
 

### 2022-02-16 paper/Paper.Rnw
reviewing edits
comments for figure adjustment
fig comments
some edits to results

@@
 %notes there indicating what the content of that section should be.
+%***JASMIN: Working on filling out diatom model set up section
 
@@
 %especially the quantitative statements at the end.
+%***JASMIN: Good 
 To display results, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
@@
 Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
-show that the ATA contributions have magnitude comparable to the 
+show that the ATA contributions have magnitudes comparable to the 
 magnitudes of other terms of the extended MCT decomposition and 
 parameters subspaces exist where ATA rescue occurs. 
+Fig.\ref{fig:fig5} plots the importance of various coexistence mechanisms for the 
+persistence of \emph{F. crotonensis} and they're sum, the IGR
+(see eq. \ref{eq:full_decomp}), for different varying temperature parameters 
+($a$, $P$, and $\theta_0$) with two of the three parameters set constant in each plot.
+Increasing amplitude and mean temperature showed increasing trends in the IGR and 
+decomposed coexistence mechanisms, while increasing periodicity showed less change. 
 Within the investigated parameters of Fig.\ref{fig:fig5}, ATA contribution, $\Delta_i^{[EC]}$, 
@@
 the contribution of varying environment, $\Delta_i^{E}$, 
-including for the original coexistence parameters specified in Gonzalez Decamps.  
-Fig.\ref{fig:fig5} plots the various contributions of coexistence (see eq.
-\ref{eq:full_decomp}) for different varying parameters with two of the three parameters set constant in each plot.
-Increasing amplitude and mean temperature showed increasing trends in the IGR and decomposed quantities, while increasing periods showed less change.
+including for the temperature parameters which \emph{F. crotonensis} 
+and \emph{C. pseudostelligera} coexisted \emph{in vitro} ? \cite{Descamps_2005}.  
+ATA rescue is revealed in Fig.\ref{fig:fig5} within parameter subspaces where the absence of 
+ATA contribution would lead to exclusion of \emph{F. crotonensis}.
+IGR decomposition of \emph{C. psuedostelligera} across the same subsets of
+temperature parameters revealed that its IGR is always positive and the ATA contribution is 
+not meaningful (see SI section). %Mean EC shows a rescue effect against variance in E
 
@@
 %have to worry about nonsense like this.
+%***JASMIN: changed labels of figures. Done.
 
@@
 %"A". Make the panel labels bigger, as well as the axis numbers and labels.
+
 %***DAN: If I interpret correctly that the blue lines on Fig. 5 are the ATA contribution (needs to be 
@@
 %in Fig. 4. Those are a couple nice wide zones of ATA rescue! Let me know if you want to talk about this.
+%***JASMIN: Fixed the line colors. Will block in the rescue areas. 
 \begin{figure}

### 2022-02-16 paper/Paper.Rnw
some edits to results

@@
 \ref{fig:fig5}, we see in Fig.
-\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence.
+\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence and the effect of periodicity of temperature fluctuation is small.
 Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{IGR}$ changes from negative to positive.
-The effect of varying periodicity of environmental fluctuations is...
 

### 2022-02-16 paper/Paper.Rnw
some edits to results

@@
 ATA contribution would lead to exclusion of \emph{F. crotonensis}.
-IGR decomposition of \emph{C. psuedostelligera} across the same subsets of
-temperature parameters revealed that its IGR is always positive and the ATA contribution is 
-not meaningful (see SI section). %Mean EC shows a rescue effect against variance in E
 
@@
 Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{IGR}$ changes from negative to positive.
+
+IGR decomposition of \emph{C. psuedostelligera} across the same subsets of
+temperature parameters revealed that its IGR is always positive; though, the ATA contribution is not meaningful to its persistence (see SI section).
+%Mean EC shows a rescue effect against variance in E
+Thus, the coexistence of the two diatom species depends only on the persistance of 
+\emph{F. crotonensis} and its ability to invade when rare. Additionally, because ATAs 
+meaningfully contribute to the IGR of \emph{F. crotonensis}, ATAs then also contribute 
+meaningfully to the coexistence of \emph{F. crotonensis} and \emph{C. psuedostelligera}.
 

### 2022-02-16 paper/Paper.Rnw
some edits to discussion

@@
 The results of our application to a simple lottery model and 
-to a competitive diatom model show that ATA contributions to storage effects
+to a competitive diatom model show that ATA contributions to storage effects and to coexistence
 are non-negligible and meaningfully contribute to the coexistence of species and thus the maintenance of biodiversity. 
@@
 be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
-Futhermore, though storage effects are typically described as the contribution
-of covariance between drivers to coexistence, commonly referred to as \emph{EC covariance},
+Though storage effects are typically described as the contribution
+of covariance between drivers to coexistence, classically \emph{EC covariance},
 \citep{Chesson_2012, Ellner_2016, Ellner_2019}, we suggest that storage effects, instead, 
@@
 
-The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
+%The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
 
@@
 \underline{ATA specialization}
-Not only is the contribution of ATA non-neglible but in some cases the presence of absence of ATA 
+Not only is the contribution of ATA non-neglible but in some cases the presence or absence of ATAs 
 can mediate the result of competition between species, i.e., coexistence or extirpation. 
@@
 coexistence – the IGR would be negative without it – deeming the scenario a case of 
-ATA rescue. For other parameter values ATA exclusion occurs where ATAs are the 
-driving mechanism of the failure of coexistence and the lack of ATAs would result in 
+ATA rescue. For other parameter values, ATA exclusion occurs where ATAs are the 
+driving mechanism of the failure to coexist (exclusion) and the lack of ATAs would result in 
 a positive IGR. Parameter
@@
 can and should be explored further. 
-Distinct increases in the magnitude of ATA contribution with increasing
-environmental fluctuations, as shown in our results and
-figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), suggests that ATAs and their effects
+The coupling of the magnitude of ATA contribution and environmental variability, 
+as shown in our results and figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), 
+suggests that ATAs and their effects
 may become more important to consider for coexistence and possibly other 
@@
 Possible changes in tail association
-patterns of environmental variables in nature can
+patterns between environmental variables in nature can
 influence the coexistence of species through the storage effect and not other mechanisms, but
@@
 \underline{E and vital rates}
+We note that \emph{E}, in our investigation and in other investigations within the MCT 
+framework, is not directly a measure of the environment, but is a measure of response to the 
+environment (vital rates). Environment and environmental response may be related linearly
+(like lottery model?); though probably real environmental variables are 
+actually related to vital rates in a non-linear and often threshold-like way. 
+Threshold relationships can produce tail association, as we showed elsewhere. 
+Changes in the means of true environmental variables may change 
+where those distributions of environmental variables are compared to thresholds, 
+which may change the nature of tail associations between the effects of 
+those environmental variables.
 %Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
 However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
-Probably real environmental variables are actually related to vital rates in a non-linear and probably often threshold-like way.
-Threshold relationships can produce tail association, as we showed elsewhere.
-Changes in the means of true environmental variables may change where those distributions of environmental variables are compared to thresholds, which may change the nature of tail associations between the effects of those environmental variables.
 
@@
 %study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
-
-From proposal:
-
 Classical analytic derivations of MCT using Taylor expansion approximations completely miss out
@@
 In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
-the contributions of correlation per se. In newer, simulation based MCT, storage effect is 
-the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, obscuring the effects of ATAs.
+the contributions of correlation \emph{per se}. 
+In newer, simulation based MCT, storage effect is 
+the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, 
+obscuring the effects of ATAs.
 

### 2022-02-18 code/MASTER.R
fixed typo

@@ -19,7 +19,7 @@ source("./diatom/Figure5.R")
 #FIGURE 6
 source("./diatom/Figure6.R")
 #SI versions of 5&6
-source(".diatom/SuppFigs5and6.R")
+source("./diatom/SuppFigs5and6.R")
 #VKQ figure (SI)
 source("./diatom/VKQplots.R")
 

### 2022-02-18 paper/SupMat.Rnw

@@
 indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
-\cite{Ellne_2016, Ellner_2019} used slightly modified values and functions,
+\cite{Ellner_2016, Ellner_2019} used slightly modified values and functions,
 indicated in \ref{fig:VQK} by filled circles and solid lines. 
-Our investigation follows that of \cite{Ellne_2016, Ellner_2019} and 
+Our investigation follows that of \cite{Ellner_2016, Ellner_2019} and 
 uses the modified functions of  $Q_i(t)$, $V_i(t)$, and $K_i(t)$ to analyze coexistence.
@@
 
+\section{\emph{Cyclotella psedosteligera} as the invader}\label{sect:Invader2}
+We generated plots analougous to Fig.
+\ref{MT-fig:fig5} and Fig.
+\ref{MT-fig:fig6} for the reverse invader-resident scenario where \emph{C.
+pseudosteligera} is the invader and \emph{F.
+crotonensis} is the resisent to compute the IGR and decompose coexistence mechanisms of \emph{C.
+pseudosteligera}.
+Fig.
+\ref{fig:fig5_2} is analogous to Fig.
+\ref{MT-fig:fig5} and shows changes in the IGR and decomposed mechanisms in response to change in one temperature parameter while the other two are held constant.
+Within the parameters of our simulations, the IGR of \emph{C.
+pseudosteligera}
+
 \clearpage
@@
 pseudostelligera}, shown in red.
-Open circles and dashed lines indicate estimated data and interpolations from \cite{Gonzalez_2005} and closed circles and solid lines indicate values and function used in our analysis as defined in \cite{Ellner_2016, Ellner_2019}.
+Open circles and dashed lines indicate estimated data and interpolations from \cite{Descamps_2005} and closed circles and solid lines indicate values and function used in our analysis as defined in \cite{Ellner_2016, Ellner_2019}.
 \label{fig:VKQ}}

### 2022-03-01 paper/Paper.Rnw
edits to the Intro to reduce length

@@
 
-%Currently 148 of an allowed 150 words for Ecology Letters
+%Currently 150 of an allowed 150 words for Ecology Letters
 \begin{abstract}
@@
 interest in ecology. Modern coexistence
-theory (MCT) identifies mechanisms of coexistence,
+theory (MCT) identifies coexistence mechanisms,
 one of which, storage effects, hinges on relationships between fluctuations
@@
 are typically measured using covariance, which does not account for 
-the possibility that environmental and competitive influences may be more related to each other
+the possibility that environment and competition may be more related to each other
 when they are strong than when weak, or vice versa. 
@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT to account for ATAs. Storage
-effects are decomposed to show the influence of ATAs. Simulations of a simple model
+We extend the MCT, decomposing storage
+effects to show the influence of ATAs. Analysis of a simple model
 and an empirical example using diatoms
@@
 difference between species coexistence and competitive exclusion.
+ATA influences are an important new mechanism of coexistence.
 \end{abstract}
@@
 that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
-the other and drive it to extinction \citep{Hardin_1960}.
+the other and drive it extinct \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
 framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
-Coexistence occurs through stabilizing factors and niche 
-differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
 MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
 in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
-applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
+applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Descamps_2005, Caceres_1997, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 and related ideas had previously been seldom 
-considered in ecology (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
-our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
-many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
-synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
+our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
+many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
+and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
+have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.
 
 %Intro to storage effects and why they might be influenced by ATAs
-Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
-have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
 It is straightforward to see, intuitively, 
 why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
 Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
@@
 example buffering mechanism \citep{Chesson_2012}. 
-When present, these three factors combine so that EC covariance provides a relative benefit to whatever species are currently rare.
+When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
 Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
@@
 occur between environment and competition variables, can be an aspect of EC covariation.
+
+%***DAN: Go though and make sure all citation callouts list papers in date order, olderst to newest,
+%except when multiple pubs by the same first author are listed they are all together.
 
@@
 It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
-heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017, Lyon2019, Keelings2020}.  
+heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
 If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
@@
 Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
-left-tail association between these variables. As we will show in this study, changes in ATAs can then
-result in different competitive outcomes between species. Although we focus here on temporal coexistence
-mechanisms, ATAs between environmental variables measured in different places may also be altered 
-by climate change, if more intensive climatic events are also more spatially extensive.
-
-%make the above abstract ideas more concrete
-To make the above ideas more concrete, we briefly mention examples from prior work to 
-which the ideas may apply. 
-\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
-are likely influenced differently by phenologically distict environmental variables: an Australian 
-grass-forb community, and an Arizona community of desert annuals. 
-The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
-examples, distinct species have different periods of germination, competition is influenced by resource 
-consumption, and population growth is buffered by overlapping generations or via seed banks.
-Temporal differences between species in germination rates suggest the possibility that 
-population growth of distinct species may be driven by rainfall signals in distinct parts
-of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-alter ATAs between environmental variables which are relevant for the coexistence
-of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
-
-%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
-Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-not been studied, the importance for coexistence of changes 
-in means and variability of environmental variables has been studied in many systems. We briefly 
-review some of that literature.
-\cite{diPaola_2012} concluded that coexistence of drought sensitive,
-flood resistant species with flood sensitive, drought resistant species in 
-Mediteranean coastal swamp forests of central Italy would be jeopardized
-by projected changes in mean rainfall. 
-\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
-higher altitudes of herbivore populations than of plant populations, newly arrived
-herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
-in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
-alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
-all species they considered, but that changes in competition would more than offset these direct benefits. 
-Numerous studies take an ``event-focused'' 
-perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
-form of extreme weather events, can alter competitive interactions. For instance, 
-\cite{White_2001} showed that single extreme heat events can alter competition 
-between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
-nutrient availability for plants, mediated by soil moisture content, can alter 
-the potential for invasive species to enter an ecosystem and establish, a process that involves
-the invader overcoming competition from resident species. 
-Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
-communities, via storage effects \citep{Adler_2006,  Angert_2009}.
+left-tail association between these variables. 
+As we will show in this study, changes in ATAs can then
+result in different competitive outcomes between species.
+\cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
+community of desert annuals that seem to have all the ingredients necessary to make effects of the
+type we describe here a possibility.
+To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+not been studied, but the importance for coexistence of changes 
+in means and variability of environmental variables has been studied in many systems 
+[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Parepa_2013, Descombes_2020, Usinowicz_2021}]. 
 Changes in variability and in extreme events are distinct from the changes in ATAs we 
@@
 of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
+
+%make the above abstract ideas more concrete - 
+%***DAN: This para commented out for length, added a sentence about the systems to the above paragraph to take the place of this text
+%To make the above ideas more concrete, we briefly mention examples from prior work to 
+%which the ideas may apply. 
+%\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
+%are likely influenced differently by phenologically distict environmental variables: an Australian 
+%grass-forb community, and an Arizona community of desert annuals. 
+%The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
+%examples, distinct species have different periods of germination, competition is influenced by resource 
+%consumption, and population growth is buffered by overlapping generations or via seed banks.
+%Temporal differences between species in germination rates suggest the possibility that 
+%population growth of distinct species may be driven by rainfall signals in distinct parts
+%of the growing season. Thus all the ingredients seem to be present by which climatic changes may
+%alter ATAs between environmental variables which are relevant for the coexistence
+%of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
+
+%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence  
+%***DAN: this para commented out for length, added a couple of the sentences to an above para to substitute
+%Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+%not been studied, the importance for coexistence of changes 
+%in means and variability of environmental variables has been studied in many systems. We briefly 
+%review some of that literature.
+%\cite{diPaola_2012} concluded that coexistence of drought sensitive,
+%flood resistant species with flood sensitive, drought resistant species in 
+%Mediteranean coastal swamp forests of central Italy would be jeopardized
+%by projected changes in mean rainfall. 
+%\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
+%higher altitudes of herbivore populations than of plant populations, newly arrived
+%herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
+%in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
+%alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
+%all species they considered, but that changes in competition would more than offset these direct benefits. 
+%Numerous studies take an ``event-focused'' 
+%perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
+%form of extreme weather events, can alter competitive interactions. For instance, 
+%\cite{White_2001} showed that single extreme heat events can alter competition 
+%between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
+%nutrient availability for plants, mediated by soil moisture content, can alter 
+%the potential for invasive species to enter an ecosystem and establish, a process that involves
+%the invader overcoming competition from resident species. 
+%Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
+%communities, via storage effects \citep{Adler_2006,  Angert_2009}.
+%Changes in variability and in extreme events are distinct from the changes in ATAs we 
+%consider here, because ATAs pertain to \emph{relationships} between environmental and other
+%variables in the extremes, which are distinct from and can be altered independently
+%of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
@@
 biodiversity, by means of the following specific goals. 
-G1) To augment MCT to show formally how ATA plays a role in coexistence. 
+G1) To augment MCT to show formally how ATAs play a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
-We address how ATA contributes to these components. 
+We address how ATAs contribute to these components. 
 We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
@@
 ATAs of environmental variables and thereby influence competition and species coexistence.
+%***DAN: Make sure we end up having such material in the Discussion
 Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
-and a means of understanding its theoretical importance and its importance for real ecosystems.
+and a means of understanding its theoretical and empirical importance.
 
 \section{Theory}\label{section:theory}
+
+%***DAN: consider sending the lottery model components of this section to sup mat, for length reasons.
 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simplicity, 
+competitive pressure on species $i$, $C_i(t)$. We assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
@@
 presentation of new ideas.
+
+%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
+%throughout the paper. Has to be done in the main text and sup mat and the figures themselves (IGR is often labeled on 
+%the figure panels themselves).
 
@@
 \includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
-\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
+\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered 
+from low densities less often and 
 spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
-species coexistence. Panel (a) shows species 1, from the lottery 
+species coexistence. Panel (a) shows species 1 from the lottery 
 model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
@@
 Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
-``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
+``noticeable coexistence'' of the species, defined to be periods when each species comprises at least $5\%$ of 
 total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
@@
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case.
-Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$.
-See Introduction and Fig.
-\ref{fig:noise} caption for model details.
+on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$, 
+$\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. 
+See Introduction and Fig. \ref{fig:noise} caption for model details. 
 \label{fig:pedagog}}

### 2022-03-01 paper/Paper.Rnw
about to remove the lottery model theory from Theory and put it in a new sup mat section

@@
 
-%***DAN: consider sending the lottery model components of this section to sup mat, for length reasons.
+%***DAN: consider sending the lottery model components of this section to sup mat, to reduce length and complexity.
+%Before doing that, make sure you understand what the consequences will be.
 
@@
 coexistence is impossible and that the case is relatively trivial (SI 
-section \ref{SI-sect:delta1}). So that case is not considered further.  
+section \ref{SI-sect:delta1}), so that case is not considered further.  
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).

### 2022-03-01 paper/Paper.Rnw
moved the lottery model material from Theory to a new sup mat section, for length reasons

@@
 
-%***DAN: consider sending the lottery model components of this section to sup mat, to reduce length and complexity.
-%Before doing that, make sure you understand what the consequences will be.
-
 We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
-plays a role in coexistence. For pedagogical clarity, we develop each step of our new theory for a general 
-two-species model, and for the lottery model (Introduction).
+plays a role in coexistence. We develop each step of our new theory for a general 
+two-species model, below, and specifically for the lottery model of the Introduction, in SI section \ref{SI-SIsect:TheoryForLottery}.
 MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
@@
 that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-So we begin by defining storage effects for a general two species model and 
-for the lottery model. Our mathematical notation parallels
+So we begin by defining storage effects. Our mathematical notation parallels
 that of \cite{Ellner_2016,Ellner_2019}. 
-
-For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
+It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
 of species $i$ can be written as an increasing function of
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. We assume, for simplicity, 
+competitive pressure on species $i$, $C_i(t)$. We also assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
+$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
 presentation of new ideas.
@@
 %throughout the paper. Has to be done in the main text and sup mat and the figures themselves (IGR is often labeled on 
-%the figure panels themselves).
-
-For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
-the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
-divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
-Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.
-
-For a general model, the storage effect for species $i$ is the contribution of covariation between $E$ and 
+%the figure panels themselves). Need to also search for "invader" since sometimes we use that word. Also think of any 
+%other related terminology that needs to be altered as part of the overall change.
+
+The storage effect for species $i$ is the contribution of covariation between $E$ and 
 $C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
@@
 influences of environment and competition on the invader or resident have been rendered independent.  
-Following \cite{Ellner_2016,Ellner_2019}, a subscript 
-$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
+A subscript $i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
 a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
-The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
 The \emph{scaling factor}
 $q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
-\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$. 
+\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two alternative definitions of $q_{ij}$
+that we use. 
 Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
@@
 Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
-Though some readers may find it slightly confusing to make simultaneous use of the variable 
-$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard and very common, so it would 
+Though some readers may find it slightly confusing to simultaneously use the variable 
+$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would 
 be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
 all operators, $\E(\cdot)$ is rendered upright.
-
-For the lottery model, when $i$ is an invader, the competition experienced by either 
-species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
-$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
-$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
-$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
-Likewise, 
-$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
-though we already knew this equals $0$. 
-An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
-by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
-Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
-$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
-$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
-$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
-We show in SI section \ref{SI-SIsect:scaling_factors}
-that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
-Since we assumed log-normally distributed $B_i$, 
-$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
-We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SI-SIsect:scaling_factors}.
-Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
-to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
-and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into (\ref{eq:storage_effects_gen}).
 
@@
 The new symbol should correspondingly be pronounced ``partial sharp''.
-The brackets in the superscripts of $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
+The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
 (\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effects, which are themselves a term within the decomposition provided by MCT \citep{Ellner_2019} -- see 
+of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
 below for additional details.
@@
 $(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
+For those familiar with copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 $C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created in an analogous way.
-
-For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
-$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
-$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
-Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
-Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
-$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
-can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
-is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
-and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
-marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
-$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
-component $\olrn_{i \bs i}$ of (\ref{eq:storage_effects_further1}) for the lottery model using the simulation 
-approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
+as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
 

### 2022-03-01 paper/SupMat.Rnw
about to remove the lottery model theory from Theory and put it in a new sup mat section
moved the lottery model material from Theory to a new sup mat section, for length reasons

@@
 
-%probably want table of contents, figure and table lists
+\section{Theory for the lottery model}\label{SIsect:TheoryForLottery}
+
+For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
+the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
+divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
+does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
+Species growth rates for the lottery model can then be written 
+$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$, which is an increasing
+function of $E_i$ and decreasing function of $C_i$.
+
+When $i$ is an invader, the competition experienced by either 
+species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
+$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
+$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
+$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
+Likewise, 
+$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
+though we already knew this equals $0$. 
+An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
+Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
+$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
+$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
+$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
+We show in SI section \ref{SIsect:scaling_factors}
+that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
+Since we assumed log-normally distributed $B_i$, 
+$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
+We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SIsect:scaling_factors}.
+Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
+and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
+
+For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
+$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
+$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
+Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
+Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
+$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
+can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
+is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
+and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
+marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
+$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
+component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
+approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
+so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
 

### 2022-03-02 paper/Paper.Rnw
adding to discussion and some edits

--- 2022-03-01 paper/Paper.Rnw
@@
 
-%Currently 150 of an allowed 150 words for Ecology Letters
+%Currently 148 of an allowed 150 words for Ecology Letters
 \begin{abstract}
@@
 interest in ecology. Modern coexistence
-theory (MCT) identifies coexistence mechanisms,
+theory (MCT) identifies mechanisms of coexistence,
 one of which, storage effects, hinges on relationships between fluctuations
@@
 are typically measured using covariance, which does not account for 
-the possibility that environment and competition may be more related to each other
+the possibility that environmental and competitive influences may be more related to each other
 when they are strong than when weak, or vice versa. 
@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT, decomposing storage
-effects to show the influence of ATAs. Analysis of a simple model
+We extend the MCT to account for ATAs. Storage
+effects are decomposed to show the influence of ATAs. Simulations of a simple model
 and an empirical example using diatoms
@@
 difference between species coexistence and competitive exclusion. 
-ATA influences are an important new mechanism of coexistence.
 \end{abstract}
@@
 that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
-the other and drive it extinct \citep{Hardin_1960}.
+the other and drive it to extinction \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
 framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
+Coexistence occurs through stabilizing factors and niche 
+differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
 MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
 in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
-applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Descamps_2005, Caceres_1997, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
+applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 and related ideas had previously been seldom 
-considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
-our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
-many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
-and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
-Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
+considered in ecology (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
+our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
+many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
+synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+
+%Intro to storage effects and why they might be influenced by ATAs
+Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
 have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
-
-%Intro to storage effects and why they might be influenced by ATAs
 It is straightforward to see, intuitively, 
 why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
 Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
@@
 occur between environment and competition variables, can be an aspect of EC covariation.
-
-%***DAN: Go though and make sure all citation callouts list papers in date order, olderst to newest,
-%except when multiple pubs by the same first author are listed they are all together.
 
@@
 It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
-heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
+heat waves, floods, hurricanes and fires \citep{Jentsch_2007, Meehl_2004, Ummenhofer_2017, Lyon2019, Keelings2020}.  
 If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
@@
 Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
-left-tail association between these variables. 
-As we will show in this study, changes in ATAs can then
-result in different competitive outcomes between species.
-\cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
-community of desert annuals that seem to have all the ingredients necessary to make effects of the
-type we describe here a possibility.
-To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-not been studied, but the importance for coexistence of changes 
-in means and variability of environmental variables has been studied in many systems 
-[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Parepa_2013, Descombes_2020, Usinowicz_2021}]. 
+left-tail association between these variables. As we will show in this study, changes in ATAs can then
+result in different competitive outcomes between species. Although we focus here on temporal coexistence
+mechanisms, ATAs between environmental variables measured in different places may also be altered 
+by climate change, if more intensive climatic events are also more spatially extensive.
+
+%make the above abstract ideas more concrete
+To make the above ideas more concrete, we briefly mention examples from prior work to 
+which the ideas may apply. 
+\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
+are likely influenced differently by phenologically distict environmental variables: an Australian 
+grass-forb community, and an Arizona community of desert annuals. 
+The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
+examples, distinct species have different periods of germination, competition is influenced by resource 
+consumption, and population growth is buffered by overlapping generations or via seed banks.
+Temporal differences between species in germination rates suggest the possibility that 
+population growth of distinct species may be driven by rainfall signals in distinct parts
+of the growing season. Thus all the ingredients seem to be present by which climatic changes may
+alter ATAs between environmental variables which are relevant for the coexistence
+of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
+
+%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
+Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+not been studied, the importance for coexistence of changes 
+in means and variability of environmental variables has been studied in many systems. We briefly 
+review some of that literature.
+\cite{diPaola_2012} concluded that coexistence of drought sensitive,
+flood resistant species with flood sensitive, drought resistant species in 
+Mediteranean coastal swamp forests of central Italy would be jeopardized
+by projected changes in mean rainfall. 
+\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
+higher altitudes of herbivore populations than of plant populations, newly arrived
+herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
+in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
+alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
+all species they considered, but that changes in competition would more than offset these direct benefits. 
+Numerous studies take an ``event-focused'' 
+perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
+form of extreme weather events, can alter competitive interactions. For instance, 
+\cite{White_2001} showed that single extreme heat events can alter competition 
+between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
+nutrient availability for plants, mediated by soil moisture content, can alter 
+the potential for invasive species to enter an ecosystem and establish, a process that involves
+the invader overcoming competition from resident species. 
+Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
+communities, via storage effects \citep{Adler_2006,  Angert_2009}.
 Changes in variability and in extreme events are distinct from the changes in ATAs we 
@@
 of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
-
-%make the above abstract ideas more concrete - 
-%***DAN: This para commented out for length, added a sentence about the systems to the above paragraph to take the place of this text
-%To make the above ideas more concrete, we briefly mention examples from prior work to 
-%which the ideas may apply. 
-%\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
-%are likely influenced differently by phenologically distict environmental variables: an Australian 
-%grass-forb community, and an Arizona community of desert annuals. 
-%The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
-%examples, distinct species have different periods of germination, competition is influenced by resource 
-%consumption, and population growth is buffered by overlapping generations or via seed banks.
-%Temporal differences between species in germination rates suggest the possibility that 
-%population growth of distinct species may be driven by rainfall signals in distinct parts
-%of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-%alter ATAs between environmental variables which are relevant for the coexistence
-%of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
-
-%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence  
-%***DAN: this para commented out for length, added a couple of the sentences to an above para to substitute
-%Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-%not been studied, the importance for coexistence of changes 
-%in means and variability of environmental variables has been studied in many systems. We briefly 
-%review some of that literature.
-%\cite{diPaola_2012} concluded that coexistence of drought sensitive,
-%flood resistant species with flood sensitive, drought resistant species in 
-%Mediteranean coastal swamp forests of central Italy would be jeopardized
-%by projected changes in mean rainfall. 
-%\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
-%higher altitudes of herbivore populations than of plant populations, newly arrived
-%herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
-%in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
-%alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
-%all species they considered, but that changes in competition would more than offset these direct benefits. 
-%Numerous studies take an ``event-focused'' 
-%perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
-%form of extreme weather events, can alter competitive interactions. For instance, 
-%\cite{White_2001} showed that single extreme heat events can alter competition 
-%between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
-%nutrient availability for plants, mediated by soil moisture content, can alter 
-%the potential for invasive species to enter an ecosystem and establish, a process that involves
-%the invader overcoming competition from resident species. 
-%Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
-%communities, via storage effects \citep{Adler_2006,  Angert_2009}.
-%Changes in variability and in extreme events are distinct from the changes in ATAs we 
-%consider here, because ATAs pertain to \emph{relationships} between environmental and other
-%variables in the extremes, which are distinct from and can be altered independently
-%of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
@@
 biodiversity, by means of the following specific goals. 
-G1) To augment MCT to show formally how ATAs play a role in coexistence. 
+G1) To augment MCT to show formally how ATA plays a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
-We address how ATAs contribute to these components. 
+We address how ATA contributes to these components. 
 We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
@@
 ATAs of environmental variables and thereby influence competition and species coexistence.
-%***DAN: Make sure we end up having such material in the Discussion
 Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
-and a means of understanding its theoretical and empirical importance.
+and a means of understanding its theoretical importance and its importance for real ecosystems.
 
 \section{Theory}\label{section:theory}
-
-%***DAN: consider sending the lottery model components of this section to sup mat, for length reasons.
 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. We assume, for simplicity, 
+competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
@@
 presentation of new ideas.
-
-%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
-%throughout the paper. Has to be done in the main text and sup mat and the figures themselves (IGR is often labeled on 
-%the figure panels themselves).
 
@@
 
-\underline{summary - include 1-2 sentences about misnomer}
 We produced and applied a mathematical extension to MCT that, for the first time, 
@@
 
-Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
+%Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
 Previous studies report storage effect as a contribution to coexistence of EC covariance; yet compute it as the contribution of the EC relationship (covariance and beyond; symmetric and asymmetric).
 
-\underline{ATA specialization}
+%\underline{ATA specialization}
 Not only is the contribution of ATA non-neglible but in some cases the presence or absence of ATAs 
@@
 subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists} - species which only persist in habitat patches where ATAs contribute to its coexistence.
+One can imagine a species whose coexistence with others is conditional on the asymmetric relationship between drivers of coexistence and whose populations may become excluded when ATAs erode. 
 %I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
@@
 
-\underline{future work: study ATAs in environmental variables and how theyre changing - tie in with specialist}
+%\underline{future work: study ATAs in environmental variables and how theyre changing - tie in with specialist}
 
@@
 possible \emph{relationships} in the extremes; however, 
-simpler changes in extreme events themselves such as
+simpler changes in [the marginal/ univariate distributions of] extreme events themselves such as
 intensity and frequency and their influence on coexistence also merit future work. 
@@
 
-\underline{E and vital rates}
+%\underline{E and vital rates}
 We note that \emph{E}, in our investigation and in other investigations within the MCT 
 framework, is not directly a measure of the environment, but is a measure of response to the 
-environment (vital rates). Environment and environmental response may be related linearly
-(like lottery model?); though probably real environmental variables are 
-actually related to vital rates in a non-linear and often threshold-like way. 
-Threshold relationships can produce tail association, as we showed elsewhere. 
+environment, measured by a population's vital rates. Environment and environmental response may be related linearly
+(like lottery model?); though real environmental variables are probable to 
+actually be related to vital rates in a non-linear and often threshold-like way. 
+Threshold relationships can produce tail association, as we showed elsewhere (explain, ref). 
 Changes in the means of true environmental variables may change 
@@
 which may change the nature of tail associations between the effects of 
-those environmental variables.
+those environmental variables. Thus, not only do the possible asymmteric association between 
+environmental variables merit future study, but the relationships between populations'
+vital rates should also be investigated for species coexistence and the maintenacne of biodiversity.
 %Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
@@
 
-\underline{the new effects are not "weak noise" effects}
-if sigma really small, ATA would not matter?
+%\underline{the new effects are not "weak noise" effects}
+%if sigma really small, ATA would not matter?
 %go back to emails
@@
 %study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
-Classical analytic derivations of MCT using Taylor expansion approximations completely miss out
-on any additions to coexistence beyond simple correlations. 
+We parse the storage effect, providing a tertiary decomposition to the simulation-based 
+approach of MCT \citep{Ellner_2016, Ellner_2019}, 
+a new approach which bypasses the mathematical assumptions and obstructions of 
+standard MCT. Classical analytic derivations of MCT using second order Taylor expansion and 
+small-variance approximations completely miss out
+on any additions to coexistence beyond simple correlations and does not include our new term, $\Delta^{[EC]}$ or $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
 In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
@@
 the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, 
-obscuring the effects of ATAs.
+obscuring the effects of ATAs. However, because standard MCT's assumption make it impossible to 
+detect interactions outide of correlation \emph{per se}, simulation-based MCT is required to
+detect ATA contributions, and our new extension to it cleary separates and quantifies the 
+relative contribution and importance of ATAs and pure correlation to the storage effect
+and subsequently species coexistence and the maintenance of biodiversity.
 
@@
 %just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
+
 We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
@@
 
-\underline{conclude: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
+\underline{conclude **NOT NEEDED**: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
 We provided another showcase of the importance of ATAs in ecology, 
@@
 \includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
-\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered 
-from low densities less often and 
+\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered from low densities less often and 
 spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
-species coexistence. Panel (a) shows species 1 from the lottery 
+species coexistence. Panel (a) shows species 1, from the lottery 
 model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
@@
 Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
-``noticeable coexistence'' of the species, defined to be periods when each species comprises at least $5\%$ of 
+``noticeable coexistence'' of the species, defined to be periods when both species each comprise at least $5\%$ of 
 total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
@@
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$, 
-$\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. 
-See Introduction and Fig. \ref{fig:noise} caption for model details. 
+on average, for the ATA case.
+Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$.
+See Introduction and Fig.
+\ref{fig:noise} caption for model details.
 \label{fig:pedagog}}

### 2022-03-02 paper/Paper.Rnw
fixing in text citation order

@@
 in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
-applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2010, Adler_2006}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Narawani_2013, Descamps_2005, Caceres_1997}, and corals \citep{Alvarez-Noriega_2020}. 
+applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Narawani_2013, Descamps_2005}, and corals \citep{Alvarez-Noriega_2020}. 
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 and related ideas had previously been seldom 
-considered in ecology (but see \citealp{Anderson_2019, deValpine_2014, Popovic_2019}), 
+considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
 our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
@@
 $(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with the language of copula statistics (see, e.g., \citealp{Ghosh_2020_AER} and \citealp{Genest_2007}), 
+For those familiar with the language of copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 considers ATAs for the coexistence of species and maintenance of biodiversity. 
-Implementing new methods of MCT \citep{Ellner_2019, Ellner_2016}, we provided a further tertiary 
+Implementing new methods of MCT \citep{Ellner_2016, Ellner_2019}, we provided a further tertiary 
 decomposition to MCT and decompose the storage effect into two new terms, 
@@
 possible to generate asymmetrically tail associated noise with imperfect correlations in 
-both halves of the distributions \citep{Ghosh_2020_AER,Nelsen2006}. The variables $(b_1,b_2)$ described in
+both halves of the distributions \citep{Nelsen2006,Ghosh_2020_AER}. The variables $(b_1,b_2)$ described in
 the Introduction and Methods were obtained by generating 

### 2022-03-04 paper/Paper.Rnw
adding to last paragraph

@@
 \underline{caveats and future directions}
+%Realism - we usede simple/extreme ATA, simple model and simple lab system, futurw work needed to assess 
+%how mcuh ATA are important in complex systems, field systems
+%what systems? what env vars?
 %Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
@@
 %Also need to talk about how this stuff will apply to data.
-we used simple/extreme ATA
-simple model and simple lab system
-future work needed to assess the extent of how much ATA are important for complex field systems
-what systems might be good? suggest systems and their important E vars.
-
-Our investigation is a simple one; 
-the ATA present is a perfect correlation is exactly one half of the relationship, 
-but varying degrees of ATA may produce a similar result as well.
-
-We used the most extreme version of asymmetric noise, i.e., perfect correlation changing to very low correlation at exactly the half point of the distribution, and a simple lottery model.
+As an introductory investigation of ATAs within the framework of coexistence theory, 
+we used simple and extreme examples, and future work should include more realism and complexity
+(assess more complex relationships, systems, and models). 
+In our theoretical exploration, we implemented a simple lottery model and assessed 
+the influence of extremely asymmetric associated noise, i.e., the noise was perfectly correlated
+in exactly one half of the distribution. However, more complex models such as structured demographic models
+or integral projection models should be explored and though similar results may be
+produced, varying deegres of ATA, less extremely shaped?/
+more moderately asymmetric / more realistic? relationships should be implemented.
+Moreover, because our empirical application makes use of a simple and controlled lab system, future 
+should be done to assess the extent of how much ATA are important for complex systems and field systems.
 
@@
 
-\underline{conclude **NOT NEEDED**: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
-We provided another showcase of the importance of ATAs in ecology, 
-building on previous studies.
-ATAs clearly have implications for coexistence and competition;
-though more research is needed in exploring more complex models 
-including spatiotemporal and multispecies models.
-Hope to see this tertiary decomposition of MCT be applied to empirical research.
-More investigations should be done to explore implications of ATAs 
-to other ecological theories.
+%\underline{conclude **NOT NEEDED**: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
+%We provided another showcase of the importance of ATAs in ecology, 
+%building on previous studies.
+%ATAs clearly have implications for coexistence and competition;
+%though more research is needed in exploring more complex models 
+%including spatiotemporal and multispecies models.
+%Hope to see this tertiary decomposition of MCT be applied to empirical research.
+%More investigations should be done to explore implications of ATAs 
+%to other ecological theories.
 

### 2022-03-04 paper/Paper.Rnw
adding to discussion

@@
 Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
+More species and more drivers= more things that can be related
 
@@
 should be done to assess the extent of how much ATA are important for complex systems and field systems.
+For example, the ATAs between rainfall timing on the coexistence of grasses and forbs 
+and their dominant years in California grasslands \citep{Hallett_2019} could be a fruitful 
+further investigation of biodiversity maintenance in California grasslands and the effects of extreme rainfall fluctuation.
 

### 2022-03-07 paper/Paper.Rnw
some sentences about multi-species models

@@
 
-We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
+%We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
 Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
-More species and more drivers= more things that can be related
+%More species and more drivers= more things that can be related
+Our further decomposition and quantification of ATA contribution to coexistence can and 
+should also be applied to multi-species models. 
+The inclusion of multiple species introduces a multitude of relationships between species'
+environmentally dependent parameters that could possibly be asymmetrically associated. 
+The greater the number of species the greater the possibilities of ATAs being more important 
+to their coexistence.
 

### 2022-03-07 paper/Paper.Rnw
readding deleted changes

@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT to account for ATAs. Storage
-effects are decomposed to show the influence of ATAs. Simulations of a simple model
+We extend the MCT, decomposing storage effect to show the influence of ATAs.
+Analysis of a simple model
 and an empirical example using diatoms
@@
 difference between species coexistence and competitive exclusion. 
+ATA influences are an important new mechanism of coexistence. 
 \end{abstract}
@@
 that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
-the other and drive it to extinction \citep{Hardin_1960}.
+the other and drive it extinct \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
 framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
-Coexistence occurs through stabilizing factors and niche 
-differentiation, in spite of the conclusions drawn for idealized scenarios by the competitive exclusion principle. 
 MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
@@
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Narawani_2013, Descamps_2005}, and corals \citep{Alvarez-Noriega_2020}. 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
 Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
@@
 to their coexistence.
+
+# notation in ellner one could also use partial sharp 
+%reread ellner paper 
+%make three different biv dis that are standard normal copula and then pair with a shuffles third
+%use to get rbar and subtract from rbar of XY#Z,
 

### 2022-03-07 paper/Paper.Rnw
readding changes intro.p2

@@
 considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
-our recent work has shown that, in fact, ATAs are extremely common in ecology \citep{Ghosh_2020_AER}, and can influence 
-many ecological phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
-synchrony, and the temporal stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
+many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
+synchrony, and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
+Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
+have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.
 
 %Intro to storage effects and why they might be influenced by ATAs
-Although ATAs are now known to be common and important, the implications of ATAs for competition and species coexistence 
-have, to our knowledge, not previously been considered; doing so is the overarching mission of this study. 
 It is straightforward to see, intuitively, 

### 2022-03-07 paper/Paper.Rnw
finished readding deletions in intro

@@
 why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and we explain why ATAs may be relevant to them.
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
 Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
@@
 occur between environment and competition variables, can be an aspect of EC covariation.
+
+%***DAN: Go through and make sure all citation callouts list papers in date order, oldest to newest,
+%except when multiple pubs by the same first author are listed they are all together.
 
@@
 left-tail association between these variables. As we will show in this study, changes in ATAs can then
-result in different competitive outcomes between species. Although we focus here on temporal coexistence
-mechanisms, ATAs between environmental variables measured in different places may also be altered 
-by climate change, if more intensive climatic events are also more spatially extensive.
-
-%make the above abstract ideas more concrete
-To make the above ideas more concrete, we briefly mention examples from prior work to 
-which the ideas may apply. 
-\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
-are likely influenced differently by phenologically distict environmental variables: an Australian 
-grass-forb community, and an Arizona community of desert annuals. 
-The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
-examples, distinct species have different periods of germination, competition is influenced by resource 
-consumption, and population growth is buffered by overlapping generations or via seed banks.
-Temporal differences between species in germination rates suggest the possibility that 
-population growth of distinct species may be driven by rainfall signals in distinct parts
-of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-alter ATAs between environmental variables which are relevant for the coexistence
-of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
-
-%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
-Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-not been studied, the importance for coexistence of changes 
-in means and variability of environmental variables has been studied in many systems. We briefly 
-review some of that literature.
-\cite{diPaola_2012} concluded that coexistence of drought sensitive,
-flood resistant species with flood sensitive, drought resistant species in 
-Mediteranean coastal swamp forests of central Italy would be jeopardized
-by projected changes in mean rainfall. 
-\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
-higher altitudes of herbivore populations than of plant populations, newly arrived
-herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
-in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
-alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
-all species they considered, but that changes in competition would more than offset these direct benefits. 
-Numerous studies take an ``event-focused'' 
-perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
-form of extreme weather events, can alter competitive interactions. For instance, 
-\cite{White_2001} showed that single extreme heat events can alter competition 
-between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
-nutrient availability for plants, mediated by soil moisture content, can alter 
-the potential for invasive species to enter an ecosystem and establish, a process that involves
-the invader overcoming competition from resident species. 
-Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
-communities, via storage effects \citep{Adler_2006,  Angert_2009}.
+result in different competitive outcomes between species. 
+\cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
+community of desert annuals that seem to have all the ingredients necessary to make effects of the
+type we describe here a possibility.
+To out knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+not been studied, but the importance for coexistence of changes 
+in means and variability of environmental variables has been studied in many systems
+[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
 Changes in variability and in extreme events are distinct from the changes in ATAs we 
@@
 of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
+
+%make the above abstract ideas more concrete - 
+%***DAN: This para commented out for length, added a sentence about the systems to the above paragraph to take the place of this text
+%To make the above ideas more concrete, we briefly mention examples from prior work to 
+%which the ideas may apply. 
+%\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
+%are likely influenced differently by phenologically distict environmental variables: an Australian 
+%grass-forb community, and an Arizona community of desert annuals. 
+%The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
+%examples, distinct species have different periods of germination, competition is influenced by resource 
+%consumption, and population growth is buffered by overlapping generations or via seed banks.
+%Temporal differences between species in germination rates suggest the possibility that 
+%population growth of distinct species may be driven by rainfall signals in distinct parts
+%of the growing season. Thus all the ingredients seem to be present by which climatic changes may
+%alter ATAs between environmental variables which are relevant for the coexistence
+%of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
+
+%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
+%**DAN: this para commetned out for length, added a couple of sentences to an aboce para to substitute
+%Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+%not been studied, the importance for coexistence of changes 
+%in means and variability of environmental variables has been studied in many systems. We briefly 
+%review some of that literature.
+%\cite{diPaola_2012} concluded that coexistence of drought sensitive,
+%flood resistant species with flood sensitive, drought resistant species in 
+%Mediteranean coastal swamp forests of central Italy would be jeopardized
+%by projected changes in mean rainfall. 
+%\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
+%higher altitudes of herbivore populations than of plant populations, newly arrived
+%herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
+%in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
+%alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
+%all species they considered, but that changes in competition would more than offset these direct benefits. 
+%Numerous studies take an ``event-focused'' 
+%perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
+%form of extreme weather events, can alter competitive interactions. For instance, 
+%\cite{White_2001} showed that single extreme heat events can alter competition 
+%between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
+%nutrient availability for plants, mediated by soil moisture content, can alter 
+%the potential for invasive species to enter an ecosystem and establish, a process that involves
+%the invader overcoming competition from resident species. 
+%Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
+%communities, via storage effects \citep{Adler_2006,  Angert_2009}.
 
@@
 ATAs of environmental variables and thereby influence competition and species coexistence.
+%***DAN: Make sure we end up having such material in the Discussion
 Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
-and a means of understanding its theoretical importance and its importance for real ecosystems.
+and a means of understanding its theoretical and empirical importance.
 

### 2022-03-07 paper/Paper.Rnw
re-removing lottery model from Paper.Rnw

@@
 We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
-plays a role in coexistence. For pedagogical clarity, we develop each step of our new theory for a general 
-two-species model, and for the lottery model (Introduction).
+plays a role in coexistence. We develop each step of our new theory for a general 
+two-species model, below, and specifically for the lottery model of the Introduction, 
+in SI section \ref{SI-SIsect:TheoryForLottery}.
 MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
@@
 that, for purely temporal variation, only one mechanism relates to tail associations: storage effects.
-So we begin by defining storage effects for a general two species model and 
-for the lottery model. Our mathematical notation parallels
-that of \cite{Ellner_2016,Ellner_2019}.
-
-For a general model, it is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
+
+So we begin by defining storage effects. Our mathematical notation parallels
+that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
 of species $i$ can be written as an increasing function of
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. As for the lottery model example (Introduction), we here assume, for simplicity, 
+competitive pressure on species $i$, $C_i(t)$. We also assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though this assumption can be circumvented, we make the assumption because it simplifies the
+$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
 presentation of new ideas.
 
-For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
-the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
-divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
-Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$.
-
-For a general model, the storage effect for species $i$ is the contribution of covariation between $E$ and 
+%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
+% throughout the paper. Has to be done in the main text and sup mat and the figures themselves.
+% Need also search for "invader" since sometimes we use that word. Also think of any other related terminology that needs to 
+% be altered as part of the overall change.
+
+The storage effect for species $i$ is the contribution of covariation between $E$ and 
 $C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
@@
 influences of environment and competition on the invader or resident have been rendered independent.  
-Following \cite{Ellner_2016,Ellner_2019}, a subscript 
+A subscript 
 $i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
 a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
-The notation $\Delta_i^{(EC)}$ for storage effects also follows \cite{Ellner_2019}.
 The \emph{scaling factor}
 $q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
-\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the definition of $q_{ij}$. 
+\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
+alternative definitions of $q_{ij}$ that we use. 
 Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
@@
 $(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with the language of copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
+For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case.
-Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$ and $\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$.
-See Introduction and Fig.
-\ref{fig:noise} caption for model details.
+on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$,
+$\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. 
+See Introduction and Fig. \ref{fig:noise} caption for model details. 
 \label{fig:pedagog}}

### 2022-03-08 paper/Paper.Rnw
changing invader growth rate to growth when rrare

@@
 Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
-invader and resident growth rates \citep{Chesson_1994,Ellner_2016}.
+rare and common species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
 Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
@@
 The storage effect for species $i$ is the contribution of covariation between $E$ and 
-$C$ to a difference between invader and resident mean growth rates, when $i$ is the invader.
+$C$ to a difference between rare and common species mean growth rates, when $i$ is the invader.
 Specifically, storage effects for species $i$ are 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``invasion growth rate'' [IGR] of species $i$, that is, the mean growth rate of 
+where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' [GWR] of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
@@
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
-influences of environment and competition on the invader or resident have been rendered independent.  
+influences of environment and competition on the either species have been rendered independent.  
 A subscript 
@@
 Storage effects for species $i$ were computed by comparing
-differences between invader and resident mean growth rates for the original model and for a 
+differences between rare and common species mean growth rates for the original model and for a 
 modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
@@
 completely independent in the modified model, storage effects quantify the contribution of the 
-whole $EC$ relationship, in all its aspects, to the invader/resident growth rate
+whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
 difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
@@
 \cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
-which equals the IGR. The decomposition, which we use below, is similar to and motivated by the 
+which equals the GWR. The decomposition, which we use below, is similar to and motivated by the 
 original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
@@
 First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
-it was then sufficient to decompose the invasion growth rate of the weaker competitor, species 1. 
+it was then sufficient to decompose the growth when rare of the weaker competitor, species 1. 
 Second, we showed that, for the 
@@
 To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
-ATAs but not without, or vice versa, we plotted the IGR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
-as well as the
-IGR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
+ATAs but not without, or vice versa, we plotted GWR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
+as well as
+GWR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
 For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
@@
 trait-difference space to be the set of 
-values $\mu_1 - \mu_2$ for which the the IGR, $\olr_{i \bs i}$, is positive. This 
+values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
 is the subspace for which the two species in the lottery model coexist in the presence of ATAs.
@@
 of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
-made it sufficient to decompose only the invasion growth rate of species 1, for this 
+made it sufficient to decompose only the growth when rare of species 1, for this 
 example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
@@
 at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
-$\Delta_i^{[EC]}/\text{IGR}$ was displayed using color, to convey the 
-importance of the ATA contribution relative to the whole IGR. Larger magnitudes
+$\Delta_i^{[EC]}/\text{GWR}$ was displayed using color, to convey the 
+importance of the ATA contribution relative to the whole GWR. Larger magnitudes
 of this quantity indicate greater importance of ATAs. The case
-$\Delta_i^{[EC]}>\text{IGR}>0$ (which produces $\Delta_i^{[EC]}/\text{IGR}>1$)
-corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{IGR}<0$
-(which also produces $\Delta_i^{[EC]}/\text{IGR}>1$) corresponds to ATA exclusion.
-The condition $\Delta_i^{[EC]}/\text{IGR}>1$ encompasses the cases for which ATA
+$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/\text{GWR}>1$)
+corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
+(which also produces $\Delta_i^{[EC]}/\text{GWR}>1$) corresponds to ATA exclusion.
+The condition $\Delta_i^{[EC]}/\text{GWR}>1$ encompasses the cases for which ATA
 exclusion or rescue occurred, though the condition does not delineate, by itself,
@@
 exclusion'' (see Methods) are possible.
-For some of the panels on Fig. \ref{fig:fig3}, the IGR line falls between the ATA
+For some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
 line and the $x$ axis for some values of $\sigma$, e.g., see panel h. 
@@
 Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
-represent the IGR when the effects of ATAs are included, and blue lines are the IGR when
+represent GWR when the effects of ATAs are included, and blue lines are GWR when
 the effects of ATAs are removed. These two lines frequently crossed the $x$ axis in 
@@
 Fig.\ref{fig:fig5} plots the importance of various coexistence mechanisms for the 
-persistence of \emph{F. crotonensis} and they're sum, the IGR
+persistence of \emph{F. crotonensis} and they're sum, the GWR
 (see eq. \ref{eq:full_decomp}), for different varying temperature parameters 
 ($a$, $P$, and $\theta_0$) with two of the three parameters set constant in each plot.
-Increasing amplitude and mean temperature showed increasing trends in the IGR and 
+Increasing amplitude and mean temperature showed increasing trends in the GWR and 
 decomposed coexistence mechanisms, while increasing periodicity showed less change. 
 Within the investigated parameters of Fig.\ref{fig:fig5}, ATA contribution, $\Delta_i^{[EC]}$, 
-is always positive, even in cases where IGR is negative, and is almost always second to 
+is always positive, even in cases where GWR is negative, and is almost always second to 
 the contribution of varying environment, $\Delta_i^{E}$, 
@@
 
-Fig.\ref{fig:fig6} displays the subsets of environmental parameters for which coexistence occurs or fails and the degree of ATA contribution to the IGR.
-Specific to this model and parameter ranges, because $\Delta_i^{[EC]}$ is always positive, all positive values of $\Delta_i^{[EC]}/\text{IGR}$ indicate coexistence and negative values of $\Delta_i^{[EC]}/\text{IGR}$ indicate failure of coexistence.
-Futhermore, ATA exlusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>\text{IGR}$.
-To preserve simplicity and readibility of the plots, $|\Delta_i^{[EC]}/\text{IGR}>1|$ were capped at $|1|$.
-Higher magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ tended to occur as the IGR became smaller and smaller, thus creating the sharp contrast in the plots where highest intensity positive spaces quickly change into highest intensity negative spaces with decreasing parameters.
+Fig.\ref{fig:fig6} displays the subsets of environmental parameters for which coexistence occurs or fails and the degree of ATA contribution to GWR.
+Specific to this model and parameter ranges, because $\Delta_i^{[EC]}$ is always positive, all positive values of $\Delta_i^{[EC]}/\text{GWR}$ indicate coexistence and negative values of $\Delta_i^{[EC]}/\text{GWR}$ indicate failure of coexistence.
+Furthermore, ATA exclusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>\text{GWR}$.
+To preserve simplicity and readability of the plots, $|\Delta_i^{[EC]}/\text{GWR}>1|$ were capped at $|1|$.
+Higher magnitudes of $\Delta_i^{[EC]}/\text{GWR}$ tended to occur as GWR became smaller and smaller, thus creating the sharp contrast in the plots where highest intensity positive spaces quickly change into highest intensity negative spaces with decreasing parameters.
 Like the result of Fig.
@@
 \ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence and the effect of periodicity of temperature fluctuation is small.
-Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{IGR}$ changes from negative to positive.
-
-IGR decomposition of \emph{C. psuedostelligera} across the same subsets of
-temperature parameters revealed that its IGR is always positive; though, the ATA contribution is not meaningful to its persistence (see SI section).
+Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{GWR}$ changes from negative to positive.
+
+GWR decomposition of \emph{C. psuedostelligera} across the same subsets of
+temperature parameters revealed that its GWR is always positive; though, the ATA contribution is not meaningful to its persistence (see SI section).
 %Mean EC shows a rescue effect against variance in E
-Thus, the coexistence of the two diatom species depends only on the persistance of 
-\emph{F. crotonensis} and its ability to invade when rare. Additionally, because ATAs 
-meaningfully contribute to the IGR of \emph{F. crotonensis}, ATAs then also contribute 
+Thus, the coexistence of the two diatom species depends only on the persistence of 
+\emph{F. crotonensis} and its ability to increase when rare. Additionally, because ATAs 
+meaningfully contribute to GWR of \emph{F. crotonensis}, ATAs then also contribute 
 meaningfully to the coexistence of \emph{F. crotonensis} and \emph{C. psuedostelligera}.
@@
 For some parameters, the contribution of ATAs was the driving mechanism of 
-coexistence – the IGR would be negative without it – deeming the scenario a case of 
+coexistence – GWR would be negative without it – deeming the scenario a case of 
 ATA rescue. For other parameter values, ATA exclusion occurs where ATAs are the 
 driving mechanism of the failure to coexist (exclusion) and the lack of ATAs would result in 
-a positive IGR. Parameter
+a positive GWR. Parameter
 subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists} - species which only persist in habitat patches where ATAs contribute to its coexistence.
@@
 \caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
-other components of the MCT decomposition. All panels show the IGR and the decomposition of
-the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
+other components of the MCT decomposition. All panels show GWR and the decomposition of
+GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
 $\delta$ and $\mu_1-\mu_2$ 
-are fixed at values specified in the margins, and the IGR and components of the decomposition
+are fixed at values specified in the margins, and GWR and components of the decomposition
 are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
 has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
-to species coexistence or failure to coexist. In some instances, the IGR line (orange) falls 
+to species coexistence or failure to coexist. In some instances, the GWR line (orange) falls 
 between the ATA line (red)
@@
 competitor: both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
-All panels show the IGR, the ATA term of the extended MCT [$\Delta_i^{[EC]}$ in equation 
+All panels show GWR, the ATA term of the extended MCT [$\Delta_i^{[EC]}$ in equation 
 (\ref{eq:full_decomp})], and the difference of these quantities. 
@@
 or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
-see Methods). Values of $\mu_1-\mu_2$ for which the IGR is positive but 
-$\text{IGR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
-values of $\mu_1-\mu_2$ for which the IGR is negative but $\text{IGR}-\Delta_i^{[EC]}$ is positive 
+see Methods). Values of $\mu_1-\mu_2$ for which GWR is positive but 
+$\text{GWR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
+values of $\mu_1-\mu_2$ for which GWR is negative but $\text{GWR}-\Delta_i^{[EC]}$ is positive 
 correspond to ATA exclusion (shaded violet). Top panels also include domains
@@
 \caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} when 
-\emph{F. crotonensis} is the invader is meaningfully large compared to 
-other components of the MCT decomposition. All panels show the IGR and the decomposition of
-the IGR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
+\emph{F. crotonensis} is rare is meaningfully large compared to 
+other components of the MCT decomposition. All panels show GWR and the decomposition of
+GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
 %***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
@@
 %***JASMIN: Done 
-the IGR and components of the MCT decomposition
+GWR and components of the MCT decomposition
 are then plotted against the other parameter. The original value of the varying parameter 
@@
 indicating that ATAs can contribute meaningfully
-to diatom coexistence in this system. In some instances, the IGR line (orange) falls 
+to diatom coexistence in this system. In some instances, the GWR line (orange) falls 
 between the ATA line (red)
 and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the invader, \emph{F. crotonensis}. 
+coexistence and exclusion of the rare species, \emph{F. crotonensis}. 
 %***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
@@
 Coexistence in this system and within 
-our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
-show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
-\emph{F. crotonensis} is the resident.}  \label{fig:fig5}
+our defined parameters only depends on GWR of \emph{F. crotonensis} so we only 
+show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is rare and 
+\emph{F. crotonensis} is common.}  \label{fig:fig5}
 \end{figure}
@@
 \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-\caption{The degree of ATA contribution to the IGR change with parameters of environmental variation, i.e., temperature 
-oscillation. The ratio $\Delta_i^{[EC]}/\text{IGR}$ constitutes the degree of ATA contribution to thee IGR with larger 
+\caption{The degree of ATA contribution to GWR change with parameters of environmental variation, i.e., temperature 
+oscillation. The ratio $\Delta_i^{[EC]}/\text{GWR}$ constitutes the degree of ATA contribution to GWR with larger 
 magnitudes signifying larger ATA contributions. In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/
-\text{IGR}$ only result from positive $\Delta_i^{[EC]}$ and positive IGR and negative $\Delta_i^{[EC]}/\text{IGR}$ only 
-result from positive $\Delta_i^{[EC]}$ and negative IGR. Thus, cyan areas indicate failure to coexist and magenta areas 
+\text{GWR}$ only result from positive $\Delta_i^{[EC]}$ and positive GWR and negative $\Delta_i^{[EC]}/\text{GWR}$ only 
+result from positive $\Delta_i^{[EC]}$ and negative GWR. Thus, cyan areas indicate failure to coexist and magenta areas 
 indicate coexistence with ATA contribution increasing with color intensity. The most intense areas of magenta indicate 
-parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{IGR}$ 
+parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{GWR}$ 
 >1 were capped at $|1|$. 
 Coexistence in this system and within 
-our defined parameters only depends on the IGR of \emph{F. crotonensis} so we only 
-show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is the invader and 
-\emph{F. crotonensis} is the resident.} \label{fig:fig6}
+our defined parameters only depends on GWR of \emph{F. crotonensis} so we only 
+show this case.} \label{fig:fig6}
 \end{figure}

### 2022-03-09 paper/Paper.Rnw
finished editing through Results

@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT, decomposing storage effect to show the influence of ATAs.
+We extend the MCT, decomposing storage effects to show the influence of ATAs.
 Analysis of a simple model
@@
 Every ecological discipline involves the study of relationships among biological and environmental variables,
-but standard methods for examining the association between two variables provide only a small 
+but standard statistical methods for examining the association between two variables provide only a small 
 portion of the overall information content of the
-relationship \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
+association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
 the three pairs of variables in Fig. \ref{fig:noise} all have the same
@@
 to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
-longer lasting increases the degree of right-tail association between early- and late-spring rainfall. 
+longer lasting increases the degree of right-tail association between early and late spring rainfall. 
 Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
@@
 type we describe here a possibility.
-To out knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
 not been studied, but the importance for coexistence of changes 
@@
 plays a role in coexistence. We develop each step of our new theory for a general 
-two-species model, below, and specifically for the lottery model of the Introduction, 
-in SI section \ref{SI-SIsect:TheoryForLottery}.
+two-species model (below), and also specifically for the lottery model of the Introduction 
+(SI section \ref{SI-SIsect:TheoryForLottery}).
 MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
@@
 that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-
 So we begin by defining storage effects. Our mathematical notation parallels
@@
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
-influences of environment and competition on the either species have been rendered independent.  
+influences of environment and competition on either species have been rendered independent.  
 A subscript 
@@
 \cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
-which equals the GWR. The decomposition, which we use below, is similar to and motivated by the 
+which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
 original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction. To pursue goal G2, below, will quantify
-the magnitude of the new ATA term in this expression, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
+which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
+to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
+the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
 
@@
 Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
-$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, with details in SI section \ref{SI-sect:noise}.
+$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, details in SI section \ref{SI-sect:noise}.
 To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
-to coexistence that come from ATA versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
+to coexistence that come from ATAs versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
 $\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
-against the independent parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and then we compared their
-values on the plots. Some mathematical shortcuts, summarized here, greatly simplified this plotting. 
+against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and we compared 
+values on plots. Some mathematical shortcuts, summarized here, simplified plotting. 
 First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
-it was then sufficient to decompose the growth when rare of the weaker competitor, species 1. 
+it was then sufficient to decompose the GWR of the weaker competitor, species 1. 
 Second, we showed that, for the 
 lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they all depend only on the difference $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we only plotted 
+instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we plotted 
 against $\mu_1 - \mu_2$. 
@@
 we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
-Relative importance of contributions was judged from the 
-plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
-the contributions of other mechanisms.
+%Relative importance of contributions was judged from the 
+%plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
+%the contributions of other mechanisms.
 Mathematical simplications for the $\delta=1$ case make it straightforward to see that
 coexistence is impossible and that the case is relatively trivial (SI 
-section \ref{SI-sect:delta1}), so that case is not considered further.  
+section \ref{SI-sect:delta1}), so that case is not plotted.  
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
-Computationally efficient means of carrying out these simulations, for the lottery model, are elaborated in 
+Computationally efficient means of performing these simulations, for the lottery model, are in 
 SI section \ref{SI-sect:efficient_computation}.
@@
 quantities is positive and the other is negative; this set of values characterizes the species which can coexist
-with ATAs, but not without, or vice versa. The one dimensional space of values $\mu$ is a species trait space,
-and $\mu_1$ and $\mu_2$ can be regarded as two points in it. But we instead define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
+with ATAs, but not without, or vice versa. We define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
 Points of the trait-difference space characterize possible combinations of traits of the two 
@@
 $\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
-trait-difference space to be the set of 
+trait-difference space to be the
 values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
-is the subspace for which the two species in the lottery model coexist in the presence of ATAs.
-We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the set of values
+is the subspace for which the two lottery model species coexist in the presence of ATAs.
+We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the values
 $\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
-This is the subspace for which the two species in the lottery model coexist in the absence of 
-ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ is intended to indicate the
+This is the subspace for which the two lottery model species coexist in the absence of 
+ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ indicates the
 absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
@@
 values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
-Similarly, we define ``ATA exclusion space'' to be the set difference 
+We define ``ATA exclusion space'' to be the set difference 
 $\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
@@
 
-%***DAN: Methods paras here for G4
+%G4
 To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
@@
 system of freshwater diatoms explored by \cite{Descamps_2005}.
-Those authors used measurements of growth rates of \emph{Cyclotella pseudostelligera} 
-and \emph{Fragilaria crotonensis} over a range of temperatures to parametrize a resource
+Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
+\emph{Cyclotella pseudostelligera} over a range of temperatures to parametrize a resource
 competition model. The model is 
@@
 of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
-made it sufficient to decompose only the growth when rare of species 1, for this 
+made it sufficient to decompose only the GWR of species 1, for this
+%***DAN: Please search for "growth when rare" and replace with GWR (except for the first instance,
+%where GWR is defined). If one defines an acronym, one then has to use it.
 example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
@@
 
-%***DAN: Jasmin, I edited the below paragraph pretty heavily, please check,
-%especially the quantitative statements at the end.
-%***JASMIN: Good 
+%***DAN: I see many instances of IGR have been replaced by GWR, but there are some IGRs remaining, e.g., in the sup mat, and
+%on the figures.
+
 To display results, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
@@
 (which also produces $\Delta_i^{[EC]}/\text{GWR}>1$) corresponds to ATA exclusion.
-The condition $\Delta_i^{[EC]}/\text{GWR}>1$ encompasses the cases for which ATA
+Thus the condition $\Delta_i^{[EC]}/\text{GWR}>1$ encompasses the cases for which ATA
 exclusion or rescue occurred, though the condition does not delineate, by itself,
@@
 tended to be negative, weakening total storage effects.
+
+%***DAN: I think ATA rescue and ATA exclusion are mislabeled in the legend of Fig. 4. I think they got switched. Please check and correct if I am right.
 
@@
 %G4
-Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
-show that the ATA contributions have magnitudes comparable to the 
-magnitudes of other terms of the extended MCT decomposition and 
-parameters subspaces exist where ATA rescue occurs. 
-Fig.\ref{fig:fig5} plots the importance of various coexistence mechanisms for the 
-persistence of \emph{F. crotonensis} and they're sum, the GWR
-(see eq. \ref{eq:full_decomp}), for different varying temperature parameters 
-($a$, $P$, and $\theta_0$) with two of the three parameters set constant in each plot.
-Increasing amplitude and mean temperature showed increasing trends in the GWR and 
-decomposed coexistence mechanisms, while increasing periodicity showed less change. 
-Within the investigated parameters of Fig.\ref{fig:fig5}, ATA contribution, $\Delta_i^{[EC]}$, 
-is always positive, even in cases where GWR is negative, and is almost always second to 
-the contribution of varying environment, $\Delta_i^{E}$, 
-including for the temperature parameters which \emph{F. crotonensis} 
-and \emph{C. pseudostelligera} coexisted \emph{in vitro} ? \cite{Descamps_2005}.  
-ATA rescue is revealed in Fig.\ref{fig:fig5} within parameter subspaces where the absence of 
-ATA contribution would lead to exclusion of \emph{F. crotonensis}.
-
-Fig.\ref{fig:fig6} displays the subsets of environmental parameters for which coexistence occurs or fails and the degree of ATA contribution to GWR.
-Specific to this model and parameter ranges, because $\Delta_i^{[EC]}$ is always positive, all positive values of $\Delta_i^{[EC]}/\text{GWR}$ indicate coexistence and negative values of $\Delta_i^{[EC]}/\text{GWR}$ indicate failure of coexistence.
-Furthermore, ATA exclusion does not occur and ATA rescue occurs in positive cases where $\Delta_i^{[EC]}>\text{GWR}$.
-To preserve simplicity and readability of the plots, $|\Delta_i^{[EC]}/\text{GWR}>1|$ were capped at $|1|$.
-Higher magnitudes of $\Delta_i^{[EC]}/\text{GWR}$ tended to occur as GWR became smaller and smaller, thus creating the sharp contrast in the plots where highest intensity positive spaces quickly change into highest intensity negative spaces with decreasing parameters.
-Like the result of Fig.
-\ref{fig:fig5}, we see in Fig.
-\ref{fig:fig6} that amplitude and mean temperature have the most influence on coexistence and the effect of periodicity of temperature fluctuation is small.
-Coexistence occurs as $\theta_0$ and $a$ increase and $\Delta_i^{[EC]}/\text{GWR}$ changes from negative to positive.
-
-GWR decomposition of \emph{C. psuedostelligera} across the same subsets of
-temperature parameters revealed that its GWR is always positive; though, the ATA contribution is not meaningful to its persistence (see SI section).
-%Mean EC shows a rescue effect against variance in E
-Thus, the coexistence of the two diatom species depends only on the persistence of 
-\emph{F. crotonensis} and its ability to increase when rare. Additionally, because ATAs 
-meaningfully contribute to GWR of \emph{F. crotonensis}, ATAs then also contribute 
-meaningfully to the coexistence of \emph{F. crotonensis} and \emph{C. psuedostelligera}.
+Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence 
+which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
+show that ATA contributions to coexistence often have magnitudes comparable to the 
+magnitudes of other terms of the extended MCT decomposition (Fig. \ref{fig:fig5}), and that 
+ATA rescue occurs for many combinations of parameters for this system (Fig. \ref{fig:fig5},
+but especially Fig. \ref{fig:fig6}).
 
@@
 
-# notation in ellner one could also use partial sharp 
+%notation in ellner one could also use partial sharp 
 %reread ellner paper 
@@
 
-The authors thank Shyamolina Ghosh for initially suggesting the research problem, and Lawrence Sheppard, Max Castorani,
+The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
 Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Steve Ellner, and Robin Snyder for helpful discussions. 
@@
 \end{figure}
-%***DAN: Jasmin, I appologize for the following pedantic request, I explain below why it is necessary. Can you please change panel
-%labels on this and all figure from the format "A", "B", "C" to the format "(a)", "(b)", "(c)" (i.e., lower case panel labels
-%with parentheses around them). The reason for this is that this is Ecology Letters style. If we do not follow their style then their
-%copy editors will make the change for us. I would prefer not to have copy editors messing around with our figures. Though Ecology Letters
-%copy editors are usually pretty good, I have had many bad experiences with copy editors, generally, and I have had at least a few bad
-%experiences with Ecology Letters copy editors. What happens is, they introduce errors into a paper while typesetting it, and one has to
-%check their work very carefully to avoid those mistakes going to press. It's hard to blame them since it must be the most boring job 
-%in the world, but anyway we do not want to have errors in our published paper as a result. The best way to prevent them introducing 
-%errors into our figures is to not given them any reason to make any changes at all to our figures. The same change has to be 
-%made to all main text and sup mat figures, and then all callouts to figures have to also be so changed, please. I will do these
-%as much as possible while I edit, but if you could please also check it will help, and I am relying on you to change the panel labels
-%in the figured themselves. I am sorry again for the tedium, here. Maybe some day they will improve the publication process so we do not 
-%have to worry about nonsense like this.
-%***JASMIN: changed labels of figures. Done.
 
@@
 of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
-probably of minimal practical importance. 
+probably of minimal importance. 
 Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
@@
 
-%***DAN: Jasmin, some changes are needed to the figure itself. Please make the text in the legend bigger.
-%It looks as though a blue line is for ATA effects, and a red line is for the effects of correlation per se
-%(that's what it says in the legend),
-%but that differs from how you did it in the other figs. Or perhaps the legend is just wrong. Please check and
-%fix so as to be consistent with the rest of the figures in which color represents which quantity and which
-%notation represents which quantity. Also, please change panel labels to be of the format "(a)" instead of 
-%"A". Make the panel labels bigger, as well as the axis numbers and labels.
-
-%***DAN: If I interpret correctly that the blue lines on Fig. 5 are the ATA contribution (needs to be 
-%changed because you use a red line on other figs), there are large areas on panels a and c where the 
-%IGR is between the ATA-contribution line and then x axis. Those are areas of ATA rescue (though you only
-%know this after looking at the analogous plot for the other species). I think in 
-%the next version of the figure you should consider highlighting them in a similar way to how you did 
-%in Fig. 4. Those are a couple nice wide zones of ATA rescue! Let me know if you want to talk about this.
-%***JASMIN: Fixed the line colors. Will block in the rescue areas. 
 \begin{figure}
 \includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{The contribution of ATAs to species coexistence in the diatom system of \cite{Descamps_2005} when 
-\emph{F. crotonensis} is rare is meaningfully large compared to 
+\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
+in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
 other components of the MCT decomposition. All panels show GWR and the decomposition of
 GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
-%***DAN: This is for one species invading the other, right? Need to specify that, and say which is 
-%invading which, and, at the end of the 
-%caption indicate that the reverse invasion scenario is in sup mat and cite the corresponding figure there.
-%***JASMIN: Done - ish. Need to make sup mat section 
 For each panel, 
-two of the three parameters defining the temperature oscillation (i.e., the amplitude, $a$; period, $P$; 
-and mean, $\theta_0$; see Methods) are fixed at the original values used by 
-\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C) and 
-%***DAN: I added units for a  and \theta_0 here, please check these are correct 
-%***JASMIN: Done 
+two of the three parameters defining the temperature oscillation (the amplitude, $a$; period, $P$; 
+and mean, $\theta_0$; see Methods) were fixed at the values used by 
+\cite{Descamps_2005} ($a=6^\circ$C, $P=60$d, $\theta_0=18^\circ$C), and 
 GWR and components of the MCT decomposition
-are then plotted against the other parameter. The original value of the varying parameter 
-is indicated by the vertical dashed magenta lines. ATA contributions 
+were then plotted against the third parameter. The original values of the varying parameters 
+are indicated by vertical dashed magenta lines. ATA contributions 
 had magnitude comparable to other components, 
-%***DAN: Make sure this is still true after the correct version of C is put in, and after we
-%resolve the potential problem with the wiggles on the plot. Reword if not.
 indicating that ATAs can contribute meaningfully
@@
 and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the rare species, \emph{F. crotonensis}. 
-%***DAN: Not the weaker competitor, actually, right, but rather whichever competitor was being considered as the invader
-%for these plots? If I'm right about that then please reword accordingly. 
-%***JASMIN: Right, Done. 
-Standard errors of plotted quantities ?.
-%***DAN: Don't forget to fill in the standard error stuff
-%***DAN: Here is where you add a statement citing another figure in the sup mat that takes the other
-%species as the invader.
-%***JASMIN: Done
-Coexistence in this system and within 
-our defined parameters only depends on GWR of \emph{F. crotonensis} so we only 
-show this case; though, an analogous plot (SI Fig. ) was generated for the reverse scenario where \emph{C. pseudostelligera} is rare and 
-\emph{F. crotonensis} is common.}  \label{fig:fig5}
+coexistence and exclusion of \emph{F. crotonensis}. 
+Standard errors of plotted quantities were always less then X, so are not shown.
+%***DAN: Don't forget to fill in the standard error stuff, please, Jasmin.
+For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
+was always positive both including and
+excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
+sufficient coexistence metric.}  \label{fig:fig5}
 \end{figure}
 
+%***DAN: Jasmin, some changes to this figure are needed. Please make axis labels and numbers bigger. Please put 
+%axis labels outside the axes instead of inside. Please make panel labels bigger. Please make coutour labels
+%slightly bigger.
 \begin{figure}
 \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-\caption{The degree of ATA contribution to GWR change with parameters of environmental variation, i.e., temperature 
-oscillation. The ratio $\Delta_i^{[EC]}/\text{GWR}$ constitutes the degree of ATA contribution to GWR with larger 
-magnitudes signifying larger ATA contributions. In this model, for the plotted parameters, positive $\Delta_i^{[EC]}/
-\text{GWR}$ only result from positive $\Delta_i^{[EC]}$ and positive GWR and negative $\Delta_i^{[EC]}/\text{GWR}$ only 
-result from positive $\Delta_i^{[EC]}$ and negative GWR. Thus, cyan areas indicate failure to coexist and magenta areas 
-indicate coexistence with ATA contribution increasing with color intensity. The most intense areas of magenta indicate 
-parameter subspaces where ATA rescue occurs. For simplicity and readability magnitudes of $\Delta_i^{[EC]}/\text{GWR}$ 
->1 were capped at $|1|$. 
-Coexistence in this system and within 
-our defined parameters only depends on GWR of \emph{F. crotonensis} so we only 
-show this case.} \label{fig:fig6}
+\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
+in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
+the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
+indicating greater importance of ATAs and values greater than $1$ corresponding to either
+ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
+$\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
+magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
+were remapped to 1 and -1, respectively, for readability.
+For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
+excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.} \label{fig:fig6}
 \end{figure}

### 2022-03-09 paper/SupMat.Rnw
finished editing through Results

@@
 
-Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$ and assuming without loss of generality that $\mu_i \leq \mu_j$ so that
-species $i$ is the weaker competitor and $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ can be regarded as a coexistence metric, we have
+Combining the decompositions of $\olr_{i \bs i}$ and $\olr_{j \bs i}$, we have
 \begin{equation}
@@
 
-\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
-
-%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.
-%It has been added but Dan needs to edit it.
-
-In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
-From section \ref{sect:math_simplifications}, 
-\begin{align}
-\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
-&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
-&= 0,
-\end{align}
-and
-\begin{align}
-\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
-&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
-&= 0. %\label{eq:decomp:j:EnatC}
-\end{align}
-Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
-\begin{equation}
-\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
-\end{equation}
-and 
-\begin{equation}
-\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
-\end{equation}
-Similarly for IGR $-\Delta_i^{[EC]}$, 
-\begin{equation}
-(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
-\end{equation}
-Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.
-
 \section{Details of diatom model setup}\label{sect:diatom_model}
@@
 
-\section{Decomposing storage effects in the diatom case study}\label{sect:Empircal_app}
+\section{     Decomposing storage effects in the diatom case study}\label{sect:Empircal_app}
 
@@
 
-\section{\emph{Cyclotella psedosteligera} as the invader}\label{sect:Invader2}
+\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
+
+%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.
+%It has been added but Dan needs to edit it.
+
+In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
+From section \ref{sect:math_simplifications}, 
+\begin{align}
+\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
+&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
+&= 0,
+\end{align}
+and
+\begin{align}
+\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
+&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
+&= 0. %\label{eq:decomp:j:EnatC}
+\end{align}
+Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
+\begin{equation}
+\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
+\end{equation}
+and 
+\begin{equation}
+\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
+\end{equation}
+Similarly for IGR $-\Delta_i^{[EC]}$, 
+\begin{equation}
+(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
+\end{equation}
+Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.
+
+\section{     \emph{Cyclotella psedosteligera} as the invader}\label{sect:Invader2}
 We generated plots analougous to Fig.

### 2022-03-10 paper/SupMat.Rnw
some changes to sup mat, mainly

@@
 
-%***DAN: Jasmin, please start on this section when you get a chance. Here are some topics that need
-%to be covered:
-Estimations of $Q_i(t)$, $V_i(t)$, and $K_i(t)$, temperature dependence functions for 
-cell quota, reproduction rate, and half saturation, respectively, were used to 
-paramaterize the competing diatom model outlined in the paper. 
-Values of and subsequent function estimations are plotted in Figure \ref{fig:VQK}. 
-\cite{Descamps_2005} obtained estimated values for $Q_i$, $V_i$, and $K_i$ from 
-single species nine-day batch experiments at four different temperatures ($6^\circ$C, 
-$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and further
-computed linear interpolations to
-functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$. These values and functions are 
-indicated in Figure \ref{fig:VKQ} by open circles and dashed lines. 
+Estimates of the temperature dependence functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$, 
+produced by earlier authors \cite{Descamps_2005,Ellner_2016,Ellner_2019} and which are part of the parameterization
+of the diatom competition model described in the main text, are now described.
+\cite{Descamps_2005} estimated values for $Q_i(t)$, $V_i(t)$, and $K_i(t)$ from 
+single-species, nine-day batch experiments at four different temperatures ($6^\circ$C, 
+$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and 
+used linear interpolations for the
+functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$. Their values and functions are 
+pictured in Fig. \ref{fig:VKQ} using open circles and dashed lines.
 \cite{Ellner_2016, Ellner_2019} used slightly modified values and functions,
-indicated in \ref{fig:VQK} by filled circles and solid lines. 
-Our investigation follows that of \cite{Ellner_2016, Ellner_2019} and 
-uses the modified functions of  $Q_i(t)$, $V_i(t)$, and $K_i(t)$ to analyze coexistence.
-
-%be very clear here. To give one example of the need for care, here, Ellner et al (2019) state on p. 6
-%of their paper that K_1 and K_2 were constant for the range of temperature considered, and they state
-%that they took constant values for these things. However, we used a larger range of temperatures. 
-%Depending on which of K1fun/K2fun versus K1flatfun/K2flatfun (from DDataAndVKQfuns.R) was actually used
-%may make a difference, so we have to specify. Note that even K1flatfun/K2flatfun were not actually flat
-%for the whole temperature range 6-24 degrees C, though they were for the range of temperature Ellner
-%et al considered. We should not say we used flat K1 and K2 functions unless we are sure we only ever
-%used temperatures in the flat ranges of these functions. Better just to describe the functions we
-%actually used, where the best way to describe them *may* be by plotting them. Anyway, what you need to
-%do is make sure you know what versions of each of these 6 functions your code actually used, and then
-%describe the functions precisely in this section.
+pictured in Fig. \ref{fig:VKQ} by filled circles and solid lines. 
+Our investigation follows \cite{Ellner_2016, Ellner_2019} and 
+uses their modified functions.
 
@@
 
-Our decomposition is defined by $E(t)=\theta(t)$, 
+Our decomposition of GWR is defined by the choices $E(t)=\theta(t)$, 
 $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_{i \bs i})=\frac{V_i(E)}{C_{i \bs i}} -D$. 
+Other choices are possible. For instance, \cite{Ellner_2016} also considers
+
 We note that our definition and selection of our "features" of interest are
@@
 coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$.
+
 Like \cite{Ellner_2016}, we set $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ to reflect 
 potential versus actual growth of the diatom populations.
-Choice of drivers and their definitions are up to the investigator,
-but should refelct real variables that influence population growth.
 
@@
 \includegraphics[width=.66\textwidth]{../results_figs/VKQfig.pdf}
-\caption[Species growth rate, silica uptake, and half sturation responses to changes in temperature.]
-{Species maximum growth rate (a), half saturation constant (b), and silica uptake (c) responses to changes in temperature.
-Species 1 is \emph{F.
-crotonensis}, shown in blue, and species 2 is \emph{C.
-pseudostelligera}, shown in red.
-Open circles and dashed lines indicate estimated data and interpolations from \cite{Descamps_2005} and closed circles and solid lines indicate values and function used in our analysis as defined in \cite{Ellner_2016, Ellner_2019}.
-\label{fig:VKQ}}
+\caption[Maximum growth rate, cellular silicate content, and half sturation constant dependences on temperature.]
+{Species maximum population growth rates, $V_i$ (a), half saturation constants for the dependence 
+of growth rates on nutrient concentration, $K_i$ (b), and cellular silicate content, $Q_i$ (c) dependend on 
+temperature in ways explored by \cite{Descamps_2005} and \cite{Ellner_2016, Ellner_2019} and 
+pictured here. Species 1 is \emph{F. crotonensis}, shown in blue, and species 2 is 
+\emph{C. pseudostelligera}, shown in red. Open circles and dashed lines indicate estimated 
+values and interpolations from \cite{Descamps_2005} and closed circles and solid lines indicate 
+values and functions used by \cite{Ellner_2016, Ellner_2019} and in our analyses. The two sets of
+estimates and functions coincide on (c).\label{fig:VKQ}}
 \end{figure}
+%***DAN: I added the last sentence indicating the two sets of data and functions coincide on (c), can you please
+%check that and let me know if it's true, Jasmin?
 

### 2022-03-10 code/MASTER.R

@@ -1,4 +1,5 @@
-#MASTER - assumes the R working directory is the directory this file is in
+#Running this script should reproduce all analyses for the paper 
+#Assumes the R working directory is the directory this file is in, i.e., the code directory for the project
 
 rm(list=ls())
 

### 2022-03-10 paper/SupMat.Rnw
modified part of the caption for the VKQ figure

@@
 values and functions used by \cite{Ellner_2016, Ellner_2019} and in our analyses. The two sets of
-estimates and functions coincide on (c).\label{fig:VKQ}}
+estimates and functions coincide on (c), except for right-most point for species 2. For that species,
+data were not available for $Q$ at $24^\circ$ from the original study of \cite{Descamps_2005}; \cite{Ellner_2016, Ellner_2019}
+assumed constancy of $Q_2$ for $t$ between $18^\circ$ and $24^\circ$.\label{fig:VKQ}}
 \end{figure}

### 2022-03-10 code/diatom/VKQplots.R
added missing point

@@ -40,7 +40,7 @@ legend("bottomleft", legend=c("original data","modified data","original function
        lty=c(NA, NA, 2, 1), pch=c(1, 19, NA, NA), bty='n', pt.cex=c(1.3, .5,NA,NA))
 plotVKQ("K", Tvals, K1data, K2data, K1flat, K2flat, K1fun, K2fun, K1flatfun, K2flatfun)
 mtext("(b)", 3, -1.5, at=6)
-plotVKQ("Q", Tvals, Q1data, Q2data, Q1data, Q2data, Q1fun, Q2fun, Q1fun, Q2fun)
+plotVKQ("Q", Tvals, Q1data, Q2data, Q1data, Q2mod, Q1fun, Q2fun, Q1fun, Q2fun)
 mtext("(c)", 3, -1.5, at=24,)
 axis(1, at=Tvals, tick=F, cex.axis=1.3)
 axis(1, at=min(Tvals):max(Tvals), labels=F)

### 2022-03-10 paper/SupMat.Rnw
further edits to the same caption

--- 2022-03-10 paper/SupMat.Rnw
@@
 values and functions used by \cite{Ellner_2016, Ellner_2019} and in our analyses. The two sets of
-estimates and functions coincide on (c).\label{fig:VKQ}}
+estimates and functions coincide on (c), except for the right-most point for species 2. For that species,
+data were not available for $Q$ at $24^\circ$C from the original study of \cite{Descamps_2005}; \cite{Ellner_2016, Ellner_2019}
+assumed constancy of $Q_2$ for temperature between $18^\circ$C and $24^\circ$C. Data were also not avaialable at
+$24^\circ$C for $K_2$, and \cite{Ellner_2016, Ellner_2019} assumed constancy of $K_2$ for teperatures from 
+$12^\circ$C through $24^\circ$C. \label{fig:VKQ}}
 \end{figure}
-%***DAN: I added the last sentence indicating the two sets of data and functions coincide on (c), can you please
-%check that and let me know if it's true, Jasmin?
 

### 2022-03-10 paper/SupMat.Rnw
finished editing the section on details of the diatom model in the sup mat

@@
 
+%The Q, K and V functions
 Estimates of the temperature dependence functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$, 
@@
 
-%2) You need to specify what choices were made for E and C. The text below, which I cut-and-pasted
-%from the main text, does some of this. Please use an edited version of that text here, edited as you
-%see fit if clarity needs to be improved and to fit the text in with the rest of this new section.
-%You ought to add a bit more exlanation to the text, so one understand why the choice of C make sense
-%and why r is what you've written there.
+%Defintion of E and C
+Our decomposition of GWR is defined by the choices $E_i(t)=\theta(t)$, 
+$C_{i}(t)=(K_i(t)+S(t))/S(t)$, which then imply $r_i(E_i,C_{i})=\frac{V_i(E_i)}{C_{i}} -D$. 
+This choice follows one of the choices made in the R computer code provided by \citep{Ellner_2016,Ellner_2019}.
+Other choices are possible. For instance, in addition to this choice, 
+\cite{Ellner_2016} also considers an alternative choice $E_i(t)=V_i(t)$,
+as well as some other possibilities discussed in their supporting materials and 
+represented in their computer code.
 
-Our decomposition of GWR is defined by the choices $E(t)=\theta(t)$, 
-$C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ and $r_i(E_i,C_{i \bs i})=\frac{V_i(E)}{C_{i \bs i}} -D$. 
-Other choices are possible. For instance, \cite{Ellner_2016} also considers
-
-We note that our definition and selection of our "features" of interest are
-options among many and features need not be limited to $E$ and $C$ \citep{Ellner_2019}. 
-For example \cite{Ellner_2019} decomposes coexistence with respect to fluctuation in
-inflow silicate concentration, $S$, and temperature, and \cite{Ellner_2016} decomposes
-coexistence using classical $E$ and $C$ drivers, but define $E=V(t)$.
-
-Like \cite{Ellner_2016}, we set $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$ to reflect 
-potential versus actual growth of the diatom populations.
-
-%3) You need to outline how the quantities E_i and C_i\i were obtained from simulations. 
-To obtain $E_i$ and $C_{i \bs i}$, the competiton model was simulated for a $T=3000$ number of timesteps to obtain $S$, the extracellular silicate concentration.
-The initial population size of invaderwas set at zero and initial population of the resident was set at 20.
-Environmental temperature fluctuations were defined by the paramters $a$, $P$, and $\theta_0$ for $\theta(t)=asin(2\pi t/P)+\theta_0$.
-Dilution was set to $D=0.09$ and inflow silicate concentration was set at $S_0=35$.
-A burn in time of $\frac{2}{3}T$ was cut from the resulting $S$ and $E_i(t)=\theta(t)$.
-Competitive pressure was calculated using $C_{i \bs i}(t)=(K_i(t)+S(t))/S(t)$.
-
-%***DAN: Some text and notes immediately below were cut and pasted from the main text, in case 
-%having this material helps in writing this new SI section
-%
-%and here, we consider varying parameters of temperature fluctuations and further decompose the 
-%coexistence of \emph{C. pseudostelligera} and \emph{F. crotonensis} to account for contributions 
-%of ATAs and ATA effects as described above. Our decomposition is defined by $E(t)=\theta(t)$, 
-%$C_j(t)=(K_j(t)+S(t))/S(t)$ and $r_j(E_j,C_j)=\frac{V_j(E)}{C_j} -D$. Thus in our exploration, 
-%environmental fluctuation is temperature fluctuation. We first considered the scenario where 
-%\emph{F. crotonensis} is the invader and emph{C. pseudostelligera} is the resident to compute 
-%the coexistence decomposition for \emph{F. crotonensis}. 
-%put reverse in sup mat?
-%***DAN: Yes, I think that would be a good idea. I forgot about that - we were able to only consider 1 invading 2 for 
-%the lottery model, because of some simplifying assumptions we were able to make, but cannot really do that here, right?
-%***JASMIN: Right. Species in the lottery model only differed in mu but here they differ in they're VKQ functions.
-%*** JASMIN: Note to to run the reverse invader-resident scenario and put in SUpMat!
-%do we need to explain how we separated the asymmetries using the sims??
-%***DAN: I am not sure what you mean by this question, Jasmin
-%***JASMIN: I mean should we explain the methodology of the code we wrote that is implemented to compute the 
-% ATA contribution. The functions alignranks, normcor...  (in partialSharp_fxns.R) 
-%***DAN: Yes, we are going to have to do that
+%outline how the quantities E_i and C_i\i were obtained from simulations. 
+To obtain $E_i$ and $C_{i \bs i}$, the competition model was simulated for $T=3000$ time steps.
+Environmental temperature fluctuations were defined by the paramters $a$, $P$, and $\theta_0$ using 
+$\theta(t)=asin(2\pi t/P)+\theta_0$, and this determined $E_i(t)$ for $i=1,2$. To obtain 
+$C_{i \bs i}$, a simulation was carried out starting from $x_i=0$ and $x_j=20$.
+Dilution was set to $D=0.09$ and inflow silicate concentration was set at $S_0=35$. A burn in time 
+of $\frac{2}{3}T$ was cut from the simulation and the $S(t)$ time series from the remainder of the simulation,
+along with $K_i(t)=K_i(\theta(t))$ were plugged in to the expression $C_{i}(t)=(K_i(t)+S(t))/S(t)$.
 

### 2022-03-10 paper/SupMat.Rnw
finished editing S10 and deleted S11 of the sup mat

@@
 
-%Defintion of E and C
+%Defintions of E and C
 Our decomposition of GWR is defined by the choices $E_i(t)=\theta(t)$, 
@@
 
-%outline how the quantities E_i and C_i\i were obtained from simulations. 
+%Outline how the quantities E_i and C_i\i were obtained from simulations. 
 To obtain $E_i$ and $C_{i \bs i}$, the competition model was simulated for $T=3000$ time steps.
@@
 
+%***DAN: Dan still needs to edit this section
 The decomposition of storage effects and quantification of contributions of ATAs depends on 
@@
 
-\section{Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
+\section{      Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
 
-%***DAN: Jasmin mentioned she had some math proving this, it needs to be added here.
-%It has been added but Dan needs to edit it.
-
-In Fig. \ref{MT-fig:fig4} we plot IGR , $\Delta_i^{[EC]}$, and IGR $-\Delta_i^{[EC]}$ (see equation \ref{eq:decomp_overall_delta}) and the text below it.  
-From section \ref{sect:math_simplifications}, 
-\begin{align}
-\olr_{j \bs i} &= \E[\ln(1-\delta+E_j/C_{j \bs i})] \\
-&= \E[\ln(1-\delta+\delta B_j/B_j)] \\
-&= 0,
-\end{align}
-and
-\begin{align}
-\olep_j^{[EC]} &= \E[r_j(E_j,C_{j \bs i})] - \E[r_j(E_j^\nat , C_{j \bs i}^\nat)] \\
-&= \olr_{j \bs i} - \olr_{j \bs i}^\nat \\
-&= 0. %\label{eq:decomp:j:EnatC}
-\end{align}
-Then $q_{ij}$ does not matter for IGR or $\Delta_i^{[EC]}$ because the second term in $\olr_{i \bs i}-q_{ij}\olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}$ are zero so 
-\begin{equation}
-\olr_{i \bs i}-q_{ij}\olr_{j \bs i}=\olr_{i \bs i}
-\end{equation}
-and 
-\begin{equation}
-\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}=\olep_i^{[EC]}
-\end{equation}
-Similarly for IGR $-\Delta_i^{[EC]}$, 
-\begin{equation}
-(\olr_{i \bs i}-q_{ij}\olr_{j \bs i})-(\olep_i^{[EC]}-q_{ij} \olep_j^{[EC]}) = \olr_{i \bs i}-\olep_i^{[EC]}.
-\end{equation}
-Thus $q_{ij}$ does not matter for anything computed for Fig. \ref{MT-fig:fig4}.
-
-\section{     \emph{Cyclotella psedosteligera} as the invader}\label{sect:Invader2}
-We generated plots analougous to Fig.
-\ref{MT-fig:fig5} and Fig.
-\ref{MT-fig:fig6} for the reverse invader-resident scenario where \emph{C.
-pseudosteligera} is the invader and \emph{F.
-crotonensis} is the resisent to compute the IGR and decompose coexistence mechanisms of \emph{C.
-pseudosteligera}.
-Fig.
-\ref{fig:fig5_2} is analogous to Fig.
-\ref{MT-fig:fig5} and shows changes in the IGR and decomposed mechanisms in response to change in one temperature parameter while the other two are held constant.
-Within the parameters of our simulations, the IGR of \emph{C.
-pseudosteligera}
+In Fig. \ref{MT-fig:fig4} of the main text we plotted GWR , $\Delta_i^{[EC]}$, and $\text{GWR}-\Delta_i^{[EC]}$ 
+(see equation \ref{eq:decomp_overall_delta} and the text below it) for the lottery model. We here show
+the resulting plot does not depend on which of the two choices for the scaling factors $q_{ij}$ is used.
+But $\text{GWR} = \olr_{i \bs i} - q_{ij} \olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij}\olep_j^{[EC]}$,
+so it suffices to show $\olr_{j \bs i}=0$ and $\olep_j^{[EC]}=0$. The first of these is known true, \emph{a priori}, 
+and the second was demonstrated in section \ref{sect:math_simplifications}.
 

### 2022-03-11 paper/Paper.Rnw
a few notes about progress editing were added
edited first para of Discussion

@@
 SI section \ref{SI-sect:Empircal_app}.
-%***DAN: Jasmin, you had already created and started on the second of these SI sections, but the first one
-%is new, and needs to be filled in. Please make a start at that when you get a chance. I've put some 
-%notes there indicating what the content of that section should be.
-%***JASMIN: Working on filling out diatom model set up section
 
-%***DAN: I see many instances of IGR have been replaced by GWR, but there are some IGRs remaining, e.g., in the sup mat, and
+%***DAN: I see many instances of IGR have been replaced by GWR, but there are some IGRs remaining, e.g., 
 %on the figures.
@@
 
-We produced and applied a mathematical extension to MCT that, for the first time, 
-considers ATAs for the coexistence of species and maintenance of biodiversity. 
-Implementing new methods of MCT \citep{Ellner_2016, Ellner_2019}, we provided a further tertiary 
-decomposition to MCT and decompose the storage effect into two new terms, 
-$\Delta_i^{[EC]}$, the contribution of ATAs, and $\Delta_i^{[E \nat C]}$, the contribution of correlation \emph{per se}. 
-The results of our application to a simple lottery model and 
-to a competitive diatom model show that ATA contributions to storage effects and to coexistence
-are non-negligible and meaningfully contribute to the coexistence of species and thus the maintenance of biodiversity. 
-Our results reinforce that common methods of correlation fail to reflect and include 
-important information on the structure of the relationship between variables and amends
-coexistence to the list of ecological theory and phenomena that ATAs have been found to 
-be important for \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
+%***DAN: Jasmin, I just want to point out the below paragraph was quite well structured as it was 
+%before I edited it. I only made changes on the sentence level. Nice job! Go aead and delete this 
+%comment once you read it.
+We produced and applied a new extension to MCT that 
+quantifies the impact of ATAs on species coexistence and the maintenance of biodiversity. 
+Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
+we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
+and the contribution of $EC$ correlation \emph{per se}.
+Applications of our extended theory to the lottery model and to models based on a laboratory 
+system of competing diatoms demonstrated that  
+ATA contributions to storage effects and to species coexistence and the maintenance of biodiversity
+can often be comparable in magnitude to the contributions of previously recognized mechanisms.
+Our results highlight that common methods of correlation and covariance fail to include 
+important information on the structure of relationships between variables, and our results add
+coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
+ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
 Though storage effects are typically described as the contribution
-of covariance between drivers to coexistence, classically \emph{EC covariance},
-\citep{Chesson_2012, Ellner_2016, Ellner_2019}, we suggest that storage effects, instead, 
-be described as the contribution of the \emph{relationship} between the drivers.
-Measures of covariance do not fully capture all aspects of associativity throughout the bivariate 
-distribution of the variables; thus, it may be misleading to describe the contribution of the whole 
-EC (or other drivers') relationship as the contribution of EC (or other drivers') covariance.
-Accordingly, we have separated the mechanisms of storage effect 
-into contributions of ATAs and contributions of correlation \emph{per se}.
+of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
+our results suggest that storage effects should be redefined to constitute the 
+contribution of \emph{associations} between $E$ and $C$, including both ATA effects and
+correlation \emph{per se}; the prior terminology, \emph{covariance}, is typically construed 
+as reflecting the linear relationship between variables, ignoring complexities of the association
+such as ATAs.
+
+%***DAN: Dan made it to here in editing and has to stop until after spring break
 

### 2022-03-20 paper/Paper.Rnw
responded to Dan's commented and moved commented out intro to Scraps.txt
adding to discussion a bit

@@
 %except when multiple pubs by the same first author are listed they are all together.
+%***JASMIN: Done
 
@@
 
-%make the above abstract ideas more concrete - 
-%***DAN: This para commented out for length, added a sentence about the systems to the above paragraph to take the place of this text
-%To make the above ideas more concrete, we briefly mention examples from prior work to 
-%which the ideas may apply. 
-%\cite{Chesson_2012} provides two examples of competitive plant communities in which different species
-%are likely influenced differently by phenologically distict environmental variables: an Australian 
-%grass-forb community, and an Arizona community of desert annuals. 
-%The communities follow the criteria required for temporal storage effects to operate \citep{Chesson_2012}. In both 
-%examples, distinct species have different periods of germination, competition is influenced by resource 
-%consumption, and population growth is buffered by overlapping generations or via seed banks.
-%Temporal differences between species in germination rates suggest the possibility that 
-%population growth of distinct species may be driven by rainfall signals in distinct parts
-%of the growing season. Thus all the ingredients seem to be present by which climatic changes may
-%alter ATAs between environmental variables which are relevant for the coexistence
-%of the example species, motivating our efforts, here, to perform a detailed study of how ATAs can influence competition.
-
-%Review of lit on changes in means and variances of climate signals and associated impacts on coexistence
-%**DAN: this para commetned out for length, added a couple of sentences to an aboce para to substitute
-%Although, to our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-%not been studied, the importance for coexistence of changes 
-%in means and variability of environmental variables has been studied in many systems. We briefly 
-%review some of that literature.
-%\cite{diPaola_2012} concluded that coexistence of drought sensitive,
-%flood resistant species with flood sensitive, drought resistant species in 
-%Mediteranean coastal swamp forests of central Italy would be jeopardized
-%by projected changes in mean rainfall. 
-%\cite{Descombes_2020} found that if warming trends promote faster expansion toward 
-%higher altitudes of herbivore populations than of plant populations, newly arrived
-%herbivores at high altitudes can inflate plant species richness by reducing dominant plant species 
-%in favor of subordinates. Contrastingly, \cite{Usinowicz_2021} concluded, also for
-%alpine plants, that upward range expansion under warming trends would be good for the growth and survival of 
-%all species they considered, but that changes in competition would more than offset these direct benefits. 
-%Numerous studies take an ``event-focused'' 
-%perspective \citep{Jentsch_2007}, examining how increased climatic variability, in the
-%form of extreme weather events, can alter competitive interactions. For instance, 
-%\cite{White_2001} showed that single extreme heat events can alter competition 
-%between C3 and C4 plants. And \cite{Parepa_2013} showed that extreme variability in soil
-%nutrient availability for plants, mediated by soil moisture content, can alter 
-%the potential for invasive species to enter an ecosystem and establish, a process that involves
-%the invader overcoming competition from resident species. 
-%Using MCT, it has been demonstrated that increased variability can promote coexistence in plant 
-%communities, via storage effects \citep{Adler_2006,  Angert_2009}.
-
-%***DAN: Jasmin, I commented the below sentence because that body of work is famously 
-%complex and sophisticated, and I did not feel we could do it justice in short space.
-%decreased growth season precipitation can decrease the competitive ability of dominant 
-%forest tree species \citep{Clark_2011}, 
-%***DAN: Jasmin, I commented the below sentence because I'm not sure the two studies 
-%cited there are actually directly about coexistence, in the MCT sense. One of them is
-%a general review on invasive species, and the other has to do with competition,
-%but coexistence is not mentioned. I searched both papers for "coexist" and found
-%no hits. OK, both are relevant, but maybe not so directly, to what we are talking about,
-%and since this paragraph could anyway stand to be shortened a bit, I took these out.
-%Alternatively, dominant species exclusion may be beneficial as rare species can increase 
-%in diversity \citep{Mariotte_2013}, while an increase in dominant species fitness can 
-%reduce community diversity, typical of invasive species \citep{Vitousek_1997}.
-
 %Goal para
@@
 % be altered as part of the overall change. 
+%***JASMIN: Done in text. Need to finish change in figures.
 
@@
 %where GWR is defined). If one defines an acronym, one then has to use it.
+%***JASMIN: Done
 example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
@@
 %on the figures.
+%***JASMIN: Figure 3 and 4 changed
 
@@
 %***DAN: I think ATA rescue and ATA exclusion are mislabeled in the legend of Fig. 4. I think they got switched. Please check and correct if I am right.
+%***JASMIN: Fixed.
 
@@
 
-%***DAN: Jasmin, I just want to point out the below paragraph was quite well structured as it was 
-%before I edited it. I only made changes on the sentence level. Nice job! Go aead and delete this 
-%comment once you read it.
 We produced and applied a new extension to MCT that 
@@
 %More species and more drivers= more things that can be related
-Our further decomposition and quantification of ATA contribution to coexistence can and 
+The extended decomposition of storage effects and quantification of ATA contribution to coexistence can and 
 should also be applied to multi-species models. 
@@
 
+The probability of ATAs being important for the coexistence of species increases as more variables are included in the system, 
+either through multi-species systems or multi-driver systems (coexistence drivers beyond just E and C). 
+When multiple species included, the interaction between different species' E's and C's introduces more relationships 
+to consider and could increase the relative importance of ATA contribution to the species' coexistence. 
+Likewise, coexistence between species can also depend on more than two drivers \citep{Ellner_2019}. 
 %notation in ellner one could also use partial sharp 

### 2022-03-22 paper/Paper.Rnw
editing discussion

@@
 The probability of ATAs being important for the coexistence of species increases as more variables are included in the system, 
-either through multi-species systems or multi-driver systems (coexistence drivers beyond just E and C). 
-When multiple species included, the interaction between different species' E's and C's introduces more relationships 
-to consider and could increase the relative importance of ATA contribution to the species' coexistence. 
+either through the addition of more species or more drivers (beyond just E and C), or both. 
+When multiple species are included in the system, the interaction between different species' E's and C's introduces more relationships 
+to consider that may or may not be asymmetrically related and could increase the relative importance of ATA contribution to the species' coexistence. 
 Likewise, coexistence between species can also depend on more than two drivers \citep{Ellner_2019}. 

### 2022-03-22 paper/Paper.Rnw
adding to discussion

@@
 
-%We note that our expansion, like the decomposition of \cite{Ellner_2019} is generalizable to multi-species and multi-driver models.
-Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
-%More species and more drivers= more things that can be related
-The extended decomposition of storage effects and quantification of ATA contribution to coexistence can and 
-should also be applied to multi-species models. 
-The inclusion of multiple species introduces a multitude of relationships between species'
-environmentally dependent parameters that could possibly be asymmetrically associated. 
-The greater the number of species the greater the possibilities of ATAs being more important 
-to their coexistence.
+%Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
 
-The probability of ATAs being important for the coexistence of species increases as more variables are included in the system, 
+%The inclusion of multiple species introduces a multitude of relationships between species'
+%environmentally dependent parameters that could possibly be asymmetrically associated. 
+%The greater the number of species the greater the possibilities of ATAs being more important 
+%to their coexistence.
+
+The chances of ATAs being important for the coexistence of species increases as more variables are included in the system, 
 either through the addition of more species or more drivers (beyond just E and C), or both. 
@@
 Likewise, coexistence between species can also depend on more than two drivers \citep{Ellner_2019}. 
-%notation in ellner one could also use partial sharp 
-%reread ellner paper 
+The storage effect decomposition presented here is generalizable to coexistence decompositions of multiple variables.
+Following the set up and notation of \cite{Ellner_2019}'s \emph{covariance decomposition}, 
+ATA contributions and correlation \emph{per se} can be decomposed from the \emph{associations} between two variables
+among many from the appropriate term with parenthesized superscripts with the # symbol denoting and separating 
+variables that are varying independently from the variables that are covarying whose relationship we are interested in decomposing. 
+For example, the term $\Delta^{(XY#Z)}$ could be decomposed into $\Delta^{[XY#Z]} + \Delta^{[X \nat Y#Z]}$. 
+As more terms and species are introduced into a system, more relationships between variables must be evaluated,
+more of those relationships may be asymmetrically tail associated, and ATAs may contribute more to the maintenance of 
+biodiversity in that system. 
 %make three different biv dis that are standard normal copula and then pair with a shuffles third

### 2022-03-22 paper/Paper.Rnw
last paragraph

@@
 %Also need to talk about how this stuff will apply to data.
-As an introductory investigation of ATAs within the framework of coexistence theory, 
-we used simple and extreme examples, and future work should include more realism and complexity
-(assess more complex relationships, systems, and models). 
+As an introductory investigation of ATAs within the framework of MCT, 
+we used simple and extreme examples, and future work should assess more complex relationships, systems, and models. 
 In our theoretical exploration, we implemented a simple lottery model and assessed 
 the influence of extremely asymmetric associated noise, i.e., the noise was perfectly correlated
-in exactly one half of the distribution. However, more complex models such as structured demographic models
-or integral projection models should be explored and though similar results may be
-produced, varying deegres of ATA, less extremely shaped?/
-more moderately asymmetric / more realistic? relationships should be implemented.
-Moreover, because our empirical application makes use of a simple and controlled lab system, future 
+in exactly one half of the distribution. However, more complex models (such as structured demographic models
+or integral projection models should be explored )and varying degrees and qualities of ATA, (copula parameters, 
+difference between tails, rate of relatedness(slope of expansion??) should be implemented and explored.
+Moreover, because our empirical application makes use of a simple and controlled lab system, future studies
 should be done to assess the extent of how much ATA are important for complex systems and field systems.
+Because modeling the dynamics of long-lived populations can be challenging, a good field system to start
+exploring the ideas presented here is a system of annual plants. 
 For example, the ATAs between rainfall timing on the coexistence of grasses and forbs 
-and their dominant years in California grasslands \citep{Hallett_2019} could be a fruitful 
-further investigation of biodiversity maintenance in California grasslands and the effects of extreme rainfall fluctuation.
+in California grasslands \citep{Hallett_2019} could be a fruitful 
+further investigation of biodiversity maintenance in California grasslands and the effects of extreme 
+rainfall fluctuation given the highly variable climate of the area.
+Moreover, our approach may provide more insight when expanded to consider fluctuations across space,
+in regards to questions about population synchrony, Moran effects, and the suggestion of possible ATA specialists.
 

### 2022-03-22 paper/Paper.Rnw
comments + edits for fig 2 script and edits to sexpr callouts in paper fig caption

@@
 total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
-was \Sexpr{fig2$coSym} time steps for panel (a) and \Sexpr{fig2$coATA} for panel (b), over the full length of the simulations; 
+was \Sexpr{fig2$coMets$Sym$co} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$co} for panel (b), over the full length of the simulations; 
 whereas the average duration of periods of
-one-species dominance was \Sexpr{fig2$domSym} time steps for panel (a) and \Sexpr{fig2$domATA} for (b). Thus, while periods of 
+one-species dominance was \Sexpr{fig2$coMets$Sym$dom} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$dom} for (b). Thus, while periods of 
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1] + fig2$params[1]}$,
-$\mu_2 = \Sexpr{-(fig2$params[1])}$, $\sigma = \Sexpr{fig2$params[2]}$, and $\delta = \Sexpr{fig2$params[3]}$. 
+on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]$,
+$\mu_2 = \Sexpr{fig2$params[2]}$, $\sigma = \Sexpr{fig2$params[3]}$, and $\delta = \Sexpr{fig2$params[4]}$. 
 See Introduction and Fig. \ref{fig:noise} caption for model details. 

### 2022-03-22 code/Figure2.R

@@ -2,7 +2,7 @@
 
 ### source funtions ###
 #script containing functions to make plot
-source("./popsims_fxns.R")
+source("./popsim_fxns.R")
 
 ### location to save results ###
 noise_loc <- "../results_numeric/noise.RData"

### 2022-03-22 code/popsim_fxns.R

@@ -9,12 +9,11 @@
 # transform noise by sigma and mu as defined by lottery model
 #ARGS:
   #b_tilde      list-3 sets bivariate noise 
-  #rho          common correlation of bivariate noise sets
   #sigma        common sd of bivariate noise sets
   #mudif        mu1-mu2
 #OUT:
   #list with each element as the environmental noise of one of the two species, 3 different types, =length 6
-transform <- function(b_tilde, u_tilde, rho, sigma, mudif){
+transform <- function(b_tilde, sigma, mudif){
   #multiply by sigma, shift second noise by mudif    
   #left tail ATA
   b_l1 <- sigma*b_tilde$l[,1] 
@@ -87,10 +86,6 @@ co.periods <- function(pop, dom, N){
   #return RLE object: list with length element and value (T/F) element 
   return(RLE)
 }
-#source("./pop_sim.R")
-#source("./transform2.R")
-#source("./co.periods.R")
-#end must be an even number
 
 ### co.pPlot ###
 #function to plot pop simulation with highlighted grey areas of noticeable coexistence
@@ -98,7 +93,7 @@ co.periods <- function(pop, dom, N){
   #co.p   RLE object (list length 2 with $length and $value)
   #pop    data.frame of N1 and N2
   #start  first time step to plot
-  #end    last time step to plot, must be even number
+  #end    last time step to plot, **must be even number**
 #OUT:
   #plot
 co.pPlot <- function(co.p, pop, start, end,...){
@@ -136,41 +131,54 @@ co.pPlot <- function(co.p, pop, start, end,...){
 }
 
 ### simsPlot ###
-
+#runs population simulation with and without ATA and
+#computes noticeable coexistence and dominance metrics
+#plots simulations using co.pPlot 
 #ARGS:
-
+	#noise_loc		location of noise .RData object in directory, 
+	#				contains b_tilde and u_tilde
+	#mudif			mean difference between species; mu1-mu2
+	#delta			death rate
+	#sigma			standard deviation of noise
+	#start			first time step to plot	
+	#end				last time step to plot, **must be even**
 #OUT:
+	#plot
+	#list of metrics, length 5
 
[…]

### 2022-03-22 paper/Paper.Rnw
continued wworking on Discussion

@@
 
-%***DAN: Go through and make sure all citation callouts list papers in date order, oldest to newest,
-%except when multiple pubs by the same first author are listed they are all together.
-%***JASMIN: Done
-
 %Lottery model and results based on it
@@
 The storage effect for species $i$ is the contribution of covariation between $E$ and 
-$C$ to a difference between rare and common species mean growth rates, when $i$ is the invader.
+$C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
 Specifically, storage effects for species $i$ are 
@@
 
-We produced and applied a new extension to MCT that 
+We produced and applied a new extension of MCT that 
 quantifies the impact of ATAs on species coexistence and the maintenance of biodiversity. 
@@
 Our results highlight that common methods of correlation and covariance fail to include 
-important information on the structure of relationships between variables, and our results add
+important information on the structure of relationships between variables. Our results add
 coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
@@
 
-%***DAN: Dan made it to here in editing and has to stop until after spring break
+%ATA specialists
+The contributions we have documented of ATAs to species coexistence open the possibility that
+``ATA specialists'' may exist in some contexts. If a stronger competitor is present across a
+landscape and the GWR of a weaker competitor %***DAN: Dan is working right here.
 
-%The example case exapanding upon coexistence decompositions of the \cite{Descamps_2005} chemostat experiment support our earlier findings that ATA contributions are not only non-zero, but are non-neglible and contribute meaningfully to coexistence.
-
-%Our expansion allows for the parsing of storage effects into contribution from a pure symmetrical correlation and structures beyond symmetric correlation.
-Previous studies report storage effect as a contribution to coexistence of EC covariance; yet compute it as the contribution of the EC relationship (covariance and beyond; symmetric and asymmetric).
-
-%\underline{ATA specialization}
-Not only is the contribution of ATA non-neglible but in some cases the presence or absence of ATAs 
-can mediate the result of competition between species, i.e., coexistence or extirpation. 
-For some parameters, the contribution of ATAs was the driving mechanism of 
-coexistence – GWR would be negative without it – deeming the scenario a case of 
-ATA rescue. For other parameter values, ATA exclusion occurs where ATAs are the 
-driving mechanism of the failure to coexist (exclusion) and the lack of ATAs would result in 
-a positive GWR. Parameter
+%Not only is the contribution of ATA non-neglible but in some cases the presence or absence of ATAs 
+%can mediate the result of competition between species, i.e., coexistence or extirpation. 
+%For some parameters, the contribution of ATAs was the driving mechanism of 
+%coexistence -- GWR would be negative without it -- deeming the scenario a case of 
+%ATA rescue. For other parameter values, ATA exclusion occurs where ATAs are the 
+%driving mechanism of the failure to coexist (exclusion) and the lack of ATAs would result in 
+%a positive GWR. Parameter
 subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists} - species which only persist in habitat patches where ATAs contribute to its coexistence.
@@
 
-%\underline{future work: study ATAs in environmental variables and how theyre changing - tie in with specialist}
-
+%future work: study ATAs in environmental variables and how theyre changing - tie in with specialist
 Our study highlights the importance of ATAs for ecological phenomena 
@@
 
-\underline{multispecies models}
+%\underline{multispecies models}
 %Need a para about multi-species models - how would things change? One difference is there are more different
@@
 ATA contributions and correlation \emph{per se} can be decomposed from the \emph{associations} between two variables
-among many from the appropriate term with parenthesized superscripts with the # symbol denoting and separating 
+among many from the appropriate term with parenthesized superscripts with the $\shp$ symbol denoting and separating 
 variables that are varying independently from the variables that are covarying whose relationship we are interested in decomposing. 
-For example, the term $\Delta^{(XY#Z)}$ could be decomposed into $\Delta^{[XY#Z]} + \Delta^{[X \nat Y#Z]}$. 
+For example, the term $\Delta^{(XY \shp Z)}$ could be decomposed into $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$. 
 As more terms and species are introduced into a system, more relationships between variables must be evaluated,
@@
 
-\underline{caveats and future directions}
+%\underline{caveats and future directions}
 %Realism - we usede simple/extreme ATA, simple model and simple lab system, futurw work needed to assess 
@@
 noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]$,
+on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]}$,
 $\mu_2 = \Sexpr{fig2$params[2]}$, $\sigma = \Sexpr{fig2$params[3]}$, and $\delta = \Sexpr{fig2$params[4]}$. 
-See Introduction and Fig. \ref{fig:noise} caption for model details. 
-\label{fig:pedagog}}
+See Introduction and Fig. \ref{fig:noise} caption for model details. \label{fig:pedagog}}
 \end{figure}

### 2022-04-05 paper/Paper.Rnw
finished the README and slight mods to the acks
minor changed to paper

@@
 
-%***DAN: I think we should replace the terminology "invasion growth rate" and "IGR" by "growth when rare" and "GWR",
-% throughout the paper. Has to be done in the main text and sup mat and the figures themselves.
-% Need also search for "invader" since sometimes we use that word. Also think of any other related terminology that needs to 
-% be altered as part of the overall change. 
-%***JASMIN: Done in text. Need to finish change in figures.
-
 The storage effect for species $i$ is the contribution of covariation between $E$ and 
@@
 made it sufficient to decompose only the GWR of species 1, for this
-%***DAN: Please search for "growth when rare" and replace with GWR (except for the first instance,
-%where GWR is defined). If one defines an acronym, one then has to use it.
-%***JASMIN: Done
 example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
@@
 SI section \ref{SI-sect:Empircal_app}.
-
-%***DAN: I see many instances of IGR have been replaced by GWR, but there are some IGRs remaining, e.g., 
-%on the figures.
-%***JASMIN: Figure 3 and 4 changed
 
@@
 tended to be negative, weakening total storage effects.
-
-%***DAN: I think ATA rescue and ATA exclusion are mislabeled in the legend of Fig. 4. I think they got switched. Please check and correct if I am right.
-%***JASMIN: Fixed.
 
@@
 %off this list and add him to the author list.
+%***DAN: DAN: if we add Ellner to the author list take him out here, and also take him out of the acks in the README.
 
@@
 %slightly bigger.
+%***DAN: Do we want to do anything with standard error for this fig?
 \begin{figure}

### 2022-04-08 paper/Paper.Rnw
minor edits to discussion

@@
 #allregres<-matrix(rnorm(10),2,5)
+res_loc <- "../results_numeric/"
 fig2 <- readRDS(file="../results_numeric/fig2data.RDS")
@@
 se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
+se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
+se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
 @
@@
 can and should be explored further. 
-The coupling of the magnitude of ATA contribution and environmental variability, 
+The coupling of ATA contribution magnitude with environmental variability, 
 as shown in our results and figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), 
@@
 may become more important to consider for coexistence and possibly other 
-ecological phenomena as environmental conditions evidently
-become more variable.
-As described in section \ref{section:introduction}, changes in the ATAs between environmental
-variables may occur as the climate becomes more variable and extreme events become more 
-expansive and extensive, causing relationships between separated events to become tighter at one end or another and thus asymmetric. Additionally, 
+ecological phenomena as environmental conditions may evidently
+become more variable (probably need refs).
+As described in section \ref{section:introduction}, ATAs between environmental
+variables may shift as the climate becomes more variable and extreme events become more 
+expansive and extensive, causing associations of distinct temporal events to become 
+more related in one extreme or tail, thus asymmetric. Additionally, 
 if ATA specialist exist, their persistance may
@@
 Our investigation brings to attention the consequences of such
-possible \emph{relationships} in the extremes; however, 
+possible [bivariate] \emph{relationships} in the extremes; however, 
 simpler changes in [the marginal/ univariate distributions of] extreme events themselves such as
@@
 influence the coexistence of species through the storage effect and not other mechanisms, but
-extreme and changing climate is probable to affect other mechanims in other ways.
+extreme and changing climate is probable to affect other mechanims in other ways (refs?).
 %In the Discussion, I intend to talk about how possible changes in tail association 
@@
 framework, is not directly a measure of the environment, but is a measure of response to the 
-environment, measured by a population's vital rates. Environment and environmental response may be related linearly
+environment, via a population's vital rates. Environment and environmental response may be related linearly
 (like lottery model?); though real environmental variables are probable to 
@@
 environmental variables merit future study, but the relationships between populations'
-vital rates should also be investigated for species coexistence and the maintenacne of biodiversity.
+vital rates should also be investigated for species coexistence and the maintenance of biodiversity.
 %Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
@@
 small-variance approximations completely miss out
-on any additions to coexistence beyond simple correlations and does not include our new term, $\Delta^{[EC]}$ or $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
+on any additions to coexistence beyond simple correlations and does not include our new ATA term, $\Delta^{[EC]}$, or variance \emph{per se}, $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
 In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
@@
 the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, 
-obscuring the effects of ATAs. However, because standard MCT's assumption make it impossible to 
+obscuring the effects of ATAs. Because standard MCT's assumption make it impossible to 
 detect interactions outide of correlation \emph{per se}, simulation-based MCT is required to
-detect ATA contributions, and our new extension to it cleary separates and quantifies the 
-relative contribution and importance of ATAs and pure correlation to the storage effect
+detect ATA contributions, and our new extension to it clarifies and cleary separates and quantifies the 
+relative importance of both ATAs and pure correlation in contribution to the storage effect
 and subsequently species coexistence and the maintenance of biodiversity.
@@
 ATA contributions and correlation \emph{per se} can be decomposed from the \emph{associations} between two variables
-among many from the appropriate term with parenthesized superscripts with the $\shp$ symbol denoting and separating 
+among many from the appropriate term with parenthesized superscripts containing the $\shp$ symbol which denotes and separates 
 variables that are varying independently from the variables that are covarying whose relationship we are interested in decomposing. 
@@
 Moreover, because our empirical application makes use of a simple and controlled lab system, future studies
-should be done to assess the extent of how much ATA are important for complex systems and field systems.
+should be done to assess the extent of how much ATAs are important for complex systems and field systems which are 
+often subject to factors beyond our control.
 Because modeling the dynamics of long-lived populations can be challenging, a good field system to start
 exploring the ideas presented here is a system of annual plants. 
-For example, the ATAs between rainfall timing on the coexistence of grasses and forbs 
+For example, the ATAs effcts of rainfall timing on the coexistence of grasses and forbs 
 in California grasslands \citep{Hallett_2019} could be a fruitful 
@@
 coexistence and exclusion of \emph{F. crotonensis}. 
-Standard errors of plotted quantities were always less then X, so are not shown.
-%***DAN: Don't forget to fill in the standard error stuff, please, Jasmin.
+Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
@@
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
-excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.} \label{fig:fig6}
+excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
+Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
 \end{figure}

### 2022-04-14 paper/Paper.Rnw
fix typo

@@
 exploring the ideas presented here is a system of annual plants. 
-For example, the ATAs effcts of rainfall timing on the coexistence of grasses and forbs 
+For example, the ATAs effects of rainfall timing on the coexistence of grasses and forbs 
 in California grasslands \citep{Hallett_2019} could be a fruitful 

### 2022-04-14 paper/Paper.Rnw
fix typo

@@
 influence the coexistence of species through the storage effect and not other mechanisms, but
-extreme and changing climate is probable to affect other mechanims in other ways (refs?).
+extreme and changing climate is probable to affect other mechanisms in other ways (refs?).
 %In the Discussion, I intend to talk about how possible changes in tail association 

### 2022-04-21 paper/Paper.Rnw
added ref for Walter et al 2022
very small things

--- 2022-04-08 paper/Paper.Rnw
@@
 #allregres<-matrix(rnorm(10),2,5)
-res_loc <- "../results_numeric/"
 fig2 <- readRDS(file="../results_numeric/fig2data.RDS")
@@
 se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
-se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
-se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
 @
@@
 correlation \emph{per se}; the prior terminology, \emph{covariance}, is typically construed 
-as reflecting the linear relationship between variables, ignoring complexities of the association
-such as ATAs.
+as reflecting the linear relationship between variables, ignoring complexities of the association,
+such as ATAs, which we now know are important.
 
 %ATA specialists
-The contributions we have documented of ATAs to species coexistence open the possibility that
+Contributions of ATAs to species coexistence open the possibility that
 ``ATA specialists'' may exist in some contexts. If a stronger competitor is present across a
@@
 can and should be explored further. 
-The coupling of ATA contribution magnitude with environmental variability, 
+The coupling of the magnitude of ATA contribution and environmental variability, 
 as shown in our results and figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), 
@@
 may become more important to consider for coexistence and possibly other 
-ecological phenomena as environmental conditions may evidently
-become more variable (probably need refs).
-As described in section \ref{section:introduction}, ATAs between environmental
-variables may shift as the climate becomes more variable and extreme events become more 
-expansive and extensive, causing associations of distinct temporal events to become 
-more related in one extreme or tail, thus asymmetric. Additionally, 
+ecological phenomena as environmental conditions evidently
+become more variable.
+As described in section \ref{section:introduction}, changes in the ATAs between environmental
+variables may occur as the climate becomes more variable and extreme events become more 
+expansive and extensive, causing relationships between separated events to become tighter at one end or another and thus asymmetric. Additionally, 
 if ATA specialist exist, their persistance may
@@
 Our investigation brings to attention the consequences of such
-possible [bivariate] \emph{relationships} in the extremes; however, 
+possible \emph{relationships} in the extremes; however, 
 simpler changes in [the marginal/ univariate distributions of] extreme events themselves such as
@@
 influence the coexistence of species through the storage effect and not other mechanisms, but
-extreme and changing climate is probable to affect other mechanims in other ways (refs?).
+extreme and changing climate is probable to affect other mechanims in other ways.
 %In the Discussion, I intend to talk about how possible changes in tail association 
@@
 framework, is not directly a measure of the environment, but is a measure of response to the 
-environment, via a population's vital rates. Environment and environmental response may be related linearly
+environment, measured by a population's vital rates. Environment and environmental response may be related linearly
 (like lottery model?); though real environmental variables are probable to 
@@
 environmental variables merit future study, but the relationships between populations'
-vital rates should also be investigated for species coexistence and the maintenance of biodiversity.
+vital rates should also be investigated for species coexistence and the maintenacne of biodiversity.
 %Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
@@
 small-variance approximations completely miss out
-on any additions to coexistence beyond simple correlations and does not include our new ATA term, $\Delta^{[EC]}$, or variance \emph{per se}, $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
+on any additions to coexistence beyond simple correlations and does not include our new term, $\Delta^{[EC]}$ or $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
 In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
@@
 the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, 
-obscuring the effects of ATAs. Because standard MCT's assumption make it impossible to 
+obscuring the effects of ATAs. However, because standard MCT's assumption make it impossible to 
 detect interactions outide of correlation \emph{per se}, simulation-based MCT is required to
-detect ATA contributions, and our new extension to it clarifies and cleary separates and quantifies the 
-relative importance of both ATAs and pure correlation in contribution to the storage effect
+detect ATA contributions, and our new extension to it cleary separates and quantifies the 
+relative contribution and importance of ATAs and pure correlation to the storage effect
 and subsequently species coexistence and the maintenance of biodiversity.
@@
 ATA contributions and correlation \emph{per se} can be decomposed from the \emph{associations} between two variables
-among many from the appropriate term with parenthesized superscripts containing the $\shp$ symbol which denotes and separates 
+among many from the appropriate term with parenthesized superscripts with the $\shp$ symbol denoting and separating 
 variables that are varying independently from the variables that are covarying whose relationship we are interested in decomposing. 
@@
 Moreover, because our empirical application makes use of a simple and controlled lab system, future studies
-should be done to assess the extent of how much ATAs are important for complex systems and field systems which are 
-often subject to factors beyond our control.
+should be done to assess the extent of how much ATA are important for complex systems and field systems.
 Because modeling the dynamics of long-lived populations can be challenging, a good field system to start
 exploring the ideas presented here is a system of annual plants. 
-For example, the ATAs effcts of rainfall timing on the coexistence of grasses and forbs 
+For example, the ATAs between rainfall timing on the coexistence of grasses and forbs 
 in California grasslands \citep{Hallett_2019} could be a fruitful 
@@
 coexistence and exclusion of \emph{F. crotonensis}. 
-Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
+Standard errors of plotted quantities were always less then X, so are not shown.
+%***DAN: Don't forget to fill in the standard error stuff, please, Jasmin.
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
@@
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
-excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
-Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
+excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.} \label{fig:fig6}
 \end{figure}

### 2022-04-21 paper/Paper.Rnw

--- 2022-04-14 paper/Paper.Rnw
@@
 actually be related to vital rates in a non-linear and often threshold-like way. 
-Threshold relationships can produce tail association, as we showed elsewhere (explain, ref). 
+Threshold relationships can produce tail association, as we showed elsewhere \citep{Walter_2022}. 
 Changes in the means of true environmental variables may change 

### 2022-04-21 paper/Paper.Rnw
added paragraph on ATA specialists

@@
 Contributions of ATAs to species coexistence open the possibility that
-``ATA specialists'' may exist in some contexts. If a stronger competitor is present across a
-landscape and the GWR of a weaker competitor %***DAN: Dan is working right here.
-
-%Not only is the contribution of ATA non-neglible but in some cases the presence or absence of ATAs 
-%can mediate the result of competition between species, i.e., coexistence or extirpation. 
-%For some parameters, the contribution of ATAs was the driving mechanism of 
-%coexistence -- GWR would be negative without it -- deeming the scenario a case of 
-%ATA rescue. For other parameter values, ATA exclusion occurs where ATAs are the 
-%driving mechanism of the failure to coexist (exclusion) and the lack of ATAs would result in 
-%a positive GWR. Parameter
-subspaces of ATA exclusion and ATA rescue suggest a new kind of specialization among species or \emph{ATA specialists} - species which only persist in habitat patches where ATAs contribute to its coexistence.
-One can imagine a species whose coexistence with others is conditional on the asymmetric relationship between drivers of coexistence and whose populations may become excluded when ATAs erode. 
-%I think ata specialists are a thing - if their competitor is ubiquitous, they can only live in places on the landscape 
-%where there are atas. Could have a discussion para on it, if it does not make the results. Or both. BTW, ATA exclusion has a 
-%similar thing going on - a species excluded by ATAs (and competition) can only live on the portions of the landscape where
-%there are no ATAs. Could also describe the temporal way of it - the ATA specialist can only survive in a era/epoch for which there
-%are ATAs in its local, otherwise it'll have to go dormant or move elsewhere.
+``ATA specialists'' may exist in some contexts. If a strong competitor is present across a
+network of habitat patches and the GWR of a weaker competitor is only positive 
+in a patch in the presence of 
+ATAs between the temporal fluctuations of two environmental variables in that patch,
+then the weaker competitor can only persist in patches for which ATAs are present,
+and can therefore be regarded as an ``ATA specialist'' in the landscape. EC-associations
+can often boil down to associations between two environmental variables, as was the case
+for the lottery model we considered. Mechanisms of coexitence should correspond to
+means by which niche differentiation is possible. Thinking about our new mechanism of 
+coexistence in terms of possible ATA specialists helps highlight this connection. Likewise,
+because we have described ATA exclusion as well as ATA rescue, it is possible to 
+imagine a scenario in which a weaker competitor can only exist in habitat patches 
+for which ATAs are absent, and is hence a specialist on scenarios with symmetric
+tail associations of environmental variables (an ``STA-specialist''). Additional 
+research will be needed to assess to what degree these possibilities are realized 
+in nature.
 

### 2022-04-21 paper/Paper.Rnw
made some progress on the Discussion

@@
 
+%Summary-type paragraph of main results
 We produced and applied a new extension of MCT that 
@@
 for which ATAs are absent, and is hence a specialist on scenarios with symmetric
-tail associations of environmental variables (an ``STA-specialist''). Additional 
+tail associations of environmental variables (an ``STA specialist''). Additional 
 research will be needed to assess to what degree these possibilities are realized 
@@
 %future work: study ATAs in environmental variables and how theyre changing - tie in with specialist
-Our study highlights the importance of ATAs for ecological phenomena 
-and specifically for coexistence and the maintenance of biodiversity. 
-The occurence of ATAs among environmental variables themselves, 
-their contribution to coexistence, and their changes through space and time
-can and should be explored further. 
-The coupling of ATA contribution magnitude with environmental variability, 
-as shown in our results and figures (Fig. \ref{fig:fig3} - \ref{fig:fig6}), 
-suggests that ATAs and their effects
-may become more important to consider for coexistence and possibly other 
-ecological phenomena as environmental conditions may evidently
-become more variable (probably need refs).
-As described in section \ref{section:introduction}, ATAs between environmental
-variables may shift as the climate becomes more variable and extreme events become more 
-expansive and extensive, causing associations of distinct temporal events to become 
-more related in one extreme or tail, thus asymmetric. Additionally, 
-if ATA specialist exist, their persistance may
-become at stake as the mechanism driving their coexistence changes with the climate.
-Thus it is of interest to further study how tail associations of environmental variables are changing and how this might 
-alter competition, coexistence, and the maintenance of biodiversity generally.
-Our investigation brings to attention the consequences of such
-possible [bivariate] \emph{relationships} in the extremes; however, 
-simpler changes in [the marginal/ univariate distributions of] extreme events themselves such as
-intensity and frequency and their influence on coexistence also merit future work. 
-Possible changes in tail association
-patterns between environmental variables in nature can
-influence the coexistence of species through the storage effect and not other mechanisms, but
-extreme and changing climate is probable to affect other mechanisms in other ways (refs?).
-%In the Discussion, I intend to talk about how possible changes in tail association 
-%patterns of env vars in nature may influence coexistence of species. It won't influence the 
-%other mechanisms, though changes in the extreme values of environmental variables, known
-%to be occurring, might. 
-%Should also mention this second point in the Discussion as something that merits further research.
-%though might be worth mentioning in the Discussion that events are getting more extreme. We have studied how *relationships*
-%in the extremes influence coexistence, but probably worth studying (or cite papers if it's been done) how changes
-%in the commonness and extremeness of extreme events themselves influence coexistence, and for that you'd have
-%to also look at the other mechanisms. May still need a sup mat laying out all the math.
+In the Introduction we explained a plausible means by which climate change may
+alter the tail association properties of phenologically separated environmental variables
+such as early- and late-spring rainfall if, as seems likely, meteological events which
+are more extreme also tend to be longer lasting. The possibility 
+that changes in the nature of environmental ATAs 
+may result in the extirpation of ATA and STA specialists indicates that it may
+be important to engage in further study of ATAs between environmental variables 
+and how they differ across space and may be changing through time due to human
+impacts on the climate. Extreme environmental events have been much studied, including
+how the nature of such events may be changing with climate change, and how such changes 
+may influence species coexistence (see citations in the Introduction). But
+studies of changes in environmental ATAs would differ from such prior work because
+ATAs pertain to relationships between variables in the extremes, a potential changes in those
+relationships, to the best of our knowledge, have not yet been examined.
 
-%Although environmental fluctuations is helpful for coexistence and storage effect, unusual extreme fluctuations can be a hinderance on coexistence by introducing or changing ATAs between environmental variables.
-%onsistent with the mechanisms of biodiversity maintenance and specifically 
-%luctuation-dependent mechanisms (cite chesson), higher values of environmental fluctuation 
-%esulted in higher IGR’s. Both of our new terms, $\Delta_i^{[EC]}$ and $\Delta_i^{[E \nat C]}$ 
-%lso increased in magnitude with increased environmental fluctuation. Increased environmental 
-%luctuation is reflected in increased $\sigma$ in the lottery model and increased $a$ in the 
-%ompetitive diatom model.   mean temp and period?
+%E and vital rates
+Whereas $E$ of the $EC$ associatons which produce storage effects 
+is often conceptualized as ``environment'', in fact it is a 
+species response to the environment. Because a directly measurable aspect of the 
+environment and a species response to it may be nonlinearly related, it may be 
+important, in future work, 
+to also study the nature of tail associations between species vital rates.
+In the lottery model, it was assumed that fecundity of species $i$ tracks the
+environment directly, $B_i(t)=E_i(t)$. But vital rates such as fecundity are 
+more likely related to actual environmental variables according to a nonlinear
+function which may often be threshold-like or sigmoidal, and thus the $E_i=B_i$
+of the lottery model is actually better conceptualized as a species response
+than as the environment, directly.
+Likewise, one of the choices of $E_i$ considered by \cite{Ellner_2016} for the
+diatom model was $E_i=V_i$, the maximum population growth rate, clearly a response
+to the environment rather than a direct measurement of the environment. 
+It was demonstrated elsewhere that nonlinear transformations can produce or alter 
+asymmetries of tail association \citep{Ghosh_2020_AER,Walter_2022}.
+Thus the nature of ATAs between directly measured environmental variables may differ from the
+nature of ATAs between the population quantities which actually matter for species coexistence.
+Vital rates are often measured directly in demongraphic studies. To our knowledge, the 
+nature of tail associations between vital rates has been studied seldom or never.
 
-%This would be a whole potential 
-%research program which presumably would start with a more detailed study of how tail associations of environmental
-%variables may be changing under global change.
+%Differences with the original, analytic version of the MCT
+%
+%***DAN: I am not so comfortable with this because I have insufficient 
+%knowledge of the original MCT. I also do not want to devote the time
+%to learning it right now because of other commitments. I've reworded, 
+%trying not to change the meaning. Jasmin, how confident are you in 
+%these statements? We will also want to try to get Steve Ellner to check 
+%them.
+Our approach to exploring the influence of ATAs on coexistence relies on the computational
+extension of the MCT of \cite{Ellner_2019},
+and our results help demonstrate strengths of that approach.
+Storage effects as computed using the original, analytic appoach to MCT 
+\citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
+which we refer to as the contributions to coexistence of correlation \emph{per se}.
+This is because the weak-noise/Taylor approximation approach used in the original 
+MCT neglects the effects of ATAs when higher-order terms
+in Taylor expansions are dropped. The term $\Delta^{(E \shp C)}$, which our theory 
+inherits from \cite{Ellner_2016}, is also dropped. These observations do not diminish 
+the major strengths of the 
+original, analytic appproach, which are well known and also elaborated by \cite{Ellner_2019};
+one such strength is the generality that analtic approaches provide.
+It may be possible to extend the mathematics of the original MCT to consider ATAs.
 
-%\underline{E and vital rates}
-We note that \emph{E}, in our investigation and in other investigations within the MCT 
-framework, is not directly a measure of the environment, but is a measure of response to the 
-environment, via a population's vital rates. Environment and environmental response may be related linearly
-(like lottery model?); though real environmental variables are probable to 
-actually be related to vital rates in a non-linear and often threshold-like way. 
-Threshold relationships can produce tail association, as we showed elsewhere \citep{Walter_2022}. 
-Changes in the means of true environmental variables may change 
-where those distributions of environmental variables are compared to thresholds, 
-which may change the nature of tail associations between the effects of 
-those environmental variables. Thus, not only do the possible asymmteric association between 
-environmental variables merit future study, but the relationships between populations'
-vital rates should also be investigated for species coexistence and the maintenance of biodiversity.
-%Discussion content: We should (in future) study tail associations between environmental variables, and how these might be changing.
-However, more than that, the variables E in the lottery model are not actual environmental variables but rather are the effects of environmental variables on the species of the model.
+%multispecies models and multifactor models
+\cite{Ellner_2019} elaborated a general theory in which more and different varying factors than
+$E$ and $C$ can be considered and used to produce a decomposition of GWR. For instance, if $X$, $Y$ and 
+$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, the decomposition
+\begin{align}
+\text{GWR}_i &= \Delta_i^0+\Delta_i^X+\Delta_i^Y+\Delta_i^Z \label{eq:9}\\
+&+\Delta_i^{(X \shp Y)} +\Delta_i^{(XY)} + \Delta_i^{(Y \shp Z)} + \Delta_i^{(YZ)} +\Delta_i^{(X \shp Z)}+\Delta_i^{(XZ)} \label{eq:10}\\
+&+ \Delta_i^{(XYZ)}+\Delta_i^{(XY \shp Z)}+\Delta_i^{(X \shp YZ)}+\Delta_i^{(XZ \shp Y)}+\Delta_i^{(X \shp Y \shp Z)}.
+\end{align}
+The terms in (\ref{eq:9}) and (\ref{eq:10}) are analogous to those of (\ref{eq:previous_decomp}).
+The new terms result from rendering one or more of $X$, $Y$, $Z$ independent of the others.
+We simply comment that this general decomposition should also be extendable to consider the effects of
+ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}), though the complexity
+of such a decomposition may present some difficulties.
+For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
+It seems reasonable to suppose that ATAs will play an increasingly important role in systems of 
+greater complexity because such systems have more pairs of temporally variable quantities that
+may exhibit asymmetries of association. Our prior work shows that ATAs are common in 
+ecological and environmental variables \citep{Ghosh_2020_AER}.
 
-%\underline{the new effects are not "weak noise" effects}
-%if sigma really small, ATA would not matter?
-%go back to emails
-%Maybe have a paragraph about how Chesson's weak noise and other assumptions make it impossible to detect this new effect
-%via that framework. It's only the less-assumption-reliant simulation approach of Ellner et al that makes it possible to 
-%study this new term in the decomp. Need to study a bit to substantiate this properly, or get Steve Ellner help.
-We parse the storage effect, providing a tertiary decomposition to the simulation-based 
-approach of MCT \citep{Ellner_2016, Ellner_2019}, 
-a new approach which bypasses the mathematical assumptions and obstructions of 
-standard MCT. Classical analytic derivations of MCT using second order Taylor expansion and 
-small-variance approximations completely miss out
-on any additions to coexistence beyond simple correlations and does not include our new ATA term, $\Delta^{[EC]}$, or variance \emph{per se}, $\Delta^{(E \shp C)}$ \citep{Chesson_1994, Chesson_2000}. 
-In fact, storage effects in analytic MCT is computationally equal to $\Delta^{[E \nat C]}$, 
-the contributions of correlation \emph{per se}. 
-In newer, simulation based MCT, storage effect is 
-the sum of the two new terms, $\Delta^{(EC)}=\Delta^{[E \nat C]}+\Delta^{[EC]}$, 
-obscuring the effects of ATAs. Because standard MCT's assumption make it impossible to 
-detect interactions outide of correlation \emph{per se}, simulation-based MCT is required to
-detect ATA contributions, and our new extension to it clarifies and cleary separates and quantifies the 
-relative importance of both ATAs and pure correlation in contribution to the storage effect
-and subsequently species coexistence and the maintenance of biodiversity.
-
-%\underline{multispecies models}
-%Need a para about multi-species models - how would things change? One difference is there are more different
-%things that might have asymmetric tail association patterns. See the section "General Theory" of Ellner et al 2019 - might
-%just want write a short para saying it seems like it should be possible to generalize what we've done along those lines.
-
-%Because asymmetric tail associations exist in many ecological variables \citep{Ghosh_2020_AER}, the more specified approach of considering unique drivers of coexistence rather than E and C and the ability to consider more than two drivers, could produce interesting results??
-
-%The inclusion of multiple species introduces a multitude of relationships between species'
-%environmentally dependent parameters that could possibly be asymmetrically associated. 
-%The greater the number of species the greater the possibilities of ATAs being more important 
-%to their coexistence.
-
-The chances of ATAs being important for the coexistence of species increases as more variables are included in the system, 
-either through the addition of more species or more drivers (beyond just E and C), or both. 
-When multiple species are included in the system, the interaction between different species' E's and C's introduces more relationships 
-to consider that may or may not be asymmetrically related and could increase the relative importance of ATA contribution to the species' coexistence. 
-Likewise, coexistence between species can also depend on more than two drivers \citep{Ellner_2019}. 
-The storage effect decomposition presented here is generalizable to coexistence decompositions of multiple variables.
-Following the set up and notation of \cite{Ellner_2019}'s \emph{covariance decomposition}, 
-ATA contributions and correlation \emph{per se} can be decomposed from the \emph{associations} between two variables
-among many from the appropriate term with parenthesized superscripts containing the $\shp$ symbol which denotes and separates 
-variables that are varying independently from the variables that are covarying whose relationship we are interested in decomposing. 
-For example, the term $\Delta^{(XY \shp Z)}$ could be decomposed into $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$. 
-As more terms and species are introduced into a system, more relationships between variables must be evaluated,
-more of those relationships may be asymmetrically tail associated, and ATAs may contribute more to the maintenance of 
-biodiversity in that system. 
-%make three different biv dis that are standard normal copula and then pair with a shuffles third
-%use to get rbar and subtract from rbar of XY#Z,
+%***DAN: stopped here 2022 04 21
 

### 2022-04-21 paper/Paper.Rnw
finished editing the Discussion

@@
 \author[a]{Pimsupa Jasmin Albert}
-\author[a,b,*]{Daniel C. Reuman}
+\author[b]{Stephen P. Ellner}
+\author[a,c,*]{Daniel C. Reuman}
 
 \affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
-\affil[b]{Laboratory of Populations, Rockefeller University}
+\affil[b]{Department of Ecology and Evolutionary Biology, Cornell University}
+\affil[c]{Laboratory of Populations, Rockefeller University}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 ATAs of environmental variables and thereby influence competition and species coexistence.
-%***DAN: Make sure we end up having such material in the Discussion
 Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
@@
 ``ATA specialists'' may exist in some contexts. If a strong competitor is present across a
-network of habitat patches and the GWR of a weaker competitor is only positive 
+collection of habitat patches and the GWR of a weaker competitor is only positive 
 in a patch in the presence of 
 ATAs between the temporal fluctuations of two environmental variables in that patch,
-then the weaker competitor can only persist in patches for which ATAs are present,
-and can therefore be regarded as an ``ATA specialist'' in the landscape. EC-associations
-can often boil down to associations between two environmental variables, as was the case
-for the lottery model we considered. Mechanisms of coexitence should correspond to
+then the weaker competitor can only persist in patches for which ATAs are present.
+The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
+can often boil down to associations between two environmental variables, as was assumed to be the case
+for the lottery model. Mechanisms of coexitence should correspond to
 means by which niche differentiation is possible. Thinking about our new mechanism of 
@@
 alter the tail association properties of phenologically separated environmental variables
-such as early- and late-spring rainfall if, as seems likely, meteological events which
-are more extreme also tend to be longer lasting. The possibility 
+such as early- and late-spring rainfall. The possibility 
 that changes in the nature of environmental ATAs 
@@
 be important to engage in further study of ATAs between environmental variables 
-and how they differ across space and may be changing through time due to human
-impacts on the climate. Extreme environmental events have been much studied, including
-how the nature of such events may be changing with climate change, and how such changes 
-may influence species coexistence (see citations in the Introduction). But
+and how they differ across space and may be changing through time. 
+Extreme environmental events have been much studied, including
+how the strength and commonness of such events may be changing with climate change, and how such changes 
+may influence coexistence (see citations in the Introduction). But
 studies of changes in environmental ATAs would differ from such prior work because
-ATAs pertain to relationships between variables in the extremes, a potential changes in those
-relationships, to the best of our knowledge, have not yet been examined.
+ATAs pertain to \emph{relationships} between variables in the extremes.
 
 %E and vital rates
-Whereas $E$ of the $EC$ associatons which produce storage effects 
-is often conceptualized as ``environment'', in fact it is a 
+Whereas the ``$E$'' of ``$EC$-covariance'' is often conceptualized as ``environment'', in fact it is a 
 species response to the environment. Because a directly measurable aspect of the 
@@
 diatom model was $E_i=V_i$, the maximum population growth rate, clearly a response
-to the environment rather than a direct measurement of the environment. 
+to the environment rather than a direct environmental measurement. 
 It was demonstrated elsewhere that nonlinear transformations can produce or alter 
@@
 Thus the nature of ATAs between directly measured environmental variables may differ from the
-nature of ATAs between the population quantities which actually matter for species coexistence.
-Vital rates are often measured directly in demongraphic studies. To our knowledge, the 
-nature of tail associations between vital rates has been studied seldom or never.
+nature of ATAs between the population responses which actually matter for coexistence.
+Vital rates are often measured directly in demongraphic studies, so ATAs between vital rates could,
+in principle, be studied. To our knowledge, this has not yet been done.
 
@@
 The terms in (\ref{eq:9}) and (\ref{eq:10}) are analogous to those of (\ref{eq:previous_decomp}).
-The new terms result from rendering one or more of $X$, $Y$, $Z$ independent of the others.
+The new terms result from rendering one or more of $X$, $Y$, and $Z$ independent of the others.
 We simply comment that this general decomposition should also be extendable to consider the effects of
 ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}), though the complexity
-of such a decomposition may present some difficulties.
+of such a decomposition may present difficulties.
 For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
@@
 may exhibit asymmetries of association. Our prior work shows that ATAs are common in 
-ecological and environmental variables \citep{Ghosh_2020_AER}.
-
-%***DAN: stopped here 2022 04 21
-
-%\underline{caveats and future directions}
-%Realism - we usede simple/extreme ATA, simple model and simple lab system, futurw work needed to assess 
-%how mcuh ATA are important in complex systems, field systems
-%what systems? what env vars?
-%Ellner et al (2016) points out that coexistence is also mediated by other factors that competition: shared enemies, 
-%mutualist, facilitation etc. (these are the exact factors Ellner lists), so future work could explore 
-%the importance of ATA for coexistence mediated by these factors.
-%Something about spatial mechanisms of coexistence? We considered only mechanisms that applied in a model with
-%solely temporal variation, but there are also spatial mechanisms. Need to learn about those and think about
-%how to say something about how future work may be able to explore the importance of ATA in that context?
-
-%Also need to talk about how this stuff will apply to data.
-As an introductory investigation of ATAs within the framework of MCT, 
-we used simple and extreme examples, and future work should assess more complex relationships, systems, and models. 
-In our theoretical exploration, we implemented a simple lottery model and assessed 
-the influence of extremely asymmetric associated noise, i.e., the noise was perfectly correlated
-in exactly one half of the distribution. However, more complex models (such as structured demographic models
-or integral projection models should be explored )and varying degrees and qualities of ATA, (copula parameters, 
-difference between tails, rate of relatedness(slope of expansion??) should be implemented and explored.
-Moreover, because our empirical application makes use of a simple and controlled lab system, future studies
-should be done to assess the extent of how much ATAs are important for complex systems and field systems which are 
-often subject to factors beyond our control.
-Because modeling the dynamics of long-lived populations can be challenging, a good field system to start
-exploring the ideas presented here is a system of annual plants. 
-For example, the ATAs effects of rainfall timing on the coexistence of grasses and forbs 
-in California grasslands \citep{Hallett_2019} could be a fruitful 
-further investigation of biodiversity maintenance in California grasslands and the effects of extreme 
-rainfall fluctuation given the highly variable climate of the area.
-Moreover, our approach may provide more insight when expanded to consider fluctuations across space,
-in regards to questions about population synchrony, Moran effects, and the suggestion of possible ATA specialists.
-
-%Delta=1 means no overlapping generations -  breaking the storage effect requirement of buffered population growth.
-
-%\underline{conclude **NOT NEEDED**: how ur work answers/informs the broad question from the intro and lay out 2-3 key priorities for next research on this question}
-%We provided another showcase of the importance of ATAs in ecology, 
-%building on previous studies.
-%ATAs clearly have implications for coexistence and competition;
-%though more research is needed in exploring more complex models 
-%including spatiotemporal and multispecies models.
-%Hope to see this tertiary decomposition of MCT be applied to empirical research.
-%More investigations should be done to explore implications of ATAs 
-%to other ecological theories.
-
-%Search for comments in the rest of the text indicating things that need to be done in the Discussion
-
-%Knowledge of stochastic structure (Shoemaker et al 2020)
+ecological and environmental variables \citep{Ghosh_2020_AER}. We have here used simple mathematical
+and laboratory models to provide an initial exploration of the potential importance of ATAs
+for coexistence. Future work should attempt to generalize lessons learned here to 
+field systems with multiple species, stage structure, spatial extent, non-competitive interactions 
+such as predation and facilitation that can also mediate coexistence, and other complexities.
 
@@
 The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
-Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Steve Ellner, and Robin Snyder for helpful discussions. 
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
 The authors were partly supported by US National
@@
 Wildlife Delta Science Program. 
-%***DAN: Jasmim, please add any other acknolwedgments here - anyone else you have discussed this with or anything like that.
-%Be generous here, add people even if they only helped a tiny bit. If Steve consents to be coauthor, we take him
-%off this list and add him to the author list.
-%***DAN: DAN: if we add Ellner to the author list take him out here, and also take him out of the acks in the README.
-
-\clearpage
-\newpage
+%***DAN: DAN: if we take Ellner off the author list add him back here, and also add him back to the acks in the README.
 
@@
 
-%***DAN: Jasmin, some changes to this figure are needed. Please make axis labels and numbers bigger. Please put 
-%axis labels outside the axes instead of inside. Please make panel labels bigger. Please make coutour labels
-%slightly bigger.
-%***DAN: Do we want to do anything with standard error for this fig?
 \begin{figure}

### 2022-04-27 paper/Paper.Rnw
spelling

@@
 according to (\ref{eq:previous_decomp}) for these same values of
-$\theta_0$, $a$ and $P$. We further decomponse
+$\theta_0$, $a$ and $P$. We further decompose
 coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
@@
 can often boil down to associations between two environmental variables, as was assumed to be the case
-for the lottery model. Mechanisms of coexitence should correspond to
+for the lottery model. Mechanisms of coexistence should correspond to
 means by which niche differentiation is possible. Thinking about our new mechanism of 
@@
 nature of ATAs between the population responses which actually matter for coexistence.
-Vital rates are often measured directly in demongraphic studies, so ATAs between vital rates could,
+Vital rates are often measured directly in demographic studies, so ATAs between vital rates could,
 in principle, be studied. To our knowledge, this has not yet been done.
@@
 and our results help demonstrate strengths of that approach.
-Storage effects as computed using the original, analytic appoach to MCT 
+Storage effects as computed using the original, analytic approach to MCT 
 \citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
@@
 the major strengths of the 
-original, analytic appproach, which are well known and also elaborated by \cite{Ellner_2019};
-one such strength is the generality that analtic approaches provide.
+original, analytic approach, which are well known and also elaborated by \cite{Ellner_2019};
+one such strength is the generality that analytic approaches provide.
 It may be possible to extend the mathematics of the original MCT to consider ATAs.

### 2022-04-28 paper/Paper.Rnw
minor tweaks

@@
 %E and vital rates
-Whereas the ``$E$'' of ``$EC$-covariance'' is often conceptualized as ``environment'', in fact it is a 
+Whereas the ``$E$'' of ``$EC$-covariance'' is often conceptualized as ``environment'', 
+in fact it is often instead a 
 species response to the environment. Because a directly measurable aspect of the 
@@
 function which may often be threshold-like or sigmoidal, and thus the $E_i=B_i$
-of the lottery model is actually better conceptualized as a species response
-than as the environment, directly.
+of the lottery model is probably better conceptualized as a species response.
 Likewise, one of the choices of $E_i$ considered by \cite{Ellner_2016} for the
-diatom model was $E_i=V_i$, the maximum population growth rate, clearly a response
+diatom model was $E_i=V_i$, the maximum population growth rate, a response
 to the environment rather than a direct environmental measurement. 
@@
 asymmetries of tail association \citep{Ghosh_2020_AER,Walter_2022}.
-Thus the nature of ATAs between directly measured environmental variables may differ from the
-nature of ATAs between the population responses which actually matter for coexistence.
-Vital rates are often measured directly in demographic studies, so ATAs between vital rates could,
-in principle, be studied. To our knowledge, this has not yet been done.
+So the nature of ATAs between directly measured environmental variables may differ from the
+nature of ATAs between the population responses which matter for coexistence.
+Vital rates are measured directly in demographic studies, so ATAs between vital rates could,
+in principle, be studied. To our knowledge, this has not been done.
 
@@
 &+\Delta_i^{(X \shp Y)} +\Delta_i^{(XY)} + \Delta_i^{(Y \shp Z)} + \Delta_i^{(YZ)} +\Delta_i^{(X \shp Z)}+\Delta_i^{(XZ)} \label{eq:10}\\
-&+ \Delta_i^{(XYZ)}+\Delta_i^{(XY \shp Z)}+\Delta_i^{(X \shp YZ)}+\Delta_i^{(XZ \shp Y)}+\Delta_i^{(X \shp Y \shp Z)}.
+&+ \Delta_i^{(XYZ)}+\Delta_i^{(XY \shp Z)}+\Delta_i^{(X \shp YZ)}+\Delta_i^{(XZ \shp Y)}+\Delta_i^{(X \shp Y \shp Z)}.\label{eq:11}
 \end{align}
 The terms in (\ref{eq:9}) and (\ref{eq:10}) are analogous to those of (\ref{eq:previous_decomp}).
-The new terms result from rendering one or more of $X$, $Y$, and $Z$ independent of the others.
+The new terms in (\ref{eq:11}) result from rendering one or 
+more of $X$, $Y$, and $Z$ independent of the others.
 We simply comment that this general decomposition should also be extendable to consider the effects of
 ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}), though the complexity
-of such a decomposition may present difficulties.
+of such a decomposition may present challenges.
 For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
@@
 Wildlife Delta Science Program. 
-%***DAN: DAN: if we take Ellner off the author list add him back here, and also add him back to the acks in the README.
+%***DAN: DAN: if we take Ellner off the author list add him back here, and also add him back to the acks in the README. Also take him off the author list in sup mat.
 

### 2022-05-15 paper/Paper.Rnw
line numbering
some small edits

@@
 
+%for dealing with line numbers bugs having to do with equations
+\let\oldequation\equation
+\let\oldendequation\endequation
+\renewenvironment{equation}
+  {\linenomathNonumbers\oldequation}
+  {\oldendequation\endlinenomath}
+\let\oldalign\align
+\let\oldendalign\endalign
+\renewenvironment{align}
+  {\linenomathNonumbers\oldalign}
+  {\oldendalign\endlinenomath}
+
 \begin{document}
@@
 
+\linenumbers
 \maketitle
@@
 Every ecological discipline involves the study of relationships among biological and environmental variables,
-but standard statistical methods for examining the association between two variables provide only a small 
-portion of the overall information content of the
+but standard statistical methods for examining the association between two variables provide only 
+%a small portion of the overall information content of the
+partial information about the nature of the 
 association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
@@
 what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
-Storage effects allow for species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
+Storage effects allow competing species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
 whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
@@
 covariation between the fluctuating quality of the environment (E)
-for a species and the strength of competition (C) experienced by that species (called EC covariance). 
+for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is common than when it is rare. 
 Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
@@
 coexistence and exclusion of \emph{F. crotonensis}. 
-Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
+Standard errors of plotted quantities were always less than \Sexpr{round(se5,5)}, so are not shown.
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
@@
 excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
-Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
+Standard errors of plotted quantities were always less than \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
 \end{figure}

### 2022-07-10 paper/Paper.Rnw
START OF INCORPORATING ELLNER EDITS IS PREVIOUS COMMIT ON 5/15..continuing here

@@
 
-\DeclareMathOperator{\E}{E}% expected value
+\DeclareMathOperator{\E}{\mathbb{E}}% expected value
 \DeclareMathOperator{\var}{var}
@@
 Every ecological discipline involves the study of relationships among biological and environmental variables,
-but standard statistical methods for examining the association between two variables provide only 
+but standard %statistical methods 
+measures for %examining the 
+measuring the strength and direction of the association between two variables provide only 
 %a small portion of the overall information content of the
@@
 what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
-Storage effects allow competing species to fluctuate over time while escaping extinction because storage effects provide a relative benefit to
-whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
+Storage effects allow competing species to fluctuate %over time while escaping extinction because storage effects provide 
+in abundance without any going extinct by providing
+a relative benefit to whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
 of coexistence \citep{Chesson_2000}.
@@
 Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
-same environmental conditions to be optimal. Third, a buffering mechanism must exist whereby species limit the influence of competition
-in times when the environment is poor. Dormant stages such as seed banks are an
-example buffering mechanism \citep{Chesson_2012}. 
+same environmental conditions to be optimal. 
+Third, a buffering mechanism %must exist whereby species limit 
+which limits the influence of competition %in times 
+when the environment is poor. 
+Dormant stages such as seed banks are an example buffering mechanism \citep{Chesson_2012}. 
 When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
@@
 %Lottery model and results based on it
-To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
-simulations using a simple two-species model, the lottery model. The simulations described in this
-paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
-precise nature of that influence is then the subject of the rest of the paper.
-The lottery model is as follows.
+To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model. 
+The simulations described in this paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; 
+the precise nature of that influence is then the subject of the rest of the paper.
+The lottery model was originally introduced as a model for coral reef fish competing for
+a set of $N$ breeding territories.
 Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
@@
 for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
-The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+The model postulates that adults die at rate $\delta$ at each time step, and are replaced by new juveniles 
 in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
@@
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
-presentation of new ideas.
+$C_i(t)$. Though the iid assumption %can be circumvented
+is not neccesary, %we make the assumption because 
+it simplifies the presentation of new ideas.
 
@@
 Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
-Though some readers may find it slightly confusing to simultaneously use the variable 
-$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would 
-be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
-all operators, $\E(\cdot)$ is rendered upright.
+%Though some readers may find it slightly confusing to simultaneously use the variable $E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would be more confusing to use alternative notation.
+Like all variables, $E_i$ is rendered in italic; like all operators, $\E(\cdot)$ is rendered upright.
 
-We now specify how to decompose storage effects into
-contributions due to ATAs and due
-to $EC$ correlation \emph{per se}. 
-Storage effects for species $i$ were computed by comparing
+We now specify how to decompose storage effects into contributions due to ATAs and 
+due to $EC$ correlation \emph{per se}. 
+Storage effect for species $i$ was computed by comparing
 differences between rare and common species mean growth rates for the original model and for a 
-modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
-were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
-Because these components were rendered 
-completely independent in the modified model, storage effects quantify the contribution of the 
-whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
-difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
+modified model for which 
+%the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
+$E$ and $C$
+were rendered independent of each other,
+for both rare and common species [see (\ref{eq:storage_effects_gen})]. 
+Because %these components 
+$E$ and $C$ 
+were rendered completely independent in the modified model, 
+storage effect quantifies the contribution of the whole $EC$ relationship, in all its aspects, 
+to the rare/common species growth rate difference. 
+But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
 $(E_j^\nat,C_{j \bs i}^\nat)$ with the same
@@
 and with some but not all aspects of the relationship 
-between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
-$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
+between %the components 
+$E$ and $C$
+altered relative to the original distributions.
+
+Specifically, asymmetries of tail association are removed -- see below for how this is done. 
+We then define $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model 
+using  $(E_i^\nat,C_{i \bs i}^\nat)$
 and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 

### 2022-07-10 paper/Paper.Rnw
continuing to incirporate Ellner's edits

@@
 How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
-$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
+correlation unchanged? 
+We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, 
+respectively, are strictly monotonic and continuous. 
 Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
-$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
-distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
-these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$. %, which are standard normally
+These are standard Normal 
+distributed, although they are not necessarily jointly bivariate normal.  
+Let $\rho_i$ be the covariance of these random variables. 
+We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
 $\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
@@
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
-in one of 
-four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
-$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. 
+The value of $\ddot{\rho}_i$ is then determined in one of four alternative ways: 
+either by setting 
+$\ddot{\rho}_i = \rho_i$; 
+or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. 
+The choice of bivariate normal 
 $(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
+
+For those familiar with copula statistics 
+(see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 $C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
+as desired. 
+The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
 
 The other mechanisms in MCT that apply in our modelling context of purely temporal variation
-make no reference to covariation between quantities, whereas ATAs are 
-defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
+make no reference to covariation between quantities, 
+(see eqn. (13) in Ellner (2019),
+whereas ATAs are %defined in a covariation context
+strictly about covariation. 
+Therefore, storage effects are the only mechanism of coexistence
 for which ATAs are relevant. Mathematical details supporting this claim are in SI section 

### 2022-07-10 paper/Paper.Rnw
stopped after methods of G2

@@
 %G2
-Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
+To achieve goal G2 from the Introduction, 
+left- or right-tail associated %noise 
+log fecundities $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
 $\sigma$, was generated as pictured in Fig. \ref{fig:noise}, details in SI section \ref{SI-sect:noise}.
-To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
-to coexistence that come from ATAs versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
+%To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
+%to coexistence that come from ATAs versus other mechanisms, 
+We then plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
 $\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
-against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and we compared 
-values on plots. Some mathematical shortcuts, summarized here, simplified plotting. 
+against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$. %; and we compared values on plots. 
+Some mathematical shortcuts, summarized here, simplified plotting. 
 First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
@@
 lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we plotted 
+instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). 
+So we plotted 
+the terms
 against $\mu_1 - \mu_2$. 
 Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
-for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
-for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
-and $\mu_1-\mu_2=0,-0.5,-2,-4$,
+for both our left- and right-tail associated cases (SI section \ref{SI-sect:math_simplifications}), 
+so we only generated plots for left-tail associated noise. 
+For each combination of $\delta=0.2, 0.4, 0.6$
+and $\mu_1-\mu_2=0,-0.5,-1$,
 we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 

### 2022-07-10 paper/Paper.Rnw
making storage effects singular, as suggested by LMH

@@
 theory (MCT) identifies coexistence mechanisms,
-one of which, storage effects, hinges on relationships between fluctuations
+one of which, the storage effect, hinges on relationships between fluctuations
 in environmental and competitive pressures. However, such relationships
@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT, decomposing storage effects to show the influence of ATAs.
+We extend the MCT, decomposing the storage effect to show the influence of ATAs.
 Analysis of a simple model
@@
 
-\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effects}
+\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effect}
 
@@
 MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
-in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
+in competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
@@
 It is straightforward to see, intuitively, 
-why ATAs may influence one mechanism of coexistence in particular: storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
-Storage effects allow competing species to fluctuate %over time while escaping extinction because storage effects provide 
+why ATAs may influence one mechanism of coexistence in particular: the storage effect. We review, conceptually, 
+what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
+Storage effect allows competing species to fluctuate %over time while escaping extinction because storage effects provide 
 in abundance without any going extinct by providing
-a relative benefit to whichever species is currently rare; this is the characteristic that makes storage effects a so-called \emph{stabilizing} mechanism 
+a relative benefit to whichever species is currently rare; this is the characteristic that makes the storage effect a so-called \emph{stabilizing} mechanism 
 of coexistence \citep{Chesson_2000}.
-Storage effects can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
+Storage effect can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
 covariation between the fluctuating quality of the environment (E)
@@
 When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
-Importantly, storage effects can be quantified by measuring to what extent EC covariance contributes to differences between 
+Importantly, the storage effect can be quantified by measuring to what extent EC covariance contributes to differences between 
 rare and common species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
-Thus ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, and ATAs, when they 
+Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
 occur between environment and competition variables, can be an aspect of EC covariation.
@@
 multiple mechanisms to coexistence. But we show below
-that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-So we begin by defining storage effects. Our mathematical notation parallels
+that, for purely temporal variation, only one mechanism relates to tail associations: the storage effect. 
+So we begin by defining the storage effect. Our mathematical notation parallels
 that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
@@
 $C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
-Specifically, storage effects for species $i$ are 
+Specifically, storage effect for species $i$ is 
 \begin{equation}
@@
 
-We now specify how to decompose storage effects into contributions due to ATAs and 
+We now specify how to decompose storage effect into contributions due to ATAs and 
 due to $EC$ correlation \emph{per se}. 
@@
 \end{align}
-The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
+The term $\Delta_i^{[EC]}$ is the component of storage effect due to 
 ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
-The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
+The term $\Delta_i^{[E \nat C]}$ is the component of storage effect due to $EC$ 
 correlation \emph{per se}, remaining after having removed the effects of ATAs.
@@
 (\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
+of storage effect, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
 below for additional details.
@@
 strictly about covariation. 
-Therefore, storage effects are the only mechanism of coexistence
+Therefore, the storage effect is the only mechanism of coexistence
 for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
@@
 of covariation in these quantities; and
-$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
-Only storage effects involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
+$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- the storage effect. 
+Only storage effect involve $EC$ covariation,
+so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is the storage effect. Combining our decomposition 
 (\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
@@
 In the $\mu_1=\mu_2$ case (panels a-d), for which the two species were equal 
-competitors, storage effects (blue plus red
-lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
-tended to be negative, weakening total storage effects.
+competitors, storage effect (blue plus red
+lines on Fig. \ref{fig:fig3}) was the only means by which coexistence could occur, and ATA effects
+tended to be negative, weakening total storage effect.
 
@@
 Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
-we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
+we decomposed the storage effect into two new mechanisms, the contribution of ATAs to coexistence,
 and the contribution of $EC$ correlation \emph{per se}.
@@
 system of competing diatoms demonstrated that  
-ATA contributions to storage effects and to species coexistence and the maintenance of biodiversity
+ATA contributions to the storage effect and to species coexistence and the maintenance of biodiversity
 can often be comparable in magnitude to the contributions of previously recognized mechanisms.
@@
 ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
-Though storage effects are typically described as the contribution
+Though storage effect is typically described as the contribution
 of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
-our results suggest that storage effects should be redefined to constitute the 
+our results suggest that storage effect should be redefined to constitute the 
 contribution of \emph{associations} between $E$ and $C$, including both ATA effects and
@@
 and our results help demonstrate strengths of that approach.
-Storage effects as computed using the original, analytic approach to MCT 
+Storage effect as computed using the original, analytic approach to MCT 
 \citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 

### 2022-07-10 paper/Paper.Rnw
reorganizing ATA paragraph as suggeseted by LMH

@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT, decomposing the storage effect to show the influence of ATAs.
+We extend MCT, decomposing the storage effect to show the influence of ATAs.
 Analysis of a simple model
@@
 partial information about the nature of the 
-association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. For instance, 
-the three pairs of variables in Fig. \ref{fig:noise} all have the same
-Pearson correlation, but nevertheless demonstrate very different types of association. Other metrics 
-of association are also often used, in addition to correlations
+association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. 
+For instance, consider the three pairs of variables in Fig. \ref{fig:noise}.
+The variables in Fig. \ref{fig:noise}a,c exhibit what we have previously referred to as ``asymmetric tail association'' 
+[ATA; \cite{Ghosh_2020_AER}], i.e., they are more strongly related to each
+other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
+termed ``right-tail association'') portions of the variables' distributions. 
+The relationships displayed in Fig. \ref{fig:noise} demonstrate very different types of association, 
+but nevertheless all have the same Pearson correlation. 
+Other metrics of association are also often used, in addition to correlations 
 (e.g., various forms of regression), but correlations are extremely common, and other
 methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
-The variables in Fig. \ref{fig:noise}a,c 
-exhibit what we have previously referred to as ``asymmetric tail association'' [ATA; \cite{Ghosh_2020_AER}], i.e., 
-they are more strongly related to each
-other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
-termed ``right-tail association'') portions of the variables' distributions. Although ATAs 
-and related ideas had previously been seldom 
+Although ATAs and related ideas had previously been seldom 
 considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 

### 2022-07-10 paper/Paper.Rnw
some more edits

@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
-Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent extensions expand testability and accessibility \citep{Ellner_2016,Ellner_2019}.
 
@@
 %Lottery model and results based on it
-To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out simulations using a simple two-species model, the lottery model. 
+%To further demonstrate the plausibility of ATAs altering coexistence between species, 
+To more robustly demonstrate the role of ATAs in structuring fluctuation-dependent coexistence, we carried out simulations using a simple two-species model, the lottery model. 
 The simulations described in this paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; 

### 2022-07-12 paper/Paper.Rnw
switching paragraphs 1 and 2

@@
 
-%Intro to competition, coexistence, MCT, etc.
-Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
-\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
-The competitive exclusion principle states 
-that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
-the other and drive it extinct \citep{Hardin_1960}.
-Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
-framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
-in competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
-applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
-%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
-Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent extensions expand testability and accessibility \citep{Ellner_2016,Ellner_2019}.
-
 %Intro to ATAs
@@
 
+%Intro to competition, coexistence, MCT, etc.
+Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
+\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
+The competitive exclusion principle states 
+that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
+the other and drive it extinct \citep{Hardin_1960}.
+Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] %is a 
+has emerged as a
+framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
+MCT breaks down coexistence into the contributions of various mechanisms, %with names such as ``relative nonlinearityin competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers.
+highlighting the importance of environmental fluctuations to species coexistence. 
+For example, one such mechanism, the storage effect, promotes coexistence when species can reap or "store" the benefits of times when environmental conditions are good to overcome competition and persist through bad conditions.
+MCT has been applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
+%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent developments expand its testability and accessibility and present the opportunity for further extensions \citep{Ellner_2016,Ellner_2019}.
+
 %Intro to storage effects and why they might be influenced by ATAs
-It is straightforward to see, intuitively, 
-why ATAs may influence one mechanism of coexistence in particular: the storage effect. We review, conceptually, 
-what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
+In this paper, we show how ATAs can also be important for competition and for the coexistence of
+competing species. In particular, we show how ATAs can strongly influence one specific coexistence mechanism:
+storage effects.
+%It is straightforward to see, intuitively, why ATAs may influence one mechanism of coexistence in particular: the storage effect. 
+%We review, conceptually, what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
 Storage effect allows competing species to fluctuate %over time while escaping extinction because storage effects provide 

### 2022-07-12 paper/Paper.Rnw
changing inva/resi

@@
 covariation between the fluctuating quality of the environment (E)
-for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is common than when it is rare. 
+for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is at stable-state abundance than when it is rare. 
 Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
@@
 When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
-Importantly, the storage effect can be quantified by measuring to what extent EC covariance contributes to differences between 
-rare and common species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
+Critically, the storage effect can be quantified in invasion analysis by measuring to what extent EC covariance contributes to differences between 
+invader (rare) and resident (stable-state) species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
 Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
@@
 The storage effect for species $i$ is the contribution of covariation between $E$ and 
-$C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
+$C$ to a difference between invader and resident species mean growth rates, when $i$ is the invader.
 Specifically, storage effect for species $i$ is 
@@
 Storage effect for species $i$ was computed by comparing
-differences between rare and common species mean growth rates for the original model and for a 
+differences between invader and resident species mean growth rates for the original model and for a 
 modified model for which 
@@
 were rendered independent of each other,
-for both rare and common species [see (\ref{eq:storage_effects_gen})]. 
+for both invader and resideent species [see (\ref{eq:storage_effects_gen})]. 
 Because %these components 
@@
 storage effect quantifies the contribution of the whole $EC$ relationship, in all its aspects, 
-to the rare/common species growth rate difference. 
+to the invader/resident growth rate difference. 
 But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 

### 2022-07-12 paper/Paper.Rnw
stopping edits for now

@@
 \cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
-which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
+which equals GWR of species $i$. The decomposition, which we use below, is analogous to and motivated by the 
 original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
+which is the extension of MCT which was goal G1 of the Introduction. One typically %(but see below) 
+must examine this quantity for each species, $i$,
 to obtain complete information about coexistence. To pursue goal G2, below, we will quantify

### 2022-07-13 paper/Paper.Rnw
editts from this morniing - addressing comments from both LMH and SPE about clarity of the same paragraph in theory section

@@
 quantify the contributions of 
-multiple mechanisms to coexistence. But we show below
+multiple mechanisms to coexistence. But we show
 that, for purely temporal variation, only one mechanism relates to tail associations: the storage effect. 
@@
 
-The storage effect for species $i$ is the contribution of covariation between $E$ and 
-$C$ to a difference between invader and resident species mean growth rates, when $i$ is the invader.
+In MCT, coexistence is assessed by the ability to increase when rare, i.e., 
+the mean ``growth when rare'' [GWR] of species $i$, 
+quantified by the weighted difference between mean growth rates when species $i$, the invader, is at negligible abundance 
+and species $j$, the resident (the other species) is at stochastic steady state.
+The GWR is computed for each species assessed as the invader of a resident community, and only when the difference is positive for all is coexistence possible.
+The weighted difference, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, can be decomposed into coexistence mechanisms. 
+A subscript $i \bs i$ indicates a quantity computed for species $i$ when that species is at negligible abundance, 
+and a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent. 
+The \emph{scaling factor} $q_{ij}$ weights the relative sensitivities to competition of the two species, 
+see \cite{Chesson_1994}, \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two alternative definitions of $q_{ij}$ that we use.
+
+The storage effect %for species $i$ 
+is the contribution of covariation between $E$ and 
+$C$ to a difference between invader and resident species mean growth rates  \cite{Ellner_2016}. %, when $i$ is the invader.
 Specifically, storage effect for species $i$ is 
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction. One typically %(but see below) 
-must examine this quantity for each species, $i$,
-to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
+which is the extension of MCT which was goal G1 of the Introduction. 
+%One typically %(but see below) must examine this quantity for each species, $i$,to obtain complete information about coexistence. 
+To pursue goal G2, below, we will quantify
 the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.

### 2022-07-30 paper/Paper.Rnw
edits

@@
 Critically, the storage effect can be quantified in invasion analysis by measuring to what extent EC covariance contributes to differences between 
-invader (rare) and resident (stable-state) species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
+invader (rare) and resident (steady state) species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
 Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
@@
 
-We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
-plays a role in coexistence. We develop each step of our new theory for a general 
-two-species model (below), and also specifically for the lottery model of the Introduction 
+We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATAs
+%plays a role in coexistence. 
+can structure coexistence. We develop each step of our new theory for a general 
+two-species model %(below)
+, and also specifically for the lottery model of the Introduction 
 (SI section \ref{SI-SIsect:TheoryForLottery}).
@@
 $C_i(t)$. Though the iid assumption %can be circumvented
-is not neccesary, %we make the assumption because 
+is not necessary, %we make the assumption because 
 it simplifies the presentation of new ideas.
@@
 were rendered independent of each other,
-for both invader and resideent species [see (\ref{eq:storage_effects_gen})]. 
+for both invader and resident species [see (\ref{eq:storage_effects_gen})]. 
 Because %these components 
@@
 altered relative to the original distributions. 
-
-Specifically, asymmetries of tail association are removed -- see below for how this is done. 
+Specifically, asymmetries of tail association are removed. 
 We then define $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model 
@@
 (\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effect, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
-below for additional details.
+of storage effect, which are themselves a term in the MCT decomposition \citep{Ellner_2019}.
 
@@
 For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
+
 It seems reasonable to suppose that ATAs will play an increasingly important role in systems of 

### 2022-07-30 code/lottery_beta/decompose_LB.R
new code suite for using beta fecundities

@@ -1,0 +1,156 @@
+numeric_results_loc <- "../results_numeric"
+betanoise_loc <- paste0(numeric_results_loc, "/betanoise.RDS")
+b<-readRDS(betanoise_loc)
+
+#load litle b's
+# B_i = sigma_i * b_i + mu_i
+
+decompose <- function(mu_1,mu_2,sigma_1,sigma_2, delta,blist,dir="LEFT") {
+  #notation follows from paper/sup mat
+  
+  b <- data.frame(blist$l)
+  b_til <- data.frame(blist$til)
+  b_um <- data.frame(blist$um)
+  if (dir=="RIGT"){
+  	b <- data.frame(blist$r)
+  }
+  
+  #length
+  M <- nrow(b$i) 
+  
+  #defining epsilons for rare species i (second list in S6)
+  #1. epsilon^0_i (eq 39)
+  e0i <- log(1-delta+delta*(mu_i/mu_j))
+  
+  #2. epsilon^E_i bar (eq 42)
+  #estimate
+  eEi_hat <- mean(log(1-delta+delta*((sigma_i*b$i + mu_i)/mu_j))) - e0i
+  #standard error
+  #eEi_se <- sd()/sqrt(M)
+  
+  #3. epsilon^C_i bar (eq 45)
+  #estimate
+  eCi_hat <- mean(log(1-delta+delta*(u_i/(sigma_j*b$j + mu_j)))) - e0i
+  #standard error
+  #eCi_se <- sd()/sqrt(M)
+  
+  #4. epsilon^(E#C) bar (eq 51)
+  #estimate
+  e_ECshi_hat <- mean(log(1-delta+delta* ((sigma_i*b_til$i+mu_i)/(sigma_j*b_til$j+mu_j))))- eEi_hat-eCi_hat - e0i
+  #standard error
+  #eECshi_se <- sd()/sqrt(M)  
+  
+  #5. epsilon^[EC] bar (eq 55)
+  #estimate
+  eECi_hat <- mean(log(1-delta+delta* ((sigma_i*b$i+mu_i)/(sigma_j*b$j+mu_j)))) - mean(log(1-delta+delta* ((sigma_i*b_um$i+mu_i)/(sigma_j*b_um$j+mu_j)))) 
+  #standard error
+  #s1 <- sd()/sqrt(M)
+  #s2 <- sd()/sqrt(M)
[…]

### 2022-07-30 code/lottery_beta/makenoise_LB.R

@@ -1,0 +1,79 @@
+#This script makes sets of noise to be used in the lottery model analysis and saves their length, M, and correlation, rho
+  #b_tilde: list of 3 different types of bivariate noise: left tail ATA, righ tail ATA, and symmetric
+  #u_tilde: univariate standard normal 
+
+##libraries used (invoked with ::): MASS, stats
+
+### location to save results ###
+numeric_results_loc <- "../results_numeric"
+
+if(dir.exists(numeric_results_loc)==FALSE){
+  dir.create(numeric_results_loc)
+}
+noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
+M_loc <- paste(numeric_results_loc, "/M.RDS", sep = "")
+rho_loc <- paste(numeric_results_loc, "/rho.RDS", sep = "")
+
+### makenoise function ###
+#follows protocol described in SupMat section 
+#makes list of length 3 containing the three different types of bivariate noise: 
+#left tail ATA, right tail ATA, symmetric, standard
+#v_l, v_r, v_um (umlaut), v_ti (tilde)
+#ARGS:
+  #M    numeric, length of noise
+#OUT:
+  #List length 4, each element is a matrix with dimensions 2 x M
+makenoise <- function(M){
+  
+  #Sigma matrix and mu to make standard normal bivariate noise
+  Sigma_norm <- matrix(c(1,0,0,1),nrow=2)
+  mu_norm <- c(0,0)
+  
+  #bivariate standard normal
+  v <- MASS::mvrnorm(M, mu_norm, Sigma_norm) #points 
+  colnames(v) <- c("i","j")
+  v_ti <- v
+  
+  #half of points will be more than zero, 50%
+  a1 <- v[,1]; a2 <- v[,2]
+  
+  A <- a1>0 
+  
+  #left tail
+  v[A,] <- (-abs(a1[A])) #50% v[,1] = v[,2] = -|a_1|
+  v[!A,] <- abs(v[!A,]) #50% v[,1] = |a_1|, v[,2] = |a_2|
+  v_l <- v
+  
+  #right tail - get by flipping left tail noise 
+  v[A,] <- -v[A,] 
+  v[!A,] <- -v[!A,]
+  v_r <- v
+  
+  #symmetric 
+  #with similar correlation as asymmetric noises
+  Sigma_sym <- stats::cor(v_l) 
[…]

### 2022-07-30 code/lottery_beta/noisetransformandnotes (Autosaved).R

@@ -1,0 +1,212 @@
+#transforrming normal noise to beta
+F <- function(noise,...){
+	qbeta(pnorm(noise),...)-0.5
+}
+
+obj <- F(b[[1]], shape1=0.5, shape2=0.5)
+hist(obj)
+
+#making noise to be used
+# A
+## 1) (v_i_r, v_j_r), (v_i_l, v_j_l): 
+	# left and right TA bivariate noise with standard normal marginals 
+## 2) (v_i_tilde, v_j_tilde): 
+	# bivaraite normal mean (0,0) and cov (1,0,0,1)
+## 3) (v_i_umlaut, v_j_umlaut):
+	# bivariate normal mean (0,0) and cov (1,rho,rho,1) 
+		# where rho is cov between v_j and v_i
+# B
+## make beta b_l, b_r, b_umlaut, b_tilde 
+	# by taking the F of the respective (v_i, v_j)'s 
+
+source("./lottery_beta/makenoise_LB.R")
+noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
+load(noise_loc) #object is called v
+hist(v$l[,1]); mean(v$l[,1]); sd(v$l[,1])
+hist(v$l[,2]); mean(v$l[,2]); sd(v$l[,2])
+hist(v$r[,1]); mean(v$r[,1]); sd(v$r[,1])
+hist(v$r[,2]); mean(v$r[,2]); sd(v$r[,2])
+hist(v$til[,1]); mean(v$til[,1]); sd(v$til[,1])
+hist(v$til[,2]); mean(v$til[,2]); sd(v$til[,2])
+hist(v$um[,1]); mean(v$um[,1]); sd(v$um[,1])
+hist(v$um[,2]); mean(v$um[,2]); sd(v$um[,2])
+plot(v$l[1:10000,1], v$l[1:10000,2]); cov(v$l)
+plot(v$r[1:10000,1], v$r[1:10000,2]); cov(v$r)
+plot(v$til[1:10000,1], v$til[1:10000,2]); cov(v$til)
+plot(v$um[1:10000,1], v$um[1:10000,2]); cov(v$um)
+b <- lapply(v, FUN=F, shape1=0.5, shape2=0.5)
+#check
+hist(b$l[,1]); mean(b$l[,1]); sd(b$l[,1])
+hist(b$l[,2]); mean(b$l[,2]); sd(b$l[,2])
[…]

### 2022-07-30 code/lottery_beta/noisetransformandnotes.R

@@ -1,0 +1,210 @@
+#transforrming normal noise to beta
+F <- function(noise,...){
+	qbeta(pnorm(noise),...)-0.5
+}
+
+obj <- F(b[[1]], shape1=0.5, shape2=0.5)
+hist(obj)
+
+#making noise to be used
+# A
+## 1) (v_i_r, v_j_r), (v_i_l, v_j_l): 
+	# left and right TA bivariate noise with standard normal marginals 
+## 2) (v_i_tilde, v_j_tilde): 
+	# bivaraite normal mean (0,0) and cov (1,0,0,1)
+## 3) (v_i_umlaut, v_j_umlaut):
+	# bivariate normal mean (0,0) and cov (1,rho,rho,1) 
+		# where rho is cov between v_j and v_i
+# B
+## make beta b_l, b_r, b_umlaut, b_tilde 
+	# by taking the F of the respective (v_i, v_j)'s 
+
+source("./lottery_beta/makenoise_LB.R")
+noise_loc <- paste(numeric_results_loc, "/noise.RData", sep = "")
+load(noise_loc) #object is called v
+hist(v$l[,1]); mean(v$l[,1]); sd(v$l[,1])
+hist(v$l[,2]); mean(v$l[,2]); sd(v$l[,2])
+hist(v$r[,1]); mean(v$r[,1]); sd(v$r[,1])
+hist(v$r[,2]); mean(v$r[,2]); sd(v$r[,2])
+hist(v$til[,1]); mean(v$til[,1]); sd(v$til[,1])
+hist(v$til[,2]); mean(v$til[,2]); sd(v$til[,2])
+hist(v$um[,1]); mean(v$um[,1]); sd(v$um[,1])
+hist(v$um[,2]); mean(v$um[,2]); sd(v$um[,2])
+plot(v$l[1:10000,1], v$l[1:10000,2]); cov(v$l)
+plot(v$r[1:10000,1], v$r[1:10000,2]); cov(v$r)
+plot(v$til[1:10000,1], v$til[1:10000,2]); cov(v$til)
+plot(v$um[1:10000,1], v$um[1:10000,2]); cov(v$um)
+b <- lapply(v, FUN=F, shape1=0.5, shape2=0.5)
+#check
+hist(b$l[,1]); mean(b$l[,1]); sd(b$l[,1])
[…]

### 2022-10-10 paper/Paper.Rnw
theory fig with color

@@
 \begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/theory_fig_col.pdf}
+\caption{Pedagogical figure to demonstrate how asymmetric tail associations are removed while maintaining original marginal distributions and correlation between variables.
+(a) the starting variables with ATA.
+(b) the variables are transformed by a normal inverse cumulative distribution function of their own cumulative distribution function.
+(with a Spearman correlation of $\rho).
+Asymmetric association structure is maintained.
+(c) a newly generated, naturally symmetric, bivariate normal with the correlation, $\ddot{\rho}=\rho.
+(d) two variables which have the same marginals and correlation as the starting variables in (a) but have a symmetric association structure, obtained by taking the inverse cumulative distribution functions of the starting variables of the normal cumulative distribution function of the normal distributions.
+For demonstration purposes we start with beta distributed variables, though variables can be distributed differently and the two can be dissimilar.
+\label{fig:theory}}
+\end{figure}
+
+How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
+correlation unchanged? 
+We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, 
+respectively, are strictly monotonic and continuous. 
+Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$. %, which are standard normally
+These are standard Normal 
+distributed, although they are not necessarily jointly bivariate normal.  
+Let $\rho_i$ be the covariance of these random variables. 
+We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
+$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
+$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. 
+The value of $\ddot{\rho}_i$ is then determined in one of four alternative ways: 
+either by setting 
+$\ddot{\rho}_i = \rho_i$; 
+or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. 
+The choice of bivariate normal 
+$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
+
+\begin{figure}
 \includegraphics[width=\textwidth]{../results_figs/fig3.pdf}

### 2022-10-10 code/theory_panel_fxn.R

@@ -1,5 +1,5 @@
 
-theoryfigpanel <- function(d,p,q,cop,x,y, gnrt_var=FALSE, samps=NULL, oldp=NULL,...){
+theoryfigpanel <- function(d,p,q,cop,x,y, gnrt_var=FALSE, samps=NULL, oldp=NULL,col=F,logpoints=F,...){
 	probdisf <- makepdf(cop,d,d,p,p)
 	xy <- expand.grid(x,y)
 	z <- probdisf(xy)
@@ -7,15 +7,36 @@ theoryfigpanel <- function(d,p,q,cop,x,y, gnrt_var=FALSE, samps=NULL, oldp=NULL,
 	
 	#layout(matrix(c(2,4,1,3), byrow=T, nrow=2), heights=c(.3,1), widths=c(1,.3))
 	#par(mar=c(0.5,0.5,0.5,0.5), oma=c(2,2,1,1), xpd=NA)
-	levs<-c(-.5,0,1,1.5)
-	contour(x,y,log10(z),xlim=range(x),ylim=range(y),nlevels=10, bty='l')
+	#contour(x,y,log10(z),xlim=range(x),ylim=range(y),nlevels=10, bty='l')
+	plot(x,y,type="n", xlab="",ylab="", bty="l")
 	if (gnrt_var == TRUE){
 		rand_var <- makerandgenrtr(cop,q,q)
-		samps <- rand_var(500)
+		samps <- rand_var(1000)
 	} else{
 		samps <- tdis(q, oldp, samps)
 	}
-	points(samps[,1],samps[,2],type="p",pch=20, cex=1, col=rgb(.74,.74,.74,.4))
+	if (col==TRUE){
+	  ncol <- 100
+	  if (logpoints==TRUE){
+	    z[z==0]<-NA; zvar <- log10(z)
+	  } else { zvar <- z }
+	  z_ran <- range(zvar, na.rm=T)
+	  zbin <- .bincode(zvar, breaks=seq(z_ran[1],z_ran[2],len=ncol+1),TRUE,TRUE)
+	  zbin <- matrix(zbin, length(x), length(y))
+	  cols <- hcl.colors(ncol, palette = "YlOrRd",rev=T)
+	  xbin <- .bincode(samps[,1], breaks=seq(min(x),max(x),len=nrow(z)+1),TRUE,TRUE)
[…]

### 2022-10-10 code/theoryfig.R

@@ -11,7 +11,11 @@ source("./theory_panel_fxn.R")
 
 #figure file location and names
 fig_loc <- "../results_figs"
+if (dir.exists(fig_loc)==F){
+  dir.create(fig_loc)
+}
 theoryfig <- paste0(fig_loc,"/theory_fig_rw.pdf")
+theoryfig_col <- paste0(fig_loc,"/theory_fig_col.pdf")
 theoryfig_qd <- paste0(fig_loc,"/theory_fig_qd.pdf")
 
 #copulas
@@ -45,6 +49,15 @@ pdf(theoryfig, height=4, width=15)
 layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
 par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
 source("./theoryfig_meat.R")
+dev.off() 
+########
+
+#colored points
+#### start pdf ####
+pdf(theoryfig_col, height=4, width=15)
+layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
+par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
+source("./theoryfig_col_meat.R")
 dev.off() 
 ########
 

### 2022-10-10 code/theoryfig_col_meat.R

@@ -1,0 +1,34 @@
+# A BETA MARGS CLAYTON COP
+sampsA <- theoryfigpanel(dbmarg, pbmarg, qbmarg, ccop, x, y, gnrt_var=TRUE,col=T,logpoints=T)
+#text(x=lblc["E","x"], y=lblc["E","y"], labels=expression('E'['i']), cex=2)
+text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX(r'($E_i$)'), cex=2)
+#text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\i']), srt=90, cex=2)
+text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX(r'($C_{i\setminus i}$)'), srt=90, cex=2)
+text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[1],")"),cex=2)
+
+# B NORM MARGS CLAY COP
+sampsB <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ccop, x, y, samps=sampsA, oldp=pbmarg,xpd=NA,col=T)
+#text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression(phi^-1*'o'*'F'['E'['i']]("E"["i"]))), cex=2)
+text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX(r'($\phi^{-1} \o F_{E_i}(E_i)$)'), cex=2)
+#text(x=lblc["C","x"], y=lblc["C","y"], labels=c(expression(phi^-1*'o'*'F'['C'['i/i']]("C"["i/i"]))), srt=90, cex=2)
+text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX(r'($\phi^{-1} \o F_{C_{i\setminus i}}(C_{i\setminus i})$)'), srt=90, cex=2)
+text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[2],")"), cex=2)
+
+# C NORM MARGS NORM COP
+sampsC <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ncop, x, y, gnrt_var=TRUE, xpd=NA,col=T)
+#text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression('e'['i'])), cex=2)
[…]

### 2023-01-25 paper/Paper.Rnw
edited the Intro and Fig 1 caption

--- 2022-07-13 paper/Paper.Rnw
@@
 %header material for paper
-\title{Relationships in the extremes and their influence on competition and coexistence}
+\title{Asymmetric relationships and their impact on coexistence}
 \date{}
@@
 %Put other possible titles below, commented, and we will select the best one at the end
+%Relationships in the extremes and their influence on competition and coexistence
 %Relationships between extreme events and their influence on competition and coexistence
@@
 %Extreme [climatic] events and their implications for biodiversity
-%Deductions to coexistence, extreme events, and asymmetric relationships
-%Extreme relationships (relationships in their extremes) and their deductions ([negative] contributions) to coexistence
 %Coexistence and relationships between environmental variables in their extremes
@@
 \author[a]{Pimsupa Jasmin Albert}
-\author[b]{Stephen P. Ellner}
-\author[a,c,*]{Daniel C. Reuman}
+\author[a,b,*]{Daniel C. Reuman}
 
 \affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
-\affil[b]{Department of Ecology and Evolutionary Biology, Cornell University}
-\affil[c]{Laboratory of Populations, Rockefeller University}
+\affil[b]{Laboratory of Populations, Rockefeller University}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 theory (MCT) identifies coexistence mechanisms,
-one of which, the storage effect, hinges on relationships between fluctuations
+one of which, storage effects, hinges on relationships between fluctuations
 in environmental and competitive pressures. However, such relationships
@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend MCT, decomposing the storage effect to show the influence of ATAs.
+We extend the MCT, decomposing storage effects to show the influence of ATAs.
 Analysis of a simple model
@@
 
-\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effect}
+\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effects}
 
@@
 
-%Intro to ATAs
-Every ecological discipline involves the study of relationships among biological and environmental variables,
-but standard %statistical methods 
-measures for %examining the 
-measuring the strength and direction of the association between two variables provide only 
-%a small portion of the overall information content of the
-partial information about the nature of the 
-association \citep{Ghosh_2020_AER,Ghosh_2020_Eco,Ghosh_2020_ecolevol,Ghosh_2021}. 
-For instance, consider the three pairs of variables in Fig. \ref{fig:noise}.
-The variables in Fig. \ref{fig:noise}a,c exhibit what we have previously referred to as ``asymmetric tail association'' 
-[ATA; \cite{Ghosh_2020_AER}], i.e., they are more strongly related to each
-other in the left (Fig. \ref{fig:noise}a, termed ``left-tail association'') or right (Fig. \ref{fig:noise}c,
-termed ``right-tail association'') portions of the variables' distributions. 
-The relationships displayed in Fig. \ref{fig:noise} demonstrate very different types of association, 
-but nevertheless all have the same Pearson correlation. 
-Other metrics of association are also often used, in addition to correlations 
-(e.g., various forms of regression), but correlations are extremely common, and other
-methods also reveal only a small portion of the information content of the relationship between variables \citep{Nelsen2006}.
-Although ATAs and related ideas had previously been seldom 
-considered in ecology (but see \citealp{deValpine_2014, Anderson_2019, Popovic_2019}), 
-our recent work shows that ATAs are extremely common \citep{Ghosh_2020_AER}, and can influence 
-many phenomena, including extinction risk \citep{Ghosh_2020_Eco}, Taylor's law \citep{Ghosh_2020_AER}, 
-synchrony, and stability of ecosystem functioning \citep{Ghosh_2020_ecolevol,Ghosh_2021}.
-Although ATAs are now known to be important, the implications of ATAs for competition and species coexistence 
-have, to our knowledge, not previously been considered; doing so is the overarching mission of this study.
+%Intro the paper by intro'ing ATAs
+At the core of ecology is the study of relationships - between organisms and between organisms and their environments. 
+Ecologists may ask, for instance, how fluctuations in local- and regional-scale environments relate to 
+population dynamics. To study such relationships and to understand how two variables interact, ecologists 
+very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
+and direction of association between variables often don't tell the whole story 
+\citep{Ghosh_2020_AER}.
+To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
+Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
+symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what we have previously referred to as 
+``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
+Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
+regression), but correlations are extremely common, and other methods also reveal only a small portion 
+of the information content of the relationship between variables \citep{Nelsen2006}.
+ATAs and related ideas had previously been seldom considered in ecology, but recently several 
+insights have been gained by studying how features of associations beyond standard measures 
+impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
+Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_AER} show that ATAs are common in ecology
+and can influence Taylor's law and Moran effects; \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
+when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
+ecosystem functioning. Fig. \ref{fig:pedagog1}g,h show contrasting examples of ATAs in nature using 
+plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
+not yet been investigated. Doing so is the overarching goal of this study.
 
 %Intro to competition, coexistence, MCT, etc.
-Competition and the maintenance of biodiversity are research topics at the core of modern ecology 
-\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}.
-The competitive exclusion principle states 
-that two species competing for the same limiting resource cannot coexist at constant densities - one will instead outcompete 
-the other and drive it extinct \citep{Hardin_1960}.
-Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] %is a 
-has emerged as a
-framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems. 
-MCT breaks down coexistence into the contributions of various mechanisms, %with names such as ``relative nonlinearityin competition'' and ``storage effect'' \citep{Chesson_1994} which will be familiar to many readers.
-highlighting the importance of environmental fluctuations to species coexistence. 
-For example, one such mechanism, the storage effect, promotes coexistence when species can reap or "store" the benefits of times when environmental conditions are good to overcome competition and persist through bad conditions.
-MCT has been applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
+Competition, coexistence, and the maintenance of biodiversity are research topics at the core of modern ecology 
+\citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
+that two species competing for the same limiting resource cannot coexist at constant densities - one will 
+instead outcompete the other and drive it extinct \citep{Hardin_1960}.
+Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
+framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
+through the niche partitioning effects of environmental variation.
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
+in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
+applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
-%Although MCT is mathematically complex to many ecologists, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
-Though MCT's original analytical form presented mathematical challenges for applications to and tests of new systems, recent developments expand its testability and accessibility and present the opportunity for further extensions \citep{Ellner_2016,Ellner_2019}.
+Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
 
 %Intro to storage effects and why they might be influenced by ATAs
-In this paper, we show how ATAs can also be important for competition and for the coexistence of
-competing species. In particular, we show how ATAs can strongly influence one specific coexistence mechanism:
-storage effects.
-%It is straightforward to see, intuitively, why ATAs may influence one mechanism of coexistence in particular: the storage effect. 
-%We review, conceptually, what the storage effect is \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs may be relevant to them.
-Storage effect allows competing species to fluctuate %over time while escaping extinction because storage effects provide 
-in abundance without any going extinct by providing
-a relative benefit to whichever species is currently rare; this is the characteristic that makes the storage effect a so-called \emph{stabilizing} mechanism 
-of coexistence \citep{Chesson_2000}.
-Storage effect can occur if three conditions are met \citep{Chesson_1994}. First, there must be positive 
-covariation between the fluctuating quality of the environment (E)
-for a species and the strength of competition (C) experienced by that species (called EC covariance), that is stronger when the species is at stable-state abundance than when it is rare. 
-Second, growth responses to the environment must be species-specific, so that distinct species do not find the 
-same environmental conditions to be optimal. 
-Third, a buffering mechanism %must exist whereby species limit 
-which limits the influence of competition %in times 
-when the environment is poor. 
-Dormant stages such as seed banks are an example buffering mechanism \citep{Chesson_2012}. 
-When present, these three factors combine so that EC covariance provides a relative benefit to whatever species is currently rare.
-Critically, the storage effect can be quantified in invasion analysis by measuring to what extent EC covariance contributes to differences between 
-invader (rare) and resident (stable-state) species growth rates \citep{Chesson_1994,Ellner_2016}. %invader and resident 
-Thus ATAs may, in principle, influence storage effect simply because storage effect hinges on EC covariation, and ATAs, when they 
-occur between environment and competition variables, can be an aspect of EC covariation.
+It is possible to see, intuitively, 
+why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
+may be relevant to them. Storage effects can allow for species to coexist while fluctuating in abundance through time
+because storage effects can provide a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
+Storage effects hinge on positive covariation between the fluctuating quality of the environment
+for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
+and on the assumption that optimal environmental conditions for distinct species are also distinct,
+so that no two species find the same conditions to be optimal. Then, good environmental
+conditions for a currently common species are paired with strong intraspecific competition, because of
+positive EC covariance and the commonness of the species. But good environmental conditions for 
+a currently rare species are not as closely accompanied by strong competition because the rarity of
+the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
+rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
+differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
+ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
+and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.
 
 %Lottery model and results based on it
-%To further demonstrate the plausibility of ATAs altering coexistence between species, 
-To more robustly demonstrate the role of ATAs in structuring fluctuation-dependent coexistence, we carried out simulations using a simple two-species model, the lottery model. 
-The simulations described in this paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; 
-the precise nature of that influence is then the subject of the rest of the paper.
-The lottery model was originally introduced as a model for coral reef fish competing for
-a set of $N$ breeding territories.
+To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
+simple simulations using a two-species model, the well-known lottery model. The simulations described in this
+paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
+precise nature of that influence is the subject of the rest of the paper.
+The lottery model is as follows.
 Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
@@
 for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
-The model postulates that adults die at rate $\delta$ at each time step, and are replaced by new juveniles 
+The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
 in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
-For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically distributed (iid) through
-time. We let $B_i = \exp(b_i)$, where $\left(b_1, b_2\right)$ is some bivariate random variable with the property 
-that $b_i$ is normally distributed with mean $\mu_i$ and variance $\sigma^2$. 
-We do \emph{not} assume that $\left(b_1, b_2\right)$ is a 
-bivariate normal distribution. 
-For proof of concept, we simulated two communities for \Sexpr{M} time 
-steps each, using symmetric and asymmetric tail associations (Fig. \ref{fig:noise}) between $b_1$ and $b_2$, respectively, in the
-two simulations. Fig. \ref{fig:pedagog} gives details of the 
-simulations and shows that ATAs substantially influenced aspects of species coexistence for this model setup.
+For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
+distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
+$B_1$ and $B_2$ were, respectively, 
+left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
+tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
+for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
+ATAs substantially influenced aspects of species coexistence (Fig. \ref{fig:pedagog1}d-f).
 
-%Climate change may alter ATAs of coexistence-relevant ATAs
-An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
-alter ATAs of environmental variables which impact coexistence, as follows.
-It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
-heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
-If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
-may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
-For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
-whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
-contributing to their coexistence, or failure to coexist, can boil down to covariation 
-between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
-to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
-longer lasting increases the degree of right-tail association between early and late spring rainfall. 
-Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
-left-tail association between these variables. As we will show in this study, changes in ATAs can then
-result in different competitive outcomes between species. 
-\cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
-community of desert annuals that seem to have all the ingredients necessary to make effects of the
-type we describe here a possibility.
-To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-not been studied, but the importance for coexistence of changes 
-in means and variability of environmental variables has been studied in many systems
-[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
-Changes in variability and in extreme events are distinct from the changes in ATAs we 
-consider here, because ATAs pertain to \emph{relationships} between environmental and other
-variables in the extremes, which are distinct from and can be altered independently
-of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
+%***DAN: I commented the below because we decided it does not belong in the Intro, but a version of it 
+%should be in the Discussion. Keeping it for now until I have a chance to move it there. Note also that
+%I now think this material needs some editing for clarity, some of that was done in a paper version I 
+%marked up in January 2023.
+% %Climate change may alter ATAs of coexistence-relevant ATAs
+% An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
+% alter ATAs of environmental variables which impact coexistence, as follows.
+% It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
+% heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
+% If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
+% may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
+% For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
+% whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
+% contributing to their coexistence, or failure to coexist, can boil down to covariation 
+% between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
+% to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
+% longer lasting increases the degree of right-tail association between early and late spring rainfall. 
+% Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
+% left-tail association between these variables. As we will show in this study, changes in ATAs can then
+% result in different competitive outcomes between species. 
+% \cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
+% community of desert annuals that seem to have all the ingredients necessary to make effects of the
+% type we describe here a possibility.
+% To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
+% not been studied, but the importance for coexistence of changes 
+% in means and variability of environmental variables has been studied in many systems
+% [e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
+% Changes in variability and in extreme events are distinct from the changes in ATAs we 
+% consider here, because ATAs pertain to \emph{relationships} between environmental and other
+% variables in the extremes, which are distinct from and can be altered independently
+% of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
 %Goal para
-We previously demonstrated \citep{Ghosh_2020_AER} that ATAs are common in nature, and we argued above 
-that it is plausible both that climate change may alter environmental ATAs, and that ATAs may influence species 
-coexistence. Thus we seek to explore the consequences of ATAs for competition, coexistence, and the maintenance of 
-biodiversity, by means of the following specific goals. 
+We previously demonstrated that ATAs are common in nature can influence ecological phenomena
+\citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
+indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
+of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
+following specific goals. 
 G1) To augment MCT to show formally how ATAs play a role in coexistence. 
@@
 We address how ATAs contribute to these components. 
-We here examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
+We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
 spatial variation can be considered in future work.
 G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
-from ATAs compared to other, previously known mechanisms of coexistence. This is to help 
-illuminate whether ATA contributions to coexistence 
+from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
+sometimes coexist with ATAs but not without, or vice versa. This is to help 
+illuminate whether ATA contributions to coexistence and biodiversity maintenance
 are likely to be negligible or meaningfully large. 
-G3) To explore, using the lottery model, whether species can coexist with
-ATAs but not without, or vice versa. This is to begin to illuminate whether ATA contributions to biodiversity maintenance 
-may be negligible or meaningfully large.
-G4) To provide an empirical example of species coexistence which highlights the role ATAs can play.
-In the Discussion, we revisit the possibility that climate change will alter
-ATAs of environmental variables and thereby influence competition and species coexistence.
-Overall, our study presents a new mechanism of species coexistence which may be influenced by climate change, 
+G3) To provide an empirical example of species coexistence which highlights the role ATAs can play.
+In the Discussion, we consider whether climate change may alter
+ATAs of environmental variables and thereby influence competition and coexistence.
+Overall, our study presents a new mechanism of species coexistence
 and a means of understanding its theoretical and empirical importance.
@@
 quantify the contributions of 
-multiple mechanisms to coexistence. But we show
-that, for purely temporal variation, only one mechanism relates to tail associations: the storage effect. 
-So we begin by defining the storage effect. Our mathematical notation parallels
+multiple mechanisms to coexistence. But we show below
+that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
+So we begin by defining storage effects. Our mathematical notation parallels
 that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
@@
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption %can be circumvented
-is not neccesary, %we make the assumption because 
-it simplifies the presentation of new ideas.
+$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
+presentation of new ideas.
 
-In MCT, coexistence is assessed by the ability to increase when rare, i.e., 
-the mean ``growth when rare'' [GWR] of species $i$, 
-quantified by the weighted difference between mean growth rates when species $i$, the invader, is at negligible abundance 
-and species $j$, the resident (the other species) is at stochastic steady state.
-The GWR is computed for each species assessed as the invader of a resident community, and only when the difference is positive for all is coexistence possible.
-The weighted difference, $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, can be decomposed into coexistence mechanisms. 
-A subscript $i \bs i$ indicates a quantity computed for species $i$ when that species is at negligible abundance, 
-and a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent. 
-The \emph{scaling factor} $q_{ij}$ weights the relative sensitivities to competition of the two species, 
-see \cite{Chesson_1994}, \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two alternative definitions of $q_{ij}$ that we use.
-
-The storage effect %for species $i$ 
-is the contribution of covariation between $E$ and 
-$C$ to a difference between invader and resident species mean growth rates  \cite{Ellner_2016}. %, when $i$ is the invader.
-Specifically, storage effect for species $i$ is 
+The storage effect for species $i$ is the contribution of covariation between $E$ and 
+$C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
+Specifically, storage effects for species $i$ are 
 \begin{equation}
@@
 Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
-%Though some readers may find it slightly confusing to simultaneously use the variable $E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would be more confusing to use alternative notation.
-Like all variables, $E_i$ is rendered in italic; like all operators, $\E(\cdot)$ is rendered upright.
+Though some readers may find it slightly confusing to simultaneously use the variable 
+$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would 
+be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
+all operators, $\E(\cdot)$ is rendered upright.
 
-We now specify how to decompose storage effect into contributions due to ATAs and 
-due to $EC$ correlation \emph{per se}. 
-Storage effect for species $i$ was computed by comparing
-differences between invader and resident species mean growth rates for the original model and for a 
-modified model for which 
-%the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
-$E$ and $C$
-were rendered independent of each other,
-for both invader and resideent species [see (\ref{eq:storage_effects_gen})]. 
-Because %these components 
-$E$ and $C$ 
-were rendered completely independent in the modified model, 
-storage effect quantifies the contribution of the whole $EC$ relationship, in all its aspects, 
-to the invader/resident growth rate difference. 
-But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
+We now specify how to decompose storage effects into
+contributions due to ATAs and due
+to $EC$ correlation \emph{per se}. 
+Storage effects for species $i$ were computed by comparing
+differences between rare and common species mean growth rates for the original model and for a 
+modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
+were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
+Because these components were rendered 
+completely independent in the modified model, storage effects quantify the contribution of the 
+whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
+difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
 $(E_j^\nat,C_{j \bs i}^\nat)$ with the same
@@
 and with some but not all aspects of the relationship 
-between %the components 
-$E$ and $C$
-altered relative to the original distributions.
-
-Specifically, asymmetries of tail association are removed -- see below for how this is done. 
-We then define $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model 
-using  $(E_i^\nat,C_{i \bs i}^\nat)$
+between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
+$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
 and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
@@
 \end{align}
-The term $\Delta_i^{[EC]}$ is the component of storage effect due to 
+The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
 ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
-The term $\Delta_i^{[E \nat C]}$ is the component of storage effect due to $EC$ 
+The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
 correlation \emph{per se}, remaining after having removed the effects of ATAs.
@@
 (\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effect, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
+of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
 below for additional details.
@@
 How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? 
-We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
-$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, 
-respectively, are strictly monotonic and continuous. 
+correlation unchanged? We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
 Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
-$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$. %, which are standard normally
-These are standard Normal 
-distributed, although they are not necessarily jointly bivariate normal.  
-Let $\rho_i$ be the covariance of these random variables. 
-We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
+distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
+these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
 $\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
@@
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. 
-The value of $\ddot{\rho}_i$ is then determined in one of four alternative ways: 
-either by setting 
-$\ddot{\rho}_i = \rho_i$; 
-or so that the Pearson, Spearman or Kendall correlation of 
-$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. 
-The choice of bivariate normal 
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
+in one of 
+four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
 $(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-
-For those familiar with copula statistics 
-(see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
+For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
 we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
@@
 $C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. 
-The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
+as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
 
 The other mechanisms in MCT that apply in our modelling context of purely temporal variation
-make no reference to covariation between quantities, 
-(see eqn. (13) in Ellner (2019),
-whereas ATAs are %defined in a covariation context
-strictly about covariation. 
-Therefore, the storage effect is the only mechanism of coexistence
+make no reference to covariation between quantities, whereas ATAs are 
+defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
 for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
@@
 \cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
-which equals GWR of species $i$. The decomposition, which we use below, is analogous to and motivated by the 
+which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
 original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
@@
 of covariation in these quantities; and
-$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- the storage effect. 
-Only storage effect involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is the storage effect. Combining our decomposition 
+$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
+Only storage effects involve $EC$ covariation,
+so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
 (\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction. 
-%One typically %(but see below) must examine this quantity for each species, $i$,to obtain complete information about coexistence. 
-To pursue goal G2, below, we will quantify
+which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
+to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
 the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
@@
 %G2
-To achieve goal G2 from the Introduction, 
-left- or right-tail associated %noise 
-log fecundities $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
+Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
 $\sigma$, was generated as pictured in Fig. \ref{fig:noise}, details in SI section \ref{SI-sect:noise}.
-%To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
-%to coexistence that come from ATAs versus other mechanisms, 
-We then plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
+To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
+to coexistence that come from ATAs versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
 $\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
-against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$. %; and we compared values on plots. 
-Some mathematical shortcuts, summarized here, simplified plotting. 
+against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and we compared 
+values on plots. Some mathematical shortcuts, summarized here, simplified plotting. 
 First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
@@
 lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). 
-So we plotted 
-the terms
+instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we plotted 
 against $\mu_1 - \mu_2$. 
 Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
-for both our left- and right-tail associated cases (SI section \ref{SI-sect:math_simplifications}), 
-so we only generated plots for left-tail associated noise. 
-For each combination of $\delta=0.2, 0.4, 0.6$
-and $\mu_1-\mu_2=0,-0.5,-1$,
+for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
+for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
+and $\mu_1-\mu_2=0,-0.5,-2,-4$,
 we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
@@
 In the $\mu_1=\mu_2$ case (panels a-d), for which the two species were equal 
-competitors, storage effect (blue plus red
-lines on Fig. \ref{fig:fig3}) was the only means by which coexistence could occur, and ATA effects
-tended to be negative, weakening total storage effect.
+competitors, storage effects (blue plus red
+lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
+tended to be negative, weakening total storage effects.
 
@@
 Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
-we decomposed the storage effect into two new mechanisms, the contribution of ATAs to coexistence,
+we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
 and the contribution of $EC$ correlation \emph{per se}.
@@
 system of competing diatoms demonstrated that  
-ATA contributions to the storage effect and to species coexistence and the maintenance of biodiversity
+ATA contributions to storage effects and to species coexistence and the maintenance of biodiversity
 can often be comparable in magnitude to the contributions of previously recognized mechanisms.
@@
 ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
-Though storage effect is typically described as the contribution
+Though storage effects are typically described as the contribution
 of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
-our results suggest that storage effect should be redefined to constitute the 
+our results suggest that storage effects should be redefined to constitute the 
 contribution of \emph{associations} between $E$ and $C$, including both ATA effects and
@@
 and our results help demonstrate strengths of that approach.
-Storage effect as computed using the original, analytic approach to MCT 
+Storage effects as computed using the original, analytic approach to MCT 
 \citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
@@
 \includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
-\caption{Simple example of bivariate random variables with standard normal marginals showing
-alternative patterns of association in the tails. The varables in (a) are ``left-tail 
-associated,'' those in (b) are symmetrically associated, and those in (c) are ``right-tail associated''. All three cases have 
-the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these very different 
-patterns of association. We used asymmetrically tail associated cases (a, c) which are perfectly 
-correlated in the lower or upper halves of the distributions, though it is also 
-possible to generate asymmetrically tail associated noise with imperfect correlations in 
-both halves of the distributions \citep{Nelsen2006,Ghosh_2020_AER}. The variables $(b_1,b_2)$ described in
-the Introduction and Methods were obtained by generating 
-noise as pictured and then multipling by $\sigma$ and adding $(\mu_1,\mu_2)$.\label{fig:noise}}
+%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
+%structure it is generated and stored, and after I have re-run Jasmin's code
+\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
+and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
+alternative patterns of association in the left and right parts of the distributions. 
+The two variables in (a) are more strongly related to each
+other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
+are symmetrically associated; and those in (c) 
+are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
+the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
+very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
+correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
+tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
+see also (g,h)]. 
+The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
+pictured and then transforming the marginals as specified in Methods.
+(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
+respectively corresponding variables (a-c; see Introduction).
+(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
+\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
+in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
+left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
 \end{figure}
@@
 coexistence and exclusion of \emph{F. crotonensis}. 
-Standard errors of plotted quantities were always less than \Sexpr{round(se5,5)}, so are not shown.
+Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
 For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
@@
 excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
-Standard errors of plotted quantities were always less than \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
+Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
 \end{figure}

### 2023-01-25 paper/SupMat.Rnw
now done editing, for now, all parts, including the sup mat, except the Disscussion
finished editing the Discussion
minor tweaks
line numbering
edited the Intro and Fig 1 caption

@@
 
-\DeclareMathOperator{\EX}{E}% expected value
-\DeclareMathOperator{\E}{E}
+\DeclareMathOperator{\E}{\mathbb{E}}% expected value
 \DeclareMathOperator{\VarX}{var}
@@
 %header material for paper
-\title{Relationships in the extremes and their influence on competition and coexistence: Supporting information}
+\title{Asymmetric relationships and their impact on coexistence: Supporting information}
 \date{}
@@
 \affil[*]{Corresponding author, reuman@ku.edu}
+
+%for dealing with line numbers bugs having to do with equations
+\let\oldequation\equation
+\let\oldendequation\endequation
+\renewenvironment{equation}
+  {\linenomathNonumbers\oldequation}
+  {\oldendequation\endlinenomath}
+\let\oldalign\align
+\let\oldendalign\endalign
+\renewenvironment{align}
+  {\linenomathNonumbers\oldalign}
+  {\oldendalign\endlinenomath}
 
@@
 %***DAN: Dan, you probably need to spend a bit more time with the cited papers to better understand this. 
-We also use $q_{ij}=1$ as an alternative value, based on communication with Steve Ellner, for the following reasons.
+We also use $q_{ij}=1$ as an alternative value, for the following reasons.
 First, as discussed in the supporting information of \cite{Ellner_2016}, and as further elaborated by 
@@
 
-%***DAN: Dan still needs to edit this section
 The decomposition of storage effects and quantification of contributions of ATAs depends on 

### 2023-01-26 paper/Paper.Rnw
small change that should have been included in the previous commit
edited Theory section and caption of fig 2 and corresponding sections of sup mat

@@
 
-We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
-plays a role in coexistence. We develop each step of our new theory for a general 
-two-species model (below), and also specifically for the lottery model of the Introduction 
-(SI section \ref{SI-SIsect:TheoryForLottery}).
+We here pursue goal G1 from the Introduction: to augment MCT to show formally how ATAs play a role in coexistence. 
+We develop our new theory for a general two-species model, later applying it to specific scenarios.
 MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
@@
 presentation of new ideas.
+
+\subsection{The storage effect}
 
@@
 Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
-Though some readers may find it slightly confusing to simultaneously use the variable 
-$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would 
-be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
-all operators, $\E(\cdot)$ is rendered upright.
+
+\subsection{Decomposing the storage effect}
 
@@
 The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
-ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
+ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
+the \emph{ATA effect on coexistence}. 
 The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
@@
 
+\subsection{Removing asymmetries of association}
+
 How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
+We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
 $C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
@@
 $\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
-distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
+distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
+Let $\rho_i$ be the covariance of
 these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
 $\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
-$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
+$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
 $E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
-in one of 
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
+The value of $\ddot{\rho}_i$ is then determined in one of 
 four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
@@
 as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
+
+\subsection{Our decomposition as an extention of modern coexistence theory}
 
@@
 to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
-the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.
+the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
+lottery model described in the Introduction.
 
@@
 Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
+The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
 The authors were partly supported by US National
-Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, and the California Department of Fish and
-Wildlife Delta Science Program. 
-%***DAN: DAN: if we take Ellner off the author list add him back here, and also add him back to the acks in the README. Also take him off the author list in sup mat.
+Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
+Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.
 
-%you store your bibliographic info in refs.bib, see that file
 \bibliographystyle{ecology_letters2}
@@
 
+%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
+%structure it is generated and stored, and after I have re-run Jasmin's code
 \begin{figure}
 \includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
-%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
-%structure it is generated and stored, and after I have re-run Jasmin's code
 \caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
@@
 
+%***DAN: I edited the caption (2023 01 26) but this is the completely the wrong figure. I need to replace it 
+%with Jasmin's new pedagogical figure once I know where in the repo 
+%structure that one is generated and stored, and after I have re-run Jasmin's code, and after she has made the
+%changes I recently asked for (see next comment).
+%***DAN: I asked Jasmin to replace the \phi with \varphi in the axis labels on b and d, and to make the marginals
+%on b and c be *standard* normal instead of just normal.
 \begin{figure}
 \includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
-\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered 
-from low densities less often and 
-spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
-species coexistence. Panel (a) shows species 1 from the lottery 
-model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
-same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}a).
-Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
-``noticeable coexistence'' of the species, defined to be periods when each species comprises at least $5\%$ of 
-total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
-was \Sexpr{fig2$coMets$Sym$co} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$co} for panel (b), over the full length of the simulations; 
-whereas the average duration of periods of
-one-species dominance was \Sexpr{fig2$coMets$Sym$dom} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$dom} for (b). Thus, while periods of 
-noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
-on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]}$,
-$\mu_2 = \Sexpr{fig2$params[2]}$, $\sigma = \Sexpr{fig2$params[3]}$, and $\delta = \Sexpr{fig2$params[4]}$. 
-See Introduction and Fig. \ref{fig:noise} caption for model details. \label{fig:pedagog}}
+\caption{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
+between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
+(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
+marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
+$\varphi^{-1}$, and 
+each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
+(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
+selected to match that of the original distribution (see Theory). (d) Two new variables, 
+$E_i^{||}$ and  $C_{i \bs i}^{||}$,
+%***DAN: I replaced \nat in the above two expressions by || because for some reason the \nat does not
+%compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
+which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
+which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
+$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
+For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
+$\log_{10}$ of probability density on all panels. 
+\label{fig:pedagog2}}
 \end{figure}

### 2023-01-26 paper/SupMat.Rnw
edited Theory section and caption of fig 2 and corresponding sections of sup mat

@@
 
-\section{Theory for the lottery model}\label{SIsect:TheoryForLottery}
-
-For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
-the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
-divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
-Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$, which is an increasing
-function of $E_i$ and decreasing function of $C_i$.
-
-When $i$ is an invader, the competition experienced by either 
-species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
-$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
-$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
-$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
-Likewise, 
-$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
-though we already knew this equals $0$. 
-An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
-by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
-Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
-$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
-$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
-$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
-We show in SI section \ref{SIsect:scaling_factors}
-that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
-Since we assumed log-normally distributed $B_i$, 
-$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
-We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SIsect:scaling_factors}.
-Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
-to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
-and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
-
-For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
-$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
-$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
-Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
-Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
-$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
-can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
-is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
-and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
-marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
-$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
-component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
-approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
-
 \section{Scaling factors}\label{SIsect:scaling_factors}
@@
 
-For the lottery model, for $i \neq j$, we have
-\begin{align}
-\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
-\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
-\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
-\end{align}
-Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
-$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
-in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
-but we instead compute $q_{ij}$ through implicit differentiation:
-\begin{align}
-\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
-&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
-\end{align}
-Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
-\begin{align}
-q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
-&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
-\end{align}
-as stated in the main text.
+%***DAN: I decided to present the general theory (not for the lottery model but for the
+%general case) in the Theory section in the main text, putting off the specific lottery
+%model application to later. For that, I need to explain scaling factors in general.
+%So this section is now for that explanation. That means the below text, which is about
+%the scaling factors for the lottery model need to be handled later. It is good, however,
+%that the below text is for the general lottery model, for either way of doing the 
+%fecundities. That will come in handy later.
+%
+% For the lottery model, for $i \neq j$, we have
+% \begin{align}
+% \sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
+% \sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
+% \sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
+% \end{align}
+% Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
+% $B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
+% in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
+% but we instead compute $q_{ij}$ through implicit differentiation:
+% \begin{align}
+% \frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
+% &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
+% \end{align}
+% Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
+% \begin{align}
+% q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
+% &= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
+% \end{align}
+% as stated in the main text.
 
-%***DAN: Change the below paragraph if needed. Some reasons it might be needed: Steve joins as a coauthor - easy
-%just means a change in wording at the top bit;
-%Steve reads this and thinks it makes sense to change it because something is factually or presentationally 
-%sub-optimal; or Dan or Jasmin actually develops a more rigorous understanding of scaling factors and based
-%on that thinks something here is sub-optimal. Anyway, the "do both" approach seems like it can't go wrong,
-%especially if Steve's prediction is correct that it won't make much difference.
-%***DAN: Dan, you probably need to spend a bit more time with the cited papers to better understand this. 
 We also use $q_{ij}=1$ as an alternative value, for the following reasons.
@@
 often not robust to arbitrary modelling choices. 
-Although the lottery model is not size structured and we do not empirically parameterize it, the above shortcomings of 
+Although the lottery model to which we apply our theory 
+is not size structured and we do not empirically parameterize it, the above shortcomings of 
 scaling factors, generally, suggest we should find alternatives to their use. The alternative 
@@
 
-We begin by decomposing 
-the invasion growth rate,
+We begin by decomposing the GWR,
 \begin{equation}
@@
 \end{equation}
-The invasion growth rate with both $E_i$ and $C_{i \bs i}$ constant at their means is 
+Replacing both $E_i$ and $C_{i \bs i}$ with their means gives 
 \begin{align}
@@
 \end{align}
-The additional contribution of variation in $E_i$ to the invasion growth rate is then
+The additional contribution of variation in $E_i$ to the GWR is then
 \begin{equation}
@@
 tertiary decomposition, i.e., in the decomposition of $\olep_i^{(EC)}$. 
-The term $\olep_i^{[EC]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on the invasion growth rate,
+The term $\olep_i^{[EC]}$ quantifies the influence of asymmetric tail associations of $E_i$ and $C_{i \bs i}$ on GWR,
 and the the term $\olep_i^{[E \nat C]}$ corresponds to $EC$ correlations \emph{per se}, after the removal of asymmetries of tail association.
@@
 asymmetric tail associations. The subscript $i$ on all these newly defined quantities $\Delta_i$ refers to species $i$.
+
+\section{Theory for the lottery model}\label{SIsect:TheoryForLottery}
+
+For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
+the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
+divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
+does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
+Species growth rates for the lottery model can then be written 
+$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$, which is an increasing
+function of $E_i$ and decreasing function of $C_i$.
+
+When $i$ is an invader, the competition experienced by either 
+species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
+$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
+$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
+$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
+Likewise, 
+$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
+though we already knew this equals $0$. 
+An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
+Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
+$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
+$(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
+$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
+We show in SI section \ref{SIsect:scaling_factors}
+that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
+Since we assumed log-normally distributed $B_i$, 
+$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
+We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SIsect:scaling_factors}.
+Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
+and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
+
+For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
+$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
+$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
+Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
+Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
+$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
+can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
+is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
+and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
+marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
+$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
+component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
+approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
+so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
 

### 2023-01-27 paper/Paper.Rnw
working on revisions to Methods and associated sections of sup mat

@@
 in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
+Notation for the whole paper is summarized in Table X.
+%***DAN: make the table
 For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
@@
 for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
-ATAs substantially influenced aspects of species coexistence (Fig. \ref{fig:pedagog1}d-f).
+ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).
 
@@
 are likely to be negligible or meaningfully large. 
-G3) To provide an empirical example of species coexistence which highlights the role ATAs can play.
-In the Discussion, we consider whether climate change may alter
+G3) To provide an empirical example of species coexistence which highlights the role ATAs can play and the
+potential importance of ATAs relative to other mechanisms of coexistence.
+The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
 ATAs of environmental variables and thereby influence competition and coexistence.
@@
 We here pursue goal G1 from the Introduction: to augment MCT to show formally how ATAs play a role in coexistence.
+
+\subsection{Theoretical setup}
+
 We develop our new theory for a general two-species model, later applying it to specific scenarios.
-MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
-quantify the contributions of 
-multiple mechanisms to coexistence. But we show below
-that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
-So we begin by defining storage effects. Our mathematical notation parallels
+Our mathematical notation parallels
 that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
@@
 an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
-competitive pressure on species $i$, $C_i(t)$. We also assume, for simplicity, 
+competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
@@
 
+MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+quantify the contributions of 
+multiple mechanisms to coexistence. But we show below
+that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
+So we begin by defining storage effects. 
 The storage effect for species $i$ is the contribution of covariation between $E$ and 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' [GWR] of species $i$, that is, the mean growth rate of 
+where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' (GWR) of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
-to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
-the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
-lottery model described in the Introduction.
+which is the extension of MCT which was goal G1 of the Introduction.
+
+\subsection{ATA rescue and exclusion}
+
+To evaluate the importance of ATAs for a given scenario,
+one can quantify
+the magnitude of the new ATA effect term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
+scenario.  One typically (but see below) must examine this quantity for each species, $i$,
+to obtain complete information about coexistence, since coexistence requires both species populations to 
+be able to grow when rare. We also defined scenarios of ``ATA rescue''
+and ``ATA exclusion''. \emph{ATA rescue} was considered to have occurred for a given scenario if coexistence
+was possible when ATAs were taken into account, i.e., GWR$>0$ for both species, but impossible otherwise, i.e.,
+the sum of the non-ATA terms in (\ref{eq:full_decomp}) was negative for at least one species. \emph{ATA exclusion} was 
+considered to have occurred if the species did not coexist when ATAs
+were taken into account, i.e., GWR$<0$ for at least one species, but would have coexisted were it not for ATAs, 
+i.e., the sum of the non-ATA terms in (\ref{eq:full_decomp})
+was positive for both species.
 
 \section{Methods}\label{section:methods}
+
+\subsection{Versions of the lottery model}
+
+To pursue goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
+comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
+of the other terms of 
+the decomposition (\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
+of the lottery model which we now specify. The versions both used the same dynamical equations
+(Introduction) and assumed temporally independent and identically distributed (iid) fecundities, but the
+versions differed in the distribution of $(B_1,B_2)$ used. Both distributions were based on 
+transformations of the distributions $(b_1,b_2)$ pictured in Fig. \ref{fig:pedagog1}a-c;
+the details of how the left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c), 
+and symmetrically (Fig. \ref{fig:pedagog1}b) tail associated alternatives for $(b_1,b_2)$ were generated 
+are in SI section \ref{SI-sect:noise}.
+
+The \emph{log-normal fecundities lottery model} used $B_i = \exp(\sigma+\mu_i b_i)$,
+and therefore had parameters $\sigma$, $\mu_1$ and $\mu_2$ in addition to the death rate parameter $\delta$
+that both versions of the lottery model use (Introduction). The parameter $\sigma$ is the variance of the
+log-fecundities and $\mu_i$ is the mean log fecundity for species $i$.
+
+The \emph{beta-fecundities lottery model}
+used $B_i = \eta_i F_\beta \circ \varphi^{-1}(b_i)$, where $\varphi$ is the CDF of the standard normal
+distribution and $F_\beta$ is the CDF of a beta distribution 
+with shape and scale parameters $0.5$ (probability density function proportional to $x^{-05}(1-x)^{-0.5}$ on
+the unit interval). So the beta-fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
+and produced fecundities with $U$-shaped distributions between $0$ and $\eta_1$, for $B_1$, and
+$\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$ and larger $\eta_i$ also 
+corresponded to more variable fecundities for species $i$.
+
+For both versions of the lottery model, 
+fecundities were taken to directly reflect 
+the environment, i.e., the $E_i$ of Theory was set equal to $B_i$. 
+Competition, $C_i(t)$, in the lottery model occurs at the stage of juveniles occupying open sites.
+For both model versions, $C_i$ was therefore taken to be the number of new offspring
+divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
+does not depend on $i$. These choices for $E$ and $C$ meet the assumptions of Theory (SI section Y).
+%***DAN: This SI reference is to the big section S$ listed below, but this material will be in the
+%first part of that section, prior to the subsections.
+
+\subsection{Plotting results for G2}
+
+%Then write a para or section that starts by saying we plotted the decomp terms against model parameters for
+%the log-normal-fecundity lottery model, and some mathematical shortcuts simplified it. Then summarize
+%the shortcuts.
+%This is similar to an existing para. Also cites a lot of sup mat stuff.
+
+%Then a similar para for the beta-fecundity lottery model.
+
+%Then a para about ATA exclusion and rescue and how that was assess for each model - different from
+%the current way of doing it with trait spaces - drop the trait space stuff.
+
+%For the sup mat, I think you will need these sections:
+%S3 - on generating the Fig. 1 noise (only)
+%S4 - as large section with subsections. About the theory for the two lottery models, but not about the effecient computation
+%   Starts with material on the general lottery model
+%   Then a sub-section with all the computations for the log-normal-fecundities lottery model
+%   Then a sub-section for the beta-fecundities lottery model
+%S5 - then a similar section, but for the efficient computations
+%There might be additional sections, see the current sections and see if those need to be maintained
+%as their own sections or can be absorbed into the above.
+%This is going to take some work.
+
+\subsection{Diatom chemostat system}
+
+%Then on to the diatom stuff.
 

### 2023-01-27 paper/SupMat.Rnw
working on revisions to Methods and associated sections of sup mat

@@
 
+\section{Noise} \label{sect:noise}
+
+To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left-tail association 
+case (Fig. \ref{MT-fig:pedagog1}a), we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal 
+distribution $N\left( \vec{0}, \Sigma \right)$, where $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$. 
+Then, for each index $i$, we randomly, with 50\% probability, either 
+(A) let $b_1^{(i)} = -|a_1^{(i)}|$ and $b_2^{(i)} = -|a_1^{(i)}|$, or 
+(B) let $b_1^{(i)} = |a_1^{(i)}|$ and $b_2^{(i)} = |a_2^{(i)}|$. 
+Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from 
+the fact that, in case (A) above, both $b_1^{(i)}$ and $b_2^{(i)}$ were assigned to the same value. 
+Right-tail associated noise (Fig. \ref{MT-fig:pedagog1}c) was generated in an analogous way.
+
+%***DAN: I think I may need the below later, in the computation section, but not here
+%For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct 
+%sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same 
+%$\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values 
+%differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
+
+When needed, noise with symmetric tail association (Fig. \ref{MT-fig:pedagog1}b) was generated by taking a 
+very large sample (\Sexpr{M}) 
+of left-tail associated noise generated as 
+described above, and then calculating the Pearson correaltion coefficient, 
+$\rho$, of the sample.
+Symmetric noise $\left(b_1^{(i)},b_2^{(i)}\right)$ for $i=1,...,M$ was then 
+generated 
+from the bivariate normal distribution $N\left(\vec{0}, \Sigma \right)$, where now
+$\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$.
+
+%***DAN: stopped here on 2023 01 27. See outline in the main text of how I need to restructure the 
+%SI sections from this point forward.
+
 \section{Theory for the lottery model}\label{SIsect:TheoryForLottery}
@@
 so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
-
-\section{Noise} \label{sect:noise}
-
-We created two cases for $(b_1,b_2)$, left- and right-tail association. 
-To generate $M$ points $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$, for the left tail association case, we first generated M points $(a_1^{(i)},a_2^{(i)})$, $i=1,...,M$, from a bivariate normal distribution $N\left( \vec{0}, \Sigma \right)$, where here $\Sigma = \begin{pmatrix} 1&0\\0&1\end{pmatrix}$.
-Then, for each index $i$, we randomly, with 50\% probability, either (A) let $\tilde b_1^{(i)} = -|a_1^{(i)}|$ and $\tilde b_2^{(i)} = -|a_1^{(i)}|$, or (B) let $\tilde b_1^{(i)} = |a_1^{(i)}|$ and $\tilde b_2^{(i)} = |a_2^{(i)}|$.
-We then let $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)} + \mu_2$. 
-Perfect association in the left tails of the resulting distribution $\left(b_1,b_2\right)$ results from the fact that, in case (A) above, both $\tilde b_1^{(i)}$ and $\tilde b_2^{(i)}$ were assigned to the same value.
-For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
-Right-tail associated noise was generated in an analogous way.
-
-When needed, noise with symmetric tail association was generated by taking a very large sample (\Sexpr{M}) 
-of left-tail associated noise generated as 
-described above, and then calculating the Pearson correaltion coefficient, $\rho$, of the sample.
-Symmetric noise $\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$ for $i=1,...,M$ was then generated 
-from the bivariate normal distribution $N\left(\vec{0}, \Sigma \right)$, where $\Sigma = \begin{pmatrix} 1&\rho \\ \rho&1\end{pmatrix}$, 
-and then we set $b_1^{(i)} = \sigma \tilde b_1^{(i)} + \mu_1$ and $b_2^{(i)} = \sigma \tilde b_2^{(i)}+ \mu_2$.
-
-Overall, our version of the lottery model was completely specified by specifying $\delta$, $\mu_1$, $\mu_2$, $\sigma$ and whether 
-noise was generated from the symmetric, left- or right-tail associated cases.
 

### 2023-01-31 code/theory_panel_fxn.R
changes for using TeX

@@ -28,12 +28,14 @@ theoryfigpanel <- function(d,p,q,cop,x,y, gnrt_var=FALSE, samps=NULL, oldp=NULL,
 	  if (log_points==T){
 	    z[z==0]<-NA; zvar <- log10(z)
 	  } else { zvar <- z }
+	  
 	  z_ran <- range(zvar, na.rm=T)
 	  zbin <- .bincode(zvar, breaks=seq(z_ran[1],z_ran[2],len=ncol+1),TRUE,TRUE)
 	  zbin <- matrix(zbin, length(x), length(y))
 	  cols <- hcl.colors(ncol, palette = "YlOrRd",rev=T)
 	  xbin <- .bincode(samps[,1], breaks=seq(min(x),max(x),len=nrow(z)+1),TRUE,TRUE)
 	  ybin <- .bincode(samps[,2], breaks=seq(min(y),max(y),len=ncol(z)+1),TRUE,TRUE)
+	  
 	  colindex <- {}
 	  for (i in 1:nrow(samps)) {colindex[i]<-zbin[xbin[i],ybin[i]]}
 	  rgb<-sapply(cols[colindex],FUN=col2rgb)
@@ -45,8 +47,8 @@ theoryfigpanel <- function(d,p,q,cop,x,y, gnrt_var=FALSE, samps=NULL, oldp=NULL,
 	  points(samps[,1],samps[,2],type="p",pch=20, cex=1, col=rgb(.74,.74,.74,.4))
 	}
 	contour(x,y,log10(z),nlevels=10, bty='l', add=T)
-	pdfhist(samps[,1],d,...)
-	pdfhist(samps[,2],d,horiz=TRUE,...)
+	pdfhist(samps[,1],x,y,d)
+	pdfhist(samps[,2],x,y,d,horiz=TRUE)
 	plot.new() 
 	return(samps)
 }
@@ -56,7 +58,7 @@ tdis <- function(q, oldp, samps){
 	samps <- q(oldp(samps))
 }
 
-pdfhist <- function(samps, d, horiz=FALSE,xpd=FALSE){
+pdfhist <- function(samps,x,y,d, horiz=FALSE,xpd=FALSE){
 	min <- min(x)
 	max <- max(y)
 	h <- hist(samps, seq(min(samps),max(samps),length.out=10), plot=FALSE)

### 2023-01-31 code/theoryfig_col_meat.R

@@ -10,33 +10,33 @@ by<-seq(from=-0.5,to=0.5,by=0.01)
 # A BETA MARGS CLAYTON COP
 sampsA <- theoryfigpanel(dbmarg, pbmarg, qbmarg, ccop, bx, by, gnrt_var=TRUE,col=T,log_points=T)
 #text(x=lblc["E","x"], y=lblc["E","y"], labels=expression('E'['i']), cex=2)
-text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX(r'($E_i$)'), cex=2)
+text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('$E_i$'), cex=2)
 #text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\i']), srt=90, cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX(r'($C_{i\setminus i}$)'), srt=90, cex=2)
+text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($C_{i\\setminus i}$)'), srt=90, cex=2)
 text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[1],")"),cex=2)
 
 # B NORM MARGS CLAY COP
 sampsB <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ccop, nx, ny, samps=sampsA, oldp=pbmarg,xpd=NA,col=T, log_points=FALSE)
 #text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression(phi^-1*'o'*'F'['E'['i']]("E"["i"]))), cex=2)
-text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX(r'($\varphi^{-1} \circ F_{E_i}(E_i)$)'), cex=2)
+text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('($\\varphi^{-1} \\circ F_{E_i}(E_i)$)'), cex=2)
 #text(x=lblc["C","x"], y=lblc["C","y"], labels=c(expression(phi^-1*'o'*'F'['C'['i/i']]("C"["i/i"]))), srt=90, cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX(r'($\varphi^{-1} \circ F_{C_{i\setminus i}}(C_{i\setminus i})$)'), srt=90, cex=2)
[…]

### 2023-01-31 paper/SupMat.Rnw
commit of ongoing edits in order to do a pull

@@
 %header material for paper
-\title{Asymmetric relationships and their impact on coexistence: Supporting information}
+\title{Asymmetric relationships and their effects on coexistence: Supporting information}
 \date{}
@@
 
-%***DAN: I decided to present the general theory (not for the lottery model but for the
-%general case) in the Theory section in the main text, putting off the specific lottery
-%model application to later. For that, I need to explain scaling factors in general.
-%So this section is now for that explanation. That means the below text, which is about
-%the scaling factors for the lottery model need to be handled later. It is good, however,
-%that the below text is for the general lottery model, for either way of doing the 
-%fecundities. That will come in handy later.
-%
-% For the lottery model, for $i \neq j$, we have
-% \begin{align}
-% \sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
-% \sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
-% \sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
-% \end{align}
-% Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
-% $B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$ 
-% in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
-% but we instead compute $q_{ij}$ through implicit differentiation:
-% \begin{align}
-% \frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
-% &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
-% \end{align}
-% Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so 
-% \begin{align}
-% q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
-% &= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta},
-% \end{align}
-% as stated in the main text.
-
 We also use $q_{ij}=1$ as an alternative value, for the following reasons.
@@
 We first review the decompositions of \cite{Ellner_2016} and \cite{Ellner_2019}, which are similar to and motivated by the 
-original decompositions of \cite{Chesson_1994}; and then we include our extensions of the decomposition.
+original decompositions of \cite{Chesson_1994}; and then we include our extension of the decomposition.
 
@@
 
-%***DAN: stopped here on 2023 01 27. See outline in the main text of how I need to restructure the 
-%SI sections from this point forward.
-
 \section{Theory for the lottery model}\label{SIsect:TheoryForLottery}
 
-For the lottery model, we take $E_i(t)=B_i(t)$, henceforth assuming that fecundity depends strictly on
-the environment. Competition, $C_i(t)$, in the lottery model is the number of new offspring
-divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. Competition in the lottery model occurs at the stage of juveniles occupying open sites. 
-Species growth rates for the lottery model can then be written 
-$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)]$, which is an increasing
-function of $E_i$ and decreasing function of $C_i$.
+\subsection{Theory that applies to both versions of the model}\label{SIsect:TheoryForLotteryBoth}
+
+Recall we took $E_i(t)=B_i(t)$, thereby assuming that fecundity depends strictly on
+the environment for the lottery model (both the log-normal fecundities and the beta fecundities versions of the model). 
+Recall also that competition, $C_i(t)$, is 
+$$C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N)$$ (again for both versions of the model), which
+does not depend on $i$. Species growth rates for the lottery model, $r_i(t)=\ln(N_{i}(t+1)/N_i(t))$, can then be written 
+$$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)],$$ 
+which is an increasing
+function of $E_i$ and decreasing function of $C_i$, as was assumed at the outset of Theory.
 
 When $i$ is an invader, the competition experienced by either 
-species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$; and likewise to get
-$\olr_{i \bs i}$ from the expression for $r_i$ for the model. Thus 
-$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta$, and 
-$\olr_{i \bs i} = \E[\ln(1-\delta+\delta B_i/B_j)]=\E[\ln(1-\delta+E_i /C_{i \bs i})]$. 
+species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$. Thus 
+$$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta.$$
+Lottery-model values of all the expressions of SI section \ref{SIsect:other_mechanisms} can be obtained 
+from the above expression for $C_{i \bs i}$ and the assumption that $E_i(t)=B_i(t)$.
+For instance, 
+$$\olr_{i \bs i} = \E[r_i(E_i,C_{i \bs i})] = \E[\ln(1-\delta+E_i /C_{i \bs i})] = \E[\ln(1-\delta+\delta B_i/B_j)].$$ 
 Likewise, 
-$\olr_{j \bs i} = \E[\ln(1-\delta+\delta B_j/B_j)]=\E[\ln(1-\delta+E_j/C_{j \bs i})]$,
-though we already knew this equals $0$. 
+$$\olr_{j \bs i} = \E[r_j(E_j,C_{j \bs i})] = \E[\ln(1-\delta+E_j/C_{j \bs i})] = \E[\ln(1-\delta+\delta B_j/B_j)]=0,$$
+though we already knew, \emph{a priori}, that this equals $0$. 
 An expression for $\olrs_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
 by a random variable with the same marginals, but independent components, $(B_i^\shp,B_j^\shp /\delta)$.
-Thus $\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)]$. Likewise,
-$\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we use 
+Thus 
+$$\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)].$$ 
+Likewise, $\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we can use 
 $(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
-$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)]$.
-We show in SI section \ref{SIsect:scaling_factors}
-that $q_{ij}=\overline{B}_i/((1-\delta)\overline{B}_j+\delta \overline{B}_i)$. 
-Since we assumed log-normally distributed $B_i$, 
-$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
-We alternatively also consider $q_{ij}=1$, for reasons explained in SI section \ref{SIsect:scaling_factors}.
-Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
-to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
-and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
+$$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)].$$
+An expression for $\olrn_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
+by a random variable with the same marginals, but with ATAs removed, $(B_i^\nat,B_j^\nat /\delta)$.
+Thus 
+$$\olrn_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\nat/B_j^\nat)].$$ 
+Likewise, $\olrn_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})=(B_j,B_j/\delta)$ in an analogous way.
+But note that $B_j$ and $B_j/\delta$ are linearly related for the lottery model, so their association is 
+already symmetric, so
+so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
+This type of reasoning is applied to the log-normal fecundities and beta fecundities versions of the 
+lottery model, specifically, in the subsequent sections.
 
-For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
-$\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
-$\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
-Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
-Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
-$\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
-can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
-is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
-and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
-marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
-$(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
-component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
-approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
+%scaling factors
+We now consider the scaling factors $q_{ij}$ for the lottery model. 
+For the lottery model, for $i \neq j$, we have
+\begin{align}
+\sC_j &= -\ln \left[ 1-\delta+\delta N \frac{\overline{B}_j}{B_1(t) N_1 (t) + B_2(t) N_2(t)}  \right]  \\
+\sC_{j \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_j}{B_j(t)} \right] \\
+\sC_{i \bs i} &= -\ln \left[ 1-\delta+\delta  \frac{\overline{B}_i}{B_j(t)} \right].
+\end{align}
+Both $\sC_{i \bs i}$ and $\sC_{j \bs i}$ are strictly monotonic functions of $B_j$, so the function
+$B_j \mapsto \sC_{j \bs i}$ is invertible, and using its inverse provides an expression of $\sC_{i \bs i}$
+in terms of $\sC_{j \bs i}$. One could write this expression explicitly and then differentiate it to calculate $q_{ij}$,
+but we instead compute $q_{ij}$ through implicit differentiation:
+\begin{align}
+\frac{\partial \sC_{i \bs i}}{\partial \sC_{j \bs i}} &= \left. \frac{\partial \sC_{i \bs i}}{\partial B_j} \middle/ \frac{\partial \sC_{j \bs i}}{\partial B_j} \right. \\
+&= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{B_j}}{1-\delta+\delta \frac{\overline{B}_i}{B_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right).
+\end{align}
+Solving $\sC_{j \bs i}=0$ gives $B_j=\overline{B}_j$, so
+\begin{align}
+q_{ij} &= \left( \frac{1-\delta +\delta \frac{\overline{B}_j}{\overline{B}_j}}{1-\delta+\delta \frac{\overline{B}_i}{\overline{B}_j}} \right) \left( \frac{\overline{B}_i}{\overline{B}_j}  \right) \\
+&= \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
+\end{align}
+This expression is made specific to the log-normal decundities and beta fecundities versions of the lottery
+model in the subsequent sections. We alternatively also consider $q_{ij}=1$, for reasons explained in SI 
+section \ref{SIsect:scaling_factors}.
 
-\section{Mathematical simplifications for plotting}\label{sect:math_simplifications}
+%***DAN: Might need this text below, in the section on the log-normal fecundities lottery model, but not here
+%Since we assumed log-normally distributed $B_i$, 
+%$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
+%Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
+%to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
+%and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
 
-We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the lottery model. 
+%***DAN: We might need this text somewhere, but not sure where. Just saving it out of caution.
+% For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
+% $\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
+% $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
+% Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
+% Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
+% $\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
+% can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
+% is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
+% and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
+% marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
+% $(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
+% component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
+% approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
+% so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
+
+\subsection{Theory and plotting simplifications for the log-normal fecundities lottery model}\label{SIsect:TheoryForLotteryLogNorm}
+
+%The first of the two growth rate decompositions
+We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the 
+log-normal fecundities lottery model. 
 One can straightforwardly show from the formulas of SI section \ref{SIsect:other_mechanisms} that
 \begin{equation}
-E_i=B_i=\exp(b_i)
+E_i=B_i=\exp(\sigma b_i +\mu_i)
 \end{equation}
 \begin{equation}
-C_i=\frac{B_1 N_1 + B_2 N_2}{\delta N}
-\end{equation}
-\begin{equation}
-C_{i \bs i} = B_j/\delta=\exp(b_j)/\delta
-\end{equation}
-\begin{equation}
-r_i=\ln(1-\delta+E_i/C_i)
+C_{i \bs i} = B_j/\delta=\exp(\sigma b_j + \mu_j)/\delta
 \end{equation}
@@
 \begin{align}
-\olr_{i \bs i} &= \E[\ln(1-\delta+E_i/C_{i \bs i})] \\
-&= \E[\ln(1-\delta+\delta B_i/B_j)] \\
-&= \E[\ln(1-\delta+\delta\exp(b_i-b_j))]\label{eq:base_IGR}
+\olr_{i \bs i} &= \E[\ln(1-\delta+\delta B_i/B_j)] \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]\label{eq:base_IGR}
 \end{align}
@@
 \begin{align}
-\olep_i^E &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i+\mu_i)/\exp(\mu_j+\sigma^2/2))]-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}
+\olep_i^E &= \E[\ln(1-\delta+\delta\exp(\sigma b_i + \mu_i)/\exp(\mu_j + \sigma^2/2))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma b_i + \mu_i - \mu_j - \sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_E}
 \end{align}
-where $\tilde{b}_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+where $b_i$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
 %contrib of variation in C
 \begin{align}
-\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(b_j))]-\epsilon_i^0 \\
-&= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma \tilde{b}_j +\mu_j))]-\epsilon_i^0 \\
-&= \E[\ln(1-\delta + \delta \exp(-\sigma \tilde{b}_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 
+\olep_i^C &= \E[\ln(1-\delta + \delta \exp(\mu_i+\sigma^2/2)/\exp(\sigma b_j + \mu_j))]-\epsilon_i^0 \\
+&= \E[\ln(1-\delta + \delta \exp(-\sigma b_j +\mu_i -\mu_j+\sigma^2/2))]-\epsilon_i^0, \label{eq:decomp:contrib_of_var_C} 
 \end{align}
-where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+where $b_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
 %last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
 \begin{align}
-\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(b_i)/\exp(b_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i +\mu_i)/\exp(\sigma \tilde{b}_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}
+\olep_i^{EC} &= \E[\ln(1-\delta+\delta\exp(\sigma b_i +\mu_i)/\exp(\sigma b_j +\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma (b_i - b_j)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:last_term_primary}
 \end{align}
-where $(\tilde{b}_i, \tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
-$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$. Continuing,
+where $(b_i, b_j)$ is distributed as in SI section \ref{sect:noise}. Continuing,
 %contrib of variation in E and C, per se  
 \begin{align}
-\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(b_i^\shp)/\exp(b_j^\shp))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma \tilde{b}_i^\shp +\mu_i)/\exp(\sigma \tilde{b}_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}
+\olep_i^{(E\shp C)} &= \E[\ln(1-\delta+\delta\exp(\sigma b_i^\shp +\mu_i)/\exp(\sigma b_j^\shp + \mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i^\shp - b_j^\shp)+\mu_i-\mu_j))]-\olep_i^E-\olep_i^C-\epsilon_i^0, \label{eq:decomp:contrib_of_EC_var_per_se}
 \end{align}
-where $\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
+where $b_i^\shp$ and $b_j^\shp$ are standard normally distributed and independent. Continuing,
 %storage effects
 \begin{equation}
-\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i - \tilde{b}_j)+\mu_i-\mu_j))]-
-\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}
+\olep_i^{(EC)} = \E[\ln(1-\delta+\delta\exp(\sigma (b_i - b_j)+\mu_i-\mu_j))]-
+\E[\ln(1-\delta+\delta\exp(\sigma (b_i^\shp - b_j^\shp)+\mu_i-\mu_j))],\label{eq:decomp:storage_effects}
 \end{equation}
-where $(\tilde{b}_i,\tilde{b}_j)$ is distributed as in SI section \ref{sect:noise}, i.e., 
-$(\tilde{b}_i,\tilde{b}_j)=\left(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$, and 
-$\tilde{b}_i^\shp$ and $\tilde{b}_j^\shp$ are standard normal and independent. Continuing,
+where $(b_i,b_j)$ is distributed as in SI section \ref{sect:noise}, and 
+$b_i^\shp$ and $b_j^\shp$ are standard normal and independent. Continuing,
 %contrib of ATA
 \begin{equation}
-\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))],\label{eq:decomp:contrib_ATA_1}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]-
+\E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j))],\label{eq:decomp:contrib_ATA_1}
 \end{equation}
-where $(\ddot{b}_i,\ddot{b}_j)$ have the same (normal) marginals and covariance as $(b_i,b_j)$ but are 
-bivariate normal. Thus
-\begin{equation}
-\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i-\tilde{b}_j)+\mu_i-\mu_j))]-
-\E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))], \label{eq:decomp:contrib_ATA_2}
-\end{equation}
-where $(\tilde{b}_i,\tilde{b}_j)=\left( \frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma} \right)$ and 
-$(\tilde{\ddot{b}}_i,\tilde{\ddot{b}}_j)=\left( \frac{\ddot{b}_i-\mu_i}{\sigma},\frac{\ddot{b}_j-\mu_j}{\sigma} \right)$.
-Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_2})
+where $(b_i^\nat,b_j^\nat)$ have the same (standard normal) marginals and covariance as $(b_i,b_j)$ but are 
+bivariate normal. Note, a choice has been made here: we could also have chosen $(b_i^\nat,b_j^\nat)$
+to have the same standard normal marginals as $(b_i,b_j)$, but with covariance chosen such that
+the Pearson, Spearman, or Kendall correlation of $E_i^\nat = \exp(\sigma b_i^\nat +\mu_i)$
+and $C_{i \bs i}^\nat = \exp(\sigma b_j^\nat +\mu_j)/\delta$ matches that of $E_i = \exp(\sigma b_i +\mu_i)$
+and $C_{i \bs i} = \exp(\sigma b_j +\mu_j)$. But the choice we made seems 
+the most natural for the log-normal fecundities lottery models (and also for the 
+beta fecundities lottery model - see below). 
+Continuing, $\olep_i^{[E \nat C]}=\olep_i^{(EC)}-\olep_i^{[EC]}$ is available by subtracting expression (\ref{eq:decomp:contrib_ATA_1})
 from expression (\ref{eq:decomp:storage_effects}), i.e., 
 \begin{equation}
-\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i-\tilde{\ddot{b}}_j)+\mu_i-\mu_j))] - 
-\E[\ln(1-\delta+\delta\exp(\sigma (\tilde{b}_i^\shp - \tilde{b}_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
+\olep_i^{[E \nat C]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j))] - 
+\E[\ln(1-\delta+\delta\exp(\sigma (b_i^\shp - b_j^\shp)+\mu_i-\mu_j))].\label{eq:decomp:contrib_EC}
 \end{equation}
-This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ for the lottery model.
+This completes the calculation of all the terms of the decomposition (\ref{eq:decomp_overall_IGR})  of $\olr_{i \bs i}$ 
+for the log-normal fecundities lottery model.
 
+%The second of the two growth rate decompositions
 We next decompose $\olr_{j \bs i}$ for the lottery model. We have 

### 2023-01-31 code/theoryfig_col_meat.R
fiddly changes to math in a figure axis

@@ -11,14 +11,14 @@ by<-seq(from=-0.5,to=0.5,by=0.01)
 sampsA <- theoryfigpanel(dbmarg, pbmarg, qbmarg, ccop, bx, by, gnrt_var=TRUE,col=T,log_points=T)
 #text(x=lblc["E","x"], y=lblc["E","y"], labels=expression('E'['i']), cex=2)
 text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('$E_i$'), cex=2)
-#text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\i']), srt=90, cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($C_{i\\setminus i}$)'), srt=90, cex=2)
+text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\\i']), srt=90, cex=2)
+#text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($C_{i \\ i}$)'), srt=90, cex=2)
 text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[1],")"),cex=2)
 
 # B NORM MARGS CLAY COP
 sampsB <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ccop, nx, ny, samps=sampsA, oldp=pbmarg,xpd=NA,col=T, log_points=FALSE)
 #text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression(phi^-1*'o'*'F'['E'['i']]("E"["i"]))), cex=2)
-text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('($\\varphi^{-1} \\circ F_{E_i}(E_i)$)'), cex=2)
+text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('$\\varphi^{-1}(F_{E_i}(E_i))$'), cex=2)
 #text(x=lblc["C","x"], y=lblc["C","y"], labels=c(expression(phi^-1*'o'*'F'['C'['i/i']]("C"["i/i"]))), srt=90, cex=2)
 text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($\\varphi^{-1} \\circ F_{C_{i\\setminus i}}(C_{i\\setminus i})$)'), srt=90, cex=2)
 text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[2],")"), cex=2)

### 2023-01-31 code/theoryfig_col_meat.R
edits

@@ -12,7 +12,7 @@ sampsA <- theoryfigpanel(dbmarg, pbmarg, qbmarg, ccop, bx, by, gnrt_var=TRUE,col
 #text(x=lblc["E","x"], y=lblc["E","y"], labels=expression('E'['i']), cex=2)
 text(x=lblc["E","x"], y=lblc["E","y"], labels=TeX('$E_i$'), cex=2)
 #text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\i']), srt=90, cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($C_{i\\setminus i}$)'), srt=90, cex=2)
+text(x=lblc["C","x"], y=lblc["C","y"], labels=TeX('($C_{i\\backslash i}$)'), srt=90, cex=2)
 text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste("(",letters[1],")"),cex=2)
 
 # B NORM MARGS CLAY COP

### 2023-01-31 paper/Paper.Rnw
small changes

@@
 %header material for paper
-\title{Asymmetric relationships and their impact on coexistence}
+\title{Asymmetric relationships and their effects on coexistence}
 \date{}
@@
 The \emph{scaling factor}
-$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
+$q_{ij}$ measures the relative sensitivities to competition of the two species. See \cite{Chesson_1994},
 \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
-alternative definitions of $q_{ij}$ that we use. 
+alternative definitions of $q_{ij}$ that we use, but one definition is $q_{ij}=1$, and to understand the new
+ideas presented here it is sufficient to use that definition. 
 Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
@@
 \end{equation}
-which is the extension of MCT which was goal G1 of the Introduction.
+which is the extension of MCT that was goal G1 of the Introduction.
 
@@
 
+%Setup - two versions, and basics which apply to both
 To pursue goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
@@
 (Introduction) and assumed temporally independent and identically distributed (iid) fecundities, but the
-versions differed in the distribution of $(B_1,B_2)$ used. Both distributions were based on 
+versions differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
 transformations of the distributions $(b_1,b_2)$ pictured in Fig. \ref{fig:pedagog1}a-c;
@@
 
+%the log-normal fecundities version
 The \emph{log-normal fecundities lottery model} used $B_i = \exp(\sigma+\mu_i b_i)$,
 and therefore had parameters $\sigma$, $\mu_1$ and $\mu_2$ in addition to the death rate parameter $\delta$
-that both versions of the lottery model use (Introduction). The parameter $\sigma$ is the variance of the
+(Introduction) that both versions of the lottery model used. The parameter $\sigma$ is the variance of the
 log-fecundities and $\mu_i$ is the mean log fecundity for species $i$.
 
+%the beta fucundities version
 The \emph{beta-fecundities lottery model}
-used $B_i = \eta_i F_\beta \circ \varphi^{-1}(b_i)$, where $\varphi$ is the CDF of the standard normal
+used $B_i = \eta_i F_\beta^{-1} \circ \varphi(b_i)$, where $\varphi$ is the CDF of the standard normal
 distribution and $F_\beta$ is the CDF of a beta distribution 
@@
 the unit interval). So the beta-fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
-and produced fecundities with $U$-shaped distributions between $0$ and $\eta_1$, for $B_1$, and
-$\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$ and larger $\eta_i$ also 
+and produced fecundities with $U$-shaped, beta distributions between $0$ and $\eta_1$, for $B_1$, and
+$\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
 corresponded to more variable fecundities for species $i$.
 
+%more stuff that applies to both
 For both versions of the lottery model, 
@@
 divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. These choices for $E$ and $C$ meet the assumptions of Theory (SI section Y).
-%***DAN: This SI reference is to the big section S$ listed below, but this material will be in the
-%first part of that section, prior to the subsections.
+does not depend on $i$. These choices for $E_i(t)$ and $C_i(t)$ fulfill the assumption of Theory 
+that the growth rate, $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
+(see SI section \ref{SI-SIsect:TheoryForLottery} for details).
 
-\subsection{Plotting results for G2}
+\subsection{How results were plotted for G2}
 
-%Then write a para or section that starts by saying we plotted the decomp terms against model parameters for
-%the log-normal-fecundity lottery model, and some mathematical shortcuts simplified it. Then summarize
-%the shortcuts.
-%This is similar to an existing para. Also cites a lot of sup mat stuff.
+%Plotting shortcuts for the log-normal fecundities model
+To explore, using the log-normal fecundities lottery model, the relative magnitudes of the contributions 
+to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
+(\ref{eq:full_decomp}) for that model against the model parameters $\mu_1$, $\mu_2$, $\sigma$ and $\delta$.
+We labeled regions of the plot which corresponded to ATA rescue or exclusion, i.e., regions for which
+GWR and GWR-$\Delta_i^{[EC]}$ had opposite signs.
+Some mathematical shortcuts, summarized here, simplified plotting. 
+First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
+it was then sufficient to decompose GWR of the weaker competitor, species 1. 
+Second, we showed that, for the log-normal fecundities
+lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
+instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). 
+%***DAN: Not sure if this will the right citation anymore after we restructure those parts of the SI
+So we plotted against $\mu_1 - \mu_2$. 
+Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
+for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). 
+%***DAN: Also check this citation
+So we only generated plots 
+for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6$
+and $\mu_1-\mu_2=0,-0.5,-2,-4$,
+we plotted the components of (\ref{eq:full_decomp}) against $\sigma$. 
+%***DAN: Jasmin mentioned that she only used \delta=0.2, .4, .6 because of a comment Ellner made where he 
+%said it was known that larger values of \delta made for very little storage effects. Read that comment 
+%again and then write something here to jstify the choice. I think I can probably then delete the section of
+%sup mat, mentioned in the commented text below, that talks about \delta=1.
+%Mathematical simplications for the $\delta=1$ case make it straightforward to see that
+%coexistence is impossible and that the case is relatively trivial (SI 
+%section \ref{SI-sect:delta1}), so that case is not plotted.  
+The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
+Computationally efficient means of performing these simulations are in 
+SI section \ref{SI-sect:efficient_computation}.
+%***DAN: check the citation after restructuring the SI
 
 %Then a similar para for the beta-fecundity lottery model.
-
-%Then a para about ATA exclusion and rescue and how that was assess for each model - different from
-%the current way of doing it with trait spaces - drop the trait space stuff.
+To explore, using the beta fecundities lottery model, the relative magnitudes of the contributions 
+to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
+(\ref{eq:full_decomp}) for that model against the model parameters $\eta_1$, $\eta_2$, and $\delta$.
+%***DAN: finish it
 
@@
 %Then on to the diatom stuff.
-
-%G2
-Left- or right-tail associated noise $(b_1,b_2)$, with $b_i$ normally distributed with mean $\mu_i$ and standard deviation 
-$\sigma$, was generated as pictured in Fig. \ref{fig:noise}, details in SI section \ref{SI-sect:noise}.
-To achieve goal G2 from the Introduction, i.e., to explore using a simple model the relative magnitudes of the contributions 
-to coexistence that come from ATAs versus other mechanisms, we plotted the terms $\Delta_i^0$, $\Delta_i^E$, $\Delta_i^C$, 
-$\Delta_i^{(E\shp C)}$, $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ [see (\ref{eq:full_decomp})] for the lottery model
-against the parameters of the model, $\mu_1$, $\mu_2$, $\sigma$ and $\delta$; and we compared 
-values on plots. Some mathematical shortcuts, summarized here, simplified plotting. 
-First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
-it was then sufficient to decompose the GWR of the weaker competitor, species 1. 
-Second, we showed that, for the 
-lottery model, none of the components in (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). So we plotted 
-against $\mu_1 - \mu_2$. 
-Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
-for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). So we only generated plots 
-for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6, 0.8$
-and $\mu_1-\mu_2=0,-0.5,-2,-4$,
-we plotted all the components of (\ref{eq:full_decomp}) against $\sigma$. 
-%Relative importance of contributions was judged from the 
-%plots to assess whether and to what extent ATA effects on coexistence were meaningfully large for the lottery model, compared to 
-%the contributions of other mechanisms.
-Mathematical simplications for the $\delta=1$ case make it straightforward to see that
-coexistence is impossible and that the case is relatively trivial (SI 
-section \ref{SI-sect:delta1}), so that case is not plotted.  
-The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
-Computationally efficient means of performing these simulations, for the lottery model, are in 
-SI section \ref{SI-sect:efficient_computation}.
 
@@
 
-%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
-%structure it is generated and stored, and after I have re-run Jasmin's code
 \begin{figure}
-\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
+\includegraphics[width=\textwidth]{../results_figs/fig1_a-h.pdf}
 \caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition

### 2023-01-31 code/makTheoryFigs.R
renaming

@@ -33,8 +33,6 @@ dnmarg<-function(x){return(dnorm(x,mean=0,sd=1))}
 pnmarg<-function(x){return(pnorm(x,mean=0,sd=1))}
 qnmarg<-function(x){return(qnorm(x,mean=0,sd=1))}
 
-x<-seq(from=-4,to=4,by=0.01)
-y<-seq(from=-4,to=4,by=0.01)
 
 shape1<-0.5;shape2<-0.5
 

### 2023-01-31 code/makTheoryFigs.R
following through with rename

@@ -57,7 +57,7 @@ ly_mat<- matrix(c (	2,4,17,6,8,17,10,12,17,14,16,
 pdf(theoryfig_col, height=4, width=15)
 layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
 par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
-source("./theoryfig_col_meat.R")
+source("./theoryFig_script.R")
 dev.off() 
 ########
 

### 2023-02-01 paper/Paper.Rnw
continued progress editing Methods and accompanying SI sections (mostly working on the SI)

@@
 lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-sect:math_simplifications}). 
-%***DAN: Not sure if this will the right citation anymore after we restructure those parts of the SI
+instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
 So we plotted against $\mu_1 - \mu_2$. 
 Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
-for both our left- and right-tail associated noises (SI section \ref{SI-sect:math_simplifications}). 
-%***DAN: Also check this citation
+for both our left- and right-tail associated noises (SI section \ref{SI-SIsect:TheoryForLottery}). 
 So we only generated plots 
@@
 SI section \ref{SI-sect:efficient_computation}.
-%***DAN: check the citation after restructuring the SI
 
@@
 %***DAN: finish it
-
-%For the sup mat, I think you will need these sections:
-%S3 - on generating the Fig. 1 noise (only)
-%S4 - as large section with subsections. About the theory for the two lottery models, but not about the effecient computation
-%   Starts with material on the general lottery model
-%   Then a sub-section with all the computations for the log-normal-fecundities lottery model
-%   Then a sub-section for the beta-fecundities lottery model
-%S5 - then a similar section, but for the efficient computations
-%There might be additional sections, see the current sections and see if those need to be maintained
-%as their own sections or can be absorbed into the above.
-%This is going to take some work.
+%***DAN: Next steps as of 2023 02 01:
+%Fill in the sup mat section for the theory for the beta fecundities lottery model
+%fill in the efficient computation section for that same model
+%write this section, here, in the main text
 

### 2023-02-01 paper/SupMat.Rnw
continued progress editing Methods and accompanying SI sections (mostly working on the SI)

@@
 
-%***DAN: Might need this text below, in the section on the log-normal fecundities lottery model, but not here
-%Since we assumed log-normally distributed $B_i$, 
-%$\overline{B}_i=\exp(\mu_i+\sigma^2/2)$, so $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$.
-%Thus $\Delta_i^{(EC)}$ for the lottery model can be computed by using simulations \citep{Ellner_2016}
-%to very accurately approximate $\olr_{i \bs i}$, $\olrs_{i \bs i}$,
-%and $\olrs_{j \bs i}$, and then plugging them and $q_{ij}$ into main text equation (\ref{MT-eq:storage_effects_gen}).
-
 %***DAN: We might need this text somewhere, but not sure where. Just saving it out of caution.
@@
 %The second of the two growth rate decompositions
-We next decompose $\olr_{j \bs i}$ for the lottery model. We have 
+We next decompose $\olr_{j \bs i}$ for the log-normal fecundities lottery model [see (\ref{eq:decomp_overall_rjbsi})]. We have 
 \begin{equation}
-E_j = B_j = \exp(b_j)
+E_j = B_j = \exp(\sigma b_j + \mu_j)
 \end{equation}
 \begin{equation}
-C_j=\frac{B_1 N_1 + B_2 N_2}{\delta N}
-\end{equation}
-\begin{equation}
-C_{j \bs i} = B_j/\delta = \exp(b_j)/\delta
-\end{equation}
-\begin{equation}
-r_j=\ln(1-\delta+E_j/C_j)
+C_{j \bs i} = B_j/\delta = \exp(\sigma b_j + \mu_j)/\delta
 \end{equation}
@@
 &= \E[\ln(1-\delta+\delta B_j/\overline{B}_j)]-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma \tilde{b}_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
-&= \E[\ln(1-\delta+\delta \exp(\sigma \tilde{b}_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}
+&= \E \left[ \ln \left(1-\delta+\delta\frac{\exp(\sigma b_j + \mu_j)}{\exp(\mu_j+\sigma^2/2)} \right) \right]-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(\sigma b_j - \sigma^2/2))]-\epsilon_j^0, \label{eq:decomp:contrib_of_var_E_j}
 \end{align}
-where $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+where $b_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
 %contrib of variation in C
@@
 &= \E \left[ \ln \left(1-\delta+\delta \frac{\overline{B}_j}{B_j} \right) \right]-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(b_j)} \right) \right]-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma \tilde{b}_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
-&= \E[\ln(1-\delta+\delta \exp(-\sigma \tilde{b}_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\mu_j+\sigma^2/2)}{\exp(\sigma b_j +\mu_j)} \right) \right]-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(-\sigma b_j+\sigma^2/2))]-\epsilon_j^0,\label{eq:decomp:contrib_of_var_C_j}
 \end{align}
-where again $\tilde{b}_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
+where again $b_j$ is standard normally distributed (see SI section \ref{sect:noise}). Continuing,
 %last term of primary decomp, later splits into storage effects and effect of variation in E and C, per se  
@@
 &= \E \left[ \ln \left(1-\delta+\delta \frac{B_j^\shp}{B_j} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(b_j^\shp)}{\exp(b_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
-&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma \tilde{b}_j^\shp + \mu_j)}{\exp(\sigma \tilde{b}_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
-&= \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}
+&= \E \left[ \ln \left(1-\delta+\delta \frac{\exp(\sigma b_j^\shp + \mu_j)}{\exp(\sigma b_j + \mu_j)} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E[\ln(1-\delta+\delta \exp(\sigma(b_j^\shp - b_j)))]-\olep_j^E-\olep_j^C-\epsilon_j^0, \label{eq:decomp:j:EshpC}
 \end{align}
 where $(E_j^\shp , C_{j \bs i}^\shp)$ has the same marginals as $(E_j,C_{j \bs i})$ but has independent components;
-$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; $b_j^\shp$ is distributed in the same
-way as $b_j$ but is independent of it; and $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and 
-independent of each other. Continuing,
+$B_j^\shp$ is distributed in the same way as $B_j$ but is independent of it; and $b_j^\shp$ is distributed in the same
+way as $b_j$ but is independent of it. Continuing,
 %storage effects
@@
 &= -\E[r_j(E_j^\shp , C_{j \bs i}^\shp)] \\
-&= -\E[\ln(1-\delta+\delta \exp(\sigma (\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:pECp}
+&= -\E[\ln(1-\delta+\delta \exp(\sigma (b_j^\shp - b_j)))].\label{eq:decomp:j:pECp}
 \end{align}
-these results obtained from (\ref{eq:decomp:j:EC}) and (\ref{eq:decomp:j:EshpC}). Again, $\tilde{b}_j$ and 
-$\tilde{b}_j^\shp$ are standard normally distributed and independent. Continuing,
+Continuing,
 \begin{align}
@@
 \end{align}
-because we showed in the Theory section of the main text that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$ due to the fact that
+because we showed in SI section \ref{SIsect:TheoryForLotteryBoth} that $\olr_{j \bs i}^\nat = \olr_{j \bs i}$, 
+due to the fact that
 $E_j$ and $C_{j \bs i}$ are perfectly linearly related to each other. Continuing,
@@
 \olep_j^{[E \nat C]} &= \olep_j^{(EC)} - \olep_j^{[EC]} \\
-&= - \E[\ln(1-\delta+\delta \exp(\sigma(\tilde{b}_j^\shp - \tilde{b}_j)))],\label{eq:decomp:j:EChard}
+&= - \E[\ln(1-\delta+\delta \exp(\sigma(b_j^\shp - b_j)))],\label{eq:decomp:j:EChard}
 \end{align}
-where again $\tilde{b}_j$ and $\tilde{b}_j^\shp$ are standard normally distributed and independent of each other.
-This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the lottery model.
+where again $b_j$ and $b_j^\shp$ are standard normally distributed and independent of each other.
+This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$ for the log-normal
+fecundities lottery model.
 
@@
 As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
-values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
+values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and a value which is shown in 
+SI section \ref{SIsect:TheoryForLotteryBoth}
+to equal, for the lottery model (both versions), the quantity
+\begin{equation}
+q_{ij} = \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
+\end{equation}
+For the log-normal fecundities lottery model, this is
 \begin{align}
@@
 \end{align}
-These also either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
-$\mu_i-\mu_j$. Thus each of the terms of the decomposition (\ref{MT-eq:full_decomp}) from the main text
+Thus the $q_{ij}$ either do not depend on $\mu_i$ or $\mu_j$ at all or they depend on the difference 
+$\mu_i-\mu_j$, rather than depending separately on $\mu_i$ or $\mu_j$. Thus each of the terms of the 
+decomposition (\ref{MT-eq:full_decomp}) from the main text
 depends only on $\mu_i-\mu_j$ and not separately on $\mu_i$ or $\mu_j$, as stated in the Methods section of the main text.
 
-We now want to show that, for the lottery model, 
+We now want to show that, for the log-normal fecundities lottery model, 
 all elements of our decompositions (\ref{eq:decomp_overall_IGR}) and (\ref{eq:decomp_overall_rjbsi}), as well
@@
 for our left- and right-tail associated cases, as was stated in Methods in the main text. Using the notation 
-$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(b_i^{(l)}),\exp(b_j^{(l)})/\delta)$ and 
-$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(b_i^{(r)}),\exp(b_j^{(r)})/\delta)$
+$(E_i^{(l)},C_{i \bs i}^{(l)})=(B_i^{(l)},B_j^{(l)}/\delta)=(\exp(\sigma b_i^{(l)} + \mu_i),
+\exp(\sigma b_j^{(l)} + \mu_j)/\delta)$ and 
+$(E_i^{(r)},C_{i \bs i}^{(r)})=(B_i^{(r)},B_j^{(r)}/\delta)=(\exp(\sigma b_i^{(r)}+\mu_i),\exp(\sigma b_j^{(r)}+\mu_j)/\delta)$
 for the left- and right-tail associated cases, respectively, we note that the marginal distributions of
-$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and likewise the marginals of 
+$(b_i^{(l)},b_j^{(l)})$ are the same as those of $(b_i^{(r)},b_j^{(r)})$, and so likewise the marginals of 
 $(B_i^{(l)},B_j^{(l)}/\delta)$ are the same as those of $(B_i^{(r)},B_j^{(r)}/\delta)$.
@@
 \begin{equation}
-\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(b_i-b_j))]-\E[\ln(1-\delta+\delta\exp(\ddot{b}_i-\ddot{b}_j))].\label{eq:rep1}
+\olep_i^{[EC]} = \E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j)]-
+\E[\ln(1-\delta+\delta\exp(\sigma(b_i^\nat-b_j^\nat)+\mu_i-\mu_j)].\label{eq:rep1}
 \end{equation}
-But $(\ddot{b}_i,\ddot{b}_j)$ is a bivariate normal random variable constructed to have the same marginals and 
+But $(b_i^\nat,b_j^\nat)$ is a bivariate normal random variable constructed to have the same marginals and 
 covariance as $(b_i,b_j)$, and the marginals and covariance of $(b_i^{(l)},b_j^{(l)})$ are the same as those of
-$(b_i^{(r)},b_j^{(r)})$, so $(\ddot{b}_i^{(l)},\ddot{b}_j^{(l)})$ is distributed in the same way as 
-$(\ddot{b}_i^{(r)},\ddot{b}_j^{(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
+$(b_i^{(r)},b_j^{(r)})$, so $(b_i^{\nat,(l)},b_j^{\nat,(l)})$ is distributed in the same way as 
+$(b_i^{\nat,(r)},b_j^{\nat,(r)})$. So the second term on the right of (\ref{eq:rep1}) is independent
 of the choice between our left- and right-tail-associated noise cases. Now examining 
-$\E[\ln(1-\delta+\delta\exp(b_i-b_j))]$, we define $\beta_1=-b_2+\mu_1+\mu_2$ and $\beta_2=-b_1+\mu_1+\mu_2$,
+$\E[\ln(1-\delta+\delta\exp(\sigma(b_i-b_j)+\mu_i-\mu_j))]$, 
+we define $\beta_1=-b_2$ and $\beta_2=-b_1$,
 and we note that $(\beta_1^{(l)},\beta_2^{(l)})$ is distributed in the same way as 
 $(b_1^{(r)},b_2^{(r)})$; likewise, $(\beta_1^{(r)},\beta_2^{(r)})$ is distributed in the same way as 
-$(b_1^{(l)},b_2^{(l)})$. But then we have 
+$(b_1^{(l)},b_2^{(l)})$. So we then have 
 \begin{align}
-\E[\ln(1-\delta+\delta\exp(b_i^{(l)}-b_j^{(l)}))] &= \E[\ln(1-\delta+\delta\exp(-\beta_j^{(l)}+\mu_1+\mu_2-(-\beta_i^{(l)}+\mu_1+\mu_2)))] \\
-&= \E[\ln(1-\delta+\delta\exp(\beta_i^{(l)}-\beta_j^{(l)}))] \\
-&= \E[\ln(1-\delta+\delta\exp(b_i^{(r)}-b_j^{(r)}))], 
+\E[\ln(1-\delta+\delta\exp(\sigma(b_i^{(l)}-b_j^{(l)})+\mu_i-\mu_j))] 
+&= \E[\ln(1-\delta+\delta\exp(\sigma(\beta_i^{(r)}-\beta_j^{(r)})+\mu_i+\mu_j))] \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma(-b_j^{(r)}+b_i^{(r)})+\mu_i-\mu_j))] \\
+&= \E[\ln(1-\delta+\delta\exp(\sigma(b_i^{(r)}-b_j^{(r)})+\mu_i-\mu_j))], 
 \end{align}
@@
 
-\section{The $\delta=1$ case}\label{sect:delta1}
+\subsection{Theory and plotting simplifications for the beta fecundities lottery model}\label{SIsect:TheoryForLotteryBeta}
+%***DAN: Fill this in.
 
-%The i terms
-Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
-\begin{equation}
-\epsilon_i^0 = \mu_i - \mu_j.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
-\begin{align}
-\olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
-&= -\sigma^2/2.
-\end{align}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
-\begin{align}
-\olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
-&= \sigma^2/2.
-\end{align}
-Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
-\begin{align}
-\olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
-&= 0.
-\end{align}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
-\begin{align}
-\olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
-&= 0.
-\end{align}
-Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
-\begin{align}
-\olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
-&= 0.
-\end{align}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
-\begin{equation}
-\olep_i^{[E \nat C]}=0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
-\begin{equation}
-\olep_i^{[EC]}=0.
-\end{equation}
-%The j terms
-Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
-\begin{equation}
-\epsilon_j^0 = 0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
-\begin{equation}
-\olep_j^E = -\sigma^2/2.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
-\begin{equation}
-\olep_j^C = \sigma^2/2.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
-\begin{equation}
-\olep_j^{EC} = 0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
-\begin{equation}
-\olep_j^{(E \shp C)} = 0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
-\begin{equation}
-\olep_j^{(EC)} = 0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
-\begin{equation}
-\olep_j^{[E \nat C]} = 0.
-\end{equation}
-Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
-\begin{equation}
-\olep_j^{[EC]}=0.
-\end{equation}
-
-For $\delta=1$, the second value of $q_{ij}$ we considered, namely
-$q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
-options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
-equations using $q_{ij}=1$ gives
-\begin{align}
-\Delta_i^0 &= \mu_i - \mu_j \\
-\Delta_i^E &= 0  \\
-\Delta_i^C &= 0  \\
-\Delta_i^{(E \shp C)} &=  0 \\
-\Delta_i^{[E \nat C]} &= 0  \\
-\Delta_i^{[EC]} &=  0. 
-\end{align}
-Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
-which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
-without loss of generality).
-So there is no robust coexistence in the $\delta=1$ case for our simple model.
+%***DAN: I am 90% sure that once I re-read the comment of Ellner saying we do not need to consider really
+%big \delta, we will be able to delete the below.
+% \section{The $\delta=1$ case}\label{sect:delta1}
+% 
+% %The i terms
+% Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
+% \begin{equation}
+% \epsilon_i^0 = \mu_i - \mu_j.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
+% \begin{align}
+% \olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
+% &= -\sigma^2/2.
+% \end{align}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
+% \begin{align}
+% \olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
+% &= \sigma^2/2.
+% \end{align}
+% Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
+% \begin{align}
+% \olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
+% &= 0.
+% \end{align}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
+% \begin{align}
+% \olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
+% &= 0.
+% \end{align}
+% Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
+% \begin{align}
+% \olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
+% &= 0.
+% \end{align}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
+% \begin{equation}
+% \olep_i^{[E \nat C]}=0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
+% \begin{equation}
+% \olep_i^{[EC]}=0.
+% \end{equation}
+% %The j terms
+% Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
+% \begin{equation}
+% \epsilon_j^0 = 0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
+% \begin{equation}
+% \olep_j^E = -\sigma^2/2.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
+% \begin{equation}
+% \olep_j^C = \sigma^2/2.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
+% \begin{equation}
+% \olep_j^{EC} = 0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
+% \begin{equation}
+% \olep_j^{(E \shp C)} = 0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
+% \begin{equation}
+% \olep_j^{(EC)} = 0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
+% \begin{equation}
+% \olep_j^{[E \nat C]} = 0.
+% \end{equation}
+% Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
+% \begin{equation}
+% \olep_j^{[EC]}=0.
+% \end{equation}
+% 
+% For $\delta=1$, the second value of $q_{ij}$ we considered, namely
+% $q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
+% options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
+% equations using $q_{ij}=1$ gives
+% \begin{align}
+% \Delta_i^0 &= \mu_i - \mu_j \\
+% \Delta_i^E &= 0  \\
+% \Delta_i^C &= 0  \\
+% \Delta_i^{(E \shp C)} &=  0 \\
+% \Delta_i^{[E \nat C]} &= 0  \\
+% \Delta_i^{[EC]} &=  0. 
+% \end{align}
+% Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
+% which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
+% without loss of generality).
+% So there is no robust coexistence in the $\delta=1$ case for our simple model.
 
@@
 
+\subsection{Efficient computation for the log-normal fecundities lottery model}\label{sect:efficient_computation_lognormal}
+
 We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
-and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the lottery model.
+and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the log-normal fecundities lottery model.
 
@@
 \item Generate $M$ independent points $u^{(k)}$, $k=1,\ldots,M$, from a standard normal distribution.
-\item Generate $M$ independent points $(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$, as in SI section \ref{sect:noise}, which are left-tail associated
-and which have standard-normal marginals. 
-\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(\tilde{\ddot{b}}_i^{(k)},\tilde{\ddot{b}}_j^{(k)})$, 
+\item Generate $M$ independent points $(b_i^{(k)},b_j^{(k)})$, $k=1,\ldots,M$, as in SI section \ref{sect:noise}, 
+which are left-tail associated and which have standard-normal marginals. 
+\item Again as in SI section \ref{sect:noise}, generate $M$ independent points $(b_i^{\nat,(k)},b_j^{\nat,(k)})$, 
 $k=1,\ldots,M$, from a bivariate normal distribution with component variances $1$ and Pearson correlation the same as 
-$(\tilde{b}_i^{(k)},\tilde{b}_j^{(k)})$.
+$(b_i^{(k)},b_j^{(k)})$.
 \end{enumerate}
@@
 \item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:decomp:contrib_EC}), as
-$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$.
+$\widehat{\olep_i^{[E \nat C]}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]-\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]$.
 The standard error of this expression is
-$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$
+$\se \left( \widehat{\olep_i^{[E \nat C]}} \right) = \sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k [\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]/\sqrt{M}$
 and $s_2=\sd_k [\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]/\sqrt{M}$.
-\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_2}), as 
-$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]-
-\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))].$ The standard
+\item Estimate $\olep_i^{[EC]}$, following (\ref{eq:decomp:contrib_ATA_1}), as 
+$\widehat{\olep_i^{[EC]}} = \mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]-
+\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))].$ The standard
 error of this expression is $\se \left( \widehat{\olep_i^{[EC]}} \right)=\sqrt{s_3^2+s_1^2}$, where
-$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
+$s_1$ is as above, and $s_3=\sd_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}.$
 \item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:base_IGR}), as 
-$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{b}_i^{(k)}-\tilde{b}_j^{(k)})+\mu_i-\mu_j))],$
+$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{(k)}-b_j^{(k)})+\mu_i-\mu_j))],$
 which has standard error $s_3$ from above.
@@
 standard error $\se \left( \widehat{\olep_j^C} \right) = \sd_k [\ln(1-\delta+\delta\exp(-\sigma u^{(k)}+\sigma^2/2))]/\sqrt{M}$.
-\item Note that $\tilde{b}_j$ and $\tilde{b}_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
-independent, so $\tilde{b}_j^\shp - \tilde{b}_j$ is normally distributed with mean $0$ and variance $2$. So we can 
+\item Note that $b_j$ and $b_j^\shp$ in (\ref{eq:decomp:j:EshpC}) are standard normally distributed and
+independent, so $b_j^\shp - b_j$ is normally distributed with mean $0$ and variance $2$. So we can 
 estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:decomp:j:EshpC}), as 
@@
 (\ref{eq:decomp_overall_delta}) can straightforwardly be etimated using our estimates elaborated 
-above, as follows. There are two alternative versions of the estiamtors for this decomposition, corresponding 
+above, as follows. There are two alternative versions of the estimators for this decomposition, corresponding 
 to the alternative values for $q_{ij}$.
@@
 \item Estimate $\Delta_i^{[E \nat C]}$ as $\widehat{\Delta_i^{[E \nat C]}}=\widehat{\olep_i^{[E \nat C]}}-q_{ij}\widehat{\olep_j^{[E \nat C]}}=
-\mean_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]-
+\mean_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]-
 \mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))]+
 q_{ij}\mean_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))],$ which one can straightforwardly show has standard error
-$\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+\delta\exp(\sigma(\tilde{\ddot{b}}_i^{(k)}-\tilde{\ddot{b}}_j^{(k)})+\mu_i-\mu_j))]/\sqrt{M}$ 
+$\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+\delta\exp(\sigma(b_i^{\nat,(k)}-b_j^{\nat,(k)})+\mu_i-\mu_j))]/\sqrt{M}$ 
 and $s_2=\sd_k[\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}+\mu_i-\mu_j))-q_{ij}\ln(1-\delta+\delta\exp(\sigma\sqrt{2}u^{(k)}))]/\sqrt{M}$.
@@
 applies for the estimators as well as for the quantities being estimated.
+
+\subsection{Efficient computation for the beta fecundities lottery model}\label{sect:efficient_computation_beta}
+%***DAN: fill in later
 
@@
 
-\section{      Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
-
-In Fig. \ref{MT-fig:fig4} of the main text we plotted GWR , $\Delta_i^{[EC]}$, and $\text{GWR}-\Delta_i^{[EC]}$ 
-(see equation \ref{eq:decomp_overall_delta} and the text below it) for the lottery model. We here show
-the resulting plot does not depend on which of the two choices for the scaling factors $q_{ij}$ is used.
-But $\text{GWR} = \olr_{i \bs i} - q_{ij} \olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij}\olep_j^{[EC]}$,
-so it suffices to show $\olr_{j \bs i}=0$ and $\olep_j^{[EC]}=0$. The first of these is known true, \emph{a priori}, 
-and the second was demonstrated in section \ref{sect:math_simplifications}.
+%***DAN: I think I do not need this section anymore, but just keeping it around to be safe.
+% \section{      Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
+% 
+% In Fig. \ref{MT-fig:fig4} of the main text we plotted GWR , $\Delta_i^{[EC]}$, and $\text{GWR}-\Delta_i^{[EC]}$ 
+% (see equation \ref{eq:decomp_overall_delta} and the text below it) for the lottery model. We here show
+% the resulting plot does not depend on which of the two choices for the scaling factors $q_{ij}$ is used.
+% But $\text{GWR} = \olr_{i \bs i} - q_{ij} \olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij}\olep_j^{[EC]}$,
+% so it suffices to show $\olr_{j \bs i}=0$ and $\olep_j^{[EC]}=0$. The first of these is known true, \emph{a priori}, 
+% and the second was demonstrated in section \ref{sect:math_simplifications}.
 

### 2023-02-02 paper/Paper.Rnw
further progress editing Methods and associated SI sections

@@
 $\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
-corresponded to more variable fecundities for species $i$.
+corresponded to more variable fecundities for species $i$. The parameter $\eta_i$ represents the upper
+bound on what fecundities could occur for species $i$.
 
@@
 
-\subsection{How results were plotted for G2}
+\subsection{How results were plotted for goal G2}
 
@@
 to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
-(\ref{eq:full_decomp}) for that model against the model parameters $\eta_1$, $\eta_2$, and $\delta$.
-%***DAN: finish it
+(\ref{eq:full_decomp}) for that model against the model parameters $\eta_1$, $\eta_2$, and $\delta$,
+again labelling regions corresponding to ATA rescue or exclusion.
+Some mathematical shortcuts, summarized here, simplified plotting. 
+First, we assumed, without loss of generality, that $\eta_1 \leq \eta_2$. To understand coexistence, 
+it was then sufficient to decompose GWR of the weaker competitor, species 1. 
+Second, we showed that, for the beta fecundities
+lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
+instead they depend only on $\eta_1/\eta_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
+So we plotted against $\eta_1/\eta_2$. 
+For each combination of $\delta=0.2, 0.4, 0.6$
+and for left- and right-tail associated noise, we plotted 
+the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_1/\eta_2$. 
+The simulation methods of \cite{Ellner_2019} were again used.
+Computationally efficient means of performing these simulations are in 
+SI section \ref{SI-sect:efficient_computation}.
+
 %***DAN: Next steps as of 2023 02 01:
-%Fill in the sup mat section for the theory for the beta fecundities lottery model
+%DONE Fill in the sup mat section for the theory for the beta fecundities lottery model
 %fill in the efficient computation section for that same model
-%write this section, here, in the main text
+%DONE write this section, here, in the main text
 

### 2023-02-02 paper/SupMat.Rnw
further progress editing Methods and associated SI sections

@@
 Recall also that competition, $C_i(t)$, is 
-$$C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N)$$ (again for both versions of the model), which
+\begin{equation}
+C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N)
+\end{equation}
+(again for both versions of the model), which
 does not depend on $i$. Species growth rates for the lottery model, $r_i(t)=\ln(N_{i}(t+1)/N_i(t))$, can then be written 
-$$r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)],$$ 
+\begin{equation}
+r_i(t)=\ln[1-\delta+\delta N B_i(t)/(B_1(t)N_1(t)+B_2(t)N_2(t))]=\ln[1-\delta+E_i(t)/C_i(t)],
+\end{equation}
 which is an increasing
@@
 species is obtained from the expression for $C_i=C_j$ by setting $N_i=0$ and $N_j=N$. Thus 
-$$C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta.$$
+\begin{equation}
+C_{i \bs i}(t)=C_{j \bs i}(t)=B_j(t)/\delta=E_j(t)/\delta.
+\end{equation}
 Lottery-model values of all the expressions of SI section \ref{SIsect:other_mechanisms} can be obtained 
@@
 For instance, 
-$$\olr_{i \bs i} = \E[r_i(E_i,C_{i \bs i})] = \E[\ln(1-\delta+E_i /C_{i \bs i})] = \E[\ln(1-\delta+\delta B_i/B_j)].$$ 
+\begin{equation}
+\olr_{i \bs i} = \E[r_i(E_i,C_{i \bs i})] = \E[\ln(1-\delta+E_i /C_{i \bs i})] = \E[\ln(1-\delta+\delta B_i/B_j)].
+\end{equation}
 Likewise, 
-$$\olr_{j \bs i} = \E[r_j(E_j,C_{j \bs i})] = \E[\ln(1-\delta+E_j/C_{j \bs i})] = \E[\ln(1-\delta+\delta B_j/B_j)]=0,$$
+\begin{equation}
+\olr_{j \bs i} = \E[r_j(E_j,C_{j \bs i})] = \E[\ln(1-\delta+E_j/C_{j \bs i})] = \E[\ln(1-\delta+\delta B_j/B_j)]=0,
+\end{equation}
 though we already knew, \emph{a priori}, that this equals $0$. 
@@
 Thus 
-$$\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)].$$ 
+\begin{equation}
+\olrs_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\shp/B_j^\shp)].
+\end{equation}
 Likewise, $\olrs_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})$ in an analogous way; we can use 
 $(E_j^\shp,C_{j \bs i}^\shp)=(B_j^\shp, B_j /\delta)$, so 
-$$\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)].$$
+\begin{equation}
+\olrs_{j \bs i}=\E[\ln(1-\delta+\delta B_j^\shp / B_j)].
+\end{equation}
 An expression for $\olrn_{i \bs i}$ is obtained by replacing $(E_i,C_{i \bs i})=(B_i,B_j/\delta)$
@@
 Thus 
-$$\olrn_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\nat/B_j^\nat)].$$ 
+\begin{equation}
+\olrn_{i \bs i} = \E[\ln(1-\delta+\delta B_i^\nat/B_j^\nat)].
+\end{equation}
 Likewise, $\olrn_{j \bs i}$ is computed by replacing $(E_j,C_{j \bs i})=(B_j,B_j/\delta)$ in an analogous way.
@@
 \subsection{Theory and plotting simplifications for the beta fecundities lottery model}\label{SIsect:TheoryForLotteryBeta}
-%***DAN: Fill this in.
+
+We first compute all terms of the decomposition (\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the 
+beta fecundities lottery model. Defining, for this section, $\beta_i=F_\beta^{-1} \circ \varphi(b_i)$ 
+where $F_\beta$ is the CDF of a beta distribution with parameters $0.5,0.5$ and $\varphi$ is the CDF of 
+a standard normal distribution, we have $B_i = \eta_i \beta_i$. Then, using the formulas of 
+SI section \ref{SIsect:other_mechanisms}, we have
+\begin{equation}
+E_i = B_i = \eta_i \beta_i 
+\end{equation}
+\begin{equation}
+B_j = \eta_j \beta_j 
+\end{equation}
+\begin{equation}
+C_{i \bs i} = \frac{B_j}{\delta} = \frac{\eta_j}{\delta}\beta_j
+\end{equation}
+\begin{equation}
+\overline{E}_i = \overline{B}_i = \frac{\eta_i}{2} 
+\end{equation}
+\begin{equation}
+\overline{B}_j = \frac{\eta_j}{2} 
+\end{equation}
+\begin{equation}
+\overline{C}_{i \bs i} = \frac{\eta_j}{2 \delta} 
+\end{equation}
+\begin{equation}
+r_{i \bs i} = \ln\left(1-\delta+\delta \frac{B_i}{B_j} \right) = \ln\left(1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}\right)
+\end{equation}
+\begin{equation}
+\epsilon_i^0 = \ln\left( 1-\delta+\delta \frac{\eta_i}{\eta_j}  \right) 
+\end{equation}
+\begin{equation}
+\olep_i^E = \E\left[ \ln\left( 1-\delta+\delta \frac{2\eta_i \beta_i}{\eta_j}  \right)  \right]-\epsilon_i^0
+\end{equation}
+\begin{equation}
+\olep_i^C \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i}{2\eta_j \beta_j}\right)  \right]-\epsilon_i^0
+\end{equation}
+\begin{equation}
+\olep_i^{EC} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]-\olep_i^E-\olep_i^C-\epsilon_i^0
+\end{equation}
+and
+\begin{equation}
+\olep_i^{(E\shp C)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right]-\olep_i^E-\olep_i^C-\epsilon_i^0,
+\end{equation}
+where $(\beta_i^\shp,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_i^\shp),F_\beta^{-1} \circ \varphi(b_j^\shp))$.
+Here, $b_i^\shp$ and $b_j^\shp$ are standard normal and independent. Continuing,
+\begin{equation}
+\olep_i^{(EC)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}  \right) \right]
+-\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i\beta_i^\shp}{\eta_j \beta_j^\shp} \right) \right].
+\end{equation}
+Continuing,
+\begin{equation}
+\olep_i^{[EC]}=\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j} \right) \right]-
+\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right],
+\end{equation}
+where $(\beta_i^\nat,\beta_j^\nat)=(F_\beta^{-1} \circ \varphi (b_i^\nat),F_\beta^{-1} \circ \varphi(b_j^\nat))$,
+and $(b_i^\nat,b_j^\nat)$ has the same (standard normal) marginals and covariance as $(b_i,b_j)$ but is
+bivariate normal. Note that a choice has been made here: we could have chosen $(b_i^\nat,b_j^\nat)$
+to have the same (standard normal) marginals as $(b_i,b_j)$, but with covariance chosen so that
+the Pearson, Spearman, or Kendall correlation of $E_i^\nat = \eta_i F_\beta^{-1} \circ \varphi (b_i^\nat)$
+and $C_{i \bs i}^\nat = \eta_j F_\beta^{-1} \circ \varphi (b_j^\nat)/\delta$ is the same as that of 
+$E_i = \eta_i F_\beta^{-1} \circ \varphi (b_i)$ and $C_{i \bs i} = \eta_j F_\beta^{-1} \circ \varphi (b_j)/\delta$.
+But the choice we made seems the most natural under the circumstances.
+Continuing, $\olep_i^{[E\nat C]}=\olep_i^{(E C)}-\olep_i^{[E C]}$ is available by subtraction, i.e.,
+\begin{equation}
+\olep_i^{[E\nat C]} = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right] - 
+\E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right].
+\end{equation}
+This completes the calculation of all the terms of the decomposition 
+(\ref{eq:decomp_overall_IGR}) of $\olr_{i \bs i}$ for the beta fecundities lottery model.
+
+We next decompose $\olr_{j \bs i}$ for the beta fecundities lottery model [see (\ref{eq:decomp_overall_rjbsi})]. We have 
+\begin{equation}
+E_j = B_j = \eta_j \beta_j
+\end{equation}
+\begin{equation}
+C_{j \bs i} = \frac{B_j}{\delta} = \frac{\eta_j \beta_j}{\delta}
+\end{equation}
+\begin{equation}
+\overline{E}_j = \frac{\eta_j}{2}
+\end{equation}
+\begin{equation}
+\overline{C}_{j \bs i} = \frac{\eta_j}{2\delta}
+\end{equation}
+\begin{equation}
+\overline{B}_j = \frac{\eta_j}{2}
+\end{equation}
+\begin{align}
+\olr_{j \bs i} &= \E\left[ \ln\left( 1-\delta+\frac{E_j}{C_{j \bs i}} \right) \right] \\
+&= \E\left[ \ln\left( 1-\delta+\frac{B_j}{B_j} \right) \right] \\
+&= 0 
+\end{align}
+\begin{equation}
+\epsilon_j^0 = \ln \left( 1-\delta+\delta\frac{\eta_j /2}{\eta_j /2} \right) = 0
+\end{equation}
+\begin{align}
+\olep_j^E &= \E\left[ \ln\left( 1-\delta+\delta\frac{2\eta_j \beta_j}{\eta_j} \right) \right] - \epsilon_j^0 \\
+&= \E\left[ \ln\left( 1-\delta+2 \delta \beta_j \right) \right]
+\end{align}
+\begin{align}
+\olep_j^C &= \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_j}{2\eta_j\beta_j} \right) \right]-\epsilon_j^0 \\
+&= \E\left[ \ln\left( 1-\delta+\frac{\delta}{2\beta_j} \right) \right]
+\end{align}
+\begin{align}
+\olep_j^{EC} &= \E\left[  \ln\left( 1-\delta+\delta\frac{B_j}{B_j} \right)\right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= -\olep_j^E - \olep_j^C
+\end{align}
+\begin{align}
+\olep_j^{(E\shp C)} &= \E\left[ \ln\left( 1-\delta+\delta\frac{B_j}{B_j^\shp} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
+&= \E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right) \right]-\olep_j^E-\olep_j^C,
+\end{align}
+where $(\beta_j,\beta_j^\shp)=(F_\beta^{-1} \circ \varphi (b_j),F_\beta^{-1} \circ \varphi (b_j^\shp))$ and
+$b_j^\shp$ is distributed in the same way as $b_j$ but is independent of it. Continuing,
+\begin{align}
+\olep_j^{(EC)} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
+\E\left[ \ln\left(1-\delta+\delta\frac{B_j}{B_j^\shp}  \right)  \right] \\
+&= -\E\left[\ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right)  \right].
+\end{align}
+Continuing,
+\begin{align}
+\olep_j^{[EC]} &= \E\left[ \ln\left(1-\delta+\delta \frac{B_j}{B_j} \right)  \right]-
+\E\left[ \ln\left(1-\delta+\delta\frac{B_j^\nat}{B_j^\nat}  \right)  \right] \\
+&= 0.
+\end{align}
+We showed already (SI section \ref{SIsect:TheoryForLotteryBoth} and 
+\ref{SIsect:TheoryForLotteryLogNorm}) that this is $0$ because $E_j$ and $C_{j \bs i}$
+are linearly related to each other. Continuing,
+\begin{align}
+\olep_j^{[E \nat C]} &= \olep_j^{(EC)}-\olep_j^{[EC]} \\
+&= -\E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp}  \right)  \right].
+\end{align}
+This completes the calculation of all the terms of the decomposition of $\olr_{j \bs i}$
+for the beta fecundities lottery model.
+
+Using the above formulas, we observe that all the terms of the decomposition of $\olr_{i \bs i}$ depend on 
+$\delta$, $\sigma$, and $\eta_i/\eta_j$, but not separately on $\eta_i$ or $\eta_j$. The terms of the decomposition
+of $\olr_{j \bs i}$ depend on $\delta$ and $\sigma$, but do not depend on $\eta_i$ or $\eta_j$ at all.
+As discussed in the main text and SI section \ref{SIsect:scaling_factors}, we consider two alternative
+values for the scaling factors, $q_{ij}$, namely $q_{ij}=1$ and 
+\begin{equation}
+q_{ij} = \frac{\overline{B}_i}{\overline{B}_j (1-\delta) +\overline{B}_i \delta}.
+\end{equation}
+For the beta fecundities lottery model, this is
+\begin{align}
+q_{ij} &= \frac{\eta_i}{\eta_j(1-\delta)+\eta_i \delta} \\
+&= \frac{\eta_i/\eta_j}{(1-\delta)+\frac{\eta_i}{\eta_j}\delta}.
+\end{align}
+Thus the $q_{ij}$ either do not depend on $\eta_i$ or $\eta_j$ at all or they depend on the quotient 
+$\eta_i/\eta_j$, rather than depending separately on $\eta_i$ or $\eta_j$. Thus each of the terms of the 
+decomposition (\ref{MT-eq:full_decomp}) from the main text
+depends only on $\eta_i/\eta_j$ and not separately on $\eta_i$ or $\eta_j$, as stated in the Methods 
+section of the main text.
 

### 2023-02-03 paper/SupMat.Rnw
continued working on the efficient computations section in sup mat

@@
 \begin{equation}
-r_{i \bs i} = \ln\left(1-\delta+\delta \frac{B_i}{B_j} \right) = \ln\left(1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}\right)
-\end{equation}
-\begin{equation}
-\epsilon_i^0 = \ln\left( 1-\delta+\delta \frac{\eta_i}{\eta_j}  \right) 
-\end{equation}
-\begin{equation}
-\olep_i^E = \E\left[ \ln\left( 1-\delta+\delta \frac{2\eta_i \beta_i}{\eta_j}  \right)  \right]-\epsilon_i^0
-\end{equation}
-\begin{equation}
-\olep_i^C \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i}{2\eta_j \beta_j}\right)  \right]-\epsilon_i^0
+r_{i \bs i} = \ln\left(1-\delta+\delta \frac{B_i}{B_j} \right) = \ln\left(1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j}\right)\label{eq:103}
+\end{equation}
+\begin{equation}
+\epsilon_i^0 = \ln\left( 1-\delta+\delta \frac{\eta_i}{\eta_j}  \right) \label{eq:104}
+\end{equation}
+\begin{equation}
+\olep_i^E = \E\left[ \ln\left( 1-\delta+\delta \frac{2\eta_i \beta_i}{\eta_j}  \right)  \right]-\epsilon_i^0 \label{eq:105}
+\end{equation}
+\begin{equation}
+\olep_i^C = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i}{2\eta_j \beta_j}\right)  \right]-\epsilon_i^0 \label{eq:106}
 \end{equation}
@@
 \begin{equation}
-\olep_i^{(E\shp C)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right]-\olep_i^E-\olep_i^C-\epsilon_i^0,
+\olep_i^{(E\shp C)} = \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right]-\olep_i^E-\olep_i^C-\epsilon_i^0,\label{eq:108}
 \end{equation}
@@
 \olep_i^{[EC]}=\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i}{\eta_j \beta_j} \right) \right]-
-\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right],
+\E\left[ \ln\left( 1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right],\label{eq:110}
 \end{equation}
@@
 \olep_i^{[E\nat C]} = \E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\nat}{\eta_j \beta_j^\nat}  \right)  \right] - 
-\E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right].
+\E\left[ \ln\left(1-\delta+\delta\frac{\eta_i \beta_i^\shp}{\eta_j \beta_j^\shp}  \right)  \right].\label{eq:111}
 \end{equation}
@@
 &= \E\left[ \ln\left( 1-\delta+\frac{B_j}{B_j} \right) \right] \\
-&= 0 
-\end{align}
-\begin{equation}
-\epsilon_j^0 = \ln \left( 1-\delta+\delta\frac{\eta_j /2}{\eta_j /2} \right) = 0
+&= 0 \label{eq:119}
+\end{align}
+\begin{equation}
+\epsilon_j^0 = \ln \left( 1-\delta+\delta\frac{\eta_j /2}{\eta_j /2} \right) = 0 \label{eq:120}
 \end{equation}
@@
 \olep_j^E &= \E\left[ \ln\left( 1-\delta+\delta\frac{2\eta_j \beta_j}{\eta_j} \right) \right] - \epsilon_j^0 \\
-&= \E\left[ \ln\left( 1-\delta+2 \delta \beta_j \right) \right]
+&= \E\left[ \ln\left( 1-\delta+2 \delta \beta_j \right) \right] \label{eq:122}
 \end{align}
@@
 \olep_j^C &= \E\left[ \ln\left( 1-\delta+\delta\frac{\eta_j}{2\eta_j\beta_j} \right) \right]-\epsilon_j^0 \\
-&= \E\left[ \ln\left( 1-\delta+\frac{\delta}{2\beta_j} \right) \right]
+&= \E\left[ \ln\left( 1-\delta+\frac{\delta}{2\beta_j} \right) \right] \label{eq:124}
 \end{align}
@@
 \olep_j^{(E\shp C)} &= \E\left[ \ln\left( 1-\delta+\delta\frac{B_j}{B_j^\shp} \right) \right]-\olep_j^E-\olep_j^C-\epsilon_j^0 \\
-&= \E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right) \right]-\olep_j^E-\olep_j^C,
+&= \E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp} \right) \right]-\olep_j^E-\olep_j^C, \label{eq:128}
 \end{align}
@@
 \E\left[ \ln\left(1-\delta+\delta\frac{B_j^\nat}{B_j^\nat}  \right)  \right] \\
-&= 0.
+&= 0.\label{eq:132}
 \end{align}
@@
 \olep_j^{[E \nat C]} &= \olep_j^{(EC)}-\olep_j^{[EC]} \\
-&= -\E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp}  \right)  \right].
+&= -\E\left[ \ln\left( 1-\delta+\delta\frac{\beta_j}{\beta_j^\shp}  \right)  \right].\label{eq:134}
 \end{align}
@@
 q_{ij} &= \frac{\eta_i}{\eta_j(1-\delta)+\eta_i \delta} \\
-&= \frac{\eta_i/\eta_j}{(1-\delta)+\frac{\eta_i}{\eta_j}\delta}.
+&= \frac{\eta_i/\eta_j}{(1-\delta)+\frac{\eta_i}{\eta_j}\delta}.\label{eq:137}
 \end{align}
@@
 \widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
-at the level of the estimators, as well
+at the level of the estimators, as well as at the level of the quantities being estimated.
 
@@
 \subsection{Efficient computation for the beta fecundities lottery model}\label{sect:efficient_computation_beta}
-%***DAN: fill in later
+
+We here elaborate how the terms in the decompositions (\ref{eq:decomp_overall_IGR}), (\ref{eq:decomp_overall_rjbsi})
+and (\ref{eq:decomp_overall_delta}) are computed rapidly, for the beta fecundities lottery model.
+
+First, for a large integer $M$ (we used $M=\Sexpr{M}$),
+carry out the following steps once:
+\begin{enumerate}
+\item Generate $M$ independent points $(\beta_i^{(k)},\beta_j^{(k)})=(F_\beta^{-1} \circ \varphi (b_i^{(k)}),F_\beta^{-1} \circ \varphi (b_j^{(k)}))$,
+$k=1,\ldots,M$, where $(b_i^{(k)},b_j^{(k)})$ are left- or right-tail associated points generated as in SI section \ref{sect:noise},
+depending on whether one is currently considering the left- or right-tail associated case.
+\item Generate $M$ independent points $(\beta_i^{\shp,(k)},\beta_j^{\shp,(k)})=(F_\beta^{-1} \circ \varphi (b_i^{\shp,(k)}),
+F_\beta^{-1} \circ \varphi (b_j^{\shp,(k)}))$, $k=1,\ldots,M$, where $b_i^{\shp,(k)}$ and $b_j^{\shp,(k)}$ are standard normal and 
+independent.
+\item Generate $M$ independent points 
+$(\beta_i^{\nat,(k)},\beta_j^{\nat,(k)})=(F_\beta^{-1} \circ \varphi (b_i^{\nat,(k)}),
+F_\beta^{-1} \circ \varphi (b_j^{\nat,(k)}))$, $k=1,\ldots,M$, where the points $(b_i^{\nat,(k)},b_j^{\nat,(k)})$
+come from a bivariate normal distribution with standard normal marginals and Pearson correlation the same
+as $(b_i^{(k)},b_j^{(k)})$ (see SI section \ref{sect:noise}). 
+\end{enumerate}
+These data are stored and used repeatedly below.
+
+Next, given values of the parameters $\delta$ and $\eta_i/\eta_j$, proceed as follows to get estimates of the terms of the 
+decomposition (\ref{eq:decomp_overall_IGR}).
+\begin{enumerate}
+%baseline
+\item Compute $\epsilon_i^0$, following (\ref{eq:104}), which is a simple formula of $\delta$ and $\eta_i/\eta_j$, not 
+requiring use of the random samples described above. This gives an exact values of $\epsilon_i^0$, not an estimate.
+%E
+\item Estimate $\olep_i^E$, following (\ref{eq:105}), as $\widehat{\olep_i^E}=
+\mean_k[ \ln( 1-\delta+2\delta \frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)}) ]-\epsilon_i^0$. This 
+estimate has standard error $\se( \widehat{\olep_i^E})=
+\sd_k[  \ln( 1-\delta+2\delta \frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)}) ]/\sqrt{M}$, where 
+$\sd (\cdot)$ denotes sample standard deviation.
+%C
+\item Estimate $\olep_i^C$, following (\ref{eq:106}), as $\widehat{\olep_i^C}=
+\mean_k[ \ln( 1-\delta+\frac{\delta}{2} \frac{\eta_i}{\eta_j}\frac{1}{\beta_j^{\shp,(k)}}) ]-\epsilon_i^0$. 
+This estimate has standard error $\se( \widehat{\olep_i^C})=
+\sd_k[  \ln( 1-\delta+\frac{\delta}{2} \frac{\eta_i}{\eta_j}\frac{1}{\beta_j^{\shp,(k)}}) ]/\sqrt{M}$.
+%E#C
+\item Estimate $\olep_i^{(E \shp C)}$, following (\ref{eq:108}), as $\widehat{\olep_i^{(E \shp C)}}=
+\mean_k[ \ln( 1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}}) ]-
+\widehat{\olep_i^E}-\widehat{\olep_i^C}-\epsilon_i^0$. The standard error of this estimate is 
+$\se( \widehat{\olep_i^{(E \shp C)}}  )=
+\sd_k [ \ln( 1-\delta+\delta \frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}})-
+\ln( 1-\delta+\frac{\delta}{2}\frac{\eta_i}{\eta_j}\frac{1}{\beta_j^{\shp,(k)}} )-
+\ln( 1-\delta+2\delta\frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)} )]/\sqrt{M}$.
+%E||C
+\item Estimate $\olep_i^{[E \nat C]}$, following (\ref{eq:111}), as $\widehat{\olep_i^{[E\nat C]}}=
+\mean_k[\ln( 1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\nat,(k)}}{\beta_j^{\nat,(k)}} )]-
+\mean_k[\ln( 1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}} )]$.
+The standard error of this expression is $\sqrt{s_1^2+s_2^2}$, where 
+$s_1=\sd_k[ \ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\nat,(k)}}{\beta_j^{\nat,(k)}}  )  ]/\sqrt{M}$
+and $s_2=\sd_k[ \ln(  1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}}) ]/\sqrt{M}$.
+%[EC]
+\item Esimate $\olep_i^{[EC]}$, following (\ref{eq:110}), as $\widehat{\olep_i^{[EC]}}=
+\mean_k[\ln( 1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{(k)}}{\beta_j^{(k)}} )]-
+\mean_k[\ln( 1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\nat,(k)}}{\beta_j^{\nat,(k)}} )]$.
+The standard error of this expression is $\sqrt{s_1^2+s_3^2}$, where $s_1$ is defined above, and 
+$s_3=\sd_k[ \ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{(k)}}{\beta_j^{(k)}}  )  ]/\sqrt{M}$.
+%r
+\item Finally, estimate $\olr_{i \bs i}$, following (\ref{eq:103}), as
+$\widehat{\olr_{i \bs i}}=\mean_k[\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{(k)}}{\beta_j^{(k)}})]$,
+which has standard error $s_3$ from above.
+\end{enumerate}
+It is easy to see that $\widehat{\olr_{i \bs i}}=\epsilon_i^0+\widehat{\olep_i^E}+\widehat{\olep_i^C}+\widehat{\olep_i^{(E \shp C)}}+
+\widehat{\olep_i^{[E \nat C]}}+\widehat{\olep_i^{[EC]}},$ i.e., the decomposition (\ref{eq:decomp_overall_IGR}) applies 
+at the level of the estimators, as well as at the level of the quantities being estimated.
+
+Next, proceed as follows to get estimates of the terms in the decomposition (\ref{eq:decomp_overall_rjbsi}). 
+\begin{enumerate}
+%baseline
+\item Use $\epsilon_j^0 = 0$, as in (\ref{eq:120}), and this is an exact value.
+%E
+\item Estimate $\olep_j^E$, following (\ref{eq:122}), as $\widehat{\olep_j^E}=\mean_k[\ln(1-\delta+2\delta\beta_j^{(k)})]$. This has
+standard error $\se(\widehat{\olep_j^E})=\sd_k[\ln(1-\delta+2\delta\beta_j^{(k)})]/\sqrt{M}$.
+%C
+\item Estimate $\olep_j^C$, following (\ref{eq:124}), as $\widehat{\olep_j^C}=
+\mean_k[\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]$, which has standard error $\se(\widehat{\olep_j^C})=
+\sd_k[\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%E#C
+\item Estimate $\olep_j^{(E \shp C)}$, following (\ref{eq:128}), as $\widehat{\olep_j^{(E \shp C)}}=
+\mean_k[\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})]-\widehat{\olep_j^E}-\widehat{\olep_j^C}$. This has
+standard error $\se(\widehat{\olep_j^{(E \shp C)}})=\sd_k[\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})-
+\ln(1-\delta+2\delta \beta_j^{(k)})-\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%E||C
+\item Estimate $\olep_j^{[E \nat C]}$, following (\ref{eq:134}), as $\widehat{\olep_j^{[E \nat C]}}=
+-\mean_k[\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})]$, which has standard error 
+$\se(\widehat{\olep_j^{[E \nat C]}})=\sd_k[\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%[EC]
+\item Use $\olep_j^{[EC]}=0$, following (\ref{eq:132}). This is exact.
+%r
+\item Finally, use $\olr_{j \bs i} = 0$, following (\ref{eq:119}), again exact.
+\end{enumerate}
+As for (\ref{eq:decomp_overall_IGR}), it is again easy to see that (\ref{eq:decomp_overall_rjbsi})
+applies for the estimates, as well as for the quantities being estimated.
+
+Next, $q_{ij}$ can be taken (SI section \ref{sect:noise}) to be, alternatively, either $1$ or the quantity displayed in 
+(\ref{eq:137}). Hence the terms of the decomposition (\ref{eq:decomp_overall_delta}) can straightforwardly be
+estimated using our estimates elaborated above, as follows. There are two versions of each estimator, corresponding
+to the alternative values of $q_{ij}$. 
+%***DAN: stopped here, 2023 02 03. You can continue from p. 8 of your hand-written version of this
 

### 2023-02-03 paper/Paper.Rnw
finished Methods and accompanying SI sections, at least for now

@@
 
-%***DAN: Next steps as of 2023 02 01:
-%DONE Fill in the sup mat section for the theory for the beta fecundities lottery model
-%fill in the efficient computation section for that same model
-%DONE write this section, here, in the main text
+%Trait different space stuff. I don't think I need it anymore but keeping around until I finish writing the Discussion,
+%in case some of it turns out to be useful there.
+% To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
+% ATAs but not without, or vice versa, we plotted GWR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
+% as well as
+% GWR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
+% For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
+% we were interested in the values of $\mu_1-\mu_2$ for which one of these
+% quantities is positive and the other is negative; this set of values characterizes the species which can coexist
+% with ATAs, but not without, or vice versa. We define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
+% Points of the trait-difference space characterize possible combinations of traits of the two 
+% competing species in our model. For fixed values of 
+% $\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
+% trait-difference space to be the
+% values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
+% is the subspace for which the two lottery model species coexist in the presence of ATAs.
+% We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the values
+% $\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
+% This is the subspace for which the two lottery model species coexist in the absence of 
+% ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ indicates the
+% absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
+% set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
+% values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
+% We define ``ATA exclusion space'' to be the set difference 
+% $\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
+% i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
+% On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
+% against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
+% $\mu_1 - \mu_2$ for all combinations of $\delta=0.2, 0.4, 0.6, 0.8$ and $\sigma=2,4,5,6$.
 
@@
 
-%Then on to the diatom stuff.
-
-%G3
-To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
-ATAs but not without, or vice versa, we plotted GWR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
-as well as
-GWR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
-For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
-we were interested in the values of $\mu_1-\mu_2$ for which one of these
-quantities is positive and the other is negative; this set of values characterizes the species which can coexist
-with ATAs, but not without, or vice versa. We define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
-Points of the trait-difference space characterize possible combinations of traits of the two 
-competing species in our model. For fixed values of 
-$\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
-trait-difference space to be the
-values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
-is the subspace for which the two lottery model species coexist in the presence of ATAs.
-We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the values
-$\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
-This is the subspace for which the two lottery model species coexist in the absence of 
-ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ indicates the
-absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
-set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
-values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
-We define ``ATA exclusion space'' to be the set difference 
-$\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
-i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
-On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
-against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
-$\mu_1 - \mu_2$ for all combinations of $\delta=0.2, 0.4, 0.6, 0.8$ and $\sigma=2,4,5,6$.
-
-%G4
-To achieve goal G4 from the Introduction, i.e., to provide an empirical example of species 
-coexistence which highlights the role ATAs can play, we use a laboratory chemostat
+To achieve goal G3 from the Introduction, i.e., to provide an empirical example of species 
+coexistence which highlights the role ATAs can play, we used a laboratory chemostat
 system of freshwater diatoms explored by \cite{Descamps_2005}.
@@
 
-To display results, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
+\subsection{How results were plotted for goal G3}
+
+%***DAN: This para will likely have to be revised depending on how Jasmin revises the diatom figure presentation
+To display results for goal G3, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
 $\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values

### 2023-02-03 paper/Paper.Rnw
small things

@@
 
+%***DAN: made it to here, 2023 02 03, 4:11 pm. Resume from here.
+
+%***DAN: copy of goal from the Intro
+%G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
+%from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
+%sometimes coexist with ATAs but not without, or vice versa.
+
 %G2 
-Fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
-of the contributions to coexistence that come from ATAs compared to other mechanisms), Fig.
+To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
+of the contributions to coexistence that come from ATAs compared to other mechanisms; and to see if ATA rescue 
+or ATA exclusion can occur), Fig.
 \ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) often, but not always, has 
@@
 
-%***DAN: I edited the caption (2023 01 26) but this is the completely the wrong figure. I need to replace it 
-%with Jasmin's new pedagogical figure once I know where in the repo 
-%structure that one is generated and stored, and after I have re-run Jasmin's code, and after she has made the
-%changes I recently asked for (see next comment).
-%***DAN: I asked Jasmin to replace the \phi with \varphi in the axis labels on b and d, and to make the marginals
-%on b and c be *standard* normal instead of just normal.
+%***DAN: This looks pretty good except the histograms for some marginals do not really look like the plotted
+%pdfs of the marginals. I think Jasmin said in an email she was fixing this.
 \begin{figure}
-\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
+\includegraphics[width=1.025\textwidth]{../results_figs/theory_fig_col.pdf}
 \caption{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 

### 2023-02-16 paper/SupMat.Rnw
finished Methods and accompanying SI sections, at least for now
fixed an error spotted by Jasmin
fixed typo in one of the sup mat formulas that I noticed while doing something else. it's not consequential.
made some aspects clearer of how efficient computations are done

@@
 \olr_{j \bs i} &= \E\left[ \ln\left( 1-\delta+\frac{E_j}{C_{j \bs i}} \right) \right] \\
-&= \E\left[ \ln\left( 1-\delta+\frac{B_j}{B_j} \right) \right] \\
+&= \E\left[ \ln\left( 1-\delta+\delta\frac{B_j}{B_j} \right) \right] \\
 &= 0 \label{eq:119}
@@
 F_\beta^{-1} \circ \varphi (b_j^{\shp,(k)}))$, $k=1,\ldots,M$, where $b_i^{\shp,(k)}$ and $b_j^{\shp,(k)}$ are standard normal and 
-independent.
+independent. Make these points also independent of the points $(\beta_i^{(k)},\beta_j^{(k)})$ generated above.
 \item Generate $M$ independent points 
@@
 come from a bivariate normal distribution with standard normal marginals and Pearson correlation the same
-as $(b_i^{(k)},b_j^{(k)})$ (see SI section \ref{sect:noise}). 
+as $(b_i^{(k)},b_j^{(k)})$ (see SI section \ref{sect:noise}). Make these points also independent of both sets of 
+points generated above.
 \end{enumerate}
@@
 to the alternative values of $q_{ij}$. 
-%***DAN: stopped here, 2023 02 03. You can continue from p. 8 of your hand-written version of this
+\begin{enumerate}
+%baseline
+\item Compute $\Delta_i^0 = \epsilon_i^0 - q_{ij} \epsilon_j^0$, which is exact. Since $\epsilon_j^0=0$, 
+we have $\Delta_i^0 = \epsilon_i^0$.
+%E
+\item Estimate $\Delta_i^E$ as $\widehat{\Delta_i^E}=\widehat{\olep_i^E}-q_{ij} \widehat{\olep_j^E}=
+\mean_k[\ln(1-\delta+2\delta\frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)})]-\epsilon_i^0-
+q_{ij} \mean_k[\ln(1-\delta+2\delta\beta_j^{(k)})]$, which has standard error 
+$\se(\widehat{\Delta_i^E})=\sqrt{s_1^2+s_2^2}$, where $s_1=\sd_k[\ln(1-\delta+2\delta\frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)})]/\sqrt{M}$
+and $s_2=q_{ij}\sd_k[\ln(1-\delta+2\delta\beta_j^{(k)})]/\sqrt{M}$.
+%C
+\item Estimate $\Delta_i^C$ as $\widehat{\Delta_i^C}=\widehat{\olep_i^C}-q_{ij}\widehat{\olep_j^C}=
+\mean_k[\ln(1-\delta+\frac{\delta}{2}\frac{\eta_i}{\eta_j}\frac{1}{\beta_j^{\shp,(k)}})]-\epsilon_i^0-q_{ij} 
+\mean_k[\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]$, which has standard error
+$\se(\widehat{\Delta_i^C})=\sd_k[\ln(1-\delta+\frac{\delta}{2}\frac{\eta_i}{\eta_j}\frac{1}{\beta_j^{\shp,(k)}})-
+q_{ij}\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%E#C
+\item Estimate $\Delta_i^{(E\shp C)}$ as $\widehat{\Delta_i^{(E\shp C)}}=\widehat{\olep_i^{(E\shp C)}}-q_{ij}
+\widehat{\olep_j^{(E\shp C)}}$. This has standard error 
+$\se(\widehat{\Delta_i^{(E\shp C)}})=\sd_k[\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}})-
+\ln(1-\delta+2\delta\frac{\eta_i}{\eta_j}\beta_i^{\shp,(k)})-\ln(1-\delta+\frac{\delta}{2}\frac{\eta_1}{\eta_2}\frac{1}{\beta_j^{\shp,(k)}})-
+q_{ij}\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})+q_{ij}\ln(1-\delta+2\delta\beta_j^{(k)})+
+q_{ij}\ln(1-\delta+\frac{\delta}{2}\frac{1}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%E||C
+\item Estimate $\Delta_i^{[E\nat C]}$ as $\widehat{\Delta_i^{[E\nat C]}}=\widehat{\olep_i^{[E\nat C]}}-
+q_{ij}\widehat{\olep_j^{[E\nat C]}}=
+\mean_k[\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\nat,(k)}}{\beta_j^{\nat,(k)}})]-
+\mean_k[\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}})]+
+q_{ij}\mean_k[\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})]$, 
+which has standard error
+$\se(\widehat{\Delta_i^{[E\nat C]}})=\sd_k[\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\nat,(k)}}{\beta_j^{\nat,(k)}})-
+\ln(1-\delta+\delta\frac{\eta_i}{\eta_j}\frac{\beta_i^{\shp,(k)}}{\beta_j^{\shp,(k)}})+
+q_{ij}\ln(1-\delta+\delta\frac{\beta_j^{(k)}}{\beta_j^{\shp,(k)}})]/\sqrt{M}$.
+%[EC]
+\item For $\Delta_i^{[EC]}$, we recall that $\olep_j^{[EC]}=0$, so $\Delta_i^{[EC]}=\olep_i^{[EC]}$.
+So we use $\widehat{\olep_i^{[EC]}}$ for $\Delta_i^{[EC]}$.
+%r
+\item Finally, we can estimate $\olr_{i \bs i}-q_{ij} \olr_{j \bs i}$ simply as $\widehat{\olr_{i \bs i}}$,
+since $\olr_{j \bs i}=0$.
+\end{enumerate}
 

### 2023-02-24 paper/Paper.Rnw
problems with the R chunk in paper.Rnw identified and commented

@@
 #place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
-#allregres<-matrix(rnorm(10),2,5)
 res_loc <- "../results_numeric/"
-fig2 <- readRDS(file="../results_numeric/fig2data.RDS")
-M <- readRDS(file="../results_numeric/M.RDS")
-se3 <- readRDS(file="../results_numeric/fig3maxse.RDS")
-se4 <- readRDS(file="../results_numeric/fig4maxse.RDS")
+#fig2 <- readRDS(file=paste0(res_loc,"fig2data.RDS"))
+M <- readRDS(file=paste0(res_loc,"M.RDS"))
+se3 <- readRDS(file=paste0(res_loc,"fig3maxse.RDS"))
+#se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
+se4<-NA
 se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))

### 2023-02-26 paper/Paper.Rnw
finished editing REsults for now, though will come back to it again after Jasmin finishes modifying the diatom figures

@@
 %the log-normal fecundities version
-The \emph{log-normal fecundities lottery model} used $B_i = \exp(\sigma+\mu_i b_i)$,
+The \emph{log-normal fecundities lottery model} used $B_i = \exp(\mu_i+\sigma b_i)$,
 and therefore had parameters $\sigma$, $\mu_1$ and $\mu_2$ in addition to the death rate parameter $\delta$
-(Introduction) that both versions of the lottery model used. The parameter $\sigma$ is the variance of the
+(Introduction) that both versions of the lottery model used. The parameter $\sigma$ is the standard deviation of the
 log-fecundities and $\mu_i$ is the mean log fecundity for species $i$.
@@
 %the beta fucundities version
-The \emph{beta-fecundities lottery model}
+The \emph{beta fecundities lottery model}
 used $B_i = \eta_i F_\beta^{-1} \circ \varphi(b_i)$, where $\varphi$ is the CDF of the standard normal
@@
 with shape and scale parameters $0.5$ (probability density function proportional to $x^{-05}(1-x)^{-0.5}$ on
-the unit interval). So the beta-fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
+the unit interval). So the beta fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
 and produced fecundities with $U$-shaped, beta distributions between $0$ and $\eta_1$, for $B_1$, and
@@
 So we only generated plots 
-for left-tail associated noise. For each combination of $\delta=0.2, 0.4, 0.6$
-and $\mu_1-\mu_2=0,-0.5,-2,-4$,
-we plotted the components of (\ref{eq:full_decomp}) against $\sigma$. 
+for left-tail associated noise.
+
+For each combination of $\delta=0.2, 0.4, 0.6$ and $\mu_1-\mu_2=0,-0.5,-2,-4$,
+we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ ranging 
+from $0$ to $7$. Some of 
+the $\mu$ differences and $\sigma$ values we considered were very unrealistic (see Results),
+so we considered after plotting whether conclusions about the importance of ATAs depended 
+on realistic values of parameters.
 %***DAN: Jasmin mentioned that she only used \delta=0.2, .4, .6 because of a comment Ellner made where he 
@@
 lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
-instead they depend only on $\eta_1/\eta_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
-So we plotted against $\eta_1/\eta_2$. 
+instead they depend only on $\eta_2/\eta_1$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
+So we plotted against $\eta_2/\eta_1$.
+
 For each combination of $\delta=0.2, 0.4, 0.6$
 and for left- and right-tail associated noise, we plotted 
-the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_1/\eta_2$. 
+the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_2/\eta_1$, using 
+the fairly realistic range $1 \leq \eta_2/\eta_1 \leq 5$. 
 The simulation methods of \cite{Ellner_2019} were again used.
@@
 
-%***DAN: made it to here, 2023 02 03, 4:11 pm. Resume from here.
-
-%***DAN: copy of goal from the Intro
-%G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
-%from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
-%sometimes coexist with ATAs but not without, or vice versa.
-
 %G2 
@@
 or ATA exclusion can occur), Fig.
-\ref{fig:fig3} shows that the ATA term of (\ref{eq:full_decomp}) often, but not always, has 
+\ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
+the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
 magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
@@
 $\mu_2$, i.e., for closer-to-equal competitors. 
-In the $\mu_1=\mu_2$ case (panels a-d), for which the two species were equal 
-competitors, storage effects (blue plus red
-lines on Fig. \ref{fig:fig3}) were the only means by which coexistence could occur, and ATA effects
-tended to be negative, weakening total storage effects.
+In the $\mu_1=\mu_2$ case (panels a-c), for which the two species were equal 
+competitors, storage effects ($\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$) 
+were the only means by which coexistence could occur, and ATA effects
+tended to be negative, weakening total storage effects. The ATA term 
+was also often comparable in magnitude to the other terms of 
+the decomposition (\ref{eq:full_decomp}) for the beta fecundities lottery model 
+(Fig. \ref{fig:fig4}).
 
-%G3
-Fulfilling goal G3 of the Introduction (to explore, using the lottery model, whether species
-can coexist with ATAs, but not without, or vice versa), our results show that ATAs can, indeed, sometimes
-make the difference between coexistence and exclusion: both ``ATA rescue'' and ``ATA
-exclusion'' (see Methods) are possible.
-For some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
-line and the $x$ axis for some values of $\sigma$, e.g., see panel h. 
+Our results also show that ATAs can make the difference between coexistence 
+and failure of two species to coexist: both ``ATA rescue'' and ``ATA
+exclusion'' (see Theory) are possible. Starting with the log-normal fecundities 
+lottery model, for some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
+line and the $x$ axis for some values of $\sigma$, e.g., see panel i.
 Such cases correspond to parameter combinations
 for which the presence or absence of ATAs determine whether the species coexist or not. In the 
-case of panel h, for values of $\sigma$ around $4$, 
+case of panel i, for values of $\sigma$ around $6$, 
 species 1 is excluded with ATAs but the two species can coexist without ATAs. 
-Fig. \ref{fig:fig4} highlights additional examples. Orange lines on Fig. \ref{fig:fig4}
-represent GWR when the effects of ATAs are included, and blue lines are GWR when
-the effects of ATAs are removed. These two lines frequently crossed the $x$ axis in 
-different places, showing that both ATA rescue space and ATA exclusion space can be non-empty
-subsets of our trait-difference space (see Methods). For the parameters we considered, 
-both ATA rescue and exclusion were more common for larger values of $\sigma$.
-ATA rescue appeared more likely for smaller $\delta$, whereas ATA exclusion appeared
-more common for larger $\delta$. Although some bands of ATA rescue or exclusion on Fig.
-\ref{fig:fig4} appear narrow, they nevertheless do occur, which means ATAs can mean the 
-difference between coexistence and extirpation of one species in some cases.
+Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model.
 
-%G4
+For the log-normal fecundities lottery model, the ATA term from (\ref{eq:full_decomp}) tended
+only to be comparable in magnitude to the other terms, and ATA exclusion or rescue tended only to
+occur, for unrealistic parameters values; but ATA terms were large and ATA rescue and exclusion occurred
+for realistic parameter values for the beta fecundities lottery model. For instance, the ATA in 
+Fig. \ref{fig:fig3} was fairly 
+small, compared to the other terms of (\ref{eq:full_decomp}), for values of $\sigma$ less than 
+about $3$. And ATA rescue and exclusion also only occurred for $\sigma$ greater than about $3$.
+The standard deviation of fecundity for species 1 when 
+$\mu_1=0$ and $\sigma=3$ was greater than $8000$, so values of $\sigma$ greater than $3$ are 
+probably unrealistic. Thus ATAs seem to be unimportant for coexistence
+for the log-normal fecundities lottery model using realistic parameter values. However, 
+all the $\eta_2/\eta_1$ ratio values we plotted were fairly realistic, corresponding to
+upper-bound fecundities that were up to $5$ times higher for species 2 than for species 1. And 
+the ATA term plotted in Fig. \ref{fig:fig4} was comparable in magnitude 
+to the other components of (\ref{eq:full_decomp}) across the whole range of $\eta_2/\eta_1$ values
+considered; and both ATA rescue and exclusion occurred for values of this ratio between $1$ and $2$.
+Apparently the question of whether ATAs are important for realistic parameter values depends on model specifics.
+
+%G4 ***DAN: Dan, this will have to be edited after Jasmin gets done editing the corresponding figures
 Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence 
@@
 
+%***DAN: Jasmin, this figure looks really great. Here are a few minor things I would like to ask you to
+%do to further optimize:
+%1) Currently there are axis labels and numbers only on the bottom and left panels, which makes sense.
+%But the axis tick marks needs to occur on all panels. I noticed that this would be an imporvement
+%when I found myself wanting to see where the ATA line was between the x axis and the GWR line, but it was 
+%a little hard to see where the x axis was becuase there were no tick marks. It is possible to have tick marks
+%but no numbers of labels in R, and that's what we need here.
+%2) The font of the x axis label, "log fcundity standard deviation, \sigma" seems slightly different from the 
+%other textual labels. Can you please make it consistent? Perhaps it is not bold and the others are bold, or 
+%something.
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig3.pdf}
-\caption{The contribution of ATAs to species coexistence is often meaningfully large compared to 
-other components of the MCT decomposition. All panels show GWR and the decomposition of
+\includegraphics[width=\textwidth]{../results_figs/decompFigLLN.pdf}
+\caption{The contribution of ATAs to species coexistence was often meaningfully large compared to 
+other components of the MCT decomposition for the log-normal fecundities lottery model.
+All panels show GWR and the decomposition of
 GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel, 
@@
 are fixed at values specified in the margins, and GWR and components of the decomposition
-are plotted against $\sigma$. The contribution of ATAs (red line) frequently 
+are plotted against $\sigma$. The contribution of ATAs ($\Delta_i^{[EC]}$) frequently 
 has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
-to species coexistence or failure to coexist. In some instances, the GWR line (orange) falls 
-between the ATA line (red)
-and the $x$ axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of the weaker competitor. Standard errors of plotted quantities 
+to species coexistence or failure to coexist. In some instances, the GWR line falls 
+between the ATA line and the $x$ axis, indicating that, for those parameter values, ATAs 
+make the difference between species
+coexistence and exclusion of the weaker competitor. These instances, which are shaded, correspond to
+ATA rescue or exclusion (Theory). Standard errors of plotted quantities 
 were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
@@
 using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
-in some details, main conclusions were the same. \label{fig:fig3}}
+in detail, main conclusions were the same. \label{fig:fig3}}
 \end{figure}
 
+%***DAN: Jasmin, panel labels(a, b, c, etc.) need to be added. Please use the same format for these labels
+%as you do in the other figures.
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig4.pdf}
-\caption{ATAs can sometimes make the difference between species coexistence and exclusion of the weaker
-competitor: both ``ATA rescue'' and ``ATA exclusion'' (see Methods) are possible. 
-All panels show GWR, the ATA term of the extended MCT [$\Delta_i^{[EC]}$ in equation 
-(\ref{eq:full_decomp})], and the difference of these quantities. 
-For each panel, $\delta$ and $\sigma$ are fixed at values specified in the 
-margins, and the three quantities are plotted against $\mu_1-\mu_2$ to isolate points in the
-``trait-difference space'' corresponding to ATA rescue 
-($\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$) 
-or ATA exclusion ($\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$; 
-see Methods). Values of $\mu_1-\mu_2$ for which GWR is positive but 
-$\text{GWR}-\Delta_i^{[EC]}$ is negative correspond to ATA rescue (shaded yellow), whereas
-values of $\mu_1-\mu_2$ for which GWR is negative but $\text{GWR}-\Delta_i^{[EC]}$ is positive 
-correspond to ATA exclusion (shaded violet). Top panels also include domains
-of ATA rescue or exclusion, but they were too narrow to be highlighted by shading and are
-probably of minimal importance. 
-Standard errors of plotted quantities were never greater than \Sexpr{round(se4,5)}, 
-so are not displayed. The dashed red line on panel (l) corresponds to the parameters
-used for Fig. \ref{fig:pedagog}. This plot uses $q_{ij}=1$. An analogous plot was generated 
-using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$, but it looked exactly the same, by mathematical necessity (see SI section \ref{SI-sect:F4andqij}).}\label{fig:fig4}
+\includegraphics[width=\textwidth]{../results_figs/decompFigLB.pdf}
+\caption{The contribution of ATAs to species coexistence was often meaningfully large compared to 
+other components of the MCT decomposition for the beta fecundities lottery model.
+All panels show GWR and the decomposition of
+GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). For each panel,
+$\delta$ is fixed at the value specified in the upper margin, 
+and GWR and components of the decomposition
+are plotted against $\eta_2/\eta_1$. The contribution of ATAs ($\Delta_i^{[EC]}$) frequently 
+has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
+to species coexistence or failure to coexist. In some instances, the GWR line falls 
+between the ATA line and the $x$ axis, indicating that, for those parameter values, ATAs make 
+the difference between species
+coexistence and exclusion of the weaker competitor. These instances, which are shaded, correspond to
+ATA rescue or exclusion (Theory). Standard errors of plotted quantities 
+were never greater than \Sexpr{round(NA,5)}, so are not displayed. 
+%***DAN: Jasmin, we need to replace the NA in the above line by a variable names, which your code should have saved.
+%The variable needs to be loaded in the R chunk near the top of this file, and then refereced here, please.
+This plot uses $q_{ij}=1$. 
+%***DAN: Right, Jasmin?
+An analogous plot (SI Fig. \ref{SI-fig:fig4_alt}) was generated 
+using $q_{ij}=\eta_i/(\eta_j(1-\delta)+\eta_i \delta)$. 
+%***DAN: Jasmin, I believe the code does not produce the alternate-qij plot yet, but eventually it needs to.
+Though results differed in detail, main conclusions were the same.
+%***DAN: make sure this is indeed the case once the plot is generated.
+}\label{fig:fig4}
 \end{figure}

### 2023-02-26 paper/SupMat.Rnw
finished editing REsults for now, though will come back to it again after Jasmin finishes modifying the diatom figures

@@
 
+%***DAN: Jasmin, panels d-l of this figure are difficult to read because the y axis extent is such that
+%you can't really see what any of the lines are doing except for the ditted and dot-dashed lines. I think we
+%need to present 2 versions of the figure, this one and another one. For the other one, the y-axis limits 
+%should be determined from the y axis extents of all the lines other than the dotted and dot-dashed lines. 
+%Can you please mak this alternative version and let me know when it is made and what the pdf file is, so I 
+%insert it? It is possible you will also need two versions of the figure below, but we won't know that 
+%until you make that figure.
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig3_qij.pdf}
+\includegraphics[width=\textwidth]{../results_figs/decompFigLLNqij.pdf}
 \caption[Analogous to Fig. \ref{MT-fig:fig3} but making a different choice for $q_{ij}$]
@@
 that figure for details. Standard errors of plotted quantities 
-were never greater than \Sexpr{round(se3qij,5)}, so are not displayed. \label{fig:fig3_alt}}
+were never greater than \Sexpr{round(se3qij,5)}, so are not displayed.}\label{fig:fig3_alt}
 \end{figure}
+
+\clearpage
+\newpage
+
+%***DAN: The code does not yet produce this figure. Eventually it needs to.
+%NOTE: I have assumed this figure will look substantially the same as the corresponding main text 
+%figure when I wrote the Results. If it does not, then some revisions will be needed.
+\begin{figure}
+%\includegraphics[width=\textwidth]{../results_figs/decompFigLBqij.pdf}
+\caption[Analogous to Fig. \ref{MT-fig:fig4} but making a different choice for $q_{ij}$]
+{Analogous to Fig. \ref{MT-fig:fig4} of the main text, but using 
+$q_{ij}=\eta_i/(\eta_j(1-\delta)+\eta_i \delta)$. See the caption of
+that figure for details. Standard errors of plotted quantities 
+were never greater than \Sexpr{round(0,5)}, so are not displayed.}\label{fig:fig4_alt}
+%***DAN: jasmin, once the se is generated and saved by your code, the chunk at the outset of this 
+%document needs to be modified to load that quantity, and then the 0 in the above expression 
+%needs to be replaced by the variable name used.
+\end{figure}
+
+\clearpage
+\newpage
 

### 2023-02-27 paper/Paper.Rnw
trivial
small adjustments to control flow for getting SEs and putting them into the paper

@@
 se3 <- readRDS(file=paste0(res_loc,"fig3maxse.RDS"))
-#se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
-se4<-NA
+se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
 se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
@@
 Notation for the whole paper is summarized in Table X.
-%***DAN: make the table
+%***DAN: Dan to make the table
 For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
@@
 ATA rescue or exclusion (Theory). Standard errors of plotted quantities 
-were never greater than \Sexpr{round(NA,5)}, so are not displayed. 
+were never greater than \Sexpr{round(se4,5)}, so are not displayed. 
 %***DAN: Jasmin, we need to replace the NA in the above line by a variable names, which your code should have saved.

### 2023-02-28 paper/Paper.Rnw
took another look at the Results I edited the other day and made some further presentational improvements

@@
 the unit interval). So the beta fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
-and produced fecundities with $U$-shaped, beta distributions between $0$ and $\eta_1$, for $B_1$, and
+and produced fecundities with $U$-shaped distributions between $0$ and $\eta_1$, for $B_1$, and
 $\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
@@
 we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ ranging 
-from $0$ to $7$. Some of 
+from $0$ to $7$. \cite{Chesson_1981} showed that coexistence is impossible for $\delta=1$, and it
+is known that the storage effect (and therefore its ATA component) is weak for larger $\delta$, so we 
+only considered values of $\delta$ up to $0.6$. Some of 
 the $\mu$ differences and $\sigma$ values we considered were very unrealistic (see Results),
@@
 on realistic values of parameters.
-%***DAN: Jasmin mentioned that she only used \delta=0.2, .4, .6 because of a comment Ellner made where he 
-%said it was known that larger values of \delta made for very little storage effects. Read that comment 
-%again and then write something here to jstify the choice. I think I can probably then delete the section of
-%sup mat, mentioned in the commented text below, that talks about \delta=1.
-%Mathematical simplications for the $\delta=1$ case make it straightforward to see that
-%coexistence is impossible and that the case is relatively trivial (SI 
-%section \ref{SI-sect:delta1}), so that case is not plotted.  
 The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
-Computationally efficient means of performing these simulations are in 
+Computationally efficient means of performing those simulations are in 
 SI section \ref{SI-sect:efficient_computation}.
@@
 the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_2/\eta_1$, using 
-the fairly realistic range $1 \leq \eta_2/\eta_1 \leq 5$. 
+the fairly realistic (see Results) range $1 \leq \eta_2/\eta_1 \leq 5$. 
 The simulation methods of \cite{Ellner_2019} were again used.
@@
 were the only means by which coexistence could occur, and ATA effects
-tended to be negative, weakening total storage effects. The ATA term 
-was also often comparable in magnitude to the other terms of 
+tended to be negative, weakening total storage effects.
+
+The ATA term 
+was also often, or always (for the parameters we considered), comparable in magnitude to the other terms of 
 the decomposition (\ref{eq:full_decomp}) for the beta fecundities lottery model 
-(Fig. \ref{fig:fig4}).
+(Fig. \ref{fig:fig4}). The term contributed negatively to coexistence for environmental noise/fecundities
+showing left-tailed association, and contributed positively for noise/fecundities showing right-tailed 
+association.
 
@@
 species 1 is excluded with ATAs but the two species can coexist without ATAs. 
-Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model.
+Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model; 
+examples were more common than for the log-normal fecundities model.
 
@@
 occur, for unrealistic parameters values; but ATA terms were large and ATA rescue and exclusion occurred
-for realistic parameter values for the beta fecundities lottery model. For instance, the ATA in 
-Fig. \ref{fig:fig3} was fairly 
-small, compared to the other terms of (\ref{eq:full_decomp}), for values of $\sigma$ less than 
-about $3$. And ATA rescue and exclusion also only occurred for $\sigma$ greater than about $3$.
+for many realistic parameter values for the beta fecundities lottery model. For instance, the ATA term in 
+Fig. \ref{fig:fig3} had fairly 
+small magnitude, compared to the other terms of (\ref{eq:full_decomp}), for values of $\sigma$ less than 
+about $3$. And ATA rescue and exclusion only occurred for $\sigma$ greater than about $3$.
 The standard deviation of fecundity for species 1 when 
@@
 probably unrealistic. Thus ATAs seem to be unimportant for coexistence
-for the log-normal fecundities lottery model using realistic parameter values. However, 
+for the log-normal fecundities lottery model with realistic parameter values. However, 
 all the $\eta_2/\eta_1$ ratio values we plotted were fairly realistic, corresponding to
@@
 the ATA term plotted in Fig. \ref{fig:fig4} was comparable in magnitude 
-to the other components of (\ref{eq:full_decomp}) across the whole range of $\eta_2/\eta_1$ values
-considered; and both ATA rescue and exclusion occurred for values of this ratio between $1$ and $2$.
-Apparently the question of whether ATAs are important for realistic parameter values depends on model specifics.
+to the other components of (\ref{eq:full_decomp}) across the whole range of parameters
+considered. And both ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
+Apparently the question of whether ATAs are important for realistic parameter values depends on specifics.
 
@@
 \includegraphics[width=\textwidth]{../results_figs/decompFigLLN.pdf}
-\caption{The contribution of ATAs to species coexistence was often meaningfully large compared to 
+\caption{The contribution of ATAs to species coexistence was sometimes, but not always, 
+meaningfully large compared to 
 other components of the MCT decomposition for the log-normal fecundities lottery model.
@@
 are fixed at values specified in the margins, and GWR and components of the decomposition
-are plotted against $\sigma$. The contribution of ATAs ($\Delta_i^{[EC]}$) frequently 
+are plotted against $\sigma$. The contribution of ATAs ($\Delta_i^{[EC]}$) sometimes, but not always, 
 has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
@@
 were never greater than \Sexpr{round(se3,5)}, so are not displayed. This plot uses $q_{ij}=1$. An 
-analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) was generated 
-using $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$. Though results differed
+analogous plot (SI Fig. \ref{SI-fig:fig3_alt}) uses 
+$q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$; though results differed
 in detail, main conclusions were the same. \label{fig:fig3}}
@@
 were never greater than \Sexpr{round(se4,5)}, so are not displayed. 
-%***DAN: Jasmin, we need to replace the NA in the above line by a variable names, which your code should have saved.
-%The variable needs to be loaded in the R chunk near the top of this file, and then refereced here, please.
 This plot uses $q_{ij}=1$. 
 %***DAN: Right, Jasmin?
-An analogous plot (SI Fig. \ref{SI-fig:fig4_alt}) was generated 
-using $q_{ij}=\eta_i/(\eta_j(1-\delta)+\eta_i \delta)$. 
+An analogous plot (SI Fig. \ref{SI-fig:fig4_alt}) uses 
+$q_{ij}=\eta_i/(\eta_j(1-\delta)+\eta_i \delta)$; 
 %***DAN: Jasmin, I believe the code does not produce the alternate-qij plot yet, but eventually it needs to.
-Though results differed in detail, main conclusions were the same.
+though results differed in detail, main conclusions were the same.
 %***DAN: make sure this is indeed the case once the plot is generated.

### 2023-02-28 paper/SupMat.Rnw
took another look at the Results I edited the other day and made some further presentational improvements
tiny things

@@
 section of the main text.
-
-%***DAN: I am 90% sure that once I re-read the comment of Ellner saying we do not need to consider really
-%big \delta, we will be able to delete the below.
-% \section{The $\delta=1$ case}\label{sect:delta1}
-% 
-% %The i terms
-% Setting $\delta=1$ in (\ref{eq:decomp:baseline}) gives 
-% \begin{equation}
-% \epsilon_i^0 = \mu_i - \mu_j.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E}) gives
-% \begin{align}
-% \olep_i^E &= \mu_i-\mu_j-\sigma^2/2-\epsilon_i^0 \\
-% &= -\sigma^2/2.
-% \end{align}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C}) gives
-% \begin{align}
-% \olep_i^C &= \mu_i - \mu_j + \sigma^2/2 -\epsilon_i^0 \\
-% &= \sigma^2/2.
-% \end{align}
-% Setting $\delta=1$ in (\ref{eq:decomp:last_term_primary}) gives
-% \begin{align}
-% \olep_i^{EC} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
-% &= 0.
-% \end{align}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_EC_var_per_se}) gives
-% \begin{align}
-% \olep_i^{(E \shp C)} &= \mu_i - \mu_j -\olep_i^E - \olep_i^C - \epsilon_i^0 \\
-% &= 0.
-% \end{align}
-% Setting $\delta=1$ in (\ref{eq:decomp:storage_effects}) gives
-% \begin{align}
-% \olep_i^{(EC)} &= \mu_i - \mu_j - (\mu_i -\mu_j) \\
-% &= 0.
-% \end{align}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_ATA_2}) gives
-% \begin{equation}
-% \olep_i^{[E \nat C]}=0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_EC}) gives
-% \begin{equation}
-% \olep_i^{[EC]}=0.
-% \end{equation}
-% %The j terms
-% Setting $\delta=1$ in (\ref{eq:decomp:baseline_j}) gives
-% \begin{equation}
-% \epsilon_j^0 = 0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_E_j}) gives
-% \begin{equation}
-% \olep_j^E = -\sigma^2/2.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:contrib_of_var_C_j}) gives
-% \begin{equation}
-% \olep_j^C = \sigma^2/2.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:j:EC}) gives
-% \begin{equation}
-% \olep_j^{EC} = 0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:j:EshpC}) gives
-% \begin{equation}
-% \olep_j^{(E \shp C)} = 0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:j:pECp}) gives
-% \begin{equation}
-% \olep_j^{(EC)} = 0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:j:EnatC}) gives
-% \begin{equation}
-% \olep_j^{[E \nat C]} = 0.
-% \end{equation}
-% Setting $\delta=1$ in (\ref{eq:decomp:j:EChard}) gives
-% \begin{equation}
-% \olep_j^{[EC]}=0.
-% \end{equation}
-% 
-% For $\delta=1$, the second value of $q_{ij}$ we considered, namely
-% $q_{ij}=\frac{\exp(\mu_i)}{(1-\delta)\exp(\mu_j)+\delta \exp(\mu_i)}$, equals $1$, so both the
-% options we considered for $q_{ij}$ are $1$ in this case. Combining the above 
-% equations using $q_{ij}=1$ gives
-% \begin{align}
-% \Delta_i^0 &= \mu_i - \mu_j \\
-% \Delta_i^E &= 0  \\
-% \Delta_i^C &= 0  \\
-% \Delta_i^{(E \shp C)} &=  0 \\
-% \Delta_i^{[E \nat C]} &= 0  \\
-% \Delta_i^{[EC]} &=  0. 
-% \end{align}
-% Thus the only contribution to coexistence for the $\delta=1$ case is $\Delta_i^0 = \mu_i - \mu_j$,
-% which does not depend on $\sigma$ and which is non-positive whenever $\mu_i \leq \mu_j$ (which we assumed,
-% without loss of generality).
-% So there is no robust coexistence in the $\delta=1$ case for our simple model.
 
@@
 Estimates of the temperature dependence functions $Q_i(t)$, $V_i(t)$, and $K_i(t)$, 
-produced by earlier authors \cite{Descamps_2005,Ellner_2016,Ellner_2019} and which are part of the parameterization
+produced by earlier authors \citep{Descamps_2005,Ellner_2016,Ellner_2019} and which are part of the parameterization
 of the diatom competition model described in the main text, are now described.
@@
 single-species, nine-day batch experiments at four different temperatures ($6^\circ$C, 
-$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \cite{Descamps_2005}) and 
+$12^\circ$C, $18^\circ$C, $24^\circ$C) (see Table 1 of \citealp{Descamps_2005}) and 
 used linear interpolations for the
@@
 $C_{i}(t)=(K_i(t)+S(t))/S(t)$, which then imply $r_i(E_i,C_{i})=\frac{V_i(E_i)}{C_{i}} -D$. 
-This choice follows one of the choices made in the R computer code provided by \citep{Ellner_2016,Ellner_2019}.
+This choice follows one of the choices made in the R computer code provided by \cite{Ellner_2016} and 
+\cite{Ellner_2019}.
 Other choices are possible. For instance, in addition to this choice, 
@@
 Environmental temperature fluctuations were defined by the paramters $a$, $P$, and $\theta_0$ using 
-$\theta(t)=asin(2\pi t/P)+\theta_0$, and this determined $E_i(t)$ for $i=1,2$. To obtain 
+$\theta(t)=a\sin(2\pi t/P)+\theta_0$, and this determined $E_i(t)$ for $i=1,2$. To obtain 
 $C_{i \bs i}$, a simulation was carried out starting from $x_i=0$ and $x_j=20$.
@@
 when the distributions of $E_i$ and $C_{i \bs i}$ are known exactly, but for the 
-diatom example explored for goal G4 of the main text (see also SI section \ref{sect:diatom_model}), 
+diatom example, 
 we only had samples from $(E_i,C_{i \bs i})$, based on simulations. We here describe how to

### 2023-02-28 code/LBdecompfig_qij.R
script for making beta decomp with qij !=1

@@ -1,0 +1,108 @@
+#qij !=1 version for beta distributed fecundities lottery model
+source("./decompose_LB.R")
+numRes_loc <- "../results_numeric/"
+betanoise_loc <- paste0(numRes_loc, "betanoise.RDS")
+B<-readRDS(betanoise_loc)
+
+
+fig_loc <- "../results_figs/"
+res_loc <- paste0(numRes_loc,"fig3LB/")
+if(dir.exists(fig_loc)==FALSE){
+  dir.create(fig_loc)
+}
+if(dir.exists(res_loc)==FALSE){
+  dir.create(res_loc)
+}
+
+fig3_LB_loc <- paste0(fig_loc,"decompFigLB.pdf")
+fig3_LB_dat_loc <- paste0(res_loc,"fig3LB_Deltas_dr")
+
+### setting up legend ###
+terms <- c(expression(Delta[i]^0), expression(Delta[i]^E),
+           expression(Delta[i]^C), expression(Delta[i]^"(E#C)"),
+           expression(Delta[i]^"[E||C]"), expression(Delta[i]^"[EC]")
+           ,expression(GWR))
+ncol <- 5
+yor <- hcl.colors(ncol, palette = "YlOrRd")
+cols <- c("black","black","black","black",yor[1:3])
+ltys <- c(1,2,3,4,2,1,1)
+lwds <- c(rep(1.5,4),2,2,3)
+
+#params
+etaj <- seq(1,5,0.1)
+etai<-1; 
+delta<-c(0.2,0.4,0.6)
+
+#start fig
+pdf(fig3_LB_loc, height=8, width=10)
+par(mfrow=c(2,3), oma=c(3,4,4,7), mar=c(1,1,1,1), mgp=c(3,1,0), xpd=NA)
+
+#panel a-c (left tail)
+maxseLT <- {}
+for (d in seq_along(delta)){
+	
+	file <- paste0(fig3_LB_dat_loc,delta[d],"_LT.RDS")
+	sefile <- paste0(fig3_LB_dat_loc,delta[d],"SE_LT.RDS")
+	DeltasLT <- plotco(etai,etaj,delta[d], B,Deltas_loc=file, qij=T)
+	maxseLT[d] <- max(DeltasLT$SE,na.rm=T)
+	mtext(paste(delta[d]), side=3, line=0.2, font=2, cex=1.5, col="gray30")
[…]

### 2023-02-28 code/LBdecompFig.R
edit filee naming

@@ -67,10 +67,10 @@ plotco <- function(etai,etaj,delta, noise,Deltas_loc=0,...){
 	}
 	
 	return(list(D=Deltas, SE=SE))
-	#title(main=paste("mu_j=",params["mu_j"], "sigma_i=",params["sigma_i"], "sigma_j=",params["sigma_j"], "delta=",params["delta"]))
+
 }
 fig_loc <- "../results_figs/"
-res_loc <- paste0(numRes_loc,"fig3LB/")
+res_loc <- paste0(numRes_loc,"LBfigdat/")
 if(dir.exists(fig_loc)==FALSE){
   dir.create(fig_loc)
 }
@@ -79,7 +79,7 @@ if(dir.exists(res_loc)==FALSE){
 }
 
 fig3_LB_loc <- paste0(fig_loc,"decompFigLB.pdf")
-fig3_LB_dat_loc <- paste0(res_loc,"fig3LB_Deltas_dr")
+fig3_LB_dat_loc <- paste0(res_loc,"Deltas_dr")
 
 ### setting up legend ###
 terms <- c(expression(Delta[i]^0), expression(Delta[i]^E),
@@ -126,7 +126,7 @@ for (d in seq_along(delta)){
 	if (file.exists(file)==FALSE){saveRDS(DeltasLT$D,file=file)}
 	if (file.exists(sefile)==FALSE){saveRDS(DeltasLT$SE,file=sefile)}
 }
-#text("LEFT-TAILED ASYMMETRY", x=5.4, y=-0.2, font=2, srt=-90, cex=1.5)
+
 
 #d-f (right-tail)
 maxseRT <- {}
@@ -149,7 +149,6 @@ for (d in 1:length(delta)){
 	if (file.exists(file)==FALSE){saveRDS(DeltasRT$D,file=file)}
 	if (file.exists(sefile)==FALSE){saveRDS(DeltasRT$SE,file=sefile)}
 }
-#text("RIGHT-TAILED ASYMMETRY", x=5.4, y=-0.2, font=2, srt=-90, cex=1.5)
 
 
 #7
@@ -159,9 +158,8 @@ excol <- rgb(38/255, 38/255, 38/255,.5)
 graphics::legend("topright", legend=terms, col = cols, lty = ltys, bty="n", cex=2, inset=c(-.36,-.72), y.intersp = 1.35, x.intersp = 0.1, seg.len=0.8, lwd=lwds)
[…]

### 2023-02-28 code/LBdecompfig_qij.R

@@ -6,7 +6,7 @@ B<-readRDS(betanoise_loc)
 
 
 fig_loc <- "../results_figs/"
-res_loc <- paste0(numRes_loc,"fig3LB/")
+res_loc <- paste0(numRes_loc,"LBqij_figdat/")
 if(dir.exists(fig_loc)==FALSE){
   dir.create(fig_loc)
 }
@@ -14,8 +14,8 @@ if(dir.exists(res_loc)==FALSE){
   dir.create(res_loc)
 }
 
-fig3_LB_loc <- paste0(fig_loc,"decompFigLB.pdf")
-fig3_LB_dat_loc <- paste0(res_loc,"fig3LB_Deltas_dr")
+fig3_LB_loc <- paste0(fig_loc,"decompFigLB_qij.pdf")
+fig3_LB_dat_loc <- paste0(res_loc,"Deltas_dr")
 
 ### setting up legend ###
 terms <- c(expression(Delta[i]^0), expression(Delta[i]^E),

### 2023-02-28 paper/Paper.Rnw
added a table for notation

@@
 in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
-Notation for the whole paper is summarized in Table X.
-%***DAN: Dan to make the table
+Notation for the whole paper is summarized in Table \ref{tab:notation}.
 For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
@@
 %G4 ***DAN: Dan, this will have to be edited after Jasmin gets done editing the corresponding figures
-Fulfilling goal G4 of the Introduction (to provide an empirical example of species coexistence 
+Fulfilling goal G3 of the Introduction (to provide an empirical example of species coexistence 
 which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
@@
 
+%***DAN: Jasmin, please add to the list anyone you talked to at UO about this, and anyone who read it and commented 
+%for you. Be generous here - even short convos merit an acknowlegement.
 The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
-Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Laurn Hallett, and 
+Robin Snyder for helpful discussions. 
 The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
@@
 \bibliography{refs}
+
+\clearpage
+\newpage
+
+\begin{table}[!h]
+\caption{Summary of frequently used notation and abbreviations.}\label{tab:notation}
+\begin{center}
+\begin{tabular}{l|l}
+Notation & Meaning \\
+\hline 
+\hline
+\multicolumn{2}{c}{General} \\
+\hline
+ATA & Asymmetric tail associations \\
+MCT & Modern coexistence theory \\
+GWR & Growth rate when rare \\
+E, C & Environment, competition \\
+$\E$ & Expected value \\ 
+ATA rescue & Coexistence occurs, but not after ATA effects are removed \\
+ATA exclusion & Coexistence does not occur, but does after ATA effects are removed \\
+ \hline
+\multicolumn{2}{c}{Lottery model} \\
+\hline
+$N_i(t)$ & Population density of species $i$, time $t$ \\
+$N$ & $N_1(t)+N_2(t)$ \\
+$B_i(t)$ & Fecundity of species $i$, time $t$ \\
+$\delta$ & Mortality rate \\
+$\sigma$ & Standard deviation of log fecundities for the log-normal fecundities model \\
+$\mu_i$ & Mean log fecundity, species $i$, log-normal fecundities model \\
+$\eta_i$ & Upper bound fecundity, species $i$, log-normal fecundities model \\
+\hline
+\multicolumn{2}{c}{Theory} \\
+\hline
+$r_i(t)$ & Species $i$ growth rate, $\ln(N_i(t+1)/N_i(t))$ \\
+$E_i(t)$ & Environment experienced by the $i^{\text{th}}$ species \\
+$C_i(t)$ & Competition experienced by the $i^{\text{th}}$ species \\
+$\olr_{i \bs i}$ & Mean GWR of species $i$ \\
+$\olr_{j \bs i}$ & Mean growth of $j$ when $i$ is absent \\
+$i \bs i$ & As subscript, refers to computing a quantity for $i$ when $i$ is rare \\
+$j \bs i$ & As subscript, refers to computing a quantity for $j$ when $i$ is absent \\
+$\shp$ & As superscript, refers to computing a quantity after rendering E and C independent \\
+$\nat$ & As superscript, refers to computing a quantity after removing ATAs between E and C \\
+$q_{ij}$ & Scaling factor \\
+$\Delta_i^{(EC)}$ & Storage effects, species $i$ \\
+$\Delta_i^{[EC]}$ & ATA effect to coexistence, species $i$ \\
+$\Delta_i^{[E \nat C]}$ & Component of storage effects due to EC correlation \emph{per se}, species $i$ \\
+\hline
+\multicolumn{2}{c}{Diatom chemostat system} \\
+\hline
+$S$ & Extracellular silicate concentration \\
+$x_1$ & Population density of \emph{F. crotonensis} \\
+$x_2$ & Population density of \emph{C. pseudostelligera} \\
+$\Theta$ & Temperature \\
+$\Theta_0$ & Average temperature \\
+$a$ & Amplitude of temperature fluctuations \\
+$P$ & Period of temperature fluctuations \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2023-03-07 code/LBdecompFig.R
se things

@@ -3,13 +3,14 @@ numRes_loc <- "../results_numeric/"
 betanoise_loc <- paste0(numRes_loc, "betanoise.RDS")
 B<-readRDS(betanoise_loc)
 
-plotco <- function(etai,etaj,delta, noise,Deltas_loc=0,...){
+plotco <- function(etai,etaj,delta, noise,Deltas_loc=0,SE_loc=0,...){
 	
 	if (Deltas_loc!=0){ #check that file exists
 		
-		if (file.exists(Deltas_loc)==FALSE){ 
+		if (file.exists(SE_loc)==FALSE){ 
 			Deltas_loc<-0
-			} else {Deltas <- readRDS(Deltas_loc)}
+			SE_loc <- 0
+			} else {Deltas <- readRDS(Deltas_loc); SE <- readRDS(Deltas_loc)}
 		}
 		
 	if (Deltas_loc==0){
@@ -23,9 +24,9 @@ plotco <- function(etai,etaj,delta, noise,Deltas_loc=0,...){
     	}
     	cat("done")
     	Deltas <- data.frame(t(sapply(store, function(X){X$Delta_i})))
-		colnames(Deltas) <- rownames(store[[1]])
-		SE <- data.frame(t(sapply(store, function(X){X$stanErr_D_i})))
-		colnames(SE) <- colnames(Deltas)
+	colnames(Deltas) <- rownames(store[[1]])
+	SE <- data.frame(t(sapply(store, function(X){X$stanErr_D_i})))
+	colnames(SE) <- colnames(Deltas)
 
 	}
 	
@@ -107,7 +108,7 @@ for (d in seq_along(delta)){
 	
 	file <- paste0(fig3_LB_dat_loc,delta[d],"_LT.RDS")
 	sefile <- paste0(fig3_LB_dat_loc,delta[d],"SE_LT.RDS")
-	DeltasLT <- plotco(etai,etaj,delta[d], B,Deltas_loc=file)
+	DeltasLT <- plotco(etai,etaj,delta[d], B,Deltas_loc=file, SE_loc=sefile)
 	maxseLT[d] <- max(DeltasLT$SE,na.rm=T)
 	mtext(paste(delta[d]), side=3, line=0.2, font=2, cex=1.5, col="gray30")
 	#axis(1, cex.axis=2,tck=-0.028, lwd.ticks=2)
[…]

### 2023-03-07 paper/SupMat.Rnw
loading in se4qij

@@
 se3qij <- readRDS(file="../results_numeric/fig3qijmaxse.RDS")
+se4qij <- readRDS(file="../results_numeric/fig4qijmaxse.RDS")
 @
@@
 \begin{figure}
-%\includegraphics[width=\textwidth]{../results_figs/decompFigLBqij.pdf}
+\includegraphics[width=\textwidth]{../results_figs/decompFigLBqij.pdf}
 \caption[Analogous to Fig. \ref{MT-fig:fig4} but making a different choice for $q_{ij}$]
@@
 that figure for details. Standard errors of plotted quantities 
-were never greater than \Sexpr{round(0,5)}, so are not displayed.}\label{fig:fig4_alt}
+were never greater than \Sexpr{round(se4qij,5)}, so are not displayed.}\label{fig:fig4_alt}
 %***DAN: jasmin, once the se is generated and saved by your code, the chunk at the outset of this 
@@
 %needs to be replaced by the variable name used.
+%***JAS:done
 \end{figure}

### 2023-03-07 code/Figure3.R
fixed right hand label

@@ -97,8 +97,8 @@ for (m in seq_along(mudif)){ #iterates across mudif values
       graphics::text(paste(mudif[m]), srt=-90, x=7.8,y=mean(ylims[[m]]),  font=2, cex=2, col="gray30")
     }
     if (n==6){
-    	graphics::text("mean log fecundity difference, ", x=9,y=2.2, font=2, cex=2.5, srt=-90, adj=0, col="gray40")
-    	graphics::text(expression(mu[1]-mu[2]), x=9,y=-3.3, font=2, cex=3, srt=-90, adj=1, col="gray40")
+    	graphics::text("mean log fecundity difference, ", x=9,y=2.45, font=2, cex=2.5, srt=-90, adj=0, col="gray40")
+    	graphics::text(expression(mu[1]-mu[2]), x=9,y=-3.95, font=2, cex=3, srt=-90, adj=1, col="gray40")
     }
     if (n==11){
     	graphics::mtext("log fecundity standard deviation, ",side=1, line=2.8, cex=1.25, at=-3.25, adj=0,col="gray40")
@@ -168,8 +168,8 @@ for (m in seq_along(mudif)){ #iterates across mudif values
       graphics::text(paste(mudif[m]), srt=-90, x=7.8,y=mean(ylims[[m]]),  font=2, cex=2, col="gray30")
     }
     if (n==6){
-    	graphics::text("mean log fecundity difference, ", x=9,y=2.2, font=2, cex=2.5, srt=-90, adj=0, col="gray40")
-    	graphics::text(expression(mu[1]-mu[2]), x=9,y=-3.3, font=2, cex=3, srt=-90, adj=1, col="gray40")
+    	graphics::text("mean log fecundity difference, ", x=9,y=2.45, font=2, cex=2.5, srt=-90, adj=0, col="gray40")
+    	graphics::text(expression(mu[1]-mu[2]), x=9,y=-3.95, font=2, cex=3, srt=-90, adj=1, col="gray40")
     }
     if (n==11){
     	graphics::mtext("log fecundity standard deviation, ",side=1, line=2.8, cex=1.25, at=-3.25, adj=0,col="gray40")

### 2023-03-20 paper/Paper.Rnw
commits I should have made a couple weeks ago

@@
 
+%One-parameters plots
+To display results for goal G3, we started by plotting the terms of 
+(\ref{eq:full_decomp}) against one of the parameters 
+$\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
+used by \cite{Descamps_2005}. We again labelled regions of parameters space corresponding to
+ATA rescue or exclusion.
+
+%Two-parameter plots
 %***DAN: This para will likely have to be revised depending on how Jasmin revises the diatom figure presentation
-To display results for goal G3, we plotted the terms of (\ref{eq:full_decomp}) against one of the parameters 
-$\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
-used by \cite{Descamps_2005}. We also generated two-dimensional plots, where
+We also generated two-dimensional plots, where
 two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
@@
 
-%G4 ***DAN: Dan, this will have to be edited after Jasmin gets done editing the corresponding figures
-Fulfilling goal G3 of the Introduction (to provide an empirical example of species coexistence 
-which highlights the role ATAs can play), Fig. \ref{fig:fig5} and Fig. \ref{fig:fig6} 
+%G3 ***DAN: check this again after Jasmin gets done making the new version of Fig. 5
+Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
+which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
 show that ATA contributions to coexistence often have magnitudes comparable to the 
-magnitudes of other terms of the extended MCT decomposition (Fig. \ref{fig:fig5}), and that 
-ATA rescue occurs for many combinations of parameters for this system (Fig. \ref{fig:fig5},
-but especially Fig. \ref{fig:fig6}).
+magnitudes of other terms of the extended MCT decomposition, and that 
+ATA rescue occurs for some combinations of parameters for this system (Fig. \ref{fig:fig5}a-c).
+
+%***DAN: finish it once Jasmin finishes d-f
+Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f show that...
 
@@
 
+%***DAN: Jasmin, can you please ensure that ATA enclusion and rescue are 
+%shaded and keyed in the legend for panels a-c in the same way as the earlier figures?
 \begin{figure}
 \includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
+\caption{ATA effects and ATA rescue and exclusion in the diatom system. 
+The contribution of ATAs to the GWR of \emph{F. crotonensis}
 in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
-other components of the MCT decomposition. All panels show GWR and the decomposition of
+other components of the MCT decomposition (a-c). Panels a-c show GWR and the decomposition of
 GWR provided by the extended MCT, equation (\ref{eq:full_decomp}). 
@@
 GWR and components of the MCT decomposition
-were then plotted against the third parameter. The original values of the varying parameters 
-are indicated by vertical dashed magenta lines. ATA contributions 
+were plotted against the third parameter. The original values of the varying parameters 
+are indicated by vertical dotted lines. 
+%***DAN: Change the above sentence if Jasmin changes the font of the line
+ATA contributions 
 had magnitude comparable to other components, 
 indicating that ATAs can contribute meaningfully
-to diatom coexistence in this system. In some instances, the GWR line (orange) falls 
-between the ATA line (red)
-and the horizontal axis, indicating that, for those parameter values, ATAs make the difference between species
-coexistence and exclusion of \emph{F. crotonensis}. 
+to diatom coexistence in this system. In some instances, the GWR line falls 
+between the ATA line
+and the horizontal axis, indicating that, for those parameter values, ATA rescue occurs. 
 Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
@@
 excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
-sufficient coexistence metric.}  \label{fig:fig5}
+sufficient coexistence metric.
+%***DAN: Once Jasmin is done with the other part of the figure (the part that used to be Fig. 6), 
+%finish this caption. Use the below caption, currently commented, to help. 
+}  \label{fig:fig5}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-\caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
-in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
-the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
-indicating greater importance of ATAs and values greater than $1$ corresponding to either
-ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
-$\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
-magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
-were remapped to 1 and -1, respectively, for readability.
-For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
-excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
-Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
-\end{figure}
+% \begin{figure}
+% \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
+% \caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
+% in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
+% the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
+% indicating greater importance of ATAs and values greater than $1$ corresponding to either
+% ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
+% $\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
+% magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
+% were remapped to 1 and -1, respectively, for readability.
+% For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
+% excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
+% Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
+% \end{figure}
 

### 2023-03-20 paper/SupMat.Rnw
commits I should have made a couple weeks ago

--- 2023-02-28 paper/SupMat.Rnw
@@
 se3qij <- readRDS(file="../results_numeric/fig3qijmaxse.RDS")
+#se4qij <- readRDS(file="../results_numeric/fig4qijmaxse.RDS")
 @
@@
 
-%***DAN: Jasmin, panels d-l of this figure are difficult to read because the y axis extent is such that
-%you can't really see what any of the lines are doing except for the ditted and dot-dashed lines. I think we
-%need to present 2 versions of the figure, this one and another one. For the other one, the y-axis limits 
-%should be determined from the y axis extents of all the lines other than the dotted and dot-dashed lines. 
-%Can you please mak this alternative version and let me know when it is made and what the pdf file is, so I 
-%insert it? It is possible you will also need two versions of the figure below, but we won't know that 
-%until you make that figure.
+%***DAN: Jasmin, please fix the collision of text that occurs in the text near the legend
 \begin{figure}
@@
 
-%***DAN: The code does not yet produce this figure. Eventually it needs to.
-%NOTE: I have assumed this figure will look substantially the same as the corresponding main text 
-%figure when I wrote the Results. If it does not, then some revisions will be needed.
 \begin{figure}
-%\includegraphics[width=\textwidth]{../results_figs/decompFigLBqij.pdf}
+\includegraphics[width=\textwidth]{../results_figs/decompFigLBqij.pdf}
 \caption[Analogous to Fig. \ref{MT-fig:fig4} but making a different choice for $q_{ij}$]

### 2023-03-23 paper/Paper.Rnw
pretty complete job editing for now, just waiting for Jasmin's figure work to complete my editing

@@
 At the core of ecology is the study of relationships - between organisms and between organisms and their environments. 
-Ecologists may ask, for instance, how fluctuations in local- and regional-scale environments relate to 
-population dynamics. To study such relationships and to understand how two variables interact, ecologists 
+Ecologists may ask, for instance, how local environment relates to 
+population growth rate. To study such relationships and to understand how two variables interact, ecologists 
 very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
-and direction of association between variables often don't tell the whole story 
+and direction of association between variables often do not tell the whole story 
 \citep{Ghosh_2020_AER}.
@@
 Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
-regression), but correlations are extremely common, and other methods also reveal only a small portion 
+regression), but correlations are extremely common, and other common methods also reveal only a small portion 
 of the information content of the relationship between variables \citep{Nelsen2006}.
@@
 what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
-may be relevant to them. Storage effects can allow for species to coexist while fluctuating in abundance through time
-because storage effects can provide a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
+may be relevant to them. Storage effects allow competing species to fluctuate in abundance without any going extinct
+by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
 Storage effects hinge on positive covariation between the fluctuating quality of the environment
@@
 precise nature of that influence is the subject of the rest of the paper.
-The lottery model is as follows.
-Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
+The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
+is as follows.
+Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
 model equations are
@@
 \end{equation}
-for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
-The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
+for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
+The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
 in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
@@
 ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).
-
-%***DAN: I commented the below because we decided it does not belong in the Intro, but a version of it 
-%should be in the Discussion. Keeping it for now until I have a chance to move it there. Note also that
-%I now think this material needs some editing for clarity, some of that was done in a paper version I 
-%marked up in January 2023.
-% %Climate change may alter ATAs of coexistence-relevant ATAs
-% An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
-% alter ATAs of environmental variables which impact coexistence, as follows.
-% It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
-% heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
-% If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
-% may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
-% For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
-% whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
-% contributing to their coexistence, or failure to coexist, can boil down to covariation 
-% between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
-% to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
-% longer lasting increases the degree of right-tail association between early and late spring rainfall. 
-% Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
-% left-tail association between these variables. As we will show in this study, changes in ATAs can then
-% result in different competitive outcomes between species. 
-% \cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
-% community of desert annuals that seem to have all the ingredients necessary to make effects of the
-% type we describe here a possibility.
-% To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
-% not been studied, but the importance for coexistence of changes 
-% in means and variability of environmental variables has been studied in many systems
-% [e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
-% Changes in variability and in extreme events are distinct from the changes in ATAs we 
-% consider here, because ATAs pertain to \emph{relationships} between environmental and other
-% variables in the extremes, which are distinct from and can be altered independently
-% of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
@@
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
+$C_i(t)$. Though the iid assumption is not necessary, it simplifies the
 presentation of new ideas.
@@
 The value of $\ddot{\rho}_i$ is then determined in one of 
-four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
 $E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
@@
 make no reference to covariation between quantities, whereas ATAs are 
-defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
+strictly about covariation. Therefore, storage effects are the only mechanism of coexistence
 for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
@@
 %Setup - two versions, and basics which apply to both
-To pursue goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
+To achieve goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
 comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
@@
 
-%G2 
+\subsection{Goal G2, lottery model results}
+
 To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
@@
 
-%G3 ***DAN: check this again after Jasmin gets done making the new version of Fig. 5
+\subsection{Goal G3, results for the diatom system}
+
+%***DAN: check this again after Jasmin gets done making the new version of Fig. 5
 Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
@@
 %Summary-type paragraph of main results
-We produced and applied a new extension of MCT that 
-quantifies the impact of ATAs on species coexistence and the maintenance of biodiversity. 
+We produced and applied a new extension of modern coexistence theory (MCT) that 
+quantifies the impact of asymmetric tail associations (ATAs) on species coexistence and the maintenance of biodiversity. 
 Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
@@
 and the contribution of $EC$ correlation \emph{per se}.
-Applications of our extended theory to the lottery model and to models based on a laboratory 
-system of competing diatoms demonstrated that  
-ATA contributions to storage effects and to species coexistence and the maintenance of biodiversity
+Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
+ATA contributions to storage effects and to species coexistence 
 can often be comparable in magnitude to the contributions of previously recognized mechanisms.
-Our results highlight that common methods of correlation and covariance fail to include 
-important information on the structure of relationships between variables. Our results add
+Our results add
 coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
@@
 our results suggest that storage effects should be redefined to constitute the 
-contribution of \emph{associations} between $E$ and $C$, including both ATA effects and
-correlation \emph{per se}; the prior terminology, \emph{covariance}, is typically construed 
+contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
+correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
 as reflecting the linear relationship between variables, ignoring complexities of the association,
-such as ATAs, which we now know are important.
+such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
+in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
+(the beta fecundities lottery model), including our empirically
+based example (the diatom system). Though future work should seek to understand precisely when ATAs are or are not important
+for coexistence, our work demonstrates the overall importance of this new mechanism.
 
@@
 Contributions of ATAs to species coexistence open the possibility that
-``ATA specialists'' may exist in some contexts. If a strong competitor is present across a
-collection of habitat patches and the GWR of a weaker competitor is only positive 
-in a patch in the presence of 
+``ATA specialists'' may exist. For instance, if a strong competitor is present across a
+collection of habitat patches on a landscape and the GWR of a weaker competitor is only positive 
+in a patch if there are 
 ATAs between the temporal fluctuations of two environmental variables in that patch,
@@
 The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
-can often boil down to associations between two environmental variables, as was assumed to be the case
-for the lottery model. Mechanisms of coexistence should correspond to
+can boil down to associations between two environmental variables, as was the case
+for the lottery model. For the lottery model, $B_i(t)=E_i(t)$ was assumed, and $C_{i \bs i}$ simplified to $E_j(t)/\delta$ (see SI section 
+\ref{SI-SIsect:TheoryForLotteryBoth}). Thus, EC-associations parallel associations between the two
+environmental variables $E_i(t)$ and $E_j(t)$ for the lottery model, and ATA specialists may exist. 
+Mechanisms of coexistence should correspond to
 means by which niche differentiation is possible. Thinking about our new mechanism of 
@@
 imagine a scenario in which a weaker competitor can only exist in habitat patches 
-for which ATAs are absent, and is hence a specialist on scenarios with symmetric
-tail associations of environmental variables (an ``STA specialist''). Additional 
+for which ATAs between environmental variables are absent, and is hence a specialist on scenarios with symmetric
+tail associations of variables (an ``STA specialist''). Additional 
 research will be needed to assess to what degree these possibilities are realized 
@@
 
-%future work: study ATAs in environmental variables and how theyre changing - tie in with specialist
-In the Introduction we explained a plausible means by which climate change may
-alter the tail association properties of phenologically separated environmental variables
-such as early- and late-spring rainfall. The possibility 
-that changes in the nature of environmental ATAs 
-may result in the extirpation of ATA and STA specialists indicates that it may
-be important to engage in further study of ATAs between environmental variables 
-and how they differ across space and may be changing through time. 
-Extreme environmental events have been much studied, including
-how the strength and commonness of such events may be changing with climate change, and how such changes 
-may influence coexistence (see citations in the Introduction). But
-studies of changes in environmental ATAs would differ from such prior work because
-ATAs pertain to \emph{relationships} between variables in the extremes.
+%Climate change may alter ATAs of coexistence-relevant ATAs
+An additional motivation for quantifying the importance of ATAs for species coexistence 
+is that there is reason to believe climate change may
+alter ATAs of environmental variables which impact coexistence, as follows.
+It is well known that climate change can alter the intensity of extreme environmental events, including
+heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.
+If extreme events are not only becoming more intense, but are also increasing in duration (e.g., heat waves 
+are not only hotter but also last longer, \citealp{Meehl_2004,Lyon2019,Keelings2020}), 
+then the nature of the covariation between phenologically separated environmental signals
+may be modified by these changes; in particular, ATAs between such variables may be altered.
+For instance, imagine a case in which early spring temperature influences the growth of a plant species, A,
+whereas late spring temperature influences plant species B. If these species are competitors, then EC covariation
+contributing to their coexistence, or failure to coexist, could boil down to covariation
+between early and late spring temperatures (e.g., if competitive dynamics are governed by the lottery model with $E_i$ representing
+early spring temperature and $E_j$ representing late spring temperature, see previous paragraph). 
+Because longer lasting heat waves are more likely to contribute
+to both early and late spring temperature, a tendency for heat waves to become both more extreme and
+longer lasting increases the degree of right-tail association between early and late spring temperature.
+Likewise, a tendency for cold snaps to become both more intense (i.e., colder temperatures) and longer lasting can increase
+left-tail association between early- and late-spring temperatures. As we showed in this study, changes in ATAs can then
+result in different competitive outcomes between species.
+To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has
+not been studied, though the importance for coexistence of changes
+in means and variability of environmental variables has been studied in many systems
+[e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
+Changes in variability and in extreme events are distinct from the changes in ATAs we
+consider here, because ATAs pertain to \emph{relationships} between environmental and other
+variables in the extremes, which are distinct from and can be altered independently
+of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
 
@@
 environment directly, $B_i(t)=E_i(t)$. But vital rates such as fecundity are 
-more likely related to actual environmental variables according to a nonlinear
+more likely related to commonly measured environmental variables according to a nonlinear
 function which may often be threshold-like or sigmoidal, and thus the $E_i=B_i$
@@
 %Differences with the original, analytic version of the MCT
-%
-%***DAN: I am not so comfortable with this because I have insufficient 
-%knowledge of the original MCT. I also do not want to devote the time
-%to learning it right now because of other commitments. I've reworded, 
-%trying not to change the meaning. Jasmin, how confident are you in 
-%these statements? We will also want to try to get Steve Ellner to check 
-%them.
 Our approach to exploring the influence of ATAs on coexistence relies on the computational
@@
 one such strength is the generality that analytic approaches provide.
-It may be possible to extend the mathematics of the original MCT to consider ATAs.
+It may be possible to expand the mathematics of the original MCT to consider ATAs.
 
@@
 $E$ and $C$ can be considered and used to produce a decomposition of GWR. For instance, if $X$, $Y$ and 
-$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, the decomposition
-\begin{align}
-\text{GWR}_i &= \Delta_i^0+\Delta_i^X+\Delta_i^Y+\Delta_i^Z \label{eq:9}\\
-&+\Delta_i^{(X \shp Y)} +\Delta_i^{(XY)} + \Delta_i^{(Y \shp Z)} + \Delta_i^{(YZ)} +\Delta_i^{(X \shp Z)}+\Delta_i^{(XZ)} \label{eq:10}\\
-&+ \Delta_i^{(XYZ)}+\Delta_i^{(XY \shp Z)}+\Delta_i^{(X \shp YZ)}+\Delta_i^{(XZ \shp Y)}+\Delta_i^{(X \shp Y \shp Z)}.\label{eq:11}
-\end{align}
-The terms in (\ref{eq:9}) and (\ref{eq:10}) are analogous to those of (\ref{eq:previous_decomp}).
-The new terms in (\ref{eq:11}) result from rendering one or 
+$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, a decomposition
+of the GWR of $i$ for which some terms were based on rendering one or 
 more of $X$, $Y$, and $Z$ independent of the others. 
 We simply comment that this general decomposition should also be extendable to consider the effects of
-ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}), though the complexity
-of such a decomposition may present challenges.
-For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
+ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}). See SI section \ref{SI-sect:threeway} for details.
+
+%close
 It seems reasonable to suppose that ATAs will play an increasingly important role in systems of 
@@
 %as you do in the other figures.
+%Also please add tick marks to all axes (see comment on previous figure)
 \begin{figure}
@@
 
-%***DAN: Jasmin, can you please ensure that ATA enclusion and rescue are 
+%***DAN: Jasmin, can you please ensure that ATA exclusion and rescue are 
 %shaded and keyed in the legend for panels a-c in the same way as the earlier figures?
@@
 %***DAN: Change the above sentence if Jasmin changes the font of the line
-ATA contributions 
+ATA contributions ($\Delta_i^{[EC]}$)
 had magnitude comparable to other components, 

### 2023-03-23 paper/SupMat.Rnw
pretty complete job editing for now, just waiting for Jasmin's figure work to complete my editing

--- 2023-03-07 paper/SupMat.Rnw
@@
 
+\section{    Three-way decompositions}\label{sect:threeway}
+
+As mentioned in the Discussion in the main text, \cite{Ellner_2019} elaborated 
+a general theory in which more and different varying factors than
+$E$ and $C$ can be considered and used to produce a decomposition of GWR. For instance, if $X$, $Y$ and 
+$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, the decomposition
+\begin{align}
+\text{GWR}_i &= \Delta_i^0+\Delta_i^X+\Delta_i^Y+\Delta_i^Z \label{eq:9}\\
+&+\Delta_i^{(X \shp Y)} +\Delta_i^{(XY)} + \Delta_i^{(Y \shp Z)} + \Delta_i^{(YZ)} +\Delta_i^{(X \shp Z)}+\Delta_i^{(XZ)} \label{eq:10}\\
+&+ \Delta_i^{(XYZ)}+\Delta_i^{(XY \shp Z)}+\Delta_i^{(X \shp YZ)}+\Delta_i^{(XZ \shp Y)}+\Delta_i^{(X \shp Y \shp Z)}.\label{eq:11}
+\end{align}
+The terms in (\ref{eq:9}) and (\ref{eq:10}) are analogous to those of main text equations (\ref{MT-eq:previous_decomp}).
+The new terms in (\ref{eq:11}) result from rendering one or 
+more of $X$, $Y$, and $Z$ independent of the others.
+We simply comment that this general decomposition should also be extendable to consider the effects of
+ATAs in the same way main text equation (\ref{MT-eq:full_decomp}) extends main text equation (\ref{MT-eq:previous_decomp}), 
+though the complexity
+of such a decomposition may present challenges.
+For example, $\Delta^{(XY \shp Z)}$ can be decomposed as $\Delta^{[XY \shp Z]} + \Delta^{[X \nat Y \shp Z]}$.
+
 \clearpage
@@
 
-%***DAN: The code does not yet produce this figure. Eventually it needs to.
-%NOTE: I have assumed this figure will look substantially the same as the corresponding main text 
-%figure when I wrote the Results. If it does not, then some revisions will be needed.
 \begin{figure}

### 2023-04-26 paper/Paper.Rnw
working on (hopefully) final edits

@@
 what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
-may be relevant to them. Storage effects allow competing species to fluctuate in abundance without any going extinct
+may be relevant to them. Storage effects allow competing species to fluctuate in abundance, without any going extinct,
 by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
@@
 Notation for the whole paper is summarized in Table \ref{tab:notation}.
-For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
-distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
+We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
+distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}d-f), 
 $B_1$ and $B_2$ were, respectively, 
-left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
-tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
+left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symmetrically (Fig. \ref{fig:pedagog1}b)
+tail associated, while the same marginal distributions for these fecundities were used in all 
+simulations (see Methods
 for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
@@
 following specific goals. 
-G1) To augment MCT to show formally how ATAs play a role in coexistence. 
+G1) We will augment MCT to show formally how ATAs play a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
@@
 spatial variation can be considered in future work.
-G2) To explore, using the simple lottery model, the relative magnitudes of the contributions to coexistence that come
+G2) We will explore, using the simple lottery model, the relative magnitudes 
+of the contributions to coexistence that come
 from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
@@
 are likely to be negligible or meaningfully large. 
-G3) To provide an empirical example of species coexistence which highlights the role ATAs can play and the
+G3) We will provide an empirical example of species coexistence which highlights the role ATAs can play and the
 potential importance of ATAs relative to other mechanisms of coexistence.
@@
 and a means of understanding its theoretical and empirical importance.
+
+%***DAN: Dan made it to here on 2022 04 26
 
@@
 %Two-parameter plots
-%***DAN: This para will likely have to be revised depending on how Jasmin revises the diatom figure presentation
 We also generated two-dimensional plots, where
@@
 at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
-$\Delta_i^{[EC]}/\text{GWR}$ was displayed using color, to convey the 
+$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color, to convey the 
 importance of the ATA contribution relative to the whole GWR. Larger magnitudes
 of this quantity indicate greater importance of ATAs. The case
-$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/\text{GWR}>1$)
+$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/|\text{GWR}|>1$)
 corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
-(which also produces $\Delta_i^{[EC]}/\text{GWR}>1$) corresponds to ATA exclusion.
-Thus the condition $\Delta_i^{[EC]}/\text{GWR}>1$ encompasses the cases for which ATA
-exclusion or rescue occurred, though the condition does not delineate, by itself,
-which of these was the case.
-
-All computations were done in R on an ordinary laptop running Ubuntu Linux. 
-Complete computer codes for the project are stored at
-\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
+(which produces $\Delta_i^{[EC]}/|\text{GWR}| < -1$) corresponds to ATA exclusion.
+Thus the value of $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
+and also whether ATA rescue or exclusion occurred.
+
+All computations were done in R on a laptop running Ubuntu Linux. 
+Complete computer codes for the project are stored at (link to be added).
+%\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
 
@@
 considered. And both ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
-Apparently the question of whether ATAs are important for realistic parameter values depends on specifics.
+Apparently the question of whether ATAs are important for realistic parameter 
+values depends on specifics.
 
@@
 
-%***DAN: check this again after Jasmin gets done making the new version of Fig. 5
 Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
@@
 magnitudes of other terms of the extended MCT decomposition, and that 
-ATA rescue occurs for some combinations of parameters for this system (Fig. \ref{fig:fig5}a-c).
-
-%***DAN: finish it once Jasmin finishes d-f
-Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f show that...
+ATA rescue occurs for some combinations of parameters for this system.
+
+Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f reinforce the 
+same conclusions. The plotted quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
+indicating that ATAs were often a substantial contributor to whether the diatom species 
+coexist. For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
+$\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1/4$, with large portions of
+Fig. \ref{fig:fig5}d-e satisfying the same criterion. Substantial portions of 
+panels d-f also show values of $\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1$,
+corresponding to ATA rescue. ATA exclusion did not occur for this system for the 
+parameter ranges we considered.
 
@@
 Robin Snyder for helpful discussions. 
-The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
+The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 
 The authors were partly supported by US National
@@
 
+%***DAN: Jasmin, could you, with reasonable effort, move panels around? Specifically, put the panels which
+%are now g ang h in the middle instead of on the right (and they will then be called panels d and e) and 
+%put what are now panels d-f on the right (and thwill then be called f-h)? The reason is, we refer to
+%g-h before we refer to d-f, and the reader may become confused by this, including possibly havig trouble 
+%finding the parts of the caption that refer to g-h while getting confused about the parts of the caption 
+%that refer to d-e, which are about something that has not even been introduced at the point at which the 
+%reader first is referred th panels g-h. 
 \begin{figure}
@@
 %but no numbers of labels in R, and that's what we need here.
-%2) The font of the x axis label, "log fcundity standard deviation, \sigma" seems slightly different from the 
+%2) The font of the x axis label, "log fecundity standard deviation, \sigma" seems slightly different from the 
 %other textual labels. Can you please make it consistent? Perhaps it is not bold and the others are bold, or 
@@
 
-%***DAN: Jasmin, can you please ensure that ATA exclusion and rescue are 
-%shaded and keyed in the legend for panels a-c in the same way as the earlier figures?
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{ATA effects and ATA rescue and exclusion in the diatom system. 
+%\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
+\caption{ATA effects and ATA rescue in the diatom system. 
 The contribution of ATAs to the GWR of \emph{F. crotonensis}
@@
 are indicated by vertical dotted lines. 
-%***DAN: Change the above sentence if Jasmin changes the font of the line
 ATA contributions ($\Delta_i^{[EC]}$)
@@
 indicating that ATAs can contribute meaningfully
-to diatom coexistence in this system. In some instances, the GWR line falls 
+to coexistence in this system. In some instances, the GWR line falls 
 between the ATA line
@@
 Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
-For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, 
-was always positive both including and
+Panels d-f show bivariate plots (Methods) for which the third parameter was 
+fixed at the values used by \cite{Descamps_2005}. ATAs are again often an important factor helping
+determine coexistence or exclusion, and ATA rescue happened for an appreciable fraction of the
+explored parameter space.
+For the parameters we considered, GWR of the other diatom species in the system, 
+\emph{C. pseudostelligera}, was always positive both including and
 excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
 sufficient coexistence metric.
-%***DAN: Once Jasmin is done with the other part of the figure (the part that used to be Fig. 6), 
-%finish this caption. Use the below caption, currently commented, to help. 
 }  \label{fig:fig5}

### 2023-05-07 paper/Paper.Rnw
Dan finished going through the paper, about to try cutting 550 words

@@
 
-\author[a]{Pimsupa Jasmin Albert}
-\author[a,b,*]{Daniel C. Reuman}
+\author[a,b]{Pimsupa Jasmin Albert}
+\author[a,c,*]{Daniel C. Reuman}
 
 \affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
-\affil[b]{Laboratory of Populations, Rockefeller University}
+\affil[b]{Environmental Studies Program and Department of Biology, Universit of Oregon}
+\affil[c]{Laboratory of Populations, Rockefeller University}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 \maketitle
+
+\noindent \textbf{Running head:} Asymmetric relationships and coexistence \\
+\noindent \textbf{Key words:} asymmetric tail association, biodiversity, coexistence, competition, copula, modern coexistence theory, niche differentiation, storage effects \\
+\noindent \textbf{Type of article:} Letter\\
+\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 5518; \textbf{Words in text boxes:} 0\\
+\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
+\noindent \textbf{Corresponding author:}
+Daniel C. Reuman,
+Department of Ecology and Evolutionary Biology,
+University of Kansas,
+2101 Constant Ave,
+Lawrence, KS 66047;
+Phone: 626-560-7084 \\
+\noindent \textbf{Author emails:} pimsupaa@uoregon.edu, reuman@ku.edu\\
+\noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
+\noindent \textbf{Data accessibility statement:} All codes associated with this study will be released
+in a Zenodo repository prior to publication.
+
+\doublespacing
 
@@
 
-\keywords{asymmetric tail association, biodiversity, coexistence, competition, competitive exclusion, modern coexistence theory, niche differentiation, storage effects}
-
 \section{Introduction}\label{section:introduction}
@@
 when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
-ecosystem functioning. Fig. \ref{fig:pedagog1}g,h show contrasting examples of ATAs in nature using 
+ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
 plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
@@
 We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
-distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}d-f), 
+distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}f-h), 
 $B_1$ and $B_2$ were, respectively, 
@@
 for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
-ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).
+ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
 
@@
 and a means of understanding its theoretical and empirical importance.
-
-%***DAN: Dan made it to here on 2022 04 26
 
@@
 for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
-\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
+\ref{SI-SIsect:other_mechanisms}. \cite{Ellner_2016} and
 \cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
 which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
-original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
-coexistence operating in a given case are illuminated. The decomposition is 
+original decomposition of \cite{Chesson_1994}, 
 \begin{equation}
@@
 $\Delta_i^C$ is a contribution to coexistence of variation in competition;
-$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
+$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation 
+in both environment and competition, but not including the effects 
 of covariation in these quantities; and
 $\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
-Only storage effects involve $EC$ covariation,
-so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
+Combining our decomposition 
 (\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
@@
 distribution and $F_\beta$ is the CDF of a beta distribution 
-with shape and scale parameters $0.5$ (probability density function proportional to $x^{-05}(1-x)^{-0.5}$ on
+with shape and scale parameters $0.5$ (probability density function proportional to $x^{-0.5}(1-x)^{-0.5}$ on
 the unit interval). So the beta fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
@@
 \end{align}
-for $j=1,2$, where $S$ is a extracellular silicate concentration in the chemostat, $x_1$ is the population density
+for $j=1,2$, where $S$ is an extracellular silicate concentration in the chemostat, $x_1$ is the population density
 of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
@@
 $\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
-used by \cite{Descamps_2005}. We again labelled regions of parameters space corresponding to
+used by \cite{Descamps_2005}. We again labelled regions of parameter space corresponding to
 ATA rescue or exclusion.
@@
 We produced and applied a new extension of modern coexistence theory (MCT) that 
-quantifies the impact of asymmetric tail associations (ATAs) on species coexistence and the maintenance of biodiversity. 
+quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
 Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
@@
 
-%***DAN: Jasmin, please add to the list anyone you talked to at UO about this, and anyone who read it and commented 
-%for you. Be generous here - even short convos merit an acknowlegement.
 The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
@@
 
-%***DAN: Jasmin, could you, with reasonable effort, move panels around? Specifically, put the panels which
-%are now g ang h in the middle instead of on the right (and they will then be called panels d and e) and 
-%put what are now panels d-f on the right (and thwill then be called f-h)? The reason is, we refer to
-%g-h before we refer to d-f, and the reader may become confused by this, including possibly havig trouble 
-%finding the parts of the caption that refer to g-h while getting confused about the parts of the caption 
-%that refer to d-e, which are about something that has not even been introduced at the point at which the 
-%reader first is referred th panels g-h. 
 \begin{figure}
@@
 tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
-see also (g,h)]. 
+see also (d,e)]. 
 The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
 pictured and then transforming the marginals as specified in Methods.
-(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
+(d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
+\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
+in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (d) showing 
+left-tail association and (e) showing right-tail association.
+(f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
 respectively corresponding variables (a-c; see Introduction).
-(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
-\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
-in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
-left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
+\label{fig:pedagog1}}
 \end{figure}
 
-%***DAN: This looks pretty good except the histograms for some marginals do not really look like the plotted
+%***DAN: This looks pretty good except one of the histograms on c does not look like the plotted
 %pdfs of the marginals. I think Jasmin said in an email she was fixing this.
+%***DAN: Jasmin, also, we use \varphi in the text, but \phi here. These need to be made consistent, since they are different to
+%a math person. The easiest way will be to change it in the figure, so I hope that can straightforwardly be done.
 \begin{figure}
@@
 $E_i^{||}$ and  $C_{i \bs i}^{||}$,
-%***DAN: I replaced \nat in the above two expressions by || because for some reason the \nat does not
+%Dan replaced \nat in the above two expressions by || because for some reason the \nat does not
 %compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
@@
 and GWR and components of the decomposition
-are plotted against $\eta_2/\eta_1$. The contribution of ATAs ($\Delta_i^{[EC]}$) frequently 
+are plotted against the upper bound ratio, $\eta_2/\eta_1$. The contribution of ATAs ($\Delta_i^{[EC]}$) frequently 
 has magnitude comparable to other components, indicating that ATAs can contribute meaningfully
@@
 This plot uses $q_{ij}=1$. 
-%***DAN: Right, Jasmin?
 An analogous plot (SI Fig. \ref{SI-fig:fig4_alt}) uses 
 $q_{ij}=\eta_i/(\eta_j(1-\delta)+\eta_i \delta)$; 
-%***DAN: Jasmin, I believe the code does not produce the alternate-qij plot yet, but eventually it needs to.
 though results differed in detail, main conclusions were the same.
@@
 \end{figure}
+%***DAN: Jasmin, please note that the standard error number that currently comes up is about 0.27, which
+%is actually really big compared to the plotted values. Is this a temporary value based on a limited set 
+%of sims, and you're planing to do a larger number? If this is the actual value based on the final number 
+%of sims that you plan to do, then we may have to figure out a way to plot standard errors. Let's discuss
+%The same problem occurs for the accompany SI figure.
 

### 2023-05-07 paper/Paper.Rnw
moved a subsection and a figure that goes with it to sup mat to save length and complexity

@@
 and with some but not all aspects of the relationship 
-between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
+between the components altered relative to the original distributions. Specifically, asymmetries of 
+tail association are removed -- see SI section \ref{SI-sect:remove_asymmetries} for how this is done. We then define 
 $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
@@
 below for additional details.
-
-\subsection{Removing asymmetries of association}
-
-How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
-We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
-$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
-Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
-$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
-distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
-Let $\rho_i$ be the covariance of
-these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
-$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
-$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
-The value of $\ddot{\rho}_i$ is then determined in one of 
-four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
-$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
-$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
-we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
-The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
-$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
-$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
 
@@
 
-%***DAN: This looks pretty good except one of the histograms on c does not look like the plotted
-%pdfs of the marginals. I think Jasmin said in an email she was fixing this.
-%***DAN: Jasmin, also, we use \varphi in the text, but \phi here. These need to be made consistent, since they are different to
-%a math person. The easiest way will be to change it in the figure, so I hope that can straightforwardly be done.
-\begin{figure}
-\includegraphics[width=1.025\textwidth]{../results_figs/theory_fig_col.pdf}
-\caption{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
-between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
-(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
-marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
-$\varphi^{-1}$, and 
-each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
-(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
-selected to match that of the original distribution (see Theory). (d) Two new variables, 
-$E_i^{||}$ and  $C_{i \bs i}^{||}$,
-%Dan replaced \nat in the above two expressions by || because for some reason the \nat does not
-%compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
-which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
-which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
-$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
-For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
-$\log_{10}$ of probability density on all panels. 
-\label{fig:pedagog2}}
-\end{figure}
-
 %***DAN: Jasmin, this figure looks really great. Here are a few minor things I would like to ask you to

### 2023-05-07 paper/SupMat.Rnw
cleaning up a bit is all
Dan finished going through the paper, about to try cutting 550 words
moved a subsection and a figure that goes with it to sup mat to save length and complexity

@@
 
+\section{Removing asymmetries of association}\label{sect:remove_asymmetries}
+
+How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
+correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
+We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
+Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
+distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
+Let $\rho_i$ be the covariance of
+these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
+$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
+$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
+The value of $\ddot{\rho}_i$ is then determined in one of 
+four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
+$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
+For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
+we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
+The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
+$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
+as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
+
 \section{Mechanisms other than storage effects}\label{SIsect:other_mechanisms}
@@
 
-%***DAN: I think I may need the below later, in the computation section, but not here
-%For speed of computation, when samples $(b_1^{(i)},b_2^{(i)})$, $i=1,...,M$ were needed for distinct 
-%sets of values of the parameters $\mu_1$, $\mu_2$, and $\sigma$, we used the same 
-%$\left(\tilde b_1^{(i)},\tilde b_2^{(i)}\right)$, $i=1,...,M$, simply transforming these values 
-%differently (using the different valus of $\mu_1$, $\mu_2$, $\sigma$) for each set of parameter values.
-
 When needed, noise with symmetric tail association (Fig. \ref{MT-fig:pedagog1}b) was generated by taking a 
@@
 section \ref{SIsect:scaling_factors}.
-
-%***DAN: We might need this text somewhere, but not sure where. Just saving it out of caution.
-% For the lottery model, recall that $E_i=B_i$ and $C_{i \bs i} = B_j/\delta$. So 
-% $\varphi^{-1} \circ F_{E_i}(E_i) = \frac{b_i-\mu_i}{\sigma}$ and 
-% $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})=\frac{b_j - \mu_j}{\sigma}$.  
-% Also, $\rho_i = \cov(\frac{b_i-\mu_i}{\sigma},\frac{b_j-\mu_j}{\sigma})$, which equals the 
-% Pearson correlation of $b_i$ and $b_j$. For the lottery model, it seems natural to make the choice
-% $\ddot{\rho}_i=\rho_i$, which straightforwardly implies that $(E_i^\nat,C_{i \bs i}^\nat)$
-% can be written as $(\exp(\ddot{b}_i),\exp(\ddot{b}_j))$, where $(\ddot{b}_i,\ddot{b}_j)$
-% is bivariate-normally distributed with mean $(\mu_i,\mu_j)$, $\var(\ddot{b}_i)=\var(\ddot{b}_j)=\sigma$,
-% and $\cov(\ddot{b}_i,\ddot{b}_j)=\cov(b_i,b_j)$. So $(\ddot{b}_i,\ddot{b}_j)$ has the same
-% marginal distributions and covariance matrix as $(b_i,b_j)$, but it is bivariate normal, whereas
-% $(b_i,b_j)$ itself was not required to be. Given these definitions, it is straightforward to compute the 
-% component $\olrn_{i \bs i}$ of main text equation (\ref{MT-eq:storage_effects_further1}) for the lottery model using the simulation 
-% approach of \cite{Ellner_2016,Ellner_2019}. Note that $E_j$ and $C_{j \bs i}$ are linearly related for the lottery model,
-% so $E_j^\nat=E_j$, $C_{j \bs i}^\nat = C_{j \bs i}$, and $\olrn_{j \bs i}=\olr_{j \bs i}$.
 
@@
 
-%***DAN: I think I do not need this section anymore, but just keeping it around to be safe.
-% \section{      Fig. \ref{MT-fig:fig4} does not depend on the value of $q_{ij}$ used}\label{sect:F4andqij}
-% 
-% In Fig. \ref{MT-fig:fig4} of the main text we plotted GWR , $\Delta_i^{[EC]}$, and $\text{GWR}-\Delta_i^{[EC]}$ 
-% (see equation \ref{eq:decomp_overall_delta} and the text below it) for the lottery model. We here show
-% the resulting plot does not depend on which of the two choices for the scaling factors $q_{ij}$ is used.
-% But $\text{GWR} = \olr_{i \bs i} - q_{ij} \olr_{j \bs i}$ and $\Delta_i^{[EC]}=\olep_i^{[EC]}-q_{ij}\olep_j^{[EC]}$,
-% so it suffices to show $\olr_{j \bs i}=0$ and $\olep_j^{[EC]}=0$. The first of these is known true, \emph{a priori}, 
-% and the second was demonstrated in section \ref{sect:math_simplifications}.
-
 \section{    Three-way decompositions}\label{sect:threeway}
@@
 
-%***DAN: Jasmin, please fix the collision of text that occurs in the text near the legend
+%***DAN: This looks pretty good except one of the histograms on c does not look like the plotted
+%pdfs of the marginals. I think Jasmin said in an email she was fixing this.
+%***DAN: Jasmin, also, we use \varphi in the text, but \phi here. These need to be made consistent, since they are different to
+%a math person. The easiest way will be to change it in the figure, so I hope that can straightforwardly be done.
+\begin{figure}
+\includegraphics[width=1.025\textwidth]{../results_figs/theory_fig_col.pdf}
+\caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
+between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
+(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
+marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
+$\varphi^{-1}$, and 
+each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
+(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
+selected to match that of the original distribution (see Theory). (d) Two new variables, 
+$E_i^{||}$ and  $C_{i \bs i}^{||}$,
+%Dan replaced \nat in the above two expressions by || because for some reason the \nat does not
+%compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
+which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
+which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
+$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
+For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
+$\log_{10}$ of probability density on all panels. 
+\label{fig:pedagog2}}
+\end{figure}
+
+\clearpage
+\newpage
+
 \begin{figure}
@@
 were never greater than \Sexpr{round(se4qij,5)}, so are not displayed.}\label{fig:fig4_alt}
-%***DAN: jasmin, once the se is generated and saved by your code, the chunk at the outset of this 
-%document needs to be modified to load that quantity, and then the 0 in the above expression 
-%needs to be replaced by the variable name used.
-%***JAS:done
 \end{figure}
+%***DAN: Jasmin, the value of se4 quoted in the main text was the same, to 5 digits, as the 
+%value of se4qij quoted here. This made me suspicious so I checked the values to 9 digits
+%and they were the same to 9 digits. This seems like it must be an error in the code. Please
+%check. The standard error expressions in SI S5.2 DO depend on qij, so this seems like it
+%should not happen. The standard errors listed in Fig. 3 and Fig. S1 are different, so that's
+%good.
 

### 2023-05-07 paper/Paper.Rnw
moved some more stuff to the SI for length reasons

@@
 \noindent \textbf{Type of article:} Letter\\
-\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 5518; \textbf{Words in text boxes:} 0\\
-\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
+\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 5013; \textbf{Words in text boxes:} 0\\
+\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 4; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
 \noindent \textbf{Corresponding author:}
@@
 
-%E and vital rates
-Whereas the ``$E$'' of ``$EC$-covariance'' is often conceptualized as ``environment'', 
-in fact it is often instead a 
-species response to the environment. Because a directly measurable aspect of the 
-environment and a species response to it may be nonlinearly related, it may be 
-important, in future work, 
-to also study the nature of tail associations between species vital rates.
-In the lottery model, it was assumed that fecundity of species $i$ tracks the
-environment directly, $B_i(t)=E_i(t)$. But vital rates such as fecundity are 
-more likely related to commonly measured environmental variables according to a nonlinear
-function which may often be threshold-like or sigmoidal, and thus the $E_i=B_i$
-of the lottery model is probably better conceptualized as a species response.
-Likewise, one of the choices of $E_i$ considered by \cite{Ellner_2016} for the
-diatom model was $E_i=V_i$, the maximum population growth rate, a response
-to the environment rather than a direct environmental measurement. 
-It was demonstrated elsewhere that nonlinear transformations can produce or alter 
-asymmetries of tail association \citep{Ghosh_2020_AER,Walter_2022}.
-So the nature of ATAs between directly measured environmental variables may differ from the
-nature of ATAs between the population responses which matter for coexistence.
-Vital rates are measured directly in demographic studies, so ATAs between vital rates could,
-in principle, be studied. To our knowledge, this has not been done.
-
 %Differences with the original, analytic version of the MCT
@@
 It may be possible to expand the mathematics of the original MCT to consider ATAs.
-
-%multispecies models and multifactor models
-\cite{Ellner_2019} elaborated a general theory in which more and different varying factors than
-$E$ and $C$ can be considered and used to produce a decomposition of GWR. For instance, if $X$, $Y$ and 
-$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, a decomposition
-of the GWR of $i$ for which some terms were based on rendering one or 
-more of $X$, $Y$, and $Z$ independent of the others. 
-We simply comment that this general decomposition should also be extendable to consider the effects of
-ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}). See SI section \ref{SI-sect:threeway} for details.
+Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:threeway}.
 
@@
 
+%***DAN: Jasmin, this currently pulls in a pdf you sent me, stored temporarily on my drive in the indicated 
+%location. Once the code produces fig 5 on a normal run, we'll have to change the below to suck that
+%version of the figure up instead of this one.
 \begin{figure}
-%\includegraphics[width=\textwidth]{../results_figs/fig5.pdf}
-\caption{ATA effects and ATA rescue in the diatom system. 
+\includegraphics[width=\textwidth]{fig5_new.pdf}
+\caption{See next page for caption.}  \label{fig:fig5}
+\end{figure}
+
+\clearpage
+\newpage
+
+\singlespacing
+
+\noindent Figure 4. ATA effects and ATA rescue in the diatom system. 
 The contribution of ATAs to the GWR of \emph{F. crotonensis}
@@
 sufficient coexistence metric.
-}  \label{fig:fig5}
-\end{figure}
 

### 2023-05-07 paper/SupMat.Rnw
moved some more stuff to the SI for length reasons

@@
 
+\section{   Asymmetric associations between population vital rates}\label{sect:vital_rates}
+
+Whereas the ``$E$'' of ``$EC$-covariance'' is often conceptualized as ``environment'', 
+in fact it is often instead a 
+species response to the environment. Because a directly measurable aspect of the 
+environment and a species response to it may be nonlinearly related, it may be 
+important, in future work, 
+to also study the nature of tail associations between species vital rates.
+In the lottery model, it was assumed that fecundity of species $i$ tracks the
+environment directly, $B_i(t)=E_i(t)$. But vital rates such as fecundity are 
+more likely related to commonly measured environmental variables according to a nonlinear
+function which may often be threshold-like or sigmoidal, and thus the $E_i=B_i$
+of the lottery model is probably better conceptualized as a species response.
+Likewise, one of the choices of $E_i$ considered by \cite{Ellner_2016} for the
+diatom model was $E_i=V_i$, the maximum population growth rate, a response
+to the environment rather than a direct environmental measurement. 
+It was demonstrated elsewhere that nonlinear transformations can produce or alter 
+asymmetries of tail association \citep{Ghosh_2020_AER,Walter_2022}.
+So the nature of ATAs between directly measured environmental variables may differ from the
+nature of ATAs between the population responses which matter for coexistence.
+Vital rates are measured directly in demographic studies, so ATAs between vital rates could,
+in principle, be studied. To our knowledge, this has not been done.
+
 \section{    Three-way decompositions}\label{sect:threeway}
+
+\cite{Ellner_2019} elaborated a general theory in which more and different varying factors than
+$E$ and $C$ can be considered and used to produce a decomposition of GWR. For instance, if $X$, $Y$ and 
+$Z$ are temporally varying factors, \cite{Ellner_2019} produced, for the $i^{\text{th}}$ species, a decomposition
+of the GWR of $i$ for which some terms were based on rendering one or 
+more of $X$, $Y$, and $Z$ independent of the others. 
+We comment that this general decomposition should also be extendable to consider the effects of
+ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}). Details follow.
 

### 2023-05-15 code/Figure3.R
edits to label and axis ticks

@@ -78,12 +78,12 @@ for (m in seq_along(mudif)){ #iterates across mudif values
   	
   	#plot
   	plotco(file,sigma,ylim=ylims[[m]])
+  	#axes
   	if (d==1){
-  		graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
-  	}
+  		graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)} else {graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2, labels=F)}
   	if (n>9){
-  		graphics::axis(1, cex.axis=1.8,tck=-0.028, lwd.ticks=2)
-  	}
+  		graphics::axis(1, cex.axis=1.8,tck=-0.028, lwd.ticks=2)} else {graphics::axis(1, cex.axis=1.8, tck=-0.028, lwd.ticks=2, labels=F)}
+  	#panel labels
     graphics::mtext(paste0("(", letters[n],")"), side=3, line=-1.7, at=0, cex=1.3, adj=0)
     
     if (n==2){
@@ -101,13 +101,14 @@ for (m in seq_along(mudif)){ #iterates across mudif values
     	graphics::text(expression(mu[1]-mu[2]), x=9,y=-3.95, font=2, cex=3, srt=-90, adj=1, col="gray40")
     }
     if (n==11){
-    	graphics::mtext("log fecundity standard deviation, ",side=1, line=2.8, cex=1.25, at=-3.25, adj=0,col="gray40")
-    	graphics::mtext(expression(sigma),side=1, line=2.8, cex=1.5, at=10.15, adj=1,col="gray40")
+    	graphics::mtext("log fecundity standard deviation, ",side=1, line=2.8, cex=1.25, at=-3.85, adj=0,col="gray40", font=2)
+    	graphics::mtext(expression(sigma),side=1, line=2.8, cex=1.5, at=10.75, adj=1,col="gray40", font=2)
     }
     n <- n+1
   }
-} 
+}
 graphics::mtext("contribution to coexistence", side=2, outer=TRUE, line=2, font=2, cex=1.7, col="gray40")
+#legend
 graphics::legend("topright", legend=terms, col = cols, lty = ltys, bty="n", cex=2, inset=c(-0.83,-2.75), y.intersp = 1.35, x.intersp = 0.1, seg.len=0.8, lwd=lwds)
[…]

### 2023-05-15 code/Figure3.R
copied axis and tick edits to qij version of fig

@@ -150,12 +150,12 @@ for (m in seq_along(mudif)){ #iterates across mudif values
   	
   	#plot
   	plotco(file,sigma,ylim=ylims[[m]])
+  	#axes
   	if (d==1){
-  		graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
-  	}
+  		graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)} else {graphics::axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2, labels=F)}
   	if (n>9){
-  		graphics::axis(1, cex.axis=1.8,tck=-0.028, lwd.ticks=2)
-  	}
+  		graphics::axis(1, cex.axis=1.8,tck=-0.028, lwd.ticks=2)} else {graphics::axis(1, cex.axis=1.8, tck=-0.028, lwd.ticks=2, labels=F)}
+	#panel labels
     graphics::mtext(paste0("(", letters[n],")"), side=3, line=-1.7, at=0, cex=1.3, adj=0)
     
     if (n==2){

### 2023-05-15 code/LBdecompFig.R
added labels and ticks

@@ -111,10 +111,13 @@ for (d in seq_along(delta)){
 	DeltasLT <- plotco(etai,etaj,delta[d], B,Deltas_loc=file, SE_loc=sefile)
 	maxseLT[d] <- max(DeltasLT$SE,na.rm=T)
 	mtext(paste(delta[d]), side=3, line=0.2, font=2, cex=1.5, col="gray30")
-	#axis(1, cex.axis=2,tck=-0.028, lwd.ticks=2)
+	#axis
+	axis(1, cex.axis=1.8,tck=-0.028, lwd.ticks=2, labels=F)
 	if (d==1){
 		axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
-	}
+	} else {axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2, labels=F)}
+	#panel labels
+    graphics::mtext(paste0("(", letters[d],")"), side=3, line=-1.7, at=4.6, cex=1.3, adj=0)
 	if (d==2){
 		graphics::mtext("adult death rate, ", line=2.5, font=2, cex=1.5, col="gray40", at=1.25, adj=0)
     	graphics::mtext(expression(delta), line=2.5, font=2, cex=2.3, col="gray40", side=3, at=5, adj=1)
@@ -128,7 +131,6 @@ for (d in seq_along(delta)){
 	if (file.exists(sefile)==FALSE){saveRDS(DeltasLT$SE,file=sefile)}
 }
 
-
 #d-f (right-tail)
 maxseRT <- {}
 for (d in 1:length(delta)){
@@ -140,6 +142,8 @@ for (d in 1:length(delta)){
 	if (d==1){
 		axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
 	}
+	#panel labels
+    graphics::mtext(paste0("(", letters[d+3],")"), side=3, line=-1.7, at=4.6, cex=1.3, adj=0)
 	if (d==2){
 		title(xlab="upper bound ratio", line=2.75, font.lab=2, cex.lab=2, col.lab="gray40")
 	}	
@@ -150,7 +154,6 @@ for (d in 1:length(delta)){
 	if (file.exists(file)==FALSE){saveRDS(DeltasRT$D,file=file)}
 	if (file.exists(sefile)==FALSE){saveRDS(DeltasRT$SE,file=sefile)}
 }
-
 
 #7
 rescol <- rgb(227/255, 211/255, 148/255,.5)

### 2023-05-15 code/LBdecompfig_qij.R
axis and lbls

@@ -46,10 +46,12 @@ for (d in seq_along(delta)){
 	DeltasLT <- plotco(etai,etaj,delta[d], B,Deltas_loc=file, SE_loc=sefile, qij=T)
 	maxseLT[d] <- max(DeltasLT$SE,na.rm=T)
 	mtext(paste(delta[d]), side=3, line=0.2, font=2, cex=1.5, col="gray30")
-	#axis(1, cex.axis=2,tck=-0.028, lwd.ticks=2)
+	axis(1, cex.axis=2,tck=-0.028, lwd.ticks=2, labels=F)
 	if (d==1){
 		axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
-	}
+	} else {axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2, labels=F)}
+	#panel labels
+    graphics::mtext(paste0("(", letters[d],")"), side=3, line=-1.7, at=4.6, cex=1.3, adj=0)
 	if (d==2){
 		graphics::mtext("adult death rate, ", line=2.5, font=2, cex=1.5, col="gray40", at=1.25, adj=0)
     	graphics::mtext(expression(delta), line=2.5, font=2, cex=2.3, col="gray40", side=3, at=5, adj=1)
@@ -75,6 +77,8 @@ for (d in 1:length(delta)){
 	if (d==1){
 		axis(2, cex.axis=1.8, tck=-0.035, lwd.ticks=2)
 	}
+	#panel labels
+    graphics::mtext(paste0("(", letters[d+3],")"), side=3, line=-1.7, at=4.6, cex=1.3, adj=0)
 	if (d==2){
 		title(xlab="upper bound ratio", line=2.75, font.lab=2, cex.lab=2, col.lab="gray40")
 	}	

### 2023-05-15 paper/Paper.Rnw
deleted some old commented material
fixed a typo
Jasmin addressing cmments

@@
 \usepackage{float}
-\usepackage{setspace}
+\usepackage{setspace}ƒ
 \usepackage{lineno}
@@
 \affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
-\affil[b]{Environmental Studies Program and Department of Biology, Universit of Oregon}
+\affil[b]{Environmental Studies Program and Department of Biology, University of Oregon}
 \affil[c]{Laboratory of Populations, Rockefeller University}
@@
 The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
-Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, Laurn Hallett, and 
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, and 
 Robin Snyder for helpful discussions. 
@@
 %something.
+%***JASMIN: Done. Added ticks to other panels and make x label bold
 \begin{figure}
@@
 %Also please add tick marks to all axes (see comment on previous figure)
+%***JASMIN: Done. Labels and ticks.
 \begin{figure}
@@
 
-% \begin{figure}
-% \includegraphics[width=.38\textwidth]{../results_figs/fig6.pdf}
-% \caption{The contribution of ATAs to GWR of \emph{F. crotonensis}
-% in the diatom system of \cite{Descamps_2005} was often meaningfully large compared to 
-% the GWR itself. The ratio $\Delta_i^{[EC]}/\text{GWR}$ is shown, larger magnitudes
-% indicating greater importance of ATAs and values greater than $1$ corresponding to either
-% ATA rescue or ATA exclusion (Methods). For this system and the plotted parameters, 
-% $\Delta_i^{[EC]}$ was always positive, so cyan indicates exclusion of \emph{F. crotonensis} and
-% magenta indicates coexistence of the two diatom species. Values greater than 1 or less than -1
-% were remapped to 1 and -1, respectively, for readability.
-% For the parameters we considered, GWR of the other species, \emph{C. pseudostelligera}, was always positive both including and
-% excluding ATAs, so GWR for \emph{F. crotonensis} is a sufficient coexistence metric.
-% Standard errors of plotted quantities were always less then \Sexpr{round(se6,5)}, so are not shown.} \label{fig:fig6}
-% \end{figure}
-
 \end{document}

### 2023-05-24 code/theoryFig_script.R
code edits for figS1

@@ -1,5 +1,4 @@
-source("./theory_panel_fxn.R")
-library(latex2exp)
+#library(latex2exp)
 
 nx<-seq(from=-4,to=4,by=0.02)
 ny<-seq(from=-4,to=4,by=0.02)

### 2023-05-24 code/theory_panel_fxn.R

@@ -59,12 +59,12 @@ tdis <- function(q, oldp, samps){
 }
 
 pdfhist <- function(samps,x,y,d, horiz=FALSE,xpd=FALSE){
+	if (sum(x==y)!=length(x)){stop("x!=y")}
+	
 	nbrks <- 15
-	if (horiz==TRUE){
-		breaks <- seq(min(y), max(y), length.out=nbrks)
-	} else {
-		breaks <- seq(min(x), max(x), length.out=nbrks)
-	}
+	breaks <- seq(min(x), max(x), length.out=nbrks)
+	
+	if (min(samps) < min(x)){samps[samps==min(samps)] <- min(x)}
 	h <- hist(samps, breaks, plot=FALSE)
 	
 	ylines <- d(x)*length(samps)* diff(h$breaks)[1]

### 2023-05-26 paper/Paper.Rnw
adderssde Dan comment
last minute tweaks

@@
 \usepackage{float}
-\usepackage{setspace}ƒ
+\usepackage{setspace}
 \usepackage{lineno}
@@
 \usepackage{tikz}
+\usepackage{Sweave}
 
@@
 se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
-se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
-se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
+#se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
+#se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
 @
@@
 
-%***DAN: Jasmin, this figure looks really great. Here are a few minor things I would like to ask you to
-%do to further optimize:
-%1) Currently there are axis labels and numbers only on the bottom and left panels, which makes sense.
-%But the axis tick marks needs to occur on all panels. I noticed that this would be an imporvement
-%when I found myself wanting to see where the ATA line was between the x axis and the GWR line, but it was 
-%a little hard to see where the x axis was becuase there were no tick marks. It is possible to have tick marks
-%but no numbers of labels in R, and that's what we need here.
-%2) The font of the x axis label, "log fecundity standard deviation, \sigma" seems slightly different from the 
-%other textual labels. Can you please make it consistent? Perhaps it is not bold and the others are bold, or 
-%something.
-%***JASMIN: Done. Added ticks to other panels and make x label bold
 \begin{figure}
@@
 
-%***DAN: Jasmin, panel labels(a, b, c, etc.) need to be added. Please use the same format for these labels
-%as you do in the other figures.
-%Also please add tick marks to all axes (see comment on previous figure)
-%***JASMIN: Done. Labels and ticks.
 \begin{figure}
@@
 though results differed in detail, main conclusions were the same.
-%***DAN: make sure this is indeed the case once the plot is generated.
 }\label{fig:fig4}
 \end{figure}
-%***DAN: Jasmin, please note that the standard error number that currently comes up is about 0.27, which
-%is actually really big compared to the plotted values. Is this a temporary value based on a limited set 
-%of sims, and you're planing to do a larger number? If this is the actual value based on the final number 
-%of sims that you plan to do, then we may have to figure out a way to plot standard errors. Let's discuss
-%The same problem occurs for the accompany SI figure.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{fig5_new.pdf}
+\includegraphics[width=\textwidth]{../results_figs/fig5_new.pdf}
 \caption{See next page for caption.}  \label{fig:fig5}
@@
 and the horizontal axis, indicating that, for those parameter values, ATA rescue occurs. 
-Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
+Standard errors of plotted quantities were always less then \Sexpr{round(1,5)}, so are not shown.
 Panels d-f show bivariate plots (Methods) for which the third parameter was 

### 2023-05-26 paper/SupMat.Rnw
addressing Dan comment
addressing Dan comments
last minute tweaks

@@
 We comment that this general decomposition should also be extendable to consider the effects of
-ATAs in the same way (\ref{eq:full_decomp}) extends (\ref{eq:previous_decomp}). Details follow.
+ATAs in the same way (\ref{MT-eq:full_decomp}) extends (\ref{MT-eq:previous_decomp}). Details follow.
 
@@
 %pdfs of the marginals. I think Jasmin said in an email she was fixing this.
+%***JASMIN: fixed!!
 %***DAN: Jasmin, also, we use \varphi in the text, but \phi here. These need to be made consistent, since they are different to
 %a math person. The easiest way will be to change it in the figure, so I hope that can straightforwardly be done.
+%***JASMIN: we said this is OK. Operating system thing - renders the same in Windows. 
 \begin{figure}
@@
 %good.
+%JASMIN: I think its okay that they are the same.
 

### 2023-05-26 paper/Paper.Rnw
trivial

@@
 se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
-#se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
+se5 <- 1 #readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
 #se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
@@
 and the horizontal axis, indicating that, for those parameter values, ATA rescue occurs. 
-Standard errors of plotted quantities were always less then \Sexpr{round(1,5)}, so are not shown.
+Standard errors of plotted quantities were always less then \Sexpr{round(se5,5)}, so are not shown.
 Panels d-f show bivariate plots (Methods) for which the third parameter was 

### 2023-06-29 paper/Paper.Rnw
trivial

@@
 se4 <- readRDS(file=paste0(res_loc,"fig4maxse.RDS"))
-se5 <- 1 #readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
+se5 <- readRDS(file=paste0(res_loc, "fig5maxse.RDS"))
 #se6 <- readRDS(file=paste0(res_loc, "fig6maxse.RDS"))
@@
 \noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 5013; \textbf{Words in text boxes:} 0\\
-\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 4; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
+\noindent \textbf{Number of references:} 37; \textbf{Number of figures:} 4; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
 \noindent \textbf{Corresponding author:}

### 2023-06-29 paper/RefereeResponses1.tex
added document for referee responses

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{Sweave}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\DeclareMathOperator{\E}{\mathbb{E}}% expected value
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\cor}{cor}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\prob}{P}

%attempt 1 at nat and sharp
%\newcommand{\nat}{\mathlarger{\natural}}
%\newcommand{\shp}{\mathlarger{\sharp}}

%attempt 2 at nat and sharp
%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
%\newcommand{\shp}{\#}

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\bs}{\backslash}

%header material for paper
\title{Responses to referees for ``Asymmetric relationships and their effects on coexistence''}
\date{July 2023}

\begin{document}

\end{document}

### 2023-06-29 paper/RefereeResponses1.tex
added a template doc within which we can reply to referee comments

@@
 \usepackage{tikz}
-\usepackage{Sweave}
+\usepackage{hyperref}
 
@@
 
+\noindent \emph{Dear Editors,} \\
+
+\noindent \emph{Thank you for the useful and very positive reviews of our manuscript, and for
+the opportunity to resubmit a revised version. We have addressed all referee comments below and in
+the manuscript, and hope the paper can now be rapidly accepted. } \\
+
+\noindent \emph{We have replied to referee comments below in italic, with each response preceded
+by ``***'', to make our responses easier to find.} \\
+
+\noindent \emph{We submit the following pdf files: ...} \\
+
+\noindent \emph{We additionally submit the following latex files, necessary for compiling the
+main text latex: ...} \\
+
+\noindent \emph{The document showing the changes we made does not include links such as citations and
+references to figures, but it does accurately represent the changes made to the text. 
+If all latex files are deposited into the same folder, it should then be possible to 
+compile the main text latex from that folder.} \\
+
+\noindent \emph{We again thank the referees and editors for their feedback and positive reviews!} \\
+
+\noindent \emph{Yours,}
+
+\noindent \emph{Dan Reuman and Jasmin Albert} \\
+
+\noindent \textbf{Comments of referee 1:} \\
+
+%The C) that precedes each comment is because we will later add numbers, C1, C2, etc., after making sure these
+%comments are divided up to our liking. Then we can refer to each comment by its number, which can be useful.
+\noindent C) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.
+Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated.
+This is one among many ways that one could characterize or decompose the full bivariate distribution between competition and environment, but the authors argue and demonstrate that ATAs of this kind are (a) prevalent in natural systems and (b) can have a meaningful and sometimes qualitative effect on the outcome of coexistence between two species.
+The combination of results for theoretical distirbutions but also for experimental data is very powerful and convincing.
+\\
+
+\noindent C) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience.
+\\
+
+\noindent C) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.
+
+My first comment related to this is that some form of this text (whcih is relatively short---around half a page in the SI) and, importantly, the Figure should be in the main text.
+I think this is such a key part of the definitions that are being used for the rest of the paper that readers will want to get an intuition for what the "partial sharp" distributions look like and how they are arrived at.
+\\
+
+\noindent C) Perhaps equally important to clarifying this process for the readership is whether this process is unique.
+I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables.
+Are there alternatives, or is the authors' procedure unique in achieving the desired outcome?
+If the former, then I would want to understand if the outcomes downstream tend to be robust to the specific method used.
+If the latter, it would be helpful to have an explanation as to why this is a unique decomposition.
+\\
+
+\noindent C) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
+$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.
+Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.
+\\
+
+\noindent C) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).
+These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species.
+Is this really a demonstration that ATAs occur for C and E?
+If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).
+Maybe I am missing something.
+\\
+
+\noindent C) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.
+I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned.
+Perhaps that could be added for completeness.
+\\
+
+\noindent C) E.g.
+line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense.
+I'd consider going through the paper carefully for any typos or small corrections of this kind.
+\\
+
+\noindent \textbf{Comments of referee 2:} \\
+
+\noindent C) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se.
+The authors explain well why they do this and why they think this may be important.
+Additionally, they show the best possible efforts to convince the readers why such a decomposition may be important and what we may learn from it.
+Overall, the paper is interesting and quite understandable.
+I think the authors have done a very good job.
+\\
+
+\noindent C) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach.
+I want to be clear that this is, in part, my personal opinion and I know experts in the field which have a very different view.
+What the authors have shown (and it appears to be correct) is that in a model where niche differentiation via different resources and predators is impossible, then ATA can be an important driver of coexistence, and I agree.
+However, in every application where both storage effect, relative non-linearity and niche differentiation have been measured, there we see that storage effect and relative non-linearity where very small compared to simple resource differentiation.
+But, one might argue, that we have barely any such empirical applications, and I would agree, which is why I label this my personal opinion.
+(The few papers that I know of are \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1002%2Fecy.2726&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=HbUTjzCNQrvGSARY5u3Eqme88DbKiNyiAVNOUGzjcbE%3D&reserved=0}{Zepeda et al.
+2019, Ecology}, \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1890%2F14-1741.1&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=KKIBuLjBN2vNGWfEpduAGGrMAt7ov9a4F53CmlZQlO0%3D&reserved=0}{Chu and Adler 2015}, and Ellner et al.
+2019, see also the new paper by \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fesajournals.onlinelibrary.wiley.com%2Fdoi%2Fepdf%2F10.1002%2Fecm.1585&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=YaqsZ6axu6rYV12kh9IiRIbES7ZUMWIN7HXgB08iPzc%3D&reserved=0}{Stump and Vasseur}, which I have not yet read fully) In my understanding, ATA is a decomposition of storage effect into even more components, therefore I would expect, on average the ATA effect to be slightly smaller than storage effect (but, as you have shown can also be larger than storage effect).
+I would therefore assume that ATA is generally weak compared to niche differentiation (but again, admittedly based on few datapoints).
+Obviously, the authors may disagree with my personal view, in this case I’d like to see a discussion of this.
+\\
+
+\noindent C) I like the conceptual idea of removing the long tail. \\
+
+\noindent C) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does.
+I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed.
+Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition.
+For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct?
+\\
+
+\noindent C) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur.
+But on a broader level you essentially decompose the EC distribution into it’s different momentums.
+One only carrying about the mean (i.e.
+first order correct), one with only carrying about the mean and covariance (i.e.
+getting the first two orders correct), and one with all other orders correct.
+Given this interpretation of your work, I might apply the same logic to the relative-nonlinearity, where I decompose the relative non-linearity into the variance term and the term with long tails.
+(one might also even decompose the actual distribution into more distributions, getting successively more and more momentums of the distribution correct, but I would argue that this is probably less useful).
+\\
+
+\noindent C) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\
+
+\noindent C) L41: The “we” is slightly confusing because the author teams share only one author \\
+
+\noindent C) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\
+
+\noindent C) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\
+
+\noindent C) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\
+
+\noindent C) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters.
+\\
+
+\noindent C) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
+
+\noindent C) L74: The optimum might actually be the same, but they need different responses Eq.
+1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\
+
+\noindent C) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\
+
+\noindent C) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\
+
+\noindent C) L126: Do you do this in the appendix? \\
+
+\noindent C) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\
+
+\noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\
+
+\noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\
+
+\noindent C) L316: So why do you show values of $\sigma >6$?
+And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA.
+In 1d it simply appears that the variance scales with the density.
+1e is quite unconvincing to me.
+That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example?
+\\
+
+\noindent C) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\
+
+\noindent C) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case.
+\\
+
+\noindent C) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\
+
+\noindent \textbf{Comments of referee 3:} \\
+
+\noindent C) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked.
+ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect.
+To illustrate these ideas the authors combine results from mathematical models, simulations, and empirical examples, and show that under a certain combination of parameters ATAs can be a significant mechanism of species coexistence.
+\\
+
+\noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\
+
+\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations.
+Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics.
+I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript.
+\\
+
+\noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
+\\
+
+\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect.
+In detail, the explanation of the storage effect is superb, but how ATAs enter in play is not well solved.
+It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented.
+\\
+
+\noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor.
+Could the authors briefly indicate how this could occur?
+\\
+
+\noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\
+
+\noindent C) The model used for the chemostat is in its formulation quite different compared to the lottery model.
+Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model.
+\\
+
+\noindent C) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
+this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones.
+\\
+
+\noindent \textbf{Comments of the editor:} \\
+
+\noindent C) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\
+
+\noindent C) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists.
+My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing.
+The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems.
+Some mathematical material could be moved to Supplementary material.
+\\
+
 \end{document}

### 2023-06-29 paper/RefereeResponses1.tex
started brainstorming responses and actions

@@
 %comments are divided up to our liking. Then we can refer to each comment by its number, which can be useful.
-\noindent C) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.
-Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated.
+\noindent C1) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.\\
+
+\noindent ***\emph{Thank you for the positive feedback!} \\
+
+\noindent C2) Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated.
 This is one among many ways that one could characterize or decompose the full bivariate distribution between competition and environment, but the authors argue and demonstrate that ATAs of this kind are (a) prevalent in natural systems and (b) can have a meaningful and sometimes qualitative effect on the outcome of coexistence between two species.
@@
 
-\noindent C) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience.
+\noindent ***\emph{Thank you for the positive feedback!} \\
+
+\noindent C3) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience.
 \\
 
-\noindent C) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.
+\noindent ***\emph{response} \\
+
+\noindent C4) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.
 
@@
 
-\noindent C) Perhaps equally important to clarifying this process for the readership is whether this process is unique.
+\noindent ***\emph{<Move SI Fig S1 and associated text back to the MT – at least two referees asked for it.>} \\
+
+\noindent C5) Perhaps equally important to clarifying this process for the readership is whether this process is unique.
 I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables.
@@
 
-\noindent C) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
+\noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\
+
+\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
 $E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.
@@
 \\
+
+\noindent ***\emph{<This one asks for possibly changing notation.
+I don’t think we should do it, but we need to explain in a response what is actually going on, why we chose the notation that we do, and then add a new SM section addressing what the referee is talking about.
+Actually it’s the bivariate distribution that changes when you apply ||, tho the marginals are not changed.
+So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C.
+This is the same issue with the # notation that has already been established.
+It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space.
+So in the end the referee does not really have a point if one wants to be formal.
+But you can explain all this in a new sup mat section.>} \\
 
@@
 
+\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing.
+Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.>} \\
+
 \noindent C) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.
@@
 
+\noindent ***\emph{<Just do this and say we did it - it's a good enough idea.>} \\
+
 \noindent C) E.g.
@@
 \\
+
+\noindent ***\emph{response} \\
 
@@
 \\
+
+\noindent ***\emph{Thank you for the very positive feedback!} \\
 
@@
 
+\noindent ***\emph{<Add a discussion, either in the Discussion or in an SM section called something like 
+``Additional discussion point". Acknowledge the viewpoint, state that insufficient studies have been done to be sure
+. 2) >} \\
+
 \noindent C) I like the conceptual idea of removing the long tail. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) L126: Do you do this in the appendix? \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\
 
+\noindent ***\emph{response} \\
+
 \noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
 \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor.
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \noindent C) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
@@
 
+\noindent ***\emph{response} \\
+
 \noindent \textbf{Comments of the editor:} \\
@@
 \noindent C) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\
+
+\noindent ***\emph{response} \\
 
@@
 
+\noindent ***\emph{response} \\
+
 \end{document}

### 2023-06-30 paper/RefereeResponses1.tex
continuing to insert thoughts from which to build responses and ms edits

@@
 \noindent ***\emph{<Add a discussion, either in the Discussion or in an SM section called something like 
-``Additional discussion point". Acknowledge the viewpoint, state that insufficient studies have been done to be sure
-. 2) >} \\
+``Additional discussion points". Acknowledge the viewpoint, state that insufficient studies have been done to be sure
+. Thank the referee for the openness that this is their own opinion - it's very welcome when a referee qualifies
+what they say in this way which permits consideration of alternative viewpoints rather than just being forced to
+do what referees say because they are in the position of power. Point out that we have not only worked with the lotery
+model, as the referee mentions, we have also worked with an empirical example, and that one definitely supported
+the importance of ATAs. Though we accept that future work may reveal that storage effects, and therefore probably
+also ATAs, are not so important in field systems, we think it is also possible that this will not be 
+revealed to be the case, and so we think it is important to not pre-judge the issue when (as the referee 
+admits) there is insufficient data. Our results add another piece of evidence to the small collection of 
+evidence as to the importance of storage effects, as well as adding the first evidence on the potential 
+importance of ATAs. Also emphasize what the referee already seems to accept - that even if storage effects are 
+small, ATA effects can be larger in magnitude if correltion effects per se counteract ATA effects.
+
+Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.
+For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition.
+However, we did not specify our vital rate regressions in a way that makes the covariation between the environment (or random year effects) and competition explicit.
+As a result, our model may not allow full expression of a temporal storage effect and the simulations may underestimate its strength.
+Such differences in model specification may explain why a previous analysis, based on a very different modeling approach, showed a strong storage effect in the Kansas community (Adler et al.
+2006) in contrast to the weak effect we observed here.
+We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."
+
+Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
+It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
+has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
+to the Ellner approach of quantifying storage effects?
+
+Finally, I notice that sometimes for our results, ATA effects and effects of correlation per se counter
+each other. If that happens, storage effects could be weak, while ATA effects are still strong. 
+>} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Thank you for the positive feedback.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{<Clearly since two referees have now mentioned it we need to put some version of Fig S1 back into the
+main text. The earlier referee wanted to see the whole figure brought back, but this referee wants to see just part, i.e.,
+a simpler version, as specified. It's not clear to me how it would help the reader to just see the three panels the referee
+suggests. Only 2 of the suggested panels are currently part of Fig. S1 - the suggestion of putting in the uncorrelated
+distribution would be new. The current Fig. S1 serves the purpose of describing exactly the process of removing ATAs. 
+The new suggested figure would not serve that purpose, it would instead maybe give the reader a sense for what these
+distributions look like when manipulated in the ways we are talking about. But it's not real clear to me that much
+intuition will be conveyed by doing that. One way to reply would be just to add a panel to the figure that shows 
+the uncorrelated case, and move the resulting 5-panel figure back in. Then we'd explain that 1) the other referee asked
+for the whole figure, 2) it's not clear to us that the simplified version serves any good purpose, it certainly 
+does not serve the purpose of explaining the process of removing ATAs, but 3) we went ahead and added the uncorrelated
+case anyway, since that's what was used previously, so it makes sense to have that point of comparison. If we do
+this, Jasmin can add the new panel to the figure.>} \\
 

### 2023-06-30 paper/RefereeResponses1.tex
continued

@@
 
-\noindent C) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).
+\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).
 These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species.
@@
 
-\noindent C) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.
+\noindent C8) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.
 I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned.
@@
 
-\noindent C) E.g.
+\noindent C9) E.g.
 line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense.
@@
 
-\noindent C) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se.
+\noindent C10) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se.
 The authors explain well why they do this and why they think this may be important.
@@
 
-\noindent C) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach.
+\noindent C11) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach.
 I want to be clear that this is, in part, my personal opinion and I know experts in the field which have a very different view.
@@
 each other. If that happens, storage effects could be weak, while ATA effects are still strong.
+
+I suppose one must always accept the possibility that future work will render one's results less
+meaningful or obsolete.
 >} \\
 
-\noindent C) I like the conceptual idea of removing the long tail. \\
+\noindent C12) I like the conceptual idea of removing the long tail. \\
 
@@
 
-\noindent C) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does.
+\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does.
 I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed.
@@
 
-\noindent C) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur.
+\noindent C14) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur.
 But on a broader level you essentially decompose the EC distribution into it’s different momentums.
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We are not sure we understand precisely what the referee is saying here because their language
+appears more suggestive than formally correct. Perhaps they mean
+``moment'' rather than ``momentum''? Perhaps they mean ``caring'' instead of ``carrying''? I think the phrase 
+``term with long tails'' does not precisely relate to what we're doing since marginal distributions
+of the bivariate distribution $(E,C)$ are unchanged in all our approaches, and hence tails are unchanged. 
+It's just relationships
+between $E$ and $C$ that are modified in the different scenarios we consider. The referee's ideas, here,
+if made precise, may be about potentially using a series of approximations to the bivariate distribution $(E,C)$
+that have the same marginals as $(E,C)$ itself, but that get successively more co-moments right, compared to the
+co-moments of $(E,C)$ itself (and the 
+remaining co-moments would be zero, we presume, if
+such distributions can be defined). Such an approach, applied to relative nonlinearity, would not give
+anything interesting because relative nonlinearity only depends on the marginal distributions, which, 
+under this framework, are always equal to the original marginals of $(E,C)$. On the other hand,
+perhaps the referee means to take a series of approximations of $(E,C)$ so that the $n$th 
+approximate distribution gets both moments and co-moments of order $\leq n$ right, with remaining
+moments and co-moments zero (if such distributions can be defined). That may be interesting, but would
+certainly be a different paper from ours, since our paper is about \emph{relationships} between $E$ and $C$
+(while leaving the univariate distributions of $E$ and $C$ themselves fixed), whereas this idea of the 
+referee would make approximations
+both to the relationship between $E$ and $C$ and to the marginals, simultaneously. Nevertheless, 
+we are pleased the referee 
+is interested in our approach to the extent that they are generating potential ideas for future research.} \\
 
-\noindent C) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\
+\noindent C15) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We modified the sentence.} \\
 
-\noindent C) L41: The “we” is slightly confusing because the author teams share only one author \\
+\noindent C16) L41: The “we” is slightly confusing because the author teams share only one author \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We changed the wording here.} \\
 
-\noindent C) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\
+\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
+extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
+Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
+a bell for most ecologists.} \\
 
-\noindent C) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\
+\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
+the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
+flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
+FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
+growth rates of local populations and high values elevate those growth rates. In that case one may
+expect that left-tail associations increase metapopulation extinction risks because then environmental
+``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
+``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
+high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
+that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
+see the paper for additional details. Thanks for your interest, by the way.} \\
 
-\noindent C) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\
+\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
+in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
+of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
+than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
+reader that storage effects would be defined below, which they are. } \\
 
-\noindent C) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters.
+\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters.
 \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce a clear paper, but, of the comments 
+C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
+and we have revised the text C19 mentions. 
+For instance, the manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
+It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
+the link to that other literature and to show that ATAs have been found to be important in ecology, generally.
+Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
+we cite, and which inspired our paper, and which were published in Ecology Letters. 
+Nevertheless, we have gone through the paper and redoubled our efforts to make the paper broadly accessible.} \\
+%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.
 

### 2023-06-30 paper/Paper.Rnw
continued

@@
 %Intro the paper by intro'ing ATAs
-At the core of ecology is the study of relationships - between organisms and between organisms and their environments. 
+At the core of ecology is the study of relationships between organisms and their environments.
 Ecologists may ask, for instance, how local environment relates to 
@@
 Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
-symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what we have previously referred to as 
+symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
 ``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
@@
 impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
-Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_AER} show that ATAs are common in ecology
-and can influence Taylor's law and Moran effects; \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
+Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
 when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``relative nonlinearity
-in competition'' and ``storage effects'' \citep{Chesson_1994} which will be familiar to many readers. MCT has been
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also defined below).
+MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 

### 2023-07-03 paper/RefereeResponses1.tex
continued working on responses

@@
 in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
-of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
+and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
 than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
-reader that storage effects would be defined below, which they are. } \\
+reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
+defined oihn Theory, which they are.} \\
 
@@
 
-\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce a clear paper, but, of the comments 
+\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
+paper possible, but, of the comments 
 C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
-and we have revised the text C19 mentions. 
-For instance, the manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
+and we have revised the text C19 mentions so as to robustly address that comment (see response above). 
+The manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
 It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
-the link to that other literature and to show that ATAs have been found to be important in ecology, generally.
+the link to that other literature, so that readers can explore it if they so choose, 
+and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
 Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
-we cite, and which inspired our paper, and which were published in Ecology Letters. 
-Nevertheless, we have gone through the paper and redoubled our efforts to make the paper broadly accessible.} \\
+we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
+specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
+central in ecology these days, and increasingly so. It makes no sense to us to sideline them into 
+"specialist journals", which few people read, especially when they are about mainstream topics such
+as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
+paper broadly accessible.} \\
 %We have to go through the paper and redoubled our efforts to make the paper broadly accessible.
 
-\noindent C) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
+\noindent C21) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We replaced the reference.} \\
+%Jasmin can you please do this?
 
-\noindent C) L74: The optimum might actually be the same, but they need different responses Eq.
+\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\
+
+\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\
+
+\noindent C23) Eq.
 1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
+First, the lottery model is a very simple model, at the same level of complexity
+as models introduced in essentially all introductory 
+undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
+the readership of Ecology Letters cannot easily understand the material presented here.
+Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
+The last author (DCR) has found, from decades of pitching 
+quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
+clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
+the kind of mental energy necessary to read the guts of the paper. So this 
+paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
+example. That goal is convincingly 
+accomplished by the paragraph. Given that the math is simple, and also more precise and definitive than 
+qualitative verbal arguments that usually take up space in Introductions, we think this material is actually
+more beneficial to the readership and more effective for the intended motivational purpose than would be more words
+(and we see based on C25 below that the referee appears to agree with us here). 
+Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
+to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
+accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
+Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
+trend of the last few decades of incorporating more quantitative approaches. We believe we
+benefit the field by challenging assumptions and biases such as categorically not putting equations 
+in the Introduction.
+Overall, we think making this suggested change would make the paper weaker, and would pander to an
+outmoded and unhelpful tradition. So we will not make the change unless the 
+editor insists on it as a condition of publication.} \\
 
-\noindent C) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\
+\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
+above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
 
-\noindent C) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\
+\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
+Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
+without first stating the lottery model, since the result uses that model. } \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
+To avoid making claims that we have not fully supported
+with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
+``probably not necessary''. Thanks for spurring us to do this.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
+Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
+the $\#$ term because that terms involves both of them varying (though independently).} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
+used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
+to do this the first time we use the hard brackets. So thanks for pointing this out.
+We do not know what the referee means by ``using curly under brackets'' so have not done anything
+in response to that part of this comment.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
+negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
+work (and potential for error) 
+of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
+After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
+paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
+trying to do here is accurately portray our results about
+when ATAs were important and when they were not. To us, the most forthright way of doing that is to
+show \emph{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
+(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
+we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
+our results seems important when such is our goal. Furthermore, there's no real cost to showing
+all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
+Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
+matters for continuing to explore when ATAs are important and when not is probably the strength
+of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
+a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
+based on such a thing.
+
+Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
+this is mentioned here,
+as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
+shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
+seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
+%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
+%the original paper and extract from p-value info, or we could come up with a better example,
+%or something else, not sure, let's discuss.
+} \\
 
@@
 
-\noindent ***\emph{response} \\
+%Jasmin, isn't there already such a line? It just gets overplotted sometimes, right? But you can see it on
+%panel (e), for instance, right? I've written a response assuming that, but let's briefly discuss.
+\noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
+It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
+for different $y$-axis extents is so that you can see the variation within each panel with maximum 
+clarity. The main point, here, is to compare the relative contributions of different components
+to coexistence, so it's important to show within-plot variation (between lines and across 
+values of the $x$-axis) clearly. We thank the referee for expressing their own uncertainty, which was
+useful to be aware of.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
+(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
+is usually one of the largest-magnitude terms. Our results as to the importance of ATAs for coexistence 
+contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
+of section 4.1.} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
+negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
+it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
+the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
+such errors. But we're happy to hear specific suggestions.} \\
 

### 2023-07-03 paper/Paper.Rnw
continued working on responses

@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also defined below).
-MCT has been
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also reviewed conceptually below and
+defined formally in Theory). MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
@@
 for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
-and on the assumption that optimal environmental conditions for distinct species are also distinct,
-so that no two species find the same conditions to be optimal. Then, good environmental
+and on the assumption that environmental responses for distinct species are also distinct,
+so that no two species respond in the same way to environmental variation. Then, good environmental
 conditions for a currently common species are paired with strong intraspecific competition, because of
@@
 The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
-in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
+in proportion to the reproductive outputs of the two species that year. Note that $N = N_1(t)+N_2(t)$ 
+is constant through time. 
 Notation for the whole paper is summarized in Table \ref{tab:notation}.
@@
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption is not necessary, it simplifies the
+$C_i(t)$. Though the iid assumption is probably not necessary, it simplifies the
 presentation of new ideas.
@@
 correlation \emph{per se}, remaining after having removed the effects of ATAs.
+Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly in our notation,
+throughout the paper - they are not interchangeable.
 The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
@@
 The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
-Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, and 
-Robin Snyder for helpful discussions. 
+Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
+Robin Snyder, and three anonymous referees and the editor for helpful feedback. 
 The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project. 

### 2023-07-03 paper/RefereeResponses1.tex
continued working on responses and revisions

@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Many thanks for the positive feedback!} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Thank you for the positive feedback!} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
+efforts to make the manuscript as accessible as possible.} \\
+%We have to redouble our efforts, as it says.
 
@@
 \noindent ***\emph{response} \\
+%Jasmin, let's discuss. I cannot, myself, say, even now, precisely *how* or *why* ATAs promote or hinder coexistence. 
+%In my view, the paper is about showing that they CAN promote or hinder coexistence, and to what extent, so that 
+%future work can continue to explore when they do and when they don't, and precisely how. So I want to write a response
+%to that effect. However, perhaps you have an intuition, which can be simply explained, as to how ATAs
+%either promote or hinder coexistence. I know that's maybe a lot to ask, but if we have such a thing we should consider
+%putting it in. Ecologists always want a two-sentence, math-less explanation for everything. If quantum mechanics were
+%part of ecology they'd want a two-sentence intuitive explanation for it, even if it was meaningless aside making
+%the reader feel better about themselves. Sometimes such a thing does not exist, but when it does exist it can,
+%at least sometimes, be useful. For the record, I think most intuitive, verbal explanations of storage effects
+%are mostly useless, but perhaps we can do better than those.
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
+lines xyz. Is that enough?>} \\
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
+them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
+we talk about why ATAs matter for one but not the other. I think this is just because the
+effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
+accentuated. But that's not unreasonable, since many species have environments which fluctuate between
+being suitable and unsuitable.>} \\
 
@@
 
+%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
+%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
+%at least for the main text) as not possible because of space constraints. This particular request does not 
+%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
+%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
+%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
+%something like that.
 \noindent ***\emph{response} \\
@@
 \noindent ***\emph{response} \\
+%I think the earlier papers already showed that, at least for the parameters used there? Check. And the other parameters
+%are experimentally manipulable, so this should be OK, but is worth discussing the extent to which ATAs were important
+%for the original parameters. The it would be worth noting that the strongest ATA contributions come for parameters
+%which were not imposed in the original experiment, so it may be worth repeating the experiment but with these other
+%parameters.
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{Thank you for pointing out the novelty.} \\
 
@@
 
+%Need to address. However, this papers seems no more technical than the earlier Ellner paper. And, furthermore, 
+%one of the referees (ref 1) actually asked us to put BACK technical content in the main text which we had 
+%moved to the SI before submission to keep it less technical. That said, we do need to do a pass thu and 
+%figure out what technical content we can reduce. 
 \noindent ***\emph{response} \\

### 2023-07-03 paper/RefereeResponses1.tex
a few last minute changes before meeting with Jasmin
changes made during meeting with Jasmin

@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We address your questions and comments below.} \\
 
@@
 \noindent ***\emph{<Move SI Fig S1 and associated text back to the MT – at least two referees asked for it.>} \\
+%Dan to do.
 
@@
 \noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\
+%Dan to do
 
@@
 So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C.
-This is the same issue with the # notation that has already been established.
+This is the same issue with the # notation that has already been established - elaborate.
 It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space.
-So in the end the referee does not really have a point if one wants to be formal.
-But you can explain all this in a new sup mat section.>} \\
+So in the end the referee does not really have a point, the notation is formally correct.
+But you can explain all this in a new sup mat section, and that's worth doing in its own right as well as to satisfy them and people like them.>} \\
+%Dan to do
 
@@
 \noindent ***\emph{<Yes, the referee is right about this, that's what we're doing.
-Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.>} \\
+Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.
+But if we change the example, as suggetsed by another reviewer Q, we should mention that here, too.>} \\
 
@@
 efforts to make the manuscript as accessible as possible.} \\
-%We have to redouble our efforts, as it says.
+%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
+%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
+%area in ecology?
 

### 2023-07-06 paper/RefereeResponses1.tex
changes I forgot to commit at the last commit

@@
 
-\noindent ***\emph{<Just do this and say we did it - it's a good enough idea.>} \\
+\noindent ***\emph{This was done already, the referee appears to have missed it. It was line 136 in the original submission,
+now line XXX in the resubmitted version.} \\
+%Don't forget to add the new line numbers at the end.
 
@@
 
-\noindent ***\emph{response} \\
+\noindent ***\emph{We changed the wording as requested, and went through the entire manuscript again looking for typos
+and grammatical errors.} \\
+%Jasmin to look at line 100 and revise as needed.
 
@@
 >} \\
+%Dan to do
 
@@
 this, Jasmin can add the new panel to the figure.>} \\
+%Jasmin to make the ammended figure, Dan to move it and the associatde text bck to main text
+%The other main thing Jasmin should do is finalize the repo - checkpoint issues.
 
@@
 see the paper for additional details. Thanks for your interest, by the way.} \\
+%DAN: maybe add to the above to point out that they don't need to know this to understand what we're doing in this paper!
 
@@
 reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
-defined oihn Theory, which they are.} \\
+defined in Theory, which they are.} \\
 
@@
 editor insists on it as a condition of publication.} \\
+%Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.
 
@@
 above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
+%Are we being too petulant, should we just do it and say we did it?
 
@@
 the $\#$ term because that terms involves both of them varying (though independently).} \\
+%Jasmin to check and make sure Dan is right , here. Maybe there is a better place to refer them to to support our claim?
 
@@
 paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
+%Jasmin to consider (don't do yet) how hard/error prone this change would be, taking into account the 
+%formulas in the SI? Unless this is MUCH easier and MUCH less error prone than Dan's hunch, I suggest 
+%we don't do it.
 
@@
 %the original paper and extract from p-value info, or we could come up with a better example,
-%or something else, not sure, let's discuss.
+%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
+%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
+%hopefully drawing people's attention to the right features of this thing.
 } \\
@@
 
-%Jasmin, isn't there already such a line? It just gets overplotted sometimes, right? But you can see it on
-%panel (e), for instance, right? I've written a response assuming that, but let's briefly discuss.
 \noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
@@
 %are mostly useless, but perhaps we can do better than those.
+%
+%Dan and Jasmin to discuss further. Jasmin to dig out old paragraph of text. Jasmin to work on this.
 
-\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect.
-In detail, the explanation of the storage effect is superb, but how ATAs enter in play is not well solved.
+\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...
+
+\noindent ***\emph{Thank you for the positive feedback!} \\
+
+\noindent C) ...but how ATAs enter in play is not well solved.
 It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented.
@@
 \noindent ***\emph{response} \\
+%Same comment again, Jasmin can work on it.
 
@@
 \noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
-lines xyz. Is that enough?>} \\
+lines xyz. Is that enough? Elaborate and re-iterate stuff from the paper.>} \\
+%Dan to do
 
@@
 being suitable and unsuitable.>} \\
+%Jasmin to consider this while brainstorming the prior comment.
 
@@
 \noindent ***\emph{response} \\
+%Jasmin to also lump this in with her thinking about the earlier ref 3 comments. 
+%Q: When you look at E and C marginals for the diatom model, do they have most of their mass in the tails, like the 
+%beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?
 

### 2023-07-07 paper/ExplanationOfATAEffectsWorkingDoc.tex
added files exploring ways to explain why/how ATAs have an effect, in an understandable manner

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=1.5cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{xr}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{Sweave}

%external documents
\externaldocument[SI-]{SupMat}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}

\DeclareMathOperator{\E}{\mathbb{E}}% expected value
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\cor}{cor}
\DeclareMathOperator{\mean}{mean}
\DeclareMathOperator{\se}{se}
\DeclareMathOperator{\sd}{sd}
\DeclareMathOperator{\prob}{P}

%attempt 1 at nat and sharp
%\newcommand{\nat}{\mathlarger{\natural}}
%\newcommand{\shp}{\mathlarger{\sharp}}

%attempt 2 at nat and sharp
%\newcommand{\nat}{\raisebox{1pt}{\mathsmaller{\mathsmaller{/\hspace{-2pt}/}}}}
%\newcommand{\shp}{\#}

%attempt 3 at nat and sharp
\newcommand{\nat}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\shp}{%
\text{\hspace{-1.5pt}
\begin{tikzpicture}[scale=1.8]%
\draw (0,.333ex) -- (1ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}
\newcommand{\test}{%
\text{
\begin{tikzpicture}[scale=1.8]%
\draw (0,0) -- (1ex,0ex);
\draw (0,0) -- (0ex,1ex);
\draw (0,1ex) -- (1ex,1ex);
\draw (1ex,0) -- (1ex,1ex);
\draw (0,.333ex) -- (2ex,.333ex);%
\draw (0,.666ex) -- (1ex,.666ex);%
\draw (.333ex,0) -- (.333ex,1ex);%
\draw (.666ex,0) -- (.666ex,1ex);
\end{tikzpicture}%
}}

\newcommand{\olr}{\overline{r}}
\newcommand{\olrs}{\overline{r}^{\shp}}
\newcommand{\olrn}{\overline{r}^{\nat}}
\newcommand{\bs}{\backslash}

%header material for paper
\title{Working document to try to explain why/how ATAs influence coexistence in an intuitive way for referee 3}
\date{}

\author[a,b]{Pimsupa Jasmin Albert}
\author[a,c,*]{Daniel C. Reuman}

\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
\affil[b]{Environmental Studies Program and Department of Biology, University of Oregon}
\affil[c]{Laboratory of Populations, Rockefeller University}
\affil[*]{Corresponding author, reuman@ku.edu}

%for dealing with line numbers bugs having to do with equations
\let\oldequation\equation
\let\oldendequation\endequation
\renewenvironment{equation}
  {\linenomathNonumbers\oldequation}
  {\oldendequation\endlinenomath}
\let\oldalign\align
\let\oldendalign\endalign
\renewenvironment{align}
  {\linenomathNonumbers\oldalign}
  {\oldendalign\endlinenomath}

\begin{document}

\maketitle

This approach is nothing analogous to attempts I have seen to intuitively explain storage effects. Part of the
reason for that is that those attempts may be rooted at least partly in the Chesson theory, which assumes
small noise. We no longer make that assumption, and we already know (or strongly suspect, based on the 
intuition provided by Steve Ellner) that ATAs will not make any difference under a small noise assumption,
because under such an assumption the approximations of Chesson are good. The approach described here also has other differences, 
discussed below. One major shortcoming is that this approach provides little ecological intuition as to what ATA
effects precisely are, even if it does (I hope) make it seem not only plausible but expected that they should
frequently matter. And it provides some mathematical intuition that could prehaps be substituted for purely
ecological intuition in some sense. I look forward to discussing this approach with Jasmin before possibly putting some version 
of it in the paper.

We mainly consider the lottery model. Though we begin with some more general notation, we quickly make an
assumption that I think will only hold for the lottery model.
We assume, as usual, that for whatever model we consider (a 2-species model), we can write $r_i$ as
a function of $E_i$ and $C_i$, and $r_i$ is an increasing function of $E_i$ and decreasing function
of $C_i$. Then ATA effects are defined as
\begin{align}
\Delta_i^{[EC]} &= (\olr_{i \bs i} - \olr_{j \bs i})-(\olrn_{i \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects1}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}) - (\olr_{j \bs i} - \olrn_{j \bs i}) \label{eq:ATAeffects2}\\
&= (\olr_{i \bs i} - \olrn_{i \bs i}).
\end{align}
The third term in (\ref{eq:ATAeffects2}) is $0$ because that represents growth of $j$ in a steady-state scenario,
and the fourth term is $0$, for the lottery model, because of reasoning in S5.1. So here is where we are 
getting specific to the lottery model, as warned above. DAN: Maybe the next iteration of this thinking should 
consider what happens when that fourth term does not vanish. Can we generalize?

For the lottery model, we have 
\begin{align}
\Delta_i^{[EC]} &= (\olr_{i \bs i} - \olrn_{i \bs i}) \\
&= \E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right] - \E \left[ \ln \left( 1-\delta+\frac{E_i^\nat}{C_{i \bs i}^\nat} \right) \right]. \label{eq:DeltaTerm1} 
\end{align}
The first term of (\ref{eq:DeltaTerm1}) is the average of the function $\ln(1-\delta+x/y)$ over the bivariate 
distribution $(E_i,C_{i \bs i})$, whereas the second term of that expression is the average of the same
function over the distribution $(E_i^\nat,C_{i \bs i}^\nat)$. We can visualize these averages pretty easily and 
effectively (see below) for specific distributions, and it helps to see how ATAs can influence coexistence.
Fig. \ref{thefig} shows this visualization. For the log-normal fecundities model (A-D), the average value of the 
function over the green points should be about the same on each panel as over the black points, 
corresponding to the fact that $\Delta_i^{[EC]}$ is close to zero for the parameters used (see Fig. 2). 
On the other hand, for the beta-fecundities model for left-tail associated noise (E, F), the average of the function 
over the black points is obviously going to be bigger than its average over the green points, corresponding to the 
fact that, for these parameters, $\Delta_i^{[EC]}$ is negative (Fig. 3).
For the beta-fecundities model for right-tail associated noise (G, H), the average of the function 
over the black points is obviously going to be smaller than its average over the green points, corresponding to the 
fact that, for these parameters, $\Delta_i^{[EC]}$ is positive (Fig. 3).

\begin{figure}
\includegraphics[width=\textwidth]{../results_figs/PostHocExplanatoryFig.jpg}
\caption{Pictorial representation of $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ and 
$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$ for the 
log-normal-fecundities lottery model (A-D) and the beta-fecundities lottery model (E-H),
for left-tail associated noise (A, B, E, F), and for right-tail associated noise (C, D, G, H).
The color and contours show $\ln(1-\delta+x/y)$, whereas the green points on each panel show a sample from
the distribution $(E_i,C_{i \bs i})$ and the black points show a sample from the distribution 
$(E_i^{||},C_{i \bs i}^{||})$. Thus the average value of the function over the green points minus its
average value over the black points is a good approximation of $\Delta_i^{[EC]}$. Here $i=1$ and
$j=2$. We used parameters $\sigma=1$, $\mu_1 = 0$, and $\mu_2 = 0.5$ for the log-normal fecundities
model; $\eta_1 = 1$ and $\eta_2 = 1.2$ for the beta-fecundities model; and $\delta=0.6$
for both models. Results are plotted on both linear (A, C, E, G) and log (B, D, F, H) 
scales for readability.}\label{thefig}
\end{figure}

Viewing $\Delta_i^{[EC]}$ as an average of a function over a set of points distributed in a certain way
(with ATAs) minus an average of the same function over a set of points distributed in another way
(with ATAs removed) makes it clear why we would expect to commonly see substantial ATA effects 
on coexistence: removing ATAs really changes the way points in a bivairate distribution are 
distributed, in a way that can interact strongly with averaging over a function representing growth of
a potential invader. In fact, it seems clear that only very special functions will give the same results 
when averaged over a typical distribution with ATAs or with ATAs removed.

Growth from rarity, $\olr_{i \bs i}$ is an average of the individual growth rates which pertain across
years, and those values depend on the details of the distribution of $E$ and $C$. If we use 
$E_i$ and $C_{i \bs i}$, well then of course we will tend to get different values
from if we use $E_i^\nat$ and $C_{i \bs i}^\nat$. The question is just how much different, which is what 
we have sought to answer with this paper for lottery model and experimental examples.

\end{document}

### 2023-07-07 code/ATAEffectDemoAttempt.R

@@ -1,0 +1,394 @@
+#This is a new script as of 2023 07 06, hoping to demo to a referee
+#and other readers some intuition behind the influence of ATAs.
+#Reuman
+#begun 2023 07 06 
+
+#***
+#Setup
+#***
+
+rm(list=ls())
+
+#data locations, copied from the scripts that generated the noise
+numRes_loc <- "../results_numeric/"
+
+noise_loc <- paste0(numRes_loc, "noise.RData")
+M_loc <- paste0(numRes_loc, "M.RDS")
+rho_loc <- paste0(numRes_loc, "rho.RDS")
+
+#locations to save resulting plots in
+resloc<-"../results_figs/"
+
+#***
+#Load noise previously generated (standard normal marginals)
+#***
+
+#load and rename
+load(noise_loc)
+rm(u)
+nl<-b[[1]]
+nr<-b[[2]]
+ns<-b[[3]]
+rm(b)
+
+# #check
+# plot(nl[1:1000,1],nl[1:1000,2],type="p")
+# plot(nr[1:1000,1],nr[1:1000,2],type="p")
+# plot(ns[1:1000,1],ns[1:1000,2],type="p")
+# cor(nl)
+# cor(nr)
+# cor(ns)
+
+#cut from 1000000 to numpoints points
+numpoints<-1000
+nl<-nl[1:1000,]
+nr<-nr[1:1000,]
+ns<-ns[1:1000,]
+
+#***
+#Parameters for a particular beta-fecundities model
+#***
+
+eta2<-1.2
+eta1<-1
+bf_delta<-0.6
+
+#***
+#Calculate E1 and C_{2\1} for the cases of right-tail, left-tail, and symmetric noise, beta-fecundities
+#***
+
+bf_E1l<-eta1*qbeta(pnorm(nl[,1]),0.5,0.5)
+bf_C1bs1l<-eta2*qbeta(pnorm(nl[,2]),0.5,0.5)/bf_delta
+
+bf_E1s<-eta1*qbeta(pnorm(ns[,1]),0.5,0.5)
+bf_C1bs1s<-eta2*qbeta(pnorm(ns[,2]),0.5,0.5)/bf_delta
+
+bf_E1r<-eta1*qbeta(pnorm(nr[,1]),0.5,0.5)
+bf_C1bs1r<-eta2*qbeta(pnorm(nr[,2]),0.5,0.5)/bf_delta
+
+#ATA contributions to coexistence, for checking against Fig. 3 
[…]

### 2023-07-10 paper/ExplanationOfATAEffectsWorkingDoc.tex
additional comments and generalization

@@
 
+\section{Initial thinking}
+
 This approach is nothing analogous to attempts I have seen to intuitively explain storage effects. Part of the
@@
 and the fourth term is $0$, for the lottery model, because of reasoning in S5.1. So here is where we are 
-getting specific to the lottery model, as warned above. DAN: Maybe the next iteration of this thinking should 
-consider what happens when that fourth term does not vanish. Can we generalize?
+getting specific to the lottery model, as warned above. Maybe the next iteration of this thinking should 
+consider what happens when that fourth term does not vanish. Can we generalize? I think we can - see below for
+some efforts to generalize which were added later.
 
@@
 distributed, in a way that can interact strongly with averaging over a function representing growth of
-a potential invader. In fact, it seems clear that only very special functions will give the same results 
+a potential invader. In fact, it seems clear that only very special functions will give the exact same results 
 when averaged over a typical distribution with ATAs or with ATAs removed.
@@
 
+\section{Efforts to generalize}
+
+Equation (\ref{eq:ATAeffects2}) is what we had prior to specifying to the lottery model. 
+Above, we've considered the first two terms, which are the average of $r_i(E_i,C_i)$ over 
+the distribution $(E_i,C_{i \bs i})$ minus the average of the same function
+over $(E_i^\nat,C_{i \bs i}^\nat)$. But the last two terms of (\ref{eq:ATAeffects2})
+are the average of $r_i(E_j,C_j)$ over the distribution $(E_j,C_{j \bs i})$ minus the 
+average of the same function over $(E_j^\nat,C_{j \bs i}^\nat)$. So the same basic
+ideas apply here as in the lottery model case, it's just that we have to make two comparisons 
+of growth functions averaged over distributions, one comparison for the invader, $i$,
+and the other for the resident, $j$. Again, essentially any model will have 
+non-zero value of (\ref{eq:ATAeffects2}) -- only models designed as counterexamples
+should typically have a precisely zero value for that expression. So ATAs will always
+matter to some extent for coexistence. It's just a question of how much, for real systems. 
+And, as far as can tell, that can only be answered by computing these quantities for
+several real systems. We've shown that, at least for some systems, the quantities 
+can be meaningfully large.
+
+\section{Additional commentary/thinking by writing}
+
+OK, it seems clear to me that basically any meanigful change to the distribution 
+$(E,C)$ will typically alter the average value of the nonlinear function $r(E,C)$ 
+over that distribution. There are many ways to decompose the growth rate 
+$\olr_{i \bs i}$ through a series of alterations of the distribution $(E,C)$. 
+Ellner et al. also make this point in their 2019 paper. The original Chesson 
+theory is probably the most natural decomposition under a weak noise assumption,
+but Ellner et al. and our paper make it clear there are other ways that may make
+more sense, in some circumstances, if noise is not weak.
+
 \end{document}

### 2023-07-11 paper/ExplanationOfATAEffectsWorkingDoc.tex
starting on ecological explanation
this for now

@@
 
+\section{Ecological explanation}
+Under Modern Coexistence Theory, the outcome of coexistence is determined by niche differences and fitness differences (Chesson 2000, Adler 2007).
+When niche differences overcome fitness differences, coexistence is possible.
+Niche differences are driven by stabalizing mechanisms (such as the storage effect), which are defined as mechanisms that cause a species to limit itself more than it limits others, i.e., the species experiences more intraspecific competition than interspecific competition.
+
 \end{document}

### 2023-07-12 paper/ExplanationOfATAEffectsWorkingDoc.tex
adding stuff before lunch

@@
 \section{Ecological explanation}
-Under Modern Coexistence Theory, the outcome of coexistence is determined by niche differences and fitness differences (Chesson 2000, Adler 2007).
+Under Modern Coexistence Theory, the outcome of coexistence is determined by 
+niche differences and fitness differences (Chesson 2000, Adler 2007). 
 When niche differences overcome fitness differences, coexistence is possible. 
-Niche differences are driven by stabalizing mechanisms (such as the storage effect), which are defined as mechanisms that cause a species to limit itself more than it limits others, i.e., the species experiences more intraspecific competition than interspecific competition.
+Niche differences are driven by stabalizing mechanisms (such as the storage effect), 
+which are defined as mechanisms that cause a species to limit itself more than it 
+limits others, i.e., the species experiences more intraspecific competition than 
+interspecific competition.
+
+The storage effect stably contributes to coexistence through 
+temporal niche partitioning with EC covariance and buffered population growth.
+Under invader-resident analysis [rare-common], EC covariance limits the resident by 
+coupling high environmental responses with high intraspecific competition. The magnitude
+of the storage effects then depends on the EC covariance of the invader. Because the 
+invader is introduced into the resident steady state community at close to zero density, 
+its C is only a result of interspecific competition from the resident. Thus, the more closely 
+related the invader's EC relationship is, the more interspecific competition it experiences, 
+destabalizing coexistence. Equivalently, invader EC relationships with more spread stably
+contribute to coexistence (through the storage effect).
+As seen in Fig.
+\ref{thefig}, ATAs in the invader EC relationship changes its bivariate distribution to both have more spread about the "central line" and be more densely situated about the "central line".
 

### 2023-07-12 paper/ExplanationOfATAEffectsWorkingDoc.tex
added a good explanation

@@
 \section{Ecological explanation}
+\subsection{Basic MCT}
 Under Modern Coexistence Theory, the outcome of coexistence is determined by 
@@
 
+\subsection{Storage effect and how ATAs can change it}
 The storage effect stably contributes to coexistence through 
@@
 destabalizing coexistence. Equivalently, invader EC relationships with more spread stably
-contribute to coexistence (through the storage effect).
-As seen in Fig.
-\ref{thefig}, ATAs in the invader EC relationship changes its bivariate distribution to both have more spread about the "central line" and be more densely situated about the "central line".
+contribute to coexistence (through the storage effect). As seen in Fig. \ref{thefig}, ATAs in 
+the invader EC relationship changes its bivariate distribution to both have more spread 
+about the "central line" and be more densely situated about the "central line".    
+When the invader's EC relationship has more spread (more density in values far away from 
+the central line) than what the invader's EC relationship would be if it was symmetric, 
+the ATA effect is positive.
+More specifically, points to the right of the central line are where environmental response is high and competition is low.
+When points fall in this plane, is when species can "store" the beneficial effects of a good environment to counteract when conditions are less ideal.
+Thus when the EC relationship is less spread, the invader is less able to recover from rare because the points are far away from the beneficial extremes.
+I think this related to the subadditivity criteria of storage effect, i.e., buffered population growth which is related to nonlinear averaging.
+Subaddivity/ buffered popualtion growth is when species growth rate responses to competition are less affected/sensitive to bad environmental conditions and more affected and sensitive when conditions are good.
+So it is nonlinear.
+So points in the beneficial extremes are when the invader can take advantage of the nonlinear averaging that comes with buffered population growth.
+ATA'd EC relationships have different distribution of points compared to one that is symmetric with the same marginals and correlation so the asymmetry changes whether or not points are in the beneficial extremes.
+
+\subsection{How can ATAs in the EC relationship arise?}
+\emph{Water limitation.} Everyone depends on water.
+When water is limited niches become less different.
+So the invader's EC relationship in low values of environmental response is more correlated to the competitive pressure it experiences from the resident since the resident is also experiencing water limitation and will be in its low environmental response values.
+When water is not the limiting factor, environmental responses take on higher values and thus, the EC relationship may be more spread.
 

### 2023-07-13 paper/RefereeResponses1.tex
working on replies - moved fig S1 back to main text

@@
 \noindent \emph{We submit the following pdf files: ...} \\
+%DAN: Fill in ...
 
@@
 main text latex: ...} \\
+%DAN: Fill in ...
 
@@
 
-%The C) that precedes each comment is because we will later add numbers, C1, C2, etc., after making sure these
-%comments are divided up to our liking. Then we can refer to each comment by its number, which can be useful.
 \noindent C1) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.\\
@@
 
-\noindent ***\emph{<Move SI Fig S1 and associated text back to the MT – at least two referees asked for it.>} \\
-%Dan to do.
+\noindent ***\emph{We have done what the referee requests, here, and we agree with this request. Thanks for
+making the request. This request is somewhat in tension with other requests, from other referees, 
+to minimize the quantitative 
+content of the main text, and to keep the overall length under the journal limit while adding a bunch of
+verbal explanations requested by the other referees. We point this tension out here, and refer back to it in
+other responses, below. We also point out that we have not only moved the figure back to the main text,
+we have also added a panel to the figure in response to the related comment CX, below.} \\
+%DAN: Waiting on Jasmin to add the new panel. In other replies, you can refer to this reply to justify when you
+%decide to add explanations to the SI instead of to the main text, and when you have to respond to requests
+%to minimize the quant. Also fill in the correct value of CX above.
 
@@
 
-\noindent ***\emph{<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>} \\
+\noindent ***\emph{This is a good question, thanks for asking it. The extent to which the process is 
+unique is now covered in a new SI section, refered to from the main text at the end of section 2.3, 
+right at the end of the text the referee asked us to move to the main text. There are some choices involved,
+but they are pretty cannonical. See the new SI section for explanation. } \\
 %Dan to do
+%<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>
 

### 2023-07-13 paper/Paper.Rnw
working on replies - moved fig S1 back to main text

@@
 between the components altered relative to the original distributions. Specifically, asymmetries of 
-tail association are removed -- see SI section \ref{SI-sect:remove_asymmetries} for how this is done. We then define 
+tail association are removed -- see below for how this is done. We then define 
 $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
@@
 below for additional details.
+
+How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
+correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
+We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
+Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
+$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
+distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
+Let $\rho_i$ be the covariance of
+these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
+$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
+$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
+and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
+The value of $\ddot{\rho}_i$ is then determined in one of 
+four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
+$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
+For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
+we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
+The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
+$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
+as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
+For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
+the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
+defined in the previous section (following \cite{Ellner_2016}).
+%DAN: Jasmin will be adding that panel.
 
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/theory_fig_col.pdf}
+\caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
+between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
+(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
+marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
+$\varphi^{-1}$, and 
+each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
+(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
+selected to match that of the original distribution (see Theory). (d) Two new variables, 
+$E_i^{\nat}$ and  $C_{i \bs i}^{\nat}$,
+which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
+which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
+$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
+Panel (e) show the distribution $(E_i^\shp,C_{i \bs i}^\shp)$, defined in Theory
+following \citep{Ellner_2016}, for comparison.
+%DAN: jasmin is adding this new panel
+For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
+$\log_{10}$ of probability density on all panels. 
+\label{fig:pedagog2}}
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=\textwidth]{../results_figs/decompFigLLN.pdf}

### 2023-07-13 paper/SupMat.Rnw
working on replies - moved fig S1 back to main text

@@
 by researchers making that choice.
-
-\section{Removing asymmetries of association}\label{sect:remove_asymmetries}
-
-How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
-correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
-We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
-$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
-Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
-$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
-distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
-Let $\rho_i$ be the covariance of
-these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
-$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
-$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
-$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
-and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
-The value of $\ddot{\rho}_i$ is then determined in one of 
-four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
-$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
-$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
-For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
-we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
-The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
-$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
-$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.
 
@@
 
-%***DAN: This looks pretty good except one of the histograms on c does not look like the plotted
-%pdfs of the marginals. I think Jasmin said in an email she was fixing this.
-%***JASMIN: fixed!!
-%***DAN: Jasmin, also, we use \varphi in the text, but \phi here. These need to be made consistent, since they are different to
-%a math person. The easiest way will be to change it in the figure, so I hope that can straightforwardly be done.
-%***JASMIN: we said this is OK. Operating system thing - renders the same in Windows. 
-\begin{figure}
-\includegraphics[width=1.025\textwidth]{../results_figs/theory_fig_col.pdf}
-\caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
-between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
-(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
-marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
-$\varphi^{-1}$, and 
-each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
-(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
-selected to match that of the original distribution (see Theory). (d) Two new variables, 
-$E_i^{||}$ and  $C_{i \bs i}^{||}$,
-%Dan replaced \nat in the above two expressions by || because for some reason the \nat does not
-%compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
-which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
-which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
-$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
-For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
-$\log_{10}$ of probability density on all panels. 
-\label{fig:pedagog2}}
-\end{figure}
-
-\clearpage
-\newpage
-
 \begin{figure}

### 2023-07-14 code/makTheoryFigs.R
new theory fig for main

@@ -49,18 +49,23 @@ ly_mat<- matrix(c (	2,4,17,6,8,17,10,12,17,14,16,
 
 #colored points
 #### start pdf ####
-pdf(theoryfig_col, height=4, width=15)
-layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
-par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
-source("./theoryFig_script.R")
-dev.off() 
+#pdf(theoryfig_col, height=4, width=15)
+#layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
+#par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
+#source("./theoryFig_script.R")
+#dev.off() 
 ########
 
-# fig for reviewers
+#new fig for reviewers
 #### start pdf ####
-reviwer <- !TRUE
-pdf(theoryfig_col2, height=4, width=15)
-layout(ly_mat, heights=hts, widths=wds)#;layout.show(n=16)
+new <- TRUE
+ly_mat <- cbind(ly_mat, c(17,17), c(19,18), c(21,20))
+lblc[2:3,"x"] <- lblc[2:3,"x"]+.17
+lblc["E","y"] <- lblc["E","y"]+.1  
+lblc["C","y"] <- lblc["C","y"]-.18  
+wds[(1:3)*3] <- 0.2
+pdf(theoryfig_col2, height=4, width=19)
+layout(ly_mat, heights=hts, widths=c(wds,wds[3:5]))#;layout.show(n=16)
 par(mar=c(0,0.25,0,0.25), oma=c(6,5,2,3), xpd=NA)
 source("./theoryFig_script.R")
 dev.off() 

### 2023-07-14 code/theoryFig_script.R

@@ -17,34 +17,25 @@ text(x=lblc["C","x"], y=lblc["C","y"], labels=expression('C'['i\\i']), srt=90, c
 	# TeX('$E_i$')	# TeX('($C_{i\\setminus i}$)')
 text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste0("(",letters[1],")"),cex=2)
 
-# B2 BETA MARGS NO COV CLAY COP
-# same marginals - no correlation (variance per se)
-sampsB2 <- theoryfigpanel(dbmarg, pbmarg, qbmarg, shcop, bx, by, samps=sampsA, method="var_perse",col=T,log_points=T, plot=reviwer)
-if (reviwer){
-text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression('E'['i']^'#')), cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=c(expression('C'['i\\i']^'#')), srt=90, cex=2)
-text(x=lblc["lab","x"], y=lblc["lab","y"],labels=paste0("(",letters[2],")"), cex=2)
-}
+
 # B NORM MARGS CLAY COP
 # EC marginals individially normalized
-sampsB <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ccop, nx, ny, samps=sampsA,oldp=pbmarg,xpd=NA,col=T, log_points=FALSE, plot=!reviwer)
-if (!reviwer){
+sampsB <- theoryfigpanel(dnmarg, pnmarg, qnmarg, ccop, nx, ny, samps=sampsA,oldp=pbmarg,xpd=NA,col=T, log_points=FALSE)
 text(x=lblc["E","x"], y=lblc["E","y"], labels=c(expression(varphi^-1*('F'['E'['i']]("E"["i"])) )), cex=2)
-text(x=lblc["C","x"], y=lblc["C","y"], labels=c(expression(varphi^-1*('F'['C'['i\\i']]("C"["i\\i"])) )), srt=90, cex=2)
+text(x=lblc["C","x"]-.05, y=lblc["C","y"], labels=c(expression(varphi^-1*('F'['C'['i\\i']]("C"["i\\i"])) )), srt=90, cex=2)
 #TeX('$\\varphi^{-1} ( F_{E_i}(E_i) )$')
 #TeX('$\\varphi^{-1} ( F_{C_{i\\setminus i}}(C_{i\\setminus i}) )$')	
[…]

### 2023-07-14 paper/RefereeResponses1.tex
looking through and numbered rest of comments

@@
 other responses, below. We also point out that we have not only moved the figure back to the main text,
-we have also added a panel to the figure in response to the related comment CX, below.} \\
+we have also added a panel to the figure in response to the related comment C13, below.} \\
 %DAN: Waiting on Jasmin to add the new panel. In other replies, you can refer to this reply to justify when you
@@
 
-\noindent C) L126: Do you do this in the appendix? \\
+\noindent C26) L126: Do you do this in the appendix? \\
 
@@
 
-\noindent C) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\
+\noindent C27) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\
 
@@
 
-\noindent C) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\
+\noindent C28) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\
 
@@
 
-\noindent C) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\
+\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\
 
@@
 
-\noindent C) L316: So why do you show values of $\sigma >6$?
+\noindent C30) L316: So why do you show values of $\sigma >6$?
 And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA.
@@
 
-\noindent C) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\
+\noindent C31) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\
 
@@
 
-\noindent C) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case.
+\noindent C32) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case.
 \\
@@
 
-\noindent C) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\
+\noindent C33) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\
 
@@
 
-\noindent C) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked.
+\noindent C34) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked.
 ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect.
@@
 
-\noindent C) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\
+\noindent C35) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\
 
@@
 
-\noindent C) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations.
+\noindent C36) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations.
 Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics.
@@
 
-\noindent C) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
+\noindent C37) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
 \\
@@
 
-\noindent C) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...
+\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...
 
@@
 
-\noindent C) ...but how ATAs enter in play is not well solved.
+\noindent C39) ...but how ATAs enter in play is not well solved.
 It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented.
@@
 
-\noindent C) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor.
+\noindent C40) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor.
 Could the authors briefly indicate how this could occur?
@@
 
-\noindent C) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\
+\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\
 
@@
 
-\noindent C) The model used for the chemostat is in its formulation quite different compared to the lottery model.
+\noindent C42) The model used for the chemostat is in its formulation quite different compared to the lottery model.
 Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model.
@@
 
-\noindent C) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
+\noindent C43) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
 this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones.
@@
 
-\noindent C) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\
+\noindent C43) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\
 
@@
 
-\noindent C) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists.
+\noindent C44) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists.
 My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing.

### 2023-07-16 code/ATAEffectDemoAttempt.R
worked on the plotting in the new figure for explaining

@@ -19,6 +19,9 @@ rho_loc <- paste0(numRes_loc, "rho.RDS")
 #locations to save resulting plots in
 resloc<-"../results_figs/"
 
+#graphical params
+numhistbins<-12
+
 #***
 #Load noise previously generated (standard normal marginals)
 #***
@@ -102,16 +105,17 @@ mean(log(1-lnf_delta+lnf_E1r/lnf_C1bs1r))-mean(log(1-lnf_delta+lnf_E1s/lnf_C1bs1
 
 #figure dimensions, inches
 xaxht<-0.5
-yaxwd<-0.5
+yaxwd<-0.6
 gap<-0.3
 totwd<-6.5
 panwd<-(totwd-4*gap-yaxwd)/4
 panht<-panwd
-totht<-2*xaxht+2*panht+gap
+totht<-3*xaxht+3*panht+gap
 cex_lab<-1.5
 cex_lab_cont<-0.5
 
 jpeg(file=paste0(resloc,"PostHocExplanatoryFig.jpg"),width=totwd,height=totht,units="in",quality=90,res=400)
+#pdf(file=paste0(resloc,"PostHocExplanatoryFig.pdf"),width=totwd,height=totht)
 
 #***
 #Now generate the distributions of r_{i\i} for left tail noise and symmetric noise,
@@ -131,27 +135,32 @@ i<-1
 j<-1
 par(fig=c((yaxwd+(j-1)*(panwd+gap))/totwd,
           (yaxwd+(j-1)*(panwd+gap)+panwd)/totwd,
-          (xaxht+(2-i)*(panht+xaxht))/totht,
-          (xaxht+(2-i)*(panht+xaxht)+panht)/totht),
+          (xaxht+(3-i)*(panht+xaxht))/totht,
+          (xaxht+(3-i)*(panht+xaxht)+panht)/totht),
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 
 #first make the color image
+lims<-c(0,120)
+dlims<-diff(lims)
 x<-seq(from=0.01,to=120.01,by=0.1)
 y<-x
 z<-outer(x,y,function(x,y){log(1-bf_delta+x/y)})
 breaks<-pretty(z,10)
 image(x,y,z,xlab="E1",ylab="C1bs1",col = hcl.colors(length(breaks)-1, "YlOrRd", rev = TRUE),
-      breaks=breaks) 
+      breaks=breaks,xlim=lims,ylim=lims) 
[…]

### 2023-07-17 paper/RefereeResponses1.tex
worked on the plotting in the new figure for explaining
continued work on responses and related edits to ms

@@
 \noindent ***\emph{Thank you for the positive feedback!} \\
+%DONE
 
@@
 \noindent ***\emph{We address your questions and comments below.} \\
+%DONE
 
@@
 content of the main text, and to keep the overall length under the journal limit while adding a bunch of
-verbal explanations requested by the other referees. We point this tension out here, and refer back to it in
+verbal explanations requested by the other referees. We point out this tension here, and refer back to it in
 other responses, below. We also point out that we have not only moved the figure back to the main text,
 we have also added a panel to the figure in response to the related comment C13, below.} \\
-%DAN: Waiting on Jasmin to add the new panel. In other replies, you can refer to this reply to justify when you
+%DONE
+%DAN: In other replies, you can refer to this reply to justify when you
 %decide to add explanations to the SI instead of to the main text, and when you have to respond to requests
-%to minimize the quant. Also fill in the correct value of CX above.
+%to minimize the quant.
 
@@
 unique is now covered in a new SI section, refered to from the main text at the end of section 2.3, 
-right at the end of the text the referee asked us to move to the main text. There are some choices involved,
-but they are pretty cannonical. See the new SI section for explanation. } \\
-%Dan to do
-%<Address the issue of whether and to what extent the process of removing ATAs is unique – probably just with a one-sentence callout in the main text, and some material in the SM (possibly a new section)>
+right at the end of the text the referee asked us to move to the main text. There is a choice involved,
+but it is pretty canonical -- certainly more canonical than the kinds of choices which are often made
+in MCT, e.g., $E$ and $C$ can sometimes be defined in more than one way. Basically the choice involves 
+using a bivariate normal distribution instead of another, more esoteric 
+distribution with standard normal marginals and symmetric tail association. The bivariate normal is by far the 
+most natural example of such a distribution. See the new SI section for further explanations. }
+
+\emph{We are reticent, at this time, to launch into a lot of robustness explorations with regard to this choice, 
+as the referee may be suggesting, for several reasons. First, other referees and the editor have urged us to
+simplify, and such explorations would complexify instead. Second, as mentioned above, the choice we made is very
+natural and quite standard, so exploring other possible choices seems therefore of 
+secondary importance. Finally, our overall goal in writing this paper
+was to show that ATAs may be important for coexistence and, as such, deserve additional exploration in this
+context by the wider research community. As long as any choices we made are reasonable (they are), our point is
+made, regardless of whether making some more esoteric choices might have altered outcomes. As we already argue in the
+Discussion, since we have now showed that ATAs \textbf{can} be important, future work should explore \textbf{when} 
+they are important, and part of 
+that exploration could be whether some more esoteric choice for $(e_i c_{i \bs i})$ might alter our results. }\\
+%DONE
 
@@
 
-\noindent ***\emph{<This one asks for possibly changing notation.
-I don’t think we should do it, but we need to explain in a response what is actually going on, why we chose the notation that we do, and then add a new SM section addressing what the referee is talking about.
-Actually it’s the bivariate distribution that changes when you apply ||, tho the marginals are not changed.
-So the referee does have a bit of a point – E^{||} by itself is actually the same distribution as E, it’s just that it’s a different random variable in the sense of being related differently to C^{||} than E is to C.
-This is the same issue with the # notation that has already been established - elaborate.
-It’s actually formally correct in the sense of random variables if we view *all* our random variables as being measurable functions from the *same* probability space.
-So in the end the referee does not really have a point, the notation is formally correct.
-But you can explain all this in a new sup mat section, and that's worth doing in its own right as well as to satisfy them and people like them.>} \\
-%Dan to do
+\noindent ***\emph{This comment has led us to put some clarifying technical comments in a new SI section 2,
+so thanks for making the comment.}
+
+\emph{We use notational conventions of random variables which are formally correct and standard, 
+as we now explain in more detail in the SI. 
+$E_i$, $C_{i \bs i}$, $E_i^{\nat}$, $C_{i \bs i}^\nat$ and the earlier 
+notation $E_i^\shp$ and $C_{i \bs i}^\shp$ all denote \textbf{random variables}, using a 
+notational approach which is standard in applications of probability. 
+Whereas the \textbf{distributions} of the random variables $E_i$, $E_i^{\nat}$ and $E_i^\shp$ (for instance) are all the same,
+these things are \textbf{not} the same \textbf{as random variables} because their relationships with other random variables 
+differ. The relationships between random variables are considered instrinsic
+to the notation. Thus it's not that somehow the expected value in $\E[r(E,C)]$ is a different definition of expected
+value than the one in $\E[r(E^{||},C^{||})]$ -- there is only one expected value operator. 
+Instead, $r(E,C)$, $r(E^{||},C^{||})$ and $r(E^\shp,C^\shp)$ differ from each other 
+as random variables because of the distinct relationships between their
+different $E$ and $C$ constituents. It would be formally 
+incorrect, and confusing to readers familiar with probability, to try to define multiple expected value 
+operators, as the referee seems to suggest, when what fundamentally differs, here, is the random variables 
+themselves. Readers familiar with
+probability already know what the one accepted definition of expected value is, would typically
+not be comfortable with providing additional definitions, and are comfortable handling these concepts
+with the notational conventions we have used. To remind the referee of some concepts from basic probability
+which illustrate these conventions, recall that if $X$ and $Y$ are random variables, the expectation of the 
+product, $\E(XY)$, depends not only on the univariate distributions of $X$ and $Y$, but also on the relationship
+between $X$ and $Y$, namely their joint distribution. In particular, $\E(XY)=\E(X)\E(Y)$ if $X$ and 
+$Y$ are independent, but not necessarily otherwise. These definitons and notation are completely 
+standard in probability theory; notice that we use the same definition of $\E$ regardless of 
+whether $X$ and $Y$ are independent or nonindependent, it's just that the product random variable
+$XY$ differs accordingly so you get different results of $\E(XY)$ in those distinct cases.}
+
+\emph{Our notation also parallels that of Ellner et al (2016, 2019). To make it
+easier for readers to understand and to reduce the potential for future confusion, it makes sense to continue
+using existing notation whenever that notation is formally correct and otherwise effective (it is, as explained above).}
+
+\emph{Nevertheless, the referee's comment has made us understand that some readers unfamiliar with random variable 
+notation may benefit from putting some
+of these explanations in the SI, so we have done so.} \\
+%DONE
 
@@
 
-\noindent ***\emph{<Yes, the referee is right about this, that's what we're doing.
-Just admit to it and justify doing it anyway because a) we were asked to include this kinda thing by poeple who read it before we submitted; b) the value is not only in showing that ATAs occur, but also in helping people understand more fully what they look like; and c) we cannot show ATAs in E and C (at least not this early in the paper), becuase no one has looked at this before so there is no prior evidence of it.
-But if we change the example, as suggetsed by another reviewer Q, we should mention that here, too.>} \\
+\noindent ***\emph{Yes, the referee is right about this, we are not illustrating ATAs in $E$ and $C$ at this stage of the 
+paper - the reader would anyway not be equiped to handle that at this stage, since we have not even formally defined 
+$E$ and $C$ at the point this figure is first cited (in the very first paragraph). 
+We were asked to include this kind of empirical illustration 
+by poeple who read the manuscript before we submitted it, so it seems important to some readers to have this. 
+The value is not only in showing that ATAs occur in nature, but is priarily  
+in helping people understand more fully what they look like when they occur by giving real examples. 
+We appreciate what the referee is saying,
+here, that perhaps one non-$EC$ example of ATAs from nature is not going
+to give much motivation for looking for ATAs in $E$ and $C$. However, we have published droves of other examples of
+ATAs in nature, and we now cite these (changing what used to be ``Fig. 1d,e show contrasting examples of ATAs in nature
+using plankton population density time series'' to ``Fig. 1d,e show contrasting examples of ATAs in nature
+using plankton population density time series; and many other examples have also been previously explored (Ghosh et al 2020)''). 
+And anyway we cannot produce a prior example of ATAs between $E$ and $C$ because there are no prior examples - this paper 
+is, to our
+knowledge, the first time anyone has looked at ATAs between $E$ and $C$. Fortunately, we think the other motivational
+content of the Introduction is strong, and should stand on its own to motivate readers to spend their mental 
+energy reading this paper. In partcular, we show in the Introduction, for one simple simulation, that ATAs between E and
+C make a visually striking difference between simulation outcomes that appears clearly to influence coexistence.} \\
+%DONE
 
@@
 \noindent ***\emph{Thank you for the very positive feedback!} \\
+%DONE
 
@@
 
-\noindent ***\emph{<Add a discussion, either in the Discussion or in an SM section called something like 
-``Additional discussion points". Acknowledge the viewpoint, state that insufficient studies have been done to be sure
-. Thank the referee for the openness that this is their own opinion - it's very welcome when a referee qualifies
-what they say in this way which permits consideration of alternative viewpoints rather than just being forced to
-do what referees say because they are in the position of power. Point out that we have not only worked with the lotery
-model, as the referee mentions, we have also worked with an empirical example, and that one definitely supported
-the importance of ATAs. Though we accept that future work may reveal that storage effects, and therefore probably
-also ATAs, are not so important in field systems, we think it is also possible that this will not be 
-revealed to be the case, and so we think it is important to not pre-judge the issue when (as the referee 
-admits) there is insufficient data. Our results add another piece of evidence to the small collection of 
-evidence as to the importance of storage effects, as well as adding the first evidence on the potential 
-importance of ATAs. Also emphasize what the referee already seems to accept - that even if storage effects are 
-small, ATA effects can be larger in magnitude if correltion effects per se counteract ATA effects.
+\noindent ***\emph{We thank the referee for this viewpoint. We also thank the referee for acknowledging that this
+is their opinion, and that data are insufficient for a clear conclusion to be reached at this point. We find that 
+such an acknowledgment is both rare from a referee and very useful for causing authors to most usefully revise their 
+paper. We have added new discussion material covering these topics and find that the new material improves the paper.
+We tried to represent both viewpoints in the new material, and we cited the papers the referee mentions.}
 
-Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.
+\emph{We point out that we worked not only
+with the lottery model, but also with an empirical diatom system. That system definitely supported the importance of 
+ATAs for coexistance, which would tend to give at least one data point in favor of the importance of these phenomena, though we 
+acknowledge that our lab systems is also a system ``where niche differentiation via ... predators is impossible.''
+It may also be a system ``where niche differentiation via different resources ... is impossible,'' though resource 
+differentiation can happen when you don't expect it. It seems to us 
+the first step in assessing the importance of a newly recognized phenomenon has to be examining it in simple systems
+such as we have used, prior to examing it in field systems, in all their complexity, where niche differentiation
+by resources and predators can more readily occur. }
+
+\emph{We also point
+out (and the referee already seems to accept this) that it is possible, \emph{a priori}, for ATAs to be 
+important for coexistence in a system even while storage effects are not.
+Storage effects are ATA effects plus what we call the effects of $EC$ correlation \emph{per se} 
+($\Delta_i^{[E \nat C]} in our notation), and either of these two summands can be negative or positive.
+So if the twe effects are oppostite in sign and nearly equal, storage effects will be unimportant but ATA
+effects will be important for coexistence.}
+
+\emph{So the referee appears to agree with us that storage effects and ATA effects may or may not turn
+out to be broadly important in ecology. Our research goals were to introduce the idea of ATA effects on
+coexistence and to show that ATAs \text{may} be broadly important for coexistence, by showing examples where they
+\textbf{are} important. The few examples we showed are all that are readily available at this time. 
+Future work can assess, broadly, how important ATAs are by looking across a larger number of systems, once
+available. These are two different goals. It seems crucial to us not to allow insufficiently tested 
+opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
+goals, so we thank the referee for not doing that.}
+
+\emph{The new Discussion material is in a new SI section which is cited from the main text Discussion section. 
+We would have liked to put some of the new material into the main text. But length limits, together with
+clear requests from other referees to add a variety of content to the main text, made doing so impossible.}
+%DAN: Write the discussion material which is promised and the citation from the Discussion.
+
+% 
+% Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.
 For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition.
@@
 We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."
-
-Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
-It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
-has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
-to the Ellner approach of quantifying storage effects?
-
-Finally, I notice that sometimes for our results, ATA effects and effects of correlation per se counter
-each other. If that happens, storage effects could be weak, while ATA effects are still strong.
-
-I suppose one must always accept the possibility that future work will render one's results less
-meaningful or obsolete.
->} \\
+% 
+% Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
+% It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
+% has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
+% to the Ellner approach of quantifying storage effects?
+} \\
 %Dan to do
@@
 
-\noindent ***\emph{<Clearly since two referees have now mentioned it we need to put some version of Fig S1 back into the
-main text. The earlier referee wanted to see the whole figure brought back, but this referee wants to see just part, i.e.,
-a simpler version, as specified. It's not clear to me how it would help the reader to just see the three panels the referee
-suggests. Only 2 of the suggested panels are currently part of Fig. S1 - the suggestion of putting in the uncorrelated
-distribution would be new. The current Fig. S1 serves the purpose of describing exactly the process of removing ATAs. 
-The new suggested figure would not serve that purpose, it would instead maybe give the reader a sense for what these
-distributions look like when manipulated in the ways we are talking about. But it's not real clear to me that much
-intuition will be conveyed by doing that. One way to reply would be just to add a panel to the figure that shows 
-the uncorrelated case, and move the resulting 5-panel figure back in. Then we'd explain that 1) the other referee asked
-for the whole figure, 2) it's not clear to us that the simplified version serves any good purpose, it certainly 
-does not serve the purpose of explaining the process of removing ATAs, but 3) we went ahead and added the uncorrelated
-case anyway, since that's what was used previously, so it makes sense to have that point of comparison. If we do
-this, Jasmin can add the new panel to the figure.>} \\
-%Jasmin to make the ammended figure, Dan to move it and the associatde text bck to main text
-%The other main thing Jasmin should do is finalize the repo - checkpoint issues.
+\noindent ***\emph{In comment C4 above, another referee asked us also to put S1 back into the main text, so
+we have done as the referee asks here. The new Fig. 2 has the three elements the referee asks for, here, as 
+its panels (a), (e) and (d). We retained the other panels as well, since those were requested by the 
+other referee. Regarding the second part of the referee's comments, 
+there is nothing in our paper for which the notation $E_i^{(EC)}$ is used, but we think what the referee must
+be requesting here is what we remind the reader that $(E_i^\shp,C_{i \bs i}^\shp)$ is used to compute storage effects,
+$\Delta_i^{(EC)}$, and that $(E_i^\nat,C_{i \bs i}^\nat)$ is used to compute ATA effects, $\Delta_i^{[EC]}$. We have 
+now added such a reminder to the figure caption.} \\
 
@@
 when ATAs were important and when they were not. To us, the most forthright way of doing that is to
-show \emph{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
+show \textbf{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
 (as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
@@
 a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
-based on such a thing.
+based on such a thing.}
 
-Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
+\emph{Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
 this is mentioned here,

### 2023-07-17 paper/Paper.Rnw
line 100 (C9)
continued work on responses and related edits to ms

@@
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
-plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
-not yet been investigated. Doing so is the overarching goal of this study.
+plankton population density time series; and many other examples have also been previously explored 
+\citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
+competition and coexistence have not yet been investigated. Doing so is the overarching goal of this study.
 
@@
 %Goal para
-We previously demonstrated that ATAs are common in nature can influence ecological phenomena
+We previously demonstrated that ATAs are common in nature and can influence ecological phenomena
 \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
@@
 The new symbol should correspondingly be pronounced ``partial sharp''.
+See SI section \ref{SI-SIsect:techcomm} for technical comments on the new notation. 
 The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
@@
 $C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
+as desired. The question of the uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
+is addressed in SI section \ref{SI-SIsect:uniqueness}.
+The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
 For comparison, Fig. \ref{fig:pedagog2} also shows, for the example explored by that figure, 
@@
 defined in the previous section (following \cite{Ellner_2016}).
-%DAN: Jasmin will be adding that panel.
 
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../results_figs/theory_fig_col.pdf}
+\includegraphics[width=\textwidth]{../results_figs/theory_fig_col2.pdf}
 \caption[Visual guide to removing asymmetries of association]{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
-between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
+between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATAs present. 
 (b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
@@
 which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
-which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
+which have symmetric association structure. These are the distributions used to compute ATA effects, 
+$\Delta_i^{[EC]}$ (Theory). 
+These variables are obtained by transforming the marginals from (c), 
 $e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
-Panel (e) show the distribution $(E_i^\shp,C_{i \bs i}^\shp)$, defined in Theory
-following \citep{Ellner_2016}, for comparison.
-%DAN: jasmin is adding this new panel
+Panel (e) shows the distribution $(E_i^\shp,C_{i \bs i}^\shp)$, defined in Theory
+following \cite{Ellner_2016}, for comparison. These are the distributions used to compute storage effects,
+$\Delta_i^{(EC)}$ (Theory).
 For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 

### 2023-07-17 paper/SupMat.Rnw
continued work on responses and related edits to ms

@@
 by researchers making that choice.
+
+\section{Technical comments on notation}\label{SIsect:techcomm}
+
+The quantities $E_i$, $C_{i \bs i}$, $E_i^\shp$, $C_{i \bs i}^\shp$, $E_i^\nat$, $C_{i \bs i}^\nat$
+and related quantities are random variables, and hence notational conventions for random variables 
+which are standard in fields of applied probability are used throughout the paper.
+Relationships are considered intrinsic to random variable notation; for instance, though  
+$E_i$, $E_i^\shp$, and $E_i^\nat$ have the same distribution, they are not the same as random variables
+because of their different relationships with other random variables specified in the paper, e.g.,
+$C_{i \bs i}$,  $C_{i \bs i}^\shp$, and $C_{i \bs i}^\nat$. Therefore the distributions
+of the bivariate random variables $(E_i,C_{i \bs i})$, $(E_i^\shp,C_{i \bs i}^\shp)$, and
+$(E_i^\nat,C_{i \bs i}^\nat)$ are all distinct, as are the random variables
+$r(E_i,C_{i \bs i})$, $r(E_i^\shp,C_{i \bs i}^\shp)$, and $r(E_i^\nat,C_{i \bs i}^\nat)$.
+
+\section{Uniqueness of our definition of ATA effects}\label{SIsect:uniqueness}
+
+The description in the main text of how the distribution $(E_j^\nat, C_{j \bs i}^\nat)$
+is constructed already reveals that a choice 
+is made as to what type of correlation (e.g., Pearson, Spearman, etc.) is to be preserved
+in the new distribution. Another choice is also made, though it is quite a canonical choice.
+The choice involves the use of a bivariate normal distribution $(e_i,c_{i \bs i})$ 
+with standard normal marginals (see Theory). Any other symmetrically tail associated bivariate 
+distribution with standard normal marginals could also be used, instead, as $(e_i,c_{i \bs i})$ 
+in our construction. For instance, one could form  $(e_i,c_{i \bs i})$ by using a
+Frank copula \citep{Nelsen2006} to unite two standard normal univariate distributions. 
+The bivariate normal distribution is a very natural choice, however, and other choices may 
+be somewhat esoteric. We note that choices were present in MCT prior to our work. For instance, 
+choices are often made in the definitions of $E$ and $C$ used \citep{Ellner_2019}. 
+Our overall research point, that ATAs
+\emph{can} be important for coexistence, is successfully made regardless of such choices, as long as 
+reasonable choices were made, which they were.
 

### 2023-07-17 paper/Paper.Rnw
plankton ATA stats in caption

@@
 res_loc <- "../results_numeric/"
+plankStats <- readRDS(file=paste0(res_loc, "plankStats.RDS"))
 #fig2 <- readRDS(file=paste0(res_loc,"fig2data.RDS"))
@@
 \emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
-in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (d) showing 
-left-tail association and (e) showing right-tail association.
+in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location: 
+(d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d,3)}, indicating left-tail association
+ and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association.
 (f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 

### 2023-07-17 code/Figure1.R

@@ -11,6 +11,7 @@ numRes_loc <- "../results_numeric/"
 noise_loc <- paste0(numRes_loc,"noise.RData") #from makenoise_LB.R
 popsim_loc <- paste0(numRes_loc,"betapopsim.RData") #from beta_popsims.R
 plankton_loc <- paste0(numRes_loc,"ceratium1x2.RData") #from plankton.R
+plankStats_loc <- paste0(numRes_loc, "plankStats.RDS")
 
 #locations where result figs are stored
 fig_loc <- "../results_figs/"
@@ -115,10 +116,11 @@ cerfus23 <- ceratium1x2$loc23$cerfus
 cerfur23 <- ceratium1x2$loc23$cerfur
 cerfus40 <- ceratium1x2$loc40$cerfus
 cerfur40 <- ceratium1x2$loc40$cerfur
-partCor(cerfus23, cerfur23, 1/2, "average")
-partCor(cerfus40, cerfur40, 1/2, "average")
-partCor(cerfus23, cerfur23, 1/3, "average")
-partCor(cerfus40, cerfur40, 1/3, 'average')
-cor(cerfus40, cerfur40, method="spearman")
+ceratiumPartCorD <- partCor(cerfus23, cerfur23, 1/2, "average")
+ceratiumPartCorE <- partCor(cerfus40, cerfur40, 1/2, "average")
+saveRDS(list(d=ceratiumPartCorD, d=ceratiumPartCorE),plankStats_loc)
+#partCor(cerfus23, cerfur23, 1/3, "average")
+#partCor(cerfus40, cerfur40, 1/3, 'average')
+#cor(cerfus40, cerfur40, method="spearman")
 
 

### 2023-07-17 paper/Paper.Rnw
added citation for stat

@@
 (d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d,3)}, indicating left-tail association
- and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association.
+ and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association \cite{Ghosh_2020_AER}.
 (f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 

### 2023-07-19 paper/RefereeResponses1.tex
continued addressing referee comments, about done with what I can do before meeting with Jasmin

@@
 is their opinion, and that data are insufficient for a clear conclusion to be reached at this point. We find that 
-such an acknowledgment is both rare from a referee and very useful for causing authors to most usefully revise their 
+such an acknowledgment is both rare from a referee and very useful for causing us, as authors, to most usefully revise our 
 paper. We have added new discussion material covering these topics and find that the new material improves the paper.
@@
 such as we have used, prior to examing it in field systems, in all their complexity, where niche differentiation
-by resources and predators can more readily occur. }
+by resources and predators can more readily occur. But once our paper is published we hope and expect that we and
+others will more broadly assess the potential importance of ATAs.}
 
@@
 important for coexistence in a system even while storage effects are not.
-Storage effects are ATA effects plus what we call the effects of $EC$ correlation \emph{per se} 
-($\Delta_i^{[E \nat C]} in our notation), and either of these two summands can be negative or positive.
-So if the twe effects are oppostite in sign and nearly equal, storage effects will be unimportant but ATA
-effects will be important for coexistence.}
+Storage effects are ATA effects plus what we called, in the manuscript, the effects of $EC$ correlation \emph{per se} 
+($\Delta_i^{[E \nat C]}$ in our notation). Either of these two summands can be negative or positive.
+So if the two effects are opposite in sign and similar in magnitude, storage effects will be unimportant for 
+coexistence, but ATA effects will be important.}
 
@@
 out to be broadly important in ecology. Our research goals were to introduce the idea of ATA effects on
-coexistence and to show that ATAs \text{may} be broadly important for coexistence, by showing examples where they
+coexistence and to show that ATAs \textbf{may} be broadly important for coexistence, by showing examples where they
 \textbf{are} important. The few examples we showed are all that are readily available at this time.
+We won't know for sure if ATAs (or storage effects) \textbf{are} important until more data can be brought to bear. 
 Future work can assess, broadly, how important ATAs are by looking across a larger number of systems, once
@@
 opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
-goals, so we thank the referee for not doing that.}
+goals, so we thank the referee for not doing that. We modified a few key points in the text to clarify 
+our viewpoint that we do not think our work definitively indicates that ATAs will certainly be a crucial
+mechanism of coexistence; instead we just think our work opens that possibility, for future exploration.
+For instance, we replaced the last sentence of the first paragraph of the Discussion, which used to be
+``Though future work should seek to understand precisely when ATAs are or are not important
+for coexistence, our work demonstrates the overall importance of this new mechanism'' with the revision
+``Future work should seek to understand precisely when ATAs are or are not important
+for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.''
+And we changed the last sentence of the abstract, which used to read ``ATA influences are an important new 
+mechanism of coexistence'' so that it now reads ``ATA influences may be an important new mechanism of coexistence.''
+We think these changes make the paper stronger, so thanks for spurring us to make them. The last
+sentence of the Introduction already read ``Overall, our study presents a new mechanism of species coexistence
+and a means of understanding its theoretical and empirical importance,'' which accurately reflects our belief
+that what we've accomplished is an indication that ATAs \textbf{may} be important for coexistence, 
+and a means of assessing
+that in future work, rather than a definitive conclusion that they are important for coexistence.}
 
-\emph{The new Discussion material is in a new SI section which is cited from the main text Discussion section. 
+\emph{The new Discussion material is in an SI section which is cited from the main text Discussion section. 
 We would have liked to put some of the new material into the main text. But length limits, together with
-clear requests from other referees to add a variety of content to the main text, made doing so impossible.}
-%DAN: Write the discussion material which is promised and the citation from the Discussion.
-
-% 
-% Note also that Chu and Adler (2015) say "we now believe that the models we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.
-For example, the temporal storage effect requires positive covariance between the effects of competition and the environment (Chesson and Huntly 1989): in years most favorable to a species, it should also be the most vulnerable to competition.
-However, we did not specify our vital rate regressions in a way that makes the covariation between the environment (or random year effects) and competition explicit.
-As a result, our model may not allow full expression of a temporal storage effect and the simulations may underestimate its strength.
-Such differences in model specification may explain why a previous analysis, based on a very different modeling approach, showed a strong storage effect in the Kansas community (Adler et al.
-2006) in contrast to the weak effect we observed here.
-We plan to address this model specification issue in future work focused on the population and community consequences of environmental variability."
-% 
-% Note that the Stump and Vasseur paper very clearly says that they don't think storage effects are important. 
-% It's a 73-page ms, I don't have time to absorb it. This will be part of acknowledging the viewpoint. Maybe Jasmin
-% has time to read it and decide how it relates to our paper? Do the assumptions they explore also apply
-% to the Ellner approach of quantifying storage effects?
-} \\
-%Dan to do
+clear requests from other referees to add a variety of content to the main text, made doing so impossible.} \\
+%DONE
 
@@
 \noindent ***\emph{Thank you for the positive feedback.} \\
+%DONE
 
@@
 now added such a reminder to the figure caption.} \\
+%DONE
 
@@
 is interested in our approach to the extent that they are generating potential ideas for future research.} \\
+%DONE
 
@@
 \noindent ***\emph{We modified the sentence.} \\
+%DONE
 
@@
 \noindent ***\emph{We changed the wording here.} \\
+%DONE
 
@@
 Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
-a bell for most ecologists.} \\
+a bell for most ecologists. It was not actually necessary to link to Taylor's law, we just though it 
+would provide an additional tie-in to another part of the literature. But that won't work if the referee is right 
+that most ecologists will become confused, so we were happy to drop that connection here.} \\
+%DONE
 
@@
 
-\noindent ***\emph{We spent 15-20 minutes or so trying to add an explanation that was succinct enough to not break
-the flow of the paper while still being informative, but we could not do it. Any more than a short sentence would break the 
-flow, and we felt that all the one-sentence explanations we could come up with were unclear. 
-FYI, in a nutshell the idea is as follows. Assume that low values of an environmental variable reduce the 
-growth rates of local populations and high values elevate those growth rates. In that case one may
-expect that left-tail associations increase metapopulation extinction risks because then environmental
-``catastrophes'' are spatially synchronized, presumably reducing the potential for rescue effects. Whereas
-``bonanzas'' (extremely good conditions, with a lot of growth) are much less correlated. The case where
-high values of the variable reduce the growth and low values elevate it is similar, but it's right-tail association
-that elevates extinction risk. This is most of the explanation, but it gets a bit more nuanced, still - 
-see the paper for additional details. Thanks for your interest, by the way.} \\
-%DAN: maybe add to the above to point out that they don't need to know this to understand what we're doing in this paper!
+\noindent ***\emph{It is not necessary for the reader to understand the mechanism we explored in that other
+paper in order to understand this current paper, we were just trying to tie the current work in to 
+another area of research in ecology, for the sake of contextualizing it a bit and motivating studies of ATAs generally. 
+So an explanation, here, is not 
+strictly necessary to the current paper. Nevertheless, we tried to satisfy the referee's request of adding a bit of explanation
+without disrupting the flow too much. Thanks for your interest, by the way, we encourage you to have a look at the
+other paper if you haven't already.} \\
+%DONE
 
@@
 defined in Theory, which they are.} \\
+%DONE
 
@@
 \noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
-paper possible, but, of the comments 
+paper possible, and we have used the comments to clarify. But, of the comments 
 C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
@@
 and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
-Furthermore, we don't think our paper is any more specialized than the papers Ellner et al (2016, 2019) which
+Furthermore, we don't think our paper is any more specialized or mathematical than the papers Ellner et al (2016, 2019) which
 we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
 specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
-central in ecology these days, and increasingly so. It makes no sense to us to sideline them into 
-"specialist journals", which few people read, especially when they are about mainstream topics such
+central in ecology these days, and increasingly so. So comments which would tend to sideline such paper
+"specialist journals", which fewer people read, seem outdated to us, especially when the papers in question 
+are about mainstream topics such
 as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
@@
 \noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\
+%DONE
 
@@
 example. That goal is convincingly 
-accomplished by the paragraph. Given that the math is simple, and also more precise and definitive than 
-qualitative verbal arguments that usually take up space in Introductions, we think this material is actually
+accomplished by the paragraph, more so than could be accomplished by a purely qualitative, words-only Introduction. 
+Given that the math is simple, and also more precise and definitive than 
+qualitative verbal arguments, we think this material is actually
 more beneficial to the readership and more effective for the intended motivational purpose than would be more words
-(and we see based on C25 below that the referee appears to agree with us here). 
+(and we see based on C25 below that the referee appears to agree with us here in some respects, at least). 
 Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
@@
 benefit the field by challenging assumptions and biases such as categorically not putting equations 
-in the Introduction.
-Overall, we think making this suggested change would make the paper weaker, and would pander to an
-outmoded and unhelpful tradition. So we will not make the change unless the 
-editor insists on it as a condition of publication.} \\
+in the Introduction.} \\
 %Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.
+%It's done except for any improvements or revisions that Jasmin may suggest.
 
@@
 above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
-%Are we being too petulant, should we just do it and say we did it?
+%DONE
 
@@
 Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
-without first stating the lottery model, since the result uses that model. } \\
+without first stating the lottery model, since the result uses that model. Math is more precise and can be
+more convincing than language, the trade off being that some readers won't understand some math. However, the 
+math we have in the Intro is so simple we think all readers of Ecology Letters will understand it, so there are
+benefits and basically no costs of including it there. } \\
+%DONE
 
@@
 To avoid making claims that we have not fully supported
-with rigorous mathematics or other formal methods, we rephrased to say the iid assumption is 
+with rigorous mathematics, we rephrased to say the iid assumption is 
 ``probably not necessary''. Thanks for spurring us to do this.} \\
+%DONE
 
@@
 in response to that part of this comment.} \\
+%DONE
 
@@
 of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
-After all, negative numbers are basic, and represent no real impediment to the average ecologist understanding the 
-paper. And we could instead spend that time working on the concepts underlying the next paper.} \\
-%Jasmin to consider (don't do yet) how hard/error prone this change would be, taking into account the 
-%formulas in the SI? Unless this is MUCH easier and MUCH less error prone than Dan's hunch, I suggest 
-%we don't do it.
+After all, negative numbers are basic, and represent no impediment to the average ecologist understanding the 
+paper.} \\
+%DONE
 
@@
 based on such a thing.}
+%DONE
 
@@
 
-\noindent ***\emph{There is already a line at $y=0$, you can see it, for instance, on panel (e). 
+\noindent ***\emph{There is already a line at $y=0$. You can see it, for instance, on panel (e). 
 It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
@@
 to coexistence, so it's important to show within-plot variation (between lines and across 
-values of the $x$-axis) clearly. We thank the referee for expressing their own uncertainty, which was
+values of the $x$-axis) clearly. We tried plotting with the same scales everywhere and it obscured some of the 
+patterns. We thank the referee for expressing their own uncertainty, which was
 useful to be aware of.} \\
+%DONE
 
@@
 (see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
-is usually one of the largest-magnitude terms. Our results as to the importance of ATAs for coexistence 
+is usually one of the largest-magnitude contributions to coexistence. Our results as to the importance of ATAs for coexistence 
 contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
-of section 4.1.} \\
+of section 4.1. That's one of the points we are making.} \\
+%DONE
 
@@
 such errors. But we're happy to hear specific suggestions.} \\
+%DONE
 
@@
 \noindent ***\emph{Many thanks for the positive feedback!} \\
+%DONE
 
@@
 
-\noindent ***\emph{<It cannot. ATAs can only influence storage effects, not the other mechanisms, as explained on 
-lines xyz. Is that enough? Elaborate and re-iterate stuff from the paper.>} \\
-%Dan to do
+\noindent ***\emph{This cannot occur. ATAs can only influence storage effects, not the other mechanisms, as stated at
+the top of section 2.4 and explained mathematically in the SI section cited there. Basically, as explained in 2.4, 
+this is because ATAs are about covariation between $E$ and $C$, whereas the mechanisms other than storage effects
+do not pertain at all to covariation. For instance, relative nonlinearity is about the effects of variation 
+in $E$ or $C$, not covariation between them.} \\
+%DONE
 
@@
 being suitable and unsuitable.>} \\
-%Jasmin to consider this while brainstorming the prior comment.
+%Jasmin to consider this while brainstorming the prior comment. Dan can help, or do this one, but 
+%I'll wait until after we've met and had a chance to discuss intutive explanations of ATA effects
+%that Jasmin has been working on.
 
@@
 %beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?
+%Let's talk more about this after Jasmin communicates to Dan the work she's done explaining the ATA effect
+%intuitively.
 

### 2023-07-19 paper/Paper.Rnw
continued addressing referee comments, about done with what I can do before meeting with Jasmin

--- 2023-07-17 paper/Paper.Rnw
@@
 difference between species coexistence and competitive exclusion. 
-ATA influences are an important new mechanism of coexistence. 
+ATA influences may be an important new mechanism of coexistence. 
 \end{abstract}
@@
 Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
-when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
+when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
+across component populations
+than are years of good environment, limiting the potential for rescue effects; 
+and \cite{Ghosh_2021} show that ATAs can alter the stability of 
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
@@
 (the beta fecundities lottery model), including our empirically
-based example (the diatom system). Though future work should seek to understand precisely when ATAs are or are not important
-for coexistence, our work demonstrates the overall importance of this new mechanism.
+based example (the diatom system). 
+Future work should seek to understand precisely when ATAs are or are not important
+for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
 
@@
 It may be possible to expand the mathematics of the original MCT to consider ATAs.
-Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:threeway}.
+Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:storage_importance}.
 

### 2023-07-19 paper/SupMat.Rnw
continued addressing referee comments, about done with what I can do before meeting with Jasmin

@@
 
+\section{   Diverse viewpoints on the potential general importance of storage effects and ATAs for coexistence}\label{sect:storage_importance}
+
+Some researchers have the opinion that storage effects are likely to be unimportant for 
+coexistence, generally, in field systems, though evidence is still insufficient to know for sure,
+and other researchers have the reverse opinion. Regardless of how that debate is eventually resolved, 
+ATAs can, \emph{a priori}, be important for coexistence even if storage effects are not.
+
+Opinions differ on the likely general importance of storage effects for coexistence, and 
+evidence so far is insufficient to clearly determine their importance.  
+\cite{Zapeda_2019} states that ``evidence for the relative importance of the different mechanisms 
+that promote species coexistence is lacking,'' though they went on to address this lack by showing that 
+niche differentiation and relative nonlinearity were more important for species coexistence 
+in a grassland system than were storage effects.   
+\cite{ChuAdler2015} argued that ``Niche differences and average fitness differences jointly determine 
+coexistence. However, little empirical information about the magnitude of these two mechanisms is available.''
+They then went on to show strong niche differences and modest average fitness differences in five 
+North American grassland and shrubland plant communities. During this process they found little 
+contribution of fluctuation-dependent stabilization mechanisms such as storage effects to coexistence,
+but they also stated that ``The temptation is to conclude that fluctuation-dependent mechanisms 
+are weak in these ecosystems (e.g., Adler et al. 2010). However, we now believe that the models 
+we fit may not be appropriate for rigorously quantifying fluctuation-dependent mechanisms.'' 
+In a very recent and thorough paper, \cite{StumpVasseur2023} state that they know of only four studies that
+used empirically quantified models to measure the relative importance of storage effects versus other 
+coexistence mechanisms \citep{Descamps_2005,Letten2018,Hallett_2019,Zapeda_2019}. \cite{StumpVasseur2023} 
+state that ``In all of these studies, the stabilizing contribution of the storage effect was on par 
+with the contribution of relative nonlinearity.'' \cite{StumpVasseur2023} interpret this to be
+evidence that storage effects are typically unimportant, because relative nonlinearity has been 
+thought to be weak -- and they spend most of their paper giving 
+a theoretical/conceptual explanation for the unimportance of storage effects. 
+But they also admit the possibility that relative nonlinearity is 
+stronger than has previously been appreciated, and that both storage effects and relative nonlinearity
+provide appreciable contributions to coexistence. \cite{StumpVasseur2023} also state that all studies
+that have measured both storage effect and variation-independent mechanisms have found that the latter
+overshadow the former, though there have only been two such studies \citep{Hallett_2019,Zapeda_2019}.
+
+Regardless of the general importance of storage effects for coexistence,  
+ATAs can, \emph{a priori}, be important for coexistence even if storage effects are not.  
+Storage effects are ATA effects plus what we have called the effects of $EC$ correlation \emph{per se} 
+($\Delta_i^{[E \nat C]}$ in our notation). Either of these two summands can be negative or positive.
+So if the two effects are opposite in sign and similar in magnitude, storage effects will be unimportant for 
+coexistence, but ATA effects will be important. 
+This is precisely what we observed in some of our lottery model results (e.g., the beta-fecundities model 
+with left-tail associations).
+It seems to us that only further research can 
+clearly resolve the question of the general
+relative importance of storage effects, ATA effects, and other mechanisms of coexistence.
+
 \clearpage

### 2023-07-19 paper/Paper.Rnw

@@
 \emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
-in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (d) showing 
-left-tail association and (e) showing right-tail association.
+in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location: 
+(d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d,3)}, indicating left-tail association
+ and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association \cite{Ghosh_2020_AER}.
 (f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 

### 2023-07-19 paper/Paper.Rnw
started a general effort to trim and make more accessible, but decided I might want to wait until after discussing Jasmin's recent explanatory efforts before proceding

@@
 extinction risk, ecosystem stability, and other phenomena. 
-We extend the MCT, decomposing storage effects to show the influence of ATAs.
+We extend MCT, decomposing storage effects to show the influence of ATAs.
 Analysis of a simple model
@@
 Ecologists may ask, for instance, how local environment relates to 
-population growth rate. To study such relationships and to understand how two variables interact, ecologists 
+population growth rate. To study such relationships, ecologists 
 very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
@@
 Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
-symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been referred to as 
+symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been called 
 ``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
 Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
-regression), but correlations are extremely common, and other common methods also reveal only a small portion 
+regression), but correlations are common, and other common methods also reveal only a small portion 
 of the information content of the relationship between variables \citep{Nelsen2006}.
@@
 across component populations
-than are years of good environment, limiting the potential for rescue effects; 
-and \cite{Ghosh_2021} show that ATAs can alter the stability of 
+than are years of good environment, limiting the potential for rescue effects. 
+And \cite{Ghosh_2021} show that ATAs can alter the stability of 
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
-plankton population density time series; and many other examples have also been previously explored 
+plankton population density time series; many other examples have also been previously explored 
 \citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
-competition and coexistence have not yet been investigated. Doing so is the overarching goal of this study.
+competition and coexistence have not been investigated. Doing so is the goal of this study.
 
@@
 following specific goals. 
-G1) We will augment MCT to show formally how ATAs play a role in coexistence. 
+G1) We augment MCT to show formally how ATAs play a role in coexistence. 
 MCT decomposes a measure of coexistence into components due to each of several mechanisms. 
@@
 spatial variation can be considered in future work.
-G2) We will explore, using the simple lottery model, the relative magnitudes 
+G2) We explore, using the simple lottery model, the relative magnitudes 
 of the contributions to coexistence that come
@@
 are likely to be negligible or meaningfully large. 
-G3) We will provide an empirical example of species coexistence which highlights the role ATAs can play and the
+G3) We provide an empirical example of species coexistence which highlights the role ATAs can play and the
 potential importance of ATAs relative to other mechanisms of coexistence.
@@
 be able to grow when rare. We also defined scenarios of ``ATA rescue''
-and ``ATA exclusion''. \emph{ATA rescue} was considered to have occurred for a given scenario if coexistence
-was possible when ATAs were taken into account, i.e., GWR$>0$ for both species, but impossible otherwise, i.e.,
-the sum of the non-ATA terms in (\ref{eq:full_decomp}) was negative for at least one species. \emph{ATA exclusion} was 
-considered to have occurred if the species did not coexist when ATAs
-were taken into account, i.e., GWR$<0$ for at least one species, but would have coexisted were it not for ATAs, 
+and ``ATA exclusion''. \emph{ATA rescue} is considered to have occurred for a given scenario if coexistence
+is possible when ATAs are taken into account, i.e., GWR$>0$ for both species, but impossible otherwise, i.e.,
+the sum of the non-ATA terms in (\ref{eq:full_decomp}) is negative for at least one species. \emph{ATA exclusion} is 
+considered to have occurred if the species do not coexist when ATAs
+are taken into account, i.e., GWR$<0$ for at least one species, but would coexist were it not for ATAs, 
 i.e., the sum of the non-ATA terms in (\ref{eq:full_decomp})
-was positive for both species.
+is positive for both species.
 
@@
 the decomposition (\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
-of the lottery model which we now specify. The versions both used the same dynamical equations
-(Introduction) and assumed temporally independent and identically distributed (iid) fecundities, but the
+of the lottery model. The versions both used the same dynamical equations
+(Introduction) and assumed temporally iid fecundities, but the
 versions differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
@@
 Our approach to exploring the influence of ATAs on coexistence relies on the computational
-extension of the MCT of \cite{Ellner_2019},
+extension of MCT of \cite{Ellner_2019},
 and our results help demonstrate strengths of that approach.

### 2023-07-20 paper/RefereeResponses1.tex
responding to C27 and C21

@@
 \newcommand{\bs}{\backslash}
+\newcommand{\olep}{\overline{\epsilon}}
 
@@
 \noindent ***\emph{We replaced the reference.} \\
-%Jasmin can you please do this?
+%Jasmin can you please do this?   %%REPLACED
 
@@
 %It's done except for any improvements or revisions that Jasmin may suggest.
+%Jasmin: I really like this response and don't think its too harsh or needs to be toned down.
 
@@
 Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
-the $\#$ term because that terms involves both of them varying (though independently).} \\
+the $\#$ term because that terms involves both of them varying (though independently).
+
+Also, see S3 "Mechanisms other than storage effect" of our Supporting information. 
+%can delete if not needed
+Since,
+\begin{align}
+\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
+&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
+\end{align}
+and 
+\begin{equation}
+\olep_i^{(E\shp C)}=\E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0
+&= \olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
+\end{equation}
+then,
+\begin{align}
+\olep_i^{(EC)} = \olep_i^{EC}-\olep_i^{(E\shp C)}
+&= (\olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) - (\olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0)
+&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0 - \olrs_{i \bs i} + \olep_i^E + \olep_i^C + \epsilon_i^0
+&= \olr_{i \bs i} - \olrs_{i \bs i}. 
+\end{align}
+So,
+\begin{align}
+\Delta_i^{(EC)}=\olep_i^{(EC)} - q_{ij}\olep_j^{(EC)}
+&= (\olr_{i \bs i} - \olrs_{i \bs i}) - q_{ij}(\olr_{j \bs i} - \olrs_{j \bs i})
+&= \olr_{i \bs i} - \olrs_{i \bs i} - q_{ij}\olr_{j \bs i} + q_{ij}\olrs_{j \bs i}
+&= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i}
+&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}).
+\end{align}
+} \\
 %Jasmin to check and make sure Dan is right , here. Maybe there is a better place to refer them to to support our claim?

### 2023-07-20 paper/Paper.Rnw
responding to C27 and C21

@@
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
-tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Narawani_2013}, and corals \citep{Alvarez-Noriega_2020}. 
+tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Letten2018}, and corals \citep{Alvarez-Noriega_2020}. 
+%Narawani_2013
 Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.

### 2023-07-21 paper/Paper.Rnw
small change for fig1 captoin

@@
 in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location: 
-(d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d,3)}, indicating left-tail association
- and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association \cite{Ghosh_2020_AER}.
+(d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d$diff,3)}, indicating left-tail association
+ and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e$diff,3)}, indicating right-tail association \cite{Ghosh_2020_AER}.
 (f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 

### 2023-07-21 code/Figure1.R

@@ -46,8 +46,8 @@ cPCD <- partCor(cerfus23, cerfur23, 1/2, "average")
 #ceratium partial correlation for panel E
 cPCE <- partCor(cerfus40, cerfur40, 1/2, "average")
 plankStats <- list(d=cPCD, e=cPCE)
-plankStatsATA <- list(d=cPCD$diff, e=cPCE$diff)
-saveRDS(plankStatsATA ,plankStats_loc)
+#plankStatsATA <- list(d=cPCD$diff, e=cPCE$diff)
+saveRDS(plankStats ,plankStats_loc)
 #partCor(cerfus23, cerfur23, 1/3, "average")
 #partCor(cerfus40, cerfur40, 1/3, 'average')
 #cor(cerfus40, cerfur40, method="spearman")

### 2023-07-24 paper/RefereeResponses1.tex
small thing
made changes to various files to address the referee desire to know that Fig 1d,e are significantly asymmetric

@@
 as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
-shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we take
-seriously the referee's comment that they are are not convinced there are ATAs in Fig. 1d-e, and ...
-%Jasmin, let's discuss what to do here. We could work back to what specific result this is in 
-%the original paper and extract from p-value info, or we could come up with a better example,
-%or something else, not sure, let's discuss. Maybe also augment the figure to make it clearer?
-%How about if Jasmin: 1) adds statistical support to caption; 2) adds to the figure some means of
-%hopefully drawing people's attention to the right features of this thing.
-} \\
+shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we took
+seriously the referee's comment that they were apparently not convinced there are ATAs in Fig. 1d-e.
+In fact, one can significance test these results and they are solidly significant - we added
+those results, along with the methods used to get them, to a new SI section which is now cited from the
+figure caption. We also added some text to the figure caption which we hope will focus the 
+readers attention on the aspects of these panels which clearly show asymmetries of tail association. } \\
 
@@
 %sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
-%area in ecology?
+%area in ecology? %food webs? networks
 

### 2023-07-24 paper/Paper.Rnw
made changes to various files to address the referee desire to know that Fig 1d,e are significantly asymmetric

--- 2023-07-20 paper/Paper.Rnw
@@
 res_loc <- "../results_numeric/"
-plankStats <- readRDS(file=paste0(res_loc, "plankStats.RDS"))
+
 #fig2 <- readRDS(file=paste0(res_loc,"fig2data.RDS"))
@@
 \emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
-in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location: 
-(d) has a partial Spearman correlation of  \Sexpr{round(plankStats$d,3)}, indicating left-tail association
- and (e) has a partial Spearman correlation of  \Sexpr{round(plankStats$e,3)}, indicating right-tail association \cite{Ghosh_2020_AER}.
+in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs were statistically
+significant and differed by location, with associations being primarily in the left tails in 
+panel d, and primarily in the right tails in panel e (SI section \ref{SI-SIsect:planktonstats}). 
+One can see these asymmetries by noting that the open points in panel d, which are in the left halves of the
+distributions, are more tightly clustered around the line $y=x$ than are the closed points,
+which are in the right halves of the distributions; and \emph{vice versa} for panel e.
 (f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 

### 2023-07-24 paper/SupMat.Rnw
made changes to various files to address the referee desire to know that Fig 1d,e are significantly asymmetric

@@
 \newcommand{\olC}{\overline{C}}
+\newcommand{\cor}{\text{cor}}
 
@@
 #place R code here for loading in necessary variables from Results
-#ordinarily this would be a readRDS command, but I'm just defining a variable here for now
 #allregres<-matrix(rnorm(10),2,5)
+
 M <- readRDS(file="../results_numeric/M.RDS")
@@
 se4qij <- readRDS(file="../results_numeric/fig4qijmaxse.RDS")
+
+res_loc<-"../results_numeric/"
+panD_stat <- readRDS(file=paste0(res_loc, "DiffPartialSpearmanPanelD.Rds"))
+panE_stat <- readRDS(file=paste0(res_loc, "DiffPartialSpearmanPanelE.Rds"))
+panD_pval <- readRDS(file=paste0(res_loc, "DiffPartialSpearman_pval_PanelD.Rds"))
+panE_pval <- readRDS(file=paste0(res_loc, "DiffPartialSpearman_pval_PanelE.Rds"))
+panD_parm <- readRDS(file=paste0(res_loc, "DiffPartialSpearman_parm_PanelD.Rds"))
+panE_parm <- readRDS(file=paste0(res_loc, "DiffPartialSpearman_parm_PanelE.Rds"))
+nsims<-readRDS(file=paste0(res_loc,"DiffPartialSpearmannsims.Rds"))
 @
@@
 \listoffigures
+
+\section{Statistical significance of the plankton example of ATAs of main text Fig. \ref{MT-fig:pedagog1}}\label{SIsect:planktonstats}
+
+For scatterplot data showing positive associations, such as those pictured in Fig. \ref{MT-fig:pedagog1}d-e, 
+\cite{Ghosh_2020_AER} defined an index which takes positive values when association
+is primarily in the left halves of the distributions, and negative values when association
+is primarily in the right halves of the distributions. See below for a restatement of the definition
+of the index of \cite{Ghosh_2020_AER}. Values of the index were \Sexpr{round(panD_stat,3)} for panel
+d, and \Sexpr{round(panE_stat,3)} for panel e. These values were both statistically different from 
+$0$ according to one-tailed tests based on an appropriate null hypothesis, $p=$\Sexpr{round(panD_pval,4)}
+and $p=$\Sexpr{round(panE_pval,4)}, respectively. See below for details of the test.
+
+Given scatterplot data, $(x_i,y_i)$ for $i=1,\ldots,N$, \cite{Ghosh_2020_AER} defined the 
+\emph{normalized rank} $u_i$ of $x_i$ as the rank of $x_i$ in the set 
+$\{x_1,\ldots,x_N\}$ divided by $N+1$, where the smallest value is taken to have rank $1$. 
+The normalized ranks $v_i$ were defined for $y_i$ in a likewise manner. 
+Then, the \emph{partial
+Spearman correlation} for the left halves of the distributions, $\cor_{\text{left}}$, was defined as 
+\begin{equation}
+\frac{\sum(u_i-\mean(u))(v_i-\mean{v})}{(N-1)\sqrt{\var(u)\var(v)}},
+\end{equation}
+where the sum is over all indices $i$ for which $u_i+v_i<1$ (though the means and variances
+in the above formula are computed over all the data). The indices satisfying $u_i+v_i<1$
+are those corresponding to points falling under the dashed diagonal lines in Fig. \ref{MT-fig:pedagog1}d-e.
+%***DAN: Note I mentioned a feature which Jasmin has not yet added, so make sure she 
+%adds it to those panels.
+The partial Spearman correlation for the right halves of the distributions,
+$\cor_{\text{right}}$, was defined using the same formula but where the sum is 
+now over all indices $i$ for which $u_i+v_i>1$. These are the points falling above the
+dashed diagonal lines in Fig. \ref{MT-fig:pedagog1}d-e. The left and right partial Spearman
+correlations typically add up to the Spearman correlation. The index 
+$\cor_{\text{left}}-\cor_{\text{right}}$ is the index used by \cite{Ghosh_2020_AER}
+mentioned in the previous paragraph.
+
+To calculate whether the index of \cite{Ghosh_2020_AER} was significantly positive for 
+Fig. \ref{MT-fig:pedagog1}d,
+we compared the index value computed for the data of that panel to values computed for an appropriate 
+symmetric distribution. We used the function `iRho' in the R package `copula'
+to determine that the Spearman correlation of a bivariate normal distribution with
+standard normal marginals and covariance parameter \Sexpr{round(panD_parm,4)}
+is the same as the Spearman correlation of the data in Fig. \ref{MT-fig:pedagog1}d.
+We simulated \Sexpr{nsims} data sets from such a distribution, each of size the same as 
+that of the data set pictured in Fig. \ref{MT-fig:pedagog1}d. For each simulated data set,
+we computed the index $\cor_{\text{left}}-\cor_{\text{right}}$ of \cite{Ghosh_2020_AER}.
+The $p$-value quoted above is the fraction of values so obtained that were 
+greater than the value $\cor_{\text{left}}-\cor_{\text{right}}$=\Sexpr{round(panD_stat,3)}
+obtained for the real data. A similar approach was used for Fig. \ref{MT-fig:pedagog1}e.
 

### 2023-07-24 code/MASTER.R

@@ -53,5 +53,6 @@ source("./diatom/fig5NEW.R")
 #VKQ figure (SI), shows assumptions of how V, K and Q depend on temperature 
 source("./diatom/VKQplots.R")
 
-
+#Stats the referee asked us to add regarding Fig 1d, e
+source("Fig1de_stats.R")
 

### 2023-07-24 code/partCor.R

@@ -1,0 +1,12 @@
+#Compute the partial Spearman correlation in the left halves and right halves 
+#of the distributions, and the difference. Based on the Ghosh et al work.
+
+partCor <- function(x,y, b, ties){
+  Tot <- length(x)
+  u <- rank(x, ties=ties)/(Tot+1); v <- rank(y, ties=ties)/(Tot+1)
+  left <- (u+v) > 0 & (u+v) < b*2
+  right <- (u+v) > b*2 & (u+v) < 2 
+  corl <- sum((u[left]-mean(u))*(v[left]-mean(v)))/((Tot-1)*sqrt(var(u)*var(v)))
+  corr <- sum((u[right]-mean(u))*(v[right]-mean(v)))/((Tot-1)*sqrt(var(u)*var(v)))	
+  return(list(left=corl, right=corr, diff=corl-corr, Sh=corl+corr, S=cor(u,v, method="spearman"), uv=cbind(u,v), bounds=cbind(left, right)))
+}

### 2023-07-26 paper/RefereeResponses1.tex
trivial
more changes to responses

@@
 Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
-the $\#$ term because that terms involves both of them varying (though independently).
+the $\#$ term because that terms involves both of them varying (though independently).}
 
-Also, see S3 "Mechanisms other than storage effect" of our Supporting information. 
+\emph{Also, see our SI section ``Mechanisms other than storage effects". 
 %can delete if not needed
@@
 and 
-\begin{equation}
-\olep_i^{(E\shp C)}=\E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0
+\begin{align}
+\olep_i^{(E\shp C)} &= \E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
 &= \olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
-\end{equation}
-then,
+\end{align}
+we have
 \begin{align}
-\olep_i^{(EC)} = \olep_i^{EC}-\olep_i^{(E\shp C)}
-&= (\olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) - (\olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0)
-&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0 - \olrs_{i \bs i} + \olep_i^E + \olep_i^C + \epsilon_i^0
+\olep_i^{(EC)} &= \olep_i^{EC}-\olep_i^{(E\shp C)} \\
+&= (\olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) - (\olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) \\
+&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0 - \olrs_{i \bs i} + \olep_i^E + \olep_i^C + \epsilon_i^0 \\
 &= \olr_{i \bs i} - \olrs_{i \bs i}. 
@@
 \begin{align}
-\Delta_i^{(EC)}=\olep_i^{(EC)} - q_{ij}\olep_j^{(EC)}
-&= (\olr_{i \bs i} - \olrs_{i \bs i}) - q_{ij}(\olr_{j \bs i} - \olrs_{j \bs i})
-&= \olr_{i \bs i} - \olrs_{i \bs i} - q_{ij}\olr_{j \bs i} + q_{ij}\olrs_{j \bs i}
-&= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i}
-&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}).
+\Delta_i^{(EC)} &= \olep_i^{(EC)} - q_{ij}\olep_j^{(EC)} \\
+&= (\olr_{i \bs i} - \olrs_{i \bs i}) - q_{ij}(\olr_{j \bs i} - \olrs_{j \bs i}) \\
+&= \olr_{i \bs i} - \olrs_{i \bs i} - q_{ij}\olr_{j \bs i} + q_{ij}\olrs_{j \bs i} \\
+&= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i} \\
+&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}),.
 \end{align} 
-} \\
-%Jasmin to check and make sure Dan is right , here. Maybe there is a better place to refer them to to support our claim?
+which is the equation the referee is asking about.} \\
+%DONE
 
@@
 negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
-work (and potential for error) 
+work, and potential for error, 
 of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
@@
 based on such a thing.}
-%DONE
 
@@
 readers attention on the aspects of these panels which clearly show asymmetries of tail association. } \\
+%DONE
 

### 2023-07-27 paper/ExplanationOfATAEffectsWorkingDoc.tex
added what I think is a good explanation of the intuition of ATAs, intended for the Discussion, and currently working on SI support of that material

@@
 about the "central line" and be more densely situated about the "central line".  
+%is central line like niche overlap line? species association line?  
 When the invader's EC relationship has more spread (more density in values far away from 
@@
 
+\emph{Pollinator competition}.
+When environmental conditions are bad, specialized pollinators are less abundant.
+Two plant species may have to depend more on the generalist species, thus niches become less different.
+(Niche overlap in bad environment leading to left tailed ATA)
+
+\subsection{Dan thoughts 2023 07 26}
+
+% %Attempt 1 - not as good, Dan thinks. May be slightly misleading. See below for attempt 2.
+% %
+% %I have in mind that the following paragraph will go i the Discussion, with maybe a sentence
+% %or two in the Intro as a preview of coming attractions.
+% Just as for storage effects, temporal niche differentiation is necessary in order for ATA
+% effects to materialize. This makes intuitive sense because temporal niche differentiation is 
+% also a prerequistite of storage effects [\cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}],
+% and ATA effects are a part of storage effects, but see
+% SI section X for a detailed explanation. 
+% %Topic sentence follows. Possibly we can state something like this in the Intro and say
+% %we explain it in the Discussion after formalizing the extent to which ATAs are important.
+% If temporal niche differentiation occurs, ATA effects essentially capture the influence of asymmetries 
+% in the extent to which storage effects can occur in environmentally good versus bad years,
+% as we here explain.
+% %
+% These asymmetries can make a big difference to total storage effects and coexistence,
+% as some of our analyses showed.
+% %
+% If temporal niche differentiation occurs
+% but is partial (as will be typical), the quantities $E$ and $C$ experienced by a currently scarce 
+% species (i.e., $E_i$ and $C_{i \bs i})$ will be positively but imperfectly correlated. 
+% %
+% Storage effects can occur if some buffering mechanism reduces losses in bad years so that
+% gains in good years can more than make up for them. 
+% %
+% Bad years happen whenever the negative effects of $C$ outweigh any 
+% positive effects of $E$; but this can occur
+% both when 1) $E$ is moderately favorable but $C$ is very strong, as well as when 2) $E$ is very unfavorable,
+% or moderately unfavorable and $C$ is not too weak.
+% Conversely, good years happen whenever positive effects of $E$ outweigh negative effects of $C$;
+% but this can occur when 1) $E$ is very favorable, or moderately favorable and $C$ is not too strong,
+% as well as when 2) $E$ is moderately unfavorable but $C$ is very weak. 
+% %
+% ATA effects essentially quantify the asymmetry of storage that occurs when 
+% $E$ and $C$ are both higher than average (the cases 1, above); versus storage that occurs 
+% when $E$ and $C$ are
+% both lower than average (the cases 2, above). 
+% %
+% The distinction is an important one because it corresponds to temporal niche differentiation
+% which occurs under two very different sets of environmental conditions, possibly via
+% different ecological or physiological mechanisms.
+% %
+% Details, applications, and examples of this intuition 
+% are in SI section X, and these ideas are also used there to provide 
+% intuitive explanations for the relative differences in the importance
+% of ATA effects between our log-normal and beta fecundities lottery models.
+% %DAN: You ought to also mention explaining differences with the diatom system,
+% %if you figure out an explanation.
+
+%Attempt 2 below - better than the above, but may need to be simplified a bit more,
+%if possible
+%
+Just as for storage effects, temporal niche differentiation is necessary for ATA
+effects. This makes intuitive sense because temporal niche differentiation is 
+also a prerequistite of storage effects [\cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}],
+and ATA effects are a part of storage effects, but see
+SI section X for a detailed explanation. 
+%
+Assuming that temporal niche differentiation is present, how do ATA effects then happen?
+We here offer an intuitive explanation. 
+%
+If temporal niche differentiation occurs
+but is partial, the $E$ and $C$ experienced by a currently scarce 
+species (i.e., $E_i$ and $C_{i \bs i})$ will be positively but imperfectly associated, 
+and the association may be asymmetric. 
+%
+Storage effects can occur if some buffering mechanism reduces losses in bad years so that
+gains in good years can more than make up for them. 
+%
+Bad years happen when the negative effects of $C$ outweigh any 
+positive effects of $E$; but this can occur
+both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is very unfavorable,
+or moderately unfavorable and $C$ is not too weak.
+Conversely, good years happen when positive effects of $E$ outweigh negative effects of $C$;
+but this can occur 1) when $E$ is very favorable, or moderately favorable and $C$ is not too strong,
+as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
+%
+It turns out that the potential for storage to occur when 
+$E$ and $C$ are both higher than average (the cases 1, above); versus the potential  
+when $E$ and $C$ are
+both lower than average (the cases 2) are not necessarily equivalent.
+Depending on the growth rate function, $r(E,C)$, storage potential may differ
+across the spectrum from 1 to 2.
+%
+ATA effects are the extent to which asymmetries of association between $E$ and 
+$C$ are aligned with the potential for storage across the dichotomy 1 versus 2.
+A comparison is made to a benchmark of symmetric association between $E$ and $C$.
+%
+ATA effects therefore help quantify different degrees of temporal niche differentiation 
+occuring under different sets of environmental conditions (such as cases 1 and 2), possibly via
+different ecological or physiological mechanisms.
+%
+Details, applications, and examples of this intuition 
+are in SI section X, and these ideas are also used there to provide 
+intuitive explanations for the relative differences in the importance
+of ATA effects between our log-normal and beta fecundities lottery models.
+%DAN: You ought to also mention explaining differences with the diatom system,
+%if you figure out an explanation.
+
+\subsection{SI section X corresponding to the thoughts above}
+
+Just as for storage effects [see \cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}], 
+temporal niche differentiation is necessary in order for ATA
+effects to materialize; we here explain why in detail. ``Lack of temporal niche differentiation'' can be interpreted
+to mean that species respond in identical ways to environmental fluctuations. 
+In a two-species model such as those we consider, this should mean a monotonically
+increasing relationship between $E_i$ and $C_{i \bs i}$. The former is the environment
+as perceived by species $i$. The latter is competition that a scarce invader $i$ experiences
+in a monoculture of $j$, and that competition comes only from $j$ (because $i$ itself
+is scarce), so should correspond to the environmental response of $j$. If 
+$E_i$ and $C_{i \bs i}$ are monotonically related, then $(E_i,C_{i \bs i})$
+and $(E_i^\nat,C_{i \bs i}^\nat)$ are distributed in the same way, so 
+$\E[r_i(E_i,c_{i \bs i})]-\E[r_i(E_i^\nat,c_{i \bs i}^\nat)]=0$.
+Therefore, 
+\begin{align}
+\Delta_i^{[EC]} &= (\olr_{i \bs i}-\olr_{j \bs i})-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
+&= \olr_{i \bs i}-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
+&= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)]+\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] \label{eq:expri}\\
+&= \E[r_j(E_j^\nat,C_{j \bs i}^\nat)].
+\end{align}
+But we already know $\E[r_j(E_j^\nat,C_{j \bs i}^\nat)]=0$
+because there is no niche differentiation between $j$ and itself, so 
+$E_i$ and $C_{j \bs i}$ should be monotonically related. So 
+$\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] = \E[r_j(E_j,C_{j \bs i})]=0$, giving
+$\Delta_i^{[EC]}=0$. So temporal
+niche differentiation is necessary for ATA effects on coexistence to 
+materialize.
+
+If temporal niche differentiation occurs but is partial, 
+then $E_i$ and $C_{i \bs i}$ will be positively but imperfectly correlated with
+each other, corresponding to similar but non-identical responses of species to
+environmental fluctuation; but how do ATA effects then happen? We here extend the intuition
+introduced in the Discussion and add detail.
+
+Storage can occur if population
+increases during good years outweigh decreases during bad years. For instance,
+if $r_i(E_i,C_{i \bs i})$ is very large when the balance of $E_i$ and 
+$C_{i \bs i}$ is favorable, but only moderately small when this balance is
+unfavorable, the expected value $\E[r_i(E_i,C_{i \bs i})]$ in (\ref{eq:expri})
+will be positive. But the potential for such storage of population growth
+may differ between when $E_i$ and $C_{i \bs i}$ are both high compared
+to when they are both low. 
+For instance, for the beta fecundities lottery model, $E_i/C_{i \bs i}$, a quantity
+which features prominently in $\E[r_i(E_i,C_{i \bs i})]=\log(1-\delta+E_i/C_{i \bs i})$,
+takes values bounded above by $\frac{2\eta_i \delta}{\eta_j}$ 
+and bounded below by $\frac{\eta_i \delta}{2 \eta_j}$ when $E_i$ and $C_{i \bs i}$
+are both in the upper halves of their assumed beta distribtions (recall that 
+these quantities were assumed to be distributed, respectively, as $\eta_1 \beta$
+and $\eta_2 \beta/\delta$ for $\beta$ a beta-distributed random variable with parameters 
+$0.5$ and $0.5$). These bounds are separated only by a factor of four. 
+On the other hand, when $E_i$ and $C_{i \bs i}$ are both in the lower halves 
+of their distributions, $E_i/C_{i \bs i}$ can take any positive value because
+then either $E_i$ or $C_{i \bs i}$ can take values arbitrarily close to $0$. 
+Because values of $E_i/C_{i \bs i}$ have the potential to be much larger when 
+$E_i$ and $C_{i \bs i}$ are both in the lower halves of their distributions
+than when these quantities are both in their upper halves, and because
+the growth rate, $\log(1-\delta+E_i/C_{i \bs i})$, buffers low values of 
+$E_i/C_{i \bs i}$ (the infemum of the growth rate is $\log(1-\delta)$, whereas the supremum is
+infinite), the storage effects contributed by the left tails of 
+$E_i$ and $C_{i \bs i}$ have the potential to be much greater than those contributed
+by the right tails. Essentially, values of $\log(1-\delta+E_i/C_{i \bs i})$
+occurring when $E_i$ and $C_{i \bs i}$ are both in the upper halves of their
+distributions can never exceed $\log(1-\delta+\frac{2\eta_i \delta}{\eta_j})$,
+whereas values ocurring when $E_i$ and $C_{i \bs i}$ are both in the lower
+halves of their distributions can be arbitrarily large (though with decreasing 
+probability for larger values).
+
+\bibliographystyle{ecology_letters2}
+\bibliography{refs}
+
 \end{document}

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
ongoing work on a set of intuitive explanations and examples for ATA effects

@@
 
-\subsection{Dan thoughts 2023 07 26}
+\subsection{Dan thoughts 2023 07 26 - planned for inclusion in the Discussion}
 
@@
 and ATA effects are a part of storage effects, but see
-SI section X for a detailed explanation. 
+SI section X1 for a detailed explanation. 
 %
@@
 $C$ are aligned with the potential for storage across the dichotomy 1 versus 2.
+If $E$ and $C$ are strongly associated in the parts of their distribution where
+storage potential is high, ATA effects will be relatively low, and vice versa.
 A comparison is made to a benchmark of symmetric association between $E$ and $C$.
@@
 Details, applications, and examples of this intuition 
-are in SI section X, and these ideas are also used there to provide 
+are in SI section X2, and these ideas are also used there to provide 
 intuitive explanations for the relative differences in the importance
@@
 
-\subsection{SI section X corresponding to the thoughts above}
+\subsection{SI section X1 corresponding to the thoughts above, call it Temporal niche differentiation is necessary for ATAs}\label{SIsect:temporal_niche_diff}
 
@@
 temporal niche differentiation is necessary in order for ATA
-effects to materialize; we here explain why in detail. ``Lack of temporal niche differentiation'' can be interpreted
+effects to materialize; we here explain why, in detail. ``Lack of temporal niche differentiation'' can be interpreted
 to mean that species respond in identical ways to environmental fluctuations. 
@@
 because there is no niche differentiation between $j$ and itself, so 
-$E_i$ and $C_{j \bs i}$ should be monotonically related. So 
+$E_j$ and $C_{j \bs i}$ should be monotonically related. So 
 $\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] = \E[r_j(E_j,C_{j \bs i})]=0$, giving
 $\Delta_i^{[EC]}=0$. So temporal
-niche differentiation is necessary for ATA effects on coexistence to 
+niche differentiation, interpreted as Spearman correlation between 
+$E$ and $C$ less than $1$, is necessary for ATA effects on coexistence to 
 materialize.
+
+\subsection{SI section X2 corresponding to the thoughts above, call it Intuition and 
+examples for ATAs}\label{SIsect:intuition_examples}
 
@@
 each other, corresponding to similar but non-identical responses of species to
-environmental fluctuation; but how do ATA effects then happen? We here extend the intuition
-introduced in the Discussion and add detail.
+environmental fluctuation (see SI section \ref{SIsect:temporal_niche_diff}); 
+but how do ATA effects then happen? We here extend the intuition
+introduced in the Discussion and provide examples.
 
@@
 For instance, for the beta fecundities lottery model, $E_i/C_{i \bs i}$, a quantity
-which features prominently in $\E[r_i(E_i,C_{i \bs i})]=\log(1-\delta+E_i/C_{i \bs i})$,
+which features prominently in $r_i(E_i,C_{i \bs i})=\log(1-\delta+E_i/C_{i \bs i})$,
 takes values bounded above by $\frac{2\eta_i \delta}{\eta_j}$ 
@@
 halves of their distributions can be arbitrarily large (though with decreasing 
-probability for larger values).
+probability for larger values). Apparently, the years with the most potential to 
+contribute to storage
+effects for the beta-fecundities lottery model are those for which the environment
+is actually poorer than average, because, for those years, competition can be very low,
+so that the ratio $E/C$ is large. Years for which the
+environment is exceptionally good actually cannot contribute much to storage effects for 
+this model, because for those years, $E/C$ cannot be large.
+It is the imbalance between $E$ and $C$ that allows for storage effects. 
+Counterintuitively, for this model, that imbalance is greatest when the environment is 
+poor. This result is counterintuitive because one may be used to thinking of storage effects
+occuring when reproductive pulses from environmentally good years sustain the population
+through environmentally poor years. But, in fact, good years for population growth may
+happen principally when the environment is poor, if those are also the times when 
+competition is very low, as for this model.
+
+Fig. \ref{fig:beta_fecundidities_explanatory} provides quantitative backing to
+the ideas of the prior paragraph. The growth function $\ln(1-\delta+E/C)$ that pertains 
+to the lottery model
+is plotted with colors and contours on Fig. \ref{fig:beta_fecundidities_explanatory}A-C.
+Samples were taken from the distribution $(E_i,C_{i \bs i})$ for the beta 
+fecundities lottery model for left- and right-tailed 
+associations between $E_i$ and $C_{i \bs i}$, and are plotted as
+points on Fig. \ref{fig:beta_fecundidities_explanatory}A and C, respectively. 
+Samples were also taken from the distribution $(E_i^\nat,C_{i \bs i}^\nat)$
+(which is the same regardless of whether the original $(E_i,C_{i \bs i})$
+distribution showed left- or right-tail association) and are plotted on 
+Fig. \ref{fig:beta_fecundidities_explanatory}B. 
+The dashed lines labelled $L_2$ on A-C separate the upper-right $50\%$
+of each distribution from the lower-left $50\%$. The dashed lines labeled
+$L_1$ separate the upper-left $50\%$ from the lower-right $50\%$.
+The quantity $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ can be
+well approximated for left- and right-tail associations by averaging function 
+values, depicted with color and contours, over the points on panels A and C, respectively. The quantity
+$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
+can be well approximated by averaging function 
+values over the points on B. So ATA effects for the model with left-tail association
+between $E_i$ and $C_{i \bs i}$ (respectively, right-tail association) are given by 
+computing the average function value over points on panel A (respectively, C)
+and then subtracting the average function value over points on panel B.
+Contributions to the storage effect of points in the left portions of 
+the $E,C$ distribution (the region labeled ``left'' on panels A and C)
+can be accurately estimated by averaging function values across
+points in that region; likewise storage contributions of points in the 
+``right'' region can be estimated by averaging function values over 
+points in that region.
+
+The details of the averaging process described above are pictured in 
+Fig. \ref{fig:beta_fecundidities_explanatory}, panels D-F;
+we now describe how. 
+Signed distances of points on Fig. \ref{fig:beta_fecundidities_explanatory} 
+A-C from the lines $L_1$ correspond to 
+the quantities $\log_{10}(E_1/C_{1 \bs 1})$ on A, C and to 
+$\log_{10}(E_1^\nat/C_{1 \bs 1}^\nat)$ on B. The distributions of these 
+quantities across points to the left of $L_2$ are pictured 
+in the lower horizontal-axis historgrams on panels D-F, respectively;
+and distributions for points to the right of $L_2$ are 
+pictured in the upper horizontal-axis histograms. 
+For the left-tailed case (A, D), distributions to the
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/PostHocExplanatoryFig_betaFecund.jpg}
+\caption{A detailed look at ATA effects for the beta fecundities lottery model.
+The figure gives a pictorial representation of the calculation of  
+$\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ (A, C, D, F) and 
+$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$ (B, E)
+for the model, considering both left- (A, D) and right-tail (C, F) associations
+between $E_i$ and $C_{i \bs i}$.
+The colors and contours on A-C $\ln(1-\delta+E/C)$, whereas the points show a sample
+from the distribution $(E_i,C_{i \bs i})$ (A, C) or $(E_i^{||},C_{i \bs i}^{||})$ (B).
+The quantity $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ can be
+well approximated for left- and right-tail associations by averaging function 
+values over the points on panels A and C, respectively. The quantity
+$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
+can be well approximated by averaging function 
+values over the points on B. Values of $\log_{10}(E/C)$ for the points on A-C correspond 
+to signed distances of the points from the upward sloping dashed lines. 
+We used parameters $\eta_1 = 1$, $\eta_2 = 1.2$, and and $\delta=0.6$. }\label{fig:beta_fecundidities_explanatory}
+\end{figure}
 

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
explanatory text pretty developed at this point, though I will edit, and may add another example using the lognormal fecundities lottery model

@@
 each other, corresponding to similar but non-identical responses of species to
-environmental fluctuation (see SI section \ref{SIsect:temporal_niche_diff}); 
-but how do ATA effects then happen? We here extend the intuition
+environmental fluctuation (SI section \ref{SIsect:temporal_niche_diff}). 
+But how do ATA effects then happen? We here extend the intuition
 introduced in the Discussion and provide examples.
 
+First, we illustrate using the beta fecundities lottery model that
+the potential for storage to occur when 
+$E$ and $C$ are both higher than average, versus the potential  
+when $E$ and $C$ are
+both lower than average, are not necessarily equivalent.
 Storage can occur if population
@@
 $E_i$ and $C_{i \bs i}$ have the potential to be much greater than those contributed
-by the right tails. Essentially, values of $\log(1-\delta+E_i/C_{i \bs i})$
-occurring when $E_i$ and $C_{i \bs i}$ are both in the upper halves of their
+by the right tails. Essentially, annual growth-when-rare rates for species $i$
+that occur when $E_i$ and $C_{i \bs i}$ are both in the upper halves of their
 distributions can never exceed $\log(1-\delta+\frac{2\eta_i \delta}{\eta_j})$,
-whereas values ocurring when $E_i$ and $C_{i \bs i}$ are both in the lower
-halves of their distributions can be arbitrarily large (though with decreasing 
-probability for larger values). Apparently, the years with the most potential to 
+whereas rates ocurring when $E_i$ and $C_{i \bs i}$ are both in the lower
+halves of their distributions can be arbitrarily large. 
+Apparently, the years with the most potential to 
 contribute to storage
@@
 is actually poorer than average, because, for those years, competition can be very low,
-so that the ratio $E/C$ is large. Years for which the
+so that the ratio $E_i/C_{i \bs i}$ is large. Years for which the
 environment is exceptionally good actually cannot contribute much to storage effects for 
-this model, because for those years, $E/C$ cannot be large.
-It is the imbalance between $E$ and $C$ that allows for storage effects. 
-Counterintuitively, for this model, that imbalance is greatest when the environment is 
-poor. This result is counterintuitive because one may be used to thinking of storage effects
-occuring when reproductive pulses from environmentally good years sustain the population
-through environmentally poor years. But, in fact, good years for population growth may
-happen principally when the environment is poor, if those are also the times when 
-competition is very low, as for this model.
+this model, because for those years, $E_i/C_{i \bs i}$ cannot be very large.
+It is the \emph{imbalance} between $E$ and $C$ that allows for storage effects,
+not solely the quality of environment. 
+Counterintuitively, that imbalance can be greater for this model when the environment is 
+sub-optimal.
 
 Fig. \ref{fig:beta_fecundidities_explanatory} provides quantitative backing to
-the ideas of the prior paragraph. The growth function $\ln(1-\delta+E/C)$ that pertains 
-to the lottery model
+the ideas of the prior paragraph, so we orient the reader 
+to the figure before using it below to further explain
+ATAs. The growth function $\ln(1-\delta+E/C)$ that pertains to the lottery model
 is plotted with colors and contours on Fig. \ref{fig:beta_fecundidities_explanatory}A-C.
@@
 The dashed lines labelled $L_2$ on A-C separate the upper-right $50\%$
-of each distribution from the lower-left $50\%$. The dashed lines labeled
-$L_1$ separate the upper-left $50\%$ from the lower-right $50\%$.
+of each distribution from the lower-left $50\%$, and those portions are labelled. 
 The quantity $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ can be
-well approximated for left- and right-tail associations by averaging function 
-values, depicted with color and contours, over the points on panels A and C, respectively. The quantity
+well approximated, for left- and right-tail association cases, by averaging values of the
+function depicted with color and contours over the points on panels A and C, 
+respectively. The quantity
 $\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
@@
 
-The details of the averaging process described above are pictured in 
-Fig. \ref{fig:beta_fecundidities_explanatory}, panels D-F;
-we now describe how. 
+The details of the averaging processes described above are pictured in 
+Fig. \ref{fig:beta_fecundidities_explanatory}, panels D-F. 
 Signed distances of points on Fig. \ref{fig:beta_fecundidities_explanatory} 
@@
 $\log_{10}(E_1^\nat/C_{1 \bs 1}^\nat)$ on B. The distributions of these 
-quantities across points to the left of $L_2$ are pictured 
-in the lower horizontal-axis historgrams on panels D-F, respectively;
-and distributions for points to the right of $L_2$ are 
+quantities across points in the ``left'' region are pictured 
+in the lower horizontal-axis histograms on panels D-F, respectively;
+and distributions for points in the ``right'' region are 
 pictured in the upper horizontal-axis histograms. 
-For the left-tailed case (A, D), distributions to the
+``Right'' histograms on the horizontal axis of D,F are always highly 
+concentrated, corresponding to the fact that, for the beta fecundities
+lottery model, $E_i/C_{i \bs i}$ cannot vary much when $E_i$ and 
+$C_{i \bs i}$ are in the upper portions of their distributions. 
+Likewise, the ``right'' histogram on E is highly concentrated because
+$E_i^\nat/C_{i \bs i}^\nat$ cannot vary much when $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ are in the upper portions of their distributions.
+In contrast, ``left'' histograms on the horizontal axis of D-F
+can be highly variable, though only when $E_i$ and $C_{i \bs i}$
+(or $E_i^\nat$ and $C_{i \bs i}^\nat$) are not tightly 
+correlated in the left portions of their distributions.
+The function pictured with a black line on panels D-F is simply the function
+$z \rightarrow \ln(1-\delta+10^z)$ that takes $z=\log_{10}(E_1,C_{1 \bs 1})$
+to $\ln(1-\delta+E_1/C_{1 \bs 1})$, representing growth rate for the model.
+Storage effects are manifest by this function, because it is flat on the left and 
+upward sloping on the right, so highly variable horizontal-axis inputs to the function
+can translate into higher average vertical-axis values than can concentrated
+horizontal-axis inputs of the same mean; e.g., the histograms labeled ``left'' on
+E,F. Red dots indicate the mean values of
+each histogram. The average vertical-axis values of the two red dots on each panel
+equal the GWR of species $i$ for the model (value displayed numerically).
+
+We now use Fig. \ref{fig:beta_fecundidities_explanatory} to explain ATAs. Coexistence is possible for 
+right-tail association (panel F, positive GWR) but impossible for left-tail association (D,
+negative GWR). Coexistence is possible for the right-tail case essentially
+because the points on the ``left'' region in C produce highly variable
+annual growth rates; the same effect does not occur for the left-tail case because 
+points in the ``left'' region on panel A are concentrated on line $L_1$ because of the
+strong left-tail association of $E_i$ and $C_{i \bs i}$ on that panel. 
+Association, or lack thereof, in right tails makes essentially no difference 
+to growth rates because $E_i/C_{i \bs i}$ cannot vary much. Overall, ATA effects 
+are positive for the right-tail case because strong association between 
+$E_i$ and $C_{i \bs i}$ is concentrated in the right tails, where there 
+is little potential for storage effects; and so disassociation between
+$E_i$ and $C_{i \bs i}$ in the left tails, where there is much potential for
+storage effects, can be stronger. For the actual quantification of ATA effects,
+a comparison is made to the case of symmetric association 
+(panels B, E), as a benchmark. What is assessed by our quantification of ATA
+effects is essentially the degree to which patterns of disassociations between 
+$E$ and $C$, which can cause storage effects, align with the portions of 
+the $E,C$ distribution for which the potential for storage effects is highest,
+with larger ATA effects indicating better alignment, such as in panels C, F
+of Fig. \ref{fig:beta_fecundidities_explanatory}.
 
@@
 for the model, considering both left- (A, D) and right-tail (C, F) associations
-between $E_i$ and $C_{i \bs i}$.
-The colors and contours on A-C $\ln(1-\delta+E/C)$, whereas the points show a sample
-from the distribution $(E_i,C_{i \bs i})$ (A, C) or $(E_i^{||},C_{i \bs i}^{||})$ (B).
-The quantity $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ can be
-well approximated for left- and right-tail associations by averaging function 
-values over the points on panels A and C, respectively. The quantity
-$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
-can be well approximated by averaging function 
-values over the points on B. Values of $\log_{10}(E/C)$ for the points on A-C correspond 
-to signed distances of the points from the upward sloping dashed lines. 
-We used parameters $\eta_1 = 1$, $\eta_2 = 1.2$, and and $\delta=0.6$. }\label{fig:beta_fecundidities_explanatory}
+between $E_i$ and $C_{i \bs i}$. ATA effects are the difference of these two values.
+The figure also illustrates how ATAs reflect the degree of alignment between where 
+in the $EC$ distributions associations are allocated and which parts of that distribution
+have high potential for storage effects. See SI section X2
+%***DAN: fix this link when you put the text into the paper
+for details. We used model parameters $\eta_1 = 1$, $\eta_2 = 1.2$, and $\delta=0.6$
+for this example.}\label{fig:beta_fecundidities_explanatory}
 \end{figure}

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
merge

--- 2023-07-27 paper/ExplanationOfATAEffectsWorkingDoc.tex
@@
 related the invader's EC relationship is, the more interspecific competition it experiences, 
-destabalizing coexistence. Equivalently, invader EC relationships with more spread stably
+destabalizing coexistence (maybe only true with small variance). Equivalently, invader EC relationships with more spread stably
 contribute to coexistence (through the storage effect). As seen in Fig. \ref{thefig}, ATAs in 
@@
 
-\subsection{Dan thoughts 2023 07 26}
-
-% %Attempt 1 - not as good, Dan thinks. May be slightly misleading. See below for attempt 2.
-% %
-% %I have in mind that the following paragraph will go i the Discussion, with maybe a sentence
-% %or two in the Intro as a preview of coming attractions.
-% Just as for storage effects, temporal niche differentiation is necessary in order for ATA
-% effects to materialize. This makes intuitive sense because temporal niche differentiation is 
-% also a prerequistite of storage effects [\cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}],
-% and ATA effects are a part of storage effects, but see
-% SI section X for a detailed explanation. 
-% %Topic sentence follows. Possibly we can state something like this in the Intro and say
-% %we explain it in the Discussion after formalizing the extent to which ATAs are important.
-% If temporal niche differentiation occurs, ATA effects essentially capture the influence of asymmetries 
-% in the extent to which storage effects can occur in environmentally good versus bad years,
-% as we here explain.
-% %
-% These asymmetries can make a big difference to total storage effects and coexistence,
-% as some of our analyses showed.
-% %
-% If temporal niche differentiation occurs
-% but is partial (as will be typical), the quantities $E$ and $C$ experienced by a currently scarce 
-% species (i.e., $E_i$ and $C_{i \bs i})$ will be positively but imperfectly correlated. 
-% %
-% Storage effects can occur if some buffering mechanism reduces losses in bad years so that
-% gains in good years can more than make up for them. 
-% %
-% Bad years happen whenever the negative effects of $C$ outweigh any 
-% positive effects of $E$; but this can occur
-% both when 1) $E$ is moderately favorable but $C$ is very strong, as well as when 2) $E$ is very unfavorable,
-% or moderately unfavorable and $C$ is not too weak.
-% Conversely, good years happen whenever positive effects of $E$ outweigh negative effects of $C$;
-% but this can occur when 1) $E$ is very favorable, or moderately favorable and $C$ is not too strong,
-% as well as when 2) $E$ is moderately unfavorable but $C$ is very weak. 
-% %
-% ATA effects essentially quantify the asymmetry of storage that occurs when 
-% $E$ and $C$ are both higher than average (the cases 1, above); versus storage that occurs 
-% when $E$ and $C$ are
-% both lower than average (the cases 2, above). 
-% %
-% The distinction is an important one because it corresponds to temporal niche differentiation
-% which occurs under two very different sets of environmental conditions, possibly via
-% different ecological or physiological mechanisms.
-% %
-% Details, applications, and examples of this intuition 
-% are in SI section X, and these ideas are also used there to provide 
-% intuitive explanations for the relative differences in the importance
-% of ATA effects between our log-normal and beta fecundities lottery models.
-% %DAN: You ought to also mention explaining differences with the diatom system,
-% %if you figure out an explanation.
-
-%Attempt 2 below - better than the above, but may need to be simplified a bit more,
-%if possible
-%
-Just as for storage effects, temporal niche differentiation is necessary for ATA
-effects. This makes intuitive sense because temporal niche differentiation is 
-also a prerequistite of storage effects [\cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}],
-and ATA effects are a part of storage effects, but see
-SI section X for a detailed explanation. 
-%
-Assuming that temporal niche differentiation is present, how do ATA effects then happen?
-We here offer an intuitive explanation. 
-%
-If temporal niche differentiation occurs
-but is partial, the $E$ and $C$ experienced by a currently scarce 
-species (i.e., $E_i$ and $C_{i \bs i})$ will be positively but imperfectly associated, 
-and the association may be asymmetric. 
-%
-Storage effects can occur if some buffering mechanism reduces losses in bad years so that
-gains in good years can more than make up for them. 
-%
-Bad years happen when the negative effects of $C$ outweigh any 
-positive effects of $E$; but this can occur
-both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is very unfavorable,
-or moderately unfavorable and $C$ is not too weak.
-Conversely, good years happen when positive effects of $E$ outweigh negative effects of $C$;
-but this can occur 1) when $E$ is very favorable, or moderately favorable and $C$ is not too strong,
-as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
-%
-It turns out that the potential for storage to occur when 
-$E$ and $C$ are both higher than average (the cases 1, above); versus the potential  
-when $E$ and $C$ are
-both lower than average (the cases 2) are not necessarily equivalent.
-Depending on the growth rate function, $r(E,C)$, storage potential may differ
-across the spectrum from 1 to 2.
-%
-ATA effects are the extent to which asymmetries of association between $E$ and 
-$C$ are aligned with the potential for storage across the dichotomy 1 versus 2.
-A comparison is made to a benchmark of symmetric association between $E$ and $C$.
-%
-ATA effects therefore help quantify different degrees of temporal niche differentiation 
-occuring under different sets of environmental conditions (such as cases 1 and 2), possibly via
-different ecological or physiological mechanisms.
-%
-Details, applications, and examples of this intuition 
-are in SI section X, and these ideas are also used there to provide 
-intuitive explanations for the relative differences in the importance
-of ATA effects between our log-normal and beta fecundities lottery models.
-%DAN: You ought to also mention explaining differences with the diatom system,
-%if you figure out an explanation.
-
-\subsection{SI section X corresponding to the thoughts above}
-
-Just as for storage effects [see \cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}], 
-temporal niche differentiation is necessary in order for ATA
-effects to materialize; we here explain why in detail. ``Lack of temporal niche differentiation'' can be interpreted
-to mean that species respond in identical ways to environmental fluctuations. 
-In a two-species model such as those we consider, this should mean a monotonically
-increasing relationship between $E_i$ and $C_{i \bs i}$. The former is the environment
-as perceived by species $i$. The latter is competition that a scarce invader $i$ experiences
-in a monoculture of $j$, and that competition comes only from $j$ (because $i$ itself
-is scarce), so should correspond to the environmental response of $j$. If 
-$E_i$ and $C_{i \bs i}$ are monotonically related, then $(E_i,C_{i \bs i})$
-and $(E_i^\nat,C_{i \bs i}^\nat)$ are distributed in the same way, so 
-$\E[r_i(E_i,c_{i \bs i})]-\E[r_i(E_i^\nat,c_{i \bs i}^\nat)]=0$.
-Therefore, 
-\begin{align}
-\Delta_i^{[EC]} &= (\olr_{i \bs i}-\olr_{j \bs i})-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
-&= \olr_{i \bs i}-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
-&= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)]+\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] \label{eq:expri}\\
-&= \E[r_j(E_j^\nat,C_{j \bs i}^\nat)].
-\end{align}
-But we already know $\E[r_j(E_j^\nat,C_{j \bs i}^\nat)]=0$
-because there is no niche differentiation between $j$ and itself, so 
-$E_i$ and $C_{j \bs i}$ should be monotonically related. So 
-$\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] = \E[r_j(E_j,C_{j \bs i})]=0$, giving
-$\Delta_i^{[EC]}=0$. So temporal
-niche differentiation is necessary for ATA effects on coexistence to 
-materialize.
-
-If temporal niche differentiation occurs but is partial, 
-then $E_i$ and $C_{i \bs i}$ will be positively but imperfectly correlated with
-each other, corresponding to similar but non-identical responses of species to
-environmental fluctuation; but how do ATA effects then happen? We here extend the intuition
-introduced in the Discussion and add detail.
-
-Storage can occur if population
-increases during good years outweigh decreases during bad years. For instance,
-if $r_i(E_i,C_{i \bs i})$ is very large when the balance of $E_i$ and 
-$C_{i \bs i}$ is favorable, but only moderately small when this balance is
-unfavorable, the expected value $\E[r_i(E_i,C_{i \bs i})]$ in (\ref{eq:expri})
-will be positive. But the potential for such storage of population growth
-may differ between when $E_i$ and $C_{i \bs i}$ are both high compared
-to when they are both low. 
-For instance, for the beta fecundities lottery model, $E_i/C_{i \bs i}$, a quantity
-which features prominently in $\E[r_i(E_i,C_{i \bs i})]=\log(1-\delta+E_i/C_{i \bs i})$,
-takes values bounded above by $\frac{2\eta_i \delta}{\eta_j}$ 
-and bounded below by $\frac{\eta_i \delta}{2 \eta_j}$ when $E_i$ and $C_{i \bs i}$
-are both in the upper halves of their assumed beta distribtions (recall that 
-these quantities were assumed to be distributed, respectively, as $\eta_1 \beta$
-and $\eta_2 \beta/\delta$ for $\beta$ a beta-distributed random variable with parameters 
-$0.5$ and $0.5$). These bounds are separated only by a factor of four. 
-On the other hand, when $E_i$ and $C_{i \bs i}$ are both in the lower halves 
-of their distributions, $E_i/C_{i \bs i}$ can take any positive value because
-then either $E_i$ or $C_{i \bs i}$ can take values arbitrarily close to $0$. 
-Because values of $E_i/C_{i \bs i}$ have the potential to be much larger when 
-$E_i$ and $C_{i \bs i}$ are both in the lower halves of their distributions
-than when these quantities are both in their upper halves, and because
-the growth rate, $\log(1-\delta+E_i/C_{i \bs i})$, buffers low values of 
-$E_i/C_{i \bs i}$ (the infemum of the growth rate is $\log(1-\delta)$, whereas the supremum is
-infinite), the storage effects contributed by the left tails of 
-$E_i$ and $C_{i \bs i}$ have the potential to be much greater than those contributed
-by the right tails. Essentially, values of $\log(1-\delta+E_i/C_{i \bs i})$
-occurring when $E_i$ and $C_{i \bs i}$ are both in the upper halves of their
-distributions can never exceed $\log(1-\delta+\frac{2\eta_i \delta}{\eta_j})$,
-whereas values ocurring when $E_i$ and $C_{i \bs i}$ are both in the lower
-halves of their distributions can be arbitrarily large (though with decreasing 
-probability for larger values).
-
-\bibliographystyle{ecology_letters2}
-\bibliography{refs}
-
 \end{document}

### 2023-07-28 code/Figure1.R
removed v & h lines from plankotn plots

@@ -115,8 +115,8 @@ for (panel in 1:2){
 	pStats <- plankStats[[panel]]
 	plankton <- pStats$uv
 	plot(plankton, type="n", bty="n", ylim=c(0,1),xlim=c(0,1), xlab=ifelse(panel==2,"C. fusus density",NA), ylab=NA, xaxt=ifelse(panel==2,"s","n"))
-	abline(v=0.5, col="lightgrey", lwd=0.5, xpd=F)
-	abline(h=0.5, col="lightgrey", lwd=0.5, xpd=F)
+	#abline(v=0.5, col="lightgrey", lwd=0.5, xpd=F)
+	#abline(h=0.5, col="lightgrey", lwd=0.5, xpd=F)
 	lines(0:1,0:1, lwd=0.5, col="darkgrey")
 	points(plankton[pStats$bounds[,"left"],], col="darkgrey")
 	points(plankton[pStats$bounds[,"right"],], col="darkgrey", pch=19)

### 2023-07-28 code/ATAEffectIntuitionDemo_betaFecund.R
added code for making new SI figures

@@ -1,0 +1,496 @@
+#This script to help explain ATA intuitively. Uses the beta-fecundities lottery 
+#model case to make a figure supporting text added to the Discussion and SI
+#following a referee request.
+#
+#Reuman
+
+#***
+#Setup
+#***
+
+rm(list=ls())
+
+#data locations, copied from the scripts that generated the noise
+numRes_loc <- "../results_numeric/"
+
+noise_loc <- paste0(numRes_loc, "noise.RData")
+M_loc <- paste0(numRes_loc, "M.RDS")
+rho_loc <- paste0(numRes_loc, "rho.RDS")
+
+#locations to save resulting plots in
+resloc<-"../results_figs/"
+
+#graphical params
+numhistbins<-20
+numpoints<-1000
+
+#***
+#Functions
+#***
+
+#A function that makes it relatively easier for me to make the barplots
+#I want to make. Adds to a plot assumed to already be open.
+#
+#Args
+#m      A vector of counts
+#bks    Breaks. Length is length(m)+1
+#col    A colors
+#base   bottoms of rectangles
+#top    top of highest rectangle
+#axis   Use 1 for plotting the bars vertically, using the x axis for breaks.
+#         Use 2 for plotting the bars horizontally, using the y axis for breaks.
+#reddot Put a red dot at this location on the base axis
+#
+#Output - none, but adds to a plot
+#
+#Notes. 
+#If the arguments bks, base, top do not comport with the axes already
+#open then the barplot will appear (invisibly) off the plotting axes.
+#
+mybarplot<-function(m,bk,col,base,top,axis,reddot)
+{
+  Mx<-max(m)
+  
+  if (axis==1)
+  {
+    for (rc in 1:length(m))
+    {
[…]

### 2023-07-28 code/ATAEffectIntuitionDemo_lognormalFecund.R

@@ -1,0 +1,498 @@
+#This script to help explain ATA intuitively. Uses the lognormal fecundities lottery 
+#model case to make a figure supporting text added to the Discussion and SI
+#following a referee request.
+#
+#Reuman
+
+#***
+#Setup
+#***
+
+rm(list=ls())
+
+#data locations, copied from the scripts that generated the noise
+numRes_loc <- "../results_numeric/"
+
+noise_loc <- paste0(numRes_loc, "noise.RData")
+M_loc <- paste0(numRes_loc, "M.RDS")
+rho_loc <- paste0(numRes_loc, "rho.RDS")
+
+#locations to save resulting plots in
+resloc<-"../results_figs/"
+
+#graphical params
+numhistbins<-20
+numpoints<-1000
+
+#***
+#Functions
+#***
+
+#A function that makes it relatively easier for me to make the barplots
+#I want to make. Adds to a plot assumed to already be open.
+#
+#Args
+#m      A vector of counts
+#bks    Breaks. Length is length(m)+1
+#col    A colors
+#base   bottoms of rectangles
+#top    top of highest rectangle
+#axis   Use 1 for plotting the bars vertically, using the x axis for breaks.
+#         Use 2 for plotting the bars horizontally, using the y axis for breaks.
+#reddot Put a red dot at this location on the base axis
+#
+#Output - none, but adds to a plot
+#
+#Notes. 
+#If the arguments bks, base, top do not comport with the axes already
+#open then the barplot will appear (invisibly) off the plotting axes.
+#
+mybarplot<-function(m,bk,col,base,top,axis,reddot)
+{
+  Mx<-max(m)
+  
+  if (axis==1)
+  {
+    for (rc in 1:length(m))
+    {
[…]

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
just added a figure for the lognormal fecundities model

--- 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
@@
 
-We now use Fig. \ref{fig:beta_fecundidities_explanatory} to explain ATAs. Coexistence is possible for 
-right-tail association (panel F, positive GWR) but impossible for left-tail association (D,
+We now relate Fig. \ref{fig:beta_fecundidities_explanatory} back to ATAs. Coexistence is possible for 
+right-tail association (panel F, positive GWR) but impossible for left-tail association (panel D,
 negative GWR). Coexistence is possible for the right-tail case essentially
-because the points on the ``left'' region in C produce highly variable
-annual growth rates; the same effect does not occur for the left-tail case because 
+because the points in the ``left'' region in C produce highly variable
+annual growth rates, which lead to storage effects; the same effect does 
+not occur for the left-tail case because 
 points in the ``left'' region on panel A are concentrated on line $L_1$ because of the
 strong left-tail association of $E_i$ and $C_{i \bs i}$ on that panel. 
-Association, or lack thereof, in right tails makes essentially no difference 
-to growth rates because $E_i/C_{i \bs i}$ cannot vary much. Overall, ATA effects 
+Association, or lack thereof, in right tails makes little difference 
+to mean growth rates because $E_i/C_{i \bs i}$ cannot vary much. Overall, ATA effects 
 are positive for the right-tail case because strong association between 
 $E_i$ and $C_{i \bs i}$ is concentrated in the right tails, where there 
-is little potential for storage effects; and so disassociation between
-$E_i$ and $C_{i \bs i}$ in the left tails, where there is much potential for
-storage effects, can be stronger. For the actual quantification of ATA effects,
+is little potential for storage effects anyway; and so disassociation between
+$E_i$ and $C_{i \bs i}$ occurs in the left tails, where there is much potential for
+storage effects. For the actual quantification of ATA effects,
 a comparison is made to the case of symmetric association 
-(panels B, E), as a benchmark. What is assessed by our quantification of ATA
-effects is essentially the degree to which patterns of disassociations between 
-$E$ and $C$, which can cause storage effects, align with the portions of 
-the $E,C$ distribution for which the potential for storage effects is highest,
-with larger ATA effects indicating better alignment, such as in panels C, F
-of Fig. \ref{fig:beta_fecundidities_explanatory}.
+(panels B, E), as a benchmark. What is assessed by ATA
+effects is the degree to which patterns of disassociation between 
+$E$ and $C$ align with the portions of $E,C$ distribution for which the potential
+for storage effects is high. Larger ATA effects indicate better alignment.
 
@@
 
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/PostHocExplanatoryFig_lognormalFecund.jpg}
+\caption{A detailed look at ATA effects for the log-normal fecundities lottery model.
+Similar to Fig. \ref{fig:beta_fecundidities_explanatory}, but for the 
+log-normal fecundities lottery model instead of the beta fecundities
+lottery model. See SI section X2
+%***DAN: fix this link when you put the text into the paper
+for details. We used model parameters $\sigma = 1$, $\mu_1 = 0$, $\mu_2=0.5$, 
+and $\delta=0.6$ for this example.}\label{fig:lognormal_fecundidities_explanatory}
+\end{figure}
+
 \bibliographystyle{ecology_letters2}

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
finished a complete draft of the new explanatory material, about to start putting it into the paper

@@
 
+Fig. \ref{fig:lognormal_fecundidities_explanatory} is similar to 
+Fig. \ref{fig:beta_fecundidities_explanatory}, but is based on the log-normal 
+fecundities lottery model instead of the beta fecundities lottery model. 
+The figure illustrates that the potential for storage events is equivalent in
+the upper and lower tails of the $EC$ distribution (labelled ``right'' and ``left'',
+respectively, on panels A-C). That is why ATA effects are fairly modest for the 
+log-normal fecundities lottery model for the both the left- and right-tail
+associated cases.
+
 \begin{figure}

### 2023-07-28 code/MASTER.R
caused MASTER to run the new codes making the explanatory figures

@@ -56,3 +56,6 @@ source("./diatom/VKQplots.R")
 #Stats the referee asked us to add regarding Fig 1d, e
 source("Fig1de_stats.R")
 
+#Figurs supporting new explanatory material referee asked for
+source("ATAEffectIntuitionDemo_betaFecund.R")
+source("ATAEffectIntuitionDemo_lognormalFecund.R")

### 2023-07-28 paper/Paper.Rnw
added an Intro para on intuition behind ATAs

@@
 impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
-Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
+Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
 when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
@@
 than are years of good environment, limiting the potential for rescue effects. 
-And \cite{Ghosh_2021} show that ATAs can alter the stability of 
+And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
@@
 %Intro to storage effects and why they might be influenced by ATAs
-It is possible to see, intuitively, 
-why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
+It is possible to see why it is plausible that
+ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
 what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
@@
 by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
-Storage effects hinge on positive covariation between the fluctuating quality of the environment
+Storage effects hinge on covariation between the fluctuating quality of the environment
 for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
@@
 ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
+
+%Some intuition
+We next give the beginnings of some intuition about how ATA effects work, with a fuller understanding
+necessarily coming only later in the paper.
+It is known that storage effects can occur if some buffering mechanism reduces losses in bad years so that
+gains in good years can more than make up for them. 
+%
+Bad years happen when the negative effects of $C$ outweigh any 
+positive effects of $E$; but this can occur
+both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is very unfavorable,
+or moderately unfavorable and $C$ is not too weak.
+Conversely, good years happen when positive effects of $E$ outweigh negative effects of $C$;
+but this can occur 1) when $E$ is very favorable, or moderately favorable and $C$ is not too strong,
+as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
+%
+It turns out that the potential for storage to occur when 
+$E$ and $C$ are both higher than average (the cases 1, above); versus the potential  
+when $E$ and $C$ are
+both lower than average (the cases 2) are not necessarily equal.
+Depending on the nature of population growth, storage potential may differ
+across the spectrum from 1 to 2.
+%
+ATA effects are essentially the extent to which asymmetries of association between $E$ and 
+$C$ are aligned with the differences in the potential for storage across the dichotomy 1 versus 2.
+If $E$ and $C$ are strongly associated in the parts of their distribution where
+storage potential is high, ATA effects will be relatively low, and vice versa.
+We spend most of the paper formalizing ATA effects in MCT and demonstrating to what 
+extent ATA effects can be strong versus weak, but we return in the Discussion to the task
+of building an intuition for how ATAs work.
 

### 2023-07-28 paper/Paper.Rnw
more edits to the Intro pertaining to the new intuition building thta I put there

@@
 is constant through time. 
-Notation for the whole paper is summarized in Table \ref{tab:notation}.
 We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
@@
 ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
+Notation for the whole paper is summarized in Table \ref{tab:notation}.
+Typesetting procedures used by publishers of ecological journals render mathematics sub-optimally,
+so we have appended a version of this main text rendered using latex to the end of the 
+Supporting Information (SI); we recommend that readers interested in the mathematical content
+of the study read that version.
+%***DAN: Don't forget to do this
 
 %Some intuition
-We next give the beginnings of some intuition about how ATA effects work, with a fuller understanding
+We next give some initial intuition about how ATA effects on coexistence work, with a fuller understanding
 necessarily coming only later in the paper.
-It is known that storage effects can occur if some buffering mechanism reduces losses in bad years so that
+It is known that storage effects can occur if some buffering mechanism reduces population 
+losses in bad years so that
 gains in good years can more than make up for them. 
@@
 positive effects of $E$; but this can occur
-both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is very unfavorable,
-or moderately unfavorable and $C$ is not too weak.
+both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is 
+unfavorable and $C$ is not too weak.
 Conversely, good years happen when positive effects of $E$ outweigh negative effects of $C$;
-but this can occur 1) when $E$ is very favorable, or moderately favorable and $C$ is not too strong,
+but this can occur 1) when $E$ is favorable and $C$ is not too strong,
 as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
 %
-It turns out that the potential for storage to occur when 
+It turns out that the potential for storage events to occur when 
 $E$ and $C$ are both higher than average (the cases 1, above); versus the potential  
@@
 ATA effects are essentially the extent to which asymmetries of association between $E$ and 
-$C$ are aligned with the differences in the potential for storage across the dichotomy 1 versus 2.
+$C$ are aligned with differences in the potential for storage across the spectrum from 1 to 2.
 If $E$ and $C$ are strongly associated in the parts of their distribution where
-storage potential is high, ATA effects will be relatively low, and vice versa.
-We spend most of the paper formalizing ATA effects in MCT and demonstrating to what 
-extent ATA effects can be strong versus weak, but we return in the Discussion to the task
-of building an intuition for how ATAs work.
+storage potential is high, ATA effects will be low, and vice versa.
 
@@
 We address how ATAs contribute to these components. 
-We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
-spatial variation can be considered in future work.
+%We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
+%spatial variation can be considered in future work.
 G2) We explore, using the simple lottery model, the relative magnitudes 
 of the contributions to coexistence that come
-from ATAs compared to other, previously known mechanisms of coexistence; and to determine whether species may 
+from ATAs compared to other, previously known mechanisms of coexistence; and we determine whether species may 
 sometimes coexist with ATAs but not without, or vice versa. This is to help 
@@
 ATAs of environmental variables and thereby influence competition and coexistence.
+And we return in the Discussion to the task
+of building intuition for how ATAs work. 
 Overall, our study presents a new mechanism of species coexistence

### 2023-07-28 paper/Paper.Rnw
took a stab at a Discussion paragraph

@@
 Typesetting procedures used by publishers of ecological journals render mathematics sub-optimally,
-so we have appended a version of this main text rendered using latex to the end of the 
-Supporting Information (SI); we recommend that readers interested in the mathematical content
-of the study read that version.
+so we have appended a latex version of this main text to the end of the 
+Supporting Information (SI). Readers interested in the mathematical content
+of the study should read that version.
 %***DAN: Don't forget to do this
@@
 %Some intuition
-We next give some initial intuition about how ATA effects on coexistence work, with a fuller understanding
-necessarily coming only later in the paper.
-It is known that storage effects can occur if some buffering mechanism reduces population 
+We next give some initial intuition about how ATA effects on coexistence work, 
+providing additional intuition in the Discussion.
+It is known that storage effects can occur if a buffering mechanism reduces population 
 losses in bad years so that
 gains in good years can more than make up for them. 
-%
 Bad years happen when the negative effects of $C$ outweigh any 
@@
 as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
-%
 It turns out that the potential for storage events to occur when 
@@
 when $E$ and $C$ are
-both lower than average (the cases 2) are not necessarily equal.
-Depending on the nature of population growth, storage potential may differ
+both lower than average (the cases 2) are not necessarily equal -- storage potential may differ
 across the spectrum from 1 to 2.
-%
-ATA effects are essentially the extent to which asymmetries of association between $E$ and 
+ATA effects quantify the extent to which asymmetries of association between $E$ and 
 $C$ are aligned with differences in the potential for storage across the spectrum from 1 to 2.
-If $E$ and $C$ are strongly associated in the parts of their distribution where
-storage potential is high, ATA effects will be low, and vice versa.
+If $E$ and $C$ are largely unassociated in the parts of their distribution where
+storage potential is high, ATA effects will be high, and vice versa.
 
@@
 ATAs of environmental variables and thereby influence competition and coexistence.
-And we return in the Discussion to the task
-of building intuition for how ATAs work. 
 Overall, our study presents a new mechanism of species coexistence
@@
 for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
+
+%Para continuing to try to build intuition about how ATAs work
+We return to the task, begun in the Introduction, of building intuition for how
+ATA effects work. 
+Just as for storage effects [\cite{Chesson_1994}, Box 2 of \cite{StumpVasseur2023}], 
+temporal niche differentiation is necessary for ATA
+effects (SI section X1). 
+%DAN: fix reference
+If temporal niche differentiation occurs
+but is partial, as will be typical, $E_i$ and $C_{i \bs i}$ for a 
+species $i$ will be positively but imperfectly associated, 
+and the association may be asymmetric. 
+Storage events occur for species $i$ when the positive effects of $E_i$ overwhelm the negative 
+effects of $C_{i \bs i}$, producing large values of $r_i(E_i,C_{i \bs i})$
+which are not completely offset by declines at other times. 
+For instance, for the lottery model,
+$r_i(E_i,C_{i \bs i})=\log(1-\delta+E_i/C_{i \bs i})$ can get large whenever
+$C_{i \bs i}$ is a small fraction of $E_i$, but can never get smaller
+than $\log(1-\delta)$. 
+The potential for storage events can differ when $E_i$ and $C_{i \bs i}$ are both
+above their median values compared to when they are both below median. For instance,
+the ratio $E_i/C_{i \bs i}$, and therefore the growth rate 
+$\log(1-\delta+E_i/C_{i \bs i})$, can both get very big, corresponding to
+rapid growth, if $C_{i \bs i}$ is close to $0$, even if $E_i$ is fairly small.
+In contrast, if $E_i$ and $C_{i \bs i}$ are both in the right portions of their distributions, 
+so that $C_{i \bs i}$ is bounded away from $0$, there may be much sharper limits on the 
+ratio $E_i/C_{i \bs i}$ and the corresponding growth rate.
+ATA effects quantify the extent to which asymmetries of association between $E$ and 
+$C$ are aligned with varying potential for storage across the $E,C$ distribution. 
+ATA effects help quantify different degrees of temporal niche differentiation 
+occurring under different sets of environmental conditions, possibly via
+different mechanisms.
+Details and examples of this intuition 
+are in SI section X2, 
+%DAN: Fix reference
+where comparisons are also made between our beta and log-normal fecundities 
+lottery models.
 

### 2023-07-28 paper/ExplanationOfATAEffectsWorkingDoc.tex
wrote an intuition para in the Discussion and put the SI material in. Need to edit. Need to shorten and simplify the whole doc as much as possible.

@@
 
-\subsection{SI section X1 corresponding to the thoughts above, call it Temporal niche differentiation is necessary for ATAs}\label{SIsect:temporal_niche_diff}
+\subsection{SI section X1 corresponding to the thoughts above, call it Temporal niche differentiation is necessary for ATA effects}\label{SIsect:temporal_niche_diff}
 
@@
 \subsection{SI section X2 corresponding to the thoughts above, call it Intuition and 
-examples for ATAs}\label{SIsect:intuition_examples}
+examples for ATA effects}\label{SIsect:intuition_examples}
 

### 2023-07-28 paper/Paper.Rnw
wrote an intuition para in the Discussion and put the SI material in. Need to edit. Need to shorten and simplify the whole doc as much as possible.

@@
 temporal niche differentiation is necessary for ATA
-effects (SI section X1). 
-%DAN: fix reference
+effects (SI section \ref{SI-SIsect:temporal_niche_diff}). 
 If temporal niche differentiation occurs
@@
 Details and examples of this intuition 
-are in SI section X2, 
-%DAN: Fix reference
-where comparisons are also made between our beta and log-normal fecundities 
-lottery models.
+are in SI section \ref{SI-SIsect:intuition_examples} and SI 
+Figs \ref{SI-fig:beta_fecundidities_explanatory}-\ref{SI-fig:lognormal_fecundidities_explanatory}, 
+where comparisons are also made between our two lottery models.
 

### 2023-07-28 paper/SupMat.Rnw
wrote an intuition para in the Discussion and put the SI material in. Need to edit. Need to shorten and simplify the whole doc as much as possible.

@@
 
+\section{   Temporal niche differentiation is necessary for ATA effects}\label{SIsect:temporal_niche_diff}
+
+Just as for storage effects [see \cite{Chesson_1994} and Box 2 of \cite{StumpVasseur2023}], 
+temporal niche differentiation is necessary in order for ATA
+effects to materialize; we here explain why, in detail. ``Lack of temporal niche differentiation'' can be interpreted
+to mean that species respond in identical ways to environmental fluctuations. 
+In a two-species model such as those we consider, this should mean a monotonically
+increasing relationship between $E_i$ and $C_{i \bs i}$. The former is the environment
+as perceived by species $i$. The latter is competition that a scarce invader $i$ experiences
+in a monoculture of $j$, and that competition comes only from $j$ (because $i$ itself
+is scarce), so should correspond to the environmental response of $j$. If 
+$E_i$ and $C_{i \bs i}$ are monotonically related, then $(E_i,C_{i \bs i})$
+and $(E_i^\nat,C_{i \bs i}^\nat)$ are distributed in the same way, so 
+$\E[r_i(E_i,c_{i \bs i})]-\E[r_i(E_i^\nat,c_{i \bs i}^\nat)]=0$.
+Therefore, 
+\begin{align}
+\Delta_i^{[EC]} &= (\olr_{i \bs i}-\olr_{j \bs i})-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
+&= \olr_{i \bs i}-(\olr_{i \bs i}^\nat - \olr_{j \bs i}^\nat) \\
+&= \E[r_i(E_i,C_{i \bs i})]-\E[r_i(E_i^\nat,C_{i \bs i}^\nat)]+\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] \label{eq:expri}\\
+&= \E[r_j(E_j^\nat,C_{j \bs i}^\nat)].
+\end{align}
+But we already know $\E[r_j(E_j^\nat,C_{j \bs i}^\nat)]=0$
+because there is no niche differentiation between $j$ and itself, so 
+$E_j$ and $C_{j \bs i}$ should be monotonically related. So 
+$\E[r_j(E_j^\nat,C_{j \bs i}^\nat)] = \E[r_j(E_j,C_{j \bs i})]=0$, giving
+$\Delta_i^{[EC]}=0$. So temporal
+niche differentiation, interpreted as Spearman correlation between 
+$E$ and $C$ less than $1$, is necessary for ATA effects on coexistence to 
+materialize.
+
+\section{   Intuition and examples for ATA effects}\label{SIsect:intuition_examples}
+
+If temporal niche differentiation occurs but is partial, 
+then $E_i$ and $C_{i \bs i}$ will be positively but imperfectly correlated with
+each other, corresponding to similar but non-identical responses of species to
+environmental fluctuation (SI section \ref{SIsect:temporal_niche_diff}). 
+But how do ATA effects then happen? We here extend the intuition
+introduced in the Discussion and provide examples.
+
+First, we illustrate using the beta fecundities lottery model that
+the potential for storage to occur when 
+$E$ and $C$ are both higher than average, versus the potential  
+when $E$ and $C$ are
+both lower than average, are not necessarily equivalent.
+Storage can occur if population
+increases during good years outweigh decreases during bad years. For instance,
+if $r_i(E_i,C_{i \bs i})$ is very large when the balance of $E_i$ and 
+$C_{i \bs i}$ is favorable, but only moderately small when this balance is
+unfavorable, the expected value $\E[r_i(E_i,C_{i \bs i})]$ in (\ref{eq:expri})
+will be positive. But the potential for such storage of population growth
+may differ between when $E_i$ and $C_{i \bs i}$ are both high compared
+to when they are both low. 
+For instance, for the beta fecundities lottery model, $E_i/C_{i \bs i}$, a quantity
+which features prominently in $r_i(E_i,C_{i \bs i})=\log(1-\delta+E_i/C_{i \bs i})$,
+takes values bounded above by $\frac{2\eta_i \delta}{\eta_j}$ 
+and bounded below by $\frac{\eta_i \delta}{2 \eta_j}$ when $E_i$ and $C_{i \bs i}$
+are both in the upper halves of their assumed beta distributions (recall that 
+these quantities were assumed to be distributed, respectively, as $\eta_1 \beta$
+and $\eta_2 \beta/\delta$ for $\beta$ a beta-distributed random variable with parameters 
+$0.5$ and $0.5$). These bounds are separated only by a factor of four. 
+On the other hand, when $E_i$ and $C_{i \bs i}$ are both in the lower halves 
+of their distributions, $E_i/C_{i \bs i}$ can take any positive value because
+then either $E_i$ or $C_{i \bs i}$ can take values arbitrarily close to $0$. 
+Because values of $E_i/C_{i \bs i}$ have the potential to be much larger when 
+$E_i$ and $C_{i \bs i}$ are both in the lower halves of their distributions
+than when these quantities are both in their upper halves, and because
+the growth rate, $\log(1-\delta+E_i/C_{i \bs i})$, buffers low values of 
+$E_i/C_{i \bs i}$ (the infemum of the growth rate is $\log(1-\delta)$, whereas the supremum is
+infinite), the storage effects contributed by the left tails of 
+$E_i$ and $C_{i \bs i}$ have the potential to be much greater than those contributed
+by the right tails. Essentially, annual growth-when-rare rates for species $i$
+that occur when $E_i$ and $C_{i \bs i}$ are both in the upper halves of their
+distributions can never exceed $\log(1-\delta+\frac{2\eta_i \delta}{\eta_j})$,
+whereas rates ocurring when $E_i$ and $C_{i \bs i}$ are both in the lower
+halves of their distributions can be arbitrarily large. 
+Apparently, the years with the most potential to 
+contribute to storage
+effects for the beta-fecundities lottery model are those for which the environment
+is actually poorer than average, because, for those years, competition can be very low,
+so that the ratio $E_i/C_{i \bs i}$ is large. Years for which the
+environment is exceptionally good actually cannot contribute much to storage effects for 
+this model, because for those years, $E_i/C_{i \bs i}$ cannot be very large.
+It is the \emph{imbalance} between $E$ and $C$ that allows for storage effects,
+not solely the quality of environment. 
+Counterintuitively, that imbalance can be greater for this model when the environment is 
+sub-optimal.
+
+Fig. \ref{fig:beta_fecundidities_explanatory} provides quantitative backing to
+the ideas of the prior paragraph, so we orient the reader 
+to the figure before using it below to further explain
+ATAs. The growth function $\ln(1-\delta+E/C)$ that pertains to the lottery model
+is plotted with colors and contours on Fig. \ref{fig:beta_fecundidities_explanatory}A-C.
+Samples were taken from the distribution $(E_i,C_{i \bs i})$ for the beta 
+fecundities lottery model for left- and right-tailed 
+associations between $E_i$ and $C_{i \bs i}$, and are plotted as
+points on Fig. \ref{fig:beta_fecundidities_explanatory}A and C, respectively. 
+Samples were also taken from the distribution $(E_i^\nat,C_{i \bs i}^\nat)$
+(which is the same regardless of whether the original $(E_i,C_{i \bs i})$
+distribution showed left- or right-tail association) and are plotted on 
+Fig. \ref{fig:beta_fecundidities_explanatory}B. 
+The dashed lines labelled $L_2$ on A-C separate the upper-right $50\%$
+of each distribution from the lower-left $50\%$, and those portions are labelled. 
+The quantity $\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ can be
+well approximated, for left- and right-tail association cases, by averaging values of the
+function depicted with color and contours over the points on panels A and C, 
+respectively. The quantity
+$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
+can be well approximated by averaging function 
+values over the points on B. So ATA effects for the model with left-tail association
+between $E_i$ and $C_{i \bs i}$ (respectively, right-tail association) are given by 
+computing the average function value over points on panel A (respectively, C)
+and then subtracting the average function value over points on panel B.
+Contributions to the storage effect of points in the left portions of 
+the $E,C$ distribution (the region labeled ``left'' on panels A and C)
+can be accurately estimated by averaging function values across
+points in that region; likewise storage contributions of points in the 
+``right'' region can be estimated by averaging function values over 
+points in that region.
+
+The details of the averaging processes described above are pictured in 
+Fig. \ref{fig:beta_fecundidities_explanatory}, panels D-F. 
+Signed distances of points on Fig. \ref{fig:beta_fecundidities_explanatory} 
+A-C from the lines $L_1$ correspond to 
+the quantities $\log_{10}(E_1/C_{1 \bs 1})$ on A, C and to 
+$\log_{10}(E_1^\nat/C_{1 \bs 1}^\nat)$ on B. The distributions of these 
+quantities across points in the ``left'' region are pictured 
+in the lower horizontal-axis histograms on panels D-F, respectively;
+and distributions for points in the ``right'' region are 
+pictured in the upper horizontal-axis histograms. 
+``Right'' histograms on the horizontal axis of D,F are always highly 
+concentrated, corresponding to the fact that, for the beta fecundities
+lottery model, $E_i/C_{i \bs i}$ cannot vary much when $E_i$ and 
+$C_{i \bs i}$ are in the upper portions of their distributions. 
+Likewise, the ``right'' histogram on E is highly concentrated because
+$E_i^\nat/C_{i \bs i}^\nat$ cannot vary much when $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ are in the upper portions of their distributions.
+In contrast, ``left'' histograms on the horizontal axis of D-F
+can be highly variable, though only when $E_i$ and $C_{i \bs i}$
+(or $E_i^\nat$ and $C_{i \bs i}^\nat$) are not tightly 
+correlated in the left portions of their distributions.
+The function pictured with a black line on panels D-F is simply the function
+$z \rightarrow \ln(1-\delta+10^z)$ that takes $z=\log_{10}(E_1,C_{1 \bs 1})$
+to $\ln(1-\delta+E_1/C_{1 \bs 1})$, representing growth rate for the model.
+Storage effects are manifest by this function, because it is flat on the left and 
+upward sloping on the right, so highly variable horizontal-axis inputs to the function
+can translate into higher average vertical-axis values than can concentrated
+horizontal-axis inputs of the same mean; e.g., the histograms labeled ``left'' on
+E,F. Red dots indicate the mean values of
+each histogram. The average vertical-axis values of the two red dots on each panel
+equal the GWR of species $i$ for the model (value displayed numerically).
+
+We now relate Fig. \ref{fig:beta_fecundidities_explanatory} back to ATAs. Coexistence is possible for 
+right-tail association (panel F, positive GWR) but impossible for left-tail association (panel D,
+negative GWR). Coexistence is possible for the right-tail case essentially
+because the points in the ``left'' region in C produce highly variable
+annual growth rates, which lead to storage effects; the same effect does 
+not occur for the left-tail case because 
+points in the ``left'' region on panel A are concentrated on line $L_1$ because of the
+strong left-tail association of $E_i$ and $C_{i \bs i}$ on that panel. 
+Association, or lack thereof, in right tails makes little difference 
+to mean growth rates because $E_i/C_{i \bs i}$ cannot vary much. Overall, ATA effects 
+are positive for the right-tail case because strong association between 
+$E_i$ and $C_{i \bs i}$ is concentrated in the right tails, where there 
+is little potential for storage effects anyway; and so disassociation between
+$E_i$ and $C_{i \bs i}$ occurs in the left tails, where there is much potential for
+storage effects. For the actual quantification of ATA effects,
+a comparison is made to the case of symmetric association 
+(panels B, E), as a benchmark. What is assessed by ATA
+effects is the degree to which patterns of disassociation between 
+$E$ and $C$ align with the portions of $E,C$ distribution for which the potential
+for storage effects is high. Larger ATA effects indicate better alignment.
+
+Fig. \ref{fig:lognormal_fecundidities_explanatory} is similar to 
+Fig. \ref{fig:beta_fecundidities_explanatory}, but is based on the log-normal 
+fecundities lottery model instead of the beta fecundities lottery model. 
+The figure illustrates that the potential for storage events is equivalent in
+the upper and lower tails of the $EC$ distribution (labelled ``right'' and ``left'',
+respectively, on panels A-C). That is why ATA effects are fairly modest for the 
+log-normal fecundities lottery model for the both the left- and right-tail
+associated cases.
+
 \section{   Asymmetric associations between population vital rates}\label{sect:vital_rates}
@@
 \end{figure}
-%***DAN: Jasmin, the value of se4 quoted in the main text was the same, to 5 digits, as the 
-%value of se4qij quoted here. This made me suspicious so I checked the values to 9 digits
-%and they were the same to 9 digits. This seems like it must be an error in the code. Please
-%check. The standard error expressions in SI S5.2 DO depend on qij, so this seems like it
-%should not happen. The standard errors listed in Fig. 3 and Fig. S1 are different, so that's
-%good.
-%JASMIN: I think its okay that they are the same.
 
@@
 \includegraphics[width=.66\textwidth]{../results_figs/VKQfig.pdf}
-\caption[Maximum growth rate, cellular silicate content, and half sturation constant dependences on temperature.]
+\caption[Maximum growth rate, cellular silicate content, and half saturation constant dependencies on temperature.]
 {Species maximum population growth rates, $V_i$ (a), half saturation constants for the dependence 
@@
 data were not available for $Q$ at $24^\circ$C from the original study of \cite{Descamps_2005}; \cite{Ellner_2016, Ellner_2019}
-assumed constancy of $Q_2$ for temperature between $18^\circ$C and $24^\circ$C. Data were also not avaialable at
-$24^\circ$C for $K_2$, and \cite{Ellner_2016, Ellner_2019} assumed constancy of $K_2$ for teperatures from 
+assumed constancy of $Q_2$ for temperature between $18^\circ$C and $24^\circ$C. Data were also not available at
+$24^\circ$C for $K_2$, and \cite{Ellner_2016, Ellner_2019} assumed constancy of $K_2$ for temperatures from 
 $12^\circ$C through $24^\circ$C. \label{fig:VKQ}}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/PostHocExplanatoryFig_betaFecund.jpg}
+\caption[A detailed look at ATA effects for the beta fecundities lottery model]{A 
+detailed look at ATA effects for the beta fecundities lottery model.
+The figure gives a pictorial representation of the calculation of  
+$\E \left[ \ln \left( 1-\delta+\frac{E_i}{C_{i \bs i}} \right) \right]$ (A, C, D, F) and 
+$\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$ (B, E)
+for the model, considering both left- (A, D) and right-tail (C, F) associations
+between $E_i$ and $C_{i \bs i}$. ATA effects are the difference of these two values.
+The figure also illustrates how ATAs reflect the degree of alignment between where 
+in the $EC$ distributions associations are allocated and which parts of that distribution
+have high potential for storage effects. See SI section \ref{SIsect:intuition_examples}
+for details. We used model parameters $\eta_1 = 1$, $\eta_2 = 1.2$, and $\delta=0.6$
+for this example.}\label{fig:beta_fecundidities_explanatory}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../results_figs/PostHocExplanatoryFig_lognormalFecund.jpg}
+\caption[A detailed look at ATA effects for the log-normal fecundities lottery model]{A 
+detailed look at ATA effects for the log-normal fecundities lottery model.
+Similar to Fig. \ref{fig:beta_fecundidities_explanatory}, but for the 
+log-normal fecundities lottery model instead of the beta fecundities
+lottery model. See SI section \ref{SIsect:intuition_examples}
+for details. We used model parameters $\sigma = 1$, $\mu_1 = 0$, $\mu_2=0.5$, 
+and $\delta=0.6$ for this example.}\label{fig:lognormal_fecundidities_explanatory}
 \end{figure}

### 2023-07-31 paper/Paper.Rnw
editted new additions explaining intuition of ATA effects, now done, moving on

@@
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
-plankton population density time series; many other examples have also been previously explored 
+plankton population density time series; many other ecological examples of ATAs have also been explored 
 \citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994} which will be familiar to some readers (storage effects are also reviewed conceptually below and
-defined formally in Theory). MCT has been
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and will be
+defined formally in the Theory section. MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
@@
 Notation for the whole paper is summarized in Table \ref{tab:notation}.
-Typesetting procedures used by publishers of ecological journals render mathematics sub-optimally,
-so we have appended a latex version of this main text to the end of the 
-Supporting Information (SI). Readers interested in the mathematical content
-of the study should read that version.
+Typesetting of mathematics by ecological journals is sub-optimal,
+so we appended a latex version of this main text to the end of the 
+Supporting Information (SI). Readers interested in the mathematical aspects should read that version.
 %***DAN: Don't forget to do this
@@
 If $E$ and $C$ are largely unassociated in the parts of their distribution where
-storage potential is high, ATA effects will be high, and vice versa.
+storage potential is high, ATA effects will be high. ATA effects quantify how temporal niche partitioning
+varies across $EC$ spectrum spanned by the cases 1-2 sketched above.
 
@@
 effects of $C_{i \bs i}$, producing large values of $r_i(E_i,C_{i \bs i})$
-which are not completely offset by declines at other times. 
+which are not completely offset by population declines at other times. 
 For instance, for the lottery model,
@@
 different mechanisms.
-Details and examples of this intuition 
+Further details and examples of this intuition 
 are in SI section \ref{SI-SIsect:intuition_examples} and SI 

### 2023-07-31 paper/SupMat.Rnw
editted new additions explaining intuition of ATA effects, now done, moving on

@@
 in a monoculture of $j$, and that competition comes only from $j$ (because $i$ itself
-is scarce), so should correspond to the environmental response of $j$. If 
-$E_i$ and $C_{i \bs i}$ are monotonically related, then $(E_i,C_{i \bs i})$
+is scarce), so should correspond to the environmental response of $j$. Therefore, if the environmental
+responses of the two species are the same, that should be reflected in a monotonic 
+relationship between $E_i$ and $C_{i \bs i}$. Then, $(E_i,C_{i \bs i})$
 and $(E_i^\nat,C_{i \bs i}^\nat)$ are distributed in the same way, so 
@@
 $\Delta_i^{[EC]}=0$. So temporal
-niche differentiation, interpreted as Spearman correlation between 
-$E$ and $C$ less than $1$, is necessary for ATA effects on coexistence to 
-materialize.
+niche differentiation, interpreted as a relationship between 
+$E$ and $C$ which is not simply monotonic, is necessary for ATA effects on coexistence to 
+materialize. Such a non-monotonic relationship occurs if  
+the Spearman correlation between 
+$E$ and $C$ is less than $1$.
 
@@
 If temporal niche differentiation occurs but is partial, 
-then $E_i$ and $C_{i \bs i}$ will be positively but imperfectly correlated with
+so that $E_i$ and $C_{i \bs i}$ are positively but imperfectly correlated with
 each other, corresponding to similar but non-identical responses of species to
-environmental fluctuation (SI section \ref{SIsect:temporal_niche_diff}). 
-But how do ATA effects then happen? We here extend the intuition
+environmental fluctuation (SI section \ref{SIsect:temporal_niche_diff}), 
+how do ATA effects then happen? We here extend the intuition
 introduced in the Discussion and provide examples.
@@
 both lower than average, are not necessarily equivalent.
-Storage can occur if population
+Recall that storage of population growth can occur if population
 increases during good years outweigh decreases during bad years. For instance,
@@
 unfavorable, the expected value $\E[r_i(E_i,C_{i \bs i})]$ in (\ref{eq:expri})
-will be positive. But the potential for such storage of population growth
+will be positive. But the potential for such storage
 may differ between when $E_i$ and $C_{i \bs i}$ are both high compared
@@
 contribute to storage
-effects for the beta-fecundities lottery model are those for which the environment
-is actually poorer than average, because, for those years, competition can be very low,
+effects for species $i$ for the beta-fecundities lottery model are those for which the environment
+is actually poorer than average for $i$, because, for those years, competition can be very low,
 so that the ratio $E_i/C_{i \bs i}$ is large. Years for which the
-environment is exceptionally good actually cannot contribute much to storage effects for 
+environment is exceptionally good for $i$ actually cannot contribute much to storage effects for 
 this model, because for those years, $E_i/C_{i \bs i}$ cannot be very large.
@@
 Counterintuitively, that imbalance can be greater for this model when the environment is 
-sub-optimal.
+sub-optimal for $i$.
 
@@
 to the figure before using it below to further explain
-ATAs. The growth function $\ln(1-\delta+E/C)$ that pertains to the lottery model
+ATA effects. The growth function $\ln(1-\delta+E/C)$ that pertains to the lottery model
 is plotted with colors and contours on Fig. \ref{fig:beta_fecundidities_explanatory}A-C.
@@
 function depicted with color and contours over the points on panels A and C, 
-respectively. The quantity
+respectively (these average values are listed as ``growth'' on panels D and F
+respectively). The quantity
 $\E \left[ \ln \left( 1-\delta+\frac{E_i^{||}}{C_{i \bs i}^{||}} \right) \right]$
 can be well approximated by averaging function 
-values over the points on B. So ATA effects for the model with left-tail association
-between $E_i$ and $C_{i \bs i}$ (respectively, right-tail association) are given by 
-computing the average function value over points on panel A (respectively, C)
-and then subtracting the average function value over points on panel B.
+values over the points on B (the average is listed as ``growth'' on E). 
+So ATA effects for the model with left-tail association
+between $E_i$ and $C_{i \bs i}$ (respectively, right-tail association) are given taking the
+value labelled ``growth'' on panel D (respectively, F) and subtracting the value ``growth''
+on E; a negative result for left-tail associations and a positive result for 
+right-tail associations. 
 Contributions to the storage effect of points in the left portions of 
@@
 The function pictured with a black line on panels D-F is simply the function
-$z \rightarrow \ln(1-\delta+10^z)$ that takes $z=\log_{10}(E_1,C_{1 \bs 1})$
-to $\ln(1-\delta+E_1/C_{1 \bs 1})$, representing growth rate for the model.
+$z \rightarrow \ln(1-\delta+10^z)$ that takes $z=\log_{10}(E_1/C_{1 \bs 1})$
+to $\ln(1-\delta+E_1/C_{1 \bs 1})$, the growth rate for the model.
 Storage effects are manifest by this function, because it is flat on the left and 
@@
 horizontal-axis inputs of the same mean; e.g., the histograms labeled ``left'' on
-E,F. Red dots indicate the mean values of
+E,F. Vertical axis histograms on D-F are the results of applying the 
+function represented by the black line on those panels to the corresponding
+horizontal-axis histograms.
+Red dots indicate the mean values of
 each histogram. The average vertical-axis values of the two red dots on each panel
-equal the GWR of species $i$ for the model (value displayed numerically).
-
-We now relate Fig. \ref{fig:beta_fecundidities_explanatory} back to ATAs. Coexistence is possible for 
+equal the GWR of species $i$ for the model (value displayed numerically,
+labelled as ``growth'').
+
+We now relate Fig. \ref{fig:beta_fecundidities_explanatory} back to ATA effects. Coexistence is possible for 
 right-tail association (panel F, positive GWR) but impossible for left-tail association (panel D,
@@
 annual growth rates, which lead to storage effects; the same effect does 
-not occur for the left-tail case because 
+not occur for the left-tail case (panels A, D) because 
 points in the ``left'' region on panel A are concentrated on line $L_1$ because of the
@@
 $E$ and $C$ align with the portions of $E,C$ distribution for which the potential
-for storage effects is high. Larger ATA effects indicate better alignment.
+for storage effects is high. Larger ATA effects come from better alignment.
 
@@
 the upper and lower tails of the $EC$ distribution (labelled ``right'' and ``left'',
-respectively, on panels A-C). That is why ATA effects are fairly modest for the 
+respectively, on panels A-C). That is why ATA effects are modest for the 
 log-normal fecundities lottery model for the both the left- and right-tail
-associated cases.
+associated cases, and are negligible for the example parameters considered in 
+Fig. \ref{fig:lognormal_fecundidities_explanatory}.
 
@@
 for the model, considering both left- (A, D) and right-tail (C, F) associations
-between $E_i$ and $C_{i \bs i}$. ATA effects are the difference of these two values.
+between $E_i$ and $C_{i \bs i}$. ATA effects are the first of these values minus the second.
 The figure also illustrates how ATAs reflect the degree of alignment between where 

### 2023-07-31 paper/RefereeResponses1.tex
added referee responses that go with the recent addition of intuitive explanations of ATAs

@@
 
-\noindent ***\emph{We have addressed the referee's specific suggestions below. We have also redoubled our
-efforts to make the manuscript as accessible as possible.} \\
+\noindent ***\emph{We have gone to great lengths to address the referee's specific suggestions below, and we have also redoubled our
+efforts to make the manuscript as accessible as possible, generally. We appreciate being spurred to do so, thanks. } \\
 %We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
@@
 
+%I wrote the following, but decided not to include it here. Might include it elsewhere:
+%The research field
+%surrounding the MCT and mechanisms of coexistence is very important within ecology, but is nevertheless technical - see, for instance, the
+%papers of Chesson, which have had enormous positive impact on the field, as well as the recent papers of Ellner which helped inspire
+%our paper. All those papers, and many other papers in the field, are unavoidably technical, but have frequently been published
+%in mainstream journals (the Ellner papers are in Ecology Letters) because of their importance and the importance of the subject
+%matter they investigate. Though we agree with all efforts to make research results as clear as possible to non-technical audiences,
+%we beg the referee
+
 \noindent C37) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
@@
 
-\noindent ***\emph{response} \\
-%Jasmin, let's discuss. I cannot, myself, say, even now, precisely *how* or *why* ATAs promote or hinder coexistence. 
-%In my view, the paper is about showing that they CAN promote or hinder coexistence, and to what extent, so that 
-%future work can continue to explore when they do and when they don't, and precisely how. So I want to write a response
-%to that effect. However, perhaps you have an intuition, which can be simply explained, as to how ATAs
-%either promote or hinder coexistence. I know that's maybe a lot to ask, but if we have such a thing we should consider
-%putting it in. Ecologists always want a two-sentence, math-less explanation for everything. If quantum mechanics were
-%part of ecology they'd want a two-sentence intuitive explanation for it, even if it was meaningless aside making
-%the reader feel better about themselves. Sometimes such a thing does not exist, but when it does exist it can,
-%at least sometimes, be useful. For the record, I think most intuitive, verbal explanations of storage effects
-%are mostly useless, but perhaps we can do better than those.
-%
-%Dan and Jasmin to discuss further. Jasmin to dig out old paragraph of text. Jasmin to work on this.
+\noindent ***\emph{The referee appears to want us to present an intuitive explanation for how ATA effects work, and that 
+is certainly idea. We have gone to great lengths to do so, and think we have succeeded. Though the referee's comment above may 
+be interpreted to suggest that they wanted us
+to add such an explanation to first paragraph of the paper, the reader is not equiped to understand the explanation at that 
+point, so we added a new paragraph later in the Introduction (it is now the
+second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added a new paragraph to the 
+Discussion (the second paragraph of the Discussion). These paragraphs are also supported by two new SI sections and two new SI figures which 
+explain, in detail, how the intuition established in the two new main text paragraphs emerges from the modelling (SI sections S11 
+and S12, SI Figs S4 and S5).
+%DAN: make sure these are the final numbers of these SI sections
+We believe that intuition such as the new material establishes should always be based in formal theoretical and 
+empirical work, so the focus of the paper remains on our expansion of the MCT and our results showing
+the *importance* of ATA effects. Though we definitely think our additions giving an intuition for ATA effects are a large
+improvement in the manuscript, so we thank the referee for spurring us to make these additions.  } \\
+%DONE except the checking as noted above.
 
-\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ...
+\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ... \\
 
@@
 
-\noindent ***\emph{response} \\
-%Same comment again, Jasmin can work on it.
+\noindent ***\emph{See our response to C37. We have added substantial material to the Introduction and Discussion that adds an
+intuitive explanation for ATAs, and we thank the referee for spurring us to do so. 
+We find it unlikely that any focussed reader will have substantial difficulties understanding the lottery model
+material in the Introduction, since the lottery model is certainly no more complicated than models such as the Lotka-Volterra 
+competition and predator-prey models, which are covered in essentially every introductory undergraduate ecology class, and we can assume
+that almost all readers of Ecology Letters have taken such a class. See also comment C25 above, which indicates that another
+referee found this material to be compelling.} \\
+%DONE
 
@@
 
-\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion why ATAs play a major role with beta fecundities and not with log-normal \\
+\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken 
+into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion 
+why ATAs play a major role with beta fecundities and not with log-normal.
+The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical
+details are left for the supplementary material, can the authors explain more in the detail in the main text how the 
+importance of ATAs can be used for this particular model.\\
 
-\noindent ***\emph{<This is a good question. We need to find a place where the models are presented to motivate
-them, AND a place later, in the Discussion (or a new SI section called ``Additional discussion points'') where
-we talk about why ATAs matter for one but not the other. I think this is just because the
-effects of extremes are accentuated in the beta one, and so the effects of associations in the extremes are
-accentuated. But that's not unreasonable, since many species have environments which fluctuate between
-being suitable and unsuitable.>} \\
-%Jasmin to consider this while brainstorming the prior comment. Dan can help, or do this one, but 
-%I'll wait until after we've met and had a chance to discuss intutive explanations of ATA effects
-%that Jasmin has been working on.
+\noindent ***\emph{This is certainly a valid question, and we have gone to some lengths to address it. 
+Partly in response to this comment and partly in response to earlier comments of referee 3, 
+we added substantial new material providing intuitive explanations of ATA effects in the Introduction,
+Discussion, and SI (see above responses). That new material also has information comparing the models
+and explaining why ATA effects are substantial for the beta-fecundities lottery model and not as strong
+for the log-normal fecundities lottery model. We believe additional explanations and intuitive 
+explorations could and should be the subject of several future papers, but we've added material to this manuscript
+that takes that process as far as possible within the scope of the current paper, given length limits and
+requests from the referees to simplify.} \\
+%DONE
 
-\noindent C42) The model used for the chemostat is in its formulation quite different compared to the lottery model.
-Although the technical details are left for the supplementary material, can the authors explain more in the detail in the main text how the importance of ATAs can be used for this particular model.
-\\
-
-%This referee is asking us to explain a lot of things in more detail. We have to be careful coming across as
-%though we have written a paper ``not for a mainstearm audience'', but we can decline some of these requests
-%at least for the main text) as not possible because of space constraints. This particular request does not 
-%make perfect sense because of grammatic problems. But overall the referee seems to want us to explain the
-%effect intuitively, whereas all we've really done is demonstrate its importance. If we want to insert an intuitive 
-%explanation, that'd be more than we've tried for. We should talk about whether we have or could make
-%something like that.
-\noindent ***\emph{response} \\
-%Jasmin to also lump this in with her thinking about the earlier ref 3 comments. 
-%Q: When you look at E and C marginals for the diatom model, do they have most of their mass in the tails, like the 
-%beta-fecundidies lottery model, or most of their mass in the middle, like the normal-fecundities lottery model?
-%Let's talk more about this after Jasmin communicates to Dan the work she's done explaining the ATA effect
-%intuitively.
-
-\noindent C43) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
+\noindent C42) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
 this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones.

### 2023-07-31 paper/RefereeResponses1.tex
working on brevity and clarity throughout. also added Jasmin's water limitation example to the Intro

@@
 \noindent ***\emph{The referee appears to want us to present an intuitive explanation for how ATA effects work, and that 
-is certainly idea. We have gone to great lengths to do so, and think we have succeeded. Though the referee's comment above may 
+is certainly a good idea. We have invested substantial efforts into doing so, and think we have succeeded. Though the referee's comment above may 
 be interpreted to suggest that they wanted us
@@
 point, so we added a new paragraph later in the Introduction (it is now the
-second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added a new paragraph to the 
-Discussion (the second paragraph of the Discussion). These paragraphs are also supported by two new SI sections and two new SI figures which 
+second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added some text to the end of
+the second paragraph that adds to the intuitive understanding of how ATAs can influence coexistence through a concrete 
+example. We also added a new paragraph to the 
+Discussion (the second paragraph of the Discussion). These new Discussion 
+paragraphs are also supported by two new SI sections and two new SI figures which 
 explain, in detail, how the intuition established in the two new main text paragraphs emerges from the modelling (SI sections S11 
@@
 
-\noindent ***\emph{Thank you for the positive feedback!} \\
+\noindent ***\emph{Thank you for the positive feedback! We have also added some text to the end of that paragraph giving a 
+concrete case (water limitation) in which one may expect an influence of ATAs. We think the new text definitely
+makes it more intuitive reasonable to explore ATA contributions to conexistence.} \\
 %DONE
@@
 
-\noindent ***\emph{This is certainly a valid question, and we have gone to some lengths to address it. 
+\noindent ***\emph{This is certainly a valid question, and we have gone to substantial lengths to address it. 
 Partly in response to this comment and partly in response to earlier comments of referee 3, 
@@
 and explaining why ATA effects are substantial for the beta-fecundities lottery model and not as strong
-for the log-normal fecundities lottery model. We believe additional explanations and intuitive 
-explorations could and should be the subject of several future papers, but we've added material to this manuscript
-that takes that process as far as possible within the scope of the current paper, given length limits and
-requests from the referees to simplify.} \\
+for the log-normal fecundities lottery model. These are very positive additions to the manuscript, so we 
+thank the referee for spurring us to add them. We also believe this is only a start -- additional explanations and intuitive 
+explorations of ATA effects could and should be the subject of future papers. We've added as much material 
+as possible on these topics to the current manuscript, given length limits and
+requests from the referees to both simplify and add a variety of other content.} \\
 %DONE
@@
 %one of the referees (ref 1) actually asked us to put BACK technical content in the main text which we had 
-%moved to the SI before submission to keep it less technical. That said, we do need to do a pass thu and 
-%figure out what technical content we can reduce. 
+%moved to the SI before submission to keep it less technical. Also, this research area is technical, though it
+%is still central to ecology as a discipline. Recycle some of your arguments above about technical aspects. 
+%That said, we do need to do a pass thu and figure out what technical content we can reduce, so we can make it
+%as clear as it can be. Note also they say the math *could* be moved, suggesting less importance than 
+%using ``a more biological/ecological language...''. So maybe let's try to do more of that, some of which
+%I've alreday done thru answering the referee comments.
 \noindent ***\emph{response} \\

### 2023-07-31 paper/Paper.Rnw
working on brevity and clarity throughout. also added Jasmin's water limitation example to the Intro

@@
 Ecologists may ask, for instance, how local environment relates to 
-population growth rate. To study such relationships, ecologists 
-very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
-and direction of association between variables often do not tell the whole story 
+population growth rate. To study relationships, ecologists 
+very commonly use metrics such as correlation. But standard measures of the strength 
+and direction of association between variables do not tell the whole story 
 \citep{Ghosh_2020_AER}.
-To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
+To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c have the same 
 Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
-symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has previously been called 
+symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what has been called 
 ``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
-Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
-regression), but correlations are common, and other common methods also reveal only a small portion 
+Metrics of association other than correlations are also often used (e.g., various forms of 
+regression), but other common methods also reveal only a small portion 
 of the information content of the relationship between variables \citep{Nelsen2006}.
-ATAs and related ideas had previously been seldom considered in ecology, but recently several 
+ATAs and related ideas had previously been seldom considered in ecology, but recently 
 insights have been gained by studying how features of associations beyond standard measures 
@@
 across component populations
-than are years of good environment, limiting the potential for rescue effects. 
+than are years of good environment, limiting rescue effects. 
 And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
 ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
-plankton population density time series; many other ecological examples of ATAs have also been explored 
+plankton population density time series; many other examples of ATAs in nature have also been explored 
 \citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
-competition and coexistence have not been investigated. Doing so is the goal of this study.
+competition have not been investigated. Doing so is our goal.
 
 %Intro to competition, coexistence, MCT, etc.
-Competition, coexistence, and the maintenance of biodiversity are research topics at the core of modern ecology 
+Competition and coexistence are research topics at the core of modern ecology 
 \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
 that two species competing for the same limiting resource cannot coexist at constant densities - one will 
-instead outcompete the other and drive it extinct \citep{Hardin_1960}.
+instead drive the other extinct \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and will be
-defined formally in the Theory section. MCT has been
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and 
+defined formally in Theory. MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
@@
 %Intro to storage effects and why they might be influenced by ATAs
-It is possible to see why it is plausible that
-ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
-what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
-may be relevant to them. Storage effects allow competing species to fluctuate in abundance, without any going extinct,
+We indicate why it is plausible that
+ATAs may influence one mechanism of coexistence in particular, storage effects, starting by conceptually reviewing 
+what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}. 
+Storage effects allow competing species to fluctuate in abundance, without any going extinct,
 by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
 Storage effects hinge on covariation between the fluctuating quality of the environment
-for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
-and on the assumption that environmental responses for distinct species are also distinct,
-so that no two species respond in the same way to environmental variation. Then, good environmental
+for a species ($E$) and the strength of competition ($C$) experienced by that species (called $EC$ covariance);
+and on the assumption that no two species respond in the same way to environmental variation. Then, good environmental
 conditions for a currently common species are paired with strong intraspecific competition, because of
-positive EC covariance and the commonness of the species. But good environmental conditions for 
+positive $EC$ covariance and the commonness of the species. But good environmental conditions for 
 a currently rare species are not as closely accompanied by strong competition because the rarity of
 the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
-rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
-differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
-ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
-and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.
+rare species. Storage effects can be quantified by measuring to what extent $EC$ covariance contributes to 
+differences between growth rates of currently rare versus currently common species \citep{Chesson_1994,Ellner_2016}.  
+ATAs may, in principle, influence storage effects simply because storage effects hinge on $EC$ covariation, 
+and ATAs can be an aspect of $EC$ covariation. One scenario in which it is reasonable to imagine that
+ATAs may be present in $EC$ covariation pertains to water limitation. Suppose two plants compete primarily for 
+water, when water is scarce, but their niches become more differentiated when water is abundant.
+Then the quality of the environment, $E$, as experienced by whichever plant is scarce, may only correspond closely to the 
+competition, $C$, imposed on it by the other, currently abundant plant, when water is scarce and $E$ is low.
 
 %Lottery model and results based on it
-To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
-simple simulations using a two-species model, the well-known lottery model. The simulations described in this
-paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
-precise nature of that influence is the subject of the rest of the paper.
-The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
+To further demonstrate the plausibility of ATAs altering coexistence, we carried out 
+simple simulations using a two-species model, the lottery model. The simulations described in this
+paragraph demonstrate that ATAs influence coexistence in some manner; the
+nature of that influence is the subject of the rest of the paper.
+The lottery model, which was introduced as a model of coral reef fish competing for a set of $N$ territories, 
 is as follows.
@@
 in proportion to the reproductive outputs of the two species that year. Note that $N = N_1(t)+N_2(t)$ 
-is constant through time. 
-We assume for simplicity that the random variables $(B_1(t),B_2(t))$ are independent and identically 
+is constant. 
+We assume that the random variables $(B_1(t),B_2(t))$ are independent and identically 
 distributed (iid) through time. In three distinct simulations (Fig. \ref{fig:pedagog1}f-h), 
@@
 left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symmetrically (Fig. \ref{fig:pedagog1}b)
-tail associated, while the same marginal distributions for these fecundities were used in all 
+tail associated, while the same marginal distributions were used in all 
 simulations (see Methods
-for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
+for details). Thus, the setup was identical in the three simulations except for different ATAs. 
 ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
-Notation for the whole paper is summarized in Table \ref{tab:notation}.
+Notation for the paper is summarized in Table \ref{tab:notation}.
 Typesetting of mathematics by ecological journals is sub-optimal,
 so we appended a latex version of this main text to the end of the 
-Supporting Information (SI). Readers interested in the mathematical aspects should read that version.
+Supporting Information (SI); readers interested in the math should read that version.
 %***DAN: Don't forget to do this
@@
 We next give some initial intuition about how ATA effects on coexistence work, 
-providing additional intuition in the Discussion.
+with more details in the Discussion.
 It is known that storage effects can occur if a buffering mechanism reduces population 
@@
 positive effects of $E$; but this can occur
-both 1) when $E$ is moderately favorable but $C$ is very strong, as well as 2) when $E$ is 
+both 1) when $E$ is moderately favorable but $C$ is very strong, and 2) when $E$ is 
 unfavorable and $C$ is not too weak.
@@
 but this can occur 1) when $E$ is favorable and $C$ is not too strong,
-as well as 2) when $E$ is moderately unfavorable but $C$ is very weak. 
+and 2) when $E$ is moderately unfavorable but $C$ is very weak. 
 It turns out that the potential for storage events to occur when 
@@
 \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}; and our arguments above
-indicate that ATAs may influence species coexistence. Thus, we here seek to quantify the consequences 
-of ATAs for competition, coexistence, and the maintenance of biodiversity, by means of the 
-following specific goals. 
+indicate that ATAs may influence coexistence. Thus, we here seek to quantify the consequences 
+of ATAs for competition and coexistence, via the following specific goals. 
 G1) We augment MCT to show formally how ATAs play a role in coexistence. 
@@
 We address how ATAs contribute to these components. 
-%We examine mechanisms of coexistence that apply in models exhibiting solely temporal variation; 
-%spatial variation can be considered in future work.
-G2) We explore, using the simple lottery model, the relative magnitudes 
+G2) We explore, using the lottery model, the relative magnitudes 
 of the contributions to coexistence that come
@@
 sometimes coexist with ATAs but not without, or vice versa. This is to help 
-illuminate whether ATA contributions to coexistence and biodiversity maintenance
+illuminate whether ATA contributions to coexistence 
 are likely to be negligible or meaningfully large. 
 G3) We provide an empirical example of species coexistence which highlights the role ATAs can play and the
-potential importance of ATAs relative to other mechanisms of coexistence.
-The example is a diatom system in a chemostat. In the Discussion, we consider whether climate change may alter
-ATAs of environmental variables and thereby influence competition and coexistence.
+potential importance of ATAs.
+The example is a laboratory diatom system. In the Discussion, we consider whether climate change may alter
+ATAs of environmental variables and thereby influence competition.
 Overall, our study presents a new mechanism of species coexistence
-and a means of understanding its theoretical and empirical importance.
+and a means of understanding its importance.
 
@@
 
-We here pursue goal G1 from the Introduction: to augment MCT to show formally how ATAs play a role in coexistence.
+We here pursue goal G1 from the Introduction: to augment MCT to show the role of ATAs for coexistence.
 
@@
 
-We develop our new theory for a general two-species model, later applying it to specific scenarios.
-Our mathematical notation parallels
+We first develop our new theory for a general two-species model.
+Our notation parallels
 that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
 of species $i$ can be written as an increasing function of
-an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
+an environmental factor $E_i(t)$ and as a decreasing function of 
 competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption is probably not necessary, it simplifies the
+$C_i(t)$. Though the iid assumption is probably unnecessary, it simplifies the
 presentation of new ideas.
@@
 
-%ATA specialists
-Contributions of ATAs to species coexistence open the possibility that
-``ATA specialists'' may exist. For instance, if a strong competitor is present across a
-collection of habitat patches on a landscape and the GWR of a weaker competitor is only positive 
-in a patch if there are 
-ATAs between the temporal fluctuations of two environmental variables in that patch,
-then the weaker competitor can only persist in patches for which ATAs are present.
-The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
-can boil down to associations between two environmental variables, as was the case
-for the lottery model. For the lottery model, $B_i(t)=E_i(t)$ was assumed, and $C_{i \bs i}$ simplified to $E_j(t)/\delta$ (see SI section 
-\ref{SI-SIsect:TheoryForLotteryBoth}). Thus, EC-associations parallel associations between the two
-environmental variables $E_i(t)$ and $E_j(t)$ for the lottery model, and ATA specialists may exist. 
-Mechanisms of coexistence should correspond to
-means by which niche differentiation is possible. Thinking about our new mechanism of 
-coexistence in terms of possible ATA specialists helps highlight this connection. Likewise,
-because we have described ATA exclusion as well as ATA rescue, it is possible to 
-imagine a scenario in which a weaker competitor can only exist in habitat patches 
-for which ATAs between environmental variables are absent, and is hence a specialist on scenarios with symmetric
-tail associations of variables (an ``STA specialist''). Additional 
-research will be needed to assess to what degree these possibilities are realized 
-in nature.
-
 %Climate change may alter ATAs of coexistence-relevant ATAs
@@
 of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.
-
-%Differences with the original, analytic version of the MCT
-Our approach to exploring the influence of ATAs on coexistence relies on the computational
-extension of MCT of \cite{Ellner_2019},
-and our results help demonstrate strengths of that approach.
-Storage effects as computed using the original, analytic approach to MCT 
-\citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
-which we refer to as the contributions to coexistence of correlation \emph{per se}.
-This is because the weak-noise/Taylor approximation approach used in the original 
-MCT neglects the effects of ATAs when higher-order terms
-in Taylor expansions are dropped. The term $\Delta^{(E \shp C)}$, which our theory 
-inherits from \cite{Ellner_2016}, is also dropped. These observations do not diminish 
-the major strengths of the 
-original, analytic approach, which are well known and also elaborated by \cite{Ellner_2019};
-one such strength is the generality that analytic approaches provide.
-It may be possible to expand the mathematics of the original MCT to consider ATAs.
-Additional discussion points are in SI sections \ref{SI-sect:vital_rates}-\ref{SI-sect:storage_importance}.
+Additional discussion points are in SI sections \ref{SI-SIsect:ATA_specialists}-\ref{SI-sect:storage_importance}.
 
@@
 
-\noindent Figure 4. ATA effects and ATA rescue in the diatom system. 
+\noindent Figure 5 caption. ATA effects and ATA rescue in the diatom system. 
 The contribution of ATAs to the GWR of \emph{F. crotonensis}

### 2023-07-31 paper/SupMat.Rnw
working on brevity and clarity throughout. also added Jasmin's water limitation example to the Intro

@@
 
+\section{   ATA specialists}\label{SIsect:ATA_specialists}
+
+Contributions of ATAs to species coexistence open the possibility that
+``ATA specialists'' may exist. For instance, if a strong competitor is present across a
+collection of habitat patches on a landscape and the GWR of a weaker competitor is only positive 
+in a patch if there are 
+ATAs between the temporal fluctuations of two environmental variables in that patch,
+then the weaker competitor can only persist in patches for which ATAs are present.
+The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
+can boil down to associations between two environmental variables, as was the case
+for the lottery model. For the lottery model, $B_i(t)=E_i(t)$ was assumed, and $C_{i \bs i}$ simplified to $E_j(t)/\delta$ (see SI section 
+\ref{SIsect:TheoryForLotteryBoth}). Thus, EC-associations parallel associations between the two
+environmental variables $E_i(t)$ and $E_j(t)$ for the lottery model, and ATA specialists may exist. 
+Mechanisms of coexistence should correspond to
+means by which niche differentiation is possible. Thinking about our new mechanism of 
+coexistence in terms of possible ATA specialists helps highlight this connection. Likewise,
+because we have described ATA exclusion as well as ATA rescue, it is possible to 
+imagine a scenario in which a weaker competitor can only exist in habitat patches 
+for which ATAs between environmental variables are absent, and is hence a specialist on scenarios with symmetric
+tail associations of variables (an ``STA specialist''). Additional 
+research will be needed to assess to what degree these possibilities are realized 
+in nature.
+
+\section{   Differences with the original, analytic version of MCT}\label{SIsect:SimsVAnalysis}
+
+Our approach to exploring the influence of ATAs on coexistence relies on the computational
+extension of MCT of \cite{Ellner_2019},
+and our results help demonstrate strengths of that approach.
+Storage effects as computed using the original, analytic approach to MCT 
+\citep{Chesson_1994,Chesson_2000} should equal our term $\Delta^{[E \nat C]}$, 
+which we refer to as the contributions to coexistence of correlation \emph{per se}.
+This is because the weak-noise/Taylor approximation approach used in the original 
+MCT neglects the effects of ATAs when higher-order terms
+in Taylor expansions are dropped. The term $\Delta^{(E \shp C)}$, which our theory 
+inherits from \cite{Ellner_2016}, is also dropped. These observations do not diminish 
+the major strengths of the 
+original, analytic approach, which are well known and also elaborated by \cite{Ellner_2019};
+one such strength is the generality that analytic approaches provide.
+It may be possible to expand the mathematics of the original MCT to consider ATAs.
+
 \section{   Asymmetric associations between population vital rates}\label{sect:vital_rates}

### 2023-08-01 paper/Paper.Rnw
continued trimming and simplifying

@@
 regression), but other common methods also reveal only a small portion 
-of the information content of the relationship between variables \citep{Nelsen2006}.
+of the information content of the relationship \citep{Nelsen2006}.
 ATAs and related ideas had previously been seldom considered in ecology, but recently 
 insights have been gained by studying how features of associations beyond standard measures 
-impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
+impact ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
 Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
@@
 \citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
-competition have not been investigated. Doing so is our goal.
+competition have not been investigated. Doing so is our goal here.
 
@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as ``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and 
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
+``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and 
 defined formally in Theory. MCT has been
@@
 water, when water is scarce, but their niches become more differentiated when water is abundant.
-Then the quality of the environment, $E$, as experienced by whichever plant is scarce, may only correspond closely to the 
-competition, $C$, imposed on it by the other, currently abundant plant, when water is scarce and $E$ is low.
+Then the quality of the environment, $E$, as experienced by whichever plant is currently rare may only correspond closely to the 
+competition, $C$, imposed on it by the currently abundant plant when water is scarce and $E$ is low.
 
@@
 storage potential is high, ATA effects will be high. ATA effects quantify how temporal niche partitioning
-varies across $EC$ spectrum spanned by the cases 1-2 sketched above.
+varies across the $EC$ spectrum spanned by the cases 1-2 sketched above.
 
@@
 
-\subsection{Theoretical setup}
-
-We first develop our new theory for a general two-species model.
+\subsection{Setup}
+
+We first develop our theory for a general two-species model.
 Our notation parallels
@@
 that the $E_i(t)$ are iid through time, and likewise for the
-$C_i(t)$. Though the iid assumption is probably unnecessary, it simplifies the
+$C_i(t)$. Though this assumption is probably unnecessary, it simplifies the
 presentation of new ideas.
@@
 
-MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
+MCT and its computational extensions \citep{Ellner_2016,Ellner_2019} 
 quantify the contributions of 
 multiple mechanisms to coexistence. But we show below
-that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
+that only one mechanism relates to ATAs: storage effects. 
 So we begin by defining storage effects. 
@@
 \end{equation}
-where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' (GWR) of species $i$, that is, the mean growth rate of 
+where: $\olr_{i \bs i}$ is the mean ``growth when rare'' (GWR) of species $i$, that is, the mean growth rate of 
 species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
-$\olr_{j \bs i}$ refers to the mean steady-state 
+$\olr_{j \bs i}$ is the mean steady-state 
 growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
-influences of environment and competition on either species have been rendered independent.  
+influences of environment and competition have been rendered independent.  
 A subscript 
@@
 \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
-alternative definitions of $q_{ij}$ that we use, but one definition is $q_{ij}=1$, and to understand the new
+alternative definitions of $q_{ij}$ that we use; but one definition is $q_{ij}=1$, and to understand the new
 ideas presented here it is sufficient to use that definition. 
@@
 differences between rare and common species mean growth rates for the original model and for a 
-modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
-were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
-Because these components were rendered 
+modified model for which $E$ and $C$
+were rendered independent [see (\ref{eq:storage_effects_gen})]. 
+Because components were rendered 
 completely independent in the modified model, storage effects quantify the contribution of the 
-whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
+whole $EC$ relationship to the rare/common species growth rate
 difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
@@
 marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
-and with some but not all aspects of the relationship 
-between the components altered relative to the original distributions. Specifically, asymmetries of 
-tail association are removed -- see below for how this is done. We then define 
+and with some but not all aspects of the $EC$ relationship 
+altered. Specifically, ATAs are removed -- see below for how. We then define 
 $\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
@@
 \end{align}
-The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
-ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
-the \emph{ATA effect on coexistence}. 
+The term $\Delta_i^{[EC]}$ is the \emph{ATA effect on coexistence}. 
 The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
 correlation \emph{per se}, remaining after having removed the effects of ATAs.
-Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly in our notation,
+Note that brackets, $[\cdot]$, and parentheses, $(\cdot)$, are used distinctly
 throughout the paper - they are not interchangeable.
-The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
-corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
-The new symbol should correspondingly be pronounced ``partial sharp''.
-See SI section \ref{SI-SIsect:techcomm} for technical comments on the new notation. 
+The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (two of its four lines), 
+corresponding to the removal of part of the relationship between $E$ and $C$ components (ATAs).
+The symbol should correspondingly be pronounced ``partial sharp''.
+See SI section \ref{SI-SIsect:techcomm} for technical comments on the notation. 
 The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
 (\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
-of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
+of storage effects, which are a term in the MCT decomposition \citep{Ellner_2019} -- see 
 below for additional details.
@@
 We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
-$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
+$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, are strictly monotonic and continuous. 
 Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
@@
 distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
-Let $\rho_i$ be the covariance of
-these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
+Let $\rho_i$ be their covariance. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
 $\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
@@
 and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
-The value of $\ddot{\rho}_i$ is then determined in one of 
+The value of $\ddot{\rho}_i$ is determined in one of 
 four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
@@
 The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
-$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
-$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
-as desired. The question of the uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
+$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is preferable. Thus $E_i^\nat$ and 
+$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se}. 
+The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
 is addressed in SI section \ref{SI-SIsect:uniqueness}.
@@
 the distribution $(E_i^\shp,C_{i \bs i}^\shp)$
-defined in the previous section (following \cite{Ellner_2016}).
+defined above (following \cite{Ellner_2016}).
 
@@
 
-The other mechanisms in MCT that apply in our modelling context of purely temporal variation
-make no reference to covariation between quantities, whereas ATAs are 
+The other mechanisms in MCT make no reference to covariation between quantities, whereas ATAs are 
 strictly about covariation. Therefore, storage effects are the only mechanism of coexistence
-for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
-\ref{SI-SIsect:other_mechanisms}. \cite{Ellner_2016} and
-\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
-which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
-original decomposition of \cite{Chesson_1994}, 
+for which ATAs are relevant. The \cite{Ellner_2019} version of MCT comprises 
+a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
+which equals GWR of species $i$, as 
 \begin{equation}
@@
 which is the extension of MCT that was goal G1 of the Introduction. 
+The derivation of this overall decomposition is in SI section 
+\ref{SI-SIsect:other_mechanisms}.
 
@@
 To evaluate the importance of ATAs for a given scenario,
-one can quantify
+we quantified
 the magnitude of the new ATA effect term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
-scenario.  One typically (but see below) must examine this quantity for each species, $i$,
-to obtain complete information about coexistence, since coexistence requires both species populations to 
-be able to grow when rare. We also defined scenarios of ``ATA rescue''
-and ``ATA exclusion''. \emph{ATA rescue} is considered to have occurred for a given scenario if coexistence
+scenario (see Methods). But we also defined scenarios of ``ATA rescue''
+and ``ATA exclusion''. \emph{ATA rescue} is considered to have occurred if coexistence
 is possible when ATAs are taken into account, i.e., GWR$>0$ for both species, but impossible otherwise, i.e.,
@@
 %Setup - two versions, and basics which apply to both
-To achieve goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
+For goal G2 of the Introduction, we applied our theory to the lottery model, 
 comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
 of the other terms of 
-the decomposition (\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
+(\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
 of the lottery model. The versions both used the same dynamical equations
-(Introduction) and assumed temporally iid fecundities, but the
-versions differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
-transformations of the distributions $(b_1,b_2)$ pictured in Fig. \ref{fig:pedagog1}a-c;
+(Introduction), but 
+differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
+transformations of the distributions $(b_1,b_2)$ of Fig. \ref{fig:pedagog1}a-c;
 the details of how the left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c), 
@@
 and therefore had parameters $\sigma$, $\mu_1$ and $\mu_2$ in addition to the death rate parameter $\delta$
-(Introduction) that both versions of the lottery model used. The parameter $\sigma$ is the standard deviation of the
+(Introduction). The parameter $\sigma$ is the standard deviation of the
 log-fecundities and $\mu_i$ is the mean log fecundity for species $i$.
@@
 distribution and $F_\beta$ is the CDF of a beta distribution 
-with shape and scale parameters $0.5$ (probability density function proportional to $x^{-0.5}(1-x)^{-0.5}$ on
-the unit interval). So the beta fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
+with shape and scale parameters $0.5$. So the beta fecundities model had parameters $\eta_1$, $\eta_2$ and $\delta$,
 and produced fecundities with $U$-shaped distributions between $0$ and $\eta_1$, for $B_1$, and
-$\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
-corresponded to more variable fecundities for species $i$. The parameter $\eta_i$ represents the upper
-bound on what fecundities could occur for species $i$.
+between $0$ and $\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
+corresponded to more variable fecundities for species $i$.
 
@@
 the environment, i.e., the $E_i$ of Theory was set equal to $B_i$. 
-Competition, $C_i(t)$, in the lottery model occurs at the stage of juveniles occupying open sites.
-For both model versions, $C_i$ was therefore taken to be the number of new offspring
+Competition, $C_i(t)$, in the lottery model occurs at the stage of juveniles occupying open sites:
+$C_i$ was taken to be the number of new offspring
 divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
 does not depend on $i$. These choices for $E_i(t)$ and $C_i(t)$ fulfill the assumption of Theory 
-that the growth rate, $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
-(see SI section \ref{SI-SIsect:TheoryForLottery} for details).
+that $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
+(SI section \ref{SI-SIsect:TheoryForLottery}).
 

### 2023-08-01 paper/Paper.Rnw
mostly trimming to get under the 5000 word length for the main text

@@
 For both versions of the lottery model, 
-fecundities were taken to directly reflect 
+fecundities directly reflected 
 the environment, i.e., the $E_i$ of Theory was set equal to $B_i$. 
@@
 To explore, using the log-normal fecundities lottery model, the relative magnitudes of the contributions 
-to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
-(\ref{eq:full_decomp}) for that model against the model parameters $\mu_1$, $\mu_2$, $\sigma$ and $\delta$.
+to coexistence of ATAs versus other mechanisms, we plotted the terms of the decomposition 
+(\ref{eq:full_decomp}) against model parameters $\mu_1$, $\mu_2$, $\sigma$ and $\delta$.
 We labeled regions of the plot which corresponded to ATA rescue or exclusion, i.e., regions for which
@@
 it was then sufficient to decompose GWR of the weaker competitor, species 1. 
-Second, we showed that, for the log-normal fecundities
-lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
+Second, we showed that none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
 instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
@@
 for both our left- and right-tail associated noises (SI section \ref{SI-SIsect:TheoryForLottery}). 
-So we only generated plots 
-for left-tail associated noise.
+So we only generated plots for left-tail associated noise.
 
 For each combination of $\delta=0.2, 0.4, 0.6$ and $\mu_1-\mu_2=0,-0.5,-2,-4$,
-we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ ranging 
+we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ 
 from $0$ to $7$. \cite{Chesson_1981} showed that coexistence is impossible for $\delta=1$, and it
-is known that the storage effect (and therefore its ATA component) is weak for larger $\delta$, so we 
-only considered values of $\delta$ up to $0.6$. Some of 
-the $\mu$ differences and $\sigma$ values we considered were very unrealistic (see Results),
-so we considered after plotting whether conclusions about the importance of ATAs depended 
+is known that the storage effect is weak for larger $\delta$, so we 
+only considered $\delta$ up to $0.6$. Some of 
+the $\mu$ differences and $\sigma$ values we considered were unrealistic (see Results),
+so we considered after plotting whether conclusions depended 
 on realistic values of parameters.
-The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
+Simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
 Computationally efficient means of performing those simulations are in 
@@
 To explore, using the beta fecundities lottery model, the relative magnitudes of the contributions 
-to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
-(\ref{eq:full_decomp}) for that model against the model parameters $\eta_1$, $\eta_2$, and $\delta$,
-again labelling regions corresponding to ATA rescue or exclusion.
-Some mathematical shortcuts, summarized here, simplified plotting. 
-First, we assumed, without loss of generality, that $\eta_1 \leq \eta_2$. To understand coexistence, 
-it was then sufficient to decompose GWR of the weaker competitor, species 1. 
-Second, we showed that, for the beta fecundities
-lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
+to coexistence of ATAs versus other mechanisms, we plotted the terms of the decomposition 
+(\ref{eq:full_decomp}) against model parameters $\eta_1$, $\eta_2$, and $\delta$,
+again labelling regions of ATA rescue or exclusion.
+Mathematical shortcuts, summarized here, simplified plotting. 
+First, we assumed, without loss of generality, that $\eta_1 \leq \eta_2$. 
+It was then sufficient to decompose GWR of the weaker competitor, species 1. 
+Second, we showed that none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
 instead they depend only on $\eta_2/\eta_1$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
@@
 the fairly realistic (see Results) range $1 \leq \eta_2/\eta_1 \leq 5$. 
-The simulation methods of \cite{Ellner_2019} were again used.
-Computationally efficient means of performing these simulations are in 
+Simulation methods of \cite{Ellner_2019} were again used, with 
+computationally efficient means of simulating in 
 SI section \ref{SI-sect:efficient_computation}.
-
-%Trait different space stuff. I don't think I need it anymore but keeping around until I finish writing the Discussion,
-%in case some of it turns out to be useful there.
-% To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
-% ATAs but not without, or vice versa, we plotted GWR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
-% as well as
-% GWR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
-% For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
-% we were interested in the values of $\mu_1-\mu_2$ for which one of these
-% quantities is positive and the other is negative; this set of values characterizes the species which can coexist
-% with ATAs, but not without, or vice versa. We define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
-% Points of the trait-difference space characterize possible combinations of traits of the two 
-% competing species in our model. For fixed values of 
-% $\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
-% trait-difference space to be the
-% values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
-% is the subspace for which the two lottery model species coexist in the presence of ATAs.
-% We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the values
-% $\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
-% This is the subspace for which the two lottery model species coexist in the absence of 
-% ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ indicates the
-% absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
-% set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
-% values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
-% We define ``ATA exclusion space'' to be the set difference 
-% $\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
-% i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
-% On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
-% against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
-% $\mu_1 - \mu_2$ for all combinations of $\delta=0.2, 0.4, 0.6, 0.8$ and $\sigma=2,4,5,6$.
 
@@
 Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
-\emph{Cyclotella pseudostelligera} over a range of temperatures to parametrize a resource
+\emph{Cyclotella pseudostelligera} over a range of temperatures to parameterize a resource
 competition model. The model is 
@@
 example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
-$i=2$, $j=1$. Further details of the model setup are in SI section \ref{SI-sect:diatom_model}
-and specifics of how our decomposition extends to this empirical case are in 
+$i=2$, $j=1$. Details of the model setup are in SI section \ref{SI-sect:diatom_model}
+and specifics of how our decomposition was computed are in 
 SI section \ref{SI-sect:Empircal_app}.
@@
 %One-parameters plots
-To display results for goal G3, we started by plotting the terms of 
+To display results for G3, we started by plotting the terms of 
 (\ref{eq:full_decomp}) against one of the parameters 
-$\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
+$\theta_0$, $a$ and $P$ at a time, keeping the other two fixed at the values
 used by \cite{Descamps_2005}. We again labelled regions of parameter space corresponding to
@@
 We also generated two-dimensional plots, where
-two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
-at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
-$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color, to convey the 
-importance of the ATA contribution relative to the whole GWR. Larger magnitudes
+two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed. 
+For those plots, the quantity  
+$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color. Larger magnitudes
 of this quantity indicate greater importance of ATAs. The case
 $\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/|\text{GWR}|>1$)
-corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
+corresponds to ATA rescue, and $\Delta_i^{[EC]}<\text{GWR}<0$
 (which produces $\Delta_i^{[EC]}/|\text{GWR}| < -1$) corresponds to ATA exclusion.
-Thus the value of $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
+Thus $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
 and also whether ATA rescue or exclusion occurred.
@@
 To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
-of the contributions to coexistence that come from ATAs compared to other mechanisms; and to see if ATA rescue 
-or ATA exclusion can occur), Fig.
+of the contributions of ATAs compared to other coexistence mechanisms; and to see if ATA rescue 
+or exclusion occur), Fig.
 \ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
 the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
-magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
+magnitude comparable to the magnitudes of other terms. 
 Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
@@
 were the only means by which coexistence could occur, and ATA effects
-tended to be negative, weakening total storage effects.
-
-The ATA term 
-was also often, or always (for the parameters we considered), comparable in magnitude to the other terms of 
-the decomposition (\ref{eq:full_decomp}) for the beta fecundities lottery model 
-(Fig. \ref{fig:fig4}). The term contributed negatively to coexistence for environmental noise/fecundities
-showing left-tailed association, and contributed positively for noise/fecundities showing right-tailed 
+tended to be negative, weakening storage effects.
+
+For the beta fecundities lottery model, the ATA term 
+was often or always, for the parameters we considered, comparable in magnitude to the other terms of 
+the decomposition (\ref{eq:full_decomp}) (Fig. \ref{fig:fig4}). The term contributed negatively to 
+coexistence for environmental noise/fecundities
+showing left-tailed association, and positively for right-tailed 
 association.
@@
 Our results also show that ATAs can make the difference between coexistence 
-and failure of two species to coexist: both ``ATA rescue'' and ``ATA
+and competitive exclusion: both ``ATA rescue'' and ``ATA
 exclusion'' (see Theory) are possible. Starting with the log-normal fecundities 
 lottery model, for some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
-line and the $x$ axis for some values of $\sigma$, e.g., see panel i.
-Such cases correspond to parameter combinations
-for which the presence or absence of ATAs determine whether the species coexist or not. In the 
-case of panel i, for values of $\sigma$ around $6$, 
-species 1 is excluded with ATAs but the two species can coexist without ATAs. 
-Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model; 
-examples were more common than for the log-normal fecundities model.
-
-For the log-normal fecundities lottery model, the ATA term from (\ref{eq:full_decomp}) tended
-only to be comparable in magnitude to the other terms, and ATA exclusion or rescue tended only to
-occur, for unrealistic parameters values; but ATA terms were large and ATA rescue and exclusion occurred
-for many realistic parameter values for the beta fecundities lottery model. For instance, the ATA term in 
-Fig. \ref{fig:fig3} had fairly 
-small magnitude, compared to the other terms of (\ref{eq:full_decomp}), for values of $\sigma$ less than 
-about $3$. And ATA rescue and exclusion only occurred for $\sigma$ greater than about $3$.
+line and the $x$ axis for some values of $\sigma$, e.g., panel i.
+For such cases, the presence or absence of ATAs determine whether the species coexist or not. 
+For panel i, for values of $\sigma$ around $6$, 
+species 1 is excluded with ATAs but the two species coexist without ATAs. 
+Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model,
+where examples were more common than for the log-normal fecundities model.
+
+Contrasting the two models, ATAs were only important for the log-normal fecundities lottery model 
+when that model was parameterized unrealistically; but ATAs were important for a wide range of 
+realistic parameter values for the beta fecundities model.
+For instance, the ATA term in Fig. \ref{fig:fig3} had large magnitude compared to the other
+terms of (\ref{eq:full_decomp}), and ATA rescue and exclusion occurred, only for 
+$\sigma$ greater than about $3$. 
 The standard deviation of fecundity for species 1 when 
 $\mu_1=0$ and $\sigma=3$ was greater than $8000$, so values of $\sigma$ greater than $3$ are 
-probably unrealistic. Thus ATAs seem to be unimportant for coexistence
-for the log-normal fecundities lottery model with realistic parameter values. However, 
-all the $\eta_2/\eta_1$ ratio values we plotted were fairly realistic, corresponding to
-upper-bound fecundities that were up to $5$ times higher for species 2 than for species 1. And 
-the ATA term plotted in Fig. \ref{fig:fig4} was comparable in magnitude 
-to the other components of (\ref{eq:full_decomp}) across the whole range of parameters
-considered. And both ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
+probably unrealistic. 
+However, all the $\eta_2/\eta_1$ ratio values we plotted for the beta fecundities model 
+were fairly realistic, corresponding to
+upper-bound fecundities that were up to $5$ times higher for species 2 than for species 1.  
+The ATA term in Fig. \ref{fig:fig4} was comparable in magnitude 
+to other components of (\ref{eq:full_decomp}) across the whole range of parameters
+considered. And ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
 Apparently the question of whether ATAs are important for realistic parameter 
@@
 which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
-show that ATA contributions to coexistence often have magnitudes comparable to the 
-magnitudes of other terms of the extended MCT decomposition, and that 
-ATA rescue occurs for some combinations of parameters for this system.
+show that ATA contributions to coexistence often have magnitudes comparable to those 
+of other terms of the extended MCT decomposition, and that 
+ATA rescue occurs for some parameters for this system.
 
 Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f reinforce the 
-same conclusions. The plotted quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
-indicating that ATAs were often a substantial contributor to whether the diatom species 
-coexist. For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
+same conclusions. The quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
+indicating that ATAs were a substantial contributor to coexistence. 
+For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
 $\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1/4$, with large portions of
@@
 corresponding to ATA rescue. ATA exclusion did not occur for this system for the 
-parameter ranges we considered.
+parameters we considered.
 
@@
 %Summary-type paragraph of main results
-We produced and applied a new extension of modern coexistence theory (MCT) that 
+We developed and applied a new extension of modern coexistence theory (MCT) that 
 quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
@@
 ATA contributions to storage effects and to species coexistence 
-can often be comparable in magnitude to the contributions of previously recognized mechanisms.
+can be comparable in magnitude to contributions of previously recognized mechanisms.
 Our results add
-coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
+coexistence to the list of ecological phenomena on which 
 ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
@@
 of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
-our results suggest that storage effects should be redefined to constitute the 
-contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
+our results suggest that storage effects should be redefined as the 
+contribution of \emph{associations} between $E$ and $C$, generally, including ATA effects and
 correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
-as reflecting the linear relationship between variables, ignoring complexities of the association,
+as reflecting the linear relationship between variables, ignoring complexities of the association
 such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
-in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
-(the beta fecundities lottery model), including our empirically
+in some contexts, but were very important in other contexts, including our empirically
 based example (the diatom system).
+Our work therefore demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
 Future work should seek to understand precisely when ATAs are or are not important
-for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
+for coexistence.
 
@@
 so that $C_{i \bs i}$ is bounded away from $0$, there may be much sharper limits on the 
-ratio $E_i/C_{i \bs i}$ and the corresponding growth rate.
+ratio $E_i/C_{i \bs i}$ and the growth rate.
 ATA effects quantify the extent to which asymmetries of association between $E$ and 
@@
 %Climate change may alter ATAs of coexistence-relevant ATAs
-An additional motivation for quantifying the importance of ATAs for species coexistence 
+An additional motivation for quantifying the importance of ATAs for coexistence 
 is that there is reason to believe climate change may
@@
 For instance, imagine a case in which early spring temperature influences the growth of a plant species, A,
-whereas late spring temperature influences plant species B. If these species are competitors, then EC covariation
-contributing to their coexistence, or failure to coexist, could boil down to covariation
+whereas late spring temperature influences species B. If these species are competitors, then $EC$ covariation
+influencing their potential coexistence could boil down to covariation
 between early and late spring temperatures (e.g., if competitive dynamics are governed by the lottery model with $E_i$ representing
-early spring temperature and $E_j$ representing late spring temperature, see previous paragraph). 
+early spring temperature and $E_j$ representing late spring temperature, see SI section \ref{SI-SIsect:ATA_specialists}). 
 Because longer lasting heat waves are more likely to contribute
 to both early and late spring temperature, a tendency for heat waves to become both more extreme and
-longer lasting increases the degree of right-tail association between early and late spring temperature.
-Likewise, a tendency for cold snaps to become both more intense (i.e., colder temperatures) and longer lasting can increase
-left-tail association between early- and late-spring temperatures. As we showed in this study, changes in ATAs can then
-result in different competitive outcomes between species.
+longer lasting increases the right-tail association between early and late spring temperature.
+As we showed in this study, changes in ATAs can then
+result in different competitive outcomes.
 To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has
@@
 %close
-It seems reasonable to suppose that ATAs will play an increasingly important role in systems of 
+It seems reasonable to suppose that ATAs may play an increasingly important role in systems of 
 greater complexity because such systems have more pairs of temporally variable quantities that
-may exhibit asymmetries of association. Our prior work shows that ATAs are common in 
-ecological and environmental variables \citep{Ghosh_2020_AER}. We have here used simple mathematical
+may exhibit ATAs. Our prior work shows that ATAs are common in 
+ecological and environmental variables \citep{Ghosh_2020_AER}. We have here used mathematical
 and laboratory models to provide an initial exploration of the potential importance of ATAs
@@
 field systems with multiple species, stage structure, spatial extent, non-competitive interactions 
-such as predation and facilitation that can also mediate coexistence, and other complexities.
+such as predation and facilitation, and other complexities.
 

### 2023-08-01 paper/SupMat.Rnw
mostly trimming to get under the 5000 word length for the main text

@@
 for the lottery model. For the lottery model, $B_i(t)=E_i(t)$ was assumed, and $C_{i \bs i}$ simplified to $E_j(t)/\delta$ (see SI section 
-\ref{SIsect:TheoryForLotteryBoth}). Thus, EC-associations parallel associations between the two
+\ref{SIsect:TheoryForLotteryBoth}). Thus, $EC$-associations parallel associations between the two
 environmental variables $E_i(t)$ and $E_j(t)$ for the lottery model, and ATA specialists may exist. 

### 2023-08-01 paper/RefereeResponses1.tex
mostly trimming to get under the 5000 word length for the main text
polishing referee responses a bit and making sure everything is addressed

@@
 \noindent \emph{We have replied to referee comments below in italic, with each response preceded
-by ``***'', to make our responses easier to find.} \\
-
-\noindent \emph{We submit the following pdf files: ...} \\
-%DAN: Fill in ...
-
-\noindent \emph{We additionally submit the following latex files, necessary for compiling the
-main text latex: ...} \\
-%DAN: Fill in ...
-
-\noindent \emph{The document showing the changes we made does not include links such as citations and
-references to figures, but it does accurately represent the changes made to the text. 
-If all latex files are deposited into the same folder, it should then be possible to 
-compile the main text latex from that folder.} \\
+by ``***'', to make our responses easier to find. We submit the following pdf files: The main text pdf, including all figures; 
+the Supporting Information; and a version of the main text with changes tracked. The document showing 
+the changes we made does not include links such as citations and
+references to figures, but it does accurately represent the changes made to the text.} \\
 
@@
 \noindent \emph{Dan Reuman and Jasmin Albert} \\
+%DONE with cover letter
 
@@
 unique is now covered in a new SI section, refered to from the main text at the end of section 2.3, 
-right at the end of the text the referee asked us to move to the main text. There is a choice involved,
-but it is pretty canonical -- certainly more canonical than the kinds of choices which are often made
-in MCT, e.g., $E$ and $C$ can sometimes be defined in more than one way. Basically the choice involves 
+near the end of the text the referee asked us to move to the main text. There is a choice involved,
+but it is a pretty canonical choice -- certainly more canonical than the kinds of choices which are often made
+in MCT, e.g., $E$ and $C$ can often be defined in more than one way. Basically the choice involves 
 using a bivariate normal distribution instead of another, more esoteric 
@@
 as the referee may be suggesting, for several reasons. First, other referees and the editor have urged us to
-simplify, and such explorations would complexify instead. Second, as mentioned above, the choice we made is very
-natural and quite standard, so exploring other possible choices seems therefore of 
+simplify, and such explorations would complexify. Second, as mentioned above, the choice we made is very
+natural and standard, so exploring other possible choices seems of 
 secondary importance. Finally, our overall goal in writing this paper
 was to show that ATAs may be important for coexistence and, as such, deserve additional exploration in this
-context by the wider research community. As long as any choices we made are reasonable (they are), our point is
-made, regardless of whether making some more esoteric choices might have altered outcomes. As we already argue in the
+context by the wider research community. As long as any choices we made are reasonable (they are, very much so), our point is
+made, regardless of whether making some esoteric choice might have altered outcomes. As we already argue in the
 Discussion, since we have now showed that ATAs \textbf{can} be important, future work should explore \textbf{when} 
@@
 \emph{We use notational conventions of random variables which are formally correct and standard, 
-as we now explain in more detail in the SI. 
+as we now explain in more detail in the SI. The alternatives the referee seems to suggest would be misleading to certain audiences 
+because they would be somewhat in conflict with existing conventions of probabilistic notation.
 $E_i$, $C_{i \bs i}$, $E_i^{\nat}$, $C_{i \bs i}^\nat$ and the earlier 
@@
 to the notation. Thus it's not that somehow the expected value in $\E[r(E,C)]$ is a different definition of expected
-value than the one in $\E[r(E^{||},C^{||})]$ -- there is only one expected value operator. 
+value than the one in $\E[r(E^{||},C^{||})]$ -- there is only one expected value operator, and we dare not 
+introduce different versions because we do not want to confuse all those who already know the commonly
+accepted definition of expected value. 
 Instead, $r(E,C)$, $r(E^{||},C^{||})$ and $r(E^\shp,C^\shp)$ differ from each other 
@@
 incorrect, and confusing to readers familiar with probability, to try to define multiple expected value 
-operators, as the referee seems to suggest, when what fundamentally differs, here, is the random variables 
+operators, when what fundamentally differs, here, is the bivariate random variables 
 themselves. Readers familiar with
-probability already know what the one accepted definition of expected value is, would typically
-not be comfortable with providing additional definitions, and are comfortable handling these concepts
-with the notational conventions we have used. To remind the referee of some concepts from basic probability
-which illustrate these conventions, recall that if $X$ and $Y$ are random variables, the expectation of the 
+probability already know what the one accepted definition of expected value is, and would typically
+not be comfortable with providing additional definitions. Some concepts from probability
+which illustrate these conventions are as follows. Recall that if $X$ and $Y$ are random variables, the expectation of the 
 product, $\E(XY)$, depends not only on the univariate distributions of $X$ and $Y$, but also on the relationship
 between $X$ and $Y$, namely their joint distribution. In particular, $\E(XY)=\E(X)\E(Y)$ if $X$ and 
-$Y$ are independent, but not necessarily otherwise. These definitons and notation are completely 
+$Y$ are independent, but not necessarily otherwise. These definitons and notation are 
 standard in probability theory; notice that we use the same definition of $\E$ regardless of 
 whether $X$ and $Y$ are independent or nonindependent, it's just that the product random variable
-$XY$ differs accordingly so you get different results of $\E(XY)$ in those distinct cases.}
+$XY$ differs accordingly so you get different results of $\E(XY)$ in those distinct cases.
+One should not attempt to instead view the difference between $\E(XY)$ for $X$ and $Y$ independent
+as distinct from $\E(XY)$ for $X$ and $Y$ non-independent by looking at two different versions
+of expected value and considering $X$ and $Y$ to be ``dummy variables'', somehow.}
 
@@
 easier for readers to understand and to reduce the potential for future confusion, it makes sense to continue
-using existing notation whenever that notation is formally correct and otherwise effective (it is, as explained above).}
+using existing notation whenever that notation is formally correct and otherwise effective (Ellner's notation is 
+both of these, as explained above).}
 
 \emph{Nevertheless, the referee's comment has made us understand that some readers unfamiliar with random variable 
-notation may benefit from putting some
-of these explanations in the SI, so we have done so.} \\
+notation may benefit from putting these explanations in the SI, so we have done so.} \\
 %DONE
@@
 by poeple who read the manuscript before we submitted it, so it seems important to some readers to have this. 
-The value is not only in showing that ATAs occur in nature, but is priarily  
-in helping people understand more fully what they look like when they occur by giving real examples. 
+The value is not only in showing that ATAs occur in nature, but is primarily  
+in helping people understand more fully what ATAs look like when they occur, by giving a real example. 
 We appreciate what the referee is saying,
@@
 using plankton population density time series'' to ``Fig. 1d,e show contrasting examples of ATAs in nature
-using plankton population density time series; and many other examples have also been previously explored (Ghosh et al 2020)''). 
-And anyway we cannot produce a prior example of ATAs between $E$ and $C$ because there are no prior examples - this paper 
-is, to our
+using plankton population density time series; many other examples of ATAs in nature have also been explored (Ghosh et al 2020)''). 
+And anyway we cannot produce a prior example of ATAs between $E$ and $C$ at this early point in the manuscript, 
+as the referee may be suggesting, because there are no prior examples - this paper is, to our
 knowledge, the first time anyone has looked at ATAs between $E$ and $C$. Fortunately, we think the other motivational
 content of the Introduction is strong, and should stand on its own to motivate readers to spend their mental 
-energy reading this paper. In partcular, we show in the Introduction, for one simple simulation, that ATAs between E and
-C make a visually striking difference between simulation outcomes that appears clearly to influence coexistence.} \\
+energy reading this paper. This motivational content has also been strengthened since the prior version - see other replies. 
+In partcular, we show in the Introduction, for one simple simulation, that ATAs between E and
+C make a visually striking difference between simulation outcomes that appears clearly to influence coexistence.
+We also provide an intuitive example with water limitation of why one might expect ATAs between $E$ and $C$, 
+a bit later in the Introduction from the point to which the referee's comment pertains.
+And we also provide a new paragraph explaining, intuitively, how ATAs work. } \\
 %DONE
@@
 \noindent ***\emph{This was done already, the referee appears to have missed it. It was line 136 in the original submission,
-now line XXX in the resubmitted version.} \\
-%Don't forget to add the new line numbers at the end.
+now line 147 in the resubmitted version.} \\
+%Don't forget to check line numbers at the end.
 
@@
 and grammatical errors.} \\
-%Jasmin to look at line 100 and revise as needed.
+%DONE
 
@@
 paper. We have added new discussion material covering these topics and find that the new material improves the paper.
-We tried to represent both viewpoints in the new material, and we cited the papers the referee mentions.}
+We tried to robustly represent the referee's viewpoints in the new material, and we cited the papers the referee mentions.}
 
@@
 by resources and predators can more readily occur. But once our paper is published we hope and expect that we and
-others will more broadly assess the potential importance of ATAs.}
+others will more broadly assess the potential importance of ATAs in such ``in the wild'' systems.}
 
@@
 opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
-goals, so we thank the referee for not doing that. We modified a few key points in the text to clarify 
+goals, since doing so would keep a lid on the basic ideas and precent their later broad testing.
+So we thank the referee for not doing that. We modified a few key points in the text to clarify 
 our viewpoint that we do not think our work definitively indicates that ATAs will certainly be a crucial
@@
 We would have liked to put some of the new material into the main text. But length limits, together with
-clear requests from other referees to add a variety of content to the main text, made doing so impossible.} \\
+clear requests from other referees to add a variety of explanatory content to the main text, made doing so impossible.} \\
 %DONE
@@
 Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
-a bell for most ecologists. It was not actually necessary to link to Taylor's law, we just though it 
+a bell for most ecologists. It was not actually necessary to link to Taylor's law, we just thought it 
 would provide an additional tie-in to another part of the literature. But that won't work if the referee is right 
-that most ecologists will become confused, so we were happy to drop that connection here.} \\
+that most ecologists will become confused, so we were happy to drop that connection here, leaving this
+section of the text to stand on the connections to extinction risk and ecosystem stability.} \\
 %DONE
@@
 So an explanation, here, is not 
-strictly necessary to the current paper. Nevertheless, we tried to satisfy the referee's request of adding a bit of explanation
+strictly necessary to the current paper. Nevertheless, we tried to satisfy the referee's request of adding a short bit of explanation
 without disrupting the flow too much. Thanks for your interest, by the way, we encourage you to have a look at the
@@
 than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
-reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
+reader that storage effects would be conceptually reviewed below (in the very next paragraph) and then 
 defined in Theory, which they are.} \\
@@
 \noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
-paper possible, and we have used the comments to clarify. But, of the comments 
+paper possible, and we have used their comments to clarify the paper. But, of the comments 
 C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
@@
 specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
-central in ecology these days, and increasingly so. So comments which would tend to sideline such paper
+central in ecology these days, and increasingly so. So comments which would tend to sideline such papers to
 "specialist journals", which fewer people read, seem outdated to us, especially when the papers in question 
 are about mainstream topics such
-as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
+as coexistence. Nevertheless, we do robustly agree that every individual paper should be made as accessible
+as possible, given the subject matter, and we have gone through the paper and redoubled our efforts to make the 
 paper broadly accessible.} \\
-%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.
+%DONE
+
+%DAN: made it to here 2023 08 01, 609am
 
@@
 \emph{Also, see our SI section ``Mechanisms other than storage effects". 
-%can delete if not needed
 Since,

### 2023-08-01 paper/RefereeResponses1.tex
pretty much done

@@
 
-%DAN: made it to here 2023 08 01, 609am
-
 \noindent C21) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
@@
 \noindent ***\emph{We replaced the reference.} \\
-%Jasmin can you please do this?   %%REPLACED
+%DONE
 
@@
 in the Introduction.} \\
-%Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.
-%It's done except for any improvements or revisions that Jasmin may suggest.
-%Jasmin: I really like this response and don't think its too harsh or needs to be toned down.
+%DONE
 
@@
 
-\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
-To avoid making claims that we have not fully supported
+\noindent ***\emph{No. We know how it would be done, but we have not formalized the mathematics
+in the write-up anywhere. And given length limits and a desire to retain as much simplicity as possible,
+we think it's best not to add such mathematics, even to the SI. 
+Therefore, to avoid making claims that we have not fully supported
 with rigorous mathematics, we rephrased to say the iid assumption is 
@@
 
-\emph{Also, see our SI section ``Mechanisms other than storage effects". 
+\emph{Also, see our SI section ``Mechanisms other than storage effects" section in SI,
+where the following mathematics are presented. 
 Since,
@@
 &= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i} \\
-&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}),.
+&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}),
 \end{align} 
@@
 negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
-work, and potential for error, 
-of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
+work, and, more importantly, potential for error, 
+of re-doing several figures, and revising many formulas (in the main text and SI).
 After all, negative numbers are basic, and represent no impediment to the average ecologist understanding the 
-paper.} \\
+paper. } \\
 %DONE
@@
 when ATAs were important and when they were not. To us, the most forthright way of doing that is to
-show \textbf{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
+show \textbf{all} the results we generated, including the values of $\sigma$ which are probably unrealistic, and mention
 (as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
@@
 matters for continuing to explore when ATAs are important and when not is probably the strength
-of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
-a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
+of the noise impacting the system. Which seems accurate since Chesson's weak-noise approximations are more likely
+to work when noise is weaker. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
+a bit of judgement call rather than a formal cutoff, so we hesitate to censor results 
 based on such a thing.}
@@
 figure caption. We also added some text to the figure caption which we hope will focus the 
-readers attention on the aspects of these panels which clearly show asymmetries of tail association. } \\
+readers attention on the visual aspects of these panels which reveal asymmetries of tail association. } \\
 %DONE
@@
 \noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
-(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
+(see first and last paragraphs of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
 is usually one of the largest-magnitude contributions to coexistence. Our results as to the importance of ATAs for coexistence 
 contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
-of section 4.1. That's one of the points we are making.} \\
+of section 4.1. The contrast is one of the points we are making -- that ATAs are sometimes, but not always, important,
+and we need additional work to determine when they are important across a larger collection of systems/models once
+those are available.} \\
 %DONE
@@
 \noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
-negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
+negative for this figure, when it is appreciably different from $0$ (the figure used to be Figure 1 in the 
+prior submission, but is now figure 3). 
+The colors of lines corresponding to different terms are in the legend. And, in the caption,
 it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
@@
 \noindent ***\emph{We have gone to great lengths to address the referee's specific suggestions below, and we have also redoubled our
-efforts to make the manuscript as accessible as possible, generally. We appreciate being spurred to do so, thanks. } \\
-%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
-%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
-%area in ecology? %food webs? networks
-
-%I wrote the following, but decided not to include it here. Might include it elsewhere:
-%The research field
-%surrounding the MCT and mechanisms of coexistence is very important within ecology, but is nevertheless technical - see, for instance, the
-%papers of Chesson, which have had enormous positive impact on the field, as well as the recent papers of Ellner which helped inspire
-%our paper. All those papers, and many other papers in the field, are unavoidably technical, but have frequently been published
-%in mainstream journals (the Ellner papers are in Ecology Letters) because of their importance and the importance of the subject
-%matter they investigate. Though we agree with all efforts to make research results as clear as possible to non-technical audiences,
-%we beg the referee
+efforts to make the manuscript as accessible as possible, throughout. We appreciate being spurred to do so, thanks. } \\
+%DONE
 
@@
 second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added some text to the end of
-the second paragraph that adds to the intuitive understanding of how ATAs can influence coexistence through a concrete 
-example. We also added a new paragraph to the 
-Discussion (the second paragraph of the Discussion). These new Discussion 
-paragraphs are also supported by two new SI sections and two new SI figures which 
-explain, in detail, how the intuition established in the two new main text paragraphs emerges from the modelling (SI sections S11 
+the third paragraph of the Introduction that adds to the intuitive understanding of how ATAs can influence coexistence through a concrete 
+example (water limitation). We also added a new paragraph to the 
+Discussion (the second paragraph of the Discussion). The new Discussion 
+paragraph is also supported by two new SI sections and two new SI figures which 
+explain how the intuition established in the new main text paragraphs emerges from the modelling (SI sections S11 
 and S12, SI Figs S4 and S5).
@@
 
-\noindent ***\emph{Thank you for the positive feedback! We have also added some text to the end of that paragraph giving a 
+\noindent ***\emph{Thank you for the positive feedback! We have also added some text to the end of that paragraph 
+(now the third paragraph of the Introduction) giving a 
 concrete case (water limitation) in which one may expect an influence of ATAs. We think the new text definitely
-makes it more intuitive reasonable to explore ATA contributions to conexistence.} \\
+makes it more intuitively reasonable to explore ATA contributions to conexistence.} \\
 %DONE
@@
 \noindent ***\emph{This cannot occur. ATAs can only influence storage effects, not the other mechanisms, as stated at
-the top of section 2.4 and explained mathematically in the SI section cited there. Basically, as explained in 2.4, 
+the top of section 2.4 and explained mathematically in the SI section cited at the end of section 2.4. Basically, as explained in 2.4, 
 this is because ATAs are about covariation between $E$ and $C$, whereas the mechanisms other than storage effects
@@
 
-\noindent ***\emph{This is certainly a valid question, and we have gone to substantial lengths to address it. 
+\noindent ***\emph{This is certainly a good question, and we have gone to substantial lengths to address it. 
 Partly in response to this comment and partly in response to earlier comments of referee 3, 
@@
 for the log-normal fecundities lottery model. These are very positive additions to the manuscript, so we 
-thank the referee for spurring us to add them. We also believe this is only a start -- additional explanations and intuitive 
-explorations of ATA effects could and should be the subject of future papers. We've added as much material 
-as possible on these topics to the current manuscript, given length limits and
-requests from the referees to both simplify and add a variety of other content.} \\
+thank the referee for spurring us to add them. } \\
 %DONE
@@
 
-\noindent ***\emph{response} \\
-%I think the earlier papers already showed that, at least for the parameters used there? Check. And the other parameters
-%are experimentally manipulable, so this should be OK, but is worth discussing the extent to which ATAs were important
-%for the original parameters. The it would be worth noting that the strongest ATA contributions come for parameters
-%which were not imposed in the original experiment, so it may be worth repeating the experiment but with these other
-%parameters.
+\noindent ***\emph{This was addressed already by the earlier study (Descamps-Julien and Gonzalez, 2005). 
+The only parameters we manipulated were the ones pertaining
+to the experimentally imposed temperature fluctuations; biological parameters determined, validated and used by Descamps-Julien and Gonzalez
+were the same as the biological parameters we used. Descamps-Julien and Gonzalez include results on the extent to
+which dynamics predicted by the model were validated by experiment; the validation was not perfect (it could
+never be), but was judged to be sufficient, and we agree - see the earlier study for details. 
+Outcomes of coexistence/exclusion were 
+correctly predicted by the model, among other validations, as described in the earlier paper. 
+All we've done is consider, in silico, fluctuations of 
+experimentally controlled temperature that sometimes
+differ from those actually imposed, in vitro, by the earlier authors. However, the temperature sensitivities of 
+biological rates determined and used by the earlier authors (reproduced in SI Fig. S3) span temperatures from 
+6-24 degrees C, and all the temperature fluctuations we considered remain within that range. So we
+have also not extrapolated beyond the temperature ranges already considered by the earlier paper.} \\
+%DONE
 
@@
 
-%Need to address. However, this papers seems no more technical than the earlier Ellner paper. And, furthermore, 
-%one of the referees (ref 1) actually asked us to put BACK technical content in the main text which we had 
-%moved to the SI before submission to keep it less technical. Also, this research area is technical, though it
-%is still central to ecology as a discipline. Recycle some of your arguments above about technical aspects. 
-%That said, we do need to do a pass thu and figure out what technical content we can reduce, so we can make it
-%as clear as it can be. Note also they say the math *could* be moved, suggesting less importance than 
-%using ``a more biological/ecological language...''. So maybe let's try to do more of that, some of which
-%I've alreday done thru answering the referee comments.
-\noindent ***\emph{response} \\
+\noindent ***\emph{We have now undertaken major additional efforts to make the paper as accessible as possible.
+Some of these efforts are summarized in responses to individual referees, above. For instance, we
+heavily revised the Introduction to include: a substantial new paragraph that introduces the intuition behind
+ATA effects (the second-to-last paragraph of the Intro); and some new material at the end of the third paragraph
+of the Intro making it clearer where ATAs can come from using a concrete example based on water limitation.
+Efforts to better present the ecological intuition behind ATA effects now also continue in the Discussion, with a new 
+paragraph there (now the second paragraph in the Discussion), which is also supported by new some SI sections. }
+
+\emph{We've also made many smaller, but important, changes to the Intro to avoid making readers think they needed
+more background knowledge to understand the paper than they actually do. In fact, not much background is
+needed. The referees made us aware of, and we have now corrected, a few places early in the Introduction that
+made it seem as though readers were expected to know all about storage effects and relative nonlinearity in
+competition in order to proceed with the paper. In fact, the readers never needed to know anything about relative 
+nonlinearity in competition (that's not needed for the paper at all), and storage effects are explained conceptually
+in the Introduction and defined perecisely in Theory, so readers also do not have to have prior knowledge about
+storage effects. We apologize for the sentences in the earlier version of the paper which made it seem as though
+this was required - we have eliminated those sentences. We have also made explanations of ATAs more accessible,
+and eliminated any passages that could be construed as an indication that prior experience with ATAs is 
+necessary to understand the paper - it is not.}
+
+\emph{We have also moved to the SI a good deal of content from the Discussion that may have been of interest
+more to specialists than generalists,
+replacing it with the new paterial helping establish an intuitive sense for how ATAs work (see above).}
+
+\emph{We thank the referees and editor for spurring us to make these changes, as the paper is improved in
+its accessibility. The paper is certainly now no less accessible than many other papers I've read recently in 
+Ecology Letters, including the two Ellner et al. papers (cited in our study) which helped inspire this 
+work and which were published in Ecology Letters.} 
+%DONE
 

### 2023-08-01 paper/SupMat.Rnw
pretty much done

@@
 in the above formula are computed over all the data). The indices satisfying $u_i+v_i<1$
-are those corresponding to points falling under the dashed diagonal lines in Fig. \ref{MT-fig:pedagog1}d-e.
-%***DAN: Note I mentioned a feature which Jasmin has not yet added, so make sure she 
-%adds it to those panels.
+are those plotted using open points in Fig. \ref{MT-fig:pedagog1}d-e.
 The partial Spearman correlation for the right halves of the distributions,
 $\cor_{\text{right}}$, was defined using the same formula but where the sum is 
-now over all indices $i$ for which $u_i+v_i>1$. These are the points falling above the
-dashed diagonal lines in Fig. \ref{MT-fig:pedagog1}d-e. The left and right partial Spearman
+now over all indices $i$ for which $u_i+v_i>1$. These are the filled points 
+in Fig. \ref{MT-fig:pedagog1}d-e. The left and right partial Spearman
 correlations typically add up to the Spearman correlation. The index 
@@
 the fact that, in case (A) above, both $b_1^{(i)}$ and $b_2^{(i)}$ were assigned to the same value. 
-Right-tail associated noise (Fig. \ref{MT-fig:pedagog1}c) was generated in an analogous way.
+Right-tail associated noise, $\left(b_1^{(i)}, b_2^{(i)}\right)$, $i=1,..,M$ (Fig. \ref{MT-fig:pedagog1}c) 
+was generated by taking the negative of left-tail noise.
 
@@
 associated cases, and are negligible for the example parameters considered in 
-Fig. \ref{fig:lognormal_fecundidities_explanatory}.
+Fig. \ref{fig:lognormal_fecundidities_explanatory}. A careful reader may notice the
+upper horizontal-axis histograms on Fig. \ref{fig:lognormal_fecundidities_explanatory}D, F
+are mirror images of each other. This was because the right-tail noise with standard normal
+marginals that we used was generated
+by simply taking the negative of our left-tail noise with standard-normal marginals (SI section 
+\ref{sect:noise}), a choice that could not realistically have influenced results in a meaningful way.
+This can also be seen by comparing main text Fig. \ref{MT-fig:pedagog1}a, c.
 

### 2023-08-02 paper/Paper.Rnw
typos

@@
 of the information content of the relationship \citep{Nelsen2006}.
-ATAs and related ideas had previously been seldom considered in ecology, but recently 
+ATAs and related ideas had previously been seldom considered in ecology, but recent 
 insights have been gained by studying how features of associations beyond standard measures 
@@
 water, when water is scarce, but their niches become more differentiated when water is abundant.
-Then the quality of the environment, $E$, as experienced by whichever plant is currently rare may only correspond closely to the 
+Then the quality of the environment, $E$, as experienced by whichever plant i. s currently rare may only correspond closely to the 
 competition, $C$, imposed on it by the currently abundant plant when water is scarce and $E$ is low.
@@
 
-\subsection{Our decomposition as an extention of modern coexistence theory}
+\subsection{Our decomposition as an extension of modern coexistence theory}
 
@@
 For goal G2 of the Introduction, we applied our theory to the lottery model, 
-comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
+comparing the magnitude of the ATA effect on coexistence, $\Delta_i^{[EC]}$, to the magnitudes 
 of the other terms of 

### 2023-08-02 paper/Paper.Rnw
acknowledgements

@@
 Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
+Madelon Case, Chhaya Werner, Lina Aoyama, Steven Haring,
 Robin Snyder, and three anonymous referees and the editor for helpful feedback. 

### 2023-08-02 paper/Paper.Rnw
acknowledgements

--- 2023-08-02 paper/Paper.Rnw
@@
 Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Jordan Rodriguez, Lauren Hallett, 
+Madelon Case, Chhaya Werner, Lina Aoyama, Steven Haring, Carmen Watkins,
 Robin Snyder, and three anonymous referees and the editor for helpful feedback. 

### 2023-08-02 paper/Paper.Rnw
small typos

@@
 water, when water is scarce, but their niches become more differentiated when water is abundant.
-Then the quality of the environment, $E$, as experienced by whichever plant i. s currently rare may only correspond closely to the 
+Then the quality of the environment, $E$, as experienced by whichever plant is currently rare may only correspond closely to the 
 competition, $C$, imposed on it by the currently abundant plant when water is scarce and $E$ is low.

### 2023-09-11 paper/RefereeResponses2.tex
started doc for second round of responses

--- 2023-07-31 paper/RefereeResponses1.tex
@@
 \title{Responses to referees for ``Asymmetric relationships and their effects on coexistence''}
-\date{July 2023}
+\date{September 2023}
 
@@
 
-\noindent C1) This is a very interesting paper that adds to our understanding of modern coexistence theory through the decomposition of the ways in which competitive pressure and environment can be correlated.\\
+\noindent C1) First, to consider introducing a discussion of the robustness to more esoteric definitions of the partial-sharp distributions, and adding this briefly to the main text.
+Primarily as the authors already suggest in their response letter, as a signpost for future work.\\
 
-\noindent ***\emph{Thank you for the positive feedback!} \\
+\noindent ***\emph{} \\
 %DONE
 
-\noindent C2) Specifically, the focus is on asymmetric tail association (ATA), by which two random variables with given, fixed marginal distributions and correlation can nevertheless differ in the way that their right and left tails are correlated.
-This is one among many ways that one could characterize or decompose the full bivariate distribution between competition and environment, but the authors argue and demonstrate that ATAs of this kind are (a) prevalent in natural systems and (b) can have a meaningful and sometimes qualitative effect on the outcome of coexistence between two species.
-The combination of results for theoretical distirbutions but also for experimental data is very powerful and convincing.
+\noindent C2) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.
+I was not suggesting that the authors include examples pertinent to their question in Figure 1---as they say, readers may not be expecting this.
+I don't think that the connection between the ATAs in Fig 1 and the rest of the project are that clear, as I already expressed.
+But I respect the authors wishes (and their other feedback) that it is worthwhile including these.
+Just a sentence or so at some point explaining that these are not directly related to the ATAs relevant in the remainder of the paper may avoid confusion.
 \\
 
-\noindent ***\emph{Thank you for the positive feedback!} \\
-%DONE
-
-\noindent C3) I have one overarching question, which I think may be important for ensuring that those points (a) and (b) are robust and then some minor comments that may improve the readability of the manuscript for the EL audience.
-\\
-
-\noindent ***\emph{We address your questions and comments below.} \\
-%DONE
-
-\noindent C4) My biggest question is related to the removal of ATAs from a given true bivariate distribution.  This is described in SI Section S2 and shown visually for an example in SI Figure S1.
-
-My first comment related to this is that some form of this text (whcih is relatively short---around half a page in the SI) and, importantly, the Figure should be in the main text.
-I think this is such a key part of the definitions that are being used for the rest of the paper that readers will want to get an intuition for what the "partial sharp" distributions look like and how they are arrived at.
-\\
-
-\noindent ***\emph{We have done what the referee requests, here, and we agree with this request. Thanks for
-making the request. This request is somewhat in tension with other requests, from other referees, 
-to minimize the quantitative 
-content of the main text, and to keep the overall length under the journal limit while adding a bunch of
-verbal explanations requested by the other referees. We point out this tension here, and refer back to it in
-other responses, below. We also point out that we have not only moved the figure back to the main text,
-we have also added a panel to the figure in response to the related comment C13, below.} \\
-%DONE
-%DAN: In other replies, you can refer to this reply to justify when you
-%decide to add explanations to the SI instead of to the main text, and when you have to respond to requests
-%to minimize the quant.
-
-\noindent C5) Perhaps equally important to clarifying this process for the readership is whether this process is unique.
-I could not quite get a sense of whether there could be alternative versions of this removal of ATA from a bivariate distribution, that still satisfy the authors' constraints on marginal distributions and overall correlation between the two variables.
-Are there alternatives, or is the authors' procedure unique in achieving the desired outcome?
-If the former, then I would want to understand if the outcomes downstream tend to be robust to the specific method used.
-If the latter, it would be helpful to have an explanation as to why this is a unique decomposition.
-\\
-
-\noindent ***\emph{This is a good question, thanks for asking it. The extent to which the process is 
-unique is now covered in a new SI section, refered to from the main text at the end of section 2.3, 
-right at the end of the text the referee asked us to move to the main text. There is a choice involved,
-but it is pretty canonical -- certainly more canonical than the kinds of choices which are often made
-in MCT, e.g., $E$ and $C$ can sometimes be defined in more than one way. Basically the choice involves 
-using a bivariate normal distribution instead of another, more esoteric 
-distribution with standard normal marginals and symmetric tail association. The bivariate normal is by far the 
-most natural example of such a distribution. See the new SI section for further explanations. }
-
-\emph{We are reticent, at this time, to launch into a lot of robustness explorations with regard to this choice, 
-as the referee may be suggesting, for several reasons. First, other referees and the editor have urged us to
-simplify, and such explorations would complexify instead. Second, as mentioned above, the choice we made is very
-natural and quite standard, so exploring other possible choices seems therefore of 
-secondary importance. Finally, our overall goal in writing this paper
-was to show that ATAs may be important for coexistence and, as such, deserve additional exploration in this
-context by the wider research community. As long as any choices we made are reasonable (they are), our point is
-made, regardless of whether making some more esoteric choices might have altered outcomes. As we already argue in the
-Discussion, since we have now showed that ATAs \textbf{can} be important, future work should explore \textbf{when} 
-they are important, and part of 
-that exploration could be whether some more esoteric choice for $(e_i c_{i \bs i})$ might alter our results. }\\
-%DONE
-
-\noindent C6) The notation by which the partial sharp variables are introduced was a little unintuitive to me.  For example, when the authors compare the expectation values 
-$E[r(E,C)]$ and $E[r(E^{||},C^{||})]$, what is really changing is the distribution over which the expectation values are evaluated.
-Maybe this is standard notation, but somehow it may be clearer to put the subscript on the E for expectation value, to indicate that the probability distribution used to generate the definition of the averaging is what differs between the two expressions, while the variables E and C themselves are in the end just dummy variables to be integrated over.
-\\
-
-\noindent ***\emph{This comment has led us to put some clarifying technical comments in a new SI section 2,
-so thanks for making the comment.}
-
-\emph{We use notational conventions of random variables which are formally correct and standard, 
-as we now explain in more detail in the SI. 
-$E_i$, $C_{i \bs i}$, $E_i^{\nat}$, $C_{i \bs i}^\nat$ and the earlier 
-notation $E_i^\shp$ and $C_{i \bs i}^\shp$ all denote \textbf{random variables}, using a 
-notational approach which is standard in applications of probability. 
-Whereas the \textbf{distributions} of the random variables $E_i$, $E_i^{\nat}$ and $E_i^\shp$ (for instance) are all the same,
-these things are \textbf{not} the same \textbf{as random variables} because their relationships with other random variables 
-differ. The relationships between random variables are considered instrinsic
-to the notation. Thus it's not that somehow the expected value in $\E[r(E,C)]$ is a different definition of expected
-value than the one in $\E[r(E^{||},C^{||})]$ -- there is only one expected value operator. 
-Instead, $r(E,C)$, $r(E^{||},C^{||})$ and $r(E^\shp,C^\shp)$ differ from each other 
-as random variables because of the distinct relationships between their
-different $E$ and $C$ constituents. It would be formally 
-incorrect, and confusing to readers familiar with probability, to try to define multiple expected value 
-operators, as the referee seems to suggest, when what fundamentally differs, here, is the random variables 
-themselves. Readers familiar with
-probability already know what the one accepted definition of expected value is, would typically
-not be comfortable with providing additional definitions, and are comfortable handling these concepts
-with the notational conventions we have used. To remind the referee of some concepts from basic probability
-which illustrate these conventions, recall that if $X$ and $Y$ are random variables, the expectation of the 
-product, $\E(XY)$, depends not only on the univariate distributions of $X$ and $Y$, but also on the relationship
-between $X$ and $Y$, namely their joint distribution. In particular, $\E(XY)=\E(X)\E(Y)$ if $X$ and 
-$Y$ are independent, but not necessarily otherwise. These definitons and notation are completely 
-standard in probability theory; notice that we use the same definition of $\E$ regardless of 
-whether $X$ and $Y$ are independent or nonindependent, it's just that the product random variable
-$XY$ differs accordingly so you get different results of $\E(XY)$ in those distinct cases.}
-
-\emph{Our notation also parallels that of Ellner et al (2016, 2019). To make it
-easier for readers to understand and to reduce the potential for future confusion, it makes sense to continue
-using existing notation whenever that notation is formally correct and otherwise effective (it is, as explained above).}
-
-\emph{Nevertheless, the referee's comment has made us understand that some readers unfamiliar with random variable 
-notation may benefit from putting some
-of these explanations in the SI, so we have done so.} \\
-%DONE
-
-\noindent C7) It was not clear to me what the reasoning was for including the correlation and ATAs in abundance for the two planktonic species in Figure 1(d) and (e).
-These ATAs in abundances don't in an obvious (to me) way translate into ATAs of the underlying environment and competition variables for each species.
-Is this really a demonstration that ATAs occur for C and E?
-If not, then it seems to be like saying ATAs exist somewhere in nature, and therefore might be present in our focal situation, which is not such a very strong reason to show (d) and (e).
-Maybe I am missing something.
-\\
-
-\noindent ***\emph{Yes, the referee is right about this, we are not illustrating ATAs in $E$ and $C$ at this stage of the 
-paper - the reader would anyway not be equiped to handle that at this stage, since we have not even formally defined 
-$E$ and $C$ at the point this figure is first cited (in the very first paragraph). 
-We were asked to include this kind of empirical illustration 
-by poeple who read the manuscript before we submitted it, so it seems important to some readers to have this. 
-The value is not only in showing that ATAs occur in nature, but is priarily  
-in helping people understand more fully what they look like when they occur by giving real examples. 
-We appreciate what the referee is saying,
-here, that perhaps one non-$EC$ example of ATAs from nature is not going
-to give much motivation for looking for ATAs in $E$ and $C$. However, we have published droves of other examples of
-ATAs in nature, and we now cite these (changing what used to be ``Fig. 1d,e show contrasting examples of ATAs in nature
-using plankton population density time series'' to ``Fig. 1d,e show contrasting examples of ATAs in nature
-using plankton population density time series; and many other examples have also been previously explored (Ghosh et al 2020)''). 
-And anyway we cannot produce a prior example of ATAs between $E$ and $C$ because there are no prior examples - this paper 
-is, to our
-knowledge, the first time anyone has looked at ATAs between $E$ and $C$. Fortunately, we think the other motivational
-content of the Introduction is strong, and should stand on its own to motivate readers to spend their mental 
-energy reading this paper. In partcular, we show in the Introduction, for one simple simulation, that ATAs between E and
-C make a visually striking difference between simulation outcomes that appears clearly to influence coexistence.} \\
-%DONE
-
-\noindent C8) The authors mention that $r_{i \bs i}$ indicates growth rate of species when $i$ is rare and $j$ is at steady state.
-I think $r_{j \bs i}$ must also assume that $j$ is at steady state, but this isn't mentioned.
-Perhaps that could be added for completeness.
-\\
-
-\noindent ***\emph{This was done already, the referee appears to have missed it. It was line 136 in the original submission,
-now line XXX in the resubmitted version.} \\
-%Don't forget to add the new line numbers at the end.
-
-\noindent C9) E.g.
-line 100 needs removal of the word "are" or the addition of the word "and" somewhere to make sense.
-I'd consider going through the paper carefully for any typos or small corrections of this kind.
-\\
-
-\noindent ***\emph{We changed the wording as requested, and went through the entire manuscript again looking for typos
-and grammatical errors.} \\
-%Jasmin to look at line 100 and revise as needed.
+\noindent ***\emph{} \\
 
@@
 
-\noindent C10) In their paper the authors decompose the “storage effect” into two separate effects, what they call the “asymmetric tail association” and the covariance term per se.
-The authors explain well why they do this and why they think this may be important.
-Additionally, they show the best possible efforts to convince the readers why such a decomposition may be important and what we may learn from it.
-Overall, the paper is interesting and quite understandable.
-I think the authors have done a very good job.
+\noindent C3) This is a novel contribution highlighting that asymmetric tail associations in correlations between two variables was a very simple aspect not properly considered before but with huge implications for the maintenance of biodiversity.
+I congratulate the authors for their work.
 \\
@@
 
-\noindent C11) Unfortunately, I’m personally not convinced that storage effect (and therefore probably ATA) are relevant for coexistence in nature and therefore I’m sceptical about this new approach.
-I want to be clear that this is, in part, my personal opinion and I know experts in the field which have a very different view.
-What the authors have shown (and it appears to be correct) is that in a model where niche differentiation via different resources and predators is impossible, then ATA can be an important driver of coexistence, and I agree.
-However, in every application where both storage effect, relative non-linearity and niche differentiation have been measured, there we see that storage effect and relative non-linearity where very small compared to simple resource differentiation.
-But, one might argue, that we have barely any such empirical applications, and I would agree, which is why I label this my personal opinion.
-(The few papers that I know of are \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1002%2Fecy.2726&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=HbUTjzCNQrvGSARY5u3Eqme88DbKiNyiAVNOUGzjcbE%3D&reserved=0}{Zepeda et al.
-2019, Ecology}, \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1890%2F14-1741.1&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=KKIBuLjBN2vNGWfEpduAGGrMAt7ov9a4F53CmlZQlO0%3D&reserved=0}{Chu and Adler 2015}, and Ellner et al.
-2019, see also the new paper by \href{https://nam10.safelinks.protection.outlook.com/?url=https%3A%2F%2Fesajournals.onlinelibrary.wiley.com%2Fdoi%2Fepdf%2F10.1002%2Fecm.1585&data=05%7C01%7Cd294r143%40ku.edu%7C13ab816eaf2a49c27ea808db784a67cf%7C3c176536afe643f5b96636feabbe3c1a%7C0%7C0%7C638236033192197676%7CUnknown%7CTWFpbGZsb3d8eyJWIjoiMC4wLjAwMDAiLCJQIjoiV2luMzIiLCJBTiI6Ik1haWwiLCJXVCI6Mn0%3D%7C3000%7C%7C%7C&sdata=YaqsZ6axu6rYV12kh9IiRIbES7ZUMWIN7HXgB08iPzc%3D&reserved=0}{Stump and Vasseur}, which I have not yet read fully) In my understanding, ATA is a decomposition of storage effect into even more components, therefore I would expect, on average the ATA effect to be slightly smaller than storage effect (but, as you have shown can also be larger than storage effect).
-I would therefore assume that ATA is generally weak compared to niche differentiation (but again, admittedly based on few datapoints).
-Obviously, the authors may disagree with my personal view, in this case I’d like to see a discussion of this.
-\\
-
-\noindent ***\emph{We thank the referee for this viewpoint. We also thank the referee for acknowledging that this
-is their opinion, and that data are insufficient for a clear conclusion to be reached at this point. We find that 
-such an acknowledgment is both rare from a referee and very useful for causing us, as authors, to most usefully revise our 
-paper. We have added new discussion material covering these topics and find that the new material improves the paper.
-We tried to represent both viewpoints in the new material, and we cited the papers the referee mentions.}
-
-\emph{We point out that we worked not only
-with the lottery model, but also with an empirical diatom system. That system definitely supported the importance of 
-ATAs for coexistance, which would tend to give at least one data point in favor of the importance of these phenomena, though we 
-acknowledge that our lab systems is also a system ``where niche differentiation via ... predators is impossible.''
-It may also be a system ``where niche differentiation via different resources ... is impossible,'' though resource 
-differentiation can happen when you don't expect it. It seems to us 
-the first step in assessing the importance of a newly recognized phenomenon has to be examining it in simple systems
-such as we have used, prior to examing it in field systems, in all their complexity, where niche differentiation
-by resources and predators can more readily occur. But once our paper is published we hope and expect that we and
-others will more broadly assess the potential importance of ATAs.}
-
-\emph{We also point
-out (and the referee already seems to accept this) that it is possible, \emph{a priori}, for ATAs to be 
-important for coexistence in a system even while storage effects are not.
-Storage effects are ATA effects plus what we called, in the manuscript, the effects of $EC$ correlation \emph{per se} 
-($\Delta_i^{[E \nat C]}$ in our notation). Either of these two summands can be negative or positive.
-So if the two effects are opposite in sign and similar in magnitude, storage effects will be unimportant for 
-coexistence, but ATA effects will be important.}
-
-\emph{So the referee appears to agree with us that storage effects and ATA effects may or may not turn
-out to be broadly important in ecology. Our research goals were to introduce the idea of ATA effects on
-coexistence and to show that ATAs \textbf{may} be broadly important for coexistence, by showing examples where they
-\textbf{are} important. The few examples we showed are all that are readily available at this time.
-We won't know for sure if ATAs (or storage effects) \textbf{are} important until more data can be brought to bear. 
-Future work can assess, broadly, how important ATAs are by looking across a larger number of systems, once
-available. These are two different goals. It seems crucial to us not to allow insufficiently tested 
-opinions about the second set of goals to obstruct the publication of results pertinent to the first set of 
-goals, so we thank the referee for not doing that. We modified a few key points in the text to clarify 
-our viewpoint that we do not think our work definitively indicates that ATAs will certainly be a crucial
-mechanism of coexistence; instead we just think our work opens that possibility, for future exploration.
-For instance, we replaced the last sentence of the first paragraph of the Discussion, which used to be
-``Though future work should seek to understand precisely when ATAs are or are not important
-for coexistence, our work demonstrates the overall importance of this new mechanism'' with the revision
-``Future work should seek to understand precisely when ATAs are or are not important
-for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.''
-And we changed the last sentence of the abstract, which used to read ``ATA influences are an important new 
-mechanism of coexistence'' so that it now reads ``ATA influences may be an important new mechanism of coexistence.''
-We think these changes make the paper stronger, so thanks for spurring us to make them. The last
-sentence of the Introduction already read ``Overall, our study presents a new mechanism of species coexistence
-and a means of understanding its theoretical and empirical importance,'' which accurately reflects our belief
-that what we've accomplished is an indication that ATAs \textbf{may} be important for coexistence, 
-and a means of assessing
-that in future work, rather than a definitive conclusion that they are important for coexistence.}
-
-\emph{The new Discussion material is in an SI section which is cited from the main text Discussion section. 
-We would have liked to put some of the new material into the main text. But length limits, together with
-clear requests from other referees to add a variety of content to the main text, made doing so impossible.} \\
-%DONE
-
-\noindent C12) I like the conceptual idea of removing the long tail. \\
-
-\noindent ***\emph{Thank you for the positive feedback.} \\
-%DONE
-
-\noindent C13) However, I have somewhat difficulties to precisely understand/be sure I correctly understood what this does.
-I suggest you move a simpler version of Fig S1 into the main text and show the actual distribution, the uncorrelated distribution of $E$ and $C$, but same marginal distribution, and the distribution with the same covariance, but the ATA removed.
-Additionally, it might be beneficial to indicate which distribution is then used to compute which part of the decomposition.
-For example, in Figure S1 d you show $E_i^{||}$, but from my understanding this distribution would be used to compute $E_i^{(EC)}$, correct?
-\\
-
-\noindent ***\emph{In comment C4 above, another referee asked us also to put S1 back into the main text, so
-we have done as the referee asks here. The new Fig. 2 has the three elements the referee asks for, here, as 
-its panels (a), (e) and (d). We retained the other panels as well, since those were requested by the 
-other referee. Regarding the second part of the referee's comments, 
-there is nothing in our paper for which the notation $E_i^{(EC)}$ is used, but we think what the referee must
-be requesting here is what we remind the reader that $(E_i^\shp,C_{i \bs i}^\shp)$ is used to compute storage effects,
-$\Delta_i^{(EC)}$, and that $(E_i^\nat,C_{i \bs i}^\nat)$ is used to compute ATA effects, $\Delta_i^{[EC]}$. We have 
-now added such a reminder to the figure caption.} \\
-%DONE
-
-\noindent C14) You focus solely on storage effect with the argument that it’s the only where ATA can actually occur.
-But on a broader level you essentially decompose the EC distribution into it’s different momentums.
-One only carrying about the mean (i.e.
-first order correct), one with only carrying about the mean and covariance (i.e.
-getting the first two orders correct), and one with all other orders correct.
-Given this interpretation of your work, I might apply the same logic to the relative-nonlinearity, where I decompose the relative non-linearity into the variance term and the term with long tails.
-(one might also even decompose the actual distribution into more distributions, getting successively more and more momentums of the distribution correct, but I would argue that this is probably less useful).
-\\
-
-\noindent ***\emph{We are not sure we understand precisely what the referee is saying here because their language
-appears more suggestive than formally correct. Perhaps they mean
-``moment'' rather than ``momentum''? Perhaps they mean ``caring'' instead of ``carrying''? I think the phrase 
-``term with long tails'' does not precisely relate to what we're doing since marginal distributions
-of the bivariate distribution $(E,C)$ are unchanged in all our approaches, and hence tails are unchanged. 
-It's just relationships
-between $E$ and $C$ that are modified in the different scenarios we consider. The referee's ideas, here,
-if made precise, may be about potentially using a series of approximations to the bivariate distribution $(E,C)$
-that have the same marginals as $(E,C)$ itself, but that get successively more co-moments right, compared to the
-co-moments of $(E,C)$ itself (and the 
-remaining co-moments would be zero, we presume, if
-such distributions can be defined). Such an approach, applied to relative nonlinearity, would not give
-anything interesting because relative nonlinearity only depends on the marginal distributions, which, 
-under this framework, are always equal to the original marginals of $(E,C)$. On the other hand,
-perhaps the referee means to take a series of approximations of $(E,C)$ so that the $n$th 
-approximate distribution gets both moments and co-moments of order $\leq n$ right, with remaining
-moments and co-moments zero (if such distributions can be defined). That may be interesting, but would
-certainly be a different paper from ours, since our paper is about \emph{relationships} between $E$ and $C$
-(while leaving the univariate distributions of $E$ and $C$ themselves fixed), whereas this idea of the 
-referee would make approximations
-both to the relationship between $E$ and $C$ and to the marginals, simultaneously. Nevertheless, 
-we are pleased the referee 
-is interested in our approach to the extent that they are generating potential ideas for future research.} \\
-%DONE
-
-\noindent C15) L34: I found this sentence confusing due to the two “betweens” and it’s not immediately clear to what they “bound” \\
-
-\noindent ***\emph{We modified the sentence.} \\
-%DONE
-
-\noindent C16) L41: The “we” is slightly confusing because the author teams share only one author \\
-
-\noindent ***\emph{We changed the wording here.} \\
-%DONE
-
-\noindent C17) L48: I do not know these on the top of my head, and likely many other readers will neither, so maybe quickly refresh our memory \\
-
-\noindent ***\emph{There is no space to intro these things, so we limited this text to just mentioning 
-extinction risks and stability of ecosystem functioning. Whereas some ecologists may not have heard of 
-Taylor's law, we feel confident that extinction risks and stability of ecosystem functioning will ring
-a bell for most ecologists. It was not actually necessary to link to Taylor's law, we just though it 
-would provide an additional tie-in to another part of the literature. But that won't work if the referee is right 
-that most ecologists will become confused, so we were happy to drop that connection here.} \\
-%DONE
-
-\noindent C18) L49: This is not immediately clear to me why (obviously, you’ve shown this in the other paper). Is it possible to briefly state why? \\
-
-\noindent ***\emph{It is not necessary for the reader to understand the mechanism we explored in that other
-paper in order to understand this current paper, we were just trying to tie the current work in to 
-another area of research in ecology, for the sake of contextualizing it a bit and motivating studies of ATAs generally. 
-So an explanation, here, is not 
-strictly necessary to the current paper. Nevertheless, we tried to satisfy the referee's request of adding a bit of explanation
-without disrupting the flow too much. Thanks for your interest, by the way, we encourage you to have a look at the
-other paper if you haven't already.} \\
-%DONE
-
-\noindent C19) L61: What if not? Or many readers have only rough understanding of it, maybe quickly refresh or refer to later \\
-
-\noindent ***\emph{It is not necessary for the reader to understand what these are at this point, and,
-in fact, the reader does not need to understand what relative nonlinearity is at all for the main text
-and main ideas of the paper. We apologize for giving the false impression, here, that more background knowledge was needed
-than is actually the case. We eliminated the reference to relative nonlinearity and indicated to the 
-reader that storage effects would be conceptually reviewed below (a couple paragraphs down) and then 
-defined in Theory, which they are.} \\
-%DONE
-
-\noindent C20) Sidenote: After writing these few comments it appears to me that you seem to have a very knowledgeable reader in mind, which may not be the case for a general journal as Ecology letters.
-\\
-
-\noindent ***\emph{Well, we appreciate the referee's effort to spur us to produce the clearest 
-paper possible, and we have used the comments to clarify. But, of the comments 
-C16-C19 above, only C19 relates at all to whether this paper is written for a general ecological audience,
-and we have revised the text C19 mentions so as to robustly address that comment (see response above). 
-The manuscript text mentioned in C17-C18 was just linking this paper to other prior work on ATAs. 
-It's not necessary for the reader to understand that work to understand this paper, the point is just to make 
-the link to that other literature, so that readers can explore it if they so choose, 
-and to show that ATAs have been found to be important in ecology, generally, to motivate the study.
-Furthermore, we don't think our paper is any more specialized or mathematical than the papers Ellner et al (2016, 2019) which
-we cite, and which inspired our paper, and which were published in Ecology Letters. Nor is it more 
-specialized than many other papers we have read in Ecology Letters. Quantitive approaches are absolutely 
-central in ecology these days, and increasingly so. So comments which would tend to sideline such paper
-"specialist journals", which fewer people read, seem outdated to us, especially when the papers in question 
-are about mainstream topics such
-as coexistence. Nevertheless, we have gone through the paper and redoubled our efforts to make the 
-paper broadly accessible.} \\
-%We have to go through the paper and redoubled our efforts to make the paper broadly accessible.
-
-\noindent C21) L64: Narwani 2013 is actually not a good reference because it did not include any temporal variation, a better one would be Letten et al. 2019 in PNAS. \\
-
-\noindent ***\emph{We replaced the reference.} \\
-%Jasmin can you please do this?   %%REPLACED
-
-\noindent C22) L74: The optimum might actually be the same, but they need different responses. \\
-
-\noindent ***\emph{Thanks for pointing this out, we have corrected our language.} \\
-%DONE
-
-\noindent C23) Eq.
-1: I would not use any equations in the introduction, and in general this part of the introduction did not feel like an introduction at all and might require a different sub-section \\
-
-\noindent ***\emph{We disagree, and would prefer not to make this change, for several reasons. 
-First, the lottery model is a very simple model, at the same level of complexity
-as models introduced in essentially all introductory 
-undergraduate ecology courses (e.g., the Lotka-Volterra competition model), so there is no argument that 
-the readership of Ecology Letters cannot easily understand the material presented here.
-Second, this material serves a clear motivating purpose, most appropriately placed in the Introduction. 
-The last author (DCR) has found, from decades of pitching 
-quantitative papers to general ecological audiences, that clear motivation is essential - it is crucial to demonstrate 
-clearly that an effect one plans to elaborate is present, and potentially important, prior to asking the reader to engage 
-the kind of mental energy necessary to read the guts of the paper. So this 
-paragraph is for showing that the effect of ATAs on coexistence can be real, at least for a simple 
-example. That goal is convincingly 
-accomplished by the paragraph, more so than could be accomplished by a purely qualitative, words-only Introduction. 
-Given that the math is simple, and also more precise and definitive than 
-qualitative verbal arguments, we think this material is actually
-more beneficial to the readership and more effective for the intended motivational purpose than would be more words
-(and we see based on C25 below that the referee appears to agree with us here in some respects, at least). 
-Third, quantitative approaches are central in ecology these days, and increasingly so. The automatic tendency
-to want to exclude mathematics of any kind from the Introduction, especially when the math is simple,
-accessible, and serves a clear motivational purpose, seems like a dated bias to us. 
-Finally, we firmly believe that ecology as a discipline gets stronger as it continues the clear
-trend of the last few decades of incorporating more quantitative approaches. We believe we
-benefit the field by challenging assumptions and biases such as categorically not putting equations 
-in the Introduction.} \\
-%Jasmin to read the above and let Dan know if it's convincing or not, and how it can be improved.
-%It's done except for any improvements or revisions that Jasmin may suggest.
-%Jasmin: I really like this response and don't think its too harsh or needs to be toned down.
-
-\noindent C24) L93: I suggest ``Note that $N = N_1(t) + N_2(T)$ is constant through time''. \\
-
-\noindent ***\emph{OK, we made this change, even though the same equation already appears just a few lines
-above this spot. Perhaps the referee is right that the reminder will benefit some readers.} \\
-%DONE
-
-\noindent C25) L95-99: I think this is a great teaser to get the readers to later dive in into your mathematics \\
-
-\noindent ***\emph{Thank you! This is exactly our purpose, here, and we are pleased it has been effective.
-Relating back to our response to C23, we think this purpose cannot really be convincingly accomplished
-without first stating the lottery model, since the result uses that model. Math is more precise and can be
-more convincing than language, the trade off being that some readers won't understand some math. However, the 
-math we have in the Intro is so simple we think all readers of Ecology Letters will understand it, so there are
-benefits and basically no costs of including it there. } \\
-%DONE
-
-\noindent C26) L126: Do you do this in the appendix? \\
-
-\noindent ***\emph{No. We can see how it would be done, but we have not formalized the mathematics. 
-To avoid making claims that we have not fully supported
-with rigorous mathematics, we rephrased to say the iid assumption is 
-``probably not necessary''. Thanks for spurring us to do this.} \\
-%DONE
-
-\noindent C27) Eq 2.: I thought we also have to remove the effect of C varying alone and E varying alone? Is this included in the $\#$ term automatically? \\
-
-\noindent ***\emph{No, the equation is correct as is. Please see equation 10 and related equations of 
-Ellner et al, 2019, Ecology Letters. Yes, the effects of $C$ and $E$ varying are included in 
-the $\#$ term because that terms involves both of them varying (though independently).}
-
-\emph{Also, see our SI section ``Mechanisms other than storage effects". 
-%can delete if not needed
-Since,
-\begin{align}
-\olep_i^{EC} &= \E[r_i(E_i,C_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
-\end{align}
-and 
-\begin{align}
-\olep_i^{(E\shp C)} &= \E[r_i(E_i^\shp,C^\shp_{i \bs i})]-\olep_i^E-\olep_i^C-\epsilon_i^0 \\
-&= \olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0,
-\end{align}
-we have
-\begin{align}
-\olep_i^{(EC)} &= \olep_i^{EC}-\olep_i^{(E\shp C)} \\
-&= (\olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) - (\olrs_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0) \\
-&= \olr_{i \bs i} - \olep_i^E - \olep_i^C - \epsilon_i^0 - \olrs_{i \bs i} + \olep_i^E + \olep_i^C + \epsilon_i^0 \\
-&= \olr_{i \bs i} - \olrs_{i \bs i}. 
-\end{align}
-So,
-\begin{align}
-\Delta_i^{(EC)} &= \olep_i^{(EC)} - q_{ij}\olep_j^{(EC)} \\
-&= (\olr_{i \bs i} - \olrs_{i \bs i}) - q_{ij}(\olr_{j \bs i} - \olrs_{j \bs i}) \\
-&= \olr_{i \bs i} - \olrs_{i \bs i} - q_{ij}\olr_{j \bs i} + q_{ij}\olrs_{j \bs i} \\
-&= \olr_{i \bs i} - q_{ij}\olr_{j \bs i} - \olrs_{i \bs i} + q_{ij}\olrs_{j \bs i} \\
-&= (\olr_{i \bs i} - q_{ij}\olr_{j \bs i}) - (\olrs_{i \bs i} - q_{ij}\olrs_{j \bs i}),.
-\end{align} 
-which is the equation the referee is asking about.} \\
-%DONE
-
-\noindent C28) Eq. 6: Maybe highlight that you use different brackets ([EC] versus (EC)), or probably even better, make 5 and 6 one equation using curly under brackets. \\
-
-\noindent ***\emph{Yes, it is a good idea to make it completely clear to the reader that these are 
-used distinctly. We now do so, though we do it above this point in the text - it is better, in our view,
-to do this the first time we use the hard brackets. So thanks for pointing this out.
-We do not know what the referee means by ``using curly under brackets'' so have not done anything
-in response to that part of this comment.} \\
-%DONE
-
-\noindent C29) L230 $\mu_1-\mu_2$ is negative, why not take $\mu_2-\mu_1$? \\
-
-\noindent ***\emph{Well, we can do this if the editor insists, but to us, avoiding the use of
-negative numbers is insufficient reason to have to undertake the substantial and time-consuming 
-work, and potential for error, 
-of re-doing several figures, and revising many formulas (in the main text and SI) which currently use the other convention.
-After all, negative numbers are basic, and represent no impediment to the average ecologist understanding the 
-paper.} \\
-%DONE
-
-\noindent C30) L316: So why do you show values of $\sigma >6$?
-And also maybe highlight this in the figure Figure 1d and e: I’m semi convinced by these distributions showing ATA.
-In 1d it simply appears that the variance scales with the density.
-1e is quite unconvincing to me.
-That’s generally fine, but figure 1 seems to be a conceptual figure, so I wonder whether you don’t have a better example?
-\\
-
-\noindent ***\emph{Regarding the first referee question listed above, well, what we are 
-trying to do here is accurately portray our results about
-when ATAs were important and when they were not. To us, the most forthright way of doing that is to
-show \textbf{all} the results, including the values of $\sigma$ which are probably unrealistic, and mention
-(as we do) that some values are unrealistic. The point of this paragraph is to outline the circumstances
-we have explored, using our lottery-model approaches, for which ATAs were important or not, and including all
-our results seems important when such is our goal. Furthermore, there's no real cost to showing
-all the $\sigma$'s we explored, it does not take up more manuscript space to do so.
-Some readers may also take the lesson, for future work, from these explorations that one of the things that probably 
-matters for continuing to explore when ATAs are important and when not is probably the strength
-of the noise impacting the system. Finally, whether a value of $\sigma$ is ``unrealistic'' is 
-a bit of judgement call rather than a formal cutoff, to we hesitate to censor results 
-based on such a thing.}
-
-\emph{Regarding the second part of the referee comment, pertaining to Figure 1d and e, it is perplexing to us that 
-this is mentioned here,
-as the paragraph at L316 is not directly related to Fig. 1d-e. That figure
-shows motivational empirical examples, distinct from the lottery model work discussed around L316. Nevertheless, we took
-seriously the referee's comment that they were apparently not convinced there are ATAs in Fig. 1d-e.
-In fact, one can significance test these results and they are solidly significant - we added
-those results, along with the methods used to get them, to a new SI section which is now cited from the
-figure caption. We also added some text to the figure caption which we hope will focus the 
-readers attention on the aspects of these panels which clearly show asymmetries of tail association. } \\
-%DONE
-
-\noindent C31) Figure 2: I recommend to add a line at $y = 0$. Also the different subgraphs have different $y$-scales which I’m not sure is beneficial, but unsure myself. \\
-
-\noindent ***\emph{There is already a line at $y=0$. You can see it, for instance, on panel (e). 
-It gets unavoidably overplotted on panels for which one of the plotted quantities is close to 0. The reason
-for different $y$-axis extents is so that you can see the variation within each panel with maximum 
-clarity. The main point, here, is to compare the relative contributions of different components
-to coexistence, so it's important to show within-plot variation (between lines and across 
-values of the $x$-axis) clearly. We tried plotting with the same scales everywhere and it obscured some of the 
-patterns. We thank the referee for expressing their own uncertainty, which was
-useful to be aware of.} \\
-%DONE
-
-\noindent C32) Figure 2 and 3: It appears that ATA is almost never the relevant factor, however I wonder whether this is because ATA is relatively weak, or whether the same analysis of say $\delta_i$ would lead to a similar pattern, it might be worth creating these graphs in the appendix to strengthen your case.
-\\
-
-\noindent ***\emph{For Fig. 2, the referee is right that ATAs are only rarely important, as explained in the text 
-(see first paragraph of section 4.1). However, for Fig 3, the referee is incorrect. In that figure, the ATA term
-is usually one of the largest-magnitude contributions to coexistence. Our results as to the importance of ATAs for coexistence 
-contrast between our log-normal and beta funcundities lottery models, as explained in the last paragraph
-of section 4.1. That's one of the points we are making.} \\
-%DONE
-
-\noindent C33) Figure 2: ATA is always positive, so how can it lead to ATA exclusion? (or ATA rescue in case I misunderstood). \\
-
-\noindent ***\emph{The referee appears to have misinterpreted the figure. The ATA terms is, in fact, usually
-negative for this figure. The colors of lines corresponding to different terms are in the legend, and, in the caption,
-it is reviewed which of these corresponds to the ATA term. Since these items are clearly labelled, and since 
-the $y=0$ line is also already present, we are unsure how to reduce the chances for other readers to make
-such errors. But we're happy to hear specific suggestions.} \\
-%DONE
-
 \noindent \textbf{Comments of referee 3:} \\
 
-\noindent C34) This is a well performed paper pointing out that asymmetric tail associations (ATA) in a correlation between the environment and competition can be an important mechanisms of coexistence previously overlooked.
-ATAs can give additional mechanisms that either promote or hinder coexistence in a similar fashion and magnitude to other previously recognised mechanisms such as the storage effect.
-To illustrate these ideas the authors combine results from mathematical models, simulations, and empirical examples, and show that under a certain combination of parameters ATAs can be a significant mechanism of species coexistence.
-\\
+\noindent C4) I thank the authors for the change and the nice discussion/replies to my questions. I especially like the new figure 2 which greatly helped my understanding of how to remove ATA. \\
 
 \noindent ***\emph{Many thanks for the positive feedback!} \\
-%DONE
-
-\noindent C35) Overall, I agree with the authors that the role of ATAs in coexistence has not been properly considered and the manuscript give convincing results of the importance of ATAs. \\
-
-\noindent ***\emph{Thank you for the positive feedback!} \\
-%DONE
-
-\noindent C36) However in order to reach a broad audience of ecologists, the manuscript needs to reach a balance between mathematical definitions and simple explanations.
-Otherwise, it can remain to technical and of little importance to ecologists interested in the role of environmental variability of community dynamics.
-I will give some examples where the message can be improved, but I would suggest the authors to do the same effort themselves in other parts of the manuscript.
-\\
-
-\noindent ***\emph{We have gone to great lengths to address the referee's specific suggestions below, and we have also redoubled our
-efforts to make the manuscript as accessible as possible, generally. We appreciate being spurred to do so, thanks. } \\
-%We have to redouble our efforts, as it says. Maybe we also want to point out somewhere that this particular
-%sub-field is already a very technical one, but is nevertheless considered a mainstream and very important
-%area in ecology? %food webs? networks
-
-%I wrote the following, but decided not to include it here. Might include it elsewhere:
-%The research field
-%surrounding the MCT and mechanisms of coexistence is very important within ecology, but is nevertheless technical - see, for instance, the
-%papers of Chesson, which have had enormous positive impact on the field, as well as the recent papers of Ellner which helped inspire
-%our paper. All those papers, and many other papers in the field, are unavoidably technical, but have frequently been published
-%in mainstream journals (the Ellner papers are in Ecology Letters) because of their importance and the importance of the subject
-%matter they investigate. Though we agree with all efforts to make research results as clear as possible to non-technical audiences,
-%we beg the referee
-
-\noindent C37) At the beginning of the introduction the authors cite previous work to show the prevalence and importance of ATAs for population dynamics, and the stability of ecological communities, and Fig.1 illustrate what ATAs mean in a clear example, yet it is not well explained or at least it was not clear to me how ATA can either promote or hinder coexistence in these initial sentences.
-\\
-
-\noindent ***\emph{The referee appears to want us to present an intuitive explanation for how ATA effects work, and that 
-is certainly a good idea. We have invested substantial efforts into doing so, and think we have succeeded. Though the referee's comment above may 
-be interpreted to suggest that they wanted us
-to add such an explanation to first paragraph of the paper, the reader is not equiped to understand the explanation at that 
-point, so we added a new paragraph later in the Introduction (it is now the
-second-to-last paragraph in the Introduction) that begins establishing such an intuition. We also added some text to the end of
-the second paragraph that adds to the intuitive understanding of how ATAs can influence coexistence through a concrete 
-example. We also added a new paragraph to the 
-Discussion (the second paragraph of the Discussion). These new Discussion 
-paragraphs are also supported by two new SI sections and two new SI figures which 
-explain, in detail, how the intuition established in the two new main text paragraphs emerges from the modelling (SI sections S11 
-and S12, SI Figs S4 and S5).
-%DAN: make sure these are the final numbers of these SI sections
-We believe that intuition such as the new material establishes should always be based in formal theoretical and 
-empirical work, so the focus of the paper remains on our expansion of the MCT and our results showing
-the *importance* of ATA effects. Though we definitely think our additions giving an intuition for ATA effects are a large
-improvement in the manuscript, so we thank the referee for spurring us to make these additions.  } \\
-%DONE except the checking as noted above.
-
-\noindent C38) Then the explanation of the importance of ATAs is done in relationship with the explanation of the storage effect. In detail, the explanation of the storage effect is superb, ... \\
-
-\noindent ***\emph{Thank you for the positive feedback! We have also added some text to the end of that paragraph giving a 
-concrete case (water limitation) in which one may expect an influence of ATAs. We think the new text definitely
-makes it more intuitive reasonable to explore ATA contributions to conexistence.} \\
-%DONE
-
-\noindent C39) ...but how ATAs enter in play is not well solved.
-It is just mentioned that ATAs can modify the EC covariation, which I believe is not enough, a reader might not understand later the explanation when the lottery model is presented.
-\\
-
-\noindent ***\emph{See our response to C37. We have added substantial material to the Introduction and Discussion that adds an
-intuitive explanation for ATAs, and we thank the referee for spurring us to do so. 
-We find it unlikely that any focussed reader will have substantial difficulties understanding the lottery model
-material in the Introduction, since the lottery model is certainly no more complicated than models such as the Lotka-Volterra 
-competition and predator-prey models, which are covered in essentially every introductory undergraduate ecology class, and we can assume
-that almost all readers of Ecology Letters have taken such a class. See also comment C25 above, which indicates that another
-referee found this material to be compelling.} \\
-%DONE
-
-\noindent C40) ATAs can not only modify the importance of the storage effect but also of other mechanisms such as relative non-linearity, which is based on how competition fluctuates in ways that favor the competitor.
-Could the authors briefly indicate how this could occur?
-\\
-
-\noindent ***\emph{This cannot occur. ATAs can only influence storage effects, not the other mechanisms, as stated at
-the top of section 2.4 and explained mathematically in the SI section cited there. Basically, as explained in 2.4, 
-this is because ATAs are about covariation between $E$ and $C$, whereas the mechanisms other than storage effects
-do not pertain at all to covariation. For instance, relative nonlinearity is about the effects of variation 
-in $E$ or $C$, not covariation between them.} \\
-%DONE
-
-\noindent C41) The lottery model is used to illustrate the importance of ATAs, and two types of fecundities are taken 
-into account (log-normal and beta), can the author explain why both are important, and later on the results and discussion 
-why ATAs play a major role with beta fecundities and not with log-normal.
-The model used for the chemostat is in its formulation quite different compared to the lottery model. Although the technical
-details are left for the supplementary material, can the authors explain more in the detail in the main text how the 
-importance of ATAs can be used for this particular model.\\
-
-\noindent ***\emph{This is certainly a valid question, and we have gone to substantial lengths to address it. 
-Partly in response to this comment and partly in response to earlier comments of referee 3, 
-we added substantial new material providing intuitive explanations of ATA effects in the Introduction,
-Discussion, and SI (see above responses). That new material also has information comparing the models
-and explaining why ATA effects are substantial for the beta-fecundities lottery model and not as strong
-for the log-normal fecundities lottery model. These are very positive additions to the manuscript, so we 
-thank the referee for spurring us to add them. We also believe this is only a start -- additional explanations and intuitive 
-explorations of ATA effects could and should be the subject of future papers. We've added as much material 
-as possible on these topics to the current manuscript, given length limits and
-requests from the referees to both simplify and add a variety of other content.} \\
-%DONE
-
-\noindent C42) The chemostat model is parameterized with information from a previous study, but does this model reproduce observed dynamics of the diatom system?
-this needs to be shown otherwise the importance of ATAs will be considered for another dynamics that do not correspond to the observed ones.
-\\
-
-\noindent ***\emph{response} \\
-%I think the earlier papers already showed that, at least for the parameters used there? Check. And the other parameters
-%are experimentally manipulable, so this should be OK, but is worth discussing the extent to which ATAs were important
-%for the original parameters. The it would be worth noting that the strongest ATA contributions come for parameters
-%which were not imposed in the original experiment, so it may be worth repeating the experiment but with these other
-%parameters.
 
@@
 
-\noindent C43) This study novelty is the focus on assymetric tail association between competition and environmental conditions as a mechanism of coexistence. \\
-
-\noindent ***\emph{Thank you for pointing out the novelty.} \\
-
-\noindent C44) The manuscript is very theoretical with a highly technical language hard to follow by a broad range of ecologists.
-My main suggestion is that the authors need to rewrite it in several parts to make it accessible and convincing.
-The authors have to make an effort to use a more biological/ecological language with reference to real world potential examples/cases/systems.
-Some mathematical material could be moved to Supplementary material.
+\noindent C5) Reviewers are enthusiastic about this paper demonstrating the importance of how variable relationships depending on when they are weak or strong influence competitive exclusion and coexistence.
 \\
 
-%Need to address. However, this papers seems no more technical than the earlier Ellner paper. And, furthermore, 
-%one of the referees (ref 1) actually asked us to put BACK technical content in the main text which we had 
-%moved to the SI before submission to keep it less technical. Also, this research area is technical, though it
-%is still central to ecology as a discipline. Recycle some of your arguments above about technical aspects. 
-%That said, we do need to do a pass thu and figure out what technical content we can reduce, so we can make it
-%as clear as it can be. Note also they say the math *could* be moved, suggesting less importance than 
-%using ``a more biological/ecological language...''. So maybe let's try to do more of that, some of which
-%I've alreday done thru answering the referee comments.
-\noindent ***\emph{response} \\
+\noindent ***\emph{Thank you for pointing out.} \\
+
+\noindent C6) The legends of figures are rather long and dense.
+Figures should be self-explanatory.
+I suggest adding to all of them the sentence at the end: “See Table 1 for used annotations and abbreviations.” \\
+
+\noindent ***\emph{We }
+
+\noindent C7) The legend in Fig.
+1 is too long and has variables not present in the graphs but mentioned in other sections of the paper.
+For example, in line 44-45: “The variables (B1, B2) described in the Introduction…”.
+I suggest revising if this information is needed at all in this Figure.
+\\
+
+\noindent ***\emph{Thank you for the very positive feedback!} \\
+
+\noindent C8) Main concepts should be described in the introduction in a way that could be understood for a broad audience.
+In this sense, I suggest adding a short non- mathematical definition for storage effects even if you go in more detail in the Theory section in lines 59-60 \\
+
+\noindent ***\emph{Thank you for the very positive feedback!} \\
+
+\noindent C9) Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994).
+Storage effects refer to…..
+Storage effects are reviewed conceptually below and defined formally in Theory.
+\\
+
+\noindent ***\emph{Thank you for the very positive feedback!} \\
+
+\noindent C10)  Page 2, Line 60: Start a new paragraph with the sentence “MCT has been applied to several systems…” \\
+
+\noindent ***\emph{Thank you for the very positive feedback!} \\
+
+\emph{We thank the referees and editor for spurring us to make these changes, as the paper is improved in
+its accessibility. The paper is certainly now no less accessible than many other papers I've read recently in 
+Ecology Letters, including the two Ellner et al. papers (cited in our study) which helped inspire this 
+work and which were published in Ecology Letters.} 
+%DONE
 

### 2023-09-11 paper/Paper.Rnw
responding to editor's comments

@@
 MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
-``storage effects'' \citep{Chesson_1994}; storage effects are reviewed conceptually below and 
-defined formally in Theory. MCT has been
+``storage effects'' \citep{Chesson_1994}. 
+Storage effects refer to ... and are reviewed conceptually below and 
+defined formally in Theory.
+
+MCT has been
 applied to several systems, including annual \citep{Angert_2009, Hallett_2019} and perennial plants \citep{Adler_2006, Adler_2010}, 
@@
 see also (d,e)]. 
-The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
-pictured and then transforming the marginals as specified in Methods.
+%The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
+%pictured and then transforming the marginals as specified in Methods.
 (d,e) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
@@
 which are in the right halves of the distributions; and \emph{vice versa} for panel e.
-(f-h) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
-respectively corresponding variables (a-c; see Introduction).
+(f-h) Lottery model population simulations% using $(B_1,B_2)$ 
+obtained by transforming the respectively corresponding variables (a-c; see Introduction).
+See Table 1 for used annotations and abbreviations.
 \label{fig:pedagog1}}
@@
 $\log_{10}$ of probability density on all panels. 
+See Table 1 for used annotations and abbreviations.
 \label{fig:pedagog2}}
@@
 $q_{ij}=\exp(\mu_i)/[(1-\delta)\exp(\mu_j)+\delta\exp(\mu_i)]$; though results differed
-in detail, main conclusions were the same. \label{fig:fig3}}
+in detail, main conclusions were the same. 
+See Table 1 for used annotations and abbreviations. \label{fig:fig3}}
 \end{figure}
@@
 though results differed in detail, main conclusions were the same.
-}\label{fig:fig4}
+See Table 1 for used annotations and abbreviations. }\label{fig:fig4}
 \end{figure}
@@
 excluding the effects of ATAs, so GWR for \emph{F. crotonensis}, decomposed here, is a 
-sufficient coexistence metric.
+sufficient coexistence metric. See Table 1 for used annotations and abbreviations.
 

### 2023-09-14 paper/RefereeResponses2.tex
responding to editor's comments
responses to reviewer1

@@
 
-\noindent C1) First, to consider introducing a discussion of the robustness to more esoteric definitions of the partial-sharp distributions, and adding this briefly to the main text.
+\noindent C1) Thank-you to the authors for their response to my comments.
+
+\noindent ***\emph{We appreciate your feedback which helped to improve our manuscript; thank you.} \\
+
+\noindent C2) I have two minor suggestions remaining which I think should be straightforward to implement.
+
+First, to consider introducing a discussion of the robustness to more esoteric definitions of the partial-sharp distributions, and adding this briefly to the main text.
 Primarily as the authors already suggest in their response letter, as a signpost for future work.\\
@@
 \noindent ***\emph{} \\
-%DONE
+%%JAS: added a sentence at the end of first paragraph of discussion.
 
-\noindent C2) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.
+\noindent C3) The only other minor suggestion is that the authors clarify that in Fig 1 they are simply showing that ATAs do occur in nature.
 I was not suggesting that the authors include examples pertinent to their question in Figure 1---as they say, readers may not be expecting this.
@@
 \noindent ***\emph{} \\
+%%JAS: is it okay to put this is the caption of fig 1?
+%%or in main, "Though not directly related to our questions at hand, 
+%and for illustrative purposes, Fig1d,e show ...
 
@@
 
-\noindent ***\emph{Thank you for the very positive feedback!} \\
+\noindent ***\emph{Thank you for the very positive feedback and previous comments which helped to improve our manuscript!} \\
 %DONE
@@
 \noindent ***\emph{We }
+%%So replace any variable explanations with this sentence? I dont feel that every caption needs this. Particularly after addressing next comment. 
+%%Sure can add but I don't feel like this replaces explanation of things and just makes caption longer..
 
@@
 
-\noindent ***\emph{Thank you for the very positive feedback!} \\
+\noindent ***\emph{. } \\
+%%Take out parts of caption that mention (B1, B2)?. As long as the main text mentions that we used noise as pictired in fig one that should okay. But double check
 
@@
 
-\noindent ***\emph{Thank you for the very positive feedback!} \\
+\noindent ***\emph{} \\
 
@@
 
-\noindent ***\emph{Thank you for the very positive feedback!} \\
+\noindent ***\emph{} \\
+%%JAS: took a stab at brief explanation but maybe should be shortende
+%"Storage effects refer to effects of [temporal] niche partitioning which allow species to "store" benefits from favorable conditions to persists in unfavorable conditions"
 
@@
 
-\noindent ***\emph{Thank you for the very positive feedback!} \\
+\noindent ***\emph{Started as new paragraph...
+But this sentence or the proceeding sentence are not good topic sentences for the next paragraph about storage effects.
+Unless editor means to make a two sentence paragraph?} \\
+%%JAS: I dont think this is a good paragraph break but I did it..
 

### 2023-09-18 paper/Paper.Rnw
responses to reviewer1
minor

--- 2023-08-02 paper/Paper.Rnw
@@
 \noindent \textbf{Type of article:} Letter\\
-\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 5013; \textbf{Words in text boxes:} 0\\
-\noindent \textbf{Number of references:} 37; \textbf{Number of figures:} 4; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
+\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4975; \textbf{Words in text boxes:} 0\\
+\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
 \noindent \textbf{Corresponding author:}

### 2023-09-18 paper/RefereeResponses2.tex

@@
 
-\noindent C3) This is a novel contribution highlighting that asymmetric tail associations in correlations between two variables was a very simple aspect not properly considered before but with huge implications for the maintenance of biodiversity.
+\noindent C4) This is a novel contribution highlighting that asymmetric tail associations in correlations between two variables was a very simple aspect not properly considered before but with huge implications for the maintenance of biodiversity.
 I congratulate the authors for their work.
@@
 
-\noindent C4) I thank the authors for the change and the nice discussion/replies to my questions. I especially like the new figure 2 which greatly helped my understanding of how to remove ATA. \\
+\noindent C5) I thank the authors for the change and the nice discussion/replies to my questions. I especially like the new figure 2 which greatly helped my understanding of how to remove ATA. \\
 
-\noindent ***\emph{Many thanks for the positive feedback!} \\
+\noindent ***\emph{Many thanks for the positive feedback which improved our manuscript to become more understandle to other readers as well!} \\
 
@@
 
-\noindent C5) Reviewers are enthusiastic about this paper demonstrating the importance of how variable relationships depending on when they are weak or strong influence competitive exclusion and coexistence.
+\noindent C6) Reviewers are enthusiastic about this paper demonstrating the importance of how variable relationships depending on when they are weak or strong influence competitive exclusion and coexistence.
 \\
 
-\noindent ***\emph{Thank you for pointing out.} \\
+\noindent ***\emph{Thank you for pointing out reviewers' positive feedback.} \\
 
-\noindent C6) The legends of figures are rather long and dense.
+\noindent C7) The legends of figures are rather long and dense.
 Figures should be self-explanatory.
@@
 
-\noindent C7) The legend in Fig.
+\noindent C8) The legend in Fig.
 1 is too long and has variables not present in the graphs but mentioned in other sections of the paper.
@@
 
-\noindent C8) Main concepts should be described in the introduction in a way that could be understood for a broad audience.
-In this sense, I suggest adding a short non- mathematical definition for storage effects even if you go in more detail in the Theory section in lines 59-60 \\
+\noindent C9) Main concepts should be described in the introduction in a way that could be understood for a broad audience.
+In this sense, I suggest adding a short non- mathematical definition for storage effects even if you go in more detail in the Theory section in lines 59-60
 
-\noindent ***\emph{} \\
-
-\noindent C9) Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994).
+Page 2, Lines 59-60 “coexistence into the contributions of various mechanisms, with names such as storage effects (Chesson1994).
 Storage effects refer to…..

### 2023-09-18 paper/Paper.Rnw

--- 2023-09-11 paper/Paper.Rnw
@@
 And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
-ecosystem functioning. Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
+ecosystem functioning. 
+%Though not directly related to our questions at hand, and for illustrative purposes,
+Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
 plankton population density time series; many other examples of ATAs in nature have also been explored 
@@
 ``storage effects'' \citep{Chesson_1994}. 
-Storage effects refer to ... and are reviewed conceptually below and 
-defined formally in Theory.
+Storage effects refer to ... 
+%effects of [temporal] niche partitioning which allow species to "store" benefits from favorable conditions to persists in unfavorable conditions
+ and are reviewed conceptually below and defined formally in Theory.
 
@@
 for coexistence. 
+%For example, the robustness of our choice in specifying a bivariate normal partial sharp distribution,
+%though a natural and canonical one, to other more esoteric specifications.
 

### 2023-09-18 paper/RefereeResponses2.tex
Dan thinks the referee responses and corresponding changes to the paper are complete

@@
 
-\noindent \emph{Thank you for the useful and very positive reviews of our manuscript, and for
-the opportunity to resubmit a revised version. We have addressed all referee comments below and in
-the manuscript, and hope the paper can now be rapidly accepted. } \\
+\noindent \emph{Thank you for the decision of ``minor revisions''. We have 
+fully addressed the referee and editor comments below and in the manuscript.
+We hope the paper can now be rapidly formally accepted. } \\
 
 \noindent \emph{We have replied to referee comments below in italic, with each response preceded
-by ``***'', to make our responses easier to find. We submit the following pdf files: The main text pdf, including all figures; 
-the Supporting Information; and a version of the main text with changes tracked. The document showing 
+by ``***'', to make our responses easier to find.} \\
+
+\noindent \emph{We submit the following pdf files: The main text pdf, including all figures for readability; 
+the Supporting Information; a version of the main text with changes tracked; and all figures as 
+separate pdfs, for the production process. The document showing 
 the changes we made does not include links such as citations and
-references to figures, but it does accurately represent the changes made to the text.} \\
+references to figures, but it does accurately represent the changes made to the text.
+We also submit...
+%***DAN: all the latex files and so on.
+%*%***DAN: make sure the stuff about putting a latex version of the MT at the end of the SM is all ready to go 
+%*%in what I submit.
+} \\
 
@@
 \noindent \emph{Dan Reuman and Jasmin Albert} \\
-%DONE with cover letter
 
@@
 
-\noindent C1) Thank-you to the authors for their response to my comments.
+\noindent C1) Thank-you to the authors for their response to my comments.\\
 
@@
 
-\noindent ***\emph{} \\
-%%JAS: added a sentence at the end of first paragraph of discussion.
+\noindent ***\emph{We added the requested material to section 2.3.} \\
 
@@
 
-\noindent ***\emph{} \\
-%%JAS: is it okay to put this is the caption of fig 1?
-%%or in main, "Though not directly related to our questions at hand, 
-%and for illustrative purposes, Fig1d,e show ...
+\noindent ***\emph{We added a sentence, as requested, in the first paragraph of the Introduction.} \\
 
@@
 \noindent ***\emph{Thank you for the very positive feedback and previous comments which helped to improve our manuscript!} \\
-%DONE
 
@@
 
-\noindent ***\emph{We }
-%%So replace any variable explanations with this sentence? I dont feel that every caption needs this. Particularly after addressing next comment. 
-%%Sure can add but I don't feel like this replaces explanation of things and just makes caption longer..
+\noindent ***\emph{We did as requested.} \\
 
@@
 
-\noindent ***\emph{. } \\
-%%Take out parts of caption that mention (B1, B2)?. As long as the main text mentions that we used noise as pictired in fig one that should okay. But double check
+\noindent ***\emph{We removed the variables, as requested. The editor was right, these were not
+needed. Apologies for missing this previously.} \\
 
@@
 
-\noindent ***\emph{} \\
-%%JAS: took a stab at brief explanation but maybe should be shortende
-%"Storage effects refer to effects of [temporal] niche partitioning which allow species to "store" benefits from favorable conditions to persists in unfavorable conditions"
+\noindent ***\emph{We did as requested.} \\
 
@@
 
-\noindent ***\emph{Started as new paragraph...
-But this sentence or the proceeding sentence are not good topic sentences for the next paragraph about storage effects.
-Unless editor means to make a two sentence paragraph?} \\
-%%JAS: I dont think this is a good paragraph break but I did it..
-
-\emph{We thank the referees and editor for spurring us to make these changes, as the paper is improved in
-its accessibility. The paper is certainly now no less accessible than many other papers I've read recently in 
-Ecology Letters, including the two Ellner et al. papers (cited in our study) which helped inspire this 
-work and which were published in Ecology Letters.} 
-%DONE
+\noindent ***\emph{This seemed, to us, to be an odd request, as the sentence which followed this one is the 
+last in the paragraph, so putting a new paragraph break where the referee suggested would make a very short, two
+sentence paragraph, with no clear topic sentence. Furthermore, the purpose of the paragraph this sentence
+is a part of is to introduce MCT, a purpose which is definitely served by these two sentences. So, to us,
+these sentences seemed to fit better where they were than in a new paragraph on their own. 
+So we thought about why the referee may have made such
+request and whether we might be able to address the underlying issue. It seemed to us the confusion might have come
+from the fact that, whereas the paragraph starts with general MCT, it then transitions to storage effects, and 
+then transitions back
+to general MCT at the point of the two sentences the referee has asked us to separate. 
+Whereas storage effects are part of MCT, so it makes sense to mention them in a paragraph about MCT,
+the transtion to storage effects and then back to general MCT might have been confusing. So we moved the
+two sentences in question to earlier in the paragraph. We think this is an improvement and should, hopefully,
+address the underlying issue as well as streamlining the transition to the following paragraph.}
 

### 2023-09-18 paper/Paper.Rnw
Dan thinks the referee responses and corresponding changes to the paper are complete

@@
 \noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
-\noindent \textbf{Data accessibility statement:} All codes associated with this study will be released
-in a Zenodo repository prior to publication.
+\noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
+public repsitory.
 
@@
 ecosystem functioning. 
-%Though not directly related to our questions at hand, and for illustrative purposes,
-Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs in nature using 
-plankton population density time series; many other examples of ATAs in nature have also been explored 
+Fig. \ref{fig:pedagog1}d,e show contrasting examples of ATAs using 
+plankton population density time series;
+these examples are intended only to empirically illustrate what ATAs look like, as they
+use data of a different type from what we focus on in this paper.
+Many other examples of ATAs in nature have also been explored 
 \citep{Ghosh_2020_AER,Ghosh_2020_ecolevol,Ghosh_2021}. To our knowledge, the implications of ATAs for 
@@
 through the niche partitioning effects of environmental variation.
-MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
-``storage effects'' \citep{Chesson_1994}. 
-Storage effects refer to ... 
-%effects of [temporal] niche partitioning which allow species to "store" benefits from favorable conditions to persists in unfavorable conditions
- and are reviewed conceptually below and defined formally in Theory.
-
+%moved the following two sentences from below
 MCT has been
@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Letten2018}, and corals \citep{Alvarez-Noriega_2020}. 
-%Narawani_2013
 Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+%
+MCT breaks down coexistence into the contributions of various mechanisms, with names such as 
+``storage effects'' \citep{Chesson_1994}. 
+Storage effects, which are reviewed conceptually below and defined formally in Theory, refer to
+a type of niche partitioning which allows species to effectively ``store" population growth 
+benefits from favorable time periods in order to persist through unfavorable periods.
 
@@
 $C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se}. 
-The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
-is addressed in SI section \ref{SI-SIsect:uniqueness}.
 The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.  
@@
 defined above (following \cite{Ellner_2016}).
+
+The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
+is addressed fully in SI section \ref{SI-SIsect:uniqueness}.
+Briefly, a choice 
+is made as to what type of correlation (e.g., Pearson, Spearman, etc.) is to be preserved. 
+Additionally, another choice involves the use of a bivariate normal distribution $(e_i,c_{i \bs i})$; 
+any other symmetrically tail associated bivariate 
+distribution with standard normal marginals could be used instead.
+But the normal distribution is a natural choice, and other choices may 
+be somewhat esoteric (SI section \ref{SI-SIsect:uniqueness}). 
+Future work could explore the consequences of this choice.
 
@@
 All computations were done in R on a laptop running Ubuntu Linux. 
-Complete computer codes for the project are stored at (link to be added).
-%\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
+Complete computer codes for the project are stored at 
+\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
 
@@
 for coexistence.
-%For example, the robustness of our choice in specifying a bivariate normal partial sharp distribution,
-%though a natural and canonical one, to other more esoteric specifications.
 
@@
 obtained by transforming the respectively corresponding variables (a-c; see Introduction).
-See Table 1 for used annotations and abbreviations.
 \label{fig:pedagog1}}

### 2023-09-18 code/diatom/plotdiatomdecomp.R
fixing things for plotting 5

@@ -112,10 +112,10 @@ fig5 <- function(dat_loc, invader){
     graphics::abline(v=orig[i], col='darkgrey', lty=3, lwd=2)
     
     #ATA effect blocking
-    resc <- varylist[[i]][varylist[[i]]$epsECbrk>varylist[[i]]$IGR & varylist[[i]]$IGR>0,]
+    resc <- vdat[vdat$ECbrk>vdat$GWR & vdat$GWR>0,]
     graphics::rect(resc[1,10], ymin*1.2, tail(resc,1)[1,10], ymax*1.2, col=rescol, border=NA)
     
-    excl <- varylist[[i]][varylist[[i]]$epsECbrk<varylist[[i]]$IGR & varylist[[i]]$IGR<0,]
+    excl <- vdat[vdat$ECbrk<vdat$GWR & vdat$GWR<0,]
     graphics::rect(excl[1,10], ymin, tail(excl,1)[1,10], ymax, col=excol, border=NA)
     
     #label

### 2023-09-18 paper/Paper.Rnw
cut main text length down below 5000

@@
 \noindent \textbf{Type of article:} Letter\\
-\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4975; \textbf{Words in text boxes:} 0\\
+\noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4998; \textbf{Words in text boxes:} 0\\
 \noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
@@
 \noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
-public repsitory.
+public repository.
 
@@
 At the core of ecology is the study of relationships between organisms and their environments.
-Ecologists may ask, for instance, how local environment relates to 
+Ecologists ask, for instance, how local environment relates to 
 population growth rate. To study relationships, ecologists 
-very commonly use metrics such as correlation. But standard measures of the strength 
+commonly use metrics such as correlation. But standard measures of the strength 
 and direction of association between variables do not tell the whole story 
@@
 ``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
-Metrics of association other than correlations are also often used (e.g., various forms of 
+Metrics of association other than correlations are also used (e.g., various forms of 
 regression), but other common methods also reveal only a small portion 
-of the information content of the relationship \citep{Nelsen2006}.
-ATAs and related ideas had previously been seldom considered in ecology, but recent 
-insights have been gained by studying how features of associations beyond standard measures 
+of the information in the relationship \citep{Nelsen2006}.
+ATAs and related ideas had previously been seldom considered in ecology, but recently, 
+insights were gained by studying how features of associations beyond standard measures 
 impact ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
-Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
+Ghosh_2020_ecolevol, Ghosh_2021}. \cite{Ghosh_2020_Eco} argued that extinction risks can become elevated 
 when metapopulations are subject to ATAs, e.g., if environmental ``catastrophes'' are more spatially synchronous 
@@
 than are years of good environment, limiting rescue effects. 
-And \cite{Ghosh_2021} showed that ATAs can alter the stability of 
+And \cite{Ghosh_2021} showed that ATAs alter the stability of 
 ecosystem functioning. 
@@
 these examples are intended only to empirically illustrate what ATAs look like, as they
-use data of a different type from what we focus on in this paper.
+use data of a different type from the focus of this paper.
 Many other examples of ATAs in nature have also been explored 
@@
 %Intro to competition, coexistence, MCT, etc.
-Competition and coexistence are research topics at the core of modern ecology 
+Competition and coexistence are topics at the core of modern ecology 
 \citep{MacArthur_1958, Hutchinson_1961, Schmida_1984}. The competitive exclusion principle states 
 that two species competing for the same limiting resource cannot coexist at constant densities - one will 
-instead drive the other extinct \citep{Hardin_1960}.
+drive the other extinct \citep{Hardin_1960}.
 Modern coexistence theory [MCT; \cite{Chesson_1994,Chesson_2000,Barabas_2018}] is a 
-framework for understanding how species nevertheless coexist, and biodiversity is maintained, in real ecosystems,
+framework for understanding how species nevertheless coexist in real ecosystems,
 through the niche partitioning effects of environmental variation.
@@
 tropical trees \citep{Usinowicz_2012}, plankton \citep{Caceres_1997, Descamps_2005, Letten2018}, and corals \citep{Alvarez-Noriega_2020}. 
-Although MCT is somewhat mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
+Although MCT is mathematically complex, recent extensions make it more accessible \citep{Ellner_2016,Ellner_2019}.
 %
@@
 a type of niche partitioning which allows species to effectively ``store" population growth 
-benefits from favorable time periods in order to persist through unfavorable periods.
+benefits from favorable periods in order to persist through unfavorable periods.
 
@@
 Storage effects hinge on covariation between the fluctuating quality of the environment
-for a species ($E$) and the strength of competition ($C$) experienced by that species (called $EC$ covariance);
-and on the assumption that no two species respond in the same way to environmental variation. Then, good environmental
+for a species ($E$) and the competition ($C$) experienced by that species (called $EC$ covariance);
+and on the assumption that no two species respond identically to environmental variation. Then, good environmental
 conditions for a currently common species are paired with strong intraspecific competition, because of
@@
 a currently rare species are not as closely accompanied by strong competition because the rarity of
-the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
-rare species. Storage effects can be quantified by measuring to what extent $EC$ covariance contributes to 
+the species limits competitive influence. The resulting asymmetry is what provides a benefit to the
+rare species. Storage effects can be quantified by measuring how much $EC$ covariance contributes to 
 differences between growth rates of currently rare versus currently common species \citep{Chesson_1994,Ellner_2016}.  
@@
 ATAs may be present in $EC$ covariation pertains to water limitation. Suppose two plants compete primarily for 
-water, when water is scarce, but their niches become more differentiated when water is abundant.
-Then the quality of the environment, $E$, as experienced by whichever plant is currently rare may only correspond closely to the 
+water, when water is scarce, but their niches become differentiated when water is abundant.
+Then the quality of the environment, $E$, as experienced by whichever plant is currently rare, may only correspond closely to the 
 competition, $C$, imposed on it by the currently abundant plant when water is scarce and $E$ is low.
@@
 To further demonstrate the plausibility of ATAs altering coexistence, we carried out 
-simple simulations using a two-species model, the lottery model. The simulations described in this
+simulations using a two-species model, the lottery model. The simulations described in this
 paragraph demonstrate that ATAs influence coexistence in some manner; the
@@
 The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
-in proportion to the reproductive outputs of the two species that year. Note that $N = N_1(t)+N_2(t)$ 
+in proportion to the reproductive outputs of the species that year. Note that $N = N_1(t)+N_2(t)$ 
 is constant. 
@@
 for details). Thus, the setup was identical in the three simulations except for different ATAs. 
-ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
+ATAs substantially influenced aspects of coexistence in these simulations (Fig. \ref{fig:pedagog1}f-h).
 Notation for the paper is summarized in Table \ref{tab:notation}.
 Typesetting of mathematics by ecological journals is sub-optimal,
-so we appended a latex version of this main text to the end of the 
-Supporting Information (SI); readers interested in the math should read that version.
+so, for mathematical readers, we appended a latex version of this main text to the end of the Supporting Information (SI).
 %***DAN: Don't forget to do this
@@
 %Some intuition
-We next give some initial intuition about how ATA effects on coexistence work, 
-with more details in the Discussion.
+We next give some intuition about how ATA effects on coexistence work, 
+with details in the Discussion.
 It is known that storage effects can occur if a buffering mechanism reduces population 
@@
 Bad years happen when the negative effects of $C$ outweigh any 
-positive effects of $E$; but this can occur
+positive effects of $E$; but this occurs
 both 1) when $E$ is moderately favorable but $C$ is very strong, and 2) when $E$ is 
@@
 Conversely, good years happen when positive effects of $E$ outweigh negative effects of $C$;
-but this can occur 1) when $E$ is favorable and $C$ is not too strong,
+but this occurs 1) when $E$ is favorable and $C$ is not too strong,
 and 2) when $E$ is moderately unfavorable but $C$ is very weak. 
@@
 when $E$ and $C$ are
-both lower than average (the cases 2) are not necessarily equal -- storage potential may differ
+both lower than average (cases 2) are not necessarily equal -- storage potential may differ
 across the spectrum from 1 to 2.
@@
 storage potential is high, ATA effects will be high. ATA effects quantify how temporal niche partitioning
-varies across the $EC$ spectrum spanned by the cases 1-2 sketched above.
+varies across the $EC$ spectrum spanned by the cases 1-2 above.
 
@@
 illuminate whether ATA contributions to coexistence 
-are likely to be negligible or meaningfully large. 
-G3) We provide an empirical example of species coexistence which highlights the role ATAs can play and the
-potential importance of ATAs.
+are negligible or meaningfully large. 
+G3) We provide an empirical example of species coexistence which highlights the role ATAs can play and their
+potential importance.
 The example is a laboratory diatom system. In the Discussion, we consider whether climate change may alter
@@
 
-We here pursue goal G1 from the Introduction: to augment MCT to show the role of ATAs for coexistence.
+We here pursue goal G1: to augment MCT to show the role of ATAs for coexistence.
 
@@
 Our notation parallels
-that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
+\cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
 of species $i$ can be written as an increasing function of
@@
 $\olr_{j \bs i}$ is the mean steady-state 
-growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
+growth rate of species $j$ when species $i$ is absent or at negligible abundance (equals $0$, \emph{a priori});
 and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
@@
 A subscript 
-$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
-a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
+$i \bs i$ indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
+$j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
 The \emph{scaling factor}
@@
 \cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
-alternative definitions of $q_{ij}$ that we use; but one definition is $q_{ij}=1$, and to understand the new
+alternative definitions of $q_{ij}$ we use; but one definition is $q_{ij}=1$, and to understand the 
 ideas presented here it is sufficient to use that definition. 
@@
 We now specify how to decompose storage effects into
-contributions due to ATAs and due
+contributions due to ATAs and 
 to $EC$ correlation \emph{per se}. 
@@
 were rendered independent [see (\ref{eq:storage_effects_gen})]. 
-Because components were rendered 
+Because components were 
 completely independent in the modified model, storage effects quantify the contribution of the 
@@
 corresponding to the removal of part of the relationship between $E$ and $C$ components (ATAs).
-The symbol should correspondingly be pronounced ``partial sharp''.
+The symbol should be pronounced ``partial sharp''.
 See SI section \ref{SI-SIsect:techcomm} for technical comments on the notation. 
@@
 of storage effects, which are a term in the MCT decomposition \citep{Ellner_2019} -- see 
-below for additional details.
+below for details.
 
@@
 The value of $\ddot{\rho}_i$ is determined in one of 
-four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
+four ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
 $E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
@@
 
-The question of uniqueness of the distribution $(E_i^\nat,C_{i \bs i}^\nat)$ 
+The question of uniqueness of $(E_i^\nat,C_{i \bs i}^\nat)$ 
 is addressed fully in SI section \ref{SI-SIsect:uniqueness}.
@@
 Additionally, another choice involves the use of a bivariate normal distribution $(e_i,c_{i \bs i})$; 
-any other symmetrically tail associated bivariate 
+any other symmetrically associated bivariate 
 distribution with standard normal marginals could be used instead.
@@
 The other mechanisms in MCT make no reference to covariation between quantities, whereas ATAs are 
-strictly about covariation. Therefore, storage effects are the only mechanism of coexistence
+about covariation. Therefore, storage effects are the only mechanism of coexistence
 for which ATAs are relevant. The \cite{Ellner_2019} version of MCT comprises 
@@
 where: $\Delta_i^0$ is a baseline;
-$\Delta_i^E$ is a contribution to coexistence of environmental variation;
-$\Delta_i^C$ is a contribution to coexistence of variation in competition;
-$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation 
+$\Delta_i^E$ is a contribution of environmental variation;
+$\Delta_i^C$ is a contribution of variation in competition;
+$\Delta_i^{(E\shp C)}$ is a contribution of having simultaneous variation 
 in both environment and competition, but not including the effects 
 of covariation in these quantities; and
-$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
+$\Delta_i^{(EC)}$ is a contribution of $EC$ covariation itself -- storage effects. 
 Combining our decomposition 
-(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
+(\ref{eq:storage_effects_further}) with (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
 \begin{equation}
@@
 \end{equation}
-which is the extension of MCT that was goal G1 of the Introduction. 
-The derivation of this overall decomposition is in SI section 
+which is the extension of MCT that was goal G1. 
+The derivation of (\ref{eq:full_decomp}) is in SI section 
 \ref{SI-SIsect:other_mechanisms}.
@@
 we quantified
-the magnitude of the new ATA effect term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
-scenario (see Methods). But we also defined scenarios of ``ATA rescue''
+the magnitude of the new ATA effect term, $\Delta_i^{[EC]}$, relative to
+the other terms (see Methods). But we also defined scenarios of ``ATA rescue''
 and ``ATA exclusion''. \emph{ATA rescue} is considered to have occurred if coexistence
 is possible when ATAs are taken into account, i.e., GWR$>0$ for both species, but impossible otherwise, i.e.,
-the sum of the non-ATA terms in (\ref{eq:full_decomp}) is negative for at least one species. \emph{ATA exclusion} is 
-considered to have occurred if the species do not coexist when ATAs
+the sum of the non-ATA terms in (\ref{eq:full_decomp}) is negative for at least one species. \emph{ATA exclusion} occurs 
+if the species do not coexist when ATAs
 are taken into account, i.e., GWR$<0$ for at least one species, but would coexist were it not for ATAs, 
@@
 between $0$ and $\eta_2$, for $B_2$. The mean fecundity for species $i$ was $\eta_i/2$, and larger $\eta_i$ also 
-corresponded to more variable fecundities for species $i$.
+corresponded to more variable fecundities.
 
@@
 divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. These choices for $E_i(t)$ and $C_i(t)$ fulfill the assumption of Theory 
+does not depend on $i$. These choices for fulfill the assumption of Theory 
 that $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
@@
 Second, we showed that none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
-instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
+they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
 So we plotted against $\mu_1 - \mu_2$. 
@@
 so we considered after plotting whether conclusions depended 
-on realistic values of parameters.
+on realistic parameter values.
 Simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
-Computationally efficient means of performing those simulations are in 
+Computationally efficient means of performing simulations are in 
 SI section \ref{SI-sect:efficient_computation}.
@@
 To achieve goal G3 from the Introduction, i.e., to provide an empirical example of species 
-coexistence which highlights the role ATAs can play, we used a laboratory chemostat
+coexistence which highlights the role ATAs can play, we used a laboratory 
 system of freshwater diatoms explored by \cite{Descamps_2005}.

### 2023-09-18 paper/RefereeResponses2.tex
typo

@@
 
-\noindent ***\emph{Many thanks for the positive feedback which improved our manuscript to become more understandle to other readers as well!} \\
+\noindent ***\emph{Many thanks for the positive feedback which improved our manuscript to become more understandable to other readers as well!} \\
 
@@
 Whereas storage effects are part of MCT, so it makes sense to mention them in a paragraph about MCT,
-the transtion to storage effects and then back to general MCT might have been confusing. So we moved the
+the transition to storage effects and then back to general MCT might have been confusing. So we moved the
 two sentences in question to earlier in the paragraph. We think this is an improvement and should, hopefully,

### 2023-09-25 paper/Paper.Rnw
changes to the readme and Reuman affiliations
fixed small ttypo

@@
 \author[a,b]{Pimsupa Jasmin Albert}
-\author[a,c,*]{Daniel C. Reuman}
+\author[a,*]{Daniel C. Reuman}
 
@@
 \affil[b]{Environmental Studies Program and Department of Biology, University of Oregon}
-\affil[c]{Laboratory of Populations, Rockefeller University}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
-does not depend on $i$. These choices for fulfill the assumption of Theory 
+does not depend on $i$. These choices fulfill the assumption of Theory 
 that $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 

### 2023-09-25 paper/Paper.Rnw
small typo

@@
 which are in the right halves of the distributions; and \emph{vice versa} for panel e.
-(f-h) Lottery model population simulations% using $(B_1,B_2)$ 
+(f-h) Lottery model population simulations % using $(B_1,B_2)$ 
 obtained by transforming the respectively corresponding variables (a-c; see Introduction).

### 2023-09-26 paper/RefereeResponses2.tex
deleted a superfluous comment and ammended the responses

@@
 
-\noindent \emph{We submit the following pdf files: The main text pdf, including all figures for readability; 
+\noindent \emph{We submit the following pdf files: The main text pdf, including all figures, embedded 
+next to their captions for readability; 
 the Supporting Information; a version of the main text with changes tracked; and all figures as 
@@
 references to figures, but it does accurately represent the changes made to the text.
-We also submit...
-%***DAN: all the latex files and so on.
-%*%***DAN: make sure the stuff about putting a latex version of the MT at the end of the SM is all ready to go 
-%*%in what I submit.
-} \\
+We also submit several latex and supporting files, for the production process.} \\
 

### 2023-09-28 paper/Paper.Rnw
deleted a superfluous comment and ammended the responses
adjustments required by the publisher to conform with data availability requirements

@@
 \noindent \textbf{Words in Abstract:} 150; \textbf{Words in main text:} 4998; \textbf{Words in text boxes:} 0\\
-\noindent \textbf{Number of references:} 39; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
+\noindent \textbf{Number of references:} 40; \textbf{Number of figures:} 5; \textbf{Number of tables:} 1; \textbf{Number of text boxes:} 0\\
 \noindent \textbf{Corresponding author:}
@@
 All computations were done in R on a laptop running Ubuntu Linux. 
-Complete computer codes for the project are stored at 
-\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence.git}.
+The only data used were the phytoplankton time series of Fig. 1, taken from
+\url{https://doi.org/10.5061/dryad.rq3jc84}.
+Complete computer codes and data are stored publicly as indicated in the Data Availability Statement.
 
@@
 
+\section{Author contributions}
+
+Both authors contributed about equally to all parts of the paper.
+
 \section{Acknowledgments}
@@
 The authors especially thank Steve Ellner for key advice at several watershed moments in the development of the project.
+
+\section{Funding information}
+
 The authors were partly supported by US National
@@
 Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.
+
+\section{Data availability statement}
+
+All code and data for the project are available at \url{https://zenodo.org/record/XXX}
+%DAN: update when you know
+and \url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence}.
 
@@
 \emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
-in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs were statistically
+in the North Sea considered by \cite{Ghosh_2020_ecolevol} and \cite{Sheppard2019}. ATAs were statistically
 significant and differed by location, with associations being primarily in the left tails in 
@@
 
-%***DAN: Jasmin, this currently pulls in a pdf you sent me, stored temporarily on my drive in the indicated 
-%location. Once the code produces fig 5 on a normal run, we'll have to change the below to suck that
-%version of the figure up instead of this one.
 \begin{figure}

### 2023-09-29 paper/Paper.Rnw
added zenodo record number
minor changes to the location of the data accessibility statement, to satisfy publisher formatting requests

@@
 \noindent \textbf{Statement of Authorship:} The authors jointly carried out all parts of the study.\\
-\noindent \textbf{Data accessibility statement:} All codes associated with this study have been released in a 
-public repository.
+\noindent \textbf{Data accessibility statement:} All code and data for the project are available 
+at \url{https://zenodo.org/record/8386993} and 
+\url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence}.
 
@@
 Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.
-
-\section{Data availability statement}
-
-All code and data for the project are available at \url{https://zenodo.org/record/XXX}
-%DAN: update when you know
-and \url{https://github.com/jasminalbert/AsymmetricRelationshipsAndCoexistence}.
 
